### 2013-06-19 HotOS'13/paper-long.tex
Commit of some old papers and in-progress PLOS.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,endnotes,multirow}
\usepackage{refstyle,amsmath,chngcntr}
\usepackage{epsfig,subfigure,framed}

% Used for code snippets
\usepackage{listings,courier}

% Used for changing the spacing after the title, sections, and subsections.
\usepackage{titlesec,titling}

% Make text in figure captions small
\usepackage{caption3} % load caption package kernel first
\DeclareCaptionOption{parskip}[]{} % disable "parskip" caption option
\usepackage[small]{caption}

% Make sure that figure numbers are continuous through the document
\counterwithout{figure}{section}
\counterwithout{figure}{subsection}

% Settings on code listings.
\lstset{language=C,
		xleftmargin=0pt,
		xrightmargin=0pt,
		framexbottommargin=0pt,
        framextopmargin=0pt,
        framesep=0pt}

% Modify spacing before/after title, sections, and subsections.
\setlength{\droptitle}{-3em} 
\posttitle{\par\end{center}\vspace{-1em}}
\titlespacing\section{0pt}{12pt plus 4pt minus 2pt}{3pt plus 2pt minus 2pt}        
\titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{3pt plus 2pt minus 2pt}

% Compressed enumerations
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{2pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}

% Compressed bibliograpgy
\let\ORIGbibliography\bibliography
\renewcommand{\bibliography}[1]{{\footnotesize \ORIGbibliography{#1}}}

% Add a horizontal rule above captions, and make captions closer to their figures
\DeclareCaptionFormat{ruled_caption}{\hrulefill\\#1#2#3}
\captionsetup[figure]{format=ruled_caption}
\let\ORIGcaption\caption
\renewcommand{\caption}[2][\compressedcaption]{%
\def\compressedcaption{#2}%
    \vspace{-12pt}%
    \ORIGcaption[#1]{#2}}

% Decrease spacing before \paragraphs
\titlespacing{\paragraph}{%
  0pt}{%              left margin
  0.5\baselineskip}{% space before (vertical)
  1em}%               space after (horizontal)

% Prevent footnotes from spanning multiple pages
\interfootnotelinepenalty=10000

% For leaving some comments in the draft.
\newcommand{\comment}[1]{}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Debugging Kernel Modules with Behavioural Watchpoints}

%for single author (just remove % characters)
\author{
{\rm Peter Goodman} \hspace{1.5em} {\rm Akshay Kumar} \hspace{1.5em} {\rm Angela Demke Brown} \hspace{1.5em} {\rm Ashvin Goel}\\
University of Toronto
} % end author

\maketitle
\subsection*{Abstract}

Detecting bugs in large systems is challenging. TODO TODO TODO

In this paper, we outline our implementation of \emph{behavioural watchpoints}: a new mechanism for implementing program analysis and debugging tools. We describe three simple applications of our mechanism: detecting buffer overflows, detecting use-after-free and read-before-write bugs, and enforcing fine-grained memory access policies.

In summary, we conclude.

%Finally, we present a brief case study on how we use behavioural watchpoints to detect misuses of the classical Read-Copy-Update (RCU) API in the Linux kernel.

%\iffalse
%	Some ideas for motivations:
%	\begin{enumerate}
%		\item Existing approaches to debugging are symptom based.
%		\item \texttt{gdb}-style breakpoints and watchpoints are tedious to use for detecting memory errors, and aren't helpful for detecting complex API misuses.
%		\item \texttt{printf}-style debugging is tedious because it requires annotating source code with print statements. This method of debugging generally ends up working, but it means inspecting logs, etc.
%		\item 
%	\end{enumerate}
%\fi

\section{Introduction}

A data \emph{watchpoint} is a powerful tool that reports on the usage of a particular memory location by a running program. Watchpoints are an integral feature of program analysis and debugging tools because they correlate watched memory locations with the program points accessing those memory locations. There are two categories of watchpoints: hardware-based and software-based. Hardware-based watchpoints are fast but scarce, which limits their usefulness as a means of perfoming large-scale program analyses \cite{UnlimitedWatchpoints}. Software-based implementations can scale to support millions of watchpoints \cite{DynamoRIOWatchpoints}, but are limited by their view of memory as an opaque sequence of bytes. This view is at odds with program analysis tools, which require contextual information about a program's memory. It is telling that watchpoint-based analysis tools must maintain separate bookkeeping mechanisms to recover contextual information about watched memory.

%However, complex program analysis tools require contextual information about a program's memory. It is telling that such tools must maintain separate bookkeeping mechanisms to recover this contextual information from their watchpoints.
%It is telling that software-based watchpoints are unsatisfactory for program analysis tools because these tools 
% tools because these tools must employ separate bookkeeping mechanisms that associate contextual information with watched memory.
%Implementing complex program analyses requires contextual information about watched memory, which must be separately maintained. 
%This extra bookkeeping
%acceptable for program analysis, but are limited by their all-or-nothing, shadow-memory-based approach. Program analysis tools that require contextual information about watched memory must implement new shadow memory schemes.
% needed by analysis tools requires different shadow memory schemes.
%Software-based watchpoints are acceptable for program analysis, but current implementations employ an all-or-nothing, shadow-memory-based approach to provide contextual information about watched memory.
% use of shadow memory to provide contextual information 
%fail to provide contextual information about watched memory.
%; however, they fail to 
%Software watchpoints typically come in two forms: coarse- and fine-grained. Coarse-grained watchpoints 
%Existing implementations of software watchpoints 
%accesses to particular memory locations with the program locations making those accesses. 
%Watchpoints are an integral feature of program analysis and debugging tools as they highlight the location where some memory of interest is used.
%Existing program analysis and debugging tools make heavy use of watchpoints

%In this paper, we address the scalability concerns of hardware-based watchpoints and the incongruency between the implementation and usage of software-based watchpoints for program analysis.

%introduce \emph{behavioural watchpoints}: a new form of software-based watchpoints that simplify the implementation of program analysis and debugging tools. 
% \comment{Behavioural watchpoints are an easy-to-use mechanism that simplify the implementation of debugging and analysis tools.}

In this paper, we introduce \emph{behavioural watchpoints}: a new form of software-based watchpoints that simplify the implementation of program analysis and debugging tools.  Behavioural watchpoints address the scalability concerns of hardware-based watchpoints and overcome the incongruency between how software implements watchpoints and how program analysis tools use watchpoints.

%implementation and usage of traditional software-based watchpoints by program analysis tools.

Two characteristics define behavioural watchpoints:
\begin{enumerate*}
	\item[i)] Context-specific information is embedded in each watchpoint. This information is directly available when a watched address is accessed.
	\item[ii)] The action taken when a watched address is accessed is a component of the context-specific information. This implies that different watchpoints can \emph{behave} differently.
\end{enumerate*}

%What does it mean to be behavioural?
%What it means is that context-specific information is embedded in each watchpoint, and the action invoked when a watched address is accessed is a component of the context-specific information.

%Behavioural watchpoints are behavioural insofar as the actions invoked when watched memory is accessed is included 
% action performed when watched memory is accessed is contex

%they combine contextual information about watched memory with actions (i.e. behaviours) that are invoked when that memory is accessed.

%are \emph{behavioural} insofar as they tie together actions (\emph{behaviours}) to perform with  contextual information when watched memory is accessed.

%A \emph{behaviour} is a programmer-specified callback function that is invoked when memory is accessed (in)directly through a watched address.
\noindent
Behavioural watchpoints have several novel features:
\begin{enumerate*}
	\item[i)] {\bf Unrestricted:} A single watchpoint watches many addresses. In \Secref{buffer_overflows}, we use this feature to detect various kinds of buffer overflows.
	\item[ii)] {\bf Extensible:} Watchpoints are easily extended by programmers. In \Secref{uninitialised_memory}, we use this feature to implement selective memory shadowing and detect reads from uninitialised memory.
	\item[iii)] {\bf Type-specific:} When a watchpoint is added to an address, the address's type determines which set of behaviours to invoke when (displacements of) that address are dereferenced. In \Secref{access_policies}, we use this feature to detect Linux kernel rootkits by implementing fine-grained memory access policies.
%	\item[iv)] {\bf Easy-to-use:} TODO TODO TODO
%as a means of tracking and providing more robust information to watchpoint behaviours.
\end{enumerate*}

%address the asymmetry between how software-based watchpoints are implemented and how they are used by program analysis tools. In the process, we introduce a powerful new framework

%. A hardware watchpoint triggers an exception when its watched address is derefenced.
%A hardware watchpoint triggers an exception when a watched address is dereferenced.
%for debugging and analysing uses of memory in a running program.

%The insight of our paper is that watchpoints provide a powerful framework for simplifying the implementation of program analysis and debugging tools.
%In this paper, we introduce behavioural watchpoints as a powerful framework that addresses the challenges of implementing DBT-based program analysis tools.
%we introduce behavioural watchpoints: a powerful framework for simplifying the implementation of DBT-based program analysis tools.

%is not widely adopted by the broader community

%While many useful DBT-based program analysis tools exist, implementing these tools typically requires expert knowledge of low-level processor architecture details.

%While useful, implementing non-trivial analysis tools with existing binary translation toolkits \cite{DynamoRIOKernel,DynamoRIO,PinOS} is challenging. Most examples

%Mostly people show simple uses of these toolkits and their performance, but they are not used widely because of the complexity of programming with them. The insight is that watchpoints provide a powerful framework for simplifying the implementation of various types of debugging and analysis tools. Hardware watchpoints are too limited and hence used sparingly. why. an infinite number of software watchpoints is a great idea, but still hard to use. why. that motivates "behavioral" watchpoints, and how they allow all kinds of cool potential applications.

% particularly when low-level architecture details are exposed.

%TODO TODO TODO
%A watched address is a memory address that encodes meta-information in its high-order bits. This meta-infomration distinguishes watched addresses from unwatched addresses, and is used by our run-time system to locate and invoke watchpoint-specific behaviours.

%The embedding of meta-information is such that any (typical) displacement of a watched address is also a watched address, and shares the same meta-information.

%A behvaioural watchpoint is a tool that invokes a callback function (called a \texttt{behaviour}) to all reads and writes within an arbitrary range of memory addresses.
%\emph{behaviour} to the accesses of 
%instrumentation of  
%Behavioural watchpoints are a mechanism for instrumenting accesses to the memory of an object. 
%Unlike a normal memory address, a watched address encodes meta-information that in its high-order bits. The run-time system uses this meta-information to distinguish between watched and unwatched addresses. If an address is watched, then the run-time 
% locate

\section{Implementation}
We implemented behavioural watchpoints using Granary: an extension of the DynamoRIO Kernel dynamic binary translation (DBT) framework \cite{DynamoRIOKernel,GranaryAtOSDI}. DBT is a useful technique for implementing dynamic program analysis and debugging tools \cite{Memcheck,DynamoRIOWatchpoints}. A dynamic binary translator is a program that rewrites the binary instructions of a ``guest" program while the guest program executes.

Granary was the DBT framework of choice for two reasons: 
\begin{enumerate*}
	\item[i)] Our goal was to analyse Linux kernel modules. Granary permits fine-grained instrumentation of the instructions of a running Linux kernel module.
	\item[ii)] Granary exposes rich type information about the Linux kernel. Granary has the ability to substitute the execution of a known function with a \emph{wrapped} version of itself. A wrapped function has the same type specification as its unwrapped counterpart and can freely modify its arguments and return value. This feature of Granary exposes type information which is used by behavioural watchpoints.
	%This information is a necessary component for making watchpoints ``behavioural".
\end{enumerate*}

%Granary permits fine-grained translation and instrumentation of the instructions of a running Linux kernel module.

%\subsection{Design}

%Although there are many DBT toolkits \cite{DynamoRIOKernel,DynamoRIO,PinOS}, DBT as a technique has not seen widespread adoption.  DBT toolkits are challenging to use because they operate at a low level: binary instructions. By necessity, behavioural watchpoints require low-level access to perform fine-grained memory instrumentation. The novelty of watchpoints is their interface, which enables high-level, context-specific reasoning about low-level memory accesses.

%This characteristic of DBT enabled the implementation of behavioural watchpoints, which perform low-level memory instrumentation. However, by design, behavioural watchpoints provide a high-level interface, which

%Although there are many binary translation toolkits \cite{DynamoRIOKernel,DynamoRIO,PinOS}, DBT as a technique has not seen widespread adoption because of the challenges involved in implementing non-trivial analysis tools.

%Behavioural watchpoints solve this problem by enabling low-level memory instrumentation

%Despite these challenges, the flexibility of DBT enabled the creation of behavioural watchpoints.

% motivated the creation of a mechanism to simplify the development of program analysis tools: behavioural watchpoints.

%This aspect of existing DBT tools was a motivation for making behavioural watchpoints accessible and easy-to-use.

%

\subsection{Architecture}

\begin{figure}
\begin{center}
\epsfig{file=watchpoints.eps}
\end{center}
\caption{\label{fig:watchpoint_descriptor_table}A watched address and its corresponding unwatched address are compared. The process for allocating and resolving the watchpoint descriptor for the watched address is shown.}
%A 64-bit unwatched address is translated into a watched address.The index into the watchpoint descriptor table is composed of 23 bits from the watched address. This index is created by concatenating bits $[20,27]$ from the unwatched address with the next available partial index counter. This counter is located in the partial index counter table at the index represented by bits $[20,27]$ from the unwatched address.
\end{figure}

The implementation of behavioural watchpoints distinguishes between watched addresses and their descriptors.

A \emph{watched address} is a pointer with an index into the \emph{watchpoint descriptor table} embedded in its bits (\Figref{watchpoint_descriptor_table}). The $23$-bit index into the descriptor table is formed by concatenating bits $[20,27]$ (called the \emph{counter index}) with bits $[48,62]$ (called the \emph{partial index}). A watched address and its unwatched counterpart share the same counter index; however, the partial index of a watched address varies\footnote{This feature of watchpoints allows for a one-to-one mapping between a watched and unwatched address, and a one-to-many mapping between a watchpoint and all addresses watched by that watchpoint. The one-to-one mapping is formed by masking the high-order bits containing the partial index.}. Partial indexes are recorded in the \emph{partial index counter table}. When a watchpoint is allocated, the current value stored in the counter table for that specific counter index is incremented and returned as the watchpoint's partial index. This allocation strategy allows for at most $2^{15}$ watchpoints per counter index or megabyte of memory\footnote{x86 has byte-addressable memory. The counter index begins at bit $20$, giving each watchpoint $1$ MB = $2^{20}$ B degrees of freedom. However, if an over/underflow across the 1 MB-aligned boundary occurs then the counter index will be corrupted. We can correct one bit of corruption by requiring that the counter index and the partial index have the same sign. An alternative solution is uses a different indexing scheme. We have successfully experimented with a different indexing scheme that solves the aforementioned overflow errors, but sacrifices the one-to-one relationship between a watched and unwatched address.}.

%For each counter index, there can be at most  partial indexes.

%and unwatched addresses share the same counter indexes, b

A \emph{watchpoint descriptor} is a data structure containing a base address, a pointer to a virtual table (vtable) of memory operations, and programmer-defined meta-information. Each vtable provides eight functions: four read and four write functions. Each function is specific to a memory operand size (1, 2, 4, or 8 bytes). A watchpoint is initialised with either a generic or a type-specific vtable pointer. Type-specific vtables are specific to the \emph{type} of the watched pointer. When invoked, a vtable function operates on the watched address and its descriptor. Behavioural watchpoints earn their name from their ability to behave differently based on the meta-information stored in the descriptor and the (type-specific) vtable function invoked.

%The extension of watchpoint descriptors with programmer-defined meta-information combine with vtable functions to 
%The combination of programmer-defined meta-information and type-specific vtables 
%Such vtables are used to great effect in our memory-related bug detection applications (\Secref{applications}).

A watchpoint is added to a previously unwatched pointer by adding an entry to the watchpoint descriptor table and embedding part of the index of the entry in the high-order bits of the pointer. The original value of the pointer is saved as the base address of the descriptor. A watchpoint is removed from a pointer by masking the high-order bits containing the partial descriptor index. This approach permits a one-to-one mapping between a watched address and its unwatched counterpart, while allowing a single descriptor to watch many addresses. Specifically, a single watchpoint can watch up to $2^{20}$ distinct addresses, and each megabyte of memory can be watched by up to $2^{16}$ distinct watchpoints.

%at most $2^{20}$ different addresses.

Watched addresses, however, do not reference valid memory. Dereferencing a watched address triggers a hardware exception. To avoid this, we use Granary to dynamically translate memory loads and stores to first check for and then resolve watched addresses to their unwatched counterparts. If a watched address is detected then a watchpoint-specific vtable function is invoked. Both the address of the watchpoint descriptor and the watched address are passed as arguments to the invoked function. Finally, the original memory instruction is emulated by one that uses the unwatched address. The choice of which vtable function to invoke depends on the size of the memory operand (1, 2, 4, or 8 bytes) and whether the operation reads or writes its operand from/to memory.

%Watchpoints are created by adding an entry to the watchpoint descriptor table, and modifying a pointer in-place to include 
%The structure of a watched address is such that any displacement of a watched address will associate with the original
%Watchpoints associate zero or more watched addresses with a single watchpoint descriptor, located in the . 
%\subsection{Optimizations}

\section{Applications\label{sec:applications}}
The following sub-sections describe three applications of behavioural watchpoints and their implementations. In these sub-sections, we use the term \emph{object} to describe any value with a named location in memory.

\subsection{Buffer Overflows \label{sec:buffer_overflows}}
A buffer overflow is a software bug where a program--in an attempt to write to some object's memory--actually writes to adjacent memory cells. Detecting buffer overflows is important because they are commonly used in remote code execution and privilege-escalation attacks against operating systems \cite{SecureProgramExecFlowTracking}. One method of detecting buffer overflows relies on the compiler to allocate ``poisoned" regions of memory around each object \cite{AddressSanitizer}. Accessing this memory is an error because poisoned memory (which is transparently added by the compiler) does not exist as a named program entity. Our approach is similar in that memory adjacent to a watched object is considered poisoned; however, we do not rely on compiler support, nor do we allocate or mark poisoned memory as such.

We employ three overflow detection policies: heap-based, type-based, and stack-based. \comment{For uniformity, the blind policy is a special case of the targeted policy.} All three policies depend on the same extension to the meta-information of watchpoint descriptors: a \emph{limit address}. Together, the base address (stored in the descriptor) and the limit address delineate an object's boundaries in memory \cite{BccFatPointers}.

\paragraph{Heap-based overflow detection \label{sec:heap_overflow}}
\begin{figure}
\begin{lstlisting}[language=C,basicstyle=\footnotesize\ttfamily]
FUNC_WRAPPER(__kmalloc, (size, flags), {
  void *addr = __kmalloc(size, flags);
  ADD_WATCHPOINT(addr, size);
  return addr;
})
\end{lstlisting}
\caption{\label{fig:kmalloc_wrapper}Definition of the \texttt{\_\_kmalloc} function wrapper in Granary. The above code expands into a function for wrapping the \texttt{\_\_kmalloc} kernel memory allocator. Calls to \texttt{\_\_kmalloc} are transparently substituted with calls to the generated wrapper. The wrapper invokes the original \texttt{\_\_kmalloc} function and returns a watched version of the allocated address.}
\end{figure}

The heap-based detection policy detects buffer overflow errors on all heap-allocated objects. We use Granary to substitute invocations of the kernel's memory allocators (e.g. \texttt{kmalloc}) with wrapped versions of the allocators (\Figref{kmalloc_wrapper}). A wrapped allocator invokes the original allocator but returns a watched version of the allocated address. The watched address returned has its descriptor initialised with the base address as the allocated address and with the limit address as the base address plus the requested allocation size. All buffer overflow detecting watchpoints are initialised with the same vtable pointer.

%\comment{TODO: move first sentence into implementation}
%To detect buffer overflows (including heap-based overflows), we use the same vtable of operations and watchpoint meta-information (limit address).
%The vtable used by buffer overflow detecting watchpoints makes use of the operand size-specific nature of vtable functions.

The vtable pointer used by all buffer overflow detecting watchpoints points to a table of memory operand size-specific functions. The operand size is necessary to detect dereferences of memory that overlap both the object and its adjacent memory, as illustrated in case \emph{(3)} of \Figref{detect_overflow}. The vtable function corresponding to the memory operand size is invoked when a watched address is dereferenced. Each vtable function is programmed to check the watched address against the base and limit addresses stored in watchpoint's descriptor (\Figref{detect_overflow}). If a buffer underflow or overflow is detected then the vtable function notifies the run-time system.

\begin{figure}
\abovedisplayskip=0pt
\belowdisplayskip=0pt
\begin{center}
	\epsfig{file=overflow.eps}
\end{center}
\iffalse
	\abovedisplayskip=0pt
	\belowdisplayskip=0pt
	\begin{align*}
		A_{base}  & > {A_{unwatched}} \tag{Underflow} \\
		A_{limit} & \leq {A_{unwatched}} \tag{Overflow} \\
		A_{limit} & < {A_{unwatched}} + S_{op} \tag{Overlap}
	\end{align*}
\fi
\caption[Caption for LOF]{\label{fig:detect_overflow}Three common buffer overflow cases are illustrated above: 1) underflow 2) overflow, and 3) overlap. Beside each illustration is the policy that detects the specific bug. When a watched address ($A_{watched}$) is dereferenced, a vtable function that is specific to the memory operand size ($S_{op}$) is invoked. This function detects a buffer overflow if the intended referenced memory address ($A_{unwatched}$) does not fall within the boundaries delineated by the base and the limit addresses ($A_{base}$ and $A_{limit}$, respectively).}
\end{figure}
%\protect\footnotemark
%\footnotetext{Policy (2) of \Figref{detect_overflow} is redundant but is illustrated for clarity.}

\paragraph{Type-based overflow detection}
The type-based detection policy is applied when the type of a pointer is known (i.e. not \texttt{void*}). The simplest use case of type information is to infer the limit address of a typed pointer. A more complex use case arises when the fields within a structure encode memory bounds information. In both cases, the vtable functions for typed and untyped memory operate in the same way.

%For example, a variable-sized array can be represented as a structure type with two fields: the number of objects in the array and a pointer to the array's objects. Out-of-bounds array accesses are detected by replacing the address of the array's objects in the structure with a watched address. The watchpoint's limit address is initialised to the address of the objects plus the number of objects in the array scaled by the size of each object. As before, the size of each object is inferred from the type of the object pointer field.

The type of an addressable object becomes known to Granary in the context of wrapped functions. Granary recursively follows argument and return value pointers based on object type specifications. These specifications tell Granary how to follow pointers and what--if any--pointers should be converted to watchpoints. To reduce programmer effort, Granary automatically generates type specifications and wrapper functions for any set of known types and functions. Some manual post-processing is required if the analysis/debugging application needs to understand semantic relationships between function arguments or structure fields. \comment{Automating this process was an important goal of Granary because of Granary wraps all (approx. 6,000) exported kernel functions.}

%If a pointer to that object is reachable from an argument to, or from the returned value of, a Granary-substituted wrapper function. Within a wrapper, Granary recursively follows argument pointers

%For example, if a function accepts a variable length array as an argument (encoded as a tuple of the size of the array and a pointer to the objects in the array) then a watchpoint can be used to detect out-of-bounds array accesses. Detecting such accesses requires replacing the address of the array's objects with the watched version of that address. The descriptor of this watchpoint is initialised with the number of objects in the array and the size of each object. Like the blind policy, the targeted detection policy performs the same addressability checks using the descriptor's meta-information.

\paragraph{Stack-based overflow detection}
Detecting stack overflows at the granularity of function activation frames requires more support from Granary's run-time system. At a high level, each activation frame is assigned a dedicated watchpoint when a function is called. The effect of an instruction that grows or shrinks an activation frame's size is reflected by a similar change to the base address\footnote{The run-time call stack on x86 grows into lower memory.\comment{ Modifying the base address instead of the limit address allows us to maintain one set of vtable functions for all three buffer overflow detection policies.}} of the frame's watchpoint descriptor. When a function returns, the descriptor of its activation frame is cleared, but remains allocated\footnote{Reclaiming/reusing allocated but unlikey-to-be-used watchpoints is challenging. One overflow-specific approach is to ensure that the next use of the watchpoint is for a buffer whose bounds do not intersect with the current use's bounds. Another approach is to use some number of context switches as a grace period in which the watchpoint cannot be reused, but after which it can be reused. We leave this as future work.} lest the address of a local variable escape the function\footnote{If the address of a stack-allocated object escapes its function activation frame, then that pointer is said to be \emph{dangling}. Dereferences of dangling pointers can cause stack overflow errors. In the kernel, data stored in unallocated stack memory (e.g. through a dangling pointer) is at risk of being clobbered by interrupt stack frames and by function activation frames.}.

We assume that a memory instruction operating on a constant displacement of the stack or frame pointer registers is well-behaved. However, if the stack or frame pointer registers participate in a memory instruction and the displacement from the register is dynamically-bound then that instruction is suspect. An instruction that copies the address in the stack or frame pointer registers is also suspect because the copied address might escape the current function or participate in a local stack overflow. All suspect instructions are translated to transparently add the frame's dedicated watchpoint to the dereferenced or copied address.

%To prevent stack overflows on stack-allocated variables that escape the current function, we 
% is translated to use the current frame's watchpoint.
%An offsetted dereference of stack or frame pointer is suspect if the offset is dynamically bound. 
%We detect buffer overflows when the stack or frame pointers are derefe
%on dynamically offsetted dereferences of the stack or frame pointers, and aliases of the stack or frame pointers.
% Stack-based protection is extended to pointers that potentially escape the current function. For example, if an instruction copies the address stored in the stack or frame pointer registers into a third location, then that instruction is translated to add the frame's watchpoint to the copied address.

This scheme segments the runtime stack into contiguous but non-overlapping regions of watched memory. A straightforward extension excludes saved return addresses and link pointers from the known bounds of activation frames, thus detecting return-oriented buffer overflow attacks.

%The typical activation frame layout on x86 
%When a function is called, a watchpoint is allocated with its base and limit addresses to the address immediately below 
%Our approach uses watchpoints to segment the runtime call stack into its activation frames and is able to detect 
%a simple representation of a variable length array contains the size of the array and a pointer to the objects of the array. 
%counterparts are invoked. The type signature of the functions being wrapped betray 
%Each wrapped function can modify the (typed) arguments and return value 
%The type of a pointer is known to Granary when a function is invoke

%Granary uses a description of the layout a typed object to discover  pointers contained within the object, as well as relations between fields of an object.

%The structure of a referenced type betrays meaningful semantic information, which can. For example, a 
%Like the blind policy, the targeted detection policy uses the same meta-information to detect buffer overflows.
%This vtable is initialised with functions 
% The vtable pointer is initialised to point to a generic 
%The first policy uses Granary to wrap the Linux kernel memory allocation routines . The wrapped versions of these routines invoke their unwrapped counterparts, but modify their return values (the address of the newly allocated memory) by converting them to watchpoint addresses.

%Our approach depends heavily on Granary's ability to wrap known functions. 
% functionality to instrument

%map to any semantically valid program references.
%erroneous program behaviour (buffer overflow) because poisoned regions of memory are transparently added by the compiler. 
%These poisoned regions, called red zones, trigger hardware exceptions when erroneously accessed.
%Adjacent memory cells can be ``poisoned" and accesses to poisened memory causes a hardware exception,

\subsection{Uninitialised Memory Bugs\label{sec:uninitialised_memory}}

In this section, we show how to use behavioural watchpoints to detect uses of uninitialised memory.  Detecting such uses is important because a program using uninitialised memory can exhibit unintended, non-deterministic behaviour.

We define uninitialised memory as either allocated memory to which no value has been written, or de-allocated memory. We selectively track the initialisation state of each byte of watched memory using \emph{shadow memory} \cite{Memcheck}. Our implementation represents shadow memory as a bitset, where the state of each bit tracks the initialisation state of a byte of watched memory. Watchpoint descriptors are augmented to include the size (in bytes) of a watched object and a pointer to the memory shadowing the object\footnote{As an optimisation, shadow memory for small objects can be embedded in the watchpoint descriptor.}.

%As an optimisation, shadow memory for small objects (size $\leq$ 56 bytes) is embedded in the seven high-order bytes of the descriptor field containing the size of the watched object. For small objects, the shadow memory pointer points into the descriptor field storing the size of the shadowed object. Shadow memory is independently allocated for large objects. This optimisation does not prevent objects larger than $2^{8}$ bytes from being tracked: the descriptor's size field serves its dual role only when the object is small.

%This optimisation does not prevent objects. 
%For large objects, . Independent allocation of shadow memory for large objects implies
%do not interfere with this optimisation because the six high-order bytes only store 
%This optimisation assumes that no object is larger than . 
% In the kernel, the original address of shadow memory is recovered by masking the high-order bits. Shadow memory for small and large objects is accessed uniformly: the shadow memory pointer is masked before it is dereferenced. For small objects, the address of shadow memory is the address of the two high-order bytes of the descriptor field storing the pointer to shadow memory. For large objects, shadow memory is independently allocated.

We apply the above shadow memory scheme to detect uses of uninitialised memory for heap-allocated objects. As in \Secref{heap_overflow}, we interpose on the kernel's memory allocators and add watchpoints to the addresses returned by those allocators. When memory is allocated, the watchpoint descriptor is initialised with the number of allocated bytes (available as an argument to the allocator) and with its own shadow memory. The bits in shadow memory are initialised to zero.

\paragraph{Read-before-write bugs}
A read-before-write bug occurs when uninitialised memory is read by a program. This bug can cause non-deterministic behaviour because uninitialised memory can contain arbitrary values. If a control-flow instruction (such as a branch) decides its target based on the uninitialised value read from memory, then the resulting control-flow can be random.

%memory is read before it is written.
%Uninitialised memory bugs often go unnoticed, especially in complex code, 
%, by luck, the previous state of the read memory cells is
%This bug can cause non-deterministic behaviour 
%if the direction that a branch instruction takes is decided by the value stored in uninitialised memory.

%expects that reads of uninitialised memory 
%depends on there being a semantically meaningful value stored at the memory location being read.

All watchpoint descriptors for detecting uses of uninitialised memory are initialised with the same vtable pointer. The read-specific vtable functions report to the run-time system when the shadow bit corresponding to one of the read bytes is zero. The write specific vtable functions set to one all shadow bits that correspond to overwritten bytes. Both groups of vtable functions perform bounds checking\footnote{This check is functionally similar to \Figref{detect_overflow}, but uses the descriptor's base address and object size instead of using a base and limit address.} on the dereferenced address.

Unlike our other memory error checking policies, the read-before-write bug policy detector can report false positives. It is common for larger-than-needed reads to be performed and for compiler-added structure padding to be read (but never written). A relaxed version of our read-before-write memory checking policy requires that at least one shadow bit is set for every read operation.

%The read-specific vtable functions look to see if any bit in the set of bits shadowing the bytes of memory being read is zero. A debugger breakpoint is triggered if a zero bit is found. The write-specific functions set the bits that shadow the bytes being overwritten to one.

\paragraph{Use-after-free bugs}
A use-after-free bug occurs when memory is accessed after it has been deallocated. This bug can cause non-deterministic behaviour if the once occupied memory of the  deallocated object is partially or fully reallocated for another purpose.

We use Granary to interpose on the kernel's deallocation functions (e.g. \texttt{kfree}). If the argument to a wrapped deallocator function is a watched address then the descriptor is cleared. Finally, the original deallocator is invoked with the unwatched address as an argument. Reads and writes to a watched address of a deallocated object fail to pass bounds checks because the object size is zero\footnote{Two positive side-effects of our approach is that it automatically detects double-free bugs and freeing of non-heap memory bugs. A double-free is detected when a watched address for an already deallocated object is passed as an argument to a deallocator. An attempt to free non-heap memory is detected when an unwatched address is passed as an argument to a deallocator.}.

%\footnote{When masked, the pointer to shadow memory for small objects points to itself. This enables uniform access to shadow memory for all object sizes.}.

%Watchpoint descriptors are augmented with a limit address and a dual-purpose 64-bit pointer to shadow memory.

%Small objects ($\leq$ 16 bytes)

%The buffer overflow policies of \Secref{buffer_overflows} augments watchpoint descriptors with shadow memory

%An uninitialised memory bug is characterized as one where memory...

\subsection{Fine-grained Access Policies\label{sec:access_policies}}
In this section, we describe the implementation and use of the field-accessor API: a novel extension to behavioural watchpoints. We use this API to detect Linux kernel rootkits. Detecting rootkits is important because rootkits are malicious programs that infect and corrupt operating systems. Rootkits are difficult to detect because they have access to  privileged functionality, which allows them to hide their existence. However, rootkits often leave hints of their existence in the form of violated data structure invariants \cite{GibraltarKernelInvariants,OSck}. For example, \Figref{field_invariant_check} discusses an invariant which, if violated, can prevent an anti-virus program from scheduling periodic scans---scans which might otherwise detect the presence of a rootkit. We use the field-accessor API to actively detect violations of data structure invariants and report them to our run-time system as potential rootkits.

%We use the field-accessor API to actively detect violations of data structure invariants, thus 
%We detect rootkits by using the field-accessor API to detect violations of data structure invariants. 
%detects a violation of a data structure invariant that 
%We show how the field-accessor API enables Granary to detect violations
%kernel data structure invariants and
% and can easily hide themselves or prevent their own detection
%Detecting kernel rootkits is challenging because the entire surface area of an operating system kernel 
%We use the field-accessor API to detect Linux kernel rootkits. We detect rootkits by encoding data structure invariants

The field-accessor API is a declarative API for interposing on memory reads and writes at the granularity of \texttt{C} structure (bit)fields. With the API, one can attach a callback function to memory reads/writes of a structure field (\Figref{field_invariant_check}). An attached read or write callback is invoked when the specific field within a watched object of the correct type is dereferenced. The type of an object is implicity resolved; watchpoint vtables are type-specific, and the field-accessor API is implemented in terms of watchpoint vtables. The code that connects vtables to attached callbacks is automatically generated. Generating this code involves parsing \texttt{C} header files and determining for each \texttt{C} structure type $T$, all fields accessed by a dereference within the memory occupied by a watched instance of $T$. If such a dereference occurs, then an attached callback function is invoked for each touched field.

%This extension is enabled by the type-specific nature of watchpoint vtables and by Granary's database of type information. We apply this extension to detect violations of data structure invariants in the Linux kernel. Detecting violations of kernel data structure invariants is important because such violations can signal the presence of kernel rootkits \cite{GibraltarKernelInvariants,OSck,LXFI}.

% The field-accessor API is an API for attaching a callback function to memory reads and/or writes of a structure field.

%Because watchpoint vtables are type-specific, The type of the object is known to the watchpoint from the time at which it was added
%The type of the object is known from the point 
%Our extension is limited by its dependency on advance knowledge of the types used by an instrumented program. This limitation does not prevent us from checking invariants on arbitrary binary Linux kernel modules. 
%Detecting violations of data structure invariants is important as such violations can be a signal of buggy code or 
%Detecting such violations is important as they signal potential kernel rootkits .

%TODO TODO TODO: Properly introduce field-accesor API as an advanced use of the type-specific vtables.

%In this section, we describe the implementation and use of the field-accessor API: a novel extension to behavioural watchpoints. The field-accessor API permits interposing on memory reads and writes at the granularity of structure fields (including bitfields). We apply this API to detect violations of data structure invariants in the Linux kernel. Detecting violations of invariants can be used to detect kernel rootkits \cite{GibraltarKernelInvariants,OSck}.

%This requirement is not overly strict, as will be shown for Linux kernel modules.
%If a read or write callback is attached to a structure's field, and that field is accessed through a watched address, then its callback is invoked when that field is read from or written to memory.
% dereference of that field through a watchpoint
% defining per-structure-field read/write access callback functions. If a field in an object with a known structure type is dereferenced through a watched address, then its corresponding callback function is invoked.

%The Linux kernel is a large software system with a complex and flexible API: exported functions, variables, and types. Extensions to the kernel, called modules, use this API to support new hardware (e.g. graphics drivers) and provide new features (e.g. file systems). However, misuses of data structures exposed through the API can lead to crashes or security breaches \cite{GibraltarKernelInvariants,OSck,LXFI}.

%The Linux kernel's exported API is defined across a set of \texttt{C} header files. We parse these files and automatically generate code that determines, for each type $T$, all fields ``touched" by a dereference within the memory occupied by a watched instance of $T$. If such a dereference occurs, then an attached callback function is invoked for each touched field.

\iffalse
	The steps taken to generate this code are described below:
	\begin{enumerate}
		\item Kernel header files are parsed. The set of all exported functions, variables, their dependent types, and the structure of those types is recorded.

		\item The recorded type information is transformed into a program that reports the bit boundaries of each structure field within its containing type. At runtime, this program creates zero-initialized instances of each type and selectively sets to one the bits of each of the type's internal fields. The bits of memory of each object are linearly scanned to determine the exact bounds of each field.

		\item The derived type layout information is processed to create a unique \texttt{C++} template class for each kernel structure type. A singleton instance of the instantiated template class for some type $T$ is the operations vtable for all watched objects of type $T$.

		The method bodies of the template class invoke attached field access callback functions. The choice of which callback(s) to invoke is statically determined for each possible dereference size (1, 2, 4, or 8 bytes) and byte offset into an object's memory. At runtime, the byte offset of a dereference of a watched object is computed by subtracting the descriptor's base address from the dereferenced address, modulo the size of the object's type.

		Our use of \texttt{C++} templates serves two purposes. First, if a watchpoint is never added to an object of type $T$, then the corresponding template is never instantiated. This reduces code bloat when Granary is compiled. Second, if a watchpoint is added to an object of type $T$, then template instantiation is what permits the correct and type-specific automatic initialisation of the vtable pointer in the watchpoint's descriptor.
	\end{enumerate} 
\fi

\Figref{field_invariant_check} shows how an invariant on the \texttt{ioctl} field of any watched object whose type is \texttt{struct file\_operations} is checked using the field-accessor API.

\begin{figure}
\begin{lstlisting}[language=C,basicstyle=\footnotesize\ttfamily]
// Invariant: rtc_fops->ioctl == &rtc_ioctl
WATCH_WRITE(struct file_operations, ioctl, {
  if(&rtc_fops  == base_address
  && &rtc_ioctl != ioctl) {
    // potential attack: prevent an anti-
    // virus scan from being scheduled!
  }
})
\end{lstlisting}
\caption{\label{fig:field_invariant_check}This code example shows how to check invariant 1(h) from \cite{GibraltarKernelInvariants} using Granary's field-accessor API. The invariant checked prevents a (potential) kernel rootkit from installing its own \texttt{ioctl} handler into the Real-Time Clock. Anti-virus programs depend on this built-in \texttt{ioctl} handler to periodically schedule virus scans. Replacing this handler can prevent such scans from being scheduled, thus allowing the rootkit to go undetected.}
\end{figure}

% watched pointer of

%it allows a programmer to define per-structure-field read/write access policies, which are invoked whenever a (third-party) binary module dereferences the fields of a watched pointer of the correct type.

%A caveat of our approach is that fine-grained

%the type of objto be watched must be known in advance. However,

%TODO: example of \texttt{get\_thread\_info}, to the task struct, to become a rootkit.
%Linux kernel modules interact with the kernel using a well-defined inteface: exported functions. Large amounts of sensitive data structures are shared over this interface, and the kernel expects that modules behave sanely.

%\section{Case Study: Detecting RCU Usage Bugs}
%We have established that behavioural watchpoints are useful for implementing memory error detection and protection policies. The extensibility of watchpoints, however, makes them useful for a broader range of applications. To that end, this section reports on some early results from our watchpoint-based tool that detects misuses of the classical Read-Copy-Update (RCU) API in the Linux kernel. Detecting misuses of the RCU API is important because RCU is increasingly (mis)used as the synchronization mechanism of choice within the Linux kernel \cite{RCUInLinux}.

%\subsection{Read-Copy-Update}
%RCU is a lock-free synchronization mechanism allowing reader and writer threads to operate concurrently on a shared data structure \cite{RCU}. RCU is applicable when the ratio of the number of reader to writer threads is high, and when reader threads can tolerate ``stale" data.

%RCU reader threads operate in a non-blocking manner and execute concurrently with writers. Because of this, some readers might observe stale data, e.g. reading the value of a pointer immediately before that pointer is modified by a writer. Writer threads serialize to prevent the coexistence of two divergent ``views" of a data structure. Serialization of writers in the Linux kernel is often implemented by having writer threads compete for a mutex or spinlock. It is typical for the winning thread to wait for all concurrently executing reader threads to complete their outstanding reads before making changes that might affect the consistency of the data structure. A reader thread announces the existence of outstanding reads to writer threads by grouping its reads of the data structure into a \emph{read-critical section}.

%\paragraph{RCU API}
%TODO

%\subsection{TODO}

\section{Evaluation}
TODO

\section{Conclusions and Future Work}
TODO

%The writer thread whose turn it is to modify the data structure waits for all concurrently executing reader threads to complete their outstanding reads of the data structure. Waiting ensures that reader threads do not observe inconsistent data.

%Reader threads signal the beginning and end of data structure reads by invoking \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock}, respectively.

%This characteristic of RCU naturally segments the liftetime of a data structure into \emph{generations}. Under this lens, the acquisition of mutual exclusion over a data structure by a writer thread is a signal that a generation is near its end. The true end of a generation occurs when the writer thread that acquired mutual exclusion waits for all reader threads to complete any outstanding reads of the data structure. When one generation ends, the next immediately begins. This implies the coexistence of multiple generations. Correct usage of the RCU API requires that

%This view is complicated by the coexistence of multiple generations; however, correct usage of the RCU API ensures that an individual reader thread never ovserves 
%Multiple generations of a RCU-protected data structure may coexist; however, proper 
%is when all concurrently \emph{active} readers at the time 
%must acquire mutual exclusion over the entirety of a protected data structure before making changes that must be visible in the most up-to-date view of the data structure.
%RCU reader threads are permitted to read the protected data structure without 
%At a high level, RCU implicitly segments the lifetime of a data structure into \emph{generations}. As one generation ends, another automatically begins.
%A grace period
%A grace period ends delimited by a writer thread acquiring the exclusive right to modify values stored in the protected structure.
% over the entirety of the protected data structure.

%approach to synchronizing concurrently executing reader and writer threads. RCU is unique
%In this section, we describe an early implementation

% The bibliography should be embedded for final submission.
\bibliographystyle{acm}
\bibliography{library}
\end{document}

### 2013-06-19 HotOS'13/paper.tex
Commit of some old papers and in-progress PLOS.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,endnotes,multirow}
\usepackage{refstyle,amsmath,chngcntr}
\usepackage{epsfig,subfigure,framed}
%\usepackage{multicol,tabularx}
\usepackage{textcomp} % for tilde

% Used for code snippets
\usepackage{listings,courier}

% Used for changing the spacing after the title, sections, and subsections.
\usepackage{titlesec,titling}

% Make text in figure captions small
\usepackage{caption3} % load caption package kernel first
\DeclareCaptionOption{parskip}[]{} % disable "parskip" caption option
\usepackage[small]{caption}

% Make sure that figure numbers are continuous through the document
\counterwithout{figure}{section}
\counterwithout{figure}{subsection}

% Settings on code listings.
\lstset{language=C,
		xleftmargin=0pt,
		xrightmargin=0pt,
		framexbottommargin=0pt,
        framextopmargin=0pt,
        framesep=0pt}

% Modify spacing before/after title, sections, and subsections.
\setlength{\droptitle}{-3em} 
\posttitle{\par\end{center}\vspace{-1.5em}}

\titlespacing\section{0pt}{5pt plus 4pt minus 2pt}{5pt plus 2pt minus 2pt}        
\titlespacing\subsection{0pt}{5pt plus 4pt minus 2pt}{5pt plus 2pt minus 2pt}

% Compressed enumerations
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{2pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}

% Compressed bibliograpgy
%\let\ORIGbibliography\bibliography
%\renewcommand{\bibliography}[1]{{\footnotesize \ORIGbibliography{#1}}}

% Add a horizontal rule above captions, and make captions closer to their figures
\DeclareCaptionFormat{ruled_caption}{#1#2#3\hrulefill}
\captionsetup[figure]{format=ruled_caption}
\let\ORIGcaption\caption
\renewcommand{\caption}[2][\compressedcaption]{%
\def\compressedcaption{#2}%
    \vspace{-12pt}%
    \ORIGcaption[#1]{#2}%
    \vspace{-12pt}}

% Decrease spacing before \paragraphs
\titlespacing{\paragraph}{%
  0pt}{%              left margin
  0.25\baselineskip}{% space before (vertical)
  1em}%               space after (horizontal)

% Prevent footnotes from spanning multiple pages
\interfootnotelinepenalty=10000

% For leaving some comments in the draft.
\newcommand{\comment}[1]{}

\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Behave or Be Watched: Debugging with Behavioural Watchpoints}

%for single author (just remove % characters)
\author{
{\rm Peter Goodman} \hspace{1.5em} {\rm Akshay Kumar} \hspace{1.5em} {\rm Angela Demke Brown} \hspace{1.5em} {\rm Ashvin Goel}\\
University of Toronto
} % end author

\maketitle
\subsection*{Abstract}

Finding, understanding, and fixing bugs in an operating system is challenging. Data watchpoints are a powerful tool that can help with this problem; however, hardware-based watchpoints are too scarce for many uses and software-based watchpoints lack contextual information about the watched memory.

In this paper, we introduce \emph{behavioural watchpoints}: a new form of software-based watchpoint that simplifies the implementation of program analysis and debugging tools. Behavioural watchpoints extend on traditional watchpoints in several novel ways: they maintain context-specific information about watched memory, they watch ranges of addresses, and they invoke programmer-defined functions when watched memory is accessed. We describe three applications of these watchpoints: detecting buffer overflows, detecting read-before-write and memory freeing bugs, and enforcing fine-grained memory access policies.

We show that behavioural watchpoints, as implemented within a kernel dynamic binary translation framework, have reasonable overheads for their intended use in analysing and debugging kernel modules.

\section{Introduction}

%A data \emph{watchpoint} is a powerful tool that reports on the usage of a particular memory location by a running program. 
Watchpoints are an integral feature of program analysis and debugging tools because they correlate watched memory locations with the program points accessing those memory locations. There are two categories of watchpoints: hardware-based and software-based. Hardware-based watchpoints are fast but scarce, which limits their usefulness as a means of perfoming large-scale program analyses \cite{UnlimitedWatchpoints}. Software-based implementations can scale to support millions of watchpoints \cite{DynamoRIOWatchpoints}, but are limited by their view of memory as an opaque sequence of bytes. This view is at odds with program analysis tools, which require contextual information about a program's memory. As a result, watchpoint-based analysis tools are harder to design and implement because they must maintain separate bookkeeping mechanisms to recover contextual information about watched memory.

In this paper, we introduce \emph{behavioural watchpoints}: a new form of software-based watchpoints that simplify the implementation of program analysis and debugging tools.  Behavioural watchpoints address the usability concerns of hardware-based watchpoints and overcome the incongruency between how software implements watchpoints and how program analysis tools use watchpoints.

Two characteristics define behavioural watchpoints:
\begin{enumerate*}
	\item[i)] Context-specific information is embedded in each watchpoint. This information is directly available when a watched address is accessed.
	\item[ii)] The action taken when a watched address is accessed is a component of the context-specific information. This implies that different watchpoints can \emph{behave} differently.
	\item[iii)] A single watchpoint can watch multiple addresses.
\end{enumerate*}

%\section{Implementation}
%DBT is a useful technique for implementing dynamic program analysis and debugging tools \cite{Memcheck,DynamoRIOWatchpoints}. A dynamic binary translator is a program that rewrites the binary instructions of a ``guest" program while the guest program executes.

%Granary was the DBT framework of choice for two reasons: 
%\begin{enumerate*}
%	\item[i)] Our goal was to analyse Linux kernel modules. Granary permits fine-grained instrumentation of the instructions of a running Linux kernel module.
%	\item[ii)] Granary exposes rich type information about the Linux kernel. Granary has the ability to substitute the execution of a known function with a \emph{wrapped} version of itself. A wrapped function has the same type specification as its unwrapped counterpart and can freely modify its arguments and return value. This feature of Granary exposes type information which is used by behavioural watchpoints.
%\end{enumerate*}

%\noindent
%Behavioural watchpoints have several novel features:
%\begin{enumerate*}
%	\item[i)] {\bf Unrestricted:} A single watchpoint watches many addresses. In \Secref{buffer_overflows}, we use this feature to detect various kinds of buffer overflows.
%	\item[ii)] {\bf Extensible:} Watchpoints are easily extended by programmers. In \Secref{uninitialised_memory}, we use this feature to implement selective memory shadowing and detect reads from uninitialised memory.
%	\item[iii)] {\bf Type-specific:} When a watchpoint is added to an address, the address's type determines which set of behaviours to invoke when (displacements of) that address are dereferenced. In \Secref{access_policies}, we use this feature to detect Linux kernel rootkits by implementing fine-grained memory access policies.
%\end{enumerate*}

\section{Design}
A behavioural watchpoint is a software-based watchpoint that triggers the invocation of a function when any watched memory is accessed. Unlike most software-based watchpoints, a behavioural watchpoint watches a \emph{range} of addresses. This design supports our goal of allowing one watchpoint to monitor all memory accesses to a single object.

To achieve this design goal, we implemented watchpoints by adding an extra level of indirection to memory addresses. An unwatched address is converted into a watched address by changing its high-order bits. These high-order bits indirectly identify context-specific information about the range of memory being watched. This information, called the watchpoint's \emph{descriptor}, contains the originating watched address and a set of functions to invoke when watched memory is dereferenced. Since watchpoint information is embedded in only the high-order bits, a typical offset of a watched address is another watched address that shares the same watchpoint-specific high-order bits. A watchpoint's descriptor is indirectly located by interpreting these high-order bits as an index into a global \emph{watchpoint descriptor table} (\Figref{watchpoint_descriptor_table}). 
%Because these bits do not change, many addresses end up indirectly referring to the same watchpoint descriptor.

%\subsection{Design Implications}
Our approach has the following design implications.

%Any unwatched address can be converted into an watched address. The extra level of indirection added by watching an address allows us to efficiently locate context-specific information about the memory being watched.

%A watchpoint is added to a program by changing the high-order order bits of an address used by the program. 
%Watchpoints work on ranges instead of individual units of memory because a watchpoint is added to a program by changing

\paragraph{A watched address must be distinguishable from an unwatched address.} All watched addresses are non-canonical addresses that trigger a hardware exception when dereferenced. We take advantage of the x86-64 architecture for implementing watched addresses. In kernel space on x86-64, canonical addresses have their 16 high-order bits set to 1. Watched addresses do not take this form. To prevent hardware exceptions, instrumentation is dynamically added to memory loads and stores. Watched addresses are detected before they are dereferenced and resolved to their unwatched counterparts (by masking the 16 high-order bits to 1). An alternative approach is to dereference a watched address and implement behavioural watchpoints in the trap handler.

\paragraph{Multiple watchpoints can watch the same range of memory.} Two copies of an address (e.g., two pointers to the same object) can be separately watched, so long as the high-order bits index different entries in the watchpoint descriptor table. This is useful for distinguishing between logically different objects that occupy the same memory. For example, this feature enables efficient detection of use-after-free bugs without preventing deallocated memory from being immediately reallocated for another use. This efficiency comes from our ability to have one watchpoint for the freed memory, and another watchpoint for newly allocated memory occupying the same space.

\paragraph{Watchpoints are viral.} If an address $A$ is watched, then every address derived from $A$ (e.g. through copying or offsetting) is also watched. This is useful for memory and taint analysis tools: a watchpoint that is added early on in the lifetime of an address (e.g., immediately before the address of some newly allocated memory is returned from an allocator) can persist and propagate until no more derived addresses exist.

\begin{figure}[t]
\begin{center}
\epsfig{file=watchpoints.eps}
\end{center}
\caption{\label{fig:watchpoint_descriptor_table}A watched address (bottom left) and its corresponding unwatched address (top left) are compared. The process of resolving the watchpoint descriptor for the watched address is shown.}
\end{figure}

\paragraph{Millions of watchpoints are supported.} Our design as described uses 15 of the 16 high-order bits (called the \emph{partial index}) of a watched address for identifying a watchpoint's descriptor. 15 bits only allows 32K watchpoints. To increase the number of watchpoints, we use an additional 8 bits (bits 20-27, called the \emph{counter index}) in the address to index into the watchpoint descriptor table (\Figref{watchpoint_descriptor_table}). This counter index extends the number of possible watchpoints to 8M and is left unchanged when converting an unwatched address into a watched address. The key advantage of our watchpoint scheme is the ability to directly map watched addresses to unwatched addresses using a simple bitmask. The main drawback of the scheme is that an offset of a watched address can cause the low-order bits to overflow into the counter index. We are exploring several solutions to this problem.

%This approach has some drawbacks when an offset of a watched address causes the low-order bits to overflow into the counter index; however, we have several solutions to this problem.

\paragraph{Watchpoint descriptors can be arbitrarily customized.} This aspect of watchpoints is possible because our design separates the allocation/management of descriptors and the addresses that they watch. Arbitrary extension of descriptors supports our goal of overcoming the incongruency between the needs of debugging an analysis tools (contextual information about watched memory), and how existing software implements watchpoints (watched memory is opaque).

%\subsection{Extensions}
%Our design includes the following extensions to watchpoints, which expand on the behavioural aspect of our watchpoint implementation.
%principally enable the \emph{behavioural} aspect of our software-based watchpoint .
%Our goal of using watchpoints to watch the memory of objects required the following

\paragraph{Watchpoints are type-specific.} When a watchpoint is added to an address, the \emph{type} of the address determines what meta-information is included in the descriptor, as well as what functions to invoke when memory watched by the watchpoint is accessed. Two powerful applications of this extension are discussed in \Secref{type_overflow} and \Secref{access_policies}.

\paragraph{Triggered watchpoint functions are polymorphic.} The function invoked when watched memory is accessed is decided using a descriptor-specific virtual table (vtable). Each vtable provides eight functions: four read and four write functions. Each function is specific to a memory operand size (1, 2, 4, or 8 bytes). A watchpoint descriptor is initialised with either a generic or a type-specific vtable. Type-specific vtables are specific to the \emph{type} of the watched address. Behavioural watchpoints earn their name from votable because they allow watchpoints to behave differently when watched memory is accessed.

%When invoked, a vtable function operates on the watched address and its descriptor. Behavioural watchpoints earn their name from their ability to behave differently based on the meta-information stored in the descriptor and the (type-specific) vtable function invoked.

\paragraph{Watchpoints remember their originating address.} The address to which a watchpoint is first added is called its \emph{base address}, and is stored in the watchpoint descriptor. We designed watchpoints to remember their base address because it helps to ``anchor" contextual information. When a watchpoint is added, \emph{something} is known about the watched address. Later in a program's execution, an offset of the watched address might be dereferenced. Little can be said about the dereferenced address in relation to the watchpoint's originating address without knowing the originating address.

%Without the former context of why or where the watchpoint was originally added, there is little that can be said about the relation between the triggered address and 
%the watched memory except that it might be arbitrarily far away from the address to which the watchpoint was originally added.
%That is, something is known about an address when the decision to add a watchpoint to that address is made. 
%If the original address is not remembered, then it is difficult to relate a triggered watchpoint
%We designed watchpoints this way so that at any point during the lifetime of a watchpoint, 
%Watchpoints were designed this way so that contextual information is always ``anchored" to something that was once known.

%This is consistent with our goal of using watchpoints to watch the memory of an object because we expect the base address to be the address in memory of a watched object.

%Watchpoint desc

%Because watchpoints are added to addresses, and

%This means that

%We can change an address into a \emph{watched address} by ensuring that a watched address is non-canonical: it cannot legally be used (on x86) without triggering a hardware exception.

%changing part of an address, not by 
%The watched range is \emph{anchored} on the address on which the watchpoint is initially added (called the base address).

%Behavioural watchpoints are implemented by changing an address to-be-watched into a non-canonical address\footnote{In kernel space on x86, a canonical virtual address has its 16 high-order bits set to 1. Current x86 processors require that the 8 high-order bits of an address match the $9^{th}$ highest order bit.}. The translation from unwatched to watched alters the unused high-order bits of an address.

%There are several implications of this design decision:
%\begin{enumerate}
%	\item 
%\end{enumerate}

%Behavioural watchpoints were designed with the following goals in mind:
%\begin{enumerate}
%	\item 
%\end{enumerate}

%\subsection{Architecture}

%The implementation of behavioural watchpoints distinguishes between watched addresses and their descriptors.

%A \emph{watched address} is a pointer with an index into the \emph{watchpoint descriptor table} embedded in its bits (\Figref{watchpoint_descriptor_table}). The $23$-bit index into the descriptor table is formed by concatenating bits $[20,27]$ (called the \emph{counter index}) with bits $[48,62]$ (called the \emph{partial index}). A watched address and its unwatched counterpart share the same counter index; however, the partial index of a watched address varies\footnote{This feature of watchpoints allows for a one-to-one mapping between a watched and unwatched address, and a one-to-many mapping between a watchpoint and all addresses watched by that watchpoint. The one-to-one mapping is formed by masking the high-order bits containing the partial index.}. Partial indexes are recorded in the \emph{partial index counter table}. When a watchpoint is allocated, the current value stored in the counter table for that specific counter index is incremented and returned as the watchpoint's partial index. This allocation strategy allows for at most $2^{15}$ watchpoints per counter index or megabyte of memory\footnote{x86 has byte-addressable memory. The counter index begins at bit $20$, giving each watchpoint $1$ MB = $2^{20}$ B degrees of freedom. However, if an over/underflow across the 1 MB-aligned boundary occurs then the counter index will be corrupted. We can correct one bit of corruption by requiring that the counter index and the partial index have the same sign. An alternative solution is uses a different indexing scheme. We have successfully experimented with a different indexing scheme that solves the aforementioned overflow errors, but sacrifices the one-to-one relationship between a watched and unwatched address.}.

%A \emph{watchpoint descriptor} is a data structure containing a base address, a pointer to a virtual table (vtable) of memory operations, and programmer-defined meta-information.

\section{Implementation}
%Granary is a DBT framework that is under development. Its primary novelty is its ability to efficiently instrument arbitrary, binary Linux kernel modules without imposing overhead on the kernel. 
We implemented behavioural watchpoints using Granary: an extension of the DynamoRIO Kernel dynamic binary translation (DBT) framework \cite{DynamoRIOKernel,GranaryAtOSDI}.  Granary instruments arbitrary, binary Linux kernel modules efficiently and without imposing overhead when the core kernel is running. We plan to use Granary to analyse and debugg kernel modules, which are a frequent source of bugs and vulnerabilities in operating systems \cite{BGI,LXFI}.

Granary is unique among DBT systems because it understands and uses program type information. For example, Granary can substitute the execution of a function with a \emph{wrapped} version of itself. A wrapped function has the same type specification as its unwrapped counterpart and can freely modify its arguments and return value. Granary can wrap some module functions in this way, even if the module source code is unavailable.

While Granary provides a framework for instrumenting kernel modules, we found it was hard to write powerful instrumentation code using low-level DBT abstractions, which motivated the design of behavioural watchpoints. Next, we describe examples of watchpoint-based debugging applications that we have applied to kernel modules.

%We are developing Granary as a framework for building kernel module analysis and debugging tools. Analysing kernel modules is important because they extend the functionality and bug count of operating systems [TODO CITATION]. However, analysing module behavior is challenging. Static analysis of module source code is difficult because of the tight interaction between modules and the kernel. Some modules, however, are only distributed in a binary format, which can make static analysis intractable (on x86).

%A watchpoint is added to a previously unwatched pointer by adding an entry to the watchpoint descriptor table and embedding part of the index of the entry in the high-order bits of the pointer. The original value of the pointer is saved as the base address of the descriptor. A watchpoint is removed from a pointer by masking the high-order bits containing the partial descriptor index. This approach permits a one-to-one mapping between a watched address and its unwatched counterpart, while allowing a single descriptor to watch many addresses. Specifically, a single watchpoint can watch up to $2^{20}$ distinct addresses, and each megabyte of memory can be watched by up to $2^{16}$ distinct watchpoints.

%Watched addresses, however, do not reference valid memory. Dereferencing a watched address triggers a hardware exception. To avoid this, we use Granary to dynamically translate memory loads and stores to first check for and then resolve watched addresses to their unwatched counterparts. If a watched address is detected then a watchpoint-specific vtable function is invoked. Both the address of the watchpoint descriptor and the watched address are passed as arguments to the invoked function. Finally, the original memory instruction is emulated by one that uses the unwatched address. The choice of which vtable function to invoke depends on the size of the memory operand (1, 2, 4, or 8 bytes) and whether the operation reads or writes its operand from/to memory.

\section{Applications\label{sec:applications}}
The following sub-sections describe three applications of behavioural watchpoints and their implementations. In these sub-sections, we use the term \emph{object} to refer to a range of memory locations that are allocated together as a unit.

%to describe a value or aggregate of values stored in memory and referenced by a memory address.

\subsection{Buffer Overflows \label{sec:buffer_overflows}}
A buffer overflow occurs when a program--in an attempt to write to some object's memory--actually writes to adjacent memory cells. One method of detecting buffer overflows relies on the compiler to allocate ``poisoned" regions of memory around each object \cite{AddressSanitizer}. Small overflows (e.g., off-by-one errors) are detected by this approach because they access poisoned memory. Big overflows that ``skip" over poisoned memory and access nearby objects in memory are not detected.

Our insight is that unrelated objects will have different base addresses (i.e. the address of one object will not be derived from the base address of another), and therefore a watched object can be adjacent to a different un/watched object without conflict.

%to an unwatched one, or one with a distinct watchpoint, without conflict.
%\emph{addresses} stored in hardware registers, like high-level programming language variable names, are typically not used to simultaneously access two unrelated objects.
%Accessing poisoned memory is an error because it does not exist as a named program entity. However, 
%Our approach is similar in that memory adjacent to a watched object is considered poisoned; however, we do not rely on compiler support, nor do we allocate or mark poisoned memory as such.

We employ three overflow detection policies: heap-based, type-based, and stack-based. % All three policies depend on the same extension to the meta-information of watchpoint descriptors: a \emph{limit address}. Together, the base address (stored in the descriptor) and the limit address delineate an object's boundaries in memory \cite{BccFatPointers}.

\paragraph{Heap-based overflow detection \label{sec:heap_overflow}}
\begin{figure}
\begin{lstlisting}[language=C,basicstyle=\footnotesize\ttfamily]
FUNC_WRAPPER(__kmalloc, (size, flags), {
  void *addr = __kmalloc(size, flags);
  ADD_WATCHPOINT(addr, size);
  return addr;
})
\end{lstlisting}
\caption{\label{fig:kmalloc_wrapper}Definition of the \texttt{\_\_kmalloc} function wrapper in Granary. The above code expands into a function for wrapping the \texttt{\_\_kmalloc} kernel memory allocator. Calls to \texttt{\_\_kmalloc} are transparently substituted with calls to the generated wrapper. The wrapper invokes the original \texttt{\_\_kmalloc} function and returns a watched version of the allocated address.}
\end{figure}

The heap-based detection policy detects buffer overflow errors on all heap-allocated objects. We use Granary to wrap the kernel's memory allocator functions (e.g., \texttt{kmalloc}) and add watchpoints to the addresses returned by those allocators (\Figref{kmalloc_wrapper}). The lifetime of a watchpoint added in this way is tied to the lifetime of the memory it watches. Each watchpoint's descriptor records bounds information about the allocated memory in the form of the object's base and limit address \cite{BccFatPointers}. A buffer overflow is detected when a dereference of a watched address occurs outside of the bounds recorded by the watchpoint's descriptor (\Figref{detect_overflow}). The memory operand size-specific feature of vtable functions is helpful in catching corner cases where memory reads or writes access both an object and its adjacent memory cells, as shown in case $(3)$ of \Figref{detect_overflow}.

%intersect an object's memory and the adjacent memory cells.
%The vtable functions used by these watchpoints detects buffer overflows when a dereference of a watched address 
%memory dereferenced in a read or write falls outside of the bounds
%We use Granary to substitute invocations of the kernel's memory allocators (e.g. \texttt{kmalloc}) with wrapped versions of the allocators (\Figref{kmalloc_wrapper}). A wrapped allocator invokes the original allocator but returns a watched version of the allocated address. The lifetime of the watchpoint is 
%The watched address returned has its descriptor initialised with the base address as the allocated address and with the limit address as the base address plus the requested allocation size. All buffer overflow detecting watchpoints are initialised with the same vtable pointer.
%The vtable pointer used by all buffer overflow detecting watchpoints points to a table of memory operand size-specific functions. The operand size is necessary to detect dereferences of memory that overlap both the object and its adjacent memory, as illustrated in case \emph{(3)} of . The vtable function corresponding to the memory operand size is invoked when a watched address is dereferenced. Each vtable function is programmed to check the watched address against the base and limit addresses stored in watchpoint's descriptor (\Figref{detect_overflow}). If a buffer underflow or overflow is detected then the vtable function notifies the run-time system.

\begin{figure}
\abovedisplayskip=0pt
\belowdisplayskip=0pt
\begin{center}
	\epsfig{file=overflow.eps}
\end{center}
\iffalse
	\abovedisplayskip=0pt
	\belowdisplayskip=0pt
	\begin{align*}
		A_{base}  & > {A_{unwatched}} \tag{Underflow} \\
		A_{limit} & \leq {A_{unwatched}} \tag{Overflow} \\
		A_{limit} & < {A_{unwatched}} + S_{op} \tag{Overlap}
	\end{align*}
\fi
\caption{\label{fig:detect_overflow}Three common buffer overflow cases are illustrated above: 1) underflow 2) overflow, and 3) overlap. Beside each illustration is the policy that detects the specific bug. When a watched address ($A_{watched}$) is dereferenced, a vtable function that is specific to the memory operand size ($S_{op}$) is invoked. This function detects a buffer overflow if the intended referenced memory address ($A_{unwatched}$) does not fall within the boundaries delineated by the base and the limit addresses ($A_{base}$ and $A_{limit}$, respectively).}
\end{figure}

\paragraph{Type-based overflow detection\label{sec:type_overflow}}
Type-based overflow detection is applied when the type of an object is known. It is distinguished from heap-based overflow detection in that it can apply both to heap and non-heap memory, and it can be used to inform the run-time system about more accurate bounds information. In all other respects, the type-based overflow detection policy detects bugs in the same way as the heap-based policy (as might be encoded within the fields of a vector-like data structure).

% If the type of an object is known then it can reveal semantic relationships which can enable further type- or heap-based overflow detection. For example, the fields within a structure might encode memory bounds information (as would be the case for a vector-like data structure)

%The type-based detection policy is applied when the type of a pointer is known (i.e. not \texttt{void*}). The simplest use case of type information is to infer the limit address of a typed pointer. A more complex use case arises when the fields within a structure encode memory bounds information. In both cases, the vtable functions for typed and untyped memory operate in the same way.

The type of an addressable object becomes known to Granary in the context of a wrapped function. The run-time system performs a type-specific initialisation of a watchpoint's descriptor when a watchpoint is added to a pointer/address with a known type. Type-specific descriptor initialisation is a powerful feature because it allows for new watchpoints to be lazily added (thus propagating our detecting capabilities) to newly discovered objects referenced \emph{within} an object of a known type. We automatically propagate type-based watchpoints using type specifications generated from parsing \texttt{C} header files. Some manual post-processing is required if one is to inform the system of semantic relationships between function arguments or structure fields.

%use a type-specific vtable instead of a generic vtable when watching a typed pointer. Granary includes a database of

%Granary recursively follows argument and return value pointers based on object type specifications. These specifications tell Granary how to follow pointers and what--if any--pointers should be converted to watchpoints. To reduce programmer effort, Granary automatically generates type specifications and wrapper functions for any set of known types and functions. Some manual post-processing is required if the analysis/debugging application needs to understand semantic relationships between function arguments or structure fields. \comment{Automating this process was an important goal of Granary because of Granary wraps all (approx. 6,000) exported kernel functions.}

\paragraph{Stack-based overflow detection}
%We detect overflows on stack allocated objects when memory outside of the bounds of the activation/call frame in which the object is allocated is accessed. This policy is distinguished from the heap- and type-based policies in that the watchpoint descriptor is managed separately from the watched memory, and the lifetime of the descriptor extends beyond the lifetime of the watched memory.

To detect stack overflows, we view the memory occupied by the activation frame of an invoked function as a dynamically-sized buffer. Like our other buffer overflow policies, we use a watchpoint to detect accesses to memory outside of this buffer. Unlike our other policies, we only \emph{associate} a descriptor with this buffer, and rely on a different mechanism to \emph{add} the watchpoint to stack addresses.

%, and manage what addresses are watched with this descriptor using a separate mechanism.

We separate adding watchpoints from allocating descriptors for stack overflow detection. In particular, we associate a descriptor with the buffer represented by the activation frame of a called function. This descriptor tracks the bounds of the frame over the lifetime of the function call. We update the recorded bounds of the frame (in the descriptor) when the frame grows or shrinks. When a function returns, the descriptor's bounds shrink to zero, but the descriptor remains allocated.

%We add an activation frame's watchpoint to a stack address under two circumstances.

Separately managing descriptors allows us to optimize for the two most common sources of stack overflows and ignore instructions that are likely to be safe. First, if we see an instruction that copies the stack or frame pointers, then we assume that the copied address can escape the function. A stack address escaping a function is a potential stack-overflow risk. Adding the frame's watchpoint to this address \emph{taints} the copied address. Future copies or displacements of the watched address implicitly propagate its taintedness because offsets of a watched address reference the same descriptor. A dereference of an escaped pointer--even one happening after the function has returned--is detected as an overflow because the watchpoint descriptor remains live. Second, if we see an indexed dereference of the stack or frame pointers that uses a dynamically bound index, then we assume that the effective memory address accessed is a potential stack-overflow risk. We translate the dereferencing instruction to add the frame's watchpoint to the effective address before the address is dereferenced.

%and addresses that we want to have watched by 
%Under this lens, we associate a watchpoint descriptor with each function call that tracks the bounds of each activation frame. 
%allocating one watchpoint descriptor for each function call. This descriptor exists to track the bounds of the activation frame over the lifetime of the function call. The run-time system dynamically adds instructions to the instrumented program that change the recorded bounds when the size of the activation frame grows or shrinks. When a function returns, the descriptor's bounds shrink to zero.

%Instrumentation is dynamically added to the program that increase/decrease the recorded size of the activation frame in the descriptor when the 
%The descriptor tracks the size of the activation frame

%Unlike the heap- and type-based overflow policies, there are numerous optimisations 
%Stack-based overflow detection is distinguished from the heap- and type-based policies in that numerous optimisations are possible, and that 
%in an access to memory outside of the object's activation frame.
%The aforementioned approaches to overflow detection 
%Stack-based overflow detection is distinguished from the heap- and type-based overflow detection policies insofar as it uses

%Detecting stack overflows is important because they are commonly used in remote code execution and privilege-escalation attacks against operating systems \cite{SecureProgramExecFlowTracking}.

% more support from Granary's run-time system. At a high level, each activation frame is assigned a dedicated watchpoint when a function is called. The effect of an instruction that grows or shrinks an activation frame's size is reflected by a similar change to the base address\footnote{The run-time call stack on x86 grows into lower memory.\comment{ Modifying the base address instead of the limit address allows us to maintain one set of vtable functions for all three buffer overflow detection policies.}} of the frame's watchpoint descriptor. When a function returns, the descriptor of its activation frame is cleared, but remains allocated\footnote{Reclaiming/reusing allocated but unlikey-to-be-used watchpoints is challenging. One overflow-specific approach is to ensure that the next use of the watchpoint is for a buffer whose bounds do not intersect with the current use's bounds. Another approach is to use some number of context switches as a grace period in which the watchpoint cannot be reused, but after which it can be reused. We leave this as future work.} lest the address of a local variable escape the function\footnote{If the address of a stack-allocated object escapes its function activation frame, then that pointer is said to be \emph{dangling}. Dereferences of dangling pointers can cause stack overflow errors. In the kernel, data stored in unallocated stack memory (e.g. through a dangling pointer) is at risk of being clobbered by interrupt stack frames and by function activation frames.}.

%We assume that a memory instruction operating on a constant displacement of the stack or frame pointer registers is well-behaved. However, if the stack or frame pointer registers participate in a memory instruction and the displacement from the register is dynamically-bound then that instruction is suspect. An instruction that copies the address in the stack or frame pointer registers is also suspect because the copied address might escape the current function or participate in a local stack overflow. All suspect instructions are translated to transparently add the frame's dedicated watchpoint to the dereferenced or copied address.

%This scheme segments the runtime stack into contiguous but non-overlapping regions of watched memory. A straightforward extension excludes saved return addresses and link pointers from the known bounds of activation frames, thus detecting return-oriented buffer overflow attacks.

\subsection{Selective Memory Shadowing\label{sec:uninitialised_memory}}

In this section, we show how to use behavioural watchpoints to shadow memory. Previous work has focused on full memory shadowing \cite{Memcheck}, while watchpoints enable \emph{selective} shadowing of watched objects. We describe how the initialisation state of each byte of watched memory is tracked using shadow memory, and how to detect bugs related to the usage of heap-allocated memory.

We use Granary to wrap kernel memory allocators and deallocators (e.g., \texttt{kmalloc}, \texttt{kfree}). Wrapped allocators add watchpoints to the addresses returned by the kernel's allocators, and wrapped deallocators remove watchpoints before invoking the kernel's deallocators. Selective shadow memory is maintained as a watchpoint descriptor-specific, variable-sized bitset\footnote{As an optimisation, shadow memory for small objects is embedded in the watchpoint descriptor.}. Each byte of allocated memory corresponds to one bit of descriptor-specific shadow memory. The bits in shadow memory are initialised to zero. Individual bits are flipped to one by write-specific vtable functions when there is a write to the bytes of memory shadowed by those bits.

%When a wrapped allocator is invoked, a watchpoint is added to the addresses returned by those allocators

% in shadow memory are initialised to zero.
%Watchpoints are added to all heap-allocated objects 
%the meta-information within each watchpoint descriptor.
%By tracking the initialisation state, we detect

%We apply the above shadow memory scheme to detect uses of uninitialised memory for heap-allocated objects. As in \Secref{heap_overflow}, we interpose on the kernel's memory allocators and add watchpoints to the addresses returned by those allocators. When memory is allocated, the watchpoint descriptor is initialised with the number of allocated bytes (available as an argument to the allocator) and with its own shadow memory.

%detect uses of uninitialised memory.  Detecting such uses is important because a program using uninitialised memory can exhibit unintended, non-deterministic behaviour.

%We define uninitialised memory as either allocated memory to which no value has been written, or de-allocated memory. We selectively track the initialisation state of each byte of watched memory using \emph{shadow memory} \cite{Memcheck}. Our implementation represents shadow memory as a bitset, where the state of each bit tracks the initialisation state of a byte of watched memory. Watchpoint descriptors are augmented to include the size (in bytes) of a watched object and a pointer to the memory shadowing the object\footnote{As an optimisation, shadow memory for small objects can be embedded in the watchpoint descriptor.}.

\paragraph{Read-before-write bugs}
Read-specific vtable functions detect read-before-write bugs by checking if the shadow bit corresponding to one of the read bytes is zero. However, this method of detection can report false positives: it is common for larger-than-needed reads to be performed and for compiler-added structure padding to be read (but never written). A relaxed version of our read-before-write memory checking policy requires that at least one shadow bit is set for every read operation.

\paragraph{Memory freeing bugs}
We detect freeing of non-heap memory when the argument to a kernel memory deallocator is not a watched address. We detect an invalid-free when the unwatched address (computed through masking) of a watched address being deallocated is different than the watchpoint descriptor's base address. We detect use-after-free bugs by marking the descriptor of a watched address being deallocated as dead. The lifetime of a dead descriptor extends beyond that of the object so that a later use of any watched address with that descriptor will report a use-after-free bug. We detect double-free bugs when the descriptor of a watched address being deallocated is already marked as dead.

%We use Granary to wrap the kernel's deallocation functions (e.g. \texttt{kfree}). Because all allocators are wrapped and add watchpoints, freeing of non-heap memory is easily detected when

%We detect freeing of non-heap memory

%If the argument to a wrapped deallocator function is a watched address then the shadow memory for the .

%Finally, the original deallocator is invoked with the unwatched address as an argument. Reads and writes to a watched address of a deallocated object fail to pass bounds checks because the object size is zero.

%Two positive side-effects of our approach is that it detects double-free bugs and freeing of non-heap memory bugs. A double-free is detected when a watched address for an already deallocated object is passed as an argument to a deallocator. An attempt to free non-heap memory is detected when an unwatched address is passed as an argument to a deallocator.

\subsection{Field-grained Access Policies\label{sec:access_policies}}

\begin{figure}[t]
\begin{lstlisting}[language=C,basicstyle=\footnotesize\ttfamily]
// Invariant: rtc_fops->ioctl == &rtc_ioctl
WATCH_WRITE(struct file_operations, ioctl, {
  if(&rtc_fops  == base_address
  && &rtc_ioctl != ioctl) {
    // potential attack: prevent an anti-
    // virus scan from being scheduled!
  }
})
\end{lstlisting}
\caption{\label{fig:field_invariant_check}This code example shows how to check invariant 1(h) from \cite{GibraltarKernelInvariants} using Granary's field-accessor API. The invariant checked prevents a (potential) kernel rootkit from installing its own \texttt{ioctl} handler into the Real-Time Clock. Anti-virus programs depend on this built-in \texttt{ioctl} handler to periodically schedule virus scans. Replacing this handler can prevent such scans from being scheduled, thus allowing the rootkit to go undetected.}
\end{figure}

In this section, we briefly describe using watchpoints at the granularity of object fields. Using a field-accessor API, we detect Linux kernel rootkits, which is challenging because rootkits actively try to hide their existence. However, rootkits often leave hints of their existence in the form of violated data structure invariants \cite{GibraltarKernelInvariants,OSck}.

\Figref{field_invariant_check} shows an example use of the field-accessor API that checks an invariant on the \texttt{ioctl} field of any watched object whose type is \texttt{struct file\_operations}. A rootkit violating this invariant can prevent an anti-virus program from scheduling periodic scans---scans which might otherwise detect the presence of a rootkit. However, a violation of this invariant is immediately detected by our run-time system because the field-accessor API enables active rather than passive checking of invariants.

The field-accessor API connects to type-specific watchpoint vtables through automatically generated code. Generating this code involves parsing \texttt{C} structure type layout information from the \texttt{C} header files of the Linux kernel.

\section{Evaluation}
%\begin{figure*}[t]
%\begin{multicols}{2}
%{\footnotesize
%\begin{tabularx}{\columnwidth}{l|l|c}
%Benchmark & Benchmark Description & Result \\
%\hline
%Microbenchmark & Tight loop of watched dereferences & 20x \\
%\texttt{rcutorture} & Slowdown of reader threads & 7.5x \\
%IOZone & Slowdown of reader threads & 7.5x \\
%Netperf & Slowdown of reader threads & 7.5x \\
%\end{tabularx}}
%\noindent\epsfig{file=iozone.eps,height=100pt}
%\end{multicols}
%\end{figure*}

We evaluated the performance of behavioural watchpoints using a microbenchmark, Netperf, and IOzone. Our tests ran on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ i7-860 2.80 GHz CPU, 8GB memory, and an Intel 82578DM Gigabit Ethernet card. Watchpoint instrumentation was enabled for every memory load and store, and buffer overflow detecting watchpoints were added to all memory returned from kernel heap allocators.

%We compare native execution, passive execution, and active execution to evaluate the performance of our watchpoint implementation. Passive and active execution both run the Granary DBT framework; however, passive execution only performs basic translation of module code, while active execution translates and adds watchpoint-specific instrumentation to module code. Our choice of network and filesystem modules depend on the fact that they are

\paragraph{Microbenchmark} The microbenchmark performed a tight-loop of memory operations and exhibits the worst-case performance overhead of watchpoints (about 20{\footnotesize$\times$}). This is expected because our instrumentation adds roughly 13 instructions to each memory load and store.

%\paragraph{\texttt{rcutorture}} We tested the \texttt{rcutorture} kernel module. It exhibited interesting behaviour: all threads experienced a minimum of 6{\footnotesize$\times$} overhead; however, readers were abnormally penalized and tended to see more stale data. We think this is because each writer synchronized with fewer readers.

\paragraph*{Netperf} We tested network throughput and latency using the \texttt{e1000e} network device driver module. CPU utilization increased by {\texttildelow}6\%, UDP throughput decreased by at most 50\%, TCP throughput was unaffected, and both the TCP and UDP request-response latency increased by no more than 50\%.

\paragraph*{IOzone} We tested the throughput of disk read and write operations with the \texttt{ext3} file system module, a file size of 512Kb written in 4Kb records, and with a single reader/writer thread in each case. The throughput for reads and writes decreased by  {\texttildelow}20\% and {\texttildelow}80\%, respectively.

%the throughput of file reads by a single reader thread decreases by , and the throughput of file writes by a single writer thread decreases by {\texttildelow}80\%.
%Our testing environment includes running our system with active (actively looking for watchpoint) and passive (null instrumentation) mode.
%On the microbenchmark, passive instrumentation has a 4\% overhead and active instrumentation has a 20x overhead. For the \texttt{rcutorture} module, 
%the overhead of passive instrumentation is 4\% and the overhead of active instrumentation is 20x
%under a worst-case scenario with a microbenc
%Our evaluation of the performance of behavioural watchpoints

\section{Related Work}
%Watchpoints are an important debugging facility that help developers track accesses to memory. Most state-of-the-art processors provide some support for hardware watchpoints, but such watchpoints are usually scarce or require specialised hardware support \cite{Mondrix,UnlimitedWatchpoints}.

\paragraph{Hardware-based}
Greathouse \emph{et al.}~\cite{UnlimitedWatchpoints} propose a hardware solution that efficiently supports an unlimited number of watchpoints. Witchel and Asanovic \cite{Mondrix} describes the implementation of memory protection domains for the Linux kernel. Protection domains are implemented using specialised hardware and enable fine- and coarse-grained memory protection using a mechanism similar to hardware watchpoints. Unlike our approach, both \cite{Mondrix} and \cite{UnlimitedWatchpoints} depend on specialised hardware and require that applications using this hardware  separately maintain context-specific information. Suh \emph{et al.} \cite{SecureProgramExecFlowTracking} proposes a method of secure program execution by tracking dynamic information flow. Memory tagging at the hardware level allows their system to track tainted data as it propagates through a running program. Behavioural watchpoints are similar insofar as a watched address is tagged, and this tag propagates through a program.

%makes the case for supporting an unlimited number of watchpoints. A hardware solution is proposed and multiple applications are described. Unlike our approach, the cited approach depends on specialised hardware and requires that applications using these watchpoints maintain their own context-specific information.

%In ,

\paragraph{Software-based}
Zhao \emph{et al.} \cite{DynamoRIOWatchpoints} describe a method of implementing an efficient and scalable DBT-based watchpoint system. His method uses page protection and indirection through a hash table to track watched memory. This approach does not support watching ranges of memory, nor does it support context-specific information. Lueck \emph{et al.} \cite{PinADX} introduce semantic watchpoints as part of the PinADX system, an extension of the PIN DBT framework. PinADX enables interactive debugging by triggering debugger breakpoints when semantic conditions are met. While similar in spirit to behavioural watchpoints, semantic watchpoints do not maintain context-specific, per-watchpoint state.

%In this respect, a semantic watchpoint is similar to a behavioural watchpoint whose vtable functions trap when a semantic condition is met.
% using DynamoRIO, a popular user-space DBT framework. The method described is very efficient, but unlike behavioural watchpoints, views memory as an opaque sequence of bytes. 
%There have been several proposals on implementing watchpoints in software. 
%Zhao introduced a method for supporting millions of software-based watchpoints using DynamoRIO, a popular user-space DBT framework \cite{DynamoRIOWatchpoints}. Their method is very efficient and does not require inspecting all memory. 
%Our approach differs from theirs in that we e
%Our approach differs from them in the sense that our watchpoint contains the contextual information which can be used for debugging. 
%In , 
% can add a data breakpoint at an address and any access of the gets checked. Unlike this our approach of behavioral watchpoint watches the range of addresses and a watchpoint monitors the access of an object.  Our approach of watchpoint is also viral and any address derived from the watched addresses will also get watched. Our watchpoint implementation bears a resemblance with {cite dynamic information flow} however we don’t require hardware support to efficiently store the tag else it gets embedded with the address itself.

\section{Conclusions and Future Work}
%We created behavioural watchpoints with the goal of simplifying the implementation of program analysis and debugging tools. We implemented behavioural watchpoints in software because hardware approaches to watchpoints do not meet the needs of program analysis tools. We identified an incongruency between how existing software implements watchpoints and how program analysis tools use watchpoints. Finally, we overcame this incongruency by developing a new form of watchpoint that maintains its own context-specific information---information that is required for large-scale program analysis. We demonstrated the usability of behavioural watchpoints by describing the implementation of three memory error detection and protection policies.

We created behavioural watchpoints with the goal of simplifying the implementation of program analysis and debugging tools. Behavioural watchpoints are a new form of software watchpoint, which address both the scalability limits of hardware watchpoints and the lack of contextual information with traditional software watchpoints.  The key innovation is to maintain context-specific information about the memory being watched---information that is required for large-scale program analysis---with the watchpoint itself.  Thus, behavioural watchpoints resolve the incongruency between how existing software implements watchpoints and how program analysis tools use watchpoints. We demonstrated the usability of behavioural watchpoints by describing the implementation of three memory error detection and protection policies.  Although our evaluation is limited at this stage, the overheads of behavioural watchpoints, as implemented in Granary, are quite reasonable for the target use of debugging kernel modules.

As future work, we will investigate the applicability of behavioural watchpoints as a means of isolating and enforcing control-flow integrity policies on Linux kernel modules.

%Behavioural watchpoints make it easier to implement program analysis and

%overcome the incongruency

%We have established that behavioural watchpoints are useful for implementing memory error detection and protection policies. In summary, we conclude.

% The bibliography should be embedded for final submission.
\bibliographystyle{acm}
\bibliography{library}
\end{document}

### 2013-06-19 OSDI'12/paper.tex
Commit of some old papers and in-progress PLOS.

\documentclass[10pt,preprint]{sigplanconf}
\usepackage{amsmath,epsfig,subfigure,multirow}

\usepackage{graphicx}
\usepackage{refstyle}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{framed}
\usepackage{lipsum}
\usepackage{multicol}
\usepackage[english]{babel}
\usepackage{numprint}
\usepackage[dvipsnames]{color}
\definecolor{shadecolor}{RGB}{250,250,250}

\begin{document}

\conferenceinfo{OSDI '12}{Feb. 23, Shenzhen.} 
\copyrightyear{2012} 
\copyrightdata{[to be supplied]}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Granary: Comprehensive Kernel Module Instrumentation}
\subtitle{}

\authorinfo{Peter Goodman\and Angela Demke Brown}
           {Department of Computer Science\\ University of Toronto}
           {\{pag, demke\}@cs.toronto.edu}

\authorinfo{Akshay Kumar\and Ashvin Goel}
           {Department of Electrical and Computer Engineering\\ University of Toronto}
           {akshayk@cs.toronto.edu\and ashvin@eecg.toronto.edu}

\maketitle

\renewcommand{\secref}[1]{\S\ref{sec:#1}}

% until we find a name for our system, the following commands should suffice :-)
\newcommand{\sysname}{Granary}
\newcommand{\Sysname}{Granary}

\newcommand{\note}[1]{}

\begin{abstract}

Kernel modules extend the functionality of operating systems. Modules are used to support new devices (e.g. network and graphics cards) and provide new features (e.g. file systems). These kernel modules often available as binary, executes and interacts with the kernel in a complex and dynamic environment. This makes the undertanding of module bahviour challenging. Static analysis is one of the method to understand their behaviour but this left many question unanswered. This makes it important to analyse the bahaviour of these modules dynamically.

This work presents {\sysname}, a framework for efficiently instrumenting Linux kernel modules. {\sysname} uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules. The goal of comprehensivness requires that {\sysname} control the execution of all module code. This is achieved by attching instrumentation when module code first gets executed and leaving it attached until the next context switch to kernel code. The second goal requires that any overhead introduced by {\sysname} only affect instrumented modules. The {\sysname} introduces overhead when it translated, caches, and executes instrumented code leaving kernel execution unaffected. The attaching and detaching of the instrumentaton framework causes the execution to switch between native and instrumented mode. This makes it challenging to maintain the module isolation and consistency of the interrupt and exceptions.

TODO : write a short description about the interrupt handling as an improvement over DRK.

Interposing on the control-flow transfers that leads to mode switching is easy when module executes under the control of {\sysname} but is challenging when kernel code executes. Page protection is one mechanism that can be used to detect the control-flow transfer to the modules, where all modules code can be made non-executable and any attempt to execute the module code results in a page fault execption that will be handled by the {\sysname } to attach the instrumentation. Considering the associated cost of page-fault handling, {\sysname} takes an alternative approach where it interposes during module loading changing the module entry point used by the kernel, pointing to the shadow module, allowing for efficient entry point interposition. The system also introduces the novel approach of type based kernel wrapper which understand the function arguments passed to the kernel and proactively wraps all the associated function pointers of the raw module pointing to the new module entry point.

Kernel wrapper recursively wrap arguments to kernel functions. However, these arguments can be deeply nested (e.g file-system module) and rewrapping them each time they cross module-kernel boundry is expensive. {\sysname} solves the problem of recursive wrapping on each kernel function call by using the taint tracking method where it wraps then on first call of the function call and stores an embedded hash map for the wrapped structure. on the subsequent calls the recursive wrapping only gets enabled when the structure value gets tainted by the module or on the fall-through where the kernel tries to execute the non-executed kernel code.

Using the same method, the kernel wrapper also introduces the dynamic wrapping mechanism where each kernel object passed over the module-kernel interface will have its own wrapper, providing a more finer control over analyzing the objects passed over the module-kernel interface.

{\Sysname} is a work in progress. It works on multi-core processors with pre-emtive kernels, and has modest overhead of 10\% for network drivers and 2x for in-RAM file systems. We have used {\sysname} to isolate and comprehensively instrument several network device drivers (e1000, e1000e, ixgbe, tg3) and file system modules (ext2, ext3, btrfs). We have used {\sysname} to develope an application which enforces partial control-flow integrity policies. These policies disallow modules from executing dangerous control-flow transfers. As a future work, we plan to implement more optimizations and develope applications which helps in analysing the bahaviour and debugging the module and to protect the kernel from its malicious behaviour.

\section {Rest}
We have applied Granary and enforced some control-flow integrity policies. These policies
disallow modules from executing dangerous control-flow transfers. As
future work, we plan to implement more optimizations and more
applications.
{\Sysname} is a work-in-progress. We have used {\sysname} to comprehensively instrument several network device drivers (e1000, e1000e, ixgbe, tg3) and file system modules (ext2, ext3, btrfs).

We have applied {\sysname} and enforced some control-flow integrity policies. These policies disallow modules from executing dangerous control-flow transfers. As future work, we plan to implement more optimizations and develope mo
TODO

uses the taint tracking method to restrict the recusive wrapping of the kernel functions argument only if gets tainted after the first wrap. this

Kernel wrappers recursively wrap arguments to kernel functions. However, these arguments can be deeply nested and linked structures (e.g. file system inodes). Continually re-wrapping deep structures each time they cross the module-kernel boundary is expensive.

A partial solution to this problem limits the depth of wrapping. In the worst case, some raw module addresses escape wrapping. If invoked by the kernel, a raw module address will trigger a memory protection exception and be handled by {\sysname}. {\Sysname} re-directs control to the shadow module, thus maintaining isolation. This solution alone is undesirable as it does not correct the root problem.

Ideally, every kernel object would contain and embedded hash. If this were the case, then shallow changes to an object would be detected when a wrapper compares the object's embedded hash against a re-computed hash. Again, deep changes might escape detection, but not break control-flow isolation. If the kernel invokes a raw module address, then {\sysname} would take control as before but change the seed of the hash function, thus invalidating every previously wrapped data-structure and correcting the root problem.

Unfortunately, embedding hashes into kernel objects is not practical, and overwriting arbitrary object fields for storing the hash can result in data loss and corruption. However, one can change a function pointer, so long as the changed function pointer behaves as before. The restriction to function pointers is not so bad: many kernel objects reference shared virtual tables (vtables) of function pointers, and continually re-wrapping these vtables is wasteful.

We track changes to objects containing at least one function pointer by designating a tracking function pointer (TFP), as shown in \Figref{tracker}. TFPs are changed to address dynamically generated code that \texttt{JMP} to the TFP's original target address, thus maintaining the same control-flow behaviour. The hash of the object is located beside the \texttt{JMP} instruction in memory. Checking for changes to an object is analogous to comparing agaisnt an embedded hash, but instead compares against the TFP hash. As before, breaches in control-flow isolation are handled by {\sysname} and invalidate all previously computed hashes.

TODO
TODO

we can achieve this goal by attaching instrumentation when some module code is first executed and leaving instrumentation attach until the next context switch. However, our other second goal requires tht any overhead introduced by {\sysname} only affect instrumentation modules. {\sysname} introduces overhead when it translates, caches, and executes instrumented code. The interactions between first two goal forces code to execute in one of the two modes: native and instrumented. Under this lens, module code executes in instrumented code and the module code instrumeted mode and kernel code executes it in the native mode. Attaching and detacing {\sysname} causes the execution to switch between native to instrumented and back into native mode.

{\Sysname} dynamically detects when control transfers between the kernel and its module and switches the mode accordingly. Detecting and interposing on control-flow transfers that should leas to mode switches challenging. The obvious way to detect when to switch modes is to inspect every control-flow transfer instructions. This is easy when executing instrumented code as the {\sysname} can detect the control transfer to the kernel and can switch mode.However, interposing on module entry points is non-trivial because {\sysname} has no control when kernel code is executing. Page protection is one mechanism that {\sysname} uses to detect control-flow tranfers to modules. All module code is made non-executable, and any attempt to execute module code results in a page fault exception. Page fault exceptions are first handled by {\sysname}, which chooses to attach instrumentation if the faulted instruction is a module address. Because of their cost, page faults are depended on only as a last resort. Instead, {\sysname} changes the module entry points used by the kernel, allowing for efficient entry point interposition.

TODO

Our extensive use of compile-time meta-programming enables efficient, dynamic analyses that are driven by static type information. Our approach is novel because {\sysname} can instrument \emph{arbitrary}, binary kernel modules without imposing any overhead on non-instrumented kernel code. {\Sysname} requires no special hardware support and requires minimal (under 100 lines) changes to the kernel. {\Sysname} works on multi-core processors with pre-emptible kernels, and has modest overhead of 10\% for network drivers.

Our first goal of comprehensiveness requires that {\sysname} control the execution of all module code. We can achieve this goal by attaching instrumentation when some module code is first executed and leaving instrumentation attached until the next context switch. However, our second goal requires that any overhead introduced by {\sysname} only affect instrumented modules. {\Sysname} introduces overhead when it translates, caches, and executes instrumented code. If {\sysname} is left attached then it would eventually instrument kernel code and fail to meet our second goal. This interaction between our first two goals forces code to execute in one of two modes: native or instrumented. Under this lens, module code executes in instrumented mode and kernel code executes in native mode. Attaching {\sysname} causes a native-to-instrumented mode switch and detaching causes an instrumented-to-native mode switch.

{\Sysname} dynamically detects when control transfers between the kernel and its modules and switches the mode accordingly. Detecting and interposing on control-flow transfers that should lead to mode switches is challenging. The obvious way to detect when to switch modes is to inspect every control-flow transfer instruction. This is easy when executing instrumented code. For example, the dispatcher can switch modes when it detects that control is transferring to the kernel. However, interposing on module entry points is non-trivial because {\sysname} has no control when kernel code is executing. Page protection is one mechanism that {\sysname} uses to detect control-flow tranfers to modules. All module code is made non-executable, and any attempt to execute module code results in a page fault exception. Page fault exceptions are first handled by {\sysname}, which chooses to attach instrumentation if the faulted instruction is a module address. Because of their cost, page faults are depended on only as a last resort. Instead, {\sysname} changes the module entry points used by the kernel, allowing for efficient entry point interposition.

{\Sysname} is a work-in-progress. We have used {\sysname} to comprehensively instrument several network device drivers (e1000, e1000e, ixgbe, tg3) and file system modules (ext2, ext3, btrfs). We have applied {\sysname} and enforced some control-flow integrity policies. These policies disallow modules from executing dangerous control-flow transfers. As future work, we plan to implement more optimizations and more applications.

% {\Sysname} is efficient because of our extensive use of compile-time meta-programming.
% {\Sysname}'s dynamic analyses are driven by kernel type information

%With this tool, we show how to prevent a privilege escalation attack against a proprietary NVidia kernel module. Our technique requires no changes to the module and maintains reasonable performance on graphics-intensive workloads.

%We use {\sysname} to implement basic control-flow integrity checkers and direct memory access monitors for arbitrary modules. With these tools, we show how to prevent a privilege escalation attack against a proprietary NVidia kernel module. Our technique requires no changes to the module and maintains reasonable performance with graphics-intensive workloads.

%Modules extend the default behaviour of operating system (OS) kernels like Linux. In commodity OSes, the kernel and its modules execute in the same environment and freely share sensitive data. This integration makes debugging modules challenging, and makes modules a prime target for security exploits \cite{cve2010}. User space programs benefit from many instrumentation and isolation tools \cite{Ford2008,NaCl,Nethercote2007a}. Unfortunately, no tools exist that can isolate the kernel from arbitrary modules. To address this problem, we created {\sysname}.

%{\Sysname} is a comprehensive framework for fine-grained instrumentation of arbitrary x86-64 binary kernel modules. {\Sysname} automatically, efficiently, and transparently isolates modules; it discovers and enforces module boundaries, and maintains isolation in the presence of interrupts, even on multi-core processors. Our implementation requires minimal changes to the kernel, and uses automatically generated code from high-level type information as a means of re-introducing semantics on to arbitrary module/kernel data and code.

%We use {\sysname} to implement a basic control flow integrity (CFI) checker \cite{Abadi2009}. Our CFI implementation restricts modules to use only kernel-exported functions, enforces some return-address consistency, and performs well on standard performance benchmarks.

%  has been used as a means of restricting the execution of potentially unsafe modules

%and requires minimal changes to the kernel
%The key to \sysname's efficiency is from its automatic detection of module boundaries based on high-level type information.

%\Sysname transparently interposes on all module execution, with no performance overhead when executing non-module kernel code. \Sysname  to automatically discover module boundaries.

%We also show how to efficiently and transparently isolate modules in the presence of interrupts and on multi-core processors.

% Once installed, kernel modules execute in the same environment and on the same data as the kernel.

% This dependence enables several attack vectors. Modern hardware features attempt to protect OS kernels against attack by executing kernel code in a privileged environment. Unfortunately, kernel modules execute in the same environment as the kernel. As a result, a misbehaving/malicious module can easily corrupt the kernel's internal data structures and compromise its integrity.

%Ensuring that arbitrary kernel modules are safe to execute is challenging. Kernel modules are distributed either as source code or as pre-compiled binaries. Proving that a non-trivial kernel module is safe requires expert knowledge in the former case and is intractable in the latter case. Existing approaches attempt to solve this problem by using hardware protection domains, virtualization, or compile time and runtime checks that enforce capabilities/invariants specified in source code. Unfortunately, hardware protection domains lack widespread support, virtualization limits the breadth of modules that can be secured, and source code annotation based approaches require changes to the kernel, to the modules, and to the compiler toolchain.

% Our approach generalizes previous approaches by operating on arbitrary x86-64 binaries without virtualization, hardware support, or compiler/source code modifications. \Sysname dynamically discovers a module's entry and exit points and instruments it's execution without changing the behaviour or state of the kernel. Modules are restricted to using only exported kernel functions, unless other functions have explicitly been made available through the exported API. High-level type and capability information is propagated into instrumented modules starting from the usages of exported kernel functions. This allows for runtime type checking of function pointers, assignment of ownership to memory, API integrity checking, and more.

%Monolithic operating systems (OSes) like Linux depend on kernel modules to extend their supported behaviours and communicate with hardware. This dependence enables several attack vectors. Modern hardware features attempt to protect OS kernels against attack by executing kernel code in a privileged environment. Unfortunately, kernel modules execute in the same environment as the kernel. As a result, a misbehaving/malicious module can easily corrupt the kernel's internal data structures and compromise its integrity.

%Ensuring that arbitrary kernel modules are safe to execute is challenging. Kernel modules are distributed either as source code or as pre-compiled binaries. Proving that a non-trivial kernel module is safe requires expert knowledge in the former case and is intractable in the latter case. Existing approaches attempt to solve this problem by using hardware protection domains, virtualization, or compile time and runtime checks that enforce capabilities/invariants specified in source code. Unfortunately, hardware protection domains lack widespread support, virtualization limits the breadth of modules that can be secured, and source code annotation based approaches require changes to the kernel, to the modules, and to the compiler toolchain.

%Our approach generalizes previous approaches by operating on arbitrary x86-64 binaries without virtualization, hardware support, or compiler/source code modifications. \Sysname dynamically discovers a module's entry and exit points and instruments it's execution without changing the behaviour or state of the kernel. Modules are restricted to using only exported kernel functions, unless other functions have explicitly been made available through the exported API. High-level type and capability information is propagated into instrumented modules starting from the usages of exported kernel functions. This allows for runtime type checking of function pointers, assignment of ownership to memory, API integrity checking, and more.

\end{abstract}

% --------------------------
% --------------------------

\section{Introduction}\label{sec:intro}

TODO

We identified the following goals:
\begin{enumerate}
	\item Comprehensively isolate and instrument all binary modules.
	\item Impose no performance overhead to non-module kernel code, or to user-level code.
	\item Require nothing of modules, and make as few changes to the kernel as possible.
	\item Automate as much as possible.
\end{enumerate}

% TODO: show how related work fails on specific goals, showing that other people didn't fully succeed
% TODO: we don't stop malicious code, but we stop stupid code
% TODO: our goal is not to protect agaisnt nvidia, but to prevent people from misusing nvidia

% goal 1: comprehensive instrumentation of modules
% goal 2: no performance overhead to non-module kernel code, or to user-level code
% goal 3: portability across OS and hardware versions: minimal changes to the kernel
% goal 4: automate as much as possible

%Kernel modules (hereafter modules) extend the functionality of monolithic operating systems (OSes) like Linux. Modules and the kernel are dynamically or statically linked together and execute with unrestricted access to kernel control and non-control data. Usually, modules interact with each other and the kernel through the kernel's module-exported API. In the case of Linux, this API leaks sensitive data and is easily circumvented. Buggy, malicious, or exploited modules that misuse this API or (un)intentionally abuse kernel data structures can compromise the integrity of the kernel. Such compromises are difficult to debug and represent potential security risks.

%That state of module-extensible user space

%Extensible user space programs suffer similar problems; however, they are easier to debug because of the abundance of debugging tools \cite{GDB,Nethercote2007a,Seward2005} and are easier to secure because of the abundance of isolation techniques \cite{Ford2008,NaCl,Amarasinghe2003}.

%Similar to existing research, our approach attempts to maintain the kernel's integrity by isolating/sandboxing modules, but differs in the mechanism of isolation. Isolation  involves confining the execution of module code so that normal behaviour preserved, and illegal behaviour is impossible or detectable (and can thus be stopped). Illegal behaviour is defined as the misuse of the kernel's module-exported API, the use of a non-exported function, or the modification of sensitive kernel control and non-control data.

%Prior research implements module isolation by compile-time and run-time checking/annotations ~\cite{Ganapathy2005,Castro2009,Erlingsson2006,Swift2005}, by hardware protection domains ~\cite{Swift2005,Witchel2005}, by moving modules into user space ~\cite{Ganapathy2005,Williams2008}, and by extending virtual machines (VMs) and hypervisors with introspection capabilities \cite{Xiong2011,Hofmann2011,Srivastava_Giffin_2011}. These approaches achieve many of the goals of this research; however, each suffers from at least one of the following two issues:
%\begin{enumerate}
%	\item Dependence on hardware support, which limits the breadth of modules that can be isolated (virtualization), or lacks widespread adoption by hardware vendors (protection domains).
%	\item Dependence on language, code, or compiler changes, thus limiting the likelihood of their adoption by OS and module vendors, and limiting their ability to secure legacy modules.
%\end{enumerate}

%Our work isolates arbitrary binary modules using dynamic binary instrumentation (DBI) \cite{Bruening2004}. DBI gives fine-grained control over program execution and provides similar isolation guarantees to that of virtualization, without limiting which modules can be instrumented. Like \emph{LXFI} \cite{YMao2011}, \emph{BGI} \cite{Castro2009}, and \emph{XFI} \cite{Erlingsson2006} before us, \sysname enforces data structure and control-flow invariants through annotations. However, our annotations are valid C++, apply only to kernel data structures and APIs, and require no modifications to either the kernel, its modules, or the compiler. Thus, \sysname can secure existing modules, regardless of how or by whom they are distributed.

%We tested \sysname on several 1Gb/s and 10Gb/s network drivers

%and the invariants provided by our annotations are enforced and propagated at runtime.

%Similar to virtualization, our work provides strong isolation guarantees on code execution, but isolates arbitrary binary modules using dynamic binary instrumentation (DBI) \cite{Bruening2004}.

%Our solution attempts to maintain the integrity of the Linux kernel from untrusted binary modules with minimal performance overhead to the kernel proper. \Sysname uses a modified version of the DynamoRIO Kernel (DRK) dynamic binary instrumentation (DBI) framework to achieve this goal ~\cite{Feiner2012,Feiner2010b}. The choice of DRK as a host platform in \sysname is motivated by its ability to provide fine-grained control over the instrumentation of the Linux kernel and arbitrary x86 binary modules.

%The most successful existing solutions to kernel integrity protection and module isolation (\emph{LXFI} \cite{YMao2011}, \emph{BGI} \cite{Castro2009}, and \emph{XFI} \cite{Erlingsson2006}) require access to kernel module source code. Our approach challenges the assumption that module source code is required in order to achieve similar levels of protection. This is a good assumption to challenge because most kernel modules are developed by third party developers and are distributed as binaries.

%The main contribution of our work is to show that arbitrary kernel modules can be secured without language, compiler, or module/kernel source code changes, and without specific hardware requirements. We show how to dynamically detect the boundaries of modules linked with the kernel and use this information to isolate modules without runtime overhead when executing non-module kernel code. Finally, we show how to combine high-level type-based and functional annotations with module boundary detection and isolation  to impose runtime invariants on module code and propagate checks and type information into the module while it is running.

%Our approach has been to start with the DynamoRIO Kernel (DRK) \cite{Feiner2012,Feiner2010b} DBI framework, which instruments all kernel and module code, and transform it to instrument only module code. This has required significant changes to DRK, which primarily revolve around efficiently attaching and detaching instrumentation when transitioning between modules and the kernel, and on handling interrupts and exceptions on multi-core processors. No other DBI frameworks met our requirements of instrumenting arbitrary kernel code without kernel modification, and with the ability to run in a virtualized environment.

%\Sysname is loaded into the kernel as a module. It takes control of the interrupt descriptor table and positions itself to intercept the module loading process. When untrusted modules are loaded by the kernel, \sysname replaces its exported function addresses. The replaced function addresses act as initial entry points into instrumented code. When the kernel initializes the module, it invokes one of the entry points, which puts execution of the module under the control of \sysname. Execution remains under our control until the module calls a kernel function, returns control to the kernel, or until execution is interrupted by a hardware interrupt or exception.

%While under our control, all direct and indirect control flow is checked. Calls to kernel functions are redirected through wrappers functions. Attempts to call non-exported kernel functions result in errors. Kernel function wrappers validate arguments according to rules specified on a function and type-specific basis. Validation involves checking and updating any function pointers that point into the module's code segment. Basic validator rules are automatically generated by a GCC plugin, and can later be extended by a programmer with domain knowledge. Wrapper functions that apply validation rules are automatically created when \sysname is compiled.

% --------------------------
% --------------------------

\section{Dynamic Binary Instrumentation} \label{sec:DBI}

In this section, we give a brief background on dynamic binary instrumentation (DBI) in general, and DynamoRIO Kernel (DRK) in particular. DBI is a form of binary translation. Binary translation is used for instruction set translation/emulation \cite{Qemu2005}, runtime optimization, and runtime instrumentation (analysis, security, debugging) \cite{Ford2008,NaCl,Nethercote2007a}. Binary translation is attractive when one wants to make as few assumptions about the nature of the applications being translated. On architectures like x86 where the mix of variably-sized instructions and data make static translation intractable, dynamic binary translation (DBT) is necessary. In this paper, we focus on comprehensive, dynamic instrumentation of all x86-64 binary kernel modules. Comprehensive instrumentation enables powerful tools such as memory, control-flow, and lock checkers, needed to debug, secure, and analyze kernel modules.

Comprehensive instrumentation of a binary by a DBI tool requires that the DBI tool be in control of all instructions executed from that binary. DBI tools do this by operating on short sequences of straight-line instructions ending in control-flow transfer instructions (CTIs), called \emph{basic blocks}. Basic blocks are translated so as to transfer control to the tool's \emph{dispatcher} at the end of their execution. The dispatcher translates and transfers control to new basic blocks, thus emulating the behaviour of the original CTIs. Control remains with the tool to the extent that only translated versions of non-CTIs (e.g. arithmetic and memory instructions) from the original binary are executed natively, while all CTIs are emulated.

Comprehensiveness also requires that a DBI tool hide and protect itself from the instrumented binary. Hiding is necessary for transparency: instrumentation should not change program semantics or behaviour. A program running with instrumentation should not be able to (easily) determine that it is being instrumented, lest it change its behaviour in response to being instrumented. Protection is necessary in the event that a module is intentionally or accidentally malicious. For example, if a malicious module subverted the control of the DBI tool then it could arbitrarily generate executable code and escape further instrumentation.

%Comprehensive DBI is challenging in user space, especially when dealing with multithreaded, multiprocessing, and asynchronous execution (signals). DBI is especially challenging in kernel space: interrupts and exceptions arbitrarily alter control flow, and interrupted kernel threads can resume their execution on different CPU cores.

We built our framework on top of DRK. DRK is a kernel-space port of the user-space DynamoRIO DBI framework \cite{Bruening2004}. DRK comprehensively instruments all Linux kernel and module code, including interrupt and exception handlers, and works on multi-core processors \cite{Feiner2012}. Further, DRK maintains the precise interrupt delivery semantics of the x86 architecture. DRK is loaded as a kernel module and thereafter controls the execution of all kernel code. Execution of kernel code happens only from within DRK's code caches. DRK caches instrumented kernel code on a per-CPU basis. When DRK's dispatcher is invoked to complete a control-flow transfer to a previously un-executed instruction, it stores the newly translated basic block in the code cache of the current core. As an optimization, direct control-flow transfers are optimized to transfer control between two basic blocks in the code cache, instead of returning to the dispatcher. As with DynamoRIO, DRK exports an API to \emph{clients}, thus allowing programmers to write instrumentation and translation policies. DRK takes care not to instrument user-space code: instrumentation \emph{attaches} when user code is interrupted or invokes a system call, and \emph{detaches} when execution returns to user space.

% DRK handles interrupts and exceptions specially. Kernel and module code is necessarily sensitive to interrupt delivery: scheduling is dependent on interrupts, and modules often depend on interrupts to communicate with hardware. Further, the kernel and its modules can arbitrarily disable/enable interrupts, thus complicating any policy that might attempt to arbitrarily delay or discard interrupts. DRK goes to great lengths to preserve the x86 architecture's precise interrupt delivery.

%manages these issues in a number of ways. Caches of translated blocks are maintained on a per-CPU basis by using CPU-private memory. This improves transparency when dealing with concurrency while also setting a fixed upper bound on the number of redundant versions of a translated basic block.

%This design choice also necessitates transparency: any code cache address is unsafe 
%This design choice improves transparency , while also setting a fixed upper-bound on the number of redundant 
%sets an upper bound on the maximum number of redundant copies of a given translated basic block, while making some concurrency issues easier to handle. T

%into ones that end by transferring control to the tool's dispatcher instead. The dispatcher emulates the intended control-flow transfer by locating the next sequence of instructions to execute, translating those instructions until it encounters another control-flow transfer instruction, and then transferring control to the newly translated instructions.

%return control to the tool's dispatcher once the control-flow transfers are reached. DBI tools then emulate the intended control-flow transfers by means of a dis

%DBI systems must control all execution of code being instrumented.

%DBT is often used in cases where one wants to assume as little about the programs to static binary translation is

% involves dynamically translating all or parts of an executable (binary) while it is running. Dynamic binary translation is used for instruction set emulation \cite{Qemu2005}, runtime optimization, and instrumentation. We concern ourselves with the latter: instrumenting binaries by injecting

%DBI is a technique typically used to implement debugging tools \cite{Bruening2010,Seward2005}, isolation tools \cite{Ford2008}, and virtual machine monitors. \todo{cite vmware?}

%DBI is used in cases where one wants to inject or rewrite code in a running program

%DBI \cite{Bruening2004} is an execution control and analysis technique that involves injecting binary code into a running program. The injected code is termed ``instrumentation''. DBI frameworks like \emph{DynamoRIO} \cite{Bruening2004}, \emph{Pin} \cite{Bungale2007}, \emph{JIFL} \cite{Olszewski2007}, \emph{KProbes} \cite{KProbes}, and \emph{KernInst} \cite{Tamches1999} provide the infrastructure needed to perform fine-grained monitoring and manipulation of existing binaries. However, none of the aforementioned frameworks provide comprehensive or exclusive instrumentation of kernel code.

%\emph{DynamoRIO Kernel} (DRK) is a port of \emph{DynamoRIO} that instruments all kernel code, including module code \cite{Feiner2012,Feiner2010b}. DRK is loaded as a kernel module and takes control of all kernel code execution. DRK is able to provide the same level of fine-grained instrumentation as its user space counterpart, \emph{DynamoRIO}.

%DRK caches instrumented kernel code on a per-CPU basis to amortize the cost of (de/re)compiling parts of the kernel code. Each DRK code cache contains instrumented basic blocks: short sequences of kernel code interleaved with instrumentation code, where the original sequence of kernel code has a single entry and exit point. A basic block is eagerly inserted into DRK's code cache when it is the target of a direct branch, and lazily inserted into DRK's code cache when that block is the target of an indirect branch. Once in the code cache, execution never leaves the code cache, except for when a special ``dispatcher" function is called to find or create the next basic block to be executed.

% --------------------------
% --------------------------

\section{Architecture and Design} \label{sec:arch}

In this section, we describe how {\sysname} meets the goals laid out in \Secref{intro}. To do so, we describe several challenges, alternative approaches, and the solutions we chose.

% --------------------------

\subsection{Mixed-mode Execution}
Our first goal of comprehensiveness requires that {\sysname} control the execution of all module code. We can achieve this goal by attaching instrumentation when some module code is first executed and leaving instrumentation attached until the next context switch. However, our second goal requires that any overhead introduced by {\sysname} only affect instrumented modules. {\Sysname} introduces overhead when it translates, caches, and executes instrumented code. If {\sysname} is left attached then it would eventually instrument kernel code and fail to meet our second goal. This interaction between our first two goals forces code to execute in one of two modes: native or instrumented. Under this lens, module code executes in instrumented mode and kernel code executes in native mode. Attaching {\sysname} causes a native-to-instrumented mode switch and detaching causes an instrumented-to-native mode switch.

{\Sysname} dynamically detects when control transfers between the kernel and its modules and switches the mode accordingly. Detecting and interposing on control-flow transfers that should lead to mode switches is challenging. The obvious way to detect when to switch modes is to inspect every control-flow transfer instruction. This is easy when executing instrumented code. For example, the dispatcher can switch modes when it detects that control is transferring to the kernel. However, interposing on module entry points is non-trivial because {\sysname} has no control when kernel code is executing. Page protection is one mechanism that {\sysname} uses to detect control-flow tranfers to modules. All module code is made non-executable, and any attempt to execute module code results in a page fault exception. Page fault exceptions are first handled by {\sysname}, which chooses to attach instrumentation if the faulted instruction is a module address. Because of their cost, page faults are depended on only as a last resort. Instead, {\sysname} changes the module entry points used by the kernel, allowing for efficient entry point interposition.

TODO

%{\Sysname} dynamically and proactively changes module entry points before they are observed and executed by the kernel. Instrumentation attaches when the kernel executes these changed entry points. Before attaching, 
%
%
%Before and after each invocation of a kernel function, {\sysname} attempts to discover and change potential module entry points.
%
%
%Potential module entry points are discovered and changed each time {\sysname} detaches because of a call 
%
%
%Several implementation strategies were available when considering how to avoid page faults:
%\begin{enumerate}
%	\item Change each indirect call in the kernel source code that has the potential to transfer control to a module. These changes can be done manually or automatically. Extensive manual changes to the kernel's source code failed to meet our third and fourth goals. Automatically changing the kernel source code (e.g. using a compiler extension) had the potential to introduce unnecessary performance overhead on indirect function calls that do not transfer control to any module.
%\end{enumerate}

%Dynamically changing entry points was motivated by the goals set out in \Secref{intro}.

%Our approach needed to work on arbitrary binary modules, and so changing how module vendors compile their modules was not an option. Our approach could not depend on significant changes to the kernel or impose unnecessary overhead to non-module kernel code. This ruled out (manually or automatically) changing how the kernel calls potential module entry points. Finally, our approach needed to be automatable.

%the compiler toolchain used by module vendors was not an option. Any approach  
%
%The same effect can be achieved by changing the source code; we decided against changing the kernel source code for several reasons:
%\begin{enumerate}
%	\item Manually changing source code does not meet our third or fourth goals.
%	\item Automatically changing the kernel source code (e.g. with a compiler) meets our fourth goal, but suffers two problems. First, such automation would require pointer analyses, which are conservative by nature. This conservatism can lead to a higher dependence on page faults for handling attaching. If the anaylses are made overly aggressive, then some invocations of kernel code by the kernel will attempt to attach {\sysname}. Both cases introduce unnecessary overhead.
%	\item Significant changes, either to the kernel or to the compilation toolchain, are unlikely to be adopted by the greater community.
%\end{enumerate}

%Our third goal of ease of portability across kernel versions and changing APIs meant that we did

%Our first goal of instrumenting arbitrary binary modules meant that we could not change modules to

%{\Sysname}'s easily detects mode-switching control-flow transfers when instrumenting a module because it's dispatcher controls the flow of execution within an instrumented module. However, when executing in native mode, {\Sysname} has no control over execution.

%First, {\sysname} is not in control of native execution. Unlike native execution, all control-flow transfers are routed through {\sysname}'s dispatcher when  
%the control-flow boundaries between modules and the kernel and interposes

%Kernel and exempt module code execute natively and all other modules execute under the control of {\sysname}.

%Comprehensiveness requires only that all module code is instrumented. As a result, {\sysname} must control the execution of all module code. Comprehensiveness 
%Any attempt to execute non-exempt module code must attach {\sysname}, thus switching from native to instrumented mode. Once attached, {\sysname} 
%Kernel code and modules exempt from instrumentation execute natively, while all other modules are instrumented. 
%Control-flow transfer between the native and instrumented code requires mode switches. 
%However, comprehensive instrumentation ensures that the overhead of instrumentation does not spill into native code: {\sysname} dynamically detects when to attach and detach instrumentation so that native code is not accidentally instrumented. {\Sysname} detects the

%Interposing on all mode boundaries is challenging. Any indirect control-flow transfer potentially requires a mode switch because function pointers are shared between modules and the kernel. Further, all interrupts are initially handled by the kernel, requiring at least two mode switches any time an instrumented modules is interrupted.

% {\Sysname} achieves this and imposes no performance overhead on non-instrumented code.  Uninstrumented code includes user-space code, kernel code, and modules exempt from isolation. In the former case, user-space code is never instrumented. In the latter two cases,

%{\sysname} dynamically detects and detaches before transferring control from instrumented code to code that should not be instrumented. Ensuring that kernel code is not instrumented is complicated by interrupts:

%To isolate a module, {\sysname} must control

%Our first goal was to allow for comprehensive instrumentation of module execution without direct overhead to kernel or user-level code. Our second goal requires such coverage be available to arbitrary binary modules, thus module source code, language, or compiler changes were not an option. Our third goal--ease of portability across different versions of the same OS and hardware--meant that our approach could not depend on significant changes to the kernel or depend on specific hardware features (e.g. hardware protection domains, hardware-assisted virtualization). Together, these three goals have motivated the use of a novel dynamic control-flow isolation technique.

%Isolating a module requires that {\sysname} takes control of the execution of all module code when invoked by the kernel, and relinquish control when execution returns to the kernel. This form of mixed-mode execution within the kernel introduces many isolation challenges, especially with interrupts and exceptions (\Secref{interrupts}). To take, maintain, and yield control of execution, direct and indirect control-flow transfers must pass through {\sysname}'s dispatcher. \note{TODO: reference a diagram that shows the dispatcher} However, taking control of execution on entry to a module requires that the kernel calls the dispatcher on our behalf. This presupposes that the kernel knows exactly which control-flow transfers enter the module. Relinquishing control to the kernel from an instrumented modules requires similar information.

%Normally, when a module is loaded into the kernel, its code segment is placed into a well-defined region of executable memory. At a minimum, a module's symbol table exports the \texttt{init} and \texttt{exit} functions used by the kernel to install and uninstall the module. These and any other functions exported by the module represent natural entry points into the module. However, the functions exported in a module's symbol table are a subset of the available entry points into a module. Typically, modules register callback functions with the kernel, and these functions are not always present in the module's symbol table. As a result, any address into a module's code segment represents a potential entry point into the module.

%Our approach to detecting and interposing on module entry points has been to partition the kernel's region for module code into a \emph{raw module region} and a \emph{shadow module region}. We call addresses into these regions \emph{raw addresses} and \emph{shadow addresses}, respectively. The kernel writes module code into the raw module region when loading modules. The shadow region--which is unknown to the kernel but is maintained by {\sysname}--occupies the same amount of memory as and is located at a constant offset from the raw region. This one-to-one correspondence between raw and shadow addresses is used by {\sysname} to efficiently take control of module execution.

%We bootstrap {\sysname}'s control over the module by replacing each raw function address in the module's symbol table with it's associated shadow address (\Secref{shadow_module}). The shadow region occupies the same

%When the kernel attempts to install the module, it does so through the shadow \texttt{init} address, which redirects execution through {\sysname}'s dispatcher, thus taking control of module execution.

% We assume that the kernel will behave in a sensible way and not attempt to call arbitrary raw module addresses. In case the kernel is misbehaving or isolation is broken, the raw module region is page protected against execution. Given this expectation, the first entry point into a module must be an address in the module's symbol table. Knowing this

%In our current implementation, we have chosen to relax transparency and allow both the kernel and module to observe shadow addresses. Further transparency might be maintained by translating all shadow addresses into raw addresses so that an instrumented modules cannot any {\sysname} memory locations. Alternatively, all module addresses, including return addresses stored on the stack, could be translated to shadow addresses. While both are apparently attractive solutions, the former trades performance for transparency and the latter fails to maintain transparency for modules that inspect their own instructions.

%{\Sysname} yields control of an instrumented module back to the kernel in three cases: when a module jumps to or calls a kernel address, when a module returns to a kernel address, or when an interrupt or exception occur while executing module code. When {\sysname} is controlling module execution, all direct and indirect control-flow transfers are routed through the dispatcher. The dispatcher detects kernel addresses and detaches when control transfers. In the second case, \texttt{RET} instructions are rewritten to behave like jumps, thus allowing us to re-use the same mechanism of detaching. Finally, in order to properly detach in the event of an interrupt or exception, {\sysname} is always in control of each CPU core's \empty{interrupt descriptor table} (IDT), allowing {\sysname} to detach when interrupted and arrange for execution to re-enter {\sysname}'s control after the kernel has handled the interrupt. Again, we have chosen to relax transparency in the case where the return address of interrupts does not represent actual module code addresses.

% --------------------------

%\subsection{Maintaining Isolation}
%The methods described in the previous section isolate module execution up to the point where the kernel calls a raw module address. This is forbidden before a module is initialized but is common after module initialization. The primary means of extending the kernel's functionality is for a module to register (with the kernel) a data structure that implements a kernel interface. Modules register these data structures by invoking special kernel-exported functions. The kernel's interfaces typically require addresses to module functions, hence the possibility of raw module addresses leaking into the kernel. One approach to handling this case would be to augment the memory protection mechanism to trap and redirect execution of raw module addresses to their shadow equivalents. This solution hides failures in the isolation mechanism and has undesirable performance implications.

%Our approach has been to recognize the importance of the kernel-exported API and its role in making raw function addresses available to the kernel. When execution transfers from the module to the kernel, {\sysname}'s dispatcher detects the uses of kernel-exported functions and redirects execution to \emph{kernel wrappers}. Kernel wrappers recursively update the data structures passed to the kernel so that they contain shadow module addresses in place of raw addresses. After wrapping all function arguments, control is transferred to the intended kernel function. Kernel wrappers dynamically maintain isolation by ``growing'' the set of known module entry points. As an optimization, we developed a method that avoids redundant wrappings when shared data has not changed (\Secref{tainted_data}). The dependence on memory protection is relegated to detecting and maintaining isolation in the presence of failures and corner cases. Such corner cases arise when modules update the shared data structures after registering them with the kernel. In practice, these cases quickly self-correct, thus minimizing the potential for excessive performance overhead.

% --------------------------

\subsection{Interrupts and Exceptions \label{sec:interrupts}}

\begin{figure}
\[
\begin{matrix}
\tt{R_1} & & & & \tt{\overset{interrupt}!} & \tt{R_2}\\
\tt{R_1} & \tt{I_1} & \tt{\underset{interrupt}{\underset{actual}{\not{!}}}} & \tt{\underset{delay}{\underset{\xrightarrow[]{}}{I_2}}} & \tt{\underset{interrupt}{\underset{emulated}!}} & \tt{R_2}
\end{matrix}
\]
\caption{%
On the first line, an interrupt occurs between raw module instructions $\tt{R_1}$ and $\tt{R_2}$. On the second line, $\tt{R_1}$ and $\tt{R_2}$ are instrumented and instructions $\tt{I_1}$ and $\tt{I_2}$ have been injected between them. We delay interrupts occuring between $\tt{I_1}$ and $\tt{I_2}$ until $\tt{R_2}$ for the following reasons: i) $\tt{I_2}$ must eventually be executed, lest it be responsible for releasing a lock; ii) $\tt{I_2}$ must be executed before the interrupt is handled, lest it not be re-entrant (e.g. CPU-private lock), and; iii) the address of $\tt{I_2}$ cannot be meaningfully translated into a corresponding raw module address, so failure to execute $\tt{I_2}$ before handling the interrupt would likely result in never executing $\tt{I_2}$ when re-attaching {\sysname} at $\tt{R_2}$ after handling the interrupt (e.g. if $\tt{I_1}$ and $\tt{I_2}$ were injected to instrument $\tt{R_1}$).}
\label{fig:interrupt_delay}
\end{figure}

Maintaining isolation in the presence of interrupts and exceptions is challenging, and had an impact on the design/implementation decisions of {\sysname} and DRK before us. Our requirements for isolating only module code requires that {\sysname} safely relinquish and eventually regain control on arrival and after handling any interrupt/exception in an isolated module.

We define interrupts to be asynchronous events, not caused by any particular instruction, and exceptions to be synchronous events, caused by the execution of a single instruction. A key distinction between interrupts and exceptions is that interrupts can technically be handled at any point after their arrival, whereas exceptions must be handled before the triggering instruction can proceed. Like DRK, we do not queue interrupts as that can lead to inconsistent machine states, limits the diversity of supported hardware devices and drivers, and changes the original execution interleaving \cite{Feiner2012}. However, as with DRK, we sometimes delay the arrival of an individual interrupt so as to present a plausible machine state to the kernel.

On x86 processors, each CPU core handles interrupts and exceptions uniformly: each kind of interrupt/exception has a handler, and handlers are found in the CPU's IDT. When an interrupt/exception arrives, the CPU locates the appropriate handler in the IDT and control transfers to the handler's address \cite{IntelVolume1}. Interrupt/exception handlers are implemented by the kernel and must not be instrumented by {\sysname}. Handlers can observe and change the state (e.g. registers, runtime stack) of interrupted code, and so presenting a plausible machine state to the kernel handlers is essential.

Handling interrupts of isolated module code is tricky; there is not a one-to-one correspondence between the instructions of instrumented code (which live in {\sysname}'s code cache) and raw module code. This is partially because {\Sysname} allows extensions (called clients) to rewrite or inject code into the code cache (thus allowing for fine-grained instrumentation of modules). Client code need not be re-entrant (but must be idempotent), and can live in-between logical module instructions. Further, complex instructions such as \texttt{CALL}, \texttt{JMP}, and \texttt{RET} are emulated by instructions that interact with the dispatcher. As a result, it is important that interrupts be delivered on logical module instruction boundaries, lest some interrupted client/{\sysname} code holding a lock cause a deadlock. This requirement necessitates that the handling of some interrupts be delayed (\Figref{interrupt_delay}).

Like DRK, {\sysname} delays interrupts in the code cache until the next logical module instruction boundary \cite{Feiner2012}. Exceptions cannot be delayed, and are errors if they occur on non-module instruction boundaries. Delaying allows {\sysname} to emulate x86's precise interrupt delivery, preserves transparency, and imposes fewer burdens on client code authors. However, delaying is complicated by mixed-mode execution. Interrupt handlers are kernel code and must not be isolated. When execution of an isolated module is interrupted, {\sysname} must detach for the kernel to handle interrupt, and re-attach upon return from the interrupt. Re-attaching can be slow because a partially redundant basic block must be translated to finish executing the originally interrupted basic block. These \emph{interrupt tails} have not been a memory/garbage issue; however, the delay caused by their construction motivated an important optimization (\Secref{fast_trans}).

%Our handling of interrupts during the transitions from {\sysname} to kernel code and from {\sysname} to instrumented code is unique because of our mixed-mode considerations. In the former case, the burden of delaying an interrupt is placed on the hardware: execution continues in the kernel after the transition is complete, at which point interrupts may safely occur. Thus, interrupts are disabled until all {\sysname}-specific state has been cleaned up the machine state of the once-isolated code has been restored.

%Re-attaching af

%Our treatment of the latter case was motivated by performance overhead resulting from the mismatch of mixed-mode execution under {\sysname} and single-mode execution under DRK. We observed a case where our original usage of DRK's interrupt delaying mechanism resulted in periods where no forward progress was made toward executing isolated module code (i.e. the kernel was stuck bouncing between {\sysname}'s dispatcher and handling interrupts). Solving this issue involved creating a new means of delaying interrupts and arranging for any hardware-queued interrupts to arrive in a safe ``catch-all'' region during transitions without having to re-enter the dispatcher.

%Delayed interrupts are forcible re-raised by patching the originally interrupted code. Returns from interrupt handlers to module code are re-directed to dispatcher to maintain module isolation. This dispatch can cause a new (and potentially redundant) basic block to be emitted to the code cache, where the new basic block is a suffix of the interrupted basic block. As with DRK, these \emph{interrupt tails} have not been an issue in practice. Finally, the dispatcher executed with interrupts disabled because it is not re-entrant.

%Consider the case where, in the process of transitioning into the code cache, an interrupt occurs. Under DRK (where the entire kernel is instrumented), the interrupt would be delayed and instead of entering the code cache for the next basic block, DRK would enter the code cache for the interrupt handler. In {\sysname}, the the extra dispatch to the interrupt handlers transitions back to native code (but with interrupts disabled). Eventually, the interrupt is handled and \texttt{IRET} is executed, which re-enables interrupts and enters the dispatcher, which again attempts to transition to the original basic block. During this transition, an interrupt arrives and no forward progress has been made toward executing the basic block.

% --------------------------
% --------------------------

\section{Implementation}

We modified DRK \cite{Feiner2012} to instrument only module code. Instrumenting only module code required that we be able to dynamically gain control on all raw module entry-points and relinquish control when control transfers to the kernel. Because of the in-memory organization of kernel code, detecting when control should be relinquished requires checking if the target of a CTI lies within a fixed range of addresses. Without heavyweight mechanisms like page protection, it appears impossible to detect when kernel code transfers control to a raw module. This has required the development of two novel techniques for discovering and efficiently attaching instrumentation at module entry points: kernel wrappers and shadow modules.

% --------------------------

\subsection{Kernel Wrappers \label{sec:kernel_wrappers}}

\begin{figure*}
\lstset{language=C, tabsize=2, stepnumber=1}
\begin{multicols}{2}
\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
struct device_driver {
	...
	int (*probe)(struct device *);
	int (*remove)(struct device *);
	void (*shutdown)(struct device *);
	int (*suspend)(struct device *, pm_message_t);
	int (*resume)(struct device *);
	...
	const struct dev_pm_ops *pm;
	...
};
\end{lstlisting}
\vfill
\columnbreak
\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
TYPE_WRAPPER(struct device_driver, {
    PRE {
        WRAP_FUNC(arg.probe);
        WRAP_FUNC(arg.remove);
        WRAP_FUNC(arg.shutdown);
        WRAP_FUNC(arg.suspend);
        WRAP_FUNC(arg.resume);
    }
    POST {
        WRAP_RECURSIVE(arg.pm);
    }
})
\end{lstlisting}
\end{multicols}
\caption[LoF entry]{Example type wrapper for the Linux \texttt{device\_driver} structure. In the above code, \texttt{arg} is an implicitly defined reference to an instance of the \texttt{device\_driver} structure. Modules typically initialize the function pointers in this structure, and so they must all be wrapped as a potential module entry points.

Function argument type wrappers created using \texttt{TYPE\_WRAPPER} apply uniformly to arguments passed by value, by pointer, and by reference. In the case of pointers, a check is automatically performed to ensure that a type-based policy wrapper is not applied to a \texttt{NULL} pointer. The \texttt{pre} and \texttt{post} sections represent policies to be applied before and after the wrapped kernel function is called, respectively. This is desirable because often the kernel will initialize fields of structures.

Function argument type wrappers are specified in Granary, and use the type information exported by our GCC plugin.}
\label{fig:type_wrapper}
\end{figure*}

\begin{figure*}
\lstset{language=C, tabsize=2, stepnumber=1}
\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
struct task_struct *kthread_create(threadfn thread_fun, void *data, char namefmt[], ...);

FUNC_WRAPPER(kthread_create, (threadfn thread_fun, void *data, const char namefmt[], ...), {
    struct kthread_create_info *create(0);
    char format_buff[sizeof create->result->comm];
    va_list args;
    va_start(args, namefmt);
    kern_vsnprintf(&(format_buff[0]), sizeof format_buff, namefmt, args);
    va_end(args);
    WRAP_FUNC(thread_fun);
    return kthread_create(thread_fun, data, format_buff);
})
\end{lstlisting}
\caption{Example function wrapper for creating Linux kernel threads. Function wrappers are useful for small functions and functions where the default, automatic argument type-based policies are insufficient. Automatic type-based policies are applied independently of one-another. In some cases, a pointer passed as a function argument is the address of the first object in an array, and the length of the array is known through another argument. These edge cases are neatly handled by function wrappers. Further, function wrappers enable stubbing-out or re-implementing kernel-exported functionality.}
\label{fig:func_wrapper}
\end{figure*}

Detecting when control might transfer from the kernel to a raw module requires proactively changing all potential raw module entry points into behaviourally equivalent {\sysname} attach points. However, detecting these entry points requires understanding what kinds of data are passed to the kernel through function calls, and wrapping those data accordingly. Information about data types is not present in either of the module or kernel binaries. However, type information can be deduced based on the kernel API functions being used.

We detect and handle potential entry points as they are exposed to the kernel through its exported API. We define type-specific argument wrapping policies and function-specific wrapping policies. Wrapper policies apply automatically when an isolated module calls a kernel API function. The policies of the wrapped API maintain isolation by discovering and changing potential raw module entry points into shadow module entry points. These policies are declared using a combination of \texttt{C++11} and the \texttt{C} pre-processor (\Figref{type_wrapper,func_wrapper}).

Our goal was to automate the process of isolating new modules. To that end, we developed a GCC plugin that emits exported function and type declarations to a separate file when compiling the Linux kernel. A script automatically transforms the emitted declarations into kernel function argument type wrappers (\Figref{type_wrapper}) and kernel function wrappers (\Figref{func_wrapper}). Minor manual adjustments are needed to impose domain-specific checking and wrapping for function wrappers. At {\sysname}'s compile time, a \texttt{C++} meta-program automatically generates concrete kernel function wrappers for each of the (approx.) \numprint{6000} kernel-exported functions from the policies. Function policies are directly transformed into wrappers. For those kernel functions without specific function policies, function wrappers are generated that apply the argument-type policies to each of the kernel function's arguments.

For efficiency, a mapping from kernel-exported function addresses to wrapper function addresses is stored in a minimal perfect hash table \cite{Belazzougui_Botelho_2008}. When emulating CTIs, the {\sysname} dispatcher first checks if the target of the CTI is a kernel address. If so, the kernel address is looked up in the hash table, and control transfers to the mapped kernel function wrapper after {\sysname} detaches. Execution of the wrapper ensures that isolation is maintained, even when new module entry points are indirectly exposed through function arguments.

Our declarative approach and use of meta-programming has enabled two novelties:
\begin{enumerate}
	\item Aggressive, type-specific tainted structure tracking using function pointers and heap-allocated structure extension (\Secref{tainted_data}).
	\item The imposition of high-level type and semantic information onto opaque module binaries and their heap-allocated data structures, without changing the kernel or the module.
\end{enumerate}

% --------------------------

\subsubsection{Dynamic Kernel Wrappers \label{sec:dynamic_wrapper}}
In some cases, wrapping only exported kernel functions is insufficient to maintain isolation. For example, when isolating network drivers, we discovered that a non-exported kernel function (\texttt{x86\_\linebreak[0]swiotlb\_\linebreak[0]alloc\_\linebreak[0]coherent}) was executed by some modules. This case was an instance of a more apparent problem: internal fields of arguments to kernel functions are sometimes initialized/altered by the kernel for explicit use by modules. To maintain isolation, it is insufficient to only wrap arguments before they are passed to the kernel. This discovery showed that we needed a way to re-wrap arguments after they were passed to a kernel function, and was the motivation for distinguishing between \texttt{PRE} and \texttt{POST} wrappers.

We handle the above case with a \texttt{POST} wrapper that dynamically compiles a partial kernel function wrapper (consistent with the type of the function pointer field). The exposed function pointer is then replaced with the address of the dynamically generated code.

% --------------------------

\subsection{Shadow Modules \label{sec:shadow_module}}

\begin{figure}
\lstset{language={[x86masm]Assembler},stepnumber=1}
\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
      CALL next
next: JMP  attach_granary
\end{lstlisting}
\caption[LoF entry]{Shadow module code template for handling module entry points. When the kernel calls a shadow address, it immediately executes another \texttt{CALL} instruction. That instruction transfers control to the next instruction and also places the address of the next instruction on the stack. When the \texttt{JMP} instruction is executed, control transfers to {\sysname}. {\Sysname} inspects the return address left on the stack by the shadow \texttt{CALL} instruction and deduces the intended raw module target address. {\Sysname} attaches and begins instrumentation from the raw address.}
\label{fig:shadow_template}
\end{figure}

Kernel wrappers must be able to efficiently detect and wrap raw module addresses, as well as ignore already wrapped raw module addresses. Further, a mechanism must be in place that--upon execution of a wrapped raw module address--discovers the original raw module address for use by {\sysname}'s dispatcher. We opted for a simple solution that meets both these needs: each raw module has a corresponding \emph{shadow module}, and there is a one-to-one mapping between raw and shadow module addresses.

%, where half occupies 536MB

We implemented shadow modules by partitioning the region of memory usually alloted for kernel modules into two contiguous halves: a raw and shadow half%
\footnote{In perspective, this is not such an extreme use of memory: typical modules are small, and each CPU core already maintains a private code cache that can grow to at least the size of the modules being instrumented.}. Wrapping and unwrapping raw/shadow module addresses takes only a few instructions: we add or subtract \texttt{0x20000000} to translate a raw module address to/from a shadow address. If a raw address has not previously been encountered, then the corresponding shadow address is initialized with dynamically generated code that attaches {\sysname} (\Figref{shadow_template}).

When the kernel installs a module, it initializes portions of the raw module region with that module's instructions and static data. As part of the normal installation of a module, the kernel learns about an entry points into the module: \texttt{init}. Normally, the kernel calls \texttt{init} to initialize the module and its data structures. {\Sysname} intercepts module initialization by first wrapping \texttt{init}. From here, all other module entry points are discovered and wrapped by kernel wrappers. Isolation is enforced by exposing only shadow module addresses to the kernel, and page protecting the entire raw module region against execution as a fall-back.

%Wrapping a raw module address involves , and placing that code at the corresponding shadow address. As kernel wrappers discover and wrap raw module entry points, more of the shadow region is filled out, and control flow isolation is maintained.

% --------------------------
% --------------------------

\section{Optimizations}

In this section, we describe two optimizations to {\sysname} made possible by our use of kernel wrappers and the mixed-mode execution environment. Together, these optimizations resulted in a worst-case 50\% decrease in overhead. The first optimization, tracking tainted data, detects shallow changes to kernel data structures to avoid redundant wrappings. The second optimization, fast detach, applies when a module directly calls or returns to the kernel.

% --------------------------

\subsection{Tracking Tainted Data \label{sec:tainted_data}}

Kernel wrappers recursively wrap arguments to kernel functions. However, these arguments can be deeply nested and linked structures (e.g. file system inodes). Continually re-wrapping deep structures each time they cross the module-kernel boundary is expensive.

A partial solution to this problem limits the depth of wrapping. In the worst case, some raw module addresses escape wrapping. If invoked by the kernel, a raw module address will trigger a memory protection exception and be handled by {\sysname}. {\Sysname} re-directs control to the shadow module, thus maintaining isolation. This solution alone is undesirable as it does not correct the root problem.

Ideally, every kernel object would contain and embedded hash. If this were the case, then shallow changes to an object would be detected when a wrapper compares the object's embedded hash against a re-computed hash. Again, deep changes might escape detection, but not break control-flow isolation. If the kernel invokes a raw module address, then {\sysname} would take control as before but change the seed of the hash function, thus invalidating every previously wrapped data-structure and correcting the root problem.

Unfortunately, embedding hashes into kernel objects is not practical, and overwriting arbitrary object fields for storing the hash can result in data loss and corruption. However, one can change a function pointer, so long as the changed function pointer behaves as before. The restriction to function pointers is not so bad: many kernel objects reference shared virtual tables (vtables) of function pointers, and continually re-wrapping these vtables is wasteful.

We track changes to objects containing at least one function pointer by designating a tracking function pointer (TFP), as shown in \Figref{tracker}. TFPs are changed to address dynamically generated code that \texttt{JMP} to the TFP's original target address, thus maintaining the same control-flow behaviour. The hash of the object is located beside the \texttt{JMP} instruction in memory. Checking for changes to an object is analogous to comparing agaisnt an embedded hash, but instead compares against the TFP hash. As before, breaches in control-flow isolation are handled by {\sysname} and invalidate all previously computed hashes.

\begin{figure}
\lstset{language=C, tabsize=2, stepnumber=1}
\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
struct device_driver {
	...
	int (*probe)(struct device *);
	...
};

SET_TAINT_TRACKER(struct device_driver, probe);
\end{lstlisting}
\caption{Example of setting a function-pointer based tainted data tracker to the \texttt{device\_driver} structure (\Figref{type_wrapper}). \texttt{PRE} and \texttt{POST} wrappers are automatically compiled to transparently handle the tainted data detection protocol. Tainted data trackers specified in Granary, and use the type information exported by our GCC plugin.}
\label{fig:tracker}
\end{figure}

% --------------------------

\subsection{Fast Attach and Detach \label{sec:fast_trans}}
Transferring control between the kernel and an isolated module is analogous to a \emph{context switch}. When the kernel invokes a shadow address, control transfers to the {\sysname} dispatcher, which executes on a CPU-private stack. Control reverts back to the kernel stack to execute instrumented code from the code cache. This stack switch--along with any dynamic code translation and lookup--is expensive but necessary for maintaining isolation. If the dispatcher executed on the kernel stack and left evidence of itself behind, then any stack-use-after-return bugs in kernel or module code could potentially corrupt {\sysname}'s internal state. Even though context switches are costly, they can be predicted and optimized.

The first optimization improves attach speeds after handling interrupts. Interrupts are handled by the kernel, and so {\sysname} must detach when instrumented code is interrupted. Unfortunately, the dynamic instructions executed to context switch and re-attach after handling an interrupt cannot be avoided. Because most of the context switch and code translation is non-preemtible, a long context switch increases the likelihood of the hardware queuing interrupts for later delivery. When instrumented execution resumes, any hardware-queued interrupts are delivered, necessitating further context switching. This is particularly concerning for interrupt-driven modules like network drivers, where enough network traffic can inhibit forward progress of the interrupted module. Our solution was to create a ``catch-all'' location (a series of \texttt{NOP}s) where queued interrupts could safely arrive and be handled natively without requiring any context switches.

The second optimization improves detach speeds when an isolated module makes a direct call to the kernel. Direct calls use the \texttt{CALL} instruction and encode their target addresses. These represent low-hanging fruits for detaching {\sysname}: at basic block translation time, a range check can detect that a direct call targets the kernel. We avoid context switches by changing the translation/emulation scheme of direct calls to quickly detach {\sysname}. This optimization also avoids repeated (but fast) hash table lookups for mapping kernel addresses to kernel wrappers, as the lookup need only be done during basic block translation.

The final optimization improves detach speeds when returning to the kernel. This optimization is similar to the direct call optimization. For each emulated \texttt{RET} instruction, we first test if the return address targets the kernel. If so, a fast-path is taken to detach.

% --------------------------
% --------------------------

\section{Evaluation}
TODO

\subsection{Microbenchmarks}
TODO: kallsyms module, as a way to estimate performance overhead of context switches.

\subsection{Macrobenchmarks}
TODO: netperf, postmark, lmbench, IObench

% --------------------------
% --------------------------

\section{Applications}
TODO

\subsection{Control-Flow Integrity}

We implemented partial control-flow integrity (CFI) enforcement for all isolated modules. Full CFI enforcement requires that control flow through a program follow only those paths available through a pre-determined control-flow graph \cite{Abadi2007}. {\Sysname} enforces the following CFI policy:
\begin{enumerate}
	\item A isolated module can arbitrarily \texttt{JMP} to any other isolated module address.
	\item A module can only call exported kernel functions. If a non-exported kernel function is called, then it must be one which was made available to the module by the kernel (\Secref{dynamic_wrapper}).
	\item If a module function (either by the module or by the kernel) is called then control must resume at the next logical module/kernel instruction when control returns from the called code.
\end{enumerate}

{\Sysname} detects violations of the CFI policy in two ways: wrapper-based detection, and stack-based detection.

\subsubsection{Wrapper-based Detection}
If a module transfers control to a non-module address, then {\Sysname} checks if a kernel wrapper for that particular address exists (\Secref{kernel_wrappers}). If no wrapper exists then that address does not correspond to an exported kernel function. The only (valid) alternative is that the address points into the region of memory containing generated code for dynamic kernel wrappers (\Secref{dynamic_wrapper}). If that is not the case, then a breach of CFI is detected and the module is terminated.

\subsubsection{Stack-based Detection}
{\Sysname} maintains a thread-private CFI stack. When the kernel \texttt{CALL}s a shadow module address or when {\sysname} emulates a \texttt{CALL} instruction within the module, an entry is pushed onto the CFI stack. Each entry of the CFI stack is a tuple containing a stack pointer and an instruction pointer. The instruction pointer is the next logical module/kernel address. This address is automatically pushed onto the runtime stack by the \texttt{CALL} instruction. The stack pointer is the value of the stack pointer register (\texttt{\%rsp}) after the \texttt{CALL} instruction completes.

When emulating a \texttt{RET} instruction, we check that the stack pointer register corresponds with the most recently saved stack pointer in the CFI stack. If they are not the same then the module is terminated. The return address on the runtime stack is compared against the most recently saved instruction pointer in the stack. Again, if they disagree, then the module is terminated.

\subsubsection{Limitations}
Our CFI-enforcement policy does not prevent a module from arbitrarily changing the runtime stack or kernel data structures. A module can alter return addresses stored on the stack but not shadowed by {\sysname}'s CFI stack. For example, suppose kernel function \texttt{A} calls kernel function \texttt{B}, which in turn calls a module function \texttt{M}. The CFI stack only enforces that \texttt{M} must return to the next logical instruction in \texttt{B}. However, a malicious or exploited module (e.g. buffer overflow) could  alter control-flow so that \texttt{B} does not return to \texttt{A}. A similar attack vector involves changing function pointers stored in kernel data structures. While the page protection of the raw module region reduces the practicality of both attacks, full enforcement of CFI would require implementing byte-granularity isolation \cite{Castro2009,YMao2011}.

% --------------------------
% --------------------------

\section{Related Work}
TODO

%{\Sysname} then protects the newly initialized raw module code against execution as a last-resort means to maintain isolation. When instrumenting an isolated module, {\sysname} translates the instructions from the raw module region.

% and is page protected against execution as a last-resort means to maintain isolation. {\Sysname} incrementally decodes and translates instructions from this region when building basic blocks.

%Each shadow address is a fixed offset  (536MB) from its corresponding raw module address, allowing for very fast translation between raw and shadow addresses.

%The shadow module region is sparsely populated with {\sysname} trampolines.

%\subsection{Kernel wrappers}
%Kernel wrappers are created when {\sysname} is compiled. The set of exported kernel functions (approx. 6,000) and their addresses are used to create a minimal perfect hash table \cite{Belazzougui_Botelho_2008} mapping kernel functions to kernel wrappers. \Sysname guarantees that control flow transfers from an instrumented module to an address associated with kernel code are instead directed to the address found through a hash table lookup.

%Kernel wrappers are generated at compile time using C++11 variadic templates, which are parameterized over kernel function addresses and kernel function types. Wrapper templates automatically and consistently wrap kernel functions. Our approach allows policies to be specified in terms of individual kernel function argument types, and have those policies automatically applied within the generated kernel wrappers. \Figref{pci_wrapper} provides an example of the type-based wrapper for PCI device drivers.

%\section{Threat Model}
%%This section discusses three different attack scenarios which can be used to compromise the kernel's integrity. 
%%\begin{enumerate}
%%	\item An attacker can gain root privileges and load a malicious module (e.g. kernel-level rootkit) into the OS kernel. 
%%	\item An attacker can exploit a vulnerability existing in an already loaded module and inject malicious code to be executed in the kernel's address space.
%%	\item A system administrator can load an unverified kernel module which itself contains malicious code. 
%%\end{enumerate}

%Modules can compromise the control flow integrity and data integrity of the kernel in several different ways: 
%\begin{enumerate}
%	\item Modules can modify kernel code or control data (e.g. system call table, interrupt descriptor table, and function pointers).
%	\item Modules can modify non-control data (e.g. device list), instruct devices to attack the kernel via direct memory access (DMA) requests.
%	\item Modules can manipulate the kernel stack (e.g. return-oriented attacks).
%\end{enumerate}
%To protect the integrity of the kernel the system needs to enforce the following properties.

%\paragraph{Kernel Code/Data Integrity} Code, statically allocated data, and dynamically allocated data of the OS kernel are all protected from against prohibited modifications (that change the value being written) by untrusted modules via direct memory access or DMA.

%\paragraph{Control Flow Integrity} Control transfers from untrusted modules to the OS kernel---including function calls, (in)direct jumps, and preemptions---are restricted to a set of kernel service functions pre-identified as trusted entry points into the kernel. Call-return consistency across the module and kernel interface must be strictly enforced. That is, a return from an module to the kernel must target the instruction following the associated call into the module.

%\paragraph{Stack Integrity} Areas of the stack must be protected from memory writes by modules. These areas include (unless permission is explicitly given) function call return addresses, saved/spilled registers, function parameters, and the storage locations of local variables to kernel functions. Prohibited modification of stack data can result in both control flow and data integrity being violated (both locally and globally).

%\section{Scope}
%As described above, kernel modules can compromise the kernel in several ways. The scope of this project is limited to enforcing security policies that would detect control flow integrity violations by misbehaving modules. We do not address data integrity or its subset, stack integrity, both of which can influence control flow integrity. We also do not address the possibility of attacks by means of DMA.

%Future goals of this project include enforcing fine-grained security policies and type safety while the module is running.

%\section{Problem}
%Detecting and interposing on control transfers between the kernel and its modules is challenging. The kernel exposes a complex interface to modules where data is freely shared. The type of data shared by the kernel and its modules can include pointers to both kernel and module code. Sometimes pointers to code sections are directly passed as function arguments; other times, these pointers are deeply embedded in other shared data. Thus, any shared data is a potential opportunity for future control transfers between the kernel and its module. Further, any shared code pointer is a potential opportunity for a malicious module to have the kernel inadvertently execute a non-exported kernel function on behalf of the module.

%\section{Approach}
%Our policies detect the use of non-exported kernel functions by modules and allow kernel developers to securely check and sanitize inputs to exported kernel functions. The mechanism used to enforce these policies is a combination of DBI and static/dynamic interface wrappers. DBI is used to interpose on all control flow transfers within kernel modules and between the kernel and its modules. The static and dynamic interface wrappers are used to attach and detach instrumentation, as well as to discover new transfer points between the kernel and its modules.

%\begin{figure}[h]
%\begin{center}
%TODO
%\includegraphics[width=4cm]{drk-flow.pdf}
%\end{center}
%\caption{Example flow of control through \sysname. Native kernel code invokes module code through module wrappers. Modules wrappers attach DRK and begin execution of the instrumented module within the code cache. The instrumented module eventually calls a kernel function and is indirected through a kernel wrapper, which detaches DRK, thus resuming native code execution.}
%\end{figure}

%\section{Approach}
%This section describes DRK and how it is used to isolate modules and interpose on control transfers between the kernel and the isolated modules.

%\pagebreak[4]
%\pagebreak[4]
%\section{OLD}

%\subsection{Overview}
%\begin{enumerate}
%	\
%	\item Direct control transfers to known illegal addresses are detected when code cache blocks are constructed.
%	\item Indirect control transfers from an instrumented module to the kernel will be detected from within the DRK dispatcher.
%	\item Indirect control transfers from the kernel to the uninstrumented module will be detected using page protection.
%	\item Arguments to kernel exported functions will be inspected to ensure that no pointers to non-exported kernel functions are passed into the kernel.
%\end{enumerate}

%\Sysname isolates kernel modules into implicit protection domains: the code caches of DRK. When executing instrumented module from within DRK's code cache, we detect illegal control-flow transfers in four ways:

%\section{Implementation}
%\Sysname isolates modules in three conceptual steps. The first step introduces kernel wrappers. The second step introduces module wrappers for module initialization and teardown. The final step dynamically introduces module wrappers as module functions are exposed to the kernel through the kernel wrappers.

%The purpose of wrappers is to attach and detach DRK to/from the module binary. While DRK is attached, the executing code is said to be isolated. Isolation is achieved by transparently checking all control flow transfers and injecting runtime checks into the code. Isolated execution stops when control transfers from isolated code (i.e. the module) to the kernel. Circumventing isolation entails control flow between the kernel and the uninstrumented module that bypasses the wrappers. Bypass attempts are detected using page protection and range detection. These techniques, as well as the wrappers, are described below.

%Kernel wrappers serve three primary purposes:
%\begin{enumerate}
%	\item They sanitize data being sent from the instrumented module into the kernel. Sanitization is important. For example, a malicious module can coerce the kernel into calling a non-exported function on its behalf by giving the kernel a pointer to a non-exported function. Under normal circumstances, \sysname would not detect such an attack, as it would be control flow initiated by the kernel. However, by invoking \texttt{WRAP\_FUNC} on function pointers (indirectly) passed down to the kernel, we can protect against such an attack by requiring that all function pointers either address module code or kernel exported functions.
%	\item They wrap module function pointers being (indirectly) passed into the kernel. Wrapping module code addresses is important to maintain isolation. If the kernel were made aware of a ``raw'' module code address then isolation would be circumvented, precisely because there would be an opportunity for the module to be executed natively.
%	\item They represent instrumentation ``detach'' points, where execution transitions from instrumented to native.
%\end{enumerate}

%\subsection{Module wrappers} 
%Conceptually, there are two kinds of module wrappers: ``root'' wrappers, and derived/dynamic wrappers. There are two root wrappers, one for module initialization, and one for module teardown. There are potentially as many dynamic wrappers as there are functions in the module. Root and dynamic wrappers differ only in how and when they are introduced.

%When \sysname is loaded, a function is registered with the kernel that is called each time a new module is loaded. \Sysname replaces the the kernel-exposed addresses of a module's \texttt{init}, \texttt{exit}, and exported functions with unique wrappers when the module is loaded but before it is initialized. Within the \texttt{pre} and \texttt{post} kernel function wrappers, the \texttt{WRAP\_FUNC} macro automatically detects when a ``raw'' module function address is being exposed to the kernel, and replaces it with the address of the unique wrapper for that function.

%\begin{figure}[h]
%\begin{center}
%TODO
%\includegraphics[width=4cm]{drk-shadow-module.pdf}
%\end{center}
%\caption{Runtime memory layout of modules and their shadows. Module memory addresses are a constant distance (\texttt{0x20000000}) from their shadow addresses, enabling quick computation of a shadow address from a raw address, or a raw address from a shadow address.}
%\label{fig:shadow_layout}
%\end{figure}

%Wrapper uniqueness is ensured by partitioning the module address space into two segments, as show in \Figref{shadow_layout}. Kernel modules are loaded within the low range, and module wrapper functions are dynamically inserted within the high range. There is a one-to-one correspondence between raw module function addresses and wrapper addresses, thus the region containing module wrappers is termed the ``shadow'' module.

%Each time a raw module function address is discovered, the same 11 bytes of code are written into the shadow module (\Figref{shadow_template}) at a fixed offset from the raw module address. The 11 bytes of code represent the module function's wrapper. When called by kernel code, the wrapper code pushes the address of its second instruction (\texttt{push}) onto the runtime stack. From this address, \sysname is able to compute the intended destination address (raw module address), and jump to that address after attaching DRK.

%When wrapped kernel functions apply the \texttt{WRAP\_FUNC} macro to function pointers

%\subsection{Changes to DRK}
%\Sysname is implemented as a DRK client. Typically, DRK and its clients instrument all kernel and module code. This is done by attaching and detaching DRK at the system call boundary. One of the goals of \sysname were to not impose any performance overheads when running non-module kernel code. As such, changes needed were made to DRK that changed is default initialization behaviour. As well, changes to the DRK ``dispatcher'', which is responsible for instrumented control flow transfers,

%Module wrappers will be introduced at module load time. When an module is loaded, it invokes special kernel functions to expose function pointers to the kernel. These function pointers are the initial or root entry points to the module. Instead of invoking these module functions, the kernel will actually invoke proxies to these functions, which enter into the DRK code cache.

%\paragraph{Dynamic wrappers} will be introduced when pointers to module code are passed from the module to the kernel through kernel wrappers. Dynamic wrappers behave like normal module wrappers, but are created when they are discovered.

%\subsection{Enforcing Control Flow Integrity}
%Once our isolation mechanism has been implemented, we will detect control flow integrity violations in two ways.

%\paragraph{Page protection} will be used to detect attempts to circumvent isolation by jumping to raw module code. The uninstrumented module will remain in memory as non-executable code. Any attempt by the instrumented module or by the kernel to execute code from the uninstrumented module will result in a page fault.

%\paragraph{Range detection} will be used in the DRK dispatcher (invoked for indirect jumps) and in the DRK compiler (invoked when compiling basic blocks to instrumented x86) to detect attempts to circumvent isolation by jumping to raw kernel code. The only kernel functions that should be available to a module should be exported functions or functions from other instrumented modules. However, because of our role in the module linking phase, those functions are never exposed to the module; instead, kernel wrappers are exposed. As such, anything in the range of valid code addresses for the kernel is considered ``out of bounds" for modules.

%\subsection{Challenges}

%\paragraph{Return address consistency} Maintaining consistent return addresses from interrupts and kernel function calls is challenging for several reasons:
%\begin{enumerate}
%	\item Modules can be interrupted at any point during their execution. In handling an interrupt, the module's thread can be re-scheduled to another CPU. This implies that it is insufficient for CPU-private code cache addresses to be used as return addresses on the runtime stack.
%	\item DRK must detach when executing interrupt handlers or kernel function calls. This implies that a return from an interrupt or from a kernel function call must re-attach DRK where execution should logically re-begin in the raw module.
%\end{enumerate}

%Return address consistency is complicated by running one's OS on a multicore machine. \Sysname only supports single core; however, we have plans on supporting multicore systems in the near future. For single core systems, we have addressed the issue of kernel functions return address consistency; however, interrupt return address consistency is as-of-yet unaddressed. This is not an issue now because there is no possibility for a thread to be re-scheduled to another CPU.

%Our plan for supporting return address consistency is to maintain a private stack of return addresses at the bottom of each kernel thread's runtime stack. This stack will contain raw module addresses that should logically be targeted by interrupt/function returns.

%\paragraph{Direct memory access} DMA represents a threat to \sysname. For example, a device with DMA support could be instructed subvert the integrity of the kernel in a way that \sysname could not intercept. We have no plans to stop such attacks.

%because modules can be interrupted at any point during their execution and because interrupts are handled by kernel code. When an instrumented module is interrupted, \sysname must be able to detach instrumentation, execute the kernel's interrupt handler natively, re-attach instrumentation, and yield execution back to the instrumented module. This process is complicated by the possibility that in the process of handling an interrupt, a module can be rescheduled to another CPU. DRK code caches are CPU-private, and so the return address from an interrupt is not guaranteed to be accurate.

%We have not yet implemented a solution to this problem; however, our approach will involve maintaining a stack of raw modules addresses and storing interrupt and return addresses on this stack.

%\paragraph{Function return address consistency} The consistency of instrumented function return addresses on the runtime call stack is challenging because of CPU-private code caches. Return addresses stored on the stack cannot point to code cache addresses, lest the module's execution be re-scheduled to another CPU. Our approach to handling this issue will be based on DRK's current method of handling function call returns.

%\paragraph{Module-to-kernel return address consistency} The consistency of the module-to-kernel return address is an important security concern and represents an early effort toward stopping return-oriented attacks against the kernel. Our approach to handling this issue will be to store the expected return address on a shadow stack, and to check the shadow stack against the runtime stack when control returns from the module to the kernel.

%\paragraph{Attach/detach consistency} Correctly handling the attach/detach protocol over the kernel-to-module interface, especially in the presence of shared data, is challenging. When module function pointers are passed down into the kernel, we have assumed that we will have the ability to detect and ``wrap" these pointers. However, a module might pass a pointer to a kernel-exported function down to the kernel. In this case, such a pointer will need to be ``unwrapped". Our approach to handling this issue will be based on range checking, i.e. seeing if an address falls in the range of addresses containing DRK code caches, in the range of wrappers, or in the range of kernel code.

%\section{Evaluation}
%We tested \sysname using a simple ``hello world'' module, a module that prints kernel symbols, the \texttt{e1000} network device driver, and the \texttt{tg3} network device driver.

%To evaluate \sysname, we isolated three different modules, a simple ``hello world'' module which walks through the kernel symbol table and prints kernel symbols, the \texttt{e1000} network device driver, and the \texttt{tg3} network device driver.

%We evaluated \sysname in both a Guest virtual machine (QEMU) and natively on a 2.93GHz Intel Core\texttrademark 2 Duo CPU. All of our tests used Linux 2.6.32.

%\subsection{Hypothesis}
%Our evaluation will test three hypotheses:
%\begin{enumerate}
%	\item DBI can be used to enforce security policies.
%	\begin{enumerate}
%		\item Security policies can be enforced with similar performance penalties to systems like \emph{LXFI} and \emph{XFI}.
%		\item Security policies can be enforced on arbitrary kernel modules.
%	\end{enumerate}
%	\item Instrumenting only kernel modules imposes less overhead compared to instrumenting the entire kernel and all of its modules.
%	\item Our tool is effective at isolating kernel modules and preventing them from calling non-exported kernel functions.
%\end{enumerate}

%\subsection{Results}
%\paragraph{Security}
%\Sysname can be used to enforce various security policies in the kernel space. However to back our claim we still need to evaluate \sysname against kernel exploits residing in the modules. We describe some possible approaches to security by means of type deduction, type checking, ownership tracking, and capabilities in our future work section. \texttt{tg3} also proved that we detect invocations of unexported kernel functions. For example, we detected the use of the unexported function \texttt{x86\_swiotlb\_alloc\_\-coherent}.

%\paragraph{Isolation} 
%\Sysname promises to achieve complete isolation between the kernel and the modules. Our isolation testing focused on the \texttt{hello world} module, \texttt{tg3} and \texttt{e1000} network device driver. As described earlier the \texttt{hello module} was a simple module which walks through the symbol table and prints all the kernel symbols. For \texttt{tg3} and \texttt{e1000} network drivers, we verified that all module functions that were downward-exposed to the kernel were correctly wrapped by \sysname and no code for these device drivers are getting executed natively by a simple and unique technique. The verification required manually inserting function calls to the the address \texttt{0xFFFFFFFFDEADBEEF} as if it were the address of \texttt{printk} into every \texttt{tg3} and \texttt{e1000} function. Within DRK's dispatcher, we replaced \texttt{0xFFFFFFFFDEADBEEF} with the address of \texttt{printk}. Wrapped functions running from the code cache successfully invoked \texttt{printk}, whereas unwrapped module functions getting executed natively invoked by the kernel resulted in segmentation faults. By creating more type wrappers and directing \sysname to wrap more exported kernel functions, we were able to achieve full isolation ensuring all execution of device driver code happens from code cache. Extracting type information from the kernel and generating the type wrappers for the kernel function was addition task as we have done it manually. Both \texttt{tg3} and \texttt{e1000} uses mostly the same kernel functions so isolating one of them solves our purpose for the other but as our future work we need to write a \texttt{gcc} extensions which will do the job of extracting type information for the kernel during build.

%\paragraph{Performance} To evaluated the performance overhead incurred by \sysname, we run \texttt{netperf}\cite{netperf_hp} to exercise the linux e1000 driver as the kernel module. We did our evaluation in the guest virtual machine (QEMU) instead of ``bare metal" device since we encountered an issue related to kernel DMA operation when running netperf with \texttt{tg3} driver on the native system. As our experimental setup we run \sysname in Guest Virtual machine (QEMU) running Linux 2.6.32 on a Host system equipped with an Intel(R) Core(TM) i7-860X 2.80 GHz CPU. For the other side of network we used two different setup: a) we run the netserver on the Host Machine and b) we used other machine equipped with 2.93GHz Intel Core\texttrademark 2 Duo CPU and running Linux 2.6.32 and is located on the same subnet mask running the server. To design our testing strategy we took the reference from \emph{LXFI} who also uses netperf to evalute their system. In our description we uesd ``TX" to represent the VM running \sysname sends packets and ``RX" to represent the VM running \sysname receive packets from the network.

%\begin{table*}[htbp]
%\begin{center}
%\begin{tabular}{|l|r|r|r|r|}
%\hline
%\multicolumn{ 1}{|c|}{Test} & \multicolumn{ 2}{c|}{Throughput} & \multicolumn{ 2}{c|}{CPU \%} \\ \cline{ 2- 5}
%\multicolumn{ 1}{|c|}{} & \multicolumn{1}{c|}{Without Isolation} & \multicolumn{1}{c|}{With Isolation} & \multicolumn{1}{c|}{Without Isolation} & \multicolumn{1}{c|}{With Isolation } \\ \hline
%TCP\_STREAM TX & 463.19M bits/sec & 344.95M bits/sec & 97.7 & 99.8 \\
%TCP\_STREAM RX & 452.85M bits/sec & 331.33M bits/sec & 97.8 & 98.1 \\
%UDP\_STREAM TX & 9.02M bits/sec & 6.91M bits/sec & 100 & 100 \\
%UDP\_STREAM RX & 9.36M bits/sec & 7.01M bits/sec & 100 & 100 \\ \hline
%\hline
%TCP\_RR & 9524.04 bits/sec & 7776.16 bits/sec & 96.2 & 96.9 \\
%UDP\_RR & 8860.22 bits/sec & 7186.20 bits/sec & 99 & 99.9 \\ \hline
%\end{tabular}
%\caption{Performance of \texttt{netperf} benchmarks with and without isolation enabled for e1000 driver}
%\label{}
%\end{center}
%\end{table*}

%Table 1 and Table 2 show the results for two different setup. Each test runs for 10 seconds and the ``CPU \%" column represents the CPU utilization by the VM running \sysname. The CPU utilization term suits better when tested with system running on ``bare metal" as incase of Vitual machine this number is not very reliable. In case of virtual machine all the network processing happens outside the context of the virtual guest and we can't acturatly say how much overhead \sysname is causing in terms of CPU utilization. In our case also the CPU utilization is always above 90\% except in the case when server runs on the system other than Host on same subnet mask.

%In the table 1, the ﬁrst test, TCP STREAM, measures the TCP throughput of the e1000 driver. The test uses a send buffer of 16,384 bytes, and a receive buffer of 87,380 bytes. The message size is also set as 16,384 bytes. The table shows the throughput for both ``TX" and ``RX" kind of workloads. The drop in the throughput of e1000 driver because of isolation is close to 25\% and there is a marginal increase in the CPU utlization as it was already close to maximum.

%The UDP STREAM measures UDP throughput. In our experiment the UDP socket size was 124,928 bytes on the send side, and 116,736 bytes on the receive side. The test sends messages was set as 64 bytes and the two performance number reports the throughput for send and receive. The testing is done for both the types of workloads ``TX" and ``RX" and we see a performance drop between 20-25\% incase of UDP also. The CPU utilization in both the cases was always maximum. we also run the TCP\_RR and UDP\_RR script to measure the impact of \sysname on the latency  using a message size of 1 byte. We see a drop in througput close to 20\% incase of both TCP and UDP connections.

%In the other setup we used the similar strategy to evaluate the system. The table 2 summerizes the test result for the setup. We are only reporting the effect of \sysname on latency using a message size of one byte. Incase of both the connections we see a drop of below 15\% in the throughput and an increase of 25\% in the CPU utilization due to \sysname. In the second setup the average drop in the throughput is less compared to the first one and this is because the CPU utilization was well below the maximum and there was scope for the system to generate more packets. A more fare conclusion could have been drawn if we have evaluated \sysname with running it on ``bare metal".

%To understand the source of overhead in \sysname, a contribution comes due to frequent context switching between the kernel and the modules incase of \texttt{e1000} device driver. For a control transfer from the kernel to the modules the control goes through the shadow module region and switches to dynamorio context before putting the basic block into code-cache and for a control transfer from the module to the kernel the context switch happens from dynamorio context to the native and then it goes through the kernel interface layer to wrap the kernel API's. Although we evaluated \sysname in VM, the overhead we are getting is according to our expectation and the added advantage we are getting is the security policy which we will implement with cause no overhead when kernel is running.

%\section{Future Work}
%Our first goal is to make \sysname work with on multicore computers. This means addressing the issue of interrupt return address consistency, among other interrupt-related issues. There are a number of places where global data is shared in \sysname for the sake of convenience; different, more local strategies will need to be employed for \sysname to work on multiple CPUs. We have previously prototyped one such strategy with success; however, due to time constraints and feature requirements, we eventually removed that code.

%Our second goal is to further automate the process of wrapping kernel exported functions. Right now, one must manually collect the necessary type information and kernel-exported function signatures into a single file. We think that a compiler extension would be particularly well suited to the task of generating the necessary type information while as a side-effect of building the Linux kernel.

%Increasing the number of wrapped functions will enable type checking of kernel exported functions. Our security policy is that modules should not be aware of non-exported kernel functions. We enforce this policy by ensuring that no non-exported function is invoked (unless it is given by the kernel to the module), and that no non-exported function is passed as a function pointer argument to a kernel-exported function. There might, however, be a case where giving an exported function's address to the kernel could compromise the kernel's integrity. We think that such scenarios can be partially mitigated by type checking all exported kernel functions being given to the kernel. This necessitates full type information about all kernel-exported functions.

%Our third goal is to extend the responsibilities of \texttt{pre}/\texttt{post} wrappers. The use of C++ as a way to do type-based wrapping has worked very well for us. We think that we can further extend this feature to do runtime type assignment, type checking, object ownership, and capability passing. For example, direction of data across the kernel interface is a strong indicator of ownership. The \texttt{pre} and \texttt{post} wrappers give us the opportunity to discover the direction that data is moving, as well as the differences between the same data before and after the kernel function is called. This information can be used to assign ownership at the granularity of data structures (e.g. pointer to a \texttt{device\_driver} is passed to the kernel; deduction: the module owns this data) or at a finer granularity (e.g. the kernel changes the \texttt{pm} field of the \texttt{device\_driver} structure; deduction: the kernel owns this field).

%Our final goal involves making tools to help kernel developers debug kernel modules.

%\begin{table*}[htbp]
%\begin{center}
%\begin{tabular}{|l|r|r|r|r|}
%\hline
%\multicolumn{ 1}{|c|}{Test} & \multicolumn{ 2}{c|}{Throughput} & \multicolumn{ 2}{c|}{CPU \%} \\ \cline{ 2- 5}
%\multicolumn{ 1}{|c|}{} & \multicolumn{1}{c|}{Without Isolation} & \multicolumn{1}{c|}{With Isolation} & \multicolumn{1}{c|}{Without Isolation} & \multicolumn{1}{c|}{With Isolation } \\ \hline
%\hline
%TCP\_RR & 3620.88 bits/sec & 3151.05 bits/sec & 42.99 & 51.3 \\
%UDP\_RR & 3579.55 bits/sec & 3147.20 bits/sec & 46.83 & 56.2 \\ \hline
%\end{tabular}
%\caption{}
%\label{}
%\end{center}
%\end{table*}

%\subsection{Milestones}
%This section describes our project milestones. Our milestones are based on incrementally implementing and gaining experience with the system.
%\begin{enumerate}
%	\item Implement a limited form of attach and detach using page protection and range detection. Page protection will protect the module binary from execution. When the kernel attempts to jump into the module, a page fault will occur, at which point we will attempt to attach DRK. Once attached, range detection will be used in the DRK dispatcher to attempt to detach when the module wants to return to the kernel. Importantly, if the module calls a kernel function or if an interrupt occurs then DRK will remain attached and parts of the kernel will be instrumented.

%\item Investigate how to incrementally add simplified kernel and module wrappers. Again, we will not attempt to detach when kernel code is executed within the scope of a module.

%\item Investigate and implement detach for interrupt handlers.

%\item Attempt to apply similar techniques to the case where the instrumented module calls the kernel, allowing for interleaved control flow of instrumented and non-instrumented code. Importantly, page protection will remain as a back up form of attaching instrumentation in the event that stray function pointers into module code are stored by the kernel.

%\item Extend the functionality of our wrappers to include checking and (un)wrapping function pointers.

%\item Depend on wrappers for maintaining module isolation and on page protection and range checking for detecting attempts to circumvent isolation.
%\end{enumerate}

%\section{Evaluation}
%\subsection{Hypotheses}
%Our evaluation will test three hypotheses:
%\begin{enumerate}
%	\item DBI can be used to enforce security policies.
%	\begin{enumerate}
%		\item Security policies can be enforced with similar performance penalties to systems like \emph{LXFI} and \emph{XFI}.
%		\item Security policies can be enforced on arbitrary kernel modules.
%	\end{enumerate}
%	\item Instrumenting only kernel modules imposes less overhead compared to instrumenting the entire kernel and all of its modules.
%	\item Our tool is effective at preventing modules from calling non-exported kernel functions, getting the kernel to call such functions on the module's behalf, and limits return-oriented attacks against the kernel.
%\end{enumerate}

%\subsection{Experiments}
%We will test hypothesis 1.a by using \emph{LXFI}'s \cite{YMao2011} evaluation strategy to design our experiment. We will use the SFI Micro-benchmarks to measure the security enforcement overhead caused by \sysname. \emph{LXFI} and \emph{XFI} use three SFI Michrobechmarks (\emph{hotlist}, \emph{lld}, and \emph{MD5}) to evaluate their systems. To evaluate the performance overhead when isolating kernel modules, \emph{LXFI} runs \emph{netperf} to exercise the Linux e1000 driver as the kernel modules. We will test hypothesis 1.b by running netperf with several other network drivers and modules.

%We will test hypothesis 2 by comparing the performance of \sysname to that of an unmodified version of DRK running a null client on common benchmark suites such as \emph{SPECweb}, \emph{SPECsfs}, and \emph{SPECmail}. Deriving meaningful conclusions from differences in performance between an unmodified version of DRK and \sysname might be difficult; however, we hope to find out if the cost of module isolation is outweighed by the speculated benefit of not instrumenting the entire kernel. Further, we hope that a comparison between \sysname, baseline DRK, and native Linux will help provide evidence for our claim that isolation can be2.93GHz Intel Core\texttrademark 2 Duo CPU achieved with no performance overhead when running kernel code.

%We will test hypothesis 3 by finding existing modules or creating modules that behave maliciously and attempt to circumvent our isolation mechanism. If time permits, we will also attempt to find existing malicious modules (e.g. rootkits) that attack the kernel in ways beyond the scope of this project and measure to what extent they are undetectable by \sysname.

%\section{Related Work}
%\Sysname draws inspiration from a variety of past research, including kernel integrity protection, kernel rootkit analysis, device driver isolation, and mandatory access control models. The goals of previous research has been to protect the kernel against, or detect the presence of, attacks that compromise the integrity of the kernel's code \cite{Srivastava_Giffin_2011, Seshadri2007}, data \cite{Srivastava2009, Baliga2008}, or control-flow \cite{Hofmann2011, Wang2009}.

%\subsection{Type Checking}
%\emph{OSck} \cite{Hofmann2011} is a hypervisor-based kernel monitoring system that uses dynamic type checking and ad-hoc checks to identify the presence of kernel rootkits. Like the virtualization-based approaches listed below, \emph{OSck} is not appropriate for non-virtualized device drivers. It's not clear to what extent \emph{OSck} depends on the availability of type information for modules, nor how a lack of availability of such information would increase the detection of false positives. Regardless, \emph{OSck} has inspired our approach in two ways. First, our wrappers  make heavy use of type information present in the kernel in order to properly detect and check for function pointers. Second, we implement similar ad-hoc integrity checks within our wrappers.

%\subsection{Externalization to User Space}
%\emph{Microdrivers} \cite{Ganapathy2005} splits modules into kernel and user space components using programmer annotations. This requires an effort from programmers to decide which part of their module they want to run in user-space. This system does not achieve full isolation as a part of a micro-driver module runs in kernel space. The performance overhead of this approach is another concern. \emph{Nexus} \cite{Williams2008} enforces stronger security policies than \emph{Microdrivers} but it requires additional hardware support.

%\subsection{Virtualization}
%Systems like \emph{Gateway} \cite{Srivastava_Giffin_2011} and \emph{HUKO} \cite{Xiong2011} use virtualization to establish strong isolation guarantees between kernel components. Unfortunately, neither system is able to isolate native device drivers, which are often distributed as loadable binary modules.

%\subsection{Hardware Protection Domains}
%\emph{Nooks} \cite{Swift2005} and \emph{Mondrix} \cite{Witchel2005} isolate modules while running them in the kernel address space. Isolation is achieved using hardware-enforced protection domains. \emph{Nooks} was the first system to provide module isolation using hardware protection but it does not protect the kernel from malicious modules. \emph{Nooks} also suffers from high performance overheads. \emph{Mondrix} provides isolation with low overhead but requires special hardware support. \emph{Mondrix} also does not protect the kernel from abusive use of the module interface.

%The design of \sysname's wrappers are heavily influenced by \emph{Nooks}; however, the mechanism used to achieve isolation is different. In our case, wrappers are used to attach and detach DRK. In the case of \emph{Nooks}, wrappers exist in an ``intermediate" zone where hardware protection can be added or removed to/from the kernel. In the case of \emph{Nooks}, there is nothing stopping a module from changing what memory the hardware protects.

%\subsection{Software Fault Isolation}
%\emph{BGI} \cite{Castro2009}, \emph{XFI} \cite{Erlingsson2006}, and \emph{LXFI} \cite{YMao2011} enforce software fault isolation (SFI) by implementing memory protection through access control and static and dynamic runtime checks. \emph{BGI} and \emph{LXFI} enforce fine-grained memory access permissions at the granularity of bytes. \emph{XFI} maintains coarse grained access permissions on some ranges of contiguous memory. \emph{XFI}, \emph{BGI}, and \emph{LXFI} are successful at isolating faults but lack transparency. \emph{XFI} depends on debugging information present in module binaries so that its binary rewriter can inject runtime checks, whereas \emph{BGI} and \emph{LXFI} require a special compiler to compile both the kernel and its modules. This compiler inserts permission checks before every memory write and control transfer operation.

%\section{Discussion}
%Our work relates to the material covered in class in terms of security concerns within OSes.

%\paragraph{Capabilities} \emph{HYDRA} \cite{Wulf1974} enforces object permissions using a capability-based system. Capabilties determine what actions are available on certain types of objects and to whom. To some extent, the state of a kernel function being exported is a capability of a module. Through this lens, \sysname attempts to detect attempts to execute code to which modules lack the capabilities.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\bibliography{library}

\end{document}

### 2013-06-19 OSDI'12/poster.tex
Commit of some old papers and in-progress PLOS.

\documentclass[10pt,preprint,nocopyrightspace]{sigplanconf}
\usepackage{amsmath,epsfig,subfigure,multirow}

\usepackage{graphicx}
\usepackage{refstyle}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{framed}
\usepackage{lipsum}
\usepackage{multicol}
\usepackage[english]{babel}
\usepackage{numprint}
\usepackage[dvipsnames]{color}
\usepackage{footmisc}
\usepackage{relsize}
\definecolor{shadecolor}{RGB}{250,250,250}

\renewcommand\footnotelayout{\larger}
\long\def\symbolfootnote[#1]#2{\begingroup%
\def\thefootnote{\fnsymbol{footnote}}\footnote[#1]{#2}\endgroup}

\begin{document}

%\conferenceinfo{OSDI'12}{Feb. 23, Shenzhen.} 
%\copyrightyear{2012} 
%\copyrightdata{[to be supplied]}

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Granary: Comprehensive Kernel Module Instrumentation}
\subtitle{University of Toronto}

\authorinfo{Peter Goodman*\and Angela Demke Brown}
           {\{pag, demke\}@cs.toronto.edu}

\authorinfo{Akshay Kumar*\and Ashvin Goel}
           {\{akshayk, ashvin \}@eecg.toronto.edu}

\maketitle

\renewcommand{\secref}[1]{\S\ref{sec:#1}}

% until we find a name for our system, the following commands should suffice :-)
\newcommand{\sysname}{Granary}
\newcommand{\Sysname}{Granary}

\newcommand{\note}[1]{}

Kernel modules extend the functionality of operating systems (OSes). Modules are used to support new devices (e.g. network and graphics cards) and provide new features (e.g. file systems). The kernel and its modules execute in a complex and dynamic environment. Understanding how modules behave in and affect this environment is important. However, analyzing module behavior is challenging. Static analysis of module source code is difficult because of the tight interaction between modules and the kernel. Some modules, however, are only distributed in a binary format, which makes static analysis intractable.

We created {\Sysname} to address the challenges of module analysis. {\Sysname} is a framework that efficiently instruments \emph{arbitrary}, binary Linux kernel modules. {\Sysname} uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules. Our extensive use of compile-time meta-programming enables efficient, dynamic analyses that are driven by static kernel type information.

While designing {\sysname}, we identified four goals for practical module analysis: i) comprehensively analyze \emph{all} modules; ii) impose no performance overheads on non-module kernel code; iii) require no changes to modules and minimal changes to the kernel, and; iv) be easily portable between different hardware and kernel versions. Prior research based on source code analysis and annotations \cite{YMao2011} fails to meet goals (i) and (iii), while work based on special hardware features or virtualization \cite{Xiong2011} fails to meet goals (i) and (iv), and work based on whole-OS or -system instrumentation/emulation \cite{Feiner2012} fails to meet goal (ii).

{\Sysname} meets all four stated goals:
\begin{enumerate}[i)]
	\item {\Sysname} is comprehensive because it controls and instruments the execution of all module code. {\Sysname} maintains control by ensuring that normal module code is never executed. Instead, only decoded and translated module code is executed. Translated module code contains instrumentation and always yields control back to {\sysname}. All modules can be instrumented in this way because dynamic binary translation operates on binaries and does not depend on any hardware features.
	\item Kernel code runs without overhead because {\sysname} relinquishes its control whenever an instrumented module executes kernel code. {\Sysname} implements a novel technique for re-gaining control when kernel code attempts to execute module code. Each time some instrumented module code invokes a kernel function, each of that function's arguments are \emph{wrapped}. Argument wrappers are type- and function-specific, and ensure that potential module entry points (e.g. module function pointers) are replaced with behaviorally-equivalent values that first yield control to {\sysname}.
	\item We have changed less than 100 LOC in the Linux kernel in order to support {\sysname}.
	\item {\Sysname}'s wrapping mechanism is portable across different kernel versions because the majority of wrappers are automatically generated by a GCC plugin and several meta-programs.
\end{enumerate}

%{\Sysname} maintains control by executing only translated versions of short sequences of module instructions ending in control-flow transfers--called \emph{basic blocks}. Translated basic blocks have the same execution behaviour as native module code, but contain additional instrumentation code and always yield control back to {\Sysname}.

%{\Sysname} enables analysis of arbitrary kernel modules by using DBT instead of special hardware features like virtualization or protection domains. DBT allows for comprehensiveness by controlling the execution of all module code. , only translated versions of short sequences of module instructions ending in control-flow transfers--called \emph{basic blocks}--can be executed. Translated module code behaves in the same way as normal module code, but contains additional instrumentation code. The additional instrumentation code enables precise module analysis.

%We identified four features needed to make module analysis practical

%The existing solution which attempts to isolate the kernel modules and understand its interactions with the kernel uses different approach like compile-time and run-time checking/annotations, hardware protection domain, and by extending virtual machines (VMs) and hypervisors with the ability to introspect and monitor a running kernel and its modules. While these approaches have already achieved many of the goals, all suffer from at least one of the following issues: i) they depend on VMs/hypervisors, and thus limit the breadth of kernel modules which can be monitored, or; ii) they require a custom compilation toolchain and/or kernel modifications, and so limit the likelihood of their integration with existing OS build processes.

%In this work we present {\sysname}, a framework for efficiently instrumenting Linux kernel modules. {\sysname} uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules. The extensive use of compile-time meta-programming driven by static type information enables us to efficiently analyze its dynamic behaviour. Our approach is novel because {\sysname} can instrument arbitrary, binary kernel modules without imposing any overhead on non-instrumented kernel code.

%The goal of comprehensivness requires that {\sysname} control the execution of all module code. This is achieved by attaching instrumentation when module code first gets executed and leaving it attached until the next context switch to kernel code. The second goal requires that any overhead introduced by {\sysname} only affect instrumented modules. The {\sysname} introduces overhead when it translated, caches, and executes instrumented code leaving kernel execution unaffected. The attaching and detaching of the instrumentation framework causes the execution to switch between native and instrumented mode. This makes it challenging to maintain the module isolation and consistency of the interrupt and exceptions.

%Instrumenting only module code requires that {\sysname} should be able to dynamically gain control on all raw kernel module entry point and relinquish control when control transfers to the kernel. Interposing on the control-flow transfers leading to mode switch is easy when module executes under the control of {\sysname} but is challenging when kernel code executes. Page protection is one method that can be used to detect the control-flow transfer to the modules, where all modules code can be made non-executable and any attempt to execute them results in a page fault exeception that will be handled by {\sysname} to attach instrumentation. Considering the cost associated with page-fault handling, {\sysname} takes an alternative approach where it interposes during module loading mechanism changing the module entry point, allowing for efficient entry point interposition.{\sysname} uses a type based kernel wrapper which pro-actively wraps all the associated function pointers passed to the kernel making it easy for itself to gain control when kernel calls raw module. However, wrapping and re-wrapping of these function pointers deeply nested into the kernel objects each time on context switch makes it expensive. {\sysname} tracks the kernel object passed over the kernel-module interface and embedding the meta-information of its last state and enabling the re-wrapping of the kernel objects on its state change. Type based kernel wrapper and dynamic tracking of the kernel objects enables the scope of dynamic wrapper for {\sysname}, which helps it to analyze the different instance of the kernel object differently. {\sysname} uses page-protection as a fall-though which ensures that raw module never gets executed.

{\Sysname} is a work in progress. It works on multi-core processors with pre-emptive kernels, and incurs a modest decrease in throughput of 10\% to 50\% for network device drivers. We have used {\sysname} to isolate and comprehensively instrument several network device drivers (e1000, e1000e, ixgbe, tg3) and file system modules (ext2, ext3). We have used {\sysname} to develop an application which enforces partial control-flow integrity policies. These policies disallow modules from executing dangerous control-flow transfers. As a future work, we plan on implementing more optimizations and applications.

\bibliographystyle{abbrvnat}
\bibliography{library}

\symbolfootnote[0]{* Student}

\end{document}

### 2013-06-19 OSDI'12/bibtex/expand_mendeley_macros.py

@@ -1,0 +1,29 @@
+#!/usr/bin/env python
+
+import json
+import re
+import sys
+
+if len(sys.argv) != 4:
+    print 'Usage: %s macro_file bib_file output_file' % (sys.argv[0])
+    sys.exit(-1)
+
+def load_macros(macro_file):
+    macros = json.loads(open(macro_file).read())
+    result = {}
+    for macro in macros:
+        result[macro.encode('ascii')] = macros[macro].encode('ascii')
+    return result
+    
+def substitute_macros(bib_file, macros):
+    '''Replaces {X} with {macros[X]}'''
+    text = open(bib_file).read()
+    for macro in macros:
+        text = text.replace('{%s}' % (macro), '{%s}' % (macros[macro]))
+
+    assert(re.match('\\{\\@.*\\}', text) == None)
+    return text
+
+macros = load_macros(sys.argv[1])
+output = substitute_macros(sys.argv[2], macros)
+open(sys.argv[3], 'w').write(output)

### 2013-07-01 PLOS'13/evaluation.tex
updated a bit of evaluation text

To evaluate our system with existing kernel instrumentation framework, we modified DRK to instrument only kernel modules. We implemented similar wrapping mechanism for DRK to attach and detach the framework at module interface. We checked the correctness of our system by running different filesystem and network modules(ext2/ext3, xfs, e1000e etc) under their control. To stress test the filesystem modules we used filesystem benchmark IOzone and ran the modules with null instrumentation client. null instrumentation client was used to avoid the overhead of instrumentation. We ran our tests on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ i7-860 2.80 GHz CPU, 8GB memory, and an Intel 82578DM Gigabit Ethernet card.

### 2013-07-02 PLOS'13/evaluation.tex
updated evaluation section

@@
-To evaluate our system with existing kernel instrumentation framework, we modified DRK to instrument only kernel modules.
+Granary is designed to comprehensively instrument kernel modules.
+We first evaluated the performance of Granary framework alone by running it with different kernel modules.
+We ran the experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ i7-860 2.80 GHz CPU, 8GB memory, and an Intel 82578DM Gigabit Ethernet card.
+We used IOzone filesystem benchmark to exercise ext3 module while running it with Granary.
+Iozone is a useful benchmark for testing data throughput with number of different access pattern.
+we avoided bottleneck for disk I/O by partitioning main memory to load ext3 module.
+To compare the throughput of Granary with existing kernel instrumentation framework, we modified DRK to instrument only kernel modules.
+We implemented similar attach/detach mechanism for DRK at module interface.
+The framework uses similar kernel wrappers for fast attach and page protection as fall back mechanism to gain control.
+For initial evaluation of Granary, we developed null policy client and enabled some of the optimizations.
+
+%wrapping mechanism for DRK to attach and detach the framework at module interface.
+
+%and jbd modules while running them under Granary.
+
+%To evaluate our system with existing kernel instrumentation framework, we modified DRK to instrument only kernel modules.
 We implemented similar wrapping mechanism for DRK to attach and detach the framework at module interface.

### 2013-07-02 PLOS'13/evaluation.tex
added evaluation result for Iozone

@@
 We first evaluated the performance of Granary framework alone by running it with different kernel modules.
-We ran the experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ i7-860 2.80 GHz CPU, 8GB memory, and an Intel 82578DM Gigabit Ethernet card.
+We ran the experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU, 4GB memory, and an Intel 82598EB 10 Gigabit Ethernet controller.
 We used IOzone filesystem benchmark to exercise ext3 module while running it with Granary.
@@
 The framework uses similar kernel wrappers for fast attach and page protection as fall back mechanism to gain control.
-For initial evaluation of Granary, we developed null policy client and enabled some of the optimizations.
+For initial evaluation of Granary, we developed null policy client and enabled wrap depth optimization by limiting it to 2.
+
+\begin{table*}[thp!]
+\caption{Iozone Throughput for different filesystem operations}
+% title of Table
+\centering
+% used for centering table
+\begin{tabular}{c c c c}
+% centered columns (4 columns)
+\hline\hline
+%inserts double horizontal lines
+Workload & Native & Granary & Modified DRK  \\ [0.5ex]
+% inserts table
+%heading
+\hline
+% inserts single horizontal line
+Write & 184669.69 & 152223.03 & 55134.09 \\
+% inserting body of the table
+Read & 1577591.69 & 1648645.02 & 1421037.96 \\
+Random Write  & 552380.48 & 412069.48 & 162847.34 \\
+Random Read & 1644573.29 & 1597321.54 & 1636472.14 \\
+
+Mixed Workload & 1168804.11 & 929383.54 & 915196.76 \\[1ex]
+% [1ex] adds vertical space
+\hline
+%inserts single line
+\end{tabular}
+\label{table:nonlin}
+% is used to refer this table in the text
+\end{table*}
 

### 2013-07-02 PLOS'13/evaluation.tex
removed two workload testing

@@
 We first evaluated the performance of Granary framework alone by running it with different kernel modules.
-We ran the experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU, 4GB memory, and an Intel 82598EB 10 Gigabit Ethernet controller.
-We used IOzone filesystem benchmark to exercise ext3 module while running it with Granary.
-Iozone is a useful benchmark for testing data throughput with number of different access pattern.
-we avoided bottleneck for disk I/O by partitioning main memory to load ext3 module.
 To compare the throughput of Granary with existing kernel instrumentation framework, we modified DRK to instrument only kernel modules.
@@
 
-\begin{table*}[thp!]
-\caption{Iozone Throughput for different filesystem operations}
+We ran the experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU, 4GB memory, and an Intel 82598EB 10 Gigabit Ethernet controller.
+We used IOzone filesystem benchmark to exercise ext3 module while running it with Granary.
+Iozone is a useful benchmark for testing data throughput with number of different access pattern.
+we avoided bottleneck for disk I/O by partitioning main memory to load ext3 module.
+We enable Iozone to run in througput mode with two active processes and evaluated the system under three different workloads: read, write and mixed.
+
+\begin{table}[thp!]
+\caption{Iozone Throughput for different filesystem operations(KBytes/Sec)}
 % title of Table
@@
 % used for centering table
-\begin{tabular}{c c c c}
+\begin{tabular}{l r r r}
 % centered columns (4 columns)
@@
 %inserts double horizontal lines
-Workload & Native & Granary & Modified DRK  \\ [0.5ex]
+Workload & Native & Granary & Modified Drk  \\ [0.5ex]
 % inserts table
@@
 Read & 1577591.69 & 1648645.02 & 1421037.96 \\
-Random Write  & 552380.48 & 412069.48 & 162847.34 \\
-Random Read & 1644573.29 & 1597321.54 & 1636472.14 \\
+%Random Write  & 552380.48 & 412069.48 & 162847.34 \\
+%Random Read & 1644573.29 & 1597321.54 & 1636472.14 \\
 
-Mixed Workload & 1168804.11 & 929383.54 & 915196.76 \\[1ex]
+Mixed & 1168804.11 & 929383.54 & 915196.76 \\[1ex]
 % [1ex] adds vertical space
@@
 % is used to refer this table in the text
-\end{table*}
+\end{table}
 

### 2013-07-09 PLOS'13/evaluation.tex
added preprint option for page numbers
evaluation updated

--- 2013-07-02 PLOS'13/evaluation.tex
@@
-Granary is designed to comprehensively instrument kernel modules.
-We first evaluated the performance of Granary framework alone by running it with different kernel modules.
-We ran the experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ i7-860 2.80 GHz CPU, 8GB memory, and an Intel 82578DM Gigabit Ethernet card.
+Granary is designed with the motivation to allow mixed-mode execution in Linux kernel.
+This is particularly important to avoid the unnecessary performance overheads incurred by instrumenting whole kernel.
+Granary uses a type-rich wrapper for attach and detach of framework at kernel/module interface.
+Based on the design of Granary, we decided to evaluate it on the following three parameters: 1) Overhead of Granary with null-instrumentation, 2) Advantage of wrappers for fast attach/detach and 3) Advantages of mixed-mode execution in the Linux kernel.
+To evaluate the performance overhead of Granary on kernel modules, we used Iozone filesystem benchmark to exercise the \texttt{ext3} filesystem driver as a kernel modules.
+we run our experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU with 4GB physical memory.
+
+%To focus our evaluation on CPU overhead of Granary and avoid the disk performance bottleneck, we created a ramdisk of 1GB and mounted \texttt{ext3} module on this.
+Iozone creates two active processes each writing to and reading from a file of size 480MB with record size 1kb.
+
+\paragraph{Overhead of Granary with null-instrumentation} For the evaluation of Granary framework alone, we created a granary client with null instrumentation policy.
+This avoids the overhead caused by the instrumentation.
+While evaluation Granary, Our focus was to measure CPU overhead and avoids the disk performnace bottleneck.
+To ensure this we created a ramdisk of 1GB and mounted \texttt{ext3} moudle on this.
+Iozone creates two processes which acts as writer and reader thread and workload includes writing to and reading from a file of size 480MB with record size of 1kb.
+We compared the performance of Granary with the existing framework by modifying DRK to instrument the kernel modules.
+we implemented similar attach/detach mechanism for DRK at the kernel/module interface including similar kernel wrappers for fast attach and page protection as the fall-back mechanism to regain control.
+Figure shows
+
+\paragraph{Fast attach/detach} Granary uses kernel wrappers for fast attach/detach at kernel/module interface and page protection as the fall-back mechanism to regain control.
+The use of kernel-wrappers provides additional advantage by intergrating the static information with the Granary.
+However, in our evaluation we tried to compared the overhead of wrapper against our fallback mechanism of page-protection.
+
+%implemented similar attach/detach mechanism for DRK at the kernel/module interface including similar kernel wrappers for fast attach and page protection as the fall-back mechanism to regain control.
+
+\paragraph{Advantage of Mixed-mode execution} We evalauted the advantage of using mixed-mode execution over full kernel instrumentation by running Iozone with disabling direct I/O effect.
+Disabling direct I/O enables the buffer cache effect for the filesystem operations and all the requests issued by Iozone doesn't go directly to the disk.
+This avoids running filesystem code for every operations.
+Figure clearly shows that for different read workloads the Granary doesn't put overhead because the filesystem code doesn't gets executed.
+However for the DRK which is instrumenting entire kernel including vfs layer and thus causes overhead even when we only try to analyse filesystem code.
+
+ %two configurations : 1) with direct I/O and 2) without direct I/O.
+Direct I/O disables the buffer cache effect and all I/O requests issued by Iozone directly goes to disk and for every such request filesystem code gets executed.
+
+ %settings.
+In first configuration, while running the Iozone benchmark we enabled the direct I/O to the disk.
+This disables the buffer cache for the ramdisk and for every I/O was happening was happening to the disk.
+Second configuration enabled buffer cache.
+The effect of buffer cache is important since it avoids running of the kernel modules and IO happens directly from the VFS layer bypassing filesystem module code.
+
+%Before evaluating the performance overhead of Granary framework over native execution, we
+
+%One of the motivation for the design of Granary is allow the mixed-mode execution of 
+%One of the motivation for the design of Granary is to a
+
+%new system is
+
+%Granary is designed to instrument kernel modules and one of the motivation for
+
+%Instrumenting all code (e.g. the whole kernel) imposes unnecessary
+%performance overheads when only a subset of the code is
+%being analysed (e.g. modules).
+
+%We evaluated the performance of framework alone by running the kernel modules with Granary under null instrumentation policy.
+We ran the experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU with 4GB physical memory.
+For the experiemnt, we used ext3 filesystem module.
+we avoided the disk I/O bottleneck by creating a ramdisk of 1GB and modunting an ext3 filesystem on it.
+This setup allows us to focus on the CPU overhead of Granary, rather than the disk performance.
+We use Iozone version 3.4 to get a fine-grained view of performance on different filesystem intensive workloads.
+The Iozone creates two active processes each writing to and reading from a file of size 480MB with record size 1kb.
+For the initial experiment, we disabled the effect of buffer cache for the device which allows reader process to read data from buffer cache completely bypassing Granary-instrumented filesystem code.
+We compared the performance of Granary with existing instrumentation framework by modifying DRK to instrument the kernel modules.
+we implemented similar attach/detach mechanism for DRK at the kernel/module interface including similar kernel wrappers for fast attach and page protection as the fall-back mechanism to regain control.
+
+%We used Iozone benchmark for read and write intensive workload which sufficiently stresses filesystem module.
+To compare the performance of Granary with existing instrumentation framework, we implemented similar attach/detach mechanism for DRK at the kernel/module interface including similar kernel wrappers for fast attach and page protection as the fall-back mechanism to regain control.
+
+%We ran Iozone benchmark with two active processes each creating a file of \~480 MB with record size of 1kb.
+
+%We first evaluated the performance of Granary framework alone by running it with different kernel modules.
+To compare the throughput of Granary with existing kernel instrumentation framework, we modified DRK to instrument only kernel modules.
+We implemented similar attach/detach mechanism for DRK at module interface.
+The framework uses similar kernel wrappers for fast attach and page protection as fall back mechanism to gain control.
+For initial evaluation of Granary, we developed null policy client and enabled wrap depth optimization by limiting it to 2.
+
+%We ran the experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU, 4GB memory, and an Intel 82598EB 10 Gigabit Ethernet controller.
 We used IOzone filesystem benchmark to exercise ext3 module while running it with Granary.
@@
 we avoided bottleneck for disk I/O by partitioning main memory to load ext3 module.
-To compare the throughput of Granary with existing kernel instrumentation framework, we modified DRK to instrument only kernel modules.
-We implemented similar attach/detach mechanism for DRK at module interface.
-The framework uses similar kernel wrappers for fast attach and page protection as fall back mechanism to gain control.
-For initial evaluation of Granary, we developed null policy client and enabled some of the optimizations.
+We enable Iozone to run in througput mode with two active processes and evaluated the system under three different workloads: read, write and mixed.
+
+\begin{figure}[ht!]
+ \subfloat{
+\epsfig{file=diagrams/iozone-write-workload.eps,width=3.4in}
+ }
+\ORIGcaption{\label{fig:iozone-write-workload}Writer procress throughput under different workloads}
+\end{figure}
+
+%\begin{table}[thp!]
+%\ORIGcaption{Iozone Throughput for different filesystem operations(KBytes/Sec)}
+% title of Table
+%\centering
+% used for centering table
+%\begin{tabular}{l r r r}
+% centered columns (4 columns)
+%\hline\hline
+%inserts double horizontal lines
+%Workload & Native & Granary & Modified Drk  \\ [0.5ex]
+% inserts table
+%heading
+%\hline
+% inserts single horizontal line
+%Write & 184669.69 & 152223.03 & 55134.09 \\
+% inserting body of the table
+%Read & 1577591.69 & 1648645.02 & 1421037.96 \\
+%Random Write  & 552380.48 & 412069.48 & 162847.34 \\
+%Random Read & 1644573.29 & 1597321.54 & 1636472.14 \\
+
+%Mixed & 1168804.11 & 929383.54 & 915196.76 \\[1ex]
+% [1ex] adds vertical space
+%\hline
+%inserts single line
+%\end{tabular}
+%\label{table:nonlin}
+% is used to refer this table in the text
+%\end{table}
 

### 2013-07-17 PLOS'13/evaluation.tex
Edited evaluation section and intro.

--- 2013-07-02 PLOS'13/evaluation.tex
@@
-Granary is designed to comprehensively instrument kernel modules.
-We first evaluated the performance of Granary framework alone by running it with different kernel modules.
-To compare the throughput of Granary with existing kernel instrumentation framework, we modified DRK to instrument only kernel modules.
-We implemented similar attach/detach mechanism for DRK at module interface.
-The framework uses similar kernel wrappers for fast attach and page protection as fall back mechanism to gain control.
-For initial evaluation of Granary, we developed null policy client and enabled wrap depth optimization by limiting it to 2.
+%\begin{figure}[ht!]
+% \subfloat{
+%\epsfig{file=diagrams/iozone-write-workload.eps,width=3.4in}
+% }
+\ORIGcaption{\label{fig:iozone-write-workload}Writer procress throughput under different workloads}
+\end{figure}
 
-We ran the experiment on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU, 4GB memory, and an Intel 82598EB 10 Gigabit Ethernet controller.
-We used IOzone filesystem benchmark to exercise ext3 module while running it with Granary.
-Iozone is a useful benchmark for testing data throughput with number of different access pattern.
-we avoided bottleneck for disk I/O by partitioning main memory to load ext3 module.
-We enable Iozone to run in througput mode with two active processes and evaluated the system under three different workloads: read, write and mixed.
-
-\begin{table}[thp!]
-\caption{Iozone Throughput for different filesystem operations(KBytes/Sec)}
+%\begin{table}[thp!]
+%\ORIGcaption{Iozone Throughput for different filesystem operations(KBytes/Sec)}
 % title of Table
-\centering
+%\centering
 % used for centering table
-\begin{tabular}{l r r r}
+%\begin{tabular}{l r r r}
 % centered columns (4 columns)
-\hline\hline
+%\hline\hline
 %inserts double horizontal lines
-Workload & Native & Granary & Modified Drk  \\ [0.5ex]
+%Workload & Native & Granary & Modified Drk  \\ [0.5ex]
 % inserts table
 %heading
-\hline
+%\hline
 % inserts single horizontal line
-Write & 184669.69 & 152223.03 & 55134.09 \\
+%Write & 184669.69 & 152223.03 & 55134.09 \\
 % inserting body of the table
-Read & 1577591.69 & 1648645.02 & 1421037.96 \\
+%Read & 1577591.69 & 1648645.02 & 1421037.96 \\
 %Random Write  & 552380.48 & 412069.48 & 162847.34 \\
@@
 
-Mixed & 1168804.11 & 929383.54 & 915196.76 \\[1ex]
+%Mixed & 1168804.11 & 929383.54 & 915196.76 \\[1ex]
 % [1ex] adds vertical space
-\hline
+%\hline
 %inserts single line
-\end{tabular}
-\label{table:nonlin}
+%\end{tabular}
+%\label{table:nonlin}
 % is used to refer this table in the text
-\end{table}
+%\end{table}
 

### 2013-09-29 thesis/abstract.tex

Finding, understanding, and fixing bugs in software systems is challenging. Dynamic binary translation (DBT) systems provide a powerful facility for building program analysis and debugging tools. However, DBT abstractions are too low-level and provide limited contextual information to instrumentation tools, making it hard to implement such tools.

In this theis, we introduce \emph{behavioral watchpoints}, a new software-based watchpoint framework that simplifies the implementation of DBT-based program analysis and debugging tools. Behavioral watchpoints have two key features: 1) they provide contextual information at the instruction level which are directly available with watchpoints and 2) they enable specializing instruction-level instrumentation with individual data structures. We describe three applications that were easily developed using our watchpoint framework: detecting buffer overflows, detecting read-before-write and memory freeing bugs and detecting memory leaks. We implemented behavioral watchpoints using Granary, a DBT framework for instrumenting operating system kernels. We evaluated the overheads of watchpoints for analyzing and debugging operating system kernel modules and show that these overheads are reasonable.

### 2013-09-29 thesis/applications.tex

In this chapter, we describe several prototype applications that we have developed using the behavioral watchpoint framework. These applications provide debugging facilities for kernel modules and demonstrate the effectiveness of the behavioral watchpoint framework. This chapter uses the term \emph{object} to refer to a range of memory locations that are allocated together as a unit.

%To demonstrate the effectiveness of behavioral watchpoint framework, we developed the prototype applications providing debugging facilities for the kernel modules. In these chapter, I used the term \emph{object} to refer to a range of memory locations that are allocated together as a unit.

%to describe a value or aggregate of values stored in memory and referenced by a memory address.

\section{Buffer Overflows \label{sec:buffer_overflows}}
A buffer overflow occurs when a program---in an attempt to write to some object's memory---actually writes to adjacent memory cells. A low-level programming language like C provides raw memory pointers, permits pointer arithmetic, and does not check bounds when accessing arrays. This can result in very efficient code, but the unfortunate side-effect is that accidentally accessing wrong memory is a very common programming error.

The most obvious example in this class of errors is exceeding the bounds of an array. However the bounds of non-array data objects such as heap blocks, C structs, and stack frames, can also be violated. We consider any memory access that falls outside its intended memory range as bound errors. These errors are not difficult to introduce, and they can cause several bugs, some of which can be extremely subtle and lurk undetected for years. Because of this, tools for preventing and identifying them are extremely useful.

Existing solutions such as Memcheck~\cite{Memcheck} and \emph{fat pointers}~\cite{BccFatPointers}, provide a powerful facility for detecting bounds errors. \emph{fat pointers} hold meta-data for each pointer with the pointer. This metadata describes the memory ranges the pointer can legitimately access and any access via a pointer that is outside its legitimate range is flagged as an error. Another method of detecting buffer overflows relies on the compiler to allocate ``poisoned" regions of memory around each object \cite{AddressSanitizer}. Small overflows (e.g., off-by-one errors) are detected by this approach because they access poisoned memory. Large overflows that ``skip" over poisoned memory and access nearby memory objects are not detected.

Our implementation of a buffer-overflow detector takes an approach similar to \emph{fat pointers} but we store the meta-information for each object in its watchpoint descriptor. Our insight is that unrelated objects will have different base addresses (i.e., the address of one object will not be derived from the base address of another), and thus each object can be distinguished and uniquely identified with a separate watchpoint address, even when objects are adjacent to each other.

%to an unwatched one, or one with a distinct watchpoint, without conflict.
%\emph{addresses} stored in hardware registers, like high-level programming language variable names, are typically not used to simultaneously access two unrelated objects.
%Accessing poisoned memory is an error because it does not exist as a named program entity. However, 
%Our approach is similar in that memory adjacent to a watched object is considered poisoned; however, we do not rely on compiler support, nor do we allocate or mark poisoned memory as such.

\begin{figure}
\begin{lstlisting}[language=C,basicstyle=\footnotesize\ttfamily]
FUNC_WRAPPER(__kmalloc, (size, flags), {
  void *addr = __kmalloc(size, flags);
  ADD_WATCHPOINT(addr, size);
  return addr;
})
\end{lstlisting}
\vspace{-10pt}
\caption[Function wrapper for kernel memory allocator.]{\label{fig:kmalloc_wrapper}Definition of the \texttt{\_\_kmalloc} function wrapper in Granary. The above code expands into a function for wrapping the \texttt{\_\_kmalloc} allocator. Calls to \texttt{\_\_kmalloc} are transparently substituted with calls to the generated wrapper. The wrapper invokes the original \texttt{\_\_kmalloc} function and returns a watched version of the allocated address.}
\end{figure}

We employ three overflow detection policies: heap-based, type-based and stack-based. All three policies depend on the same extension to the meta-information associated with watchpoint descriptors: a \emph{limit address}. Together, the base address (stored in the descriptor) and the limit address delineate an object's boundaries in memory \cite{BccFatPointers}.

\subsection{Heap-based overflow detection\label{sec:heap_overflow}}
The heap-based detection policy detects buffer overflow errors on all heap-allocated objects. We use Granary to wrap the kernel's memory allocators (e.g., \texttt{kmalloc}) and add watchpoints to the addresses returned by those allocators, as shown in \Figref{kmalloc_wrapper}. The lifetime of an added watchpoint is tied to the lifetime of the memory it watches. Each watchpoint's descriptor records bounds information about the allocated memory in the form of the object's base and limit address \cite{BccFatPointers}. A buffer overflow is detected when a dereference of a watched address occurs outside of the bounds recorded by the watchpoint's descriptor (\Figref{detect_overflow}). 
The memory operand-size-specific vtable functions help catch corner cases where memory reads or writes access both an object and its adjacent memory cells, as shown in case $(3)$ of \Figref{detect_overflow}.

%intersect an object's memory and the adjacent memory cells.
%The vtable functions used by these watchpoints detects buffer overflows when a dereference of a watched address 
%memory dereferenced in a read or write falls outside of the bounds
We use Granary to substitute invocations of the kernel's memory allocators (e.g. \texttt{kmalloc}) with wrapped versions of the allocators (\Figref{kmalloc_wrapper}). A wrapped allocator invokes the original allocator but returns a watched version of the allocated address. The lifetime of the watchpoint is associated with the lifetime of the objects and watchpoint descriptor gets collected when objects get freed. 
The watched address returned has its descriptor initialised with the base address as the allocated address and with the limit address as the base address plus the requested allocation size. All buffer overflow detecting watchpoints are initialised with the same vtable pointer, which points to a table of memory operand size-specific functions. The operand size is necessary to detect dereferences of memory that overlap both the object and its adjacent memory, as illustrated in case \emph{(3)} of \Figref{detect_overflow}. The vtable function corresponding to the memory operand size is invoked when a watched address is dereferenced. Each vtable function is programmed to check the watched address against the base and limit addresses stored in watchpoint's descriptor (\Figref{detect_overflow}). If a buffer underflow or overflow is detected then the vtable function notifies the run-time system.

\begin{figure}
\abovedisplayskip=0pt
\belowdisplayskip=0pt
\begin{center}
	\includegraphics[width=4.5in]{overflow.pdf}
\end{center}
\iffalse
	\abovedisplayskip=0pt
	\belowdisplayskip=0pt
	\begin{align*}
		A_{base}  & > {A_{unwatched}} \tag{Underflow} \\
		A_{limit} & \leq {A_{unwatched}} \tag{Overflow} \\
		A_{limit} & < {A_{unwatched}} + S_{op} \tag{Overlap}
	\end{align*}
\fi
\caption[Buffer overflow detection]{\label{fig:detect_overflow}Three common buffer overflow cases are illustrated above: 1) underflow 2) overflow, and 3) overlap. Beside each illustration is the policy that detects the specific bug. When a watched address ($A_{watched}$) is dereferenced, a vtable function that is specific to the memory operand size ($S_{op}$) is invoked. This function detects a buffer overflow if the intended referenced memory address ($A_{unwatched}$) does not fall within the boundaries delineated by the base and the limit addresses ($A_{base}$ and $A_{limit}$, respectively).}
\end{figure}

\subsection{Type-based overflow detection\label{sec:type_overflow}}
Type-based overflow detection is applied when the type of an object is known. It is distinguished from heap-based overflow detection in that it can apply both to heap and non-heap memory, and it can be used to inform the run-time system about more accurate bounds information. In all other respects, the type-based overflow detection policy detects bugs in the same way as the heap-based policy (as might be encoded within the fields of a vector-like data structure). If the type of an object is known then it can reveal semantic relationships which can enable further type- or heap-based overflow detection. For example, the fields within a structure might encode memory bounds information (as would be the case for a vector-like data structure)

%The type-based detection policy is applied when the type of a pointer is known (i.e., not \texttt{void*}). 
The simplest use case of type information is to infer the limit address of a typed pointer. A more complex use case arises when the fields within a structure encode memory bounds information. In both cases, the vtable functions for typed and untyped memory operate in the same way.
The type of an addressable object becomes known to Granary in the context of a wrapped function. The run-time system performs a type-specific initialisation of a watchpoint's descriptor when a watchpoint is added to a pointer/address with a known type. Type-specific descriptor initialisation is a powerful feature because it allows for new watchpoints to be lazily added (thus propagating our detection capabilities) to newly discovered objects referenced \emph{within} an object of a known type. We automatically propagate type-based watchpoints using type specifications generated from parsing \texttt{C} header files. %Some manual post-processing is required if one is to inform the system of semantic relationships between function arguments or structure fields.

%use a type-specific vtable instead of a generic vtable when watching a typed pointer. Granary includes a database of

Granary recursively follows argument and return value pointers based on object type specifications. These specifications tell Granary how to follow pointers and what--if any--pointers should be converted to watchpoints. To reduce programmer effort, Granary automatically generates type specifications and wrapper functions for any set of known types and functions. Some manual post-processing is required if the analysis/debugging application needs to understand semantic relationships between function arguments or structure fields. %\comment{Automating this process was an important goal of Granary because of Granary wraps all (approx. 6,000) exported kernel functions.}.

\subsection{Stack-based overflow detection}
Detecting stack overflows is important because they are commonly used in remote code execution and privilege-escalation attacks against operating systems \cite{SecureProgramExecFlowTracking}.
We detect overflows on stack allocated objects when memory outside of the bounds of the activation/call frame in which the object is allocated is accessed. This policy is distinguished from the heap- and type-based policies in that the watchpoint descriptor is managed separately from the watched memory, and the lifetime of the descriptor extends beyond the lifetime of the watched memory.

To detect stack overflows, we view the memory occupied by the activation frame of an invoked function as a dynamically-sized buffer. Like our other buffer overflow policies, we use a watchpoint to detect accesses to memory outside of this buffer. Unlike our other policies, we only \emph{associate} a descriptor with this buffer, and rely on a different mechanism to \emph{add} the watchpoint to stack addresses.
%, and manage what addresses are watched with this descriptor using a separate mechanism.
We separate adding watchpoints from allocating descriptors for stack overflow detection. In particular, we associate a descriptor with the buffer represented by the activation frame of a called function. This descriptor tracks the bounds of the frame over the lifetime of the function call.

We update the bounds of the frame (in the descriptor) when the frame grows or shrinks. When a function returns, the descriptor's bounds shrink to zero, but the descriptor remains allocated. We detect the two most common sources of stack overflows. First, if we see an instruction that copies the stack or frame pointers, then we assume that the copied address can escape the function. A stack address escaping a function is a potential stack-overflow risk. Adding the frame's watchpoint to this address \emph{taints} the copied address. Future copies or displacements of the watched address implicitly propagate its taintedness because offsets of a watched address reference the same descriptor. A dereference of an escaped pointer---even one happening after the function has returned---is detected as an overflow because the watchpoint descriptor remains live. Second, if we see an indexed dereference of the stack or frame pointers that uses a dynamically bound index, then we assume that the effective memory address accessed is a potential stack-overflow risk. We instrument the dereferencing instruction to add the frame's watchpoint to the effective address before the address is dereferenced.

%more support from Granary's run-time system. At a high level, each activation frame is assigned a dedicated watchpoint when a function is called. The effect of an instruction that grows or shrinks an activation frame's size is reflected by a similar change to the base address\footnote{The run-time call stack on x86 grows into lower memory.\comment{ Modifying the base address instead of the limit address allows us to maintain one set of vtable functions for all three buffer overflow detection policies.}} of the frame's watchpoint descriptor. When a function returns, the descriptor of its activation frame is cleared, but remains allocated\footnote{Reclaiming/reusing allocated but unlikey-to-be-used watchpoints is challenging. One overflow-specific approach is to ensure that the next use of the watchpoint is for a buffer whose bounds do not intersect with the current use's bounds. Another approach is to use some number of context switches as a grace period in which the watchpoint cannot be reused, but after which it can be reused. We leave this as future work.} lest the address of a local variable escape the function\footnote{If the address of a stack-allocated object escapes its function activation frame, then that pointer is said to be \emph{dangling}. Dereferences of dangling pointers can cause stack overflow errors. In the kernel, data stored in unallocated stack memory (e.g., through a dangling pointer) is at risk of being clobbered by interrupt stack frames and by function activation frames.}.

%We assume that a memory instruction operating on a constant displacement of the stack or frame pointer registers is well-behaved. However, if the stack or frame pointer registers participate in a memory instruction and the displacement from the register is dynamically-bound then that instruction is suspect. An instruction that copies the address in the stack or frame pointer registers is also suspect because the copied address might escape the current function or participate in a local stack overflow. All suspect instructions are translated to transparently add the frame's dedicated watchpoint to the dereferenced or copied address.

%This scheme segments the runtime stack into contiguous but non-overlapping regions of watched memory. A straightforward extension excludes saved return addresses and link pointers from the known bounds of activation frames, thus detecting return-oriented buffer overflow attacks.

\begin{figure}[t]
\begin{center}
\includegraphics[width=5.0in]{watchpoint_overflow.pdf}
\end{center}
%\vspace{-15pt}
\caption[Performance impact of buffer overflow detector.]{\label{fig:watchpoint_overflow}Performance overhead (Throughput in MB/sec of common file system operations) of buffer overflow detector when used to detect the buffer overflow in \texttt{ext3} file system module. The buffer overflow detector uses code centric instrumentation for the module code and data centric instrumentation for the kernel code. The overflow detector watches all the memory allocated by the \texttt{ext3} and \texttt{jdb} modules.}
%for the watched address is shown.}
\end{figure}

\begin{table*}
\begin{center}
\begin{tabular}{|l|c c c c|}
  \hline
  Benchmark & Native & Watchpoint Null & Everything watched & Buffer overflow\\
  \hline
  Microbenchmark  & 1x & 2.7x & 3.8x & 20x\\
  \hline
\end{tabular}
\caption[Performance impact of buffer overflow detector on microbenchmark]{\label{table:overflow-benchmark}The overhead of watchpoint instrumentation for buffer overflow detector on microbenchmark shown in \Figref{microbenchmark}. The buffer overflow detector uses instrumentation for bounds checking on the access of watched objects.}
\end{center}
\end{table*}

\subsection{Evaluation}
We evaluated the performance of our buffer-overflow detector using a the synthetic microbenchmark shown in \Figref{microbenchmark} and measured the performance of common filesystem operations on an in-memory disk using the \emph{iozone} filesystem benchmark. We used the same experimental setup as described in Section~\ref{sec:approach_eval}. The overhead of the buffer-overflow detector on synthetic microbenchmark is shown in Table~\ref{table:overflow-benchmark}. The high overhead of detector is due to the instrumentation needed for precise bounds checking. \Figref{watchpoint_overflow} shows the overhead of the buffer-overflow detector on the throughput of common filesystem operations issued by the \emph{iozone} file system benchmark. We saw {\texttildelow}80\% drop in the throughput of file system operations. Much of these overhead comes because of code centric instrumentation and the overflow detector watches all the module allocated objects. Watching selected objects for buffer overflow detection will improve its performance.

%\subsection{Real World Results}
%overhead was \texttildelow 21{\footnotesize$\times$}, which is caused by the instrumentation needed for precise bounds checking.

%and measured the performance of common filesystem operations on an in-memory disk.

\section{Selective Memory Shadowing\label{sec:uninitialised_memory}}

Shadow memory enables a program analysis tool to track the history of every memory location and/or value in memory. Existing tools that use shadow memory such as Memcheck~\cite{Memcheck}, TaintCheck~\cite{NewsomeS05}, TaintTrace~\cite{Cheng:2006:TEF:1157733.1157903}, and Eraser~\cite{Savage:1997:EDD:265924.265927} demonstrate that shadow memory is a powerful and widely used method for developing program debugging applications.

Memcheck~\cite{Memcheck} uses shadow memory to track what allocation/deallocation operations have affected each memory location, and can thus detect accesses to unaddressable memory. It also tracks the undefined memory locations (uninitialised or derived from undefined locations) and can therefore detect dangerous uses of undefined memory. TaintCheck~\cite{NewsomeS05} and TaintTrace~\cite{Cheng:2006:TEF:1157733.1157903} are security tools. They track which values are from untrusted (tainted) sources and which values were subsequently derived from them, and can thus detect dangerous uses of tainted values.
%It also remembers which values are undefined (uninitialised or derived from undefined values) and can therefore detect dangerous uses of undefined values. TaintCheck is a security tool. It remembers which values are from untrusted (tainted) sources and which values were subsequently derived from them, and can thus detect dangerous uses of tainted values.
Eraser~\cite{Savage:1997:EDD:265924.265927} uses shadow memory to track which locks are held when a memory location is accessed, and can thus detect when a memory location is accessed without a consistent lock-set, which may imply a data race. In the above applications, shadow memory provides the ability to detect critical errors, which otherwise could go undetected.

The implementation of shadow memory in these tools is inherently expensive because all memory needs to be shadowed~\cite{Nethercote:2007:SBM:1254810.1254820}. The tools need to maintain extra state for each byte of memory and instrument all loads and stores to keep this state up-to-date. These requirements increase the total amount of code that is run and the program's memory footprint, thus significantly affecting performance. Instead, behavioral watchpoints enable implementing a novel shadowing scheme that allows tracking objects selectively, which decreases the amount of code that needs to be
run and the program's memory footprint.

 %The solution of the problem associated with full memory shadowing is selective shadowing. Selective shadowing allows you to track only the objects of your interested thus decreasing the memory footprints and amount of code required to be run.

%This makes it important to have an approach which allows you the selective memory shadowing. Selective 
%The previous work on shadow memory is mostly focused on full memory shadowing.

%is implemented in software. However implementing shadow memory correctly is hard because of following three reasons:

%First, the implementation of shadow memory is inherently expensive. A tool need to maintain large amounts of extra state and one shadow byte per byte of live original memory is typical. Most or all loads and stores must be instrumented to keep the shadow memory state up-to-date.% as must operations that affect large regions of memory, such as allocations and deallocations (on the heap or stack ), reads/writes of large areas by system calls, and the loading of the program image into memory at start-up.
%These requirements unavoidably increase the total amount of code that is run, increase a program’s memory footprint, and degrade the locality of its memory accesses. %Shadow memory tools thus typically slow down programs by a factor of 10–100, and shadow memory operations cause much more of this slow-down than other well-studied aspects of tools built with DBI frameworks such as hot trace formation [1] and code cache management [7]

%Second, The robustness of a shadow memory implementation is also important, arguably even more so than its speed. Real world tools must cope with large, uncooperative programs, and if they are to be portable, cannot rely on particular operating system characteristics such as memory layouts. Shadow memory is hard to implement robustly. The shadow memory must be squeezed into the address space alongside the
%original memory in a way that does not conflict with it and does not change the program’s behaviour. This requires considerable flexibility in the shadow memory structure and layout. It also unavoidably reduces the amount of address space a program can use itself,
%which is an important issue on embedded platforms and even 32-bit machines. Obviously, this becomes less of an issue if shadow memory
%can be made smaller, so compact shadow memory is desirable.

%In this section, we show how to use behavioral watchpoints to shadow memory. %Previous work has focused on full memory shadowing \cite{Memcheck}, while
%This section describes about the use of behavioral watchpoints in enabling \emph{selective} shadowing for watched objects. It also describes how the initialisation state of each byte of watched memory is tracked using shadow memory, and how to detect bugs related to the usage of heap-allocated memory. It also discusses the use of selective shadowing in tracking the memory access pattern of the kernel modules.

%Granary provides support to wrap kernel memory allocators and deallocators (e.g., \texttt{kmalloc}, \texttt{kfree}). Wrapped allocators add watchpoints to allocated memory, and wrapped deallocators remove watchpoints before invoking the kernel's deallocators. 
We implement selective shadow memory by maintaining a watchpoint descriptor-specific, variable-sized bitset. %\footnote{As an optimisation, shadow memory for small objects is embedded in the watchpoint descriptor.}. 
Each byte of allocated memory corresponds to one bit of descriptor-specific shadow memory. The bits in shadow memory are initialised to zero. Individual bits are flipped to one by write-specific vtable functions or read-specific vtable functions when a write/read operation happens to the bytes of memory shadowed by those bits.

%When a wrapped allocator is invoked, a watchpoint is added to the addresses returned by those allocators

% in shadow memory are initialised to zero.
%Watchpoints are added to all heap-allocated objects 
%the meta-information within each watchpoint descriptor.
%By tracking the initialisation state, we detect

We use this shadow memory scheme to detect memory error bugs (e.g., uses of uninitialised memory, memory freeing bugs for heap-allocated objects) and to characterize the types of objects accessed by different kernel modules. As described in Section~\ref{sec:heap_overflow}, Granary interposes on the kernel's memory allocators and adds watchpoints to the addresses returned by those allocators. When memory is allocated, the watchpoint descriptor is initialised with the number of allocated bytes (available as an argument to the allocator) and with its own shadow memory.

%detect uses of uninitialised memory.  Detecting such uses is important because a program using uninitialised memory can exhibit unintended, non-deterministic behaviour.

Uninitialised memory is defined as either allocated memory to which no value has been written, or deallocated memory. We selectively track the initialisation state of each byte of watched memory using \emph{shadow memory} \cite{Memcheck}. Our implementation represents shadow memory as a bitset, where the state of each bit tracks the initialisation state of a byte of watched memory. Watchpoint descriptors are augmented to include the size (in bytes) of a watched object and a pointer to the memory shadowing the object\footnote{As an optimisation, shadow memory for small objects can be embedded in the watchpoint descriptor.}.

\subsection{Read-before-write bugs}
Read-specific vtable functions detect read-before-write bugs by checking if the shadow bit corresponding to one of the read bytes is zero. However, this method of detection can report false positives: it is common for larger-than-needed reads to be performed and for compiler-added structure padding to be read (but never written). A relaxed read-before-write memory checking policy requires that at least one shadow bit is set for every read operation.

\subsection{Memory freeing bugs}
We detect freeing of non-heap memory when the argument to a kernel memory deallocator is not a watched address. We detect an invalid free, such as freeing an offset of an allocated object, when the unwatched address of a watched object being deallocated is different from the watchpoint descriptor's base address. We detect use-after-free bugs by marking the descriptor of a watched address being deallocated as dead. The lifetime of a dead descriptor extends beyond that of the object so that a later use of any watched address with that descriptor will report a use-after-free bug. We detect double-free bugs when the descriptor of a watched address being deallocated is already marked as dead.

%We use Granary to wrap the kernel's deallocation functions (e.g., \texttt{kfree}). Because all allocators are wrapped and add watchpoints, freeing of non-heap memory is easily detected when

%We detect freeing of non-heap memory

%If the argument to a wrapped deallocator function is a watched address then the shadow memory for the .

%Finally, the original deallocator is invoked with the unwatched address as an argument. Reads and writes to a watched address of a deallocated object fail to pass bounds checks because the object size is zero.

Two positive side-effects of our approach is that it detects double-free bugs and freeing of non-heap memory bugs. A double-free is detected when a watched address for an already deallocated object is passed as an argument to a deallocator. An attempt to free non-heap memory is detected when an unwatched address is passed as an argument to a deallocator.

\subsection{Fine-grained access pattern}
Kernel modules execute in the kernel address space with unbounded privileges. These modules are implicitly trusted and interact with each other and the kernel through well defined interfaces and by sharing data in an uncontrolled manner. Unfortunately, the assumed trust leaves commodity OSes vulnerable to malicious and misbehaving kernel extensions\cite{BGI,LXFI, Xu:2004:DEC:1038254.1038305}. We detect the access patterns of the kernel modules using selective shadowing. Behavioral watchpoints tracks the objects of interest and any read/write access to these objects by both the kernel or the module code is captured by the shadow memory. 
The shadow memory information is logged at every control transfer from the kernel to the module. This information is useful for characterizing the access patterns of kernel modules, when may then be used to classify the module to detect module vulnerabilities.
%in a prototype and also detect the vulnerability in the kernel modules.

\section{Memory Leak Detection\label{sec:memory_leaks}}
%Memory leak is a major resource issue which could lead to the malfunctions and negative performance of the system. 
A memory leak occurs when a program does not free dynamically allocated memory or when memory can no longer be referenced. Memory leaks are especially a problem in long running programs, and in operating systems, where memory is lost until the next boot. Existing tools such as Purify~\cite{Rs_purify:fast}, Valgrind~\cite{Seward:2005:UVD:1247360.1247362} and Dr Memory~\cite{Bruening:2011:PMC:2190025.2190067} provide an effective solution for detecting memory leaks in user-space programs but they are not available for the kernel. Other systems such as \emph{kmemleak}~\cite{kmemleak} and SystemTap~\cite{EIgler05architectureof} provide support for debugging memory leaks in the kernel but they operate on the kernel as a whole and not on specific modules of interest.

%one module particularly selected by the user(s).

Implementing a leak detector only for kernel modules is challenging for three reasons.

First, a leak detector for kernel modules should follow all objects allocated in the module context. Many of these objects should be deallocated by the module. For example, a file system module may make a call to \texttt{create\_workqueue} and allocate a \texttt{work\_struct} that can be used to schedule the delayed work. It is the responsibility of the module to deallocate the object before exit. Keeping track of such objects for a leak detector operating on the entire kernel, such as \emph{kmemleak}, is easy since they track all the memory allocation and deallocation but it is challenging for a leak detector operating only on the kernel module.

Second, a leak detector needs to perform reachability analysis on all objects allocated in the module context to detect the lost references. Existing leak detectors use different garbage collector algorithm such as reference counting and tracing-based algorithm to detect the live objects. Performing such analysis in only module context is challenging because module regularly lose internal references to the objects that they allocate. For example, a network module can allocate objects and pass with the \texttt{net\_device} structure to the kernel, without retaining references to those objects. The network module can later indirectly access these objects using the kernel's \texttt{netdev\_priv} interface. A whole kernel leak detector, such as \emph{kmemleak}, operating on kernel heaps \& static data can detect such lost references by scanning all allocated objects but a module based leak detector, having limited knowledge of objects allocated in the module context, can not detect the references of such objects and will report them as a leak.

Third, the existing leak detector in the kernel needs to trace all the memory allocator and deallocator functions in order to track the allocated objects. It also needs to trace the \texttt{container\_of} macros to get the addresses generated from a pre-allocated objects. \emph{kmemleak}, being tightly integrated with the kernel source code, performs this by interposing at the memory allocators/ deallocator (e.g, \texttt{kmalloc}, \texttt{vmalloc}, \texttt{kmem\_cache\_alloc} and other friend functions) and tracing the \texttt{container\_of} operations on pre-allocated objects. Performing such operations by a leak detector, having no knowledge of the kernel activities and operating only on the module binaries is not possible.

%suffers from the false negative because the allocated object doesn't maintain the state information about the memory block. This makes it difficult to identify the transient presence of freed objects in the static or the heap memory. The existing solution of leak detector based on garbage collection algorithm also can not identify the memory leaks due to the increase in stale objects. These stale objects gets allocated by the module but they never gets used and the module also doesn't lose their references. Identifying such objects without tracking the access of each memory block is not possible.

%Tracking the liveness of these objects requires a view on kernel execution which is not normally provided by module-only instrumentation.

 %are challenging because not all module-owned objects are directly allocated by modules. For example, a network driver module may make a call to an skb allocator (to allocate a \texttt{struct sk\_buff}) in an interrupt context when the module code is not running but it is the responsibility of the network module to deallocate them.

%should follow the module-owned objects and reports leak if and when module loses the reference of any of them. An object is defined as module-owned if it is allocated in the module context and it is the responsibility of the module to deallocate the object. However, keeping track of all such objects are challenging because not all module-owned objects are directly allocated by modules. For example, a network driver module may make a call to an skb allocator (to allocate a \texttt{struct sk\_buff}) in an interrupt context when the module code is not running but it is the responsibility of the network module to deallocate them.

%driver module may make a call to an skb allocator (to allocate a struct sk_buff) in an interrupt context when the module code is not running

%We built a memory leak detector for the kernel modules. The leak detector for the kernel modules follow the module owned objects and reports if module loses the references of them. A module owns the objects allocated by them and it is the module's responsibility to deallocate them. However, keeping track of all such objects is challenging because not all module-owned objects are directly allocated by modules. For example, \texttt{sk\_buff} objects used by network modules are allocated by kernel interrupt handlers, but must be deallocated by the network modules.

%Second, a leak detector performs reachability analysis on all module-owned objects to detect the lost references of the objects. Leak detector uses different garbage collector algorithm such as reference counting and tracing-based algorithm to detect the live object. Performing such analysis in only module context is challenging because module regularly lose internal references to the objects that they own. For example, a network module can allocate objects and pass with the \texttt{net\_device} structure to the kernel, without retaining references to those objects. The network module can later indirectly access these objects using the kernel's \texttt{netdev\_priv} interface. Tracking the liveness of these objects requires a view on kernel execution which is not normally provided by module-only instrumentation.

%Third, the existing leak detector suffers from the false negative because the allocated object doesn't maintain the state information about the memory block. This makes it difficult to identify the transient presence of freed objects in the static or the heap memory. The existing solution of leak detector based on garbage collection algorithm also can not identify the memory leaks due to the increase in stale objects. These stale objects gets allocated by the module but they never gets used and the module also doesn't lose their references. Identifying such objects without tracking the access of each memory block is not possible.

 %thus the transient presence of deallocated objects in the  
%In the implementation of leak detector there is a need to track the state of the allocated objects to reduce the false negative. The existing memory allocator/deallocator doesn't maintain a meta-information with each objects. Thus making it difficult to identify the freed objects thus reporting false negative.

We implemented the leak detector for the kernel modules using the watchpoint framework. The challenges involved in implementing a leak detector only for the kernel modules make behavioral watchpoints a suitable solution. Behavioral watchpoints solve the challenges involved in three ways:
\begin{enumerate}[i)]
	\item The watchpoint framework ensures that all objects allocated in the module context are watched. These watched addresses are easily disambiguated from normal memory addresses and integers that look like watched addresses. This is done by taking over all kernel memory allocators and tracking the execution context of the module running under its control. This helps the framework identify the context of every memory allocation and adds watchpoints on all these objects. However, in the process the watchpoint framework may add watchpoints on the objects not related to the module. Adding extra watchpoints on objects does not affect the correctness of the leak detector. We detect such objects during the collector scan and suppress the false positive that may get reported due to them.

	\item Behavioral watchpoint solves the problem of losing internal reference of objects allocated in the module context by providing leak detector complete visibility of the watched objects. Any read/write accesses performed on these watched objects get detected by the watchpoint instrumentation or the hardware traps which attaches the instrumentation framework with running code. The read/ write operations on watched addresses mark these objects live and the collector scan considers these objects as reachable. The ability to track the accessed objects is particularly helpful in detecting stale objects that are potentially leaked. The leak detector based on garbage collector algorithm such as \emph{kmemleak} can not detect such objects.

	\item Behavioral watchpoints help associate the allocated object with the descriptors. The descriptors hold meta-information for each objects and help identifying the state of a memory block without much of the effort. This helps leak detector limit the scope of scanning to objects that are allocated and live. The behavioral watchpoints are also viral and every addresses derived from a watched addresses (e.g., through copying or offsetting) is also watched. The watchpoint framework does not need to trace the pointer arithmetic used to derive addresses from the allocated objects. \emph{kmemleak} reports false positive if the program uses other than \texttt{container\_of} operations to generate derived addresses. %to track the derived addresses from the% It also helps in tracking the read/write access of the watched objects by kernel or the module code.
\end{enumerate}

%Behavioral watchpoints are viral and every addresses derived from a watched addresses (e.g., through copying or offsetting) is also watched.

%Our approach ensures that all module-owned objects are watched when allocated. Leak detector ensures this by taking over all the kernel allocators and tracking every entry and exit from the kernel to the module and vice-versa. This helps us in identifying if the allocation is happening in the module context or not. We add watchpoint with all the objects allocated in the module context. 
%\texttt{call} from kernel to the module code and \texttt{ret} from the module to the kernel.   
%We ensure this by hot-patching all the kernel allocators and tracking all the calls from the module to the kernel which reaches to the allocators and add watchpoint with those objects. 
%Behavioral watchpoints solve this problem of losing the internal reference of the module-owned object by giving the leak detector visibility into kernel accesses to watched objects, which trigger hardware traps that attach instrumentation to the kernel code. This instrumentation marks kernel-accessed watched objects as live. The implementation of leak detector benefits from using watchpoints in three ways: 
%\begin{enumerate}[i)]
	%\item watched addresses are easily disambiguated from normal memory addresses and integers that look like watched addresses;
%        \item the scope of scanning can be limited using the meta-information provided by descriptors
	%\item descriptors provide meta-information that allows our system to limit the scope of scanning; and
	%\item scanning can stop if all watched objects are reached.
%\end{enumerate}

%A conventional approach of implementing leak detector is to perform the reachability analysis on the module-owned objects. The existing solution, \emph{kmemleak} uses mark/sweep algorithm \cite{Boehm:1991:MPG:113445.113459} on kernel memory to detect if/when module-owned objects become unreachable starting from a set of ``root'' objects. This is challenging because modules regularly lose internal references to the objects that they own. For example, a network module can allocate objects and pass with the \texttt{net\_device} structure to the kernel, without retaining references to those objects. The network module can later indirectly access these objects using the kernel's \texttt{netdev\_priv} interface. Tracking the liveness of these objects requires a view on kernel execution not normally provided by module-only instrumentation.

%We used behavioral watchpoints to implement memory leak detector. Our approach ensures that all module-owned objects are watched when allocated. Behavioral watchpoints solve this problem of losing the internal reference of the module-owned object by giving the leak detector visibility into kernel accesses to watched objects, which trigger hardware traps that attach instrumentation to the kernel code. This instrumentation marks kernel-accessed watched objects as live. The implementation of leak detector benefits from using watchpoints in three ways: 
%\begin{enumerate}[i)]
%	\item watched addresses are easily disambiguated from normal memory addresses and integers that look like watched addresses;
%        \item the scope of scanning can be limited using the meta-information provided by descriptors
%	\item descriptors provide meta-information that allows our system to limit the scope of scanning; and
%	\item scanning can stop if all watched objects are reached.
%\end{enumerate}

%More operationally, a dynamically-live heap object is one
%that can be reached by following pointers that will be dereferenced in the future of
%the computation (dynamically-live pointers). In order to retain only dynamically-live
%objects, the ideal garbage collector must be able to exactly identify what memory locations contain dynamically-live pointers.

\subsection{Object Liveness Analysis}\label{sec:object_liveness}
The effectiveness of a leak detector in identifying the object leaks depend on the ``accuracy'' of the liveness analysis. An ideal leak detector identifies all heap-allocated objects that are not dynamically live and reports them as leak. A dynamically-live heap object is one that can be reached by following pointers that may be dereferenced in the future.

Existing leak detector solutions such as \emph{kmemleak}~\cite{kmemleak}, Purify \cite{Rs_purify:fast} and Dr Memory \cite{Bruening:2011:PMC:2190025.2190067} use a garbage collection algorithm to perform object liveness analysis. They use the mark/sweep collector algorithm \cite{Boehm:1991:MPG:113445.113459} to detect reachability of the allocated objects starting from a set of ``root-set'' objects. Their root-set includes system registers and non-heap addressable memory.

They perform reachability analysis periodically at intermediate position and it involves stopping the client when collector\footnote{A low priority daemon thread which runs intermittently in the background performing reachability analysis on the allocated objects} is in process. However, the solution is not very efficient when client operates on a large number of heap-allocated objects, such as is typical in operating systems. It stops the client for longer time.

One solution for reducing the pause time is to use generational or parallel collection. Generational collection avoids stop-the-world by running the collector thread occasionally but it doesn't remove the problem completely. However the parallel collectors take an orthogonal approach and reduce the time of pause by running the collector in parallel to the client.

In our implementation of leak detector we used the parallel collector to perform object liveness analysis. Keeping in view of the challenges involved in performing object liveness analysis, we used three different policies to perform the leak scan.

%took middle approach and used mostly-parallel collector~\cite{Boehm:1991:MPG:113445.113459, Barabash:2005:PIM:1108970.1108972} to perform the object liveness analysis. It is called as mostly parallel collector because we update the ``root-set'' in stop the world collector and perform the reachability analysis from these ``root-sets'' by running the collector parallel to mutator (client). 
%The aim of our leak detector is to detect the memory leaks in the kernel modules. These kernel modules execute in a complex and dynamic environment, sharing the data over irregular interface. Many of these shared objects are allocated directly or indirectly by the modules. Behavioral watchpoints provide an efficient solution by providing complete visibility to the module-owned object and making these watched objects easily identifiable, thus limiting the scan space but they does not provide any solution for performing the reachability analysis on all these watched objects. Performing reachability analysis on these objects is challenging because of two reasons. First, these module allocated objects get shared with the kernel over complex and irregular interface making them difficult to track and many of them are reachable only through the kernel internal data structure. The reachability analysis of these objects without accounting the kernel allocated memory is non-trivial and it will lead to false positives. Second, our leak detector framework gets loaded as the kernel modules and it comes into the picture long time after the system ``boot-up''. Without having account of these memory block we can not claim to perform the complete leak scan. Keeping in view of these challenges we used three different policy for performing the leak scan.

%A conventional approach of implementing leak detector is to perform the liveness analysis on the allocated objects. Existing solution such as \emph{kmemleak}~\cite{kmemleak}, Purify \cite{Rs_purify:fast} and Dr Memory \cite{Bruening:2011:PMC:2190025.2190067} uses garbage collection algorithm to perform object liveness analysis. \emph{Kmemleak} uses mark/sweep algorithm \cite{Boehm:1991:MPG:113445.113459} on kernel memory to detect the reachability of the allocated objects starting from a set of ``root'' objects. The effectiveness of leak detector in identifying the lost objects depends on the accuracy of their object liveness analysis. 
%Existing solution such as Purify \cite{Rs_purify:fast} and Dr Memory \cite{Bruening:2011:PMC:2190025.2190067} leak detector uses garbage collection algorithm to perform liveness analysis. 
%A straightforward implementation of garbage collection algorithm is to implement tracing based collector. It involves stopping the client when collector is in the process performing liveness-analysis. However, this solution is not very efficient when client operates on very large heap-memory as it stops the clients for longer time. The solution to reduce the pause time in tracing algorithm based leak detector is generational or parallel collection. Generational collection avoids stop-the-world time by running the collector thread occasionally, thus doesn't remove the problem completely. However the parallel collectors takes orthogonal approach and reduces the time of pause by running collector parallel to client. We took similar approach of mostly-parallel collector to perform the object liveness analysis.

%In Linux kernel, the module and kernel interacts in a complex ways and there is frequent transfer of shared objects between the kernel and the modules. Many of these shared objects are allocated by the module. Keeping track of the references of these objects with only module view of the objects is non-trivial. This makes it important for the leak detector to track both kernel and the module objects. It was important for us to study the transfer or the ownership of the objects. In our implementation of leak detector we uses three different scan policy to perform the object liveness analysis.

 %following three different liveness analysis policy for the leak detector:

\paragraph{Scanning of accessed objects:} This is an aggressive scanning policy for detecting memory leaks. The approach is similar to bookmarking collectors~\cite{Hertz:2005:GCW:1065010.1065028} and Melt~\cite{Bond:2008:TML:1449764.1449774} which track the accesses of allocated memory blocks to detect stale objects at page or object granularity. The approach considers all objects that are accessed between two consecutive scans as live objects and performs reachability analysis considering them as ``root-sets''. Any object which is not reachable from these objects is considered as stale or a potentially leaked object.

The leak detector tracks all accesses of the watched objects and updates the meta-information associated with their descriptors. The scanner thread (collector) periodically scans the descriptors stored in a global descriptor table and marks the objects that were accessed in the last epoch as live. The scanner thread later performs reachability analysis considering these objects as ``root-sets''.  The leak detector also tracks the sources of memory allocation in the descriptors and a constant increase in the stale objects from one source is reported a possible leak.

This approach reduces the overhead of leak scan which is performed by the collector periodically and it is also helpful in finding stale objects (dead but reachable) which the existing garbage collector algorithm based leak detector can not detect. However, the approach performs limited reachability analysis, thus resulting in increase in the false positive especially for objects which does not get accessed for longer interval of time. It also can't find the objects with live leak i.e the allocated objects are getting accessed periodically but these accesses are nonetheless useless. Detecting such leaks with any leak detector is challenging.

 %to find the objects which is accessed in last epoch and mark them dead accounting them as live during the scan interval. The approach also tracks the source of each allocation and the constant increase in the stale objects from one source increase the possibility of memory leaks.

%In this scan policy, every access of the watchpoint objects also updates the meta-information associated with them. The scanner thread periodically scans the descriptor to find the objects which is accessed in last epoch and mark them dead accounting them as live during the scan interval. The approach also tracks the source of each allocation and the constant increase in the stale objects from one source increase the possibility of memory leaks.

%The approach decreases the cost of performing object liveness analysis by the collector and it is also helpful in finding the leaks of stale objects (dead but reachable) which can not be identified using garbage collector algorithm. However, the approach is limited in performing reachability analysis, thus it results in the false positive especially for the objects which doesn't get accessed for long interval of time. The approach also can't find the objects with live leak where the allocated objects gets accessed periodically but these accesses are nonetheless useless. Detecting such leaks in the module is difficult.

%and mark them dead after accounting them as live during the scan interval. The leak detector makes decision after 10 scans.

%The approach deceases the cost of object liveness analysis, but it results in the false positive for those objects which doesn't gets accessed for long period of time.

\paragraph{Scanning of module reachable objects:} 
We define module reachable objects as objects which a module can access directly or has accessed in the past. It includes static data for both the kernel and the modules, system registers, threads stack executing module code and all heap-allocated memory viewed by the module directly or indirectly. Module viewed objects are the objects allocated by the kernel but passed to the module during its course of execution. We keep track of such objects by interposing at the kernel/module interface and collecting all the kernel objects passed to the modules. The type information about these objects helps us in associating a type-specific scanner with each object, which performs deep scanning of these objects during a leak scan.

In this leak scan policy, the collector considers all module reachable objects as the ``root-set'' and recursively scans them to perform liveness analysis. This scan policy uses the mostly-parallel collector~\cite{Boehm:1991:MPG:113445.113459, Barabash:2005:PIM:1108970.1108972} instead of the parallel collector for updating the ``root-set'' before performing the leak scan. This is important to have a consistent snapshot of system registers and thread stack running the module code.  The parallel scanning of kernel objects requires maintaining consistency and avoid the cyclic dependency of dependencies across data structures. We detect cyclic dependencies by hashing each memory block during a scan and storing them in a data structure. The scan of a memory block is stopped if its hash is found. We also use the scan depth as a fall back mechanism to avoid walking a long list of objects in a data structure. In our implementation we used a scan depth of ``6'' which we find is appropriate for detecting all watched objects in any kernel data structure. We are also using relaxed consistency policy during the scan. This is because our leak detector makes a decision about a memory block leak after multiple scans.

This approach reduces false positives compared to the first approach because it includes module-viewed objects in the root set. However, it has a higher cost because it requires maintaining bookkeeping information for the module-viewed objects and tracking their deallocation. The approach does not remove false positives completely as it is not able to track references to objects that are allocated by the module but passed to the kernel, after which the kernel uses internal data structures to maintain references to the objects.

%kernel and kernel uses internal data structure to maintain its references.

%of object scan by scanning them twice before making a decision. We consider our consistency approach an appropriate because the leak detector makes decision about a memory block after multiple scans.

%where the ``root-sets'' for scan gets updated in stop-the-world collector. This is required to have a consistent snapshot of system registers and thread's stack running the module code. The parallel scan of kernel objects requires to maintain consistency and avoid the cyclic dependency of data structure. We avoid the cyclic dependency by hashing the memory block after every scan and storing them in a data structure. The scan of a memory block gets suspend if its hash is found in the data structure. We also use scan depth as the fall back to avoid walking of the list objects in a data structure. In our implementation we used the scan depth as ``6'' which we find appropriate for detecting the watched objects in the data structure. We relaxed the consistency of object scan by scanning them twice before making a decision. We consider our consistency approach an appropriate because the leak detector makes decision about a memory block after multiple scans.

%The module reachable objects includes the static data for both the kernel and modules, system registers, stack of threads executing module code and all heap-allocated memory viewed directly or indirectly by the module. We defined the module viewed objects as the objects allocated by the kernel but passed to the module during its course of execution. We keep track of all such objects by interposing at the kernel/module interface and each object gets associated with a corresponding type scanner function based on the object types. These type scanner function is used by the collector to recursively scan the object.

%This scanning policy uses mostly-parallel collector where the ``root-sets'' for scan gets updated in stop-the-world collector. This is required to have a consistent snapshot of system registers and thread's stack running the module code. The parallel scan of kernel objects requires to maintain consistency and avoid the cyclic dependency of data structure. We avoid the cyclic dependency by hashing the memory block after every scan and storing them in a data structure. The scan of a memory block gets suspend if its hash is found in the data structure. We also use scan depth as the fall back to avoid walking of the list objects in a data structure. In our implementation we used the scan depth as ``6'' which we find appropriate for detecting the watched objects in the data structure. We relaxed the consistency of object scan by scanning them twice before making a decision. We consider our consistency approach an appropriate because the leak detector makes decision about a memory block after multiple scans.

%The approach decreases the false positive by including the module viewed kernel objects in the ``root-sets''. However, the approach requires to maintain book-keeping of kernel objects passed to the module which has additional cost. The approach also doesn't remove the false positive completely as it is not able to track the references of objects which are allocated by the module but passed to the kernel and kernel uses local data structure to maintain its references.

%the state data of both the kernel and modules and all the kernel and module heap-allocated object which is viewed directly or indirectly by the module.

%In the initial implementation of leak detector, we decided to track all the in-memory kernel data structure. we track all the module viewed objects. We defined the module viewed objects as the objects passed to the module during its course of execution. We keep track of all such objects and scan them periodically looking for the watchpoints. To maintain the consistency during the data structure we scan them twice and to avoid the circular dependency on the data structure we hash the data structure.

%This policy decreases the false positive but requires to have a lot of book-keeping for the kernel objects. The approach is also not able to track the references of objects which are allocated by the module but passed to the kernel and kernel uses local data structure to maintain the reference of the objects for long time.

%we studied the ownership transfer of the module owned objects. We track all the object viewed by the module and scan them periodically looking for the watchpoints.

\paragraph{Scanning of kernel heaps and static data:} This is the most conservative policy and used as the fall back approach for performing a leak scan. This policy scans the static data for both kernel and the modules and kernel heaps to find any references to watched objects. A watched object that is not found during the scan is definitely considered a leak.

For scanning kernel pages, we used the kernel page tables and the kernel data structure \texttt{init\_level4\_pgt} to access kernel pages. The scanner thread (collector) divides the kernel pages based on the memory regions and types and scans them based on priority. The scanner thread stops scanning the kernel pages once it finds at-least one reference to watched objects.

This approach avoids any false positives but scanning the kernel heap memory is expensive. Also, the scanning of kernel pages can cause false negatives because leaked objects may have internal pointers containing watched addresses. We reduced the problem by removing the memory block of watched objects that are not live from the scan list.

 %and pointers associated with kernel data structure which is not zeroed out.

%We reduce the false negative by ignoring the scan of pages associated with watched object and they get scanned only when a reference of that object is found in other pages. This is particularity required as we found in many cases the internal pointers of watched object was initialized with the watchpoint address and scanning of these pages or memory block will provide the false negative even if the object is considered as lost.

%for implementing leak detector. In this policy we scan the entire kernel heap and module \& kernel static data for finding the references of the watchpoints. Any watchpoints not found during the scan is considered as leak. For scanning the kernel heaps and static data structure we walk over the kernel page table using the page table data structure \texttt{init\_level4\_page}.

%The scanning the kernel heap and static data reduces the chances of false positive but the cost of scanning them is costly. However, In our approach we device the kernel address space in different region and we scan them based on the priority. The scanner thread also stops scanning the memory once it finds at least one reference of all the watched objects.

%The disadvantage of scanning the kernel heaps is that

%In this can policy we scan the entire heap memory and static data by walking over the page table looking for the watched objects.

%Other solutions such as (...) considers objects staleness as an indicator of problematic behavior.

%The existing leak detector tracks the behavior of the arbitrary objects

%track behaviors
%of arbitrary objects and report problems when tracked objects
%become suspicious. One major category of work [8, 9, 10, 18, 33]
%considers objects’ staleness (i.e., time elapsed since the program
%last used these objects) as an indicator of problematic behavior,
%while another category [20, 23] treats objects as suspicious if instances
%of their types exhibit sustained growth

%Our approach of leak detector is based on the liveness analysis of heap objects, which are directly or indirectly allocated by the modules. Similar to existing solution like Purify \cite{Rs_purify:fast} and Dr Memory \cite{Bruening:2011:PMC:2190025.2190067}, leak detector uses garbage collection algorithm to detect memory leaks. A straightforward implementation of garbage collection algorithm is to implement tracing based collector. It involves stopping the client when collector is in the process performing liveness-analysis. However, this solution is not very efficient when client operates on very large heap-memory as it stops the clients for longer time. The solution to reduce the pause time in tracing algorithm based leak detector is generational or parallel collection. Generational collection avoids stop-the-world time by running the collector thread occasionally, thus doesn't remove the problem completely. However the parallel collectors takes orthogonal approach and reduces the time of pause by running collector parallel to client. We took similar approach of mostly-parallel collector to perform the object liveness analysis.

%In our approach of liveness analysis, we use three different analysis policy:

%The leak detector scans kernel memory using the conventional mark/sweep algorithm \cite{Boehm:1991:MPG:113445.113459} to detect if/when module-owned objects become unreachable starting from a set of ``root'' objects. This is challenging because modules regularly lose internal references to the objects that they own. For example, a network module can allocate objects and pass with the \texttt{net\_device} structure to the kernel, without retaining references to those objects. The network module can later indirectly access these objects using the kernel's \texttt{netdev\_priv} interface. Tracking the liveness of these objects requires a view on kernel execution not normally provided by module-only instrumentation. Behavioral watchpoints solve this problem by giving the leak detector visibility into kernel accesses to watched objects, which trigger hardware traps that attach instrumentation to the kernel code. This instrumentation marks kernel-accessed watched objects as live.

%The leak detector benefits from using watchpoints in three ways: %\begin{inparaenum}[i)]
%	\item watched addresses are easily disambiguated from normal memory addresses and integers that look like watched addresses;
%        \item the scope of scanning can be limited using the meta-information provided by descriptors
%	\item descriptors provide meta-information that allows our system to limit the scope of scanning; and
%	\item scanning can stop if all watched objects are reached.
%\end{inparaenum}

%Two key benefits of our watchpoints solution is that watched addresses are easily disambiguated from normal memory addresses, and that we can stop scanning memory if all watched objects are live and reachable.

%Detecting leaks of such objects with a limited view of module is challenging. Behavioral watchpoints provide a way to ensure the visibility of such objects by triggering a hardware trap on its access and mark them as live.
%on memory access of these objects by enabling demand-based translation and marking them as live.

% during device registration.

%The object becomes invisible to the module and gets accessed through the kernel interface \texttt{netdev\_priv}. This makes it difficult to detect leaks based on module's limited view. Bhavioural watchpoints provides a way to ensure complete visibility on memory access of these objects by enabling demand-based translation and marking them as live.

%All module owned objects are not directly allocated by the module and kernel can allocate it on behalf of the module. 
%All module-owned objects are not directly allocated by the module. For example, a network driver module may make a call to an skb allocator (to allocate a \texttt{sk\_buff} structure) in an interrupt context. Once the buffer is send over the network, it is the responsibility of the network module to deallocate the buffer. In Granary, we use wrappers to add watchpoints at relevant allocators such as \texttt{\_\_alloc\_skb}. The leak detector keeps track of these watched objects and reports possible leaks.

%Module holds partial view of the kernel and to detect leaked object there is a need to perform reachability analysis on both kernel and module objects. Behavioral watchpoints helps in identifying the objects owned by the module, thus limiting the type of objects which needs to be tracked. Watched object holds descriptor information in high-order bits, making it easy to track the memory state information. This is helpful in identifying live objects and reduces the chance of false positive. The leak detector uses the conventional mark and sweep algorithm and mostly-parallel collector~\cite{Boehm:1991:MPG:113445.113459} to perform the rechability analysis.

%Watched addresses are also more easily disambiguated

%Behavioral watchpoints are important for implementing the leak detector because they help in identifying the objects owned by the module, thus limiting the type of objects that need to be tracked. Each watched object holds the memory state information embedded in its descriptor, making it easy to identify objects that are live and reachable. % from one of the root-sets\footnote{The root-set for leak detector reachability-analysis includes both kernel and module data structures.}. 
%The leak detector uses the conventional mark and sweep algorithm and mostly-parallel collector~\cite{Boehm:1991:MPG:113445.113459} to perform the reachability analysis.

%are easily identifiable because they hold the descriptor table index in high-order bits and 3) Each watched object holds the memory state information with it. Thus making it easy to identify the the objects which are live and reachable from one of the root-sets. The leak detector uses the conventional mark and sweep algorithm and mostly-parallel collector~\cite{Boehm:1991:MPG:113445.113459} to perform the rechability analysis.

%there is a need to perform rechability analysis of all kernel objects, which will have significant cost.

%then gets accessed through the kernel interface and become invisible to the module and is reachable only through the \texttt{net\_device} structure. Bhavioural watchpoints ensures complete visibility of these objects and enables on-demand instrumentation on its access, marking them live.

%Developing a leak detector only for kernel modules is challenging because of following three reasons:
%Our approach is based on performing reachability analysis on module owned objects. 
%Memory leak is a common problem in Linux kernel and its modules \cite{Boehm96simplegarbage-collector-safety}. A serious memory leak in kernel modules can easily make the system unstable this is because every time kernel losses a piece of memory, it can't be reclaimed until next boot. 
%Our approach of detecting memory leaks in kernel module is to perform rechability analysis on the module owned objects. 
%This is challenging because:

%1) All module owned objects are not directly allocated by the module. For example, a network driver module may make a call to an skb allocator (to allocate a \texttt{sk\_buff} structure) in an interrupt context when the module code is not running. Once the buffer is send over the network, it is the responsibility of the network module to deallocate the buffer.

%2) Module allocates many independent objects which they share with the kernel. Module losses its visibility once such objects are passed to the kernel. For example, a network module during device registration can allocate an object and pass it with \texttt{net\_device} structure. This object then become invisible to the module and is reachable only through the \texttt{net\_device} structure. Detecting leak for such objects requires complete visibility of the kernel. 
%an object and pass it with \texttt{net\_device} structure. 
%till it gets its reference 
% passed to the kernel till it gets his reference back. 
%frequently share with the kernel. Module looses the visibility for such objects till it gets back their references. For example, filesystem module \texttt{ext3} allocates object of type \texttt{inode} and passes it to the kernel. The module holds no reference of the object once it is passed and only 
%loose its visibility for module code till the time kernel passes them back to the module. For example,   
%There is frequent transfer of module allocated objects at kernel/module interface and module may not hold the reference of objects passed to the kernel.  For example, file-system module \texttt{ext3} allocates object of type \texttt{struct inode} and passes it to the kernel. Once it is passed to the kernel module doesn't hold reference of \texttt{struct inode}. Thus having limited view of the module doesn't help in detecting leaks.

%3) Module only holds partial view of the kernel and much of the module allocated objects are reachable only through the kernel objects. To detect the leaked object there is a need to perform rechability analysis of all kernel objects, which will have significant cost.

%and detect the leaked objects it is needed to perform the rechability analysis on

%In Granary, we use wrappers to add watchpoints to the relevant allocators (e.g., \texttt{\_\_alloc\_skb, skb\_clone}). The leak detector keeps track of these watched objects and reports possible leak.

%Behavioral watchpoints are important for implementing the leak detector because: 1) They help in identifying the objects owned by the module, thus limiting the type of objects which needs to be tracked, 2) The watched addresses are easily identifiable because they hold the descriptor table index in high-order bits and 3) Each watched object holds the memory state information with it. Thus making it easy to identify the the objects which are live and reachable from one of the root-sets. The leak detector uses the conventional mark and sweep algorithm and mostly-parallel collector~\cite{Boehm:1991:MPG:113445.113459} to perform the rechability analysis. % from its root-sets. 
%Based on this leak detector divides the watchpoints into three different categories: 
%\begin{enumerate*}
%   \item[i)] Watched objects which are reachable from one of its rootset. They are not considered as leak.  
%   \item[ii)] Watched objects that are not reachable from any of the rootsets are considered as potentially garbage.

   %Watched objects that are not reachable and never owned by the modules are considered as live. This include objects which are allocated in the module context but never reached to the module. Leak detector keeps tracks of such objects by for further analysis.
    %These includes objects which are allocated in module context but are not owned by the modules. Leak detector keep tracks of such objects by marking them as live. 
%   \item[iii)] Module owned objects that are not reachable from any of the rootsets are considered as potentially garbage.

   %Determining the garbage objects with partial view of kernel and scanning of kernel objects passed to the module is challenging. However, to detemine this we follow the ownship history of the objects and any watched objects which is owned by the modules previous and are not rechable from any of the root-sets is reported as leak. Root-sets are set of object references which are prior reachable and for our leak detector it includes kernel and module static data and kernel objects passed to the module.  
%\end{enumerate*}

%in performing reachability analysis, as the addresses are identifiable and it stores the state of a memory

%allocation of \texttt{struct sk_buff} in interrupted context and it is the responsibility of the module to deallocate it.

%call \texttt{__alloc_skb} in interrupted context to allocated sk

%which is passed to the module.

%Leak detector identifies the module ownership of an object based on its context of allocation and access pattern. Performing rechability-analysis on these objects are challenging because not all module owned objects are directly allocated by them. Granary identifies the

%The module ownership of an object is identified by its context of allocation and access pattern.

%Behavioral watchpoints track the ownership of an object by identifying its source of allocation and access pattern.

%However,

%this is challenging because not all objects owned by the module are directly allocated by them.

%Every time kernel losses a piece of memory, it is gone until the next boot and a serious memory leak in the kernel modules can easily make the system unstable. One method of detecting memory leak in kernel modules is to perform rechability analysis on all allocated objects. This is challenging because not all module owned objects are directly allocated by the module and there is frequent transfer in ownership of the objects at kernel/module interface.

%Behavioral watchpoints track the ownership of an object by identifying its source of allocation and access pattern. To detect the memory leak, Leak detector performs rechability-based analysis only on the objects owned by the module. Study of object ownership also helps in fixing the responsibility for deallocating the object and any default in this is considered as memory leak.

%We use Granary to wrap the kernel’s memory allocator functions (e.g., \texttt{\_\_kmalloc}) and add watchpoints to the objects allocated in the module context. It decides this based on the call graph of the allocator functions. Granary also tracks the kernel objects passed to the module at interface and mark them as module known objects. Leak detector use these module known objects along with kernel and module static data as rootset to perform rechability analysis. Other alternative for rechability-analysis is to scan entire kernel heap.

%Granary also keeps track of all the objects passed to the module at interface and cosiders them as root-set to perform reachability analysis. However, this approach provides leak detector limited view of the kernel and based on that it divides the allocated objects into three categories:
%\begin{enumerate*}
   %\item[i)] Module owned objects that are live and rechable from one of the prior roots are not considerd as leak. 
   %\item[ii)] Watched objects that are not reachable and never owned by the modules are considered as live. This include objects which are allocated in the module context but never reached to the module. Leak detector keeps tracks of such objects by for further analysis.
    %These includes objects which are allocated in module context but are not owned by the modules. Leak detector keep tracks of such objects by marking them as live. 
   %\item[iii)] Module owned objects that are not reachable from any of the rootsets are considered as potentially garbage.

   %Determining the garbage objects with partial view of kernel and scanning of kernel objects passed to the module is challenging. However, to detemine this we follow the ownship history of the objects and any watched objects which is owned by the modules previous and are not rechable from any of the root-sets is reported as leak. Root-sets are set of object references which are prior reachable and for our leak detector it includes kernel and module static data and kernel objects passed to the module.  
%\end{enumerate*}

%for the module-owned objects.

%is to perform the rechability analysis on all kernel allocated objects.

%Memory leak bugs are a common problem with C and C++ based applications \cite{Boehm96simplegarbage-collector-safety}. These bugs are difficult to identify or track-down without the help of a dedicated tool \cite{Memcheck, Bruening:2011:PMC:2190025.2190067}. Memory leak can be a problem especially in long running application and is worse in the kernel because every time a piece of memory is lost, it is gone until the next boot.
%Memory leaks can be a problem in applications, especially those which run for a long time
%Memory leak bugs are common in Linux kernel and its modules which are mostly developed in C. These bugs are difficult to identify or track-down without the help of a dedicated tool. 
%Memory leaks in kernel is worse because of its long running nature and every time a piece of memory is lost, it is gone until the next boot. 
%Linux kernel modules works as an extension of Operating system providing additional features, while running in the privilege mode and a serious memory leak in the kernel modules can make the operating system unstable. In this section we will describe our approach of using behavioral watchpoints in detecting memory leak bugs in kernel modules.

%\paragraph{Liveness Analysis} Our approch of leak detector is based on the liveness analysis of heap objects, which are directly or indirectly allocated by the modules. Similar to existing solution like Purify \cite{Rs_purify:fast} and Dr Memory \cite{Bruening:2011:PMC:2190025.2190067}, leak detector uses garbage collection algorithm to detect memory leaks. %Behavioral Watchpoints being able to hold context-specific information with each objects and are equally efficient for both the methods
%It uses conventional mark and sweep algorithm and recursively follows potential pointers from the data, stack segment and heaps to mark the objects live and reachable. Many implementation of mark and sweep based tracing algorithm exist and a straightforward implementation invloves stopping the client(mutator) to perform the liveness analysis. This solution is not very efficient when client operates on very large heap-memory as it will increase the pause-time and touching heap memory will blow away the cache, seriously affecting the system performance. One solution to reduce the pause time is using generational collection based algorithm. However, this avoids the problem by scheduling the collector thread less frequently, thus doesn't remove the problem completely. In our leak detector we remove this problem by using mostly parallel collector. Mostly-parallel collectors takes an orthogonal approach and reduces the problem by having collector doing most of the work parallel to the client and using stop-the-world to prepare the system.

%running collector parallel to client

%However this doesn't completely remove the

%. The solution to reduce the pause time in tracing algorithm based leak detector is generational or parallel collection. Generational collection avoids stop-the-world time by running the collector thread occasionally, thus doesn't remove the problem completely. However the parallel collectors takes orthogonal approach and reduces the time of pause by running collector parallel to client. We took similar approach of mostly-parallel collector to perform the object liveness analysis.

%A straigntforward implementation of tracing based collector involves stoping the client when collector is in the process performing liveness-analysis. This solution is not very efficient when client operates on very large heap-memory as it stops the clients for longer time. The solution to reduce the pause time in tracing algorithm based leak detector is generational or parallel collection. Generational collection avoids stop-the-world time by running the collector thread occasionally, thus doesn't remove the problem completely. However the parallel collectors takes orthogonal approach and reduces the time of pause by running collector parallel to client. We took similar approach of mostly-parallel collector to perform the object liveness analysis.

%purify uses an algorithm similar to the conventional mark and sweep. In the mark
%phase,i lyify-recursively follows potential pointers from the data and stack segmentsin to the heap and marks all
%blocks referencedin the standard"conservativea"n d "pessimistic"

%embed context-specific information with each objects and are equally efficient for implementing both direct and indirect method of leak detection but we decided to use the tracing algorithm which has no overhead cost for maintaining and updating the records for each objects.

%both direct and indirect method of leak detection but we decided to use the tracing algorithm which has no overhead cost for maintaining and updating the records for each objects.

%The existing solution for memory leak detection uses two methods to perform livness analysis: the reference counting \cite{Maebe_precisedetection} and tracing methods \cite{Bruening:2011:PMC:2190025.2190067}.

%The existing solution for memory leak detection uses two methods to perform the liveness analysis: the direct method and the indirect method. Direct method requires that a record be associated with each object in the heap on behalf of the objects which gets updated based on if a pointer is set to refer to this object or the reference of the object is lost. Reference counting is one of the common direct method for detecting leaks. The indirect method includes the tracing algorithm where the live-ness analysis is performed at later point of time independent of client. Behavioral Watchpoints embed context-specific information with each objects and are equally efficient for implementing both direct and indirect method of leak detection but we decided to use the tracing algorithm which has no overhead cost for maintaining and updating the records for each objects.
%
%\paragraph{Leak Scan}
%A common tracing algorithm used to implement garbage collector is mark-and-sweep operation. We used similar method to perform the liveness analysis on module allocated objects. A straigntforward implementation of tracing based collector involves stoping the client when collector is in the process performing liveness-analysis. This solution is not very efficient when client operates on very large heap-memory as it stops the clients for longer time. The solution to reduce the pause time in tracing algorithm based leak detector is generational or parallel collection. Generational collection avoids stop-the-world time by running the collector thread occasionally, thus doesn't remove the problem completely. However the parallel collectors takes orthogonal approach and reduces the time of pause by running collector parallel to client. We took similar approach of mostly-parallel collector to perform the object liveness analysis.

\paragraph{Types of Leaks}
The three scanning policies for performing liveness analysis on watched objects detect leaks at different levels. Based on the leaks reported by these policies, the detector divides allocated memory objects into three categories:
%Implementing memory leak detector for kernel modules without having full knowledge of kernel is non-trivial. This is because not all module owned objects are directly allocated by the module and there is frequent change in the ownership of the objects at kernel/module interface. Leak detector makes use of wrapper interface of Granary to add watchpoints with all the objects allocated in module context and lazily tracks their ownership at collection. This is particularly important to reduce the false positive. However, our approach of lazily determining object ownership require us to keep track of all the objects passed to the modules but this avoids the scanning of kernel heaps for liveness-analysis. Based on limited kernel view and lazy collection of objects allocated by the modules, leak detector divides them into three categories: 
\begin{enumerate}
   \item[i)] Live objects: These objects have been accessed between leak scans and are thus live and reachable from one of the roots. Objects are marked live after the scanning of accessed objects
   %Objects that are definitely live and reachable from one of the roots. It includes the objects which is getting accessed between the different leak scans. These objects are not stale and can not be considered as leak. The decision about such objects are made after scanning of the accessed objects.
   \item[ii)] Stale objects: These objects have not been accessed for a small number of scans but there is a possibility that they are live. Objects are marked as stale after the scanning of module reachable objects. The decision about these objects are made after performing the full system scan of kernel heaps and static data.
   %Objects that are stale but potentially live. These include objects that do not get accessed for a longer period of time, but there is still a possibility for such objects to be live. An object is moved into this category if it doesn't get accessed for a long period of time but it is not causing memory bloat. The decision about the leak of such objects is taken after scanning the module reachable objects or performing the full system scan of kernel heaps and static data.
   %of such type is not causing the memory bloat. If there is a memory bloat due to the staleness of such objects it is reported as leaked objects. %which are allocated in module context but are not owned by the modules. Leak detector keep tracks of such objects by marking them as live. 
   \item[iii)] Garbage objects: Objects that are not reachable and are certainly garbage. The decision about such objects is made after scanning the kernel heaps and static data of the kernel and its module.

   %Determining the garbage objects with partial view of kernel and scanning of kernel objects passed to the module is challenging. However, to determine this we follow the own ship history of the objects and any watched objects which is owned by the modules previous and are not reachable from any of the root-sets is reported as leak. Root-sets are set of object references which are prior reachable and for our leak detector it includes kernel and module static data and kernel objects passed to the module.  
\end{enumerate}

\paragraph{Meta-information}
The leak detector maintains two kinds of meta-information to track the state of a memory block and to detect leaked objects. This meta-information is added in the descriptor of an object. % whose index gets embedded  with the watchpoint. 
%The analysis requires maintenance of two kinds of metadata. First is the record of the segments in the program, each of which has the following form.
\begin{itemize}
\item \emph{leak\_state :} This includes the state of the memory block. It tracks if the memory is allocated and is accessed in the last epoch. It also counts the number of epochs since the object not been accessed. Our current implementation of the leak detector is aggressive and it performs a full system scan for an object if it does not get accessed in the last two epochs.
%It tracks if the memory is being accessed in last epoch and it also counts the number of epochs since the memory is not accessed.
%\item \emph{source :} This includes the source of allocation and type of allocator used for allocating the memory block and the call-stack for the allocation.
\item \emph{memory\_info :} This includes basic information about a memory blocks such as base address, size and type and source of allocation.
\end{itemize}

%X, a segment-type, describes a segment; this includes its base address, size, location (heap,
%stack, or static), and status (in-use or freed).

%Each segment is given a segment-type X.
%Second is the location metadata shadowing each register and memory word, describing the runtime
%type of the word. This metadata has one of the following three forms.

%n, a non-pointer-type, describes a value which is known to be a non-pointer.
%p(X), a pointer-type, describes a value which is known to be a pointer to a segment X.
%u, an unknown-type, describes a value for which the type is unknown.
%Each word-sized value is shadowed by one of n, u or p(X). In principle, every value produced, of
%any size, can be assigned a type. In practice, the shadow value tracking can be done at word-sized
%granularity because all memory accesses are through word-sized pointers.
%A root-set is de-
%fined by the set of references of objects which are prior
%reachable.

%These are the objects that never passed directly to the kernel and also not rechable from the rootsets. Leak detector reports them as potentially leaked objects.

% kernel and modules frequently changes the ownership of objects at complex interface.

%This is because kernel and modules frequently changes object ownership at complex interface and not all the

%since the modules are tightly integrated with the kernel exchanging objects at complex interface. This makes it challeging to perform the liveness-analysis. Leak detector uses partial kernel information embedded in the Granary to create the root-sets. A root-set is defined by the set of references of objects which are prior reachable. We take a conservative approach in defining root-sets and apart from the static data of the module and kernel, we include all the dynamically allocated kernel objects which are ever passed to the modules as root-set. This avoids the scanning of entire heap but the leak detector has to maintain bookkeeping of such objects. Based on limited kernel view and scanning of root-set objects, leak detector divides the heap allocated objects into three categories: i) Objects that are reachable and not garbage. These are the objects which are reachable from the root-set and not considerd as leak. ii) Objects that are not reachable and potentially garbage. This includes the objects which are directly or indirectly allocated by the modules but are passed to the kernel at some point. Leak detector consider it as a potential object for leak and reports them if large number of such objects exist. iii) Objects that are not reachable and are certainly garbage. These are the objects that never passed directly to the kernel and also not rechable from the rootsets. Leak detector reports them as potentially leaked objects.

%Memory chunks that are certainly garbage and need to be reclaimed.
%Memory chunks that are potentially garbage.
%Memory chunks that are not garbage.

 %livenes-analysis on the heap allocated objects divides the scanning objects into three catergories: i) Memory that is reachable by the root-sets. This is considered as live and acts as the root-set to perform the rechability on other objects.

 %Many applications do not explicitly free memory whose lifetime matches the process lifetime and this is not considered an error.

%Performing rechability analysis on the 
%For performing the reachability analysis on the heap allocated objects, the leak detector divides the scanning objects into three categories: i) Memory that is still reachable by the application. This is not considered a leak. Many applications do not explicitly free memory whose lifetime matches the process lifetime and this is not considered an error. ii) Memory that is definitely not reachable by the application
%(at least, not by an aligned pointer to the start or middle of the allocated block). This is called a leak as there is no way for the application to free this memory: it has lost all references to it. iii) Memory that is reachable only via pointers to the middle of the allocation, rather than the head. This is called a possible leak. These may or may not be legitimate pointers to that allocation.

 %The parallel collector imposes overhead on the mutator but it eliminates long pause due to stop the world. While designing the collector we took middle approach and used mostly parallel tracing collector to perform the reachability analysis. We update the root-set in stop-the-world collector and perform the rechability analysis on these rootsets by running collector parallel to the mutators.

 %concentrate on reclaiming recently allocated objects and it reduces the stop-the-world pause time by running the collector occasionally in-order to perform the reachability analysis on older objects. Thus, the problem of stop-the-world is not completely removed.

 %Generational collection concentrate on reclaiming recently allocated objects and it reduces the stop-the-world pause time by running the collector occasionally in-order to perform the reachability analysis on older objects. Thus, the problem of stop-the-world is not completely removed. Parallel collectors takes orthogonal approach and reduces the time of pause by running it parallel to the mutator(client). The parallel collector imposes overhead on the mutator but it eliminates long pause due to stop the world. While designing the collector we took middle approach and used mostly parallel tracing collector to perform the reachability analysis. We update the root-set in stop-the-world collector and perform the rechability analysis on these rootsets by running collector parallel to the mutators.

%However the tracing technique also suffers from some drawbacks and one of them is long delay in performing the object liveness analysis. One of the approach to smoothly detect the memory leak using tracing technique is to concurrently perform the liveness analysis.

%or root that object is reachable. A most common direct method of garbage detection/collection is the reference counting. In this method each object is associated with the reference count which gets incremented when a pointer is set to refer to this object and is decremented when a reference of the object gets deleted. The reference counting mechanism is advantageous in the sense that the memory management overhead in this case gets distributed throughout the program execution but it is weak in resolving cycles and collecting the cyclic data structure.  The other indirect method includes the tracing algorithm for garbage collection. The tracing algorithm based garbage collector maintains a set of roots and periodically perform the reachability analysis to check the liveness of the objects. The tracing technique has an advantage over the reference counting is that no precaution is needed while handling cycles and also the overhead for maintaining and updating the records(reference counting) is not there.

% developed in C or C++ and can be difficult to identify without the help of a dedicated tool. The problem is particularly important in applications which run for a long time. Memory leak in kernel is worse because of its long running nature and every time a piece of memory is lost, it is gone until next boot. In this section we will describe our approach of using behavioral watchpoints in detecting memory leak bugs in kernel modules. Linux kernel module is a software component which extends the functionality of Operating system. Once loaded these modules executes in privilege mode having full control on the kernel resources and a serious memory leak in the kernel modules can make the system unstable.

%Our approch of leak detection is based on the reachability analysis of heap objects, directly or indirectly allocated by the modules, as opposed to considering unfreed memory as a leak. This is similar to tracing garbage collector’s mark-and-sweep operation. A straigntforward implementation of tracing garbage collector algorithm involves stoping the client when collector is in the process performing liveness-analysis. This solution when applied with mutators operating on very large heap is not very efficient. This is one of the primary argument against using the tracing algorithm in memory leak detection. The other solution to reduce the pause time in tracing algorithm based leak detector is: i) generational collection ii) parallel collection.  Generational collection concentrate on reclaiming recently allocated objects and it reduces the stop-the-world pause time by running the collector occasionally in-order to perform the reachability analysis on older objects. Thus, the problem of stop-the-world is not completely removed. Parallel collectors takes orthogonal approach and reduces the time of pause by running it parallel to the mutator(client). The parallel collector imposes overhead on the mutator but it eliminates long pause due to stop the world. While designing the collector we took middle approach and used mostly parallel tracing collector to perform the reachability analysis. We update the root-set in stop-the-world collector and perform the rechability analysis on these rootsets by running collector parallel to the mutators.

%Performing rechability analysis on the 
%For performing the reachability analysis on the heap allocated objects, the leak detector divides the scanning objects into three categories: i) Memory that is still reachable by the application. This is not considered a leak. Many applications do not explicitly free memory whose lifetime matches the process lifetime and this is not considered an error. ii) Memory that is definitely not reachable by the application
%(at least, not by an aligned pointer to the start or middle of the allocated block). This is called a leak as there is no way for the application to free this memory: it has lost all references to it. iii) Memory that is reachable only via pointers to the middle of the allocation, rather than the head. This is called a possible leak. These may or may not be legitimate pointers to that allocation.

%Each of the leak and possible leak categories is further
%broken down into direct and indirect leaks. An indirect leak is
%a heap object that is reachable by a pointer to its start address,
%but with all such pointers originating in leaked objects. Dr.
%Memory reports the number of leaks, possible leaks, and stillreachable
%allocations, along with the allocation callstack for
%each.

 %can allocate and deallocate them dynamically and any misuse of these resources can lead to memory leaks in the kernel. Previous work on memory leaks is focused to detect leaks in the entire kernel which is not particularly required if we leak is happening in the modules code.

%To detect memory leaks in the kernel modules we perform reachability-based analysis, where a leak is defined as a heap memory that is no longer has any pointer to it as considering any unfreed memory as a leak.

 %to wrap kernel memory allocators and deallocators (e.g., \texttt{kmalloc}, \texttt{kfree}). Wrapped allocators add watchpoints to the addresses returned by the kernel's allocators, and wrapped deallocators remove watchpoints before invoking the kernel's deallocators. Selective shadow memory is maintained as a watchpoint descriptor-specific, variable-sized bitset\footnote{As an optimisation, shadow memory for small objects is embedded in the watchpoint descriptor.}. Each byte of allocated memory corresponds to one bit of descriptor-specific shadow memory. The bits in shadow memory are initialised to zero. Individual bits are flipped to one by write-specific vtable functions when there is a write to the bytes of memory shadowed by those bits. 
\subsection{Evaluation}
We evaluated the performance of the leak detector with the synthetic microbenchmark (\Figref{microbenchmark}) and the macrobenchmark discussed in Section~\ref{sec:approach_eval}. The overhead of the watchpoint instrumentation for the leak detector is shown in Table~\ref{table:leak_detctor-benchmark}. The leak detector adds extra instrumentation to mark objects as having been accessed (i.e live) between the two scans. This adds overhead to the watchpoint instrumentation. 
%This causes extra overhead of the watchpoint instrumentation.

\begin{table*}
\begin{center}
\begin{tabular}{|l|c c c c|}
  \hline
  Benchmark & Native & Watchpoint Null & Everything watched & Leak detector\\
  \hline
  Microbechmark  & 1x & 2.7x & 3.8x & 12x\\
  \hline
\end{tabular}
\caption[Performance impact of leak detector on microbenchmark.]{\label{table:leak_detctor-benchmark}The overhead of watchpoint instrumentation for the leak detector on our microbechmark in shown \Figref{microbenchmark}. The leak detector uses instrumentation to mark live objects between the scans.}
\end{center}
\end{table*}

\begin{table*}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  & Descriptors scan & Module reachable objects & Kernel heaps \& static data\\
  \hline
   CPU cycles & 591822 & 3000269 & 18859648171\\
  \hline
\end{tabular}
\caption[Performance impact of leak scan policies on the collector thread.]{\label{table:leak_detctor-scan_cost}The cost of different scanning policies for the collector. The cost of scanning module reachable objects depends on the number and size of the objects in hash table. At the time of evaluation the hash table was having ``192'' objects with the total scan size of {\texttildelow}4Mb.}
\end{center}
\end{table*}

Section~\ref{sec:object_liveness} discusses three scan policies for implementing the leak scan. We also evaluated the cost of each of the scanning policies. Table~\ref{table:leak_detctor-scan_cost} shows the overhead of the three scan policies in terms of the CPU cycles needed to complete the scan. The cost of scanning module reachable objects depends on the number and size of objects the collector needs to scan. %The cost of other scanning policy will also vary with the number of descriptors, accessed objects and the number of allocated pages \& static data.

We evaluated the effectiveness of each scanning policies in detecting leaks by introducing memory leak in the kernel module. We removed calls to deallocator functions  (\texttt{kfree} and \texttt{kmem\_cache\_free}) from the \texttt{ext3} file system module and we consider all the objects that should be deallocated by \texttt{ext3} module as leaked. We used same experimental setup as discussed in Section~\ref{sec:approach_eval} and mounted \texttt{ext3} module on a RAMDisk of size 1 GB. We used the macrobenchmark with system utilities \texttt{cp} and \texttt{tar} as the workload which was operating on a source tree of \texttt{openssh-6.2}. %Table~\ref{table:leak_detctor-leak_eval} shows the effect of three policies in detecting memory leaks over the period of four scans.

\begin{table*}
\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}
  \hline
      \multicolumn{7}{|c|}{sub-policy 1: Accessed objects are live but are not ``root-sets''}  \\
  \hline
    \multirow{2}{*}{scan} & \multirow{2}{*}{Allocated objects} & \multirow{2}{*}{Accessed objects} & \multicolumn{2}{|c|}{Reachable objects} & \multicolumn{2}{|c|}{Unreachable objects} \\
     \hhline{~~~----}
  & & & Not Leaked & Leaked & Leaked & Not Leaked\\
\hline
  1 & 25 & 24 & 13 & 11 & 0 & 0 \\
  \hline
  2 & 305 & 240 & 145 & 95 & 35 & 30 \\
  \hline
  3 & 2053 & 1545 & 23 & 1522 & 480 & 28 \\
    \hline
  4 & 2069 & 38 & 23 & 15 & 2010 & 21 \\ \hline
   \multicolumn{7}{c}{} \\ \hline
     \multicolumn{7}{|c|}{sub-policy 2: Accessed objects are live and added into ``root-sets'' for reachability analysis  }  \\
  \hline
  1 & 25 & 24 & 13 & 11 & 0 & 0 \\
  \hline
  2 & 540 & 370 & 165 & 283 & 88 & 4 \\
  \hline
  3 & 2129 & 1120 & 40 & 1221 & 862 & 6 \\
    \hline
  4 & 2314 & 39 & 24 & 318 & 1968 & 3 \\ \hline
\end{tabular}
\caption[Profile of allocated objects with leak scan policies. The collector performs scan on objects accessed in last epoch.]{\label{table:leak_detctor-leak_eval}The profile of the allocated objects and detected memory leaks when the scanner thread (collector) considers the accessed objects in last epoch as live and performs reachability analysis. The reachable objects which are not leaked is less than the accessed object in each epoch. This is because many of the allocated objects were transient and should have deallocated before the next scan starts.}
\end{center}
\end{table*}

\paragraph{Scanning of accessed objects:} This policy performs a leak scan only on the objects that are accessed in the last epoch. The knowledge of accessed objects in each epoch allows us to implement two different scanning sub policies. The first sub policy acts as the baseline and considers only the accessed objects as live. It is aggressive in nature and may lead to many false positives.

The second sub policy considers all accessed object as live and include them in the ``root-set'' to perform reachability analysis on watched objects. It reduces the number of false positives by performing shallow scan on these live objects.

 %and include them in ``root-set" to perform reachability analysis on all watched objects. This is a shallow scan of all watched memory reachable from the accessed objects. This reduces the number of stale objects after each scan and there will be less probability to go for full system scan (scanning of kernel heaps and static data). The second sub policy is more aggressive in nature and considers only accessed objects as live and reachable.

Table~\ref{table:leak_detctor-leak_eval} shows the effect of both these sub policies in detecting leaks. It shows there is an increase in false positives with sub policy 1. This is because of its aggressive nature in detecting memory leaks. This reported leak is valid for a kind of workload and may vary with different workloads.

%policy involves the scanning of descriptor table to get the accessed objects in last epoch. These objects are considered as live and are not the leaked objects. As discussed in previous section, this scanning policy is helpful in detecting the stale objects and an objects not getting accessed from the longer time are considered as stale and are potentially leaked objects. The knowledge of accessed objects in each epoch provides the opportunity of two scanning sub policies. The first sub policy considers all the accessed objects in last epoch as live and considering them as ``rootset" to perform the reachability analysis on each watched objects. This decrease the number of stale objects after each scan and there will be less probability to go for full system scan (scanning of kernel heaps and static data). The second sub policy considers only the accessed objects as live and reachable. We studied the effect of both the sub policies in detecting memory leaks.

%It shows the number of allocated objects and the detected leaked objects over the period of four scans.

We also performed the staleness analysis on the allocated objects using this policy. We used an aggressive policy in which objects that are not accessed between two epochs are considered stale. The stale objects are potentially leaked and there is need to perform full system scans if the number of stale objects increases constantly. \Figref{leak_detector-stale} shows an increasing number of stale objects with an increase in the allocated objects. This is because of our aggressive policy of detecting stale objects. The number of these stale objects also depends on the type of workload used for analysis. In this case we used workload which compile the source tree of \texttt{openssh-6.2} and copy the source tree of \texttt{gcc-4.8} on RAMDisk mounting \texttt{ext3} module.

\begin{figure}[t]
\begin{center}
\includegraphics[width=6.0in]{stale_object.pdf}
\end{center}
%\vspace{-15pt}
\caption[Profile of stale object during leak scan.]{\label{fig:leak_detector-stale}The number of stale objects with increase in the number of allocated objects. The scanner thread was scheduled after every 100ms and the workload ran for 17 scan iterations.}
\end{figure}

\begin{table*}
\begin{center}
\begin{tabular}{|r|r|r|r|r|r|}
  \hline
  \multirow{2}{*}{scan} & \multirow{2}{*}{Allocated objects} & \multicolumn{2}{|c|}{Reachable objects} & \multicolumn{2}{|c|}{Unreachable objects} \\
  \hhline{~~----}
  & & Not Leaked & Leaked & Leaked & Not Leaked\\
  \hline
  1 & 27 & 19 & 0 & 8 & 0 \\
  \hline
  2 & 705 & 65 & 11 & 629 & 8 \\
  \hline
  3 & 1426 & 125 & 11 & 1290 & 18 \\
    \hline
  4 & 1824 & 221 & 10 & 1584 & 9 \\ \hline
  \hline
\end{tabular}
\caption[Profile of allocated objects with leak scan policies. The collector performs scan on all module reachable objects.]{\label{table:leak_detctor-leak_eval-policy2}The profile of the allocated objects and detected memory leaks when the scanner thread (collector) uses module reachable objects as ``root-set'' and performs reachability analysis.}
\end{center}
\end{table*}

\paragraph{Scanning of module reachable objects:} The policy considers all objects which are reachable from the module as ``root-set''. We used the same evaluation strategy for this policy and all memory deallocator functions from the \texttt{ext3} file system module are removed. Table~\ref{table:leak_detctor-leak_eval-policy2} shows that the policy detects actual leaks with less false positives and negatives.

\paragraph{Scanning of kernel heaps and static data:} This is the fall back policy for scanning and is not aggressive in nature.  Table~\ref{table:leak_detctor-leak_eval-policy3} shows that the policy reports a lot of false negative. However, it is fairly accurate in detecting leaks and an objects not found during this scan is definitely a leak.

\begin{table*}
\begin{center}
\begin{tabular}{|r|r|r|r|r|r|}
  \hline
  \multirow{2}{*}{scan} & \multirow{2}{*}{Allocated objects} & \multicolumn{2}{|c|}{Reachable objects} & \multicolumn{2}{|c|}{Unreachable objects} \\
  \hhline{~~----}
  & & Not Leaked & Leaked & Leaked & Not Leaked\\
  \hline
  1 & 59 & 52 & 28 & 7 & 0 \\
  \hline
  2 & 789 & 30 & 734 & 25 & 0 \\
  \hline
  3 & 805 & 30 & 305 & 470 & 0 \\
    \hline
  4 & 1623 & 24 & 797 & 802 & 0 \\ \hline
  \hline
\end{tabular}
\caption[Profile of allocated objects with leak scan policies. The collector performs scan on all kernel allocated pages.]{\label{table:leak_detctor-leak_eval-policy3}The profile of the allocated objects and detected memory leaks when the scanner thread (collector) scans all the kernel allocated pages.}
\end{center}
\end{table*}

\paragraph{compare with \emph{kmemleak}}
We also compared the performance of our leak detector with the overhead incurred by \emph{kmemleak}, a leak detector which operates on the whole kernel. We used server benchmarks from Filebench to evaluate performance. \Figref{leak_detector-kmemleak} shows that the performance of our leak detector is comparable to \emph{kmemleak} but our system has overhead only when the module is used where \emph{kmemleak} has overhead on the entire kernel. Also our leak scan policies are customizable and use fall back policy only when there is a constant increase in stale objects. \emph{kmemleak} always uses the fall back mechanism of scanning which is costly.

\begin{figure}[t]
\begin{center}
\includegraphics[width=6.0in,height=3.0in]{kmemleak.pdf}
\end{center}
\caption[Performance impact of leak detector on Filebench server benchmarks.]{\label{fig:leak_detector-kmemleak}The performance overhead of our leak detector when using the Filebench benchmark. It also compares the overhead of leak detector with \emph{kmemleak}, which operates on the entire kernel.}
\end{figure}

### 2013-09-29 thesis/approach.tex

Behavioral watchpoint provides an efficient software-based watchpoint framework that simplifies the implementation of DBT-based program analysis and debugging tools. This chapter describes the design of behavioral watchpoints and the various features enabled by our design. We present the challenges raised by our approach and then describe our implementation of behavioral watchpoints. 
%Behavioral watchpoint leverages the advances in binary instrumentation and code manipulation tool to provide an efficient debugging framework that can substantially increase the feature set of standard off the shelf debuggers. 
Two characteristics that defines the watchpoints are:
\begin{enumerate}[i)]
	\item Context-specific information is embedded in each watchpoint. This information is directly available when a watched address is accessed, providing significant versatility in monitoring a large number of memory addresses.
	%\item The action taken when a watched address is accessed is a component of the context-specific information. This implies that different watchpoints can \emph{behave} differently.
	\item A behavioral watchpoint watches a \emph{range} of addresses, enabling object-granularity watchpoints, i.e., one watchpoint can watch an entire object or memory block.
%watchpoints. This enables the feature where one watchpoint can watch the entire object or memory block.%(i.e., one watchpoint watches an entire object). 
\end{enumerate}

%. This design supports our goal of allowing one watchpoint to monitor all memory accesses to a single object.

%To achieve this design goal, we implemented
\section{Design}\label{sec:design}
The design of the behavioral watchpoint framework is motivated by our aim to provide context-specific information on memory accesses, which helps provide significant versatility when monitoring these accesses. This context specific information is stored in an in-memory data structure and accessed before memory read/write operations. %which is referred before every read/write operations on memory blocks. and accessed before memory read/write operations. 
%
%
Our design is based on the key observation that the pointers in 64 bit architectures have spare bits. Both AMD64 and Intel x86-64 processors use a 48 bit implementation leaving 16 spare bits that can be used to store pointer metadata information. 
%which can be used to store the information about the pointers.

We implement watchpoints by adding an extra level of indirection to memory addresses. An unwatched address is converted into a watched address by changing its spare high-order bits. The high order bits of an unwatched address have the value \texttt{0xffff}. We call this value the canonical value. For watched addresses, the high-order bits are converted to a non-canonical value that helps identify context-specific information about the range of memory being watched. This information, called the watchpoint's \emph{descriptor}, contains the originating watched address, meta-information and a set of functions (\emph{vtable}) to invoke when watched memory is dereferenced. Since watchpoint information is embedded in the high-order bits, a typical offset of a watched address is another watched address that shares the same descriptor.

The design of behavioral watchpoints is shown in \Figref{watchpoint_descriptor_table}. Our design uses 15 high-order bits (called the \emph{counter index}) and an additional 8 bits (bits 20-27, called the \emph{inherited index}) of a watched address to identify the index into a global \emph{watchpoint descriptor table}. The \emph{watchpoint descriptor table} stores a pointer to the watchpoint's descriptor. The key advantage of our design scheme is i) the ability to map watched addresses to unwatched addresses using a simple bitmask and  ii) the ability to easily access a watchpoint's descriptor when a watched address is accessed. The high-order 15 bits counter index allows us to use 32K watchpoints at a time and the additional 8 bits of inherited index extends the number of possible watchpoints to 8M. The inherited index is left unchanged when converting an unwatched address into watched.

However, one drawback of our design is that an offset of a watched address can cause the low-order bits to overflow into the inherited index and this will lead a watched address to point to an incorrect descriptor. One approach to deal with this issue is to assign  multiple descriptors for the watched objects holding the same meta-information and putting them in adjacent indices or duplicating the same descriptor entry across adjacent indices. This is possible because inherited indexes make the descriptor table more sparse allowing them to have duplicate entries.

%15 bits only allows 32K watchpoints. To increase the number of watchpoints, we use
%an additional 8 bits (bits 20-27, called the counter index)
%in the address to index into the watchpoint descriptor table (Figure 1). This counter index  and is left unchanged when
%converting an unwatched address into a watched address.
%The key advantage of our watchpoint scheme is the ability to directly map watched addresses to unwatched addresses using a simple bitmask. The main drawback of
%the scheme is that an offset of a watched address can

%e 48 out of 64 bits in pointers,
%and Windows further limit this to 43 bits for user space
%programs. Thus 21 bits in the pointer representation are
%not used. Next we describe two uses for these spare bits,
%and present a performance evaluation on AMD64.

%Behavioral watchpoints is designed by adding an extra level of indirection to memory addresses. An unwatched address is converted into a watched address by changing its high-order bits. These high-order bits indirectly identify context-specific information about the range of memory being watched. This information, called the watchpoint's \emph{descriptor}, contains the originating watched address, meta-information and a set of functions (\emph{vtable}) to invoke when watched memory is dereferenced. Since watchpoint information is embedded in the high-order bits, a typical offset of a watched address is another watched address that shares the same descriptor.

\begin{figure}[t]
\begin{center}
%\epsfig{file=watchpoints.eps, width=3.0in, height=1.4in}
\includegraphics[width=6.0in]{watchpoints.pdf}
\end{center}
\vspace{-15pt}
\caption[Design of behavioral watchpoints.]{\label{fig:watchpoint_descriptor_table}A watched address (bottom left) and its corresponding unwatched address (top left) are compared. The watched address takes the form of non-canonical address that are not addressable in kernel space. The watchpoint framework uses a \emph{counter index} and an \emph{inherited index} to store the descriptor information in a global descriptor table. The process of resolving the watchpoint descriptor is shown.}
%for the watched address is shown.}
\end{figure}

Our design of behavioral watchpoints is geared to the 64 bit x86 (x86-64) architecture. In kernel space on the x86-64 architecture, the addressable memory includes the canonical form of addresses with their 16 high-order bits set to 1. Tagging these higher order bits with the descriptor information converts them to a non-canonical address that triggers a hardware exception when dereferenced. The watchpoint framework takes advantage of the exception to implement behavioral watchpoints.

%The design of behavioral watchpoint is shown in \Figref{watchpoint_descriptor_table}. It uses 15 high-order bits (called the \emph{counter index}) and an additional 8 bits (bits 20-27, called the \emph{inherited index}) of a watched address to identify the index into a global \emph{watchpoint descriptor table} which stores the pointer to the watchpoint's descriptor. %This extends the number of possible watchpoints to 8M. 
%The key advantage of our watchpoint scheme is the ability to directly map watched addresses to unwatched addresses using a simple bitmask. One drawback of the scheme is that an offset of a watched address can cause the low-order bits to overflow into the inherited index. We overcome this issue by assigning multiple descriptors for the watched objects holding the same meta-information and by putting them in adjacent indices.

%is to manage multiple descriptors for a watched object.

%The watched addresses are non-canonical addresses that trigger a hardware exception when dereferenced. 
%We take advantage of the x86-64 architecture for implementing watched addresses. In kernel space on x86-64, canonical addresses have their 16 high-order bits set to 1. Watched addresses do not take this form; they are non-cannonical addresses that trigger a hardware exception when dereferenced. 
Our watchpoint framework uses two approaches to perform memory operations on watched objects. First, when watchpoints are expected to be triggered frequently, it dynamically adds instrumentation at every memory load and store to avoid hardware exceptions. Watched addresses are detected before they are dereferenced and resolved to their unwatched counterparts (by masking the 16 high-order bits to 1). The approach is called code-centric instrumentation because the decision about the code translation and instrumentation is taken based on the execution path. 
Second, when watchpoints are unlikely to be triggered, the alternative approach is to dereference a watched address and implement behavioral watchpoints in the trap handler. This enables on-demand binary translation and allows adding instrumentation only when a watchpoint is triggered. We call this approach data-centric instrumentation because code translation and instrumentation is based on watchpoint accesses.

%We call it the data-centric instrumentation which enables the feature of adding selective instrumentation. The approach takes decision about code translation and instrumentation based on the access of watchpoints.

%, which lowers the overhead of binary translation. % when watchpoints are fewer in number and unlikely to be triggered.

The design of behavioral watchpoints provides the following benefits:
\begin{enumerate}[i)]
\item \emph{\textbf{Multiple watchpoints can watch the same range of memory:}} 
Two copies of an address (e.g., two pointers to the same object) can be watched separately, so long as the high-order bits index different entries in the watchpoint descriptor table. %they manage different descriptors. 
This is useful for distinguishing between logically different objects that occupy the same memory region. For example, this enables efficient detection of use-after-free bugs without preventing deallocated memory from being immediately reallocated for another use. This efficiency comes from our ability to have one watchpoint for the freed object and another watchpoint for the newly allocated memory occupying the same space.

\item \emph{\textbf{Watchpoint descriptors are context specific:}} 
Our design separates the allocation and management of descriptors from the watchpoint framework. It is the responsibility of each client to manage its descriptors. When a watchpoint gets added to an object, the client determines the \emph{vtable}, \emph{type} and \emph{meta-information} that needs to be stored in the descriptor, as shown in \Figref{watchpoint_descriptor_table}. The vtable determines the function that is invoked when watched memory is accessed. Each vtable provides eight functions: four read and four write functions. Each function is specific to a memory operand size (1, 2, 4, or 8 bytes). A watchpoint descriptor is initialised with either a generic or a type-specific vtable, which is specific to the \emph{type} of the watched address. The meta-information allows the descriptors to be arbitrarily customized or extended based on the needs of the client.

\item \emph{\textbf{Watchpoints are viral:}}
Behavioral watchpoints can be used virally. If an address $A$ is watched, then every address derived from $A$ (e.g., through copying or offsetting) is also watched. This is useful for memory and taint analysis tools. For instance, a watchpoint that is added early in the lifetime of an address (e.g., immediately before the address of newly allocated memory is returned from an allocator) can persist and propagate until no more derived addresses exist.
\end{enumerate}

%Behavioral watchpoints earn their name from vtable because they allow watchpoints to behave differently when watched memory is accessed.

%A watchpoint's descriptor is indirectly located by interpreting these high-order bits as an index into a global \emph{watchpoint descriptor table} which stores the pointer to watchpoint's descriptor(\Figref{watchpoint_descriptor_table}). 
%Because these bits do not change, many addresses end up indirectly referring to the same watchpoint descriptor.

%\subsection{Design Implications}
%Our approach has the following design implications.

%Any unwatched address can be converted into an watched address. The extra level of indirection added by watching an address allows us to efficiently locate context-specific information about the memory being watched.

%A watchpoint is added to a program by changing the high-order order bits of an address used by the program. 
%Watchpoints work on ranges instead of individual units of memory because a watchpoint is added to a program by changing

%\paragraph{A watched address must be distinguishable from an unwatched address.}

%Behavioral watchpoint framework provides API \emph{add\_watchpoints()} which is used to add watchpoints to an arbitrory object. It takes the obeject address and size to creates the descriptor for the object. The descriptor can be based on the type or size of the object. The descriptor can be changed dynamically if required. The descriptor is then stored into descriptor table at the next free index generated by counter and partial index. If some object doesn't need descriptor they are only assigned the watched address with corrosponding index in the descriptor table assigned NULL.

%Behavioral watchpoint provides support where two copies of an address (e.g., two pointers to the same object) can be separately watched, so long as they manage different descriptors, the high-order bits index different entries in the watchpoint descriptor table. 
%This is useful for distinguishing between logically different objects that occupy the same memory. For example, this feature enables efficient detection of use-after-free bugs without preventing deallocated memory from being immediately reallocated for another use. This efficiency comes from our ability to have one watchpoint for the freed memory, and another watchpoint for newly allocated memory occupying the same space.

%\paragraph{Multiple watchpoints can watch the same range of memory.} Two copies of an address (e.g., two pointers to the same object) can be separately watched, so long as they manage different descriptors. %the high-order bits index different entries in the watchpoint descriptor table. 
%This is useful for distinguishing between logicallywatchpoint framework  different objects that occupy the same memory. For example, this feature enables efficient detection of use-after-free bugs without preventing deallocated memory from being immediately reallocated for another use. This efficiency comes from our ability to have one watchpoint for the freed memory, and another watchpoint for newly allocated memory occupying the same space.

%\paragraph{Millions of watchpoints are supported.} Our design as described uses 15 high-order bits (called the \emph{counter index}) and additional 8 bits (bits 20-27, called the \emph{inherited index}) of a watched address for identifying a watchpoint's descriptor.  This extends the number of possible watchpoints to 8M.  %and is left unchanged when converting an unwatched address into a watched address.
%The key advantage of our watchpoint scheme is the ability to directly map watched addresses to unwatched addresses using a simple bitmask. The main drawback of the scheme is that an offset of a watched address can cause the low-order bits to overflow into the inherited index. One solution to overcome this problem is to manage multiple descriptors for a watched object.

%Our design allows the same range of memory to be watched differently. For example, two pointers to the same object can be watched separately, so long as they manage different descriptors. This feature is useful for distinguishing logically different objects that occupy the same memory. For example, this feature enables efficient detection of use-after-free bugs without preventing deallocated memory from being immediately reallocated for use. Having one watchpoint for the freed memory, and another watchpoint for newly allocated memory occupying the same space, is critical for this application. %Watchpoint gets added with a call to \texttt{ADD\_WATCHPOINT} which takes the reference of object and corrosponding meta-information as parameter. \texttt{REMOVE\_WATCHPOINT} is used to remove watchpoint from a watched object.

%For example, this feature enables efficient detection of use-after-free bugs without preventing deallocated memory from being immediately reallocated for another use. This efficiency comes from our ability to have one watchpoint for the freed memory, and another watchpoint for newly allocated memory occupying the same space.

%We are exploring several solutions to this problem.

% This counter index extends the number of possible watchpoints to 8M and is left unchanged when converting an unwatched address into a watched address. The key advantage of our watchpoint scheme is the ability to directly map watched addresses to unwatched addresses using a simple bitmask. The main drawback of the scheme is that an offset of a watched address can cause the low-order bits to overflow into the counter index. We are exploring several solutions to this problem.

% and 8 bits (bits 20-27, called the \emph{inherited index}) for identifying a watchpoint's descriptor

%of a watched address for identifying a watchpoint's descriptor. 15 bits only allows 32K watchpoints. To increase the number of watchpoints, we use an additional 8 bits (bits 20-27, called the \emph{inherited index}) in the address to index into the watchpoint descriptor table (\Figref{watchpoint_descriptor_table}). This counter index extends the number of possible watchpoints to 8M and is left unchanged when converting an unwatched address into a watched address. The key advantage of our watchpoint scheme is the ability to directly map watched addresses to unwatched addresses using a simple bitmask. The main drawback of the scheme is that an offset of a watched address can cause the low-order bits to overflow into the counter index. We are exploring several solutions to this problem.

%\paragraph{Millions of watchpoints are supported.} Our design as described uses 15 of the 16 high-order bits (called the \emph{counter index}) of a watched address for identifying a watchpoint's descriptor. 15 bits only allows 32K watchpoints. To increase the number of watchpoints, we use an additional 8 bits (bits 20-27, called the \emph{inherited index}) in the address to index into the watchpoint descriptor table (\Figref{watchpoint_descriptor_table}). This counter index extends the number of possible watchpoints to 8M and is left unchanged when converting an unwatched address into a watched address. The key advantage of our watchpoint scheme is the ability to directly map watched addresses to unwatched addresses using a simple bitmask. The main drawback of the scheme is that an offset of a watched address can cause the low-order bits to overflow into the counter index. We are exploring several solutions to this problem.

%This approach has some drawbacks when an offset of a watched address causes the low-order bits to overflow into the counter index; however, we have several solutions to this problem.

%\paragraph{Watchpoint descriptors are context-specific.} Our design separates the allocation and management of descriptors from the watchpoint framework. It is the responsibility of each client to manage its descriptors. %, which % and declare the corrosponding descriptor type for the framework. provides complete flexibility in handling it. 
%When a watchpoint gets added to an object, the client determines the \emph{type}, \emph{meta-information} and \emph{vtable} which needs to be stored in the descriptor. These descriptors can be arbitrarily customized or extended based on the needs of the client. %One powerful application of this extension is discussed in section~\ref{sec:access_policies}.

%This aspect of watchpoints is possible because our design separates the allocation/management of descriptors and the addresses that they watch. When a watchpoint is added to an address, the \emph{type} or \emph{size} of the address determines what meta-information is included in the descriptor, as well as what function to invoke when memory watched by the watchpoint is accessed. These descriptors can be arbitrary customized or extended based on the needs of program analysis tools. Two powerful applications of this extension are discussed in \Secref{type_overflow} and \Secref{access_policies}.

%Arbitrary extension of descriptors supports our goal of overcoming the incongruency between the needs of debugging an analysis tools (contextual information about watched memory), and how existing software implements watchpoints (watched memory is opaque).

%\subsection{Extensions}
%Our design includes the following extensions to watchpoints, which expand on the behavioral aspect of our watchpoint implementation.
%principally enable the \emph{behavioral} aspect of our software-based watchpoint .
%Our goal of using watchpoints to watch the memory of objects required the following

%\paragraph{Watchpoints are type-specific.} When a watchpoint is added to an address, the \emph{type} of the address determines what meta-information is included in the descriptor, as well as what functions to invoke when memory watched by the watchpoint is accessed. Two powerful applications of this extension are discussed in \Secref{type_overflow} and \Secref{access_policies}.

%\paragraph{Triggered watchpoint functions are polymorphic.} The function invoked when watched memory is accessed is decided using a descriptor-specific \emph{vtable}. Each vtable provides eight functions: four read and four write functions. Each function is specific to a memory operand size (1, 2, 4, or 8 bytes). A watchpoint descriptor is initialised with either a generic or a type-specific vtable, which is specific to the \emph{type} of the watched address. %Behavioral watchpoints earn their name from vtable because they allow watchpoints to behave differently when watched memory is accessed.

%When invoked, a vtable function operates on the watched address and its descriptor. Behavioral watchpoints earn their name from their ability to behave differently based on the meta-information stored in the descriptor and the (type-specific) vtable function invoked.

%\paragraph{Watchpoints remember their originating address.} The address to which a watchpoint is first added is called its \emph{base address}, and is stored in the watchpoint descriptor. We designed watchpoints to remember their base address because it helps to ``anchor" contextual information. When a watchpoint is added, \emph{something} is known about the watched address. Later in a program's execution, an offset of the watched address might be dereferenced. Little can be said about the dereferenced address in relation to the watchpoint's originating address without knowing the originating address.

%Without the former context of why or where the watchpoint was originally added, there is little that can be said about the relation between the triggered address and 
%the watched memory except that it might be arbitrarily far away from the address to which the watchpoint was originally added.
%That is, something is known about an address when the decision to add a watchpoint to that address is made. 
%If the original address is not remembered, then it is difficult to relate a triggered watchpoint
%We designed watchpoints this way so that at any point during the lifetime of a watchpoint, 
%Watchpoints were designed this way so that contextual information is always ``anchored" to something that was once known.

%This is consistent with our goal of using watchpoints to watch the memory of an object because we expect the base address to be the address in memory of a watched object.

%Watchpoint desc

%Because watchpoints are added to addresses, and

%This means that

%We can change an address into a \emph{watched address} by ensuring that a watched address is non-canonical: it cannot legally be used (on x86) without triggering a hardware exception.

%changing part of an address, not by 
%The watched range is \emph{anchored} on the address on which the watchpoint is initially added (called the base address).

%Behavioral watchpoints are implemented by changing an address to-be-watched into a non-canonical address\footnote{In kernel space on x86, a canonical virtual address has its 16 high-order bits set to 1. Current x86 processors require that the 8 high-order bits of an address match the $9^{th}$ highest order bit.}. The translation from unwatched to watched alters the unused high-order bits of an address.

%There are several implications of this design decision:
%\begin{enumerate}
%	\item 
%\end{enumerate}

%Behavioral watchpoints were designed with the following goals in mind:
%\begin{enumerate}
%	\item 
%\end{enumerate}

%\subsection{Design Challenges}
%The design of behavioral watchpoint put following restrictions in the way it can be used:
%\begin{enumerate*}
%\item[i)] A watchpoint must be added at the object source. Adding watchpoints at arbitrary location must be avoided since there is a possibility of another unwatched copy of the same object existing in the program data. This will introduce incosistency and object will be partially watched.
%\item[ii)] Behavioral watchpoints can't be used in applications playing with high-order bits. Such operation will introduce incosistency in descriptor information attached with watched addresses. Applications doing similar things like depending on sign-extension of the 64-bit addresses or handling of 32-bit addresses is also not supported. Identifying and ignoring such cases is a problem and we are exploring solutions to this.
%\item[iii)] Behavioral watchpoints are implemented for Linux kernel and one should be careful when using page-table lookup for watched addresses. Operations such as \texttt{\_\_pfn\_to\_page} and \texttt{\_\_page\_to\_pfn} can lead to wrong \texttt{page} structure or loss of descriptor information.
%One should also be careful when using kernel functions for page-table lookup on watched addresses.
%We found some cases where fast method of mapping virtual addresses to physical addresses for page table lookup was depending on sign extension of 64-bit addresses which is not the case when it is changed to watched addresses.

 %can't be used in applications playing with high-order bits. Such operation will introduce incosistency in descriptor information attached with watched addresses. Applications doing similar things like depending on sign-extension of the 64-bit addresses or handling of 32-bit addresses is also not supported. Identifying and ignoring such cases is a problem and we are exploring solutions to this.
%\end{enumerate*}

%put some restrictions on the range of applications which it can handle. Behavioral watchpoint tags the higer-order bits with the descriptor information and it can't handle the application which is also doing similar things. Such operation will corrupt the embedded descriptor information and once lost we don't have a way to recover it. Identifying applications doing such operation is challenging. Applications which are only handling with 32-bit address is also not supported. We designed behavioral watchpoints for the Linux kernel but all the functionalities in the kernel is not very suitable for the approach. We found some cases where fast method of mapping virtual addresses to physical addresses for page table lookup was depending on sign extension of 64-bit addresses which is not the case when it is changed to watched addresses.

%has the following design implications.

%are designed for Linux kernel and generated by providing an extra-level of indirection and storing the descriptor information in higher order bits of the address. This puts restrictions on the range of applications which behavioral watchpoints can handle. The applications storing tagging information with the high-order bits of the address works against the design of behavioral watchpoints. Identifying such operations in an application is an open problem for us. Our solution of embedding descriptor information in object address is not very friendly with some of the kernel functions.

%However identifying such operations in an application is an

%However all the applications

%The operation is not very friendly with some of the kernel functions. Linux kernel uses fast method of mapping virtual addresses to physical address. The mapping functions performs bit-operations in the

%and virtual address into struct pages for lookup in the page table.

 %to the physical address. These operations

%There is a requirement for Linux to have a fast method of mapping virtual addresses to physical addresses and for mapping struct pages to their physical address. Linux achieves this by knowing where, in both virtual and physical memory, the global mem_map array is because the global array has pointers to all struct pages representing physical memory in the system. All architectures achieve this with very similar mechanisms, but, for illustration purposes, we will only examine the x86 carefully. This section will first discuss how physical addresses are mapped to kernel virtual addresses and then what this means to the mem_map array.

%\subsection{Architecture}

%The implementation of behavioral watchpoints distinguishes between watched addresses and their descriptors.

%A \emph{watched address} is a pointer with an index into the \emph{watchpoint descriptor table} embedded in its bits (\Figref{watchpoint_descriptor_table}). The $23$-bit index into the descriptor table is formed by concatenating bits $[20,27]$ (called the \emph{counter index}) with bits $[48,62]$ (called the \emph{partial index}). A watched address and its unwatched counterpart share the same counter index; however, the partial index of a watched address varies\footnote{This feature of watchpoints allows for a one-to-one mapping between a watched and unwatched address, and a one-to-many mapping between a watchpoint and all addresses watched by that watchpoint. The one-to-one mapping is formed by masking the high-order bits containing the partial index.}. Partial indexes are recorded in the \emph{partial index counter table}. When a watchpoint is allocated, the current value stored in the counter table for that specific counter index is incremented and returned as the watchpoint's partial index. This allocation strategy allows for at most $2^{15}$ watchpoints per counter index or megabyte of memory\footnote{x86 has byte-addressable memory. The counter index begins at bit $20$, giving each watchpoint $1$ MB = $2^{20}$ B degrees of freedom. However, if an over/underflow across the 1 MB-aligned boundary occurs then the counter index will be corrupted. We can correct one bit of corruption by requiring that the counter index and the partial index have the same sign. An alternative solution is uses a different indexing scheme. We have successfully experimented with a different indexing scheme that solves the aforementioned overflow errors, but sacrifices the one-to-one relationship between a watched and unwatched address.}.

%A \emph{watchpoint descriptor} is a data structure containing a base address, a pointer to a virtual table (vtable) of memory operations, and programmer-defined meta-information.

\section{Implementation}
Behavioral watchpoints are implemented using Granary, a dynamic binary translation (DBT) framework \cite{GranaryAtOSDI} described in Chapter~\ref{sec:Granary}. %Granary instruments arbitrary, binary Linux kernel modules efficiently and without imposing overhead when the core kernel is running. Our implementation uses Granary because it allows us to analyze and debug kernel modules, which are a frequent source of bugs and vulnerabilities in operating systems \cite{BGI,LXFI}.
%
%
%is to use Granary to analyse and debug kernel modules, which are a frequent source of bugs and vulnerabilities in operating systems \cite{BGI,LXFI}.
%
%Granary is unique among DBT systems because it provides three important features: i) mixed-mode execution, ii) policy-driven instrumentation, and iii) reifying instrumentation. The \emph{mixed-mode execution} allows the watchpoint framework to switch the execution between instrumented and native mode implicitly or with the wrapper interface. Granary achieves this by substituting the execution of a function with a \emph{wrapped} version of itself. A wrapped function has the same type specification as its unwrapped counterpart and can freely modify its arguments and return value. Granary can wrap some module functions in this way, even if the module source code is unavailable.
%
%The \emph{policy-driven instrumentation} enables the watchpoint framework track and specialise the instrumentation with the execution context. The knowledge of the execution context helps the framework make decision on adding new watchpoints or collecting the added watchpoints. It also allows the framework to efficiently manage the \emph{context-specific} information in the watchpoint's descriptor.
%
%The \emph{reifying instrumentation} provides the high-level static analysis information to the watchpoint framework. This provides the framework ability to specialise the instruction-level instrumentation with higher-level abstractions. Writing such powerful instrumentation code using low-level DBT abstractions of Granary is hard and it motivated the design of behavioral watchpoint framework.
%
%
%it analyzes and uses static type information of the program. For example, Granary can substitute the execution of a function with a \emph{wrapped} version of itself. 
%A wrapped function has the same type specification as its unwrapped counterpart and can freely modify its arguments and return value. Granary can wrap some module functions in this way, even if the module source code is unavailable. 
%
%While Granary provides a framework for instrumenting kernel modules, we found it was hard to write powerful instrumentation code using low-level DBT abstractions, which motivated the design of behavioral watchpoints. %Next, we describe examples of watchpoint-based debugging applications developed for kernel modules.
%
%\vspace{0.5em}
%Behavioral watchpoints take the form of non-canonical addresses. To implement the behavioral watchpoint, the framework writes the watchpoint instrumentation code for every memory accesses. Granary inserts the watchpoint instrumentation code before every memory load and store instructions before putting them into code-cache. 
We implement behavioral watchpoints by adding watchpoint instrumentation before every memory read (load) and write (store) instruction in the code cache. The watchpoint instrumentation actively looks for non-canonical addresses being used as a source or destination address and triggers watchpoint handling code if it encounters them. The watchpoint framework provides an interface for the client to write the watchpoint handling code efficiently. Our approach of implementing the watchpoint framework has three advantages: %There is two advantages our current approach of watchpoint implementation:

%In order to implement the behavioral watchpoint framework, there is a need to write watchpoint instrumentation for every memory read and write operations. Granary inserts this watchpoint instrumentation code before every memory load and store instructions before putting them into code-cache. These watchpoint instrumentation code actively looks for the non-canonical addresses being used as one of the source or destination target address and triggers watchpoint handling code when it encounters the watched addresses. The current approach of implementing behavioral watchpoints has two advantages: %re is two advantages our current approach of watchpoint implementation:
 %checking code before every memory load and store instructions before putting them in code-cache. The watchpoint checking code actively looks for the non-canonical addresses being used as source or destination addresses and triggers the watchpoint handling code when it encounters the watched addresses. There is two advantages our current approach of watchpoint implementation:
\begin{enumerate}[i)]
	\item The framework makes it easy to add watchpoints with the object. It provides an indirection to the memory address of the unwatched object and tags it with the counter index. It separates the allocation and management of descriptors and provides the client an interface to manage its own descriptors. The framework only manages the global descriptor table and assigns the descriptors to the corresponding indices. %The framework manages the global descriptor table and assign the descriptors in corresponding indices. %  It also sets the descriptor for the watched objects and  %It needs to set the counter index tag with the associated address of the watched objects and sets the 
	 %makes it easy to add watchpoints with the object of interest. We only need to set the counter index tag with the associate address of a watched object and update the corresponding information about the watched objects in the descriptor and global descriptor table. The watchpoint framework takes care of inserting instrumentation before every load and store before putting them in code-cache. Any further execution of the module happens from the code-cache.
	\item The direct mapping of watchpoints with the descriptors make the descriptor information available when the watched objects are accessed. This is helpful in specialising the instrumentation based on the descriptor's information.
	\item The runtime overhead of watchpoint implementation comes from the watchpoint instrumentation. Our implementation of watchpoint instrumentation has a fixed cost and the overhead does not increase drastically with the increase in the number of watchpoints.
\end{enumerate}

\begin{figure*}
\begin{multicols}{2}
\lstset{language=[x64]Assembler, numbers=left, label=Original}
\begin{lstlisting}[basicstyle=\footnotesize\ttfamily, caption=Original instructions]
mov    %rbx,0x340(%r13)
mov    $0x400,%esi
mov    %rax,0x80(%rbx)
\end{lstlisting}
%\label{list:Original}
\lstset{language=[x64]Assembler, numbers=left}
\begin{lstlisting}[basicstyle=\footnotesize\ttfamily, caption=Translated instructions]
lea    0x340(%r13),%rsi
bt     $0x30,%rsi
jb     addr_not_watched_2
bt     $0x2f,%rsi
jae    addr_not_watched_2
callq  granary_bounds_check_8_rsi
bswap  %rsi
mov    $0xffff,%si
bswap  %rsi
LABEL: addr_not_watched_2
mov    %rbx,(%rsi)
mov    $0x400,%esi
push   %rcx
lea    0x80(%rbx),%rcx
bt     $0x30,%rcx
jb     addr_not_watched_3
bt     $0x2f,%rcx
jae    addr_not_watched_3
callq  granary_bounds_check_8_rcx
bswap  %rcx
mov    $0xffff,%cx
bswap  %rcx
LABEL: addr_not_watched_3
mov    %rax,(%rcx)
pop    %rcx
\end{lstlisting}
\label{fig:translated}
\columnbreak
\lstset{language=[x64]Assembler, numbers=left,  label=Overflow}
\begin{lstlisting}[basicstyle=\footnotesize\ttfamily, caption=Overflow detector]
granary_bounds_check_8_rsi :
pushfq 
push   %rax
push   %rdi
push   %r8
mov    %rsi,%rdi
mov    %rsi,%r8
mov    %rsi,%rax
shl    $0x24,%r8
shr    $0x38,%r8
shr    $0x31,%rdi
shl    $0x8,%rdi
or     %r8,%rdi
lea    0x192c98(%rip),%r8   		# 0xffffffffa03cf280 <client::wp::DESCRIPTORS>
lea    (%r8,%rdi,8),%rdi
mov    (%rdi),%rdi
cmp    (%rdi),%eax
jl     overflow_detected
mov    %rax,%r8
add    $0x7,%r8
cmp    %r8d,0x4(%rdi)
jle    overflow_detected
jmp    no_overflow_detected
LABEL : overflow_detected
callq  buffer_overflow_handler
LABEL : no_overflow_detected
pop    %r8
pop    %rdi
pop    %rax
popfq  
retq  
\end{lstlisting}
\end{multicols}
\caption[The baseline watchpoint instrumentation and the instrumentation for buffer overflow detector.]{The native and instrumented version of the instructions performing the basic memory operations. The translated instruction (Listing~\ref{Original}) shows the watchpoint instrumentation required for detecting watchpoints and performs the memory operation at the watched addresses. The buffer overflow detector (Listing~\ref{Overflow}) gets triggered only if the addresses are watched. The overflow detector makes a call to \texttt{buffer\_overflow\_handler} if the overflow is detected. %and sample code for checking read critical section. It checks the violation of Rule0 and verfify that RCU protected data is accessed inside read critical section.
}
\label{fig:mem-write}
\end{figure*}

%The watchpoint framework enables two approach of implementing behavioral watchpoints: the code-centric and the data-centric approach. The code-centric approach provides comprehensive instrumentation by following all execution path, translating and adding the basic blocks in the code cache lazily. The data centric approach selectively perform the instrumentation on demand. Both the approaches takes advantage of the features of granary and provides powerful infrastructure for developing analysis tools. Each approach also has its advantages and disadvantages which is discussed below.

%Behavioral watchpoints framework is implemented to provides support for three instrumentation approaches: the code-centric, the data-centric and mixed instrumentation. Each of them has its own advantages and disadvantages. The program analysis tools developed using watchpoint framework uses these approaches to optimize their performance.

\subsection{Code-Centric Instrumentation}
The Code-centric instrumentation dynamically adds watchpoint instrumentation at every memory reference before putting each basic block in code-cache. Code-centric watchpoint instrumentation detects the dereference of watched addresses and resolves them to their unwatched counterparts before performing the memory operation, thus avoiding any hardware traps that would be generated if a watched address is accessed directly. Code-centric instrumentation has low overhead when compared with trap based instrumentation when watchpoints are expected to be triggered frequently.
%The code-centric instrumentation provides comprehensiveness by following the execution path and translating the basic blocks. It avoids the hardware traps by executing the instrumented version of code from the code-cache. The watchpoint instrumentation detects the dereference of watched addresses and resolve them into their unwatched counterparts before performing the memory operations. 
%The approach benefits from the low cost of instrumentations when the watchpoints are expected to be triggered frequently. The comprehensive instrumentation also benefits from the warm code-cache effect avoiding the cost of translation and instrumentation for future execution.

We implemented behavioral watchpoints using Granary, which comprehensively instruments all module code. Granary detaches itself at the kernel interface thus providing support for code-centric instrumentation only for module code. A module analysis tool developed using the watchpoint framework may need to track the behavior of objects that are shared between the kernel and a module using watchpoints. 
For these shared objects, we use code-centric instrumentation for the module code, and trap-based or data centric instrumentation (with the transitive policy, as described in
the next section) for the kernel code.

%These watched objects often gets leaked to the kernel and dereference of such addresses results in hardware traps. The Granary with its code-centric instrumentations can not avoid these traps.

%In our implementation, the watchpoint framework also provides the code-centric instrumentation only for the module code. It avoids the hardware traps in the kernel code using on-demand data centric instrumentation. Our policy for kernel code uses transitive instrumentation which minimizes the overhead of hardware traps.

%re is high probability of dereferencing watchpoint addresses.

%takes control of the module code and dynamically add
%Behavioral watchpoints take the form of non-canonical addresses that trigger a hardware exception when dereferenced. This design of behavioral watchpoint enables the opportunity of data-centric instrumentation. This enable adding instrumentation selectively by attaching and detaching instrumentation framework on demand, introducing overhead only when they are needed. This approach benefits from the lower overhead of binary translation when several watchpoints are likely to be triggered, and the lower overhead of infrequent traps when watchpoints are unlikely to be triggered. In our implementation of behavioral watchpoints for the kernel modules we used binary translation for the module code and trap-based instrumentation for the kernel code. The instrumentation strategy is based on the assumption that the watchpoint addresses leaked to the kernel are less likely to be get accessed and for such watchpoints, instrumenting entire kernel code will be costly.

\subsection{Data-Centric Instrumentation}\label{sec:data-centric}
Data-centric instrumentation is triggered by hardware traps when a watched address is dereferenced. At this point, Granary is attached so that code can be instrumented. We used three policies for detaching instrumentation: i) basic-block instrumentation, ii) function-only instrumentation, and iii) transitive instrumentation. % and iv) thread instrumentation.

The basic block instrumentation policy detaches instrumentation at the end of the current basic block. This policy translates the minimum amount of code but causes the highest number of traps.

The function-only policy retains the same instrumentation policy across the body of a function and does not allow the transfer of policy across control-transfer instructions. The framework stops instrumenting and detaches itself at the end of function body (i.e, with the \texttt{ret} instruction). Any \texttt{call} to other functions detaches the framework temporarily and then the framework is attached once that function returns. The policy instruments a moderate amount of code with a corresponding decrease in the number of hardware traps, compared to the basic block policy.% the increase in the number of hardware traps.

The transitive policy allows the framework to follow the code-centric instrumentation transferring its instrumentation policy across the \texttt{call} or \texttt{jmp} instructions; the framework detaches itself when the function returns. The transitive policy instruments code aggressively based on the assumption that once the watched address is dereferenced in an execution path, there is high probability of encountering the watchpoints again in the same function or the functions called by this function. Our evaluation result supports this assumption.

The ability to add selective instrumentation by attaching and detaching the watchpoint framework on demand introduces overhead only when instrumentation is needed. The data-centric approach benefits from the lower overhead of infrequent traps when watched addresses are less likely to be dereferenced, e.g., when watchpoints are added selectively for a few objects.

% . The data-centric instrumentation suffers from the high overhead of hardware traps and particularly useful for watching the selective objects.

%dereference, e.g., when watchpoints are added selectively for a few objects.

 %and the framework detaches itself when there is a policy change at the indirect control-transfer. The policy translates and instruments maximum code and put them into code-cache. This avoids the future hardware traps due to the access of watched objects. We designed this policy based on the assumption that once the watched address is dereferenced in an execution path, there is high probability of encountering the watchpoints again in the same function or the functions called by it. Our evaluation result supports this assumption.

%The transitive policy

%The design of behavioral watchpoint enables the opportunity of data-centric instrumentation. The data-centric approach allows the framework add watchpoint instrumentation selectively on demand. The data-centric instrumentation gets triggered by the hardware traps, which happens due to the dereference of watched addresses. The hardware traps allows Granary to attach the instrumentation framework on a memory dereference of watched objects and detaches itself on the next control-transfer instructions.

%The data-centric approach provides three different instrumentation policies: i) transitive instrumentation, ii) function-only instrumentation, and iii) basic block instrumentation.

%The transitive instrumentation allows the framework to transfer its instrumentation policy across the \texttt{call} or \texttt{jmp} instructions and the framework detaches itself when there is a policy change at the indirect control-transfer. The policy translates and instruments maximum code and put them into code-cache. This avoids the future hardware traps due to the access of watched objects. We designed this policy based on the assumption that once the watched address is dereferenced in an execution path, there is high probability of encountering the watchpoints again in the same function or the functions called by it. Our evaluation result supports this assumption.

%The function-only instrumentation retains the same instrumentation policy across the body of a function and does not allow the transfer of policy across control-transfer instructions. The framework stops instrumenting and detaches itself at the end of function body (i.e, with the \texttt{ret} instruction). Any \texttt{call} to other functions detaches the framework temporarily which gets attached once that function returns. The policy instruments moderate amount of code with the increase in the number of hardware traps.

%The basic-block instrumentation policy is conservative in its approach of dynamically adding the translated basic block in code-cache. The policy is true in its nature of on-demand instrumentation and translates a basic block at a time when required. It also does not allow the transfer of instrumentation policy across the control transfer and the instrumentation framework detaches itself at the end of basic block. The policy translates minimum amount of code and handles maximum number of hardware traps. 
 %and instrumentation framework detaches itself on the first encounter of the control transfer instructions.

%The ability to add selective instrumentation by attaching and detaching the watchpoint framework on demand introduces overhead only when instrumentation is needed. The data-centric approach benefits from the lower overhead of infrequent traps when watched addresses are less likely to be dereferenced. The data-centric instrumentation suffers from the high overhead of hardware traps and particularly useful for watching the selective objects.

%During our evaluation we noticed that the selective instrumentation also suffers from the high overhead due to cold code-cache effect. However, this could be avoided by following the transitive instrumentation policy. The data-centric approach provides an opportunity for optimising the code-centric instrumentation where the hardware traps can be used to switch policy between null and watcpoint instrumentation. We have not yet implemented this optimisation for our code-centric instrumentation approach.

%does not allow the transfer of instrumentation policies across a control transfer and instrumentation framework detaches itself on the first encounter of the control transfer instructions. The policy instruments minimum of the module or the kernel code to avoid the watchpoint instrumentation cost. However it suffers from the high cost of traps as the number of hardware-exceptions encountered in this  policy increases significantly.

%policy does not allow the transfer of instrumentation policies on control-transfer instruction but it retains the same policy for instrumenting the function body. The policy only instruments the body of the functions to avoid any further traps by the watched objects.

%high likely of encountering the watchpoints again in the same function or in any function called by it.

%This policy instruments the maximum code once the framework is attached to avoid the high cost of hardware exception. The policy is designed based on the assumption that once the watched address is encountered there is high likely of encountering the watchpoints again in the same function or in any function called by it.

 %by attaching and detaching watchpoint framework on demand introduces overhead only when they are needed. The approach benefits from the lower overhead of infrequent traps when watched addresses are less likely to be dereferenced.

%and it gets detached on next control-transfer instructions or the end of function-body based on the instrumentation policy.

%----

%Behavioral watchpoints are implemented using an extra-level of indirection to the memory addresses. The watched addresses take the form of non-canonical address that trigger a hardware exception when dereferenced. This design of behavioral watchpoint enables the opportunity of data-centric instrumentation. The hardware traps allows Granary to attach the instrumentation framework on a memory dereference of watched objects and it gets detached on next control-transfer instructions or the end of function-body based on the instrumentation policy.

%The ability to add instrumentation selectively by attaching and detaching watchpoint framework on demand introduces overhead only when they are needed. The approach benefits from the lower overhead of infrequent traps when watched addresses are less likely to be dereferenced.

%The data-centric instrumentation provides three policies for instrumenting the module and the kernel code : i) transitive policy, ii) function-body policy, and iii) basic block policy.

%Transitive policy allows the framework to transfer the policy across \texttt{call} or \texttt{jmp} and the framework detaches only on the indirect calls. This policy instruments the maximum code once the framework is attached to avoid the high cost of hardware exception. The policy is designed based on the assumption that once the watched address is encountered there is high likely of encountering the watchpoints again in the same function or in any function called by it.

%Function-only policy does not allow the transfer of instrumentation policies on control-transfer instruction but it retains the same policy for instrumenting the function body. The policy only instruments the body of the functions to avoid any further traps by the watched objects.

%Basic-block policy does not allow the transfer of instrumentation policies across a control transfer and instrumentation framework detaches itself on the first encounter of the control transfer instructions. The policy instruments minimum of the module or the kernel code to avoid the watchpoint instrumentation cost. However it suffers from the high cost of traps as the number of hardware-exceptions encountered in this  policy increases significantly.

 %In our implementation of behavioral watchpoints for the kernel modules we used binary translation for the module code and trap-based instrumentation for the kernel code. The instrumentation strategy is based on the assumption that the watchpoint addresses leaked to the kernel are less likely to be get accessed and for such watchpoints, instrumenting entire kernel code will be costly.or

%\subsection{Mixed Instrumentation}
%Mixed instrumentation enables the opportunity for both code-centric and data-centric instrumentation. The approach benefits from the low overhead of dynamic instrumentation when then there is high likely hood of encountering watchpoints and low overhead of data-centric instrumentation when the watchpoints are less likely to be dereferenced.

\subsection {Instrumentation Optimizations}
The implementation of behavioral watchpoints essentially requires monitoring of all memory read (load) and write (store) operations. A basic monitoring framework using dynamic binary instrumentation inserts new instructions before every memory references looking for the watchpoint addresses. The framework also inserts watchpoint instrumentation before every memory operation. A naive implementation of watchpoint instrumentation performs the following operations:
\begin{enumerate}[i)]
	\item Save and restore the registers, including flags registers, that are used or affected by watchpoint instrumentation. We use the stack to spill these registers. 
% (stack is used to spill these registers). Also spill the flags if watchpoint instrumentation touches any of the flags.
	\item Calculate the referenced address and check if this is one of the non-canonical addresses. This is done by inspecting high-order 16 bits of the addresses. Inject a callback function, if the address is in the non-canonical form. This requires the masking of both user and kernel space addresses. The 47th and 48th bit of the address is reserved and used for this purpose.
	\item Emulate the actual instruction with corrected address to perform the normal load/store operation and continue execution after restoring registers and flags. The instruction emulation is done by replacing the original instruction or by creating a new instruction and providing an indirection.
	\item Continue the normal execution after restoring the registers and flags, if the address is not one of the watched addresses.
\end{enumerate}

The naive implementation of watchpoint instrumentation suffers from high runtime overhead. Granary allows implementing several optimizations for reducing runtime overhead. The watchpoint instrumentation uses the following optimizations:
\begin{enumerate}[i)]
	\item \emph{\textbf{Dead Registers Analysis}:} The instrumentation system needs scratch registers for creating and injecting new instructions. These registers can be obtained by spilling them to the stack or to memory locations. The frequent spilling of these registers on stack or memory location increases the cost of instrumentation significantly~\cite{Probst02registerliveness, Muth98registerliveness}. Granary provides support for managing live registers. It goes over the basic block, performs the register liveness analysis and collects all the registers that can be safely used without spilling them on the stack. The register manager traverses all the instructions in a basic block moving upward and collecting the source and destination registers making all the non-source and non-base-displacement registers (whole destination registers) as dead.

	\item \emph{\textbf{Flag Liveness Analysis}:} In an instrumentation system, the newly injected instructions should not affect the state of the processor. The watchpoint instrumentation ensures this by saving and restoring the status registers before and after the injected instructions. This is costly and so the framework reduces this cost by performing eflag liveness analysis~\cite{Dynst2011}. The watchpoint instrumentation saves and restores the flags register only if it is alive.

	\item \emph{\textbf{Eliminate Stack Operations}:} In the x86 architecture, any operation on the stack or an operation involving the stack pointer is also a memory operation. Watchpoint instrumentation is not required for these memory operations unless it is specified by the client. The watchpoint framework identifies all such instructions and prevents them from getting instrumented. 	Applications such as the stack-based overflow detector require instrumenting any operation involving stack pointers and they need to specifically add such instrumentation. 
	 %Watchpoint instrumentation system performs the register liveness analysis on each basic block to identify registers that can be safely used without requiring it to spill on the stack. For register liveness analysis, instrumentation system starts instrumenting from the end of the basic block moving upward and collecting all the source and destination registers making all the non source and non base-displacement registers (whole destination registers) as dead. Watchpoint instrumentation also performs the eflag Liveness analysis~\cite{Dynst2011} and save and restore flags only if it is alive.
\end{enumerate}

Figure~\ref{fig:mem-write} shows the native and instrumented versions of the instructions performing memory operations. The DBT system provides several other optimizations such as \emph{Group Checks}, which consolidate the two consecutive memory reference checks into a single check if there are no intervening instructions that affect address generation, and \emph{Merge Checks} which exploit the locality of memory references and merge the instrumentation for instructions accessing different members of the same object in the same basic block. We have not implemented all these optimizations in the watchpoint framework because our purpose is not to be exhaustive but rather to demonstrate that watchpoint instrumentation can perform online monitoring of memory references with reasonable overhead.

\section{Limitations}
The design of behavioral watchpoints introduces non addressable memory in the system. The use of non-canonical addresses puts the following restrictions on the way watchpoints can be used. Some of these limitations are application specific and depend on how the program analysis tool uses them.
\begin{enumerate}[i)]
	\item The behavioral watchpoint framework suggests adding watchpoints early in the life of the objects. Adding watchpoints at an arbitrary location should be avoided as it can introduce inconsistency in the program. A program running with both watched and unwatched versions of the same object will provide only partial information about the object. However some program analysis tools such as RCU debugger takes advantage of this limitation and generates two versions of the same objects, each of them getting watched differently.
	%A watchpoint must be added at the object source. Adding watchpoints at arbitrary location must be avoided since there is a possibility of another unwatched copy of the same object existing in the program data. This will introduce incosistency and object will be partially watched.
	\item Behavioural watchpoints uses high-order bits to store the descriptor information. The descriptor provides the \emph{meta-informations} about the object being watched. This puts restriction on the use of behavioral watchpoints in analysing applications which play with the high-order bits of its addresses. Such application will destroy the descriptor information stored with the watchpoints. Applications doing the similar operations such as sign extension of the 64-bit addresses or handling of only 32-bit addresses also loses the descriptor information. The watchpoint framework has no mechanism to recover the lost descriptor information from the watchpoints. Identifying and ignoring such applications or the operations in an application is an important challenge for the watchpoint framework.

%Identifying and ignoring such cases is a problem and we are exploring solutions to this.
	\item Behavioural watchpoints are implemented for the Linux kernel. The Linux kernel uses bitwise operations for the fast page-table lookup. These bitwise operations converts the virtual page addresses into the page frame number and assumes that the 16 high-order bits of the addresses are all one. Linux kernel also maintains different address space regions and perform different functions for the page table-lookup. The design of behavioral watchpoints are not friendly with these operations.

	%and one should be careful when using page-table lookup for watched addresses. Operations such as pfn to page and page to pfn can lead to wrong page structure or loss of descriptor information.
\end{enumerate}

\section{Evaluation}\label{sec:approach_eval}
In this section, we evaluate the performance overhead of behavioral watchpoints. The watchpoint framework implements behavioral watchpoints by injecting instrumentation for every memory read (load) and write (store) operations, introducing runtime overhead in the system. We evaluated the cost of watchpoint instrumentation, including the cost of handling hardware traps for data-driven instrumentation, using synthetic microbenchmarks.

%The behavioral watchpoints also introduces the concept of data-driven instrumentation that triggers the selective instrumentation on hardware traps. We used microbenchmark to evaluate the cost of handling hardware traps.

We also measured the overhead of using behavioral watchpoints on the throughput of common file I/O operations using the \emph{iozone}~\cite{citeulike:919086} file system benchmark and the performance of real-world workloads using several file system utilities and the Filebench benchmarks. 
%We also evaluated the overhead of the watchpoint framework on file system utilities and measure the performance of real-world workload using filebench.
We ran all our experiments on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU with 4GB physical memory. In our experimental setup, we used the \texttt{ext3} file system module that was mounted on a 1GB RAMDisk (mounting loads the \texttt{ext3} and \texttt{jbd} journaling kernel modules). The watchpoint framework wraps the kernel memory allocators to add watchpoints for all newly allocated objects.

%\textbf{TODO}: explain why we are using filesystem module and iozone for the evaluation

%We also evaluated the overhead of using behavioral watchpoint framework on common file operations.

%Microbenchmark is used to evaluate the cost of watchpoint instrumentation which is required for identifying the watchpoints and performing the memory operations at the watched addresses. We used \emph{Iozone}~\cite{citeulike:919086} filesystem benchmark for measuring the overhead of behavioral watchpoints on common filesystem operations. In our experimental setup, the \texttt{ext3} filesystem was mounted on a 1GB RAMDisk (mounting loads the \texttt{ext3} and \texttt{jbd} journaling kernel modules). We also evaluated the overhead of using behavioral watchpoint framework on common file operations.
%We implemented behavioral watchpoint framework which provides support to instrument module code. Hence, behavioral watchpoints provides support for two approaches: data-driven and mixed instrumentation.

\begin{table*}
\begin{center}
%\begin{centering}
\begin{tabular}{|l|c c c|}
  \hline
  Optimizations & Native & Watch Null & Watch All \\
  \hline
  Code-driven (with BB optimizations)  & 1x & 2.7x & 3.8x \\
  \hline
  Code-driven (without BB optimizations) & 1x & 5.0x &	6.2x \\
  \hline
  Data-driven approach  & 1x & 1x & 283x \\
  \hline
\end{tabular}
\caption[Performance impact of watchpoint instrumentation on microbenchmark.]{\label{table:microbenchmark_optimizations}The performance overhead of watchpoint instrumentation on a synthetic microbenchmark (\Figref{microbenchmark}). The code-driven approach evaluates the cost of using watchpoint instrumentation and the benefit of basic block optimizations. The overhead of the data-driven approach is dominated by the cost of handling hardware traps.}
\end{center}
\end{table*}

\begin{figure}
\begin{lstlisting}[language=C,basicstyle=\footnotesize\ttfamily]

typedef unsigned long long cycles_t;
cycles_t currentcycles() {
    unsigned cycles_low, cycles_high;
     __asm__ __volatile__ (
			"cpuid\n\t"
			"rdtsc\n\t"
			"mov %%edx, %0\n\t"
			"mov %%eax, %1\n\t": "=r" (cycles_high), 
			"=r"(cycles_low):: "%rax", "%rbx", "%rcx", "%rdx");

    return ((unsigned long long)cycles_low) 
		| (((unsigned long long)cycles_high) << 32);
}
enum {
	NUM_OUTER = 10,
	NUM_INNER = 100000
};

/* Initialize the LKM */
volatile unsigned long x[2] = {100, 200};
int init_module() {
    int i,j;
    unsigned long addr;
    unsigned long flags;
    volatile struct foo_test *ptr = 
		(struct foo_test*)kmalloc(sizeof(struct foo_test), GFP_KERNEL);

    cycles_t before[NUM_OUTER], after[NUM_OUTER];

    for(j=0; j < NUM_OUTER ; j++)
    {
	preempt_disable();
	raw_local_irq_save(flags);
    	before[j] = currentcycles();
    	for(i = 0; i < NUM_INNER; ++i) {
		__asm__  volatile("mov %rax, %rax;");
		(ptr)->l1 += x[i % 2];
		__asm__  volatile("mov %rax, %rax;");
    	}
    	after[j] = currentcycles();
	raw_local_irq_restore(flags);
	preempt_enable();
    }

    return 0;
}
\end{lstlisting}
\vspace{-10pt}
\caption[The synthetic microbenchmark used to evaluate the watchpoint framework.]{\label{fig:microbenchmark}Synthetic microbenmark used to evaluate the watchpoint instrumentation. It performs basic memory operation inside a tight-loop.}
\end{figure}

\subsection{Microbenchmark}
Our microbenchmark consists of a tight-loop of memory operations that exhibits the worst-case overhead of using watchpoint instrumentation. \Figref{microbenchmark} shows the synthetic microbenchmark we used to evaluate the overhead of watchpoint instrumentation and hardware traps. We also evaluated the benefits of our optimization schemes such as register-liveness analysis and flag-liveness analysis used for watchpoint instrumentation. We call this basic block optimizations.

Table~\ref{table:microbenchmark_optimizations} shows the maximum cost of using watchpoint instrumentation and handling a hardware trap. \emph{Watch None} represents the overhead of baseline watchpoint instrumentation when no watchpoints have been added, and \emph{Watch All} shows the overhead when all module allocated objects are being watched. The added overhead of \emph{Watch All} comes due to the indirection required for masking watched addresses and emulating the original instructions.

Table~\ref{table:microbenchmark_optimizations} also shows the high overhead of handling hardware traps. In the data-centric approach, the hardware trap triggers the instrumentation of the module code. The high overhead of a hardware trap shows that the approach is useful for selective instrumentation when watchpoints are less likely to be triggered.

%The \emph{watchpoint\_null} does not add any overhead because the hardware traps never gets triggered and instrumented code never gets executed.

\begin{figure}[t]
\begin{center}
\includegraphics[width=4.5in]{thesis_code_driven.pdf}
\end{center}
\caption[Performance impact of code centric instrumentation.]{\label{fig:watchpoint_performance_code_driven}Throughput in MB/sec of common file system operations for code-centric instrumentation. Direct IO is enabled to bypass the effect of the OS buffer cache on read requests. It compares the overhead of the watchpoint framework with the performance overhead of Granary and with native system.}
\end{figure}

\subsection{Iozone filesystem benchmark}
We used the \emph{iozone}~\cite{citeulike:919086} file system benchmark to measure the overhead of behavioral watchpoints on the throughput of common file system operations. In our experimental setup, we used the \texttt{ext3} file system module that was mounted on a RAMDisk of size 1GB. We enabled direct IO to avoid the effect of buffer cache on file system. The \emph{iozone}, in throughput mode, created two processes (reader, writer) that perform the file I/O operations on a file of size 480Mb with a record size of 4Kb. The watchpoint framework wraps the two most commonly used memory allocators in \texttt{ext3} and \texttt{jbd}: \texttt{\_\_kmalloc} and \texttt{kmem\_cache\_alloc}, to add watchpoints on the allocated objects. We evaluated both the code-centric \& data-centric approaches using \emph{iozone} and compared their performance.

%----

%We evaluated both code-centric and data-centric approach using \emph{iozone} and compared their performance in terms of the number of active watchpoints. 
%In data-centric instrumentation approach the behavioral watchpoint does not uses wrappers for attaching and detaching Granary at the interface. The framework attaches on traps raised by dereferencing of watched addresses and it detaches at the end of basic block or when the function returns. In this case we wrap the allocators by hot-patching kernel text and add watchpoints only when the allocation happens by the module code.

\begin{table*}
\begin{center}
\vspace{1em}
\begin{tabular}{|l|r|r|}
  \hline
  \multicolumn{3}{|c|}{Watchpoint statistics for the code-centric instrumentations}  \\ \hline
  \hline
  & Watch None & Watch All \\
  \hline
  Number of basic blocks & 2249 & 5495\\
  \hline
  Number of basic blocks with watched memory operations & 0 & 1208\\
  \hline
  Number of executed basic blocks & 238782485 &  825836778 \\
  \hline
  Number of dynamic memory operations & 568643857 & 1693113417 \\
  \hline
  Number of watched memory operations & 0 &199495293 \\
  \hline
  Number of kernel hardware traps & 0 & 12630779 \\
  \hline
\end{tabular}
\caption[Watchpoint statistics for code centric instrumentation. The watchpoints are added on all module allocated objects.]{\label{table:code-centric-watchpoint_stats} The statistics of the watchpoint framework for code-driven instrumentation when module allocated objects are watched. It also shows the effect of watched objects leaked to the kernel. The increase in the number of hardware traps causes an increase in the number of basic blocks and the number of executed basic blocks.}
\end{center}
\end{table*}

\paragraph{Code-centric instrumentation:}
The code-centric approach comprehensively instruments the module code and dynamically adds watchpoint instrumentation at every memory reference. This baseline watchpoint instrumentation causes overhead even when there are no added watchpoints. In the code-centric approach the module always runs under the control of Granary and executes code from the code-cache. It uses the kernel and module function wrappers for fast attaching and detaching at the interface. %This causes an extra overhead on any interactions between the kernel and the modules.

In the code-centric instrumentation, we first evaluated the overhead of watchpoint instrumentation and the cost of using Granary as the underlying DBT system. This is important to understand the baseline cost of using the watchpoint framework.

\Figref{watchpoint_performance_code_driven} represents the overhead of code-centric instrumentation on the throughput of file I/O operations. The overhead of the watchpoint framework increases to {\texttildelow}70\% when all objects allocated by the module are watched. This is because many of these watched objects are shared and leak to the kernel. When these objects are dereferenced in the kernel, they cause hardware traps because Granary detaches itself at the kernel interface and the watchpoint instrumentation is no longer added to the memory references. These hardware traps are costly and reattach the watchpoint framework which then instruments the kernel code transitively.

\begin{table*}
\begin{center}
%\caption{Performance of macrobenchmark}
\begin{tabular}{ |l||r|r|r| }
\hline
\multicolumn{4}{ |c| }{Access pattern of module allocated objects} \\ \hline
\hline
Memory Allocator & Size & Module Accesses Count  & Kernel Accesses Count\\ \hline
\multirow{5}{*}{$\_\_kmalloc$} & 8 & 1233865 & 0 \\ \cline{2-4}
 & 50 & 720 & 288\\ \cline{2-4}
 & 51 & 648 & 240\\ \cline{2-4}
 & 59 & 792 & 528\\ \cline{2-4}
 & 4096 & 23452 & 0 \\ \hline
 \hline
 \multirow{7}{*}{$kmem\_cache\_alloc$} & 16 & 31533 & 0 \\ \cline{2-4}
 & 24 & 25019188 & 0\\ \cline{2-4}
 & 32 & 2348 & 0\\ \cline{2-4}
 & 64 & 2469028 & 3099\\ \cline{2-4}
 & 112 & 21404553 & 0 \\ \cline{2-4}
 & 192 & 19902520 & 4916028 \\ \cline{2-4}
 & 768 & 57760206 & 37373718 \\ \cline{2-4}
 & 1024 & 34865132 & 7940485 \\ \cline{2-4}
 & 8196 & 299 & 821648 \\ \hline
 \hline
 \multirow{1}{*}{$\_get\_free\_page$} & 4096 & 820273 & 2973 \\ \hline
\end{tabular}
\caption[Memory access pattern of module allocated objects.]{\label{table:access_pattern_shared_objects}The memory accesses patten of the module allocated objects. It represents the number of times an object is getting accessed by the module and the kernel code. The objects are classified based on its size and the memory allocator used for allocation. %The post analysis  shows that file system \texttt{inode} objects with memory size 768 are most accessed object by the kernel. A watchpoint on \texttt{inode} object will generate maximum number of hardware traps.
}
\end{center}
\end{table*}

\begin{table*}
\begin{center}
\vspace{1em}
\begin{tabular}{|l|r|r|}
  \hline
  \multicolumn{3}{|c|}{Watchpoint statistics for the code-centric instrumentations}  \\ \hline
  \hline
  & Watch none \texttt{inode} & Watch All \\
  \hline
  Number of basic blocks & 2874 & 5495\\
  \hline
  Number of basic blocks with watched memory operations & 732 & 1208\\
  \hline
  Number of executed basic blocks & 266083908 &  825836778 \\
  \hline
  Number of dynamic memory operations & 594195510 & 1693113417 \\
  \hline
  Number of watched memory operations & 92600064 &199495293 \\
  \hline
  Number of kernel hardware traps & 5285701 & 12630779 \\
  \hline
\end{tabular}
\caption[Watchpoint statistics for code centric instrumentation. The watchpoints are added on all the module allocated objects except file \texttt{inode}s.]{\label{table:code-centric-node-inode-watchpoint_stats} The watchpoint statistics of the framework for code-driven instrumentation when all objects allocated by the modules (except \texttt{inode}s) are watched.}
\end{center}
\end{table*}

Table~\ref{table:code-centric-watchpoint_stats} represents the number of hardware traps encountered and number of executed basic blocks for the code centric instrumentation. The additional increase in the number of basic blocks ({\texttildelow}2.5{\footnotesize$\times$}) and the number of executed basic blocks ({\texttildelow}3.5{\footnotesize$\times$}) comes due to the kernel code instrumentation on hardware traps. It also shows that the one out of every sixteen watched memory operations causes hardware traps.

For understanding the source of these hardware traps, we developed a tool using the watchpoint framework which tracks the accesses of module allocated objects. We typed these objects based on their size and the memory allocator used for its allocation. Table~\ref{table:access_pattern_shared_objects} shows the access pattern of the module allocated objects. The post processing analysis shows that file system \texttt{inode} objects with its size ``768'' gets accessed by the kernel maximum number of times and adding watchpoint on \texttt{inode} objects cause maximum number of hardware traps.

We verified this by removing watchpoints from all \texttt{inode} objects and running \emph{iozone} in throughput mode. \Figref{watchpoint_performance_code_driven_none_inode} shows that the overhead for \emph{Watch All} decreases by {\texttildelow}50\% and is close to the overhead of baseline instrumentation i.e, \emph{Watch None}. The small differences in the overhead of \emph{Watch All} and \emph{Watch None} are because there were still some shared objects which was getting watched and causing hardware traps. Table~\ref{table:code-centric-node-inode-watchpoint_stats} shows the details about such objects.

\begin{figure}[t]
\begin{center}
\includegraphics[width=4.5in]{thesis_code_none_inode.pdf}
\end{center}
\caption[Performance impact of code centric instrumentation. The watchpoints are added on all module allocated objects except file \texttt{inode}s.]{\label{fig:watchpoint_performance_code_driven_none_inode}Throughput in MB/sec of common file system operations for code-centric instrumentation when \texttt{inode} objects are not getting watched. The overhead of \emph{Watch All} decrease because of the less number of hardware traps and executed basic blocks.}
\end{figure}

The evaluation of code-centric instrumentation shows that the hardware traps on the kernel code are the major source of overhead in implementing behavioral watchpoints. These hardware traps can be removed by providing full kernel support to the code-centric instrumentation. Granary, being the underlying DBT-system, currently does not provide support for whole kernel instrumentation.

%To understand the source of these hardware traps, we developed a tool using the watchpoint framework

%We also studied the source of these hardware traps

%The evaluation shows that a major contribution in the overhead of code-centric instrumentation comes due to Granary and its wrapping mechanism.

 %\Figref{watchpoint_performance_code_driven} shows the drop in throughput of filesystem operations when using the watchpoints instrumentation. As mentioned earlier \emph{watchpoint\_null} represents the overhead of baseline watchpoints instrumentation with no added watchpoints, and \emph{everything\_watched} shows the overhead where module-allocated objects are watched. We used code-centric instrumentation approach for the evaluation.

%In the previous section, we discussed that due to the limitations of Granary to instrument only the module code, our implementation of code-centric instrumentation is limited. We use code-centric instrumentation for executing the module code and data-centric instrumentation for the kernel code if the watchpoint leaks to the kernel. We also use three different policy for the code-centric instrumentation: transitive instrumentation policy, function-only instrumentation policy and basic-block instrumentation policy. The detail about these policies are discussed in section. Before evaluating the complete data-centric approach, we evaluated the compared the effect of three policies on code-centric instrumentation.

%\Figref{watchpoint_performance_code_driven} shows the overhead on the throughput of different filesystem operations. We also collected the watchpoint statistics for the the three policies. Table~\ref{table:data-centric-table} shows the number of hardware-exception encountered and the ration of executed basic block containing watched and unwatched addresses.

\begin{figure*}[t]
\begin{center}
\includegraphics[width=4.5in]{thesis_data_driven.pdf}
\end{center}
\caption[Performance impact of data centric instrumentation.]{\label{fig:watchpoint_performance_data_driven}Throughput in MB/sec of common file system operations for data-centric instrumentation. It shows the overhead of three data-centric policies: transitive policy, function only, and basic block only policy.}
%for the watched address is shown.}
\end{figure*}

\paragraph{Data-centric instrumentation:}
We implemented data-centric instrumentation with three detach policies: i) transitive policy, ii) function-only policy and iii) basic block policy. Section~\ref{sec:data-centric} discusses each of the three policies in detail. %The data-centric instrumentation adds watchpoints on newly allocated objects through hot-patching and wrapping the kernel memory allocators.

\begin{table*}
\begin{center}
\vspace{1em}
\begin{tabular}{|l|r|r|r|}
  \hline
  \multicolumn{4}{|c|}{Watchpoint statistics for data-centric instrumentation (with different detach policies)}  \\ \hline
  \hline
  & transitive & function & block \\
  \hline
  Number of basic blocks & 5755 & 3335 & 1144 \\
  \hline
  Number of basic blocks with watched memory operations & 1479 & 1218 & 1141\\
  \hline
  Number of executed basic blocks &  833167407  & 108965164 & 37378155\\
  \hline
  Number of dynamic memory operations & 1702295819 & 565666812 & 251985370\\
  \hline
  Number of watched memory operators & 336149753 & 234340514 & 206334774\\
  \hline
  Number of kernel hardware traps & 13403663 & 42259345 & 97518601\\
  \hline
  Number of module hardware traps & 112172 & 23496407 & 42902684\\
  \hline
\end{tabular}
\caption[Watchpoint statistics for data centric instrumentation.]{\label{table:data-centric-table} The statistics of the watchpoint framework for data-centric instrumentation when all the objects allocated by the modules are watched. It compares the three detach policies used by the data-centric approach.}
\end{center}
\end{table*}

We evaluated the overhead of data-centric instrumentation with \emph{iozone}, using the same experimental setup as with the code-centric instrumentations. We disabled the kernel and the module wrappers since they allow Granary to take complete control over the module code and prohibit the module code from running native. In data-centric instrumentation, the module code gets executed natively when there are no added watchpoints.

Before comparing the code-centric and data-centric approaches, we first evaluated the three detach policies of data-centric instrumentation. \Figref{watchpoint_performance_data_driven} shows the overhead of each policy on the throughput of file I/O operations. The transitive policy instruments code aggressively and performs better than the function only and basic block detach policies. Table~\ref{table:data-centric-table} shows the number of executed basic blocks and the number of hardware traps encountered when each of the three detach policies are used. The transitive policy encounters the minimum number of hardware traps and executes the largest number of basic blocks. %The transitive policy also has large number of basic block with watched memory operations and this is because it adds more number of watchpoints on the newly allocated objects.

% executes the module code natively when no watchpoints get added and does not cause any runtime overhead in the system. It also uses three policies to reduce the cost of handling hardware traps. We first evaluated the three policies of data-centric instrumentation and compared their performance overhead along with the number of basic block executed and the hardware traps encountered by each of them. \Figref{watchpoint_performance_data_driven} shows the overhead of each policies on the throughput of file I/O operations for \texttt{ext3} module. The transitive policy has slight advantage over the function only and basic block policy because of the decrease in the number of hardware traps it needs to handle. There is also an increase in the number of basic block executed and this adds to the overhead of the transitive policy. Table~\ref{table:data-centric-table} shows how the number of executed basic blocks and the number of hardware traps encountered varies across the three policies. It also shows that the number of hardware traps encountered in module code is always less than the hardware traps at the kernel code. It represents that a majority of watched objects are shared and leaks to the kernel. These hardware traps also causes a significant overhead in the code-centric instrumentation and our approach of code-centric is more biased towards the data-centric.

\begin{figure}[t]
\begin{center}
\includegraphics[width=6.0in]{thesis_data_code.pdf}
\end{center}
\caption[Performance impact of Data centric Vs Code centric instrumentation.]{\label{fig:watchpoint_performance}Throughput in MB/sec of common file system operations for the code-centric and data-centric approaches. The watchpoints are added on all objects allocated by the module.}
%for the watched address is shown.}
\end{figure}

%of the two reasons: i) the data-centric instrumentation handles more hardware traps which is costly, and ii) the number of watchpoints added in transitive detach policy is more than the code-centric instrumentations.

%In spite of increase in the number of hardware traps in the data-centric approach, it is performing better than the code-centric approach. This is because of the increase in the number of executed basic blocks. Also the code-centric approach suffers significantly due to the large number of hardware traps in the kernel code. The evaluation of three policies of data-centric approach also shows that the increase in the overhead due to hardware traps is getting compensated with the less number of basic block being executed. Table~\ref{table:data-centric-table} shows that the number of hardware traps encountered in transitive policy is 5x less than the basic block policy but it is executing 10x more basic block and thus not getting much advantages in terms of performance. The code centric approach is also using the kernel and the module wrappers which has an additional cost.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.0in]{thesis_watch_inode.pdf}
\end{center}
%\end{figure}

%\begin{figure}
\begin{center}
\vspace{1em}
\begin{tabular}{|l|r|r|}
  \hline
  \multicolumn{3}{|c|}{Watchpoint statistics for selective instrumentation in data centric instrumentation }  \\ \hline
  \hline
  & Watch \texttt{inode}s & Watch all \\
  \hline
  Number of basic blocks & 4217 & 5755\\
 % \hline
  Number of basic blocks with watched memory operations & 562 & 1479 \\
  \hline
  Number of executed basic blocks &  735367411 &  833167407 \\
  \hline
  Number of dynamic memory operations & 1519057148 & 1702295819 \\
  \hline
  Number of watched memory operators & 87433900 & 336149753 \\
  \hline
  Number of kernel hardware traps & 5522314 &  13403663 \\
  \hline
  Number of module hardware traps & 11234 & 112172 \\
  \hline
\end{tabular}
%\caption{\label{fig:watchpoint_inode_compare} Throughput in MB/sec of common file system operations and the watchpoint statistics when watchpoints are added selectively and only on \texttt{inode} objects.}
\caption[Performance impact of selective instrumentations for data-centric approach.]{\label{fig:watchpoint_inode}Throughput in MB/sec of common file system operations and the watchpoint statistics for data-centric instrumentation when watchpoints are added selectively and only for \texttt{inode} objects. The evaluation is done for transitive detach policy.}
%for the watched address is shown.}
\end{center}
\end{figure}

\Figref{watchpoint_performance} compares the overhead of code centric and data centric instrumentation. It shows that the code centric instrumentation performs better than the data centric instrumentation when all objects allocated by the module are watched. This is because the code centric instrumentation handles less hardware traps, which is costly and affects the performance of the system. Our evaluation compares the code centric instrumentation with the data centric instrumentation using transitive detach policy.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.0in]{thesis_watch_inode_compare.pdf}
\end{center}
%\caption{\label{fig:watchpoint_inode_compare}Throughput in MB/sec of common file system operations when watchpoints are added selectively and only for \texttt{inode} objects. The evaluation is done with transitive instrumentation policy.}
%for the watched address is shown.}
%\end{figure}

%\begin{table*}
\begin{center}
\vspace{1em}
\begin{tabular}{|l|r|r|}
  \hline
  \multicolumn{3}{|c|}{Watchpoint statistics for selective instrumentation (watch \texttt{inode} objects)}  \\ \hline
  \hline
  & Code centric & Data centric \\
  \hline
  Number of basic blocks & 5405 & 4217 \\
 % \hline
  %Number of basic blocks with watched memory operations & 563 & 563 \\
  \hline
  Number of executed basic blocks &  814496751  & 735367411 \\
  \hline
  Number of dynamic memory operations & 1684738365 & 1519057148\\
  \hline
  Number of watched memory operators & 91699675 & 87433900 \\
  \hline
  Number of kernel hardware traps & 5446745 & 5522314 \\
  \hline
  Number of module hardware traps & 0 & 11234 \\
  \hline
\end{tabular}
\caption[Performance impact of selective instrumentations. The watchpoints are added on all file \texttt{inode} objects.]{\label{fig:watchpoint_inode_compare} Throughput in MB/sec of common file system operations and the watchpoint statistics when watchpoints are added selectively and only on \texttt{inode} objects.}
\end{center}
%\end{table*}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.0in]{thesis_selective_watch_no_inode.pdf}
\end{center}
\vspace{1em}
\begin{center}
\begin{tabular}{|l|r|r|}
  \hline
  \multicolumn{3}{|c|}{Watchpoint statistics for selective instrumentation (no hardware traps on kernel code)}  \\ \hline
  \hline
  & Code centric & Data centric \\
  \hline
  Number of basic blocks & 2259 & 1171 \\
  \hline
  Number of basic blocks with watched memory operations & 179 & 189 \\
  \hline
  Number of executed basic blocks &  238740476  & 105345884 \\
  \hline
  Number of dynamic memory operations & 568556578 & 221904835\\
  \hline
  Number of watched memory operators & 45649624 & 45637811\\
  \hline
  Number of module hardware traps & 0 & 10244480 \\
  \hline
\end{tabular}
\caption[Performance impact of selective instrumentations. The watchpoints are added on objects that are not getting accessed by the kernel code]{\label{fig:watchpoint_none_inode_compare} Throughput in MB/sec of common file I/O operations and the statistics of the watchpoint framework when watchpoints are added to the objects selectively which are not getting accessed by the kernel. This removes the effect of hardware traps from the code-centric instrumentations.}
\end{center}
%\end{table*}
\end{figure}
%on the throughput of file operations when only inode gets watched. In this case we used transitive policy for the evaluation.

\paragraph{Selective instrumentation:}
The data centric approach enables selective instrumentation by adding watchpoints on selected objects. We first evaluated the performance of selective instrumentation by adding watchpoints only on file \texttt{inode} objects. \Figref{watchpoint_inode} shows the overhead of selective instrumentation and compares it with the overhead of data-centric instrumentation when all the module allocated objects are watched. Selective instrumentation performs better since it adds less watchpoints thus encountering fewer hardware traps, and executing fewer instrumented basic blocks.

We also compared the performance of selective instrumentation when using code-centric and data-centric approaches. \Figref{watchpoint_inode_compare} compares the overhead of selective instrumentation for both the approaches. It shows that selective instrumentation using data centric instrumentation has less overhead than with code centric instrumentation. This is because fewer basic blocks gets executed with data-centric instrumentation. The number of hardware traps encountered in both the approaches are also approximately same, causing similar overhead.

The evaluation of code-centric approach shows that a major source of overhead in code-centric instrumentation comes due to adding watchpoints on file \texttt{inode} objects. These objects gets accessed mostly by the kernel code causing hardware traps. We removed the effect of these hardware traps from code-centric instrumentation by adding watchpoints selectively on the objects accessed only by the modules. We took the help of table~\ref{table:access_pattern_shared_objects} for adding watchpoints on the objects selectively.

 \Figref{watchpoint_none_inode_compare} represents the overhead of selective instrumentations for data-centric approach and compares it with the code-centric instrumentation. It shows that the selective instrumentation reduces the overhead of using the watchpoint framework to {\texttildelow}5\% where as the baseline instrumentation for code-centric approach causes an overhead of {\texttildelow}40\%. This is because the selective instrumentation executes less number of basic blocks ({\texttildelow}2.3{\footnotesize$\times$}). The evaluation also shows that the performance of sequential write operation in case of data-centric approach is less than the code-centric. We analyzed the behavioral of write operations and found that {\texttildelow}50\% of the total hardware traps was only happening during sequential write operations. This could be a reason of sequential write not performing better.

 The evaluation of both, the code-centric and data-centric approaches for implementing behavioral watchpoints, shows an interesting trade-off between the amount of instrumented code executed from the code-cache and how much of that code actually needs to be instrumented. The code-centric approach always suffers from the overhead of baseline instrumentation since it always executes instrumented basic block from code cache. The data-centric instrumentation overcome this by adding instrumentation on-demand. However, it suffers from the high cost of the hardware traps which triggers the instrumentation. The number of executed basic blocks also increases with the hardware traps which causes additional overhead.

 We also saw that the number of hardware traps depend on the type of the watched objects. Adding watchpoints on frequently accessed shared objects such as file \texttt{inode}s, causes a large number of traps affecting the performance of the watchpoint framework where as watching relatively less accessed objects such as \texttt{ext3\_block\_alloc\_info} causes less traps improving the performance of the data-centric instrumentation.

 The advantages and disadvantaged of both the approaches opens up the space to further explore the possibility of switching between the two approaches at runtime. This will reduce both the overhead of baseline instrumentation and the cost of hardware traps, thus improving the performance of the system. The evaluation of code-centric instrumentation also shows its need to provide the whole kernel instrumentation support. We plan to take this as an immediate future work.

 %The evaluation of the watchpoint framework using \emph{iozone} shows the

 %It shows that the selective instrumentation using data-centric approach reduces the overhead of using the watchpoint framework

 %for both the code-centric and data-centric approaches.

 %by adding watchpoints on such objects. The data centric instrumentation in such case performs much better than the code centric approach and the performance of data centric instrumentation is close to that of the native system.

%and added watchpoints selectively on objects which does not get accessed by the kernel.

%, we saw that a major source of overhead in code centric approach is the number of hardware traps. Much of these hardware traps occur due to adding watchpoints on \texttt{inode} objects and the code centric instrumentation was performing worse when it was getting watched. To remove the effect of hardware traps in selective instrumentation, we took the help of table~\ref{table:access_pattern_shared_objects} and added watchpoints selectively on objects which does not get accessed by the kernel. \Figref{watchpoint_none_inode_compare} compares the overhead of selective instrumentations by adding watchpoints on such objects. The data centric instrumentation in such case performs much better than the code centric approach and the performance of data centric instrumentation is close to that of the native system.

%the number of instrumented basic block executed in case of data-centric approach is less than the code-centric approach. 
%However, it also shows that there are more hardware traps encountered with data-centric approach.

%The watchpoint gets added to the objects selectively. We evaluated the performance of selective instrumentation by adding watchpoints only to the \texttt{inode}s. \Figref{watchpoint_inode} compares the overhead on the throughput of file operations when only inode gets watched. In this case we used transitive policy for the evaluation.   
% the overhead of code-centric and data-centric approach

%using complete data-centric approach.

%from providing the complete data-centric approach by wrapping all the function pointers at the interface for faster attach and detach.

%and we also used the same experimental setup for running the benchmark. We also disabled the wrappers for fast attach and detach since it prohibits the framework from providing the complete data-centric approach by wrapping all the function pointers at the interface for faster attach and detach.

%\subsection{Filebench}
%We evaluated the real workload using Filebench file system benchmark.

\subsection{Macrobenchmark}
We further evaluated the overhead of the behavioral watchpoint framework using a file system macrobenchmark consisting common system utilities. We used the same experimental setup and mounted the \texttt{ext3} module on a RAMDisk of size 1GB. The macrobenchmark operated on the Linux source tree (\texttt{linux-3.2.50}). Table~\ref{table:system_utils-benchmark} shows the overhead of behavioral watchpoints on the standard system utilities. For both the code-centric and data-centric approaches, the framework adds watchpoints selectively on all \texttt{inode} objects allocated from the look-aside buffer. The evaluation shows data-centric instrumentation is performing better than code-centric instrumentation when used for watching objects selectively all the \texttt{inode} objects.

\begin{table*}
\begin{center}
\vspace{1em}
\begin{tabular}{|l|r|r|r|}
  \hline
   & Native execution & Data centric approach & Code centric approach \\
  \hline
  cp & 6.64 & 10.82 & 11.61\\
  \hline
  tar & 2.64 & 3.84 & 4.16\\
  \hline
  stat & 3.90 & 5.12 & 5.44\\
  \hline
  grep & 3.94 & 5.36 & 5.54\\
  \hline
\end{tabular}
\caption[Performance impact of the watchpoint framework on macrobenchmark]{\label{table:system_utils-benchmark}The CPU (system \& user) time for the standard system utilities performing file system operations on a RAMDisk of size 1GB. The framework adds watchpoints selectively on all \texttt{inode} objects of the \texttt{ext3} module.}
\end{center}
\end{table*}

\begin{figure*}[t!]
\begin{multicols}{2}
\begin{center}
\includegraphics[width=3.5in]{hardware_traps.pdf}
\end{center}
\columnbreak
\begin{center}
\includegraphics[width=3.5in]{executed_bb_data.pdf}
\end{center}
\end{multicols}
\caption[Cost profile of the watchpoint framework in terms of number of added watchpoints.]{\label{fig:traps_profile}The changes in the number of hardware traps and the number of executed basic blocks with the changes in number of added watchpoints.}
\end{figure*}

In both the code-centric and data centric approaches, the major factors which affects the performance are the number of basic block executed and the number of hardware traps encountered during the execution. We also used microbenchmark to study the changes in the number of hardware traps and the basic block executed with the change in the number of watchpoints. \Figref{traps_profile} shows that both the code centric instrumentation and data centric instrumentation encounters same number of hardware traps. This is because we add watchpoints selectively on all \texttt{inode} objects which gets accessed both by the module and the kernel code equally causing hardware traps. We also see fewer occurrences where the number of traps in case of data centric is less than the code centric instrumentation. One reason for this could be that the code centric instrumentation always detaches itself at the interface where as the data centric instrumentation, with transitive detach policy, does not always detaches itself at the interface.

\Figref{traps_profile} also shows that with fewer watchpoints the code centric instrumentation still executes many more basic blocks than the data centric approach. This causes increased overhead for code centric instrumentation even when fewer watchpoints are added.

% and causes more overhead even when the

% We also discussed previously that data centric instrumentation is useful when watchpoints are added on objects

\begin{table*}
\begin{center}
\vspace{1em}
\begin{tabular}{|l|l|l|}
  \hline
   Workload & Setting & Data Size \\
  \hline
  Fileserver & nfiles=10K & 1.2GB \\
  \hline
  Webserver & nfiles=1K, & 14.76MB \\
  \hline
  Webproxy & nfiles=10K & 154MB \\
  \hline
  Varmail & nfiles=1K & 14.76MB \\
  \hline
\end{tabular}
\caption[Filebench server benchmark workload characteristics.]{\label{table:filebench-benchmark}Benchmark characteristics.}
\end{center}
\end{table*}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.0in]{filebench.pdf}
\end{center}
\caption[Performance impact of the watchpoint framework on Filebench server benchmark workloads.]{\label{fig:sever_performance}Throughput in file system operations/sec for the filebench workload personalities shown in Table~\ref{table:filebench-benchmark}.}
%for the watched address is shown.}
\end{figure}

\subsection{Server Benchmark}
We used the Linux port of Filebench version 1.4.9, with four of the standard server workload personalities, to evaluate the watchpoint framework. We used the same experimental setup as with the \emph{iozone} and enabled default settings to define the workload characteristics. The relevant characteristics for the workloads are shown in Table~\ref{table:filebench-benchmark}. With the default characteristics, the datasets easily fit in the RAMDisk used for evaluation. This does not limit the workload with the performance of I/O operations.

\Figref{sever_performance} shows the throughput of the different Filebench server workloads. For evaluation we used data-centric instrumentation with all file \texttt{inode} objects getting watched. The data centric instrumentation reduces the throughput for all the workload personalities by ~40\%. We also evaluated the code-centric instrumentation using Filebench and noticed similar drop in the throughput but we are not showing it here.

%Null Client reduces
%throughput by 3X for fileserver, and by 4.5X and 4.2X for webserver
%and webproxy, respectively. There is only a small additional
%reduction in throughput with Instruction Count. We can see that
%the drop in throughput is correlated with the number of threads
%used in the workload, with webserver and webproxy both using
%100 threads. The overhead for varmail, which syncs its data to disk
%frequently, is much lower, with only a 1.4X drop in throughput.

%We also evaluated the performance of behavioral watchpoints using real workloads. We mounted \texttt{ext3} file system module on a disk of size 20GB and used Filebench to simulate the real workloads. Table~\ref{table:filebench-benchmark} shows the characteristics of the workload we used for the evaluation.

%We used same experimental setup and mounted \texttt{ext3} module on RAMDisk of size 1GB. The macrobenchmark operated on the source tree of \texttt{linux-3.2.50} branch. We used the following five benchmarks to evaluate the performance overhead of the watchpoint framework:

%\begin{enumerate}
%	\item \emph{MakeDir:} It reads the directories subtree of source program and constructs the identical target directories subtree. It uses \texttt{mkdir} utility for creating the directories recursively.
%	\item \emph{Copy:} It copies all the files from source files subtree to the target files subtree. We used \texttt{rsync} command to copy the files from the source directory to the target directory.
%	\item \emph{ScanDir:} It recursively goes to all the files in target directory and performs the stat operation on each file.
%	\item \emph{ReadAll:} It reads every byte of the every file in the target directory.
%	\item \emph{Compile:} It compiles and links all the files for the source program.
%\end{enumerate}

%Behavioral watchpoints uses non-canonical address to add watchpoints with the allocated objects. This provides the opportunity to implement data-centric instrumentation. Data-centric instrumentation is particularly important when watchpoints are unlikely to be triggered. We implemented three different policy for data-centric instrumentation: i) transitive policy, ii) function-only policy and iii) basic block policy. The details about it is provided in section.

%Data-centric instrumentation also uses three different instrumentation policies and the details of which is provided in section.
%We used \emph{iozone} to evaluate the performance of data-centric instrumentation and we also used the same experimental setup for running the benchmark. We also disabled the wrappers for fast attach and detach since it prohibits the framework from providing the complete data-centric approach by wrapping all the function pointers at the interface for faster attach and detach.

%In this experiment, We studied the effect of hardware exceptions and watchpoint instrumentation overhead on the performance of filesystem throughput. We used the three data-centric policies for the evaluation and compared their performance in terms of the number of hardware exception encountered and the number of instrumented basic block executed.

%Figure~\ref{fig:watchpoint_performance_data_driven} shows the throughput of common filesystem operations for data-centric instrumentation approach. Table~\ref{table:data-centric-table} shows the number of traps encountered along with the number of executed basic block and number of memory operations performed. The evaluation result shows the performance decrease with the number of traps but it gets compensated by executing less number of basic blocks and performing less number of memory operation which has an additional cost of watchpoint instrumentation.

% We also evaluated the number of hardware exception encountered and the number of basic block executed for each policies.

 %We also varies the number of watchpoints to see how the cost of data-centric instrumentation varies with the number of added watchpoints.

%These allocated objects when gets accessed causes hardware exceptions or traps. Behavioral watchpoint framework uses this to implement watchpoint instrumentation on trap. This approach allows selective instrumentation and can take advantage of low overhead when watchpoints are unlikely to be triggered. The trap-based instrumentation approach uses three different instrumentation policy: policy 1 includes the transitive instrumentation where the framework attaches itself on trap and instruments till it finds the indirect call or return and detaches itself. Policy 2 instruments the body of the function which encounters the trap and changes policy to go native on call and return. Policy 3 is the most conservative policy and tries to instruments minimum of the code. The instrumentation framework attaches itself on trap and detaches and goes to native execution on any control transfer, instrumenting only single basic block at a time. We evaluated all the three instrumentation policy by running the \emph{Iozone}~\cite{citeulike:919086} and studying the throughput of common file system operations. In trap-based instrumentation approach behavioral watchpoint is implemented on trap. We also used microbenchmark to evaluate the cost of taking trap and studied how the number of these traps increases over different policy and affects the performance.

%One of the important feature of data-driven instrumentation is the ability to watch objects selectively. The selective addition of watchpoints will have less instrumentation cost and also the number of hardware exception encountered will be less. We evaluated the performance of selective instrumentation by adding watchpoints to only \texttt{inode} objects. Figure~\ref{fig:watchpoint_inode} the overhead when only inode gets watched and compares it with everything watched. However the selective instrumentation of objects may suffer from the cold code-cache effect and it can cause high overhead.

%We also studied the change in throughput of file operations for data-driven instrumentation with the increase in number of watchpoints. With the increase in the number of watchpoints the number of executed basic block and hardware traps increases. Figure shows how the throughput decreases with the increase in number of watchpoints.

%\subsubsection{Iozone}
%In our experimental setup, the \texttt{ext3} filesystem was mounted on a 1GB RAMDisk (mounting loads the \texttt{ext3} and \texttt{jbd} journaling kernel modules). IOzone creates two processes (reader, writer) performing common file operations on a file of size 480Mb with a record size of 1Kb. Watchpoints were added to addresses returned by the two most-used allocators in \texttt{ext3} and \texttt{jbd}: \texttt{\_\_kmalloc} and \texttt{kmem\_cache\_alloc}. \Figref{iozone-workload-overhead} shows the drop in throughput of filesystem operations when using the watchpoints instrumentation; \emph{watchpoint\_null} represents the overhead of baseline watchpoints instrumentation with no added watchpoints, and \emph{watchpoint\_watched} shows the overhead where module-allocated objects are watched. \Figref{iozone-workload-overhead} also shows the overhead of the heap-allocated buffer overflow detector using behavioral watchpoints. Our implementation of the buffer-overflow detector does not yet use various optimizations offered by Granary at the instrumentation and wrapper layer, and thus exhibits high overhead.

%\subsubsection{File operations}

%\paragraph{Code-centric Vs Data-centric Instrumentation}
%We also compared the performance of Code-centric and Trap-based instrumentation approach by running them under best case and worse case scenarios. We used Iozone as the filesystem benchmark and used the similar experimental setup for running filesystem module \texttt{ext3}. Trivially the best case is when there are no watchpoints. To evaluate the worse case we add watchpoints with every allocated memory blocks while Iozone. During the experiment the total number of allocations done by \texttt{ext3} module was ~2500.

%\begin{table*}
%\begin{center}
%\caption{Performance of macrobenchmark}
%\begin{tabular}{ |c|c|r|r|r| }
%\hline
%\multicolumn{5}{ |c| }{Performance of Synthetic Macrobenchmark} \\
%\hline
%\hline
%Macrobenchmarks & time &Native & Data-centric & Code-centric\\ \hline
%\multirow{3}{*}{MakeDir} & user & 0.07 & 0.06 & 0.05 \\ \cline{2-5}
% & sys & 0.13 & 0.08 & 0.08\\ \cline{2-5}
% & real & 1.28 & 0.73 & 0.73\\ \hline
% \hline
%\multirow{3}{*}{Copy}& user & 3.03 & 0.09  & 0.20\\ \cline{2-5}
% & sys & 6.25 & 9.70 & 10.42\\ \cline{2-5}
% & real & 26.35 & 26.45 & 36.66\\ \hline
% \hline
%\multirow{3}{*}{ScanDir} & user & 3.44 & 0.0 & 0.0\\ \cline{2-5}
% & sys & 2.62 & 0.0 & 0.0\\ \cline{2-5}
% & real & 43.10 & 0.0 & 0.0\\ \cline{2-5}
% \hline
% \hline
%\multirow{3}{*}{ReadAll} & user & 2.73 & 0.0 & 0.0\\ \cline{2-5}
% & sys & 2.46 & 0.0 & 0.0\\  \cline{2-5}
% & real & 22.64 & 0.0 & 0.0\\ \cline{2-5}
%\hline
%\hline
%\multirow{3}{*}{Compile} & user & 551.59 & 0.0 & 0.0\\ \cline{2-5}
% & sys & 49.78 & 0.0 & 0.0\\  \cline{2-5}
% & real & 651.30 & 0.0 & 0.0\\ \cline{2-5}
%\hline
%\end{tabular}
%\label{table:compile-benchmark}
%\end{center}
%\end{table*}

%\subsection{Macrobenchmark}
%We further evaluated the performance of behavioral watchpoint framework with the synthetic macrobenchmark representing the real-world workloads. The benchmark uses different system utility functions and operates on a directory subtree containing the source code of a program. The experimental setup was same and we used \texttt{linux-3.2.50} source code for the analysis. The details about the program source code if provided in table. We used following five benchmarks to evaluate the system:

%We created the synthetic benchmark with the reference of Andrew File system benchmark and it perform the following five basic file system operations.
%\begin{enumerate}
%	\item \emph{MakeDir:} It reads the directories subtree of source program and constructs the identical target directories subtree. It uses \texttt{mkdir} utility for creating the directories recursively.
%	\item \emph{Copy:} It copies all the files from source files subtree to the target files subtree. We used \texttt{rsync} command to copy the files from the source directory to the target directory.
%	\item \emph{ScanDir:} It recursively goes to all the files in target directory and performs the stat operation on each file.
%	\item \emph{ReadAll:} It reads every byte of the every file in the target directory.
%	\item \emph{Compile:} It compiles and links all the files for the source program.
%\end{enumerate}

%We used RAMdisk mounted with \texttt{ext3} filesystem as the target directory. For Compile benchmark, we used similar configuration across all builds. We modified the default configuration to remove the loadable module from the build process and we calculated the time for configuration and compilation separately. The result only shows the CPU time for compilation phase. We also used \texttt{strace} to trace the system calls during the build process. Table~\ref{table:compile-benchmark} shows moderate overhead of Granary and behavioral watchpoints with the increase in \texttt{cpu} time.
%Table~\ref{table:compile-benchmark} shows the performance of the synthetic macrobenchmarks. We run these benchmarks on the vanilla Linux kernel version 3.2.50. The details about the kernel source code is given in table.

%For running the benchmark we used Ramdisk mounted with \texttt{ext3} module as the target subtree directory. We created the benchmark to be CPU bound

### 2013-09-29 thesis/related.tex

In this chapter, we discuss research topics that are related to our work. We first focus on the different solutions of implementing software and hardware watchpoints. We later look into the existing applications that can be used to detect memory errors.

\section{Watchpoints}
Watchpoints are an important debugging facility that help users identify data corruption bugs. There have been several proposals in the past on how to implement hardware and software watchpoints.

%We discuss several hardware and software watchpoints approaches below.
%This importance has been given due recognition and there has also been several proposals in the past on how to implement hardware and software watchpoints.
%in the form of hardware debug registers for watching memory locations implemented in almost all state-of-the-art processors. There has also been several proposals in the past on how to implement hardware and software watchpoints.

Greathouse \emph{et al.}~\cite{UnlimitedWatchpoints} propose a hardware solution that efficiently supports an unlimited number of watchpoints. Watchpoints are stored in main memory and utilize two different on-chip cache: a bitmap and a range cache, to accelerate performance. The bitmap lookaside buffer stores fine-grained watchpoints while the range cache can efficiently hold large contiguous regions of watchpoints and is useful for setting watchpoints on ranges of memory. 
%The design of low-overhead unlimited watchpoints helps support the varied needs of the wide range of dynamic analysis tools.
%It introduces two different on-chip caches: a bitmap and a range cache to handle spatial meradata pattern which gives it the benefits in terms of performance. 
Witchel and Asanovic~\cite{Mondrix} describe the implementation of the Mondriaan (also spelled Mondrian) memory protection domain system for the Linux kernel. Mondriaan was designed with fine-grained inter-process protection mechanism in mind, and is optimized for applications that do not perform frequent updates. The protection information is stored in main memory and is cached in a protection lookaside buffer (PLB). The implementation of protection domains using hardware enables fine- and coarse-grained memory protection, the mechanism which is similar to hardware watchpoints.

Zhou \emph{et al.}~\cite{Zhou:2004:IEA:998680.1006720} introduce intelligent Watcher (also known as iWatcher), which provides architectural support for monitoring program execution with minimum overhead. iWatcher stores per-word watchpoints alongside the cache lines that contain the watched data. These bits are initially set by hardware and are temporarily stored into a victim table on cache evictions. The hardware falls back to virtual memory watchpoints if this table overflows. iWatcher can watch a small number of ranges, which must be pinned in physical memory. If this range hardware overflows, the system falls back to setting a large number of per-word watchpoints. In general, this system is inadequate for tools that require more than a small number of large ranges.

Unlike behavioral watchpoints, these approaches depend on specialised hardware and require that applications using this hardware maintain context-specific information separately.

Suh \emph{et al.} \cite{SecureProgramExecFlowTracking} propose a method of secure program execution by tracking dynamic information flow. Their method uses a one-bit tag to indicate whether the corresponding data block is \emph{authentic} or \emph{spurious}. The scheme can be extended to use multiple-bit tags if there are many types or sources of data.
Memory tagging at the hardware level allows their system to track tainted data as it propagates through a running program. Behavioral watchpoints are similar insofar as a watched address is tagged, and this tag propagates through a program.

%makes the case for supporting an unlimited number of watchpoints. A hardware solution is proposed and multiple applications are described. Unlike our approach, the cited approach depends on specialised hardware and requires that applications using these watchpoints maintain their own context-specific information.

%In ,

%\paragraph{Software-based}
Zhao \emph{et al.} \cite{DynamoRIOWatchpoints} describe a method for implementing an efficient and scalable DBT-based watchpoint system. It uses on-demand based dynamic instrumentation to accelerate the software debugger and supports more than a million watchpoints. It uses page protection and indirection through a hash table to track watched memory. This approach does not supports watching ranges of memory, or context-specific information.

Lueck \emph{et al.} \cite{PinADX} introduce semantic watchpoints as part of the PinADX system, an extension of the PIN DBT framework. 
It enables interactive debugging by triggering debugger breakpoints when semantic conditions are met. While similar in spirit to behavioral watchpoints, semantic watchpoints do not maintain context-specific, per-watchpoint state.

Wahbe \emph{et al.}~\cite{Wahbe:1992} also proposes the implementation of software watchpoint using code patching and static analysis. %Copperman and Thomas extended the work Wahbe and use a post-loading technique to insert checks into an executable to solve the shared library issue.
Another interesting approach, proposed by Keppel. \emph{et al.}~\cite{Keppel:93a} is to use checkpoints for memory updates. However all these approaches are valid for userspace programs only. The latest patches for MemCheck in Valgrind~\cite{Seward:2005} also introduce
support of adding watchpoint. They use data structure to maintain the watchlist and track the usage of watchpoints. This puts a severe restriction on their performance.

\section{Memory Debugging}
There are many software, hardware or hybrid approaches proposed for detecting memory bugs. Purify~\cite{Rs_purify:fast} and MemCheck~\cite{Nethercote:2007:SBM:1254810.1254820} are two widely used software tools for detecting memory usage problems. Purify was one of the first tool to combine detection of memory leaks with detection of use-after-free errors. Purify uses a technique called object code insertion (OCI) to insert checking code around every instruction in an application that references, allocates, and deallocates memory. It uses a memory-coloring scheme to keep track of the state of every byte of memory in the application. The link time instrumentation and reporting reads of uninitialized errors immediately result in false positives. MemCheck~\cite{Nethercote:2007:SBM:1254810.1254820} is build on the Valgrind dynamic instrumentation framework. Dr. Memory is another memory checking tool developed using the DynamoRio instrumentation system. Both Memcheck and Dr Memory use a similar approach for tracking memory usage and both of them use Purify’s basic leak detection approach to detect memory leak errors.

The memory debugging tools developed using behavioral watchpoints also maintain the state of a memory block in its descriptor. The descriptor information gets encoded with the memory address and is directly available when the address is accessed.

Parallel Inspector~\cite{reference/parallel/Petersen11a} is a commercial tool built on the Pin~\cite{PinOS} dynamic instrumentation platform that combines data race detection with memory checking. But the details about its implementation are not publicly available. Insure++~\cite{Parasoft00INSU} is another commercial memory checking tool. It supports inserting instrumentation at various points, including in the source code prior to compile time, at link time, and at runtime, but its more advanced features require source code instrumentation. Insure++ also uses DBT-system to insert instrumentation at runtime. Unlike Insure++, Behavioral watchpoints does not require source code support for detecting memory errors. It uses a technique called reifying instrumentation to provide the benefits of high-level static analysis information in specialising instruction level instrumentation.

There have been several proposals to extend current hardware support for debugging memory related errors. DISE~\cite{Corliss:2005:LID:1042442.1043429} (Dynamic Instruction Steam Editing) is a general hardware mechanism for interactive debugging. It adds dynamic instructions for checking memory references into the instruction stream during execution. Other solutions such as SafeMem~\cite{Qin:2005:SEE:1042442.1043428}, iWatcher~\cite{Zhou:2004:IEA:998680.1006720}, and MemTracker~\cite{Venkataramani:2007:MEP:1317533.1318083} have also been proposed for detecting inappropriate usage of memory with low overhead. Unlike behavioral watchpoints, they require specialised hardware to implement their solutions.

%For example, SafeMem exploits the use of ECC (Error-Correcting Code) for detecting memory leaks and memory corruption with low-overhead. But this approach relies on a write-allocate cache policy, i.e., the block is loaded on a write miss. One important drawback of these hardware approaches is that it requires customized, fixed-functionality hardware.

%Beside requiring customized hardware, it does not
%scale well as the number of checks increases when more
%watchpoints are set.

%HeapMon~\cite{Shetty:2006:HHA:1143264.1143273} is hardware/software approach for detecting memory bugs with the helper thread. Hardware such as SafeMem~\cite{Qin:2005:SEE:1042442.1043428}, iWatcher~\cite{Zhou:2004:IEA:998680.1006720}, and MemTracker~\cite{Venkataramani:2007:MEP:1317533.1318083} have also been proposed for detecting in appropriate usage of memory with low overhead. For example, SafeMem exploits the use of ECC (Error-Correcting Code) for detecting memory leaks and memory corruption with low-overhead. But this approach relies on a write-allocate cache policy, i.e., the block is loaded on a write miss. One important drawback of these hardware approaches is that it requires customized, fixed-functionality hardware.

%BoundsChecker~\cite{Jones97backwards-compatiblebounds} monitors Windows heap library calls and detects memory leaks and unaddressable accesses. It does
%not detect uninitialized reads. Some leak detection tools, including LeakTracer~\cite{Hauswirth:2004:LML:1024393.1024412} and mprof [6], only report memory that has not been freed at the
%end of execution. For these tools to be usable, the application
%must free all of its memory prior to exiting, even though it
%may have data whose lifetime is the process lifetime where it is
%more efficient to let the operating system free those resources.
%Reachability-based leak detection, in contrast, uses a memory
%scan that is similar to a mark-and-sweep garbage collector [3]
%to identify orphaned memory allocations that can no longer be
%accessed. This type of leak detection is used by most modern
%memory checking tools, including ours.

%There have been several hardware proposals to extend
%current hardware support for debugging. DISE~\cite{Corliss:2005:LID:1042442.1043429} (Dynamic
%Instruction Steam Editing) is a general hardware mechanism
%for interactive debugging. It adds dynamic instructions for
%checking memory references into the instruction stream during
%execution. Beside requiring customized hardware, it does not
%scale well as the number of checks increases when more
%watchpoints are set.

### 2013-10-24 VEE-14/paper.tex
Initial commit for VEE paper.

\documentclass[preprint]{sigplanconf}
\usepackage{cleveref,amsmath,chngcntr}
\usepackage{textcomp} % for tilde
\usepackage{paralist} % for in-paragraph lists
\usepackage{url}
\usepackage{graphicx}

% For drawing FSMs
%\usepackage{tikz}
%\usetikzlibrary{arrows,automata}

% Used for code snippets
\usepackage{listings,courier}
\usepackage{subfig,epsfig}
\DeclareCaptionType{copyrightbox}
\usepackage{epsfig}

% Settings on code listings.
\lstset{language=C,
		xleftmargin=0pt,
		xrightmargin=0pt,
		framexbottommargin=0pt,
        framextopmargin=0pt,
        framesep=0pt}
\usepackage{enumitem}

% Add listing language for assembly
\lstdefinelanguage
   [x64]{Assembler}     % add a "x64" dialect of Assembler
   [x86masm]{Assembler} % based on the "x86masm" dialect
   % with these extra keywords:
   {morekeywords={CDQE,CQO,CMPSQ,CMPXCHG16B,JRCXZ,LODSQ,MOVSXD, %
                  POPFQ,PUSHFQ,SCASQ,STOSQ,IRETQ,RDTSCP,SWAPGS, %
                  rax,rdx,rcx,rbx,rsi,rdi,rsp,rbp, %
                  r8,r8d,r8w,r8b,r9,r9d,r9w,r9b}} % etc.

% For leaving some comments in the draft.
\newcommand{\comment}[1]{}

% Customize cleverref
\crefname{section}{Section}{Sections}

\begin{document}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{Granary: A Sane Framework for Instrumenting an Insane Environment}

%for single author (just remove % characters)
\authorinfo{Peter Goodman \and Akshay Kumar \and Angela Demke Brown \and Ashvin Goel}
{University of Toronto}{}
% end authorinfo

\maketitle
\subsection*{Abstract}
Kernel modules extend the functionality of operating systems. They represent the bulk of new kernel code in development and contain the majority of new kernel bugs. Unfortunately, analysing and debugging modules is very challenging. Static analysis of module source code is difficult because of the tight interaction between modules and the kernel. Some modules, however, are only distributed in a binary format, which makes static analysis intractable. Existing dynamic analysis tools capable of instrumenting modules are not suited toward module analysis. They either impose unnecessary performance overheads, lack flexibility and are too coarse-grained, or do not provide a means of understanding the ``big picture'' view of what an arbitrary module does as it executes.

We created Granary to address the challenges of module analysis. Granary is a framework for creating flexible and efficient tools that analyse or debug arbitrary, binary Linux kernel modules. Granary uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules. Granary makes it easy to create flexible tools by supporting context-aware runtime code specialisation, and by integrating high-level static analysis information with low-level instruction manipulation. Module analysis tools built on Granary are efficient because of Granary's adoption of a relaxed transparency model, and because of Granary's ability to instrument module code without imposing overhead on non-module kernel code.

\section{Introduction}\label{sec:intro}

Operating system (OS) kernels present an apparently insane environment for a transparent binary instrumentation framework.  For example, hardware interrupts can redirect control flow at any instruction, and interrupt delivery cannot be delayed past the execution of memory instructions that might alter the interrupt delivery state~\cite{DRK}. Preserving interrupt transparency, however, requires that interrupts be delayed until the end of any injected instrumentation instructions.  Together, these requirements make it very difficult to write instrumentation tools.  The insane environment of OS kernels also creates several opportunities for analysis tools. For example, interrupts can be ``tamed'' and put to new uses; privileged hardware features can be used; and the kernel ABI and source code can be relied upon to be a good predictor of module behaviour at the kernel/module boundary. We also found that the kernel is largely insensitive to transparency issues, allowing us to relax transparency in exchange for both improved performance and greater visibility into the execution of instrumented code.  This increased visibility is particularly important for analysis and debugging tools.

The challenges and opportunities of the kernel environment motivated the creation of Granary. Granary is a new binary instrumentation framework for the Linux kernel environment, designed specifically for analysing kernel modules. Granary targets modules because they represent the bulk of new bugs and new code in development in the Linux kernel \cite{FaultsInLinux}. Granary allows developers to create flexible and dynamic analysis tools for arbitrary Linux kernel module binaries. Granary has two main goals: \begin{inparaenum}[i)]
	\item make it easy to create efficient, low-level analysis tools; and
	\item provide high-level static analysis information to the analysis tools.
\end{inparaenum} Granary achieves the first goal by using dynamic binary translation (DBT) to provide low-level access to instructions, memory, and interrupts. Granary achieves the second goal by giving tools direct access to the results of its own static analysis of the Linux kernel source code. By meeting these goals, Granary tools are able to do things like  figure out which data structure fields are associated with each memory access.

Granary addresses three limitations of existing program analysis systems: \begin{inparaenum}[i)]
	\item they have focused on instrumenting all code (the whole kernel: DRK; or whole system: QEMU, PinOS);
	\item they do not allow for flexible runtime code specialisation (Pin, DynamoRIO); and
	\item they do not provide static type information to the instrumented code (Pin, DynamoRIO).
\end{inparaenum}

Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analysed (e.g., modules). In practice, we would like to target instrumentation at specific modules for efficiency, while also allowing rich and detailed instrumentation of these modules and any other relevant code. This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary to instrument module code while non-module kernel code executes with zero overhead. This is challenging because Granary must be comprehensive: all code that is meant to be instrumented should be instrumented. In other words, relinquishing control to execute kernel code natively must not cause Granary to miss the execution of any module code.

Another limitation of existing systems is that they are unable to perform efficient runtime code specialisation, i.e. they are unable to change how code is instrumented based on the context in which that code will run. For example, Granary has a tool that  dynamically constructs inter- and intra-procedural control flow graphs (CFGs). The outputs of this tool enable Granary to do better runtime register allocation, which improves tool efficiency. Part of building CFGs requires tracking whether or not a given basic block is the entrypoint to a function. This is challenging with existing DBT systems because some optimised functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs. Granary solves this problem using demand-based runtime code specialisation: every entrypoint basic block is instrumented in one way, and later re-executions of any entrypoint basic blocks are specialised and instrumented in another way.

In Granary, code is specialised according to the context in which that code will execute. In the previous example, executing code is classified into two contexts: entrypoint and non-entrypoint basic blocks. Granary tracks execution contexts using a technique we call \emph{policy-driven instrumentation}. In a Granary tool, each tracked execution context is associated with its own instrumentation policy. An instrumentation policy decides how to instrument code that will execute within the associated context. In the case of our CFG tool, two policies track the two execution contexts: $P_{\mathit{call\_entry}}$ and $P_{\mathit{after\_entry}}$. $P_{\mathit{call\_entry}}$ applies to the first execution of the first basic block of each function, and connects functions in the inter-procedural CFG.  $P_{\mathit{after\_entry}}$ applies to all other executed basic blocks within each function, including re-executions of the first basic block. $P_{\mathit{after\_entry}}$ connects basic blocks in the intra-procedural CFG. Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks. This avoids typical problems with tracking context in just-in-time DBT systems (e.g., concurrency, arbitrary pre-emption, context-switch\-ing, inability to predict later execution contexts at translation time).

Finally, an important missing feature of existing low-level instrumentation systems is the ``big picture'' view that one gets from static analysis information. Similarly, static analysis tools lack access to instructions, memory, and processor-specific behaviours that are only available to low-level runtime instrumentation tools. Existing mixed static/dynamic analysis tools \cite{NaCl,AddressSanitizer,ThreadSanitizer} appear to get the best of both worlds, but sacrifice on runtime flexibility by committing to a single instrumentation policy at a program's compile time. The benefits of high-level static analysis motivated its inclusion into Granary using a technique we call \emph{reifying instrumentation}. Granary statically analyses the Linux kernel source code to bootstrap learning about  the interfaces that enable modules and the kernel to interact. Granary exposes this information to instrumentation tools in the form of type and function \emph{wrappers}. Tools can use wrappers to apply context-specific policies when instrumenting module code as well as to inspect and manipulate parts of kernel/module memory in a type-safe way.

For example, we are actively developing a tool that uses static type information to learn about all the fields of all kernel data structures that are read from or written to by modules. This information allows us to model things like how a typical file system operates on kernel data structures in the process of opening a file. From these models, we can construct behavioural patterns about how modules use kernel data structures and functions. These patterns can be used to classify modules \cite{DeviceDriverClassification} and identify spurious behaviour \cite{LXFI}.

The rest of the paper describes Granary in more detail. \Cref{sec:modes} describes how to efficiently implement mixed-mode execution. \Cref{sec:policies} describes a new method for dynamically changing how code is instrumented. \Cref{sec:reify} describes how to integrate static analysis information into a DBT system. Finally, \Cref{sec:eval} evaluates Granary's performance.

\section{Dynamic Binary Translation}\label{sec:dbt}

Granary uses dynamic binary translation (DBT) to instrument Linux kernel modules. DBT is used for instruction set translation/emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimisation}, and runtime instrumentation (analysis \cite{DynamoRIO}, security \cite{Vx32,NaCl}, and debugging \cite{Valgrind}). We chose DBT because \begin{inparaenum}[i)]
	\item static binary analysis tools are unable to cope with \texttt{x86}'s mix of variably-sized instructions and data; 
	\item some modules are only distributed in a binary format; and
	\item virtualisation-based approaches are unable to analyse non-virtualisable device drivers \cite{DRK}.
\end{inparaenum}

Granary interposes on the Linux kernel's module loading process. When a kernel module is loaded, Granary bootstraps by translating the first basic block of the module's initialisation function.  It then replaces the pointer to that function with a pointer to the translated basic block. The translation process continues when the kernel initialises the module by invoking the translated module code.

Granary translates and instruments module binaries one basic block at a time. That is, new basic blocks are decoded and translated as execution ``discovers'' those basic blocks. Translated basic blocks are linked together and stored in a globally accessible \emph{code cache}. In Granary, a basic block is a sequence of instructions ending in a conditional branch, \texttt{ret}, or \texttt{jmp}, but not a \texttt{call} instruction.

Granary's just-in-time translation approach means that code executing from the code cache may yield control to Granary to request the address of the next basic block to execute. When instrumented code yields to Granary, a ``context switch'' occurs that transfers execution to a CPU-private stack where Granary operates. Granary context-switches back to the code cache when the next basic block has been found or translated so that instrumented execution may continue. Similar to other DBT systems, Granary uses caching and hotpatching to reduce the number of context switches. In our experience, modules stabilise very quickly: context switches stop happening after most module basic blocks have been translated.

Basic blocks in Granary's code cache contain \texttt{x86-64} binary instructions and meta-data describing those instructions. The meta-data that Granary records in each basic block includes: \begin{inparaenum}[i)]
	\item the length in bytes of the original and translated basic blocks;
	\item the address of the first instruction in the original basic block; and
	\item the policy information (e.g., $P_{\mathit{call\_entry}}$, $P_{\mathit{after\_entry}}$) used to instrument the basic block.
\end{inparaenum} This meta-data can be queried and extended by Granary instrumentation tools. For example, Granary's CFG-building tool represents each node in the inter- and intra-procedural CFGs as an extension of the meta-data stored for each basic block. The meta-data is also queried by interrupt handlers when deciding how to handle interrupts in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block.

%\begin{figure*}[t!]
%\lstset{language=C, tabsize=2, stepnumber=1}
%\begin{multicols}{2}
%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
%struct device_driver {
%	...
%	int (*probe)(struct device *);
%	int (*remove)(struct device *);
%	void (*shutdown)(struct device *);
%	int (*suspend)(struct device *, pm_message_t);
%	int (*resume)(struct device *);
%	...
%	const struct dev_pm_ops *pm;
%	...
%};
%\end{lstlisting}
%\columnbreak
%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
%TYPE_WRAPPER(struct device_driver, {
%    PRE_OUT {
%        ABORT_IF_FUNCTION_IS_WRAPPED(arg.probe)
%        WRAP_FUNCTION(arg.probe);
%        WRAP_FUNCTION(arg.remove);
%        ...
%    }
%    POST_OUT {
%        POST_WRAP(arg.pm);
%    }
%})
%\end{lstlisting}
%\end{multicols}
%\vspace{-1em}
%\caption{The Linux device driver structure is shown on the left. The automatically generated type wrapper for this structure is shown on the right. In the wrapper code, \texttt{arg} is a reference to a \texttt{struct device\_driver} object passed as, or referenced by, an argument to a kernel or module wrapper. Code in the \texttt{PRE\_OUT} section is applied to arguments of the wrapped type before a kernel wrapper is invoked. Similarly, code in the \texttt{POST\_OUT} section is applied to arguments of the wrapped type after a kernel wrapper is invoked. \texttt{POST\_WRAP} invokes the type wrapper that is specific to the value to which it is applied (\texttt{arg.pm}). Type wrappers also support \texttt{\_IN} suffixes instead of \texttt{\_OUT} suffixes, which apply to data going into modules (i.e., over module wrappers). Finally, the \texttt{RETURN\_} prefix is used to apply some code to return values of either kernel or module wrappers.}
%\label{fig:type_wrapper}
%\end{figure*}

\section{Mixed-Mode Execution}\label{sec:modes}
Granary supports two modes of execution: instrumented and native. Module code is instrumented and executes from Granary's code cache, which is under Granary's control. Non-module kernel code runs natively. A mode switch occurs when execution transfers between native and instrumented code. Some mode switches happen naturally (e.g., when instrumented code returns to native code) and other mode switches are mediated by Granary (e.g., when instrumented module code invokes a kernel function).

The mode switch from instrumented module code to native kernel code is easier to detect since the instrumented code runs under Granary's control. Granary treats all kernel functions as \emph{detach} points, where a mode switch from instrumented to native code occurs. At these points, Granary stops executing.

However, Granary needs a way to regain control when native kernel code invokes module code. Our fallback solution for regaining control uses hardware page protection to trap attempts by the kernel to execute native module code. We handle these traps by redirecting execution to instrumented module code. While comprehensive, this approach is not ideal because: \begin{inparaenum}[i)]
	\item trapping on every execution attempt introduces overhead; and
	\item the trap does not provide sufficient information about which interfaces were being used by the kernel to invoke the module.
\end{inparaenum}

Granary's main approach to regaining control is based on the observation that modules tell the kernel about their interfaces by registering functions with the kernel. We expect that at least some of the registered functions will be executed by the kernel because this is the mechanism by which modules extend the kernel's functionality. To Granary, registered module functions represent potential future \emph{attach} points, where a mode switch from native to instrumented code will occur. Furthermore, Granary discovers additional attach points at detach points by observing pointers to module functions that are passed by modules to the kernel.

Granary uses static analysis to dynamically discover attach points by wrapping the kernel/module interface. Kernel functions are wrapped by \emph{kernel wrappers} that inspect and traverse argument pointers in search of pointers to module functions. \Cref{fig:type_wrapper} shows an example of a type wrapper for the Linux device driver structure. If a pointer to a module function (a future attach point) is discovered, then Granary replaces that pointer with a function-specific \emph{module wrapper}. Granary inspects and modifies arguments to module wrappers in the same way as for kernel wrappers. This allows Granary to discover kernel entry points that will cause instrumentation to detach. Finally, Granary redirects execution to the appropriate kernel or instrumented module function after wrapping has occurred.

Kernel and module wrappers invoke \emph{type wrappers} to find and wrap pointers to module functions that are directly or indirectly referenced by kernel/module function arguments. A type wrapper is a function that recursively  traverses the in-memory object graph and converts pointers to module code into pointers to wrapped module functions. Type and kernel wrappers are automatically generated at Granary's compile time by scripts that statically analyse the kernel source code. Granary automatically matches any variable in a kernel wrapper to a type wrapper if the base type (absent pointers, specifiers, and qualifiers) of that variable matches the type wrapper's wrapped type. Similarly, Granary automatically generates module wrappers using a combination of compile-time meta-programming and runtime code generation to match type wrappers to the declared arguments of \texttt{C} function pointer types.  Granary's wrapper approach has two benefits: (i) it is more efficient than the trap-based alternative, and (ii) it gives instrumentation tools more information about the executing module (\Cref{sec:reify}).

After Granary bootstraps on the module initialisation process, \emph{attaching} occurs in one of three ways: \begin{enumerate}
	\item {\bf Implicit attaching:} the kernel returns to instrumented module code in the code cache.
	\item {\bf Fast attaching:} the kernel invokes a wrapped module function.
	\item {\bf Slow attaching:} the kernel invokes unwrapped module code. If the kernel executes a module function which was passed to the kernel in a type-unsafe manner, then the processor will raise a fault because Granary uses hardware page protection to prevent module code from being executed. Granary handles these faults by returning execution to the instrumented version of the faulting module code.
\end{enumerate}

Granary \emph{detaches} when control transfers from instrumented code to native (uninstrumented) code. Detaching occurs in one of two ways: \begin{enumerate}
	\item {\bf Implicit detaching:} instrumented module code returns to the original kernel, or is interrupted (initial interrupt handling is done by the kernel).
	\item {\bf Wrapped detaching:} instrumented module code invokes a kernel wrapper, which later transfers control to the kernel.
\end{enumerate}

\subsection{Transparency}\label{sec:transparency}

An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}. Granary includes configurable levels of transparency to address the trade-off between transparency and overhead. By default, Granary uses a relaxed transparency model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility. By default, Granary exposes the following three artifacts: code cache addresses as function and interrupt return addresses, and wrapped module function addresses in shared data structures.

\vspace{-3pt}\paragraph{Function return addresses}\label{para:return_address_transparency} Unlike most DBT frameworks, Granary defaults to inlining \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses. Inlining \texttt{call} instructions avoids return-address mispredictions and unnecessary control-flow by building longer basic blocks. If Granary is configured to use transparent return addresses, then implicit attaching and detaching are disabled because they depend directly on returning to instrumented module or native kernel code. Granary will fall back to slow attaching to handle kernel-to-module returns, and indirect branch lookups for module-to-module returns.

\vspace{-3pt}\paragraph{Interrupt return addresses} Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts. A special case arises for page fault exceptions that occur within specific kernel functions that access user space data. Granary handles this special case in a similar way to the kernel. Thus, maintaining return address transparency with respect to interrupts in instrumented module code has no value.

\vspace{-3pt}\paragraph{Module function wrappers} Granary does not maintain transparency when its type wrappers replace pointers to module functions with pointers to wrapped module functions. This is perhaps the riskiest break in transparency within Granary because a module might treat a function pointer as being representative of the module being in some particular state.\footnote{We have not yet encountered a module where changing pointers to module functions into module wrappers altered the module's behaviour under instrumentation.} If Granary is configured to use transparent module function pointer addresses in shared data structures then fast attaching is disabled and tools will not have access to the same level of static program information.

%\lstset{language=[x64]Assembler}
%\newsavebox\asmbox
%\begin{lrbox}{\asmbox}
%\begin{minipage}[b]{4cm}
%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
%1: mov    (%rdi), %rax
%   jmp    2
%2: mov    %rax, (%rsi)
%   test   %rax
%   jz     4
%3: inc    %rsi
%   inc    %rdi
%   jmp    1
%4: ret
%\end{lstlisting}
%\end{minipage}
%\end{lrbox}
%
%\begin{figure*}[ht!]
%\vspace{-2em}
%\centering
%\subfloat[Policy Switching Protocol]{
%\begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=3cm,scale=0.8, every node/.style={scale=0.8}]
%  \node[initial,state]          (C)   {$P_{\mathit{call\_entry}}$}; 
%  \node[state]          (A)[right of=C]      {$P_{\mathit{after\_entry}}$};
%  \path[->]     (A)     edge    [loop above]    node {{\footnotesize\ttfamily\bf jmp}, {\footnotesize\ttfamily\bf jz}} (A)                                edge         [bend right]        node {{\footnotesize\ttfamily\bf call}} (C)                        (C) edge    [bend right]     node {{\footnotesize\ttfamily\bf jmp}, {\footnotesize\ttfamily\bf jz}} (A)                        (C)     edge    [loop above]    node {{\footnotesize\ttfamily\bf call}} (C);
%\end{tikzpicture}
%}
%\hfill
% \subfloat[Uninstrumented Code]{\usebox\asmbox}
%\hfill
% \subfloat[Instrumented Basic Blocks]{
%\includegraphics[width=2.5in]{diagrams/policy-switching-entry-tracking.pdf}
% }
%\ORIGcaption{\label{fig:policy_switching}Example policy switching protocol that ensures that instrumentation ($P_{\mathit{call\_entry}}$) is performed only once on entry to every function, regardless of if that basic block is re-executed by a non-\texttt{call} instruction. Basic block (1) is instrumented by both $P_{\mathit{call\_entry}}$ and $P_{\mathit{after\_entry}}$, and basic blocks (2) and (3) are instrumented by $P_{\mathit{after\_entry}}$.}
%\end{figure*}

\section{Policy-Driven Instrumentation}\label{sec:policies}

A natural extension of mixed-mode execution is to give Granary tools the ability to dynamically switch the policy used to instrument code. Instrumenting only module code allows Granary to specialise the execution of native code. In the same way, policy switching allows tools to specialise the execution of instrumented code. In fact, optimising the performance of an early Granary tool was the original motivation for policy switching.

We developed a Granary tool that detects several Read-Copy-Update (RCU) API misuses in module code. Our tool focused on read-side critical sections (delimited by calls to \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock} in the code), however, most code executes outside of read-side critical sections. As an optimisation, we wanted our heavyweight API-checking instrumentation to apply only when the code was executing within a read-side critical section. Implementing this optimisation was challenging because each basic block was only translated once, and we had no way to know whether or not the heavyweight instrumentation should be applied to it. Knowing the state (within or outside a read-side critical section) at translation time was not sufficient, since the same translated block could later be executed in the opposite context. That is, code executing \emph{within} a read-side critical section may have originally been translated while executing \emph{outside} of a read-side critical section, and would thus omit the heavyweight API-checking instrumentation. This omission could cause our tool to miss bugs (i.e., it would not be comprehensive) when the basic block executes within a read-side critical section.

To implement this optimisation, Granary enables tailoring instrumentation to the context in which the instrumented code will execute. To do so, Granary allows different versions of the same module code to co-exist within Granary's code cache. That is, a different version of each of a module's basic blocks exists in Granary's code cache  for each encountered execution context. In the case of our RCU checker tool, there were two execution contexts: within and outside of a read-side critical section. If the same module basic block is  executed in the two contexts, then Granary's code cache would contain two different instrumented versions of that basic block. The way that Granary distinguishes between different execution contexts is with \emph{instrumentation policies}.

An instrumentation policy is both a name for an execution context, as well as a function that decides how to instrument basic blocks that will execute within that context. All tools define an initial policy that Granary uses to instrument module code. Tools are not limited to one policy though: any policy can declare a policy switch that will take effect when a selected control-transferring instruction (CTI) is executed. For example, our RCU tool invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.

Because policies name an execution context, they also represent states in a finite state machine. That is, a module's execution is in the state named by a policy if the code executing was instrumented by that policy. A state transition occurs when control transfers from code instrumented by one policy to code instrumented by another policy. A limitation with this approach is that a single state does not encode the sequence of previous states that led to execution being in the current state. For example, RCU permits nested read-side critical sections. If two read-side critical sections are nested then switching from $P_{\mathit{read\_critical}}$ to $P_{\mathit{null}}$ on the first \texttt{call} to \texttt{rcu\_read\_unlock} meant that our tool would lose track of being in the context of the outer read-side critical section. We solved this limitation by observing that, in most cases, function returns are natural policy reverting points. This is because the runtime call stack encodes both the context in which the current code is executing, as well as the continuation of the current executing code.

The effect of a tool specifying a policy switch on a CTI is that the basic block(s) targeted by that CTI will be instrumented according to the specified policy. CTIs with unspecified policies inherit their policies from their containing basic blocks. As hinted at above, \texttt{ret} instructions cannot explicitly switch policies, which allows a function instrumented by a different policy than its caller to return to the caller's context (i.e., policy). The asymmetry between \texttt{call}s and \texttt{ret}s is intentional: \texttt{call}s place contextual breadcrumbs (in the form of return addresses) on the runtime call stack, and \texttt{ret}s read these breadcrumbs to return to a previous context. Under this lens, the instruction pointer tracks our current policy, and return addresses form a stack of previous policies. Policy switches behave similarly to state transitions in a pushdown automaton; \texttt{call} instructions push a new state onto the stack for each function, \texttt{ret} instructions pop the current function's state from the stack, and other CTIs induce state transitions within the current function by altering control flow.

Granary implements policy tracking and switching by encoding policy information into the meta-data and CTIs of basic blocks. If a policy switch is not specified on an instrumented CTI then that CTI inherits the policy used to instrument the basic block containing the CTI. When an instrumented CTI executes for the first time, it yields control to Granary with the CTI target and policy information as inputs. Granary decodes and instruments the targeted instructions according to the input policy information. Because every non-\texttt{ret} CTI encodes policy information, and because Granary's translation mechanism depends only on CTI policy and target information, Granary is able to ensure that policy information is never lost or corrupted, even in the face of concurrent executions of the same module code, arbitrary pre-emption, and arbitrary resumption.

One example of a tool that uses policy switching is Granary's CFG tool, which dynamically constructs runtime inter- and intra-procedural control-flow graphs (CFGs). A single policy approach for call graph construction might add edges to the call graph only at every \texttt{call} site. However, this approach is insufficient in the face of mixed-mode execution. If an instrumented function $F$ is only invoked by a native function (which contains no instrumentation), and $F$ contains no function calls, then $F$ will not appear as a root node in the call graph. A two-policy solution easily solves the missing root problem by instrumenting both function call sites as well as function entrypoints. The CFG tool distinguishes between the first execution of a function's first basic block ($P_{\mathit{call\_entry}}$), and all other basic blocks executed within the function, including later executions of the first basic block ($P_{\mathit{after\_entry}}$). Function call sites are instrumented to set up the source node of an edge in the call graph. On entry to a function, we identify the current function as either a sink node connected to a source node (instrumented call), or as a root node if no source is present (native call). The distinction between entrypoints and non-entrypoints is achieved with minor bloat: the first basic block appears twice in Granary's code cache, but instrumented by two different policies, as shown in \Cref{fig:policy_switching}.

\section{Reifying Instrumentation}\label{sec:reify}

Granary uses a technique we call \emph{reifying instrumentation} to provide the benefits of high-level static analysis information to dynamic instrumentation tools. Reifying instrumentation bootstraps on Granary's mixed-mode execution approach, which exposes static type information to instrumentation tools. For example, we are actively using type information present in wrappers to create a tool that generates models of typical module behaviour. The wrapper information serves two roles in our tool. First, we use type wrappers to assign type-specific IDs to module-allocated memory that is shared across the module/kernel interface. Type ID assignment is critical because it allows us to match memory reads and writes to specific kernel data structure fields. Second, we use the kernel and module function wrappers to generate call graphs of a module's execution. This call graph models module behaviour (according to the kernel) because we label module code nodes with kernel data structure and function pointer field names (derived from module wrappers). This labelling allows us to generalise across similar modules. For example, code reached by calling the \texttt{ext4\_mount} and \texttt{btrfs\_mount} functions from the \texttt{ext4} and \texttt{btrfs} modules, respectively, are both labelled as \texttt{file\_system\_type::mount}, because the function addresses are stored (and later replaced by module wrappers) in the \texttt{mount} field of a \texttt{file\_system\_type} data structure. The combined records from multiple ``trusted'' modules of the same class (e.g., mature, open-source file systems) will model the behaviour of typical kernel modules. We hope that these models will help us build tools that classify and identify spurious module behaviour.

\section{Evaluation}\label{sec:eval}

We evaluated Granary by measuring the throughput of common file system operations performed by the \texttt{iozone} file system benchmark running on an \texttt{ext3} file system.
To focus on the CPU overhead of Granary and avoid the disk performance bottleneck, we mounted \texttt{ext3} on a 1 GB Ramdisk.  The benchmarks were run on an otherwise idle desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU with 4GB of RAM.  In our experiment, \texttt{iozone} mounts an \texttt{ext3} file system (which installs two kernel modules: \texttt{ext3} and \texttt{jbd}) and creates two processes (reader and writer) that read from and write to a 480MB file 1KB at a time.

\Cref{fig:perf} shows the decrease in throughput due to Granary (instrumenting only \texttt{ext3} and \texttt{jbd}), DynamoRIO Kernel (DRK; instrumenting the whole kernel and all modules), and ModifiedDRK (DRK modified to instrument only \texttt{ext3} and \texttt{jbd}, but using strict transparency) relative to native execution.  We run tests both using direct I/O, to bypass the OS buffer cache and force all file system requests to go through the \texttt{ext3}/\texttt{jbd} code, and with the OS buffer cache enabled.  Performance bars are overlaid, not stacked (e.g., buffered reads achieve nearly native throughput with both Granary and ModifiedDRK, so the native bar is covered).  Buffering highlights the benefits of mixed-mode execution: Granary and ModifiedDRK introduce almost no overhead for reads because most file system requests are served from the kernel's buffer cache. Direct requests highlight the impact of Granary's relaxed transparency model, which gives Granary lower overheads than ModifiedDRK.

%\begin{figure}[t!]
%\includegraphics[width=\columnwidth]{diagrams/stacked-performance.pdf}
%\caption{\label{fig:perf}Throughput in MB/sec of common file system operations, for Direct IO (\textbf{D}) where requests bypass the OS buffer cache and Buffered IO (\textbf{B}) where the OS buffer cache is enabled. }
%\end{figure}

\section{Related Work}\label{sec:related}
Prior work can be classified as either whole-kernel/system DBT, or probe-based kernel instrumentation. We discuss how Granary differs from these approaches.
\vspace{-3pt}\paragraph{Whole-kernel/system DBT} DRK \cite{DRK} is a kernel space port of the DynamoRIO \cite{DynamoRIO} DBT framework. DRK instruments the entire kernel, including modules, and follows a strict transparency model, which limits the flexibility of instrumentation and increases overheads.  PinOS \cite{PinOS} is a port of the Pin \cite{Pin} DBT framework that performs whole-system instrumentation. It has high overheads and depends on hardware virtualisation, which prevents it from instrumenting non-virtualisable kernel modules.

\vspace{-3pt}\paragraph{Probe-based Instrumentation} Several systems support injecting code at specific locations within kernel code. For example, KernInst \cite{KernInst} can inject code at almost any location in an unmodified commodity OS. Other examples with similar or more restricted functionality include LTTng \cite{LTTng}, KProbes \cite{KProbes}, DProbes \cite{DProbes}, and ftrace \cite{ftrace}. These systems are unable to perform fine-grained instrumentation of instructions or memory.

\section{Conclusion}\label{sec:conclusion}
We created Granary to address the challenges of instrumenting binary kernel modules.  Granary provides mixed-mode execution to remove the overheads of DBT for uninstrumented kernel code.  A relaxed transparency model further improves performance and allows greater visibility into the interactions between module and kernel code.  Granary also supports policy-driven instrumentation, which allows tools to specialise their instrumentation based on the context in which code executes.  Finally, Granary exposes high-level static analysis information to dynamic instrumentation tools, making it easy to match low-level memory reads with specific fields in data structures at the source code level. 
Together, Granary's features simplify the development of powerful kernel module analysis tools, while delivering lower overheads than previous kernel dynamic binary instrumentation solutions.

\bibliographystyle{acm}
\bibliography{library}

\end{document}

### 2013-10-27 VEE-14/paper.tex
Minor modifications to start things off.

@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{Granary: A Sane Framework for Instrumenting an Insane Environment}
+\title{Granary weevil eats binary Linux kernels for breakfast}
 
 %for single author (just remove % characters)
-\authorinfo{Peter Goodman \and Akshay Kumar \and Angela Demke Brown \and Ashvin Goel}
+\authorinfo{Peter Goodman \and Angela Demke Brown \and Ashvin Goel}
 {University of Toronto}{}
@@
 \subsection*{Abstract}
-Kernel modules extend the functionality of operating systems.
+%Kernel modules extend the functionality of operating systems.
 They represent the bulk of new kernel code in development and contain the majority of new kernel bugs.
@@
 
-We created Granary to address the challenges of module analysis.
+%We created Granary to address the challenges of module analysis.
 Granary is a framework for creating flexible and efficient tools that analyse or debug arbitrary, binary Linux kernel modules.
@@
 
-Operating system (OS) kernels present an apparently insane environment for a transparent binary instrumentation framework.
+%Operating system (OS) kernels present an apparently insane environment for a transparent binary instrumentation framework.
 For example, hardware interrupts can redirect control flow at any instruction, and interrupt delivery cannot be delayed past the execution of memory instructions that might alter the interrupt delivery state~\cite{DRK}.
@@
 
-The challenges and opportunities of the kernel environment motivated the creation of Granary.
+%The challenges and opportunities of the kernel environment motivated the creation of Granary.
 Granary is a new binary instrumentation framework for the Linux kernel environment, designed specifically for analysing kernel modules.
@@
 Granary has two main goals: \begin{inparaenum}[i)]
-	\item make it easy to create efficient, low-level analysis tools; and
-	\item provide high-level static analysis information to the analysis tools.
-\end{inparaenum} Granary achieves the first goal by using dynamic binary translation (DBT) to provide low-level access to instructions, memory, and interrupts.
+%	\item make it easy to create efficient, low-level analysis tools; and
+%	\item provide high-level static analysis information to the analysis tools.
+%\end{inparaenum} Granary achieves the first goal by using dynamic binary translation (DBT) to provide low-level access to instructions, memory, and interrupts.
 Granary achieves the second goal by giving tools direct access to the results of its own static analysis of the Linux kernel source code.
@@
 Granary addresses three limitations of existing program analysis systems: \begin{inparaenum}[i)]
-	\item they have focused on instrumenting all code (the whole kernel: DRK; or whole system: QEMU, PinOS);
-	\item they do not allow for flexible runtime code specialisation (Pin, DynamoRIO); and
-	\item they do not provide static type information to the instrumented code (Pin, DynamoRIO).
+	\item they have focused on instrumenting all code (the whole kernel: DRK, btkernel; or whole system: QEMU, PinOS);
+	\item they do not allow for flexible runtime code specialisation; and
+	\item they do not provide static type information to the instrumented code.
 \end{inparaenum}
 
-Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analysed (e.g., modules).
-In practice, we would like to target instrumentation at specific modules for efficiency, while also allowing rich and detailed instrumentation of these modules and any other relevant code.
-This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary to instrument module code while non-module kernel code executes with zero overhead.
-This is challenging because Granary must be comprehensive: all code that is meant to be instrumented should be instrumented.
+Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analysed (e.g., modules, functions).
+In practice, we would like to target instrumentation at specific code for efficiency, while also allowing rich and detailed instrumentation of the targeted code.
+This requirement motivates the design of \emph{mixed-mode execution} and \emph{probe-based instrumentation}, which allows Granary to instrument specific module code or functions while the majority of kernel code executes with zero overhead.
+
+%This is challenging because Granary must be comprehensive: all code that is meant to be instrumented should be instrumented.
 In other words, relinquishing control to execute kernel code natively must not cause Granary to miss the execution of any module code.
@@
 they are unable to change how code is instrumented based on the context in which that code will run.
-For example, Granary has a tool that  dynamically constructs inter- and intra-procedural control flow graphs (CFGs).
+
+%For example, Granary has a tool that  dynamically constructs inter- and intra-procedural control flow graphs (CFGs).
 The outputs of this tool enable Granary to do better runtime register allocation, which improves tool efficiency.
@@
 In Granary, code is specialised according to the context in which that code will execute.
-In the previous example, executing code is classified into two contexts: entrypoint and non-entrypoint basic blocks.
 Granary tracks execution contexts using a technique we call \emph{policy-driven instrumentation}.
@@
 An instrumentation policy decides how to instrument code that will execute within the associated context.
+Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks.
+This avoids typical problems with tracking context in just-in-time DBT systems (e.g., concurrency, arbitrary pre-emption, context-switch\-ing, inability to predict later execution contexts at translation time).
+
+% In the previous example, executing code is classified into two contexts: entrypoint and non-entrypoint basic blocks.
 In the case of our CFG tool, two policies track the two execution contexts: $P_{\mathit{call\_entry}}$ and $P_{\mathit{after\_entry}}$.
@@
 $P_{\mathit{after\_entry}}$ connects basic blocks in the intra-procedural CFG.
-Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks.
-This avoids typical problems with tracking context in just-in-time DBT systems (e.g., concurrency, arbitrary pre-emption, context-switch\-ing, inability to predict later execution contexts at translation time).
 
@@
 
-Granary uses dynamic binary translation (DBT) to instrument Linux kernel modules.
+Granary uses dynamic binary translation (DBT) to instrument the Linux kernel.
 DBT is used for instruction set translation/emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimisation}, and runtime instrumentation (analysis \cite{DynamoRIO}, security \cite{Vx32,NaCl}, and debugging \cite{Valgrind}).
@@
 	\item static binary analysis tools are unable to cope with \texttt{x86}'s mix of variably-sized instructions and data; 
-	\item some modules are only distributed in a binary format; and
+	\item some kernel code is only distributed in a binary format (e.g. proprietary device drivers); and
 	\item virtualisation-based approaches are unable to analyse non-virtualisable device drivers \cite{DRK}.
@@
 
+Granary instruments code according to one of three strategies: \begin{inparaenum}[i)]
+	\item full-kernel instrumentation;
+	\item module-only instrumentation, and;
+	\item probe-based instrumentation.
+\end{inparaenum}
+
+\paragraph{Full-kernel instrumentation}
+Granary interposes on the system call and interrupt entrypoints, located in the x86-64 \texttt{MSR\_LSTAR} model-specific register and interrupt descriptor table (IDT), respectively.
+Granary boostraps by duplicating the early entrypoint instructions up until the kernel's native entrypoint switches from a user-space to a kernel-space stack.
+Granary translates the first basic block immediately following the stack switch, and adds a control-flow transfer instruction from the duplicated entrypoint instructions into the translated basic block.
+The translation process continues when a system call or interrupt occurs.
+Our choice to defer translation to a point where execution operates off of a kernel stack simplifies instrumentation tools that want to save and restore transient data on the runtime call stack.
+
+\paragraph{Module-only instrumentation}
 Granary interposes on the Linux kernel's module loading process.
@@
 
-Granary translates and instruments module binaries one basic block at a time.
+\paragraph{Probe-based instrumentation}
+Granary can interpose on the execution of specific Linux kernel functions in three ways:
+\begin{enumerate}
+	\item {\bf Wrapping:} Granary can substitute the execution of a kernel function when that function is invoked from instrumented code.
+A different function can be substituted depending on the context from which the original function is invoked (instrumented kernel code, instrumented module code).
+We call this type of interposition ``wrapping'' because it is commonly used to inspect the substituted function's arguments and return value, while still invoking the original function.
+	\item {\bf Substituting:}
+	\item {\bf Instrumented Wrapping/Substituting:}
+\end{enumerate}
+
+Granary translates and instruments kernel binaries one basic block at a time.
 That is, new basic blocks are decoded and translated as execution ``discovers'' those basic blocks.
@@
 
-Basic blocks in Granary's code cache contain \texttt{x86-64} binary instructions and meta-data describing those instructions.
+Basic blocks in Granary's code cache contain x86-64 binary instructions and meta-data describing those instructions.
 The meta-data that Granary records in each basic block includes: \begin{inparaenum}[i)]

### 2013-11-10 VEE-14/paper.tex
Working on getting the message right, and cutting out PLOS stuff that doesn't quite fit.

@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{Granary weevil eats binary Linux kernels for breakfast}
+\title{Flexible Runtime Code Specialization for the Kernel}
 
 %for single author (just remove % characters)
-\authorinfo{Peter Goodman \and Angela Demke Brown \and Ashvin Goel}
+\authorinfo{Peter Goodman \and Akshay Kumar \and Angela Demke Brown \and Ashvin Goel}
 {University of Toronto}{}
@@
 They represent the bulk of new kernel code in development and contain the majority of new kernel bugs.
-Unfortunately, analysing and debugging modules is very challenging.
+Unfortunately, analyzing and debugging modules is very challenging.
 Static analysis of module source code is difficult because of the tight interaction between modules and the kernel.
@@
 Granary uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules.
-Granary makes it easy to create flexible tools by supporting context-aware runtime code specialisation, and by integrating high-level static analysis information with low-level instruction manipulation.
+Granary makes it easy to create flexible tools by supporting context-aware runtime code specialization, and by integrating high-level static analysis information with low-level instruction manipulation.
 Module analysis tools built on Granary are efficient because of Granary's adoption of a relaxed transparency model, and because of Granary's ability to instrument module code without imposing overhead on non-module kernel code.
@@
 \section{Introduction}\label{sec:intro}
+
+Granary is a framework for creating flexible and efficient tools that instrument the Linux kernel.
+Granary's key novelty is its \emph{runtime code specialization} feature, which enables tools to decide what code to instrument, and how to instrument that code.
+Flexible specialization allows tools to be picky about instrumentation.
+For example, a tool can choose to instrument every load and store operation, but only inside nested critical sections that are executing in module code.
+Granary's runtime code specialization feature addresses two limitations of existing kernel analysis systems: \begin{inparaenum}[i)]
+	\item they have focused on instrumenting all code \cite{DRK,btkernel,QEMU}; and
+	\item they can only instrument code using a single policy.
+\end{inparaenum}
 
@@
 %The challenges and opportunities of the kernel environment motivated the creation of Granary.
-Granary is a new binary instrumentation framework for the Linux kernel environment, designed specifically for analysing kernel modules.
+Granary is a new binary instrumentation framework for the Linux kernel environment, designed specifically for analyzing kernel modules.
 Granary targets modules because they represent the bulk of new bugs and new code in development in the Linux kernel \cite{FaultsInLinux}.
@@
 
-Granary addresses three limitations of existing program analysis systems: \begin{inparaenum}[i)]
-	\item they have focused on instrumenting all code (the whole kernel: DRK, btkernel; or whole system: QEMU, PinOS);
-	\item they do not allow for flexible runtime code specialisation; and
-	\item they do not provide static type information to the instrumented code.
-\end{inparaenum}
+%Granary addresses three limitations of existing program analysis systems: \begin{inparaenum}[i)]
+%	\item they have focused on instrumenting all code \cite{DRK,btkernel,QEMU};
+%	\item they do not allow for flexible runtime code specialization; and
+%	\item they do not provide static type information to the instrumented code.
+%\end{inparaenum}
 
-Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analysed (e.g., modules, functions).
+Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analyzed (e.g., modules, functions).
 In practice, we would like to target instrumentation at specific code for efficiency, while also allowing rich and detailed instrumentation of the targeted code.
-This requirement motivates the design of \emph{mixed-mode execution} and \emph{probe-based instrumentation}, which allows Granary to instrument specific module code or functions while the majority of kernel code executes with zero overhead.
+This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary tools to instrument only specific code while the majority of code executes with zero overhead.
 
@@
 
-Another limitation of existing systems is that they are unable to perform efficient runtime code specialisation, i.e.
+Another limitation of existing systems is that they are unable to perform efficient runtime code specialization, i.e.
 they are unable to change how code is instrumented based on the context in which that code will run.
+For example, Granary's Read-Copy-Update (RCU) debugging tool, \texttt{RCUdbg}, specializes instrumentation at different stages of use of the RCU synchronization API.
+Correct usage of the RCU API allows programmers to create scalable, shared data structures that permit concurrent access and modification.
+The ``trick" of RCU is that readers might not see the latest write to the data structure, and writers must synchronize to modify the data structure.
+\texttt{RCUdbg} distinguishes between code executing inside \emph{read-side} and \emph{write-side} critical sections, and instruments each code context differently.
+In read-side critical sections, \texttt{RCUdbg} detects bugs when readers access potentially non-existent data, do not the correct accessor API, or attempt to write to shared data.
+In write-side critical sections, \texttt{RCUdbg} detects when writers write to shared data without using the correct API.
+
+%is a synchronization mechanism that permits readers to operate concurrently with writers, but does not guarantee that readers observe the most up-to-date changes (by writers) to shared state.
+RCU readers
+
+%When used correctly, RCU readers must declare that they will operate on shared state by entering into a read-side critical section, delimited by \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock}, respectively.
+When inside a read-side critical section, a pointer to shared state \emph{must} be obtained by invoking \texttt{rcu\_dereference}.
+Thereafter, a reader can freely access the shared state through this pointer, but only up until the
+
+%In reader-side code
 
@@
 This is challenging with existing DBT systems because some optimised functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs.
-Granary solves this problem using demand-based runtime code specialisation: every entrypoint basic block is instrumented in one way, and later re-executions of any entrypoint basic blocks are specialised and instrumented in another way.
+Granary solves this problem using demand-based runtime code specialization: every entrypoint basic block is instrumented in one way, and later re-executions of any entrypoint basic blocks are specialized and instrumented in another way.
 
-In Granary, code is specialised according to the context in which that code will execute.
-Granary tracks execution contexts using a technique we call \emph{policy-driven instrumentation}.
+The \texttt{RCUdbg} tool is an example of \emph{context-driven code specialization}: code executing inside of a read-side critical section is instrumented differently from code executing inside of a write-side critical section, and code execution outside of both contexts executes with minimal instrumentation.
 In a Granary tool, each tracked execution context is associated with its own instrumentation policy.
@@
 Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks.
-This avoids typical problems with tracking context in just-in-time DBT systems (e.g., concurrency, arbitrary pre-emption, context-switch\-ing, inability to predict later execution contexts at translation time).
+This avoids typical poblems with tracking context in just-in-time DBT systems (e.g., maintaining runtime context-tracking state, re-entrancy, and the inability to predict later execution contexts at translation time).
 
@@
 
-Finally, an important missing feature of existing low-level instrumentation systems is the ``big picture'' view that one gets from static analysis information.
+%Finally, an important missing feature of existing low-level instrumentation systems is the ``big picture'' view that one gets from static analysis information.
 Similarly, static analysis tools lack access to instructions, memory, and processor-specific behaviours that are only available to low-level runtime instrumentation tools.
@@
 
-For example, we are actively developing a tool that uses static type information to learn about all the fields of all kernel data structures that are read from or written to by modules.
+%For example, we are actively developing a tool that uses static type information to learn about all the fields of all kernel data structures that are read from or written to by modules.
 This information allows us to model things like how a typical file system operates on kernel data structures in the process of opening a file.
@@
 Granary uses dynamic binary translation (DBT) to instrument the Linux kernel.
-DBT is used for instruction set translation/emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimisation}, and runtime instrumentation (analysis \cite{DynamoRIO}, security \cite{Vx32,NaCl}, and debugging \cite{Valgrind}).
+DBT is used for emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimisation}, and runtime instrumentation (analysis \cite{DynamoRIO, DRK, btkernel, ProfilingSimics}, security \cite{Vx32,NaCl,ProgramShepherding}, and debugging \cite{Valgrind}).
 We chose DBT because \begin{inparaenum}[i)]
@@
 	\item some kernel code is only distributed in a binary format (e.g. proprietary device drivers); and
-	\item virtualisation-based approaches are unable to analyse non-virtualisable device drivers \cite{DRK}.
+	\item virtualization-based approaches are unable to analyse non-virtualizable device drivers \cite{DRK}.
 \end{inparaenum}
 
-Granary instruments code according to one of three strategies: \begin{inparaenum}[i)]
-	\item full-kernel instrumentation;
-	\item module-only instrumentation, and;
-	\item probe-based instrumentation.
-\end{inparaenum}
-
-\paragraph{Full-kernel instrumentation}
-Granary interposes on the system call and interrupt entrypoints, located in the x86-64 \texttt{MSR\_LSTAR} model-specific register and interrupt descriptor table (IDT), respectively.
-Granary boostraps by duplicating the early entrypoint instructions up until the kernel's native entrypoint switches from a user-space to a kernel-space stack.
-Granary translates the first basic block immediately following the stack switch, and adds a control-flow transfer instruction from the duplicated entrypoint instructions into the translated basic block.
-The translation process continues when a system call or interrupt occurs.
-Our choice to defer translation to a point where execution operates off of a kernel stack simplifies instrumentation tools that want to save and restore transient data on the runtime call stack.
-
-\paragraph{Module-only instrumentation}
-Granary interposes on the Linux kernel's module loading process.
+%Granary translates and instruments code at one of three default granularities: kernel, module, and function (also called probe-based instrumentation). 
+%
+%At runtime, the granularity of instrumentation can be altered based on the execution context.
+For example, a tool analyzing modules for bugs can choose not to instrument a particular function (e.g.
+whitelist the function) if that function is known to not contain bugs, or known to generate many false-positives.
+%
+%\paragraph{Kernel}
+%Granary interposes on the system call and interrupt entrypoints, located in the x86-64 \texttt{MSR\_LSTAR} model-specific register and interrupt descriptor table (IDT), respectively.
+Granary boostraps by duplicating the early entrypoint instructions that are responsible for switching execution onto a kernel stack, and then injects a jump to translated kernel code after the stack switch.
+The translation process continues when a system call or interrupt occurs and follows the jump from duplicated into translated code.
+We only translate code that is known to operate on a kernel stack so that instrumentation tools can safely save transient data in a re-entrant way by spilling that data to the stack.
+%
+%\paragraph{Module}
+%Granary interposes on the Linux kernel's module loading process.
 When a kernel module is loaded, Granary bootstraps by translating the first basic block of the module's initialisation function.
@@
 The translation process continues when the kernel initialises the module by invoking the translated module code.
-
-\paragraph{Probe-based instrumentation}
-Granary can interpose on the execution of specific Linux kernel functions in three ways:
-\begin{enumerate}
-	\item {\bf Wrapping:} Granary can substitute the execution of a kernel function when that function is invoked from instrumented code.
+%
+%\paragraph{Function}
+%
+%Granary can interpose on the execution of specific kernel functions in three ways:
+%\begin{enumerate}
+%	\item {\bf Wrapping:} Granary can substitute the execution of a kernel function when that function is invoked from instrumented code.
 A different function can be substituted depending on the context from which the original function is invoked (instrumented kernel code, instrumented module code).
 We call this type of interposition ``wrapping'' because it is commonly used to inspect the substituted function's arguments and return value, while still invoking the original function.
-	\item {\bf Substituting:}
-	\item {\bf Instrumented Wrapping/Substituting:}
-\end{enumerate}
+%	\item {\bf Substituting:} Granary can substitutes the execution of a kernel function, even when that function is executed by native kernel code.
+This is useful when tools only instrument part of the kernel (e.g.
+modules or functions) but require visibility on specific kernel events happening outside of the instrumented code (e.g.
+particular memory allocations).
+%	\item {\bf Instrumented Wrapping/Substituting:} Granary can optionally instrument wrapped or substituted code.
+This is useful when tools want to observe function arguments/return values using high-level C/C++ code, but where that function must be instrumented for visibility.
+%\end{enumerate}
 
 Granary translates and instruments kernel binaries one basic block at a time.
-That is, new basic blocks are decoded and translated as execution ``discovers'' those basic blocks.
+In Granary, a basic block is a sequence of instructions ending in a conditional branch, \texttt{ret}, or \texttt{jmp}, but not a \texttt{call} instruction.
+Entrypoint basic blocks (e.g.
+interrupt vectors, system call entrypoints, module initializers) are translated ahead-of-time, and all other basic blocks are translated just-in-time as execution ``discovers'' them.
 Translated basic blocks are linked together and stored in a globally accessible \emph{code cache}.
-In Granary, a basic block is a sequence of instructions ending in a conditional branch, \texttt{ret}, or \texttt{jmp}, but not a \texttt{call} instruction.
 
@@
 Granary context-switches back to the code cache when the next basic block has been found or translated so that instrumented execution may continue.
-Similar to other DBT systems, Granary uses caching and hotpatching to reduce the number of context switches.
-In our experience, modules stabilise very quickly: context switches stop happening after most module basic blocks have been translated.
+Similar to other DBT systems, Granary uses caching and hot code patching to reduce the number of context switches.
+In our experience, kernel code stabilises very quickly: context switches stop happening after the first few seconds of translation.
 
@@
 The meta-data is also queried by interrupt handlers when deciding how to handle interrupts in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block.
-
-%\begin{figure*}[t!]
-%\lstset{language=C, tabsize=2, stepnumber=1}
-%\begin{multicols}{2}
-%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
-%struct device_driver {
-%	...
-%	int (*probe)(struct device *);
-%	int (*remove)(struct device *);
-%	void (*shutdown)(struct device *);
-%	int (*suspend)(struct device *, pm_message_t);
-%	int (*resume)(struct device *);
-%	...
-%	const struct dev_pm_ops *pm;
-%	...
-%};
-%\end{lstlisting}
-%\columnbreak
-%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
-%TYPE_WRAPPER(struct device_driver, {
-%    PRE_OUT {
-%        ABORT_IF_FUNCTION_IS_WRAPPED(arg.probe)
-%        WRAP_FUNCTION(arg.probe);
-%        WRAP_FUNCTION(arg.remove);
-%        ...
-%    }
-%    POST_OUT {
-%        POST_WRAP(arg.pm);
-%    }
-%})
-%\end{lstlisting}
-%\end{multicols}
-%\vspace{-1em}
-%\caption{The Linux device driver structure is shown on the left.
-The automatically generated type wrapper for this structure is shown on the right.
-In the wrapper code, \texttt{arg} is a reference to a \texttt{struct device\_driver} object passed as, or referenced by, an argument to a kernel or module wrapper.
-Code in the \texttt{PRE\_OUT} section is applied to arguments of the wrapped type before a kernel wrapper is invoked.
-Similarly, code in the \texttt{POST\_OUT} section is applied to arguments of the wrapped type after a kernel wrapper is invoked.
-\texttt{POST\_WRAP} invokes the type wrapper that is specific to the value to which it is applied (\texttt{arg.pm}).
-Type wrappers also support \texttt{\_IN} suffixes instead of \texttt{\_OUT} suffixes, which apply to data going into modules (i.e., over module wrappers).
-Finally, the \texttt{RETURN\_} prefix is used to apply some code to return values of either kernel or module wrappers.}
-%\label{fig:type_wrapper}
-%\end{figure*}
 
@@
 
-%\lstset{language=[x64]Assembler}
-%\newsavebox\asmbox
-%\begin{lrbox}{\asmbox}
-%\begin{minipage}[b]{4cm}
-%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
-%1: mov    (%rdi), %rax
-%   jmp    2
-%2: mov    %rax, (%rsi)
-%   test   %rax
-%   jz     4
-%3: inc    %rsi
-%   inc    %rdi
-%   jmp    1
-%4: ret
-%\end{lstlisting}
-%\end{minipage}
-%\end{lrbox}
-%
-%\begin{figure*}[ht!]
-%\vspace{-2em}
-%\centering
-%\subfloat[Policy Switching Protocol]{
-%\begin{tikzpicture}[>=stealth',shorten >=1pt,auto,node distance=3cm,scale=0.8, every node/.style={scale=0.8}]
-%  \node[initial,state]          (C)   {$P_{\mathit{call\_entry}}$}; 
-%  \node[state]          (A)[right of=C]      {$P_{\mathit{after\_entry}}$};
-%  \path[->]     (A)     edge    [loop above]    node {{\footnotesize\ttfamily\bf jmp}, {\footnotesize\ttfamily\bf jz}} (A)                                edge         [bend right]        node {{\footnotesize\ttfamily\bf call}} (C)                        (C) edge    [bend right]     node {{\footnotesize\ttfamily\bf jmp}, {\footnotesize\ttfamily\bf jz}} (A)                        (C)     edge    [loop above]    node {{\footnotesize\ttfamily\bf call}} (C);
-%\end{tikzpicture}
-%}
-%\hfill
-% \subfloat[Uninstrumented Code]{\usebox\asmbox}
-%\hfill
-% \subfloat[Instrumented Basic Blocks]{
-%\includegraphics[width=2.5in]{diagrams/policy-switching-entry-tracking.pdf}
-% }
-%\ORIGcaption{\label{fig:policy_switching}Example policy switching protocol that ensures that instrumentation ($P_{\mathit{call\_entry}}$) is performed only once on entry to every function, regardless of if that basic block is re-executed by a non-\texttt{call} instruction.
-Basic block (1) is instrumented by both $P_{\mathit{call\_entry}}$ and $P_{\mathit{after\_entry}}$, and basic blocks (2) and (3) are instrumented by $P_{\mathit{after\_entry}}$.}
-%\end{figure*}
-
 \section{Policy-Driven Instrumentation}\label{sec:policies}
@@
 A natural extension of mixed-mode execution is to give Granary tools the ability to dynamically switch the policy used to instrument code.
-Instrumenting only module code allows Granary to specialise the execution of native code.
-In the same way, policy switching allows tools to specialise the execution of instrumented code.
+Instrumenting only module code allows Granary to specialize the execution of native code.
+In the same way, policy switching allows tools to specialize the execution of instrumented code.
 In fact, optimising the performance of an early Granary tool was the original motivation for policy switching.
@@
 
-One example of a tool that uses policy switching is Granary's CFG tool, which dynamically constructs runtime inter- and intra-procedural control-flow graphs (CFGs).
-A single policy approach for call graph construction might add edges to the call graph only at every \texttt{call} site.
-However, this approach is insufficient in the face of mixed-mode execution.
-If an instrumented function $F$ is only invoked by a native function (which contains no instrumentation), and $F$ contains no function calls, then $F$ will not appear as a root node in the call graph.
-A two-policy solution easily solves the missing root problem by instrumenting both function call sites as well as function entrypoints.
-The CFG tool distinguishes between the first execution of a function's first basic block ($P_{\mathit{call\_entry}}$), and all other basic blocks executed within the function, including later executions of the first basic block ($P_{\mathit{after\_entry}}$).
-Function call sites are instrumented to set up the source node of an edge in the call graph.
-On entry to a function, we identify the current function as either a sink node connected to a source node (instrumented call), or as a root node if no source is present (native call).
-The distinction between entrypoints and non-entrypoints is achieved with minor bloat: the first basic block appears twice in Granary's code cache, but instrumented by two different policies, as shown in \Cref{fig:policy_switching}.
-
-\section{Reifying Instrumentation}\label{sec:reify}
-
-Granary uses a technique we call \emph{reifying instrumentation} to provide the benefits of high-level static analysis information to dynamic instrumentation tools.
+%\section{Reifying Instrumentation}\label{sec:reify}
+%
+%Granary uses a technique we call \emph{reifying instrumentation} to provide the benefits of high-level static analysis information to dynamic instrumentation tools.
 Reifying instrumentation bootstraps on Granary's mixed-mode execution approach, which exposes static type information to instrumentation tools.
@@
 
-We evaluated Granary by measuring the throughput of common file system operations performed by the \texttt{iozone} file system benchmark running on an \texttt{ext3} file system.
-To focus on the CPU overhead of Granary and avoid the disk performance bottleneck, we mounted \texttt{ext3} on a 1 GB Ramdisk.
-The benchmarks were run on an otherwise idle desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU with 4GB of RAM.
-In our experiment, \texttt{iozone} mounts an \texttt{ext3} file system (which installs two kernel modules: \texttt{ext3} and \texttt{jbd}) and creates two processes (reader and writer) that read from and write to a 480MB file 1KB at a time.
-
-\Cref{fig:perf} shows the decrease in throughput due to Granary (instrumenting only \texttt{ext3} and \texttt{jbd}), DynamoRIO Kernel (DRK; instrumenting the whole kernel and all modules), and ModifiedDRK (DRK modified to instrument only \texttt{ext3} and \texttt{jbd}, but using strict transparency) relative to native execution.
-We run tests both using direct I/O, to bypass the OS buffer cache and force all file system requests to go through the \texttt{ext3}/\texttt{jbd} code, and with the OS buffer cache enabled.
-Performance bars are overlaid, not stacked (e.g., buffered reads achieve nearly native throughput with both Granary and ModifiedDRK, so the native bar is covered).
-Buffering highlights the benefits of mixed-mode execution: Granary and ModifiedDRK introduce almost no overhead for reads because most file system requests are served from the kernel's buffer cache.
-Direct requests highlight the impact of Granary's relaxed transparency model, which gives Granary lower overheads than ModifiedDRK.
-
-%\begin{figure}[t!]
-%\includegraphics[width=\columnwidth]{diagrams/stacked-performance.pdf}
-%\caption{\label{fig:perf}Throughput in MB/sec of common file system operations, for Direct IO (\textbf{D}) where requests bypass the OS buffer cache and Buffered IO (\textbf{B}) where the OS buffer cache is enabled.
-}
-%\end{figure}
-
 \section{Related Work}\label{sec:related}
@@
 PinOS \cite{PinOS} is a port of the Pin \cite{Pin} DBT framework that performs whole-system instrumentation.
-It has high overheads and depends on hardware virtualisation, which prevents it from instrumenting non-virtualisable kernel modules.
+It has high overheads and depends on hardware virtualization, which prevents it from instrumenting non-virtualizable kernel modules.
 
@@
 A relaxed transparency model further improves performance and allows greater visibility into the interactions between module and kernel code.
-Granary also supports policy-driven instrumentation, which allows tools to specialise their instrumentation based on the context in which code executes.
+Granary also supports policy-driven instrumentation, which allows tools to specialize their instrumentation based on the context in which code executes.
 Finally, Granary exposes high-level static analysis information to dynamic instrumentation tools, making it easy to match low-level memory reads with specific fields in data structures at the source code level.

### 2013-11-10 VEE-14/paper.tex
Working on the DBT section

@@
 
-The rest of the paper describes Granary in more detail.
-\Cref{sec:modes} describes how to efficiently implement mixed-mode execution.
+The rest of the paper describes Granary in more detail. \Cref{sec:dbt} describes how Granary is implemented.
+
+% \Cref{sec:modes} describes how to efficiently implement mixed-mode execution.
 \Cref{sec:policies} describes a new method for dynamically changing how code is instrumented.
@@
 Entrypoint basic blocks (e.g.
-interrupt vectors, system call entrypoints, module initializers) are translated ahead-of-time, and all other basic blocks are translated just-in-time as execution ``discovers'' them.
+interrupt vectors, system call entrypoints, module initializers) are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as execution ``discovers'' them.
 Translated basic blocks are linked together and stored in a globally accessible \emph{code cache}.
@@
 
-Basic blocks in Granary's code cache contain x86-64 binary instructions and meta-data describing those instructions.
-The meta-data that Granary records in each basic block includes: \begin{inparaenum}[i)]
+Basic blocks in Granary's code cache contain x86-64 binary instructions.
+Associated with each basic block is meta-data describing the instructions of the basic block, and the instrumentation policy used to instrument those instructions.
+Basic block meta-data includes: \begin{inparaenum}[i)]
 	\item the length in bytes of the original and translated basic blocks;
 	\item the address of the first instruction in the original basic block; and
-	\item the policy information (e.g., $P_{\mathit{call\_entry}}$, $P_{\mathit{after\_entry}}$) used to instrument the basic block.
-\end{inparaenum} This meta-data can be queried and extended by Granary instrumentation tools.
-For example, Granary's CFG-building tool represents each node in the inter- and intra-procedural CFGs as an extension of the meta-data stored for each basic block.
-The meta-data is also queried by interrupt handlers when deciding how to handle interrupts in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block.
+	\item the instrumentation policy.
+\end{inparaenum} Basic block meta-data is queried and extended by Granary instrumentation tools.
+For example, Granary's control-flow graph build tool, \texttt{cfg}, extends the basic block meta-data to track per-block execution counts.
+Block meta-data is also queried by interrupt handlers when deciding how to handle interrupts and exceptions in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block.
+
+\subsection{Transparency}\label{sec:transparency}
+
+An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}.
+Granary includes configurable levels of transparency to address the trade-off between transparency and overhead.
+By default, Granary uses a relaxed transparency model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility.
+By default, Granary exposes the following two artifacts: code cache addresses as function and interrupt return addresses.
+
+\paragraph{Function return addresses}\label{para:return_address_transparency} Unlike most DBT frameworks, Granary inlines \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses.
+Inlining \texttt{call} instructions avoids return-address mispredictions and unnecessary control-flow by building longer basic blocks.
+
+\paragraph{Interrupt return addresses} Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts.
+A special case arises for page fault exceptions that occur within specific kernel functions that access user space data \cite{btkernel}.
+The Linux kernel records ranges of code addresses that are permitted to fault in an ``exception table'' data structure, and on a page fault exception, checks if the interrupt return address belongs to one of the pre-defined ranges.
+Granary records any exception table entries as part of each basic block's meta-data, so that a page fault in a basic block can be mapped to the correct native faulting program counter.
+
+%\paragraph{Module function wrappers} Granary does not maintain transparency when its type wrappers replace pointers to module functions with pointers to wrapped module functions.
+This is perhaps the riskiest break in transparency within Granary because a module might treat a function pointer as being representative of the module being in some particular state.\footnote{We have not yet encountered a module where changing pointers to module functions into module wrappers altered the module's behaviour under instrumentation.} If Granary is configured to use transparent module function pointer addresses in shared data structures then fast attaching is disabled and tools will not have access to the same level of static program information.
+
+\subsection{Re-entrancy}
+
+\paragraph{Code cache}
+Granary's code cache and JIT translation mechanism is fully re-entrant.
+Context-switches into Granary disable interrupts and switch execution onto a CPU-private stack.
+Context switches out of Granary and into the code cache or native code returns executuin to the native stack and re-enables interrupts.
+JIT translation, however, presents a problem: context-switches save or restore the entire register state, so as not to disturb the execution of the instrumented program.
+This raises the question: when resolving the target of a direct or indirect control-flow instruction, how should execution be directed to the correct location?
+
+Granary's solution to this problem is two-fold. Like \texttt{btkernel}, Granary generates ``edge'' code for each direct control-flow instruction (e.g. \texttt{call}, \texttt{jmp}, \texttt{jcc}).
+
+\paragraph{Instrumentation}
 
@@
 
-\subsection{Transparency}\label{sec:transparency}
-
-An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}.
-Granary includes configurable levels of transparency to address the trade-off between transparency and overhead.
-By default, Granary uses a relaxed transparency model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility.
-By default, Granary exposes the following three artifacts: code cache addresses as function and interrupt return addresses, and wrapped module function addresses in shared data structures.
-
-\vspace{-3pt}\paragraph{Function return addresses}\label{para:return_address_transparency} Unlike most DBT frameworks, Granary defaults to inlining \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses.
-Inlining \texttt{call} instructions avoids return-address mispredictions and unnecessary control-flow by building longer basic blocks.
-If Granary is configured to use transparent return addresses, then implicit attaching and detaching are disabled because they depend directly on returning to instrumented module or native kernel code.
-Granary will fall back to slow attaching to handle kernel-to-module returns, and indirect branch lookups for module-to-module returns.
-
-\vspace{-3pt}\paragraph{Interrupt return addresses} Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts.
-A special case arises for page fault exceptions that occur within specific kernel functions that access user space data.
-Granary handles this special case in a similar way to the kernel.
-Thus, maintaining return address transparency with respect to interrupts in instrumented module code has no value.
-
-\vspace{-3pt}\paragraph{Module function wrappers} Granary does not maintain transparency when its type wrappers replace pointers to module functions with pointers to wrapped module functions.
-This is perhaps the riskiest break in transparency within Granary because a module might treat a function pointer as being representative of the module being in some particular state.\footnote{We have not yet encountered a module where changing pointers to module functions into module wrappers altered the module's behaviour under instrumentation.} If Granary is configured to use transparent module function pointer addresses in shared data structures then fast attaching is disabled and tools will not have access to the same level of static program information.
-
-\section{Policy-Driven Instrumentation}\label{sec:policies}
+\section{Context-Driven Instrumentation}\label{sec:policies}
 
@@
 \section{Conclusion}\label{sec:conclusion}
-We created Granary to address the challenges of instrumenting binary kernel modules.
+%We created Granary to address the challenges of instrumenting binary kernel modules.
 Granary provides mixed-mode execution to remove the overheads of DBT for uninstrumented kernel code.
@@
 Finally, Granary exposes high-level static analysis information to dynamic instrumentation tools, making it easy to match low-level memory reads with specific fields in data structures at the source code level.
-Together, Granary's features simplify the development of powerful kernel module analysis tools, while delivering lower overheads than previous kernel dynamic binary instrumentation solutions.
+
+%Together, Granary's features simplify the development of powerful kernel module analysis tools, while delivering lower overheads than previous kernel dynamic binary instrumentation solutions.
 

### 2013-11-10 VEE-14/paper.tex
Working on the WHAT TO INSTRUMENT section of the paper.

@@
 \usepackage{graphicx}
+\usepackage{multicol}
 
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{Flexible Runtime Code Specialization for the Kernel}
+\title{Flexible Binary Instrumentation for the Kernel}
 
@@
 %We created Granary to address the challenges of module analysis.
-Granary is a framework for creating flexible and efficient tools that analyse or debug arbitrary, binary Linux kernel modules.
+Granary is a framework for creating flexible and efficient tools that analyze or debug arbitrary, binary Linux kernel modules.
 Granary uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules.
@@
 The insane environment of OS kernels also creates several opportunities for analysis tools.
-For example, interrupts can be ``tamed'' and put to new uses; privileged hardware features can be used; and the kernel ABI and source code can be relied upon to be a good predictor of module behaviour at the kernel/module boundary.
+For example, interrupts can be ``tamed'' and put to new uses; privileged hardware features can be used; and the kernel ABI and source code can be relied upon to be a good predictor of module behavior at the kernel/module boundary.
 We also found that the kernel is largely insensitive to transparency issues, allowing us to relax transparency in exchange for both improved performance and greater visibility into the execution of instrumented code.
@@
 Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks.
-This avoids typical poblems with tracking context in just-in-time DBT systems (e.g., maintaining runtime context-tracking state, re-entrancy, and the inability to predict later execution contexts at translation time).
+This avoids typical poblems with tracking context in just-in-time DBT systems (e.g., maintaining runtime context-tracking state, reentrancy, and the inability to predict later execution contexts at translation time).
 
@@
 %Finally, an important missing feature of existing low-level instrumentation systems is the ``big picture'' view that one gets from static analysis information.
-Similarly, static analysis tools lack access to instructions, memory, and processor-specific behaviours that are only available to low-level runtime instrumentation tools.
+Similarly, static analysis tools lack access to instructions, memory, and processor-specific behaviors that are only available to low-level runtime instrumentation tools.
 Existing mixed static/dynamic analysis tools \cite{NaCl,AddressSanitizer,ThreadSanitizer} appear to get the best of both worlds, but sacrifice on runtime flexibility by committing to a single instrumentation policy at a program's compile time.
 The benefits of high-level static analysis motivated its inclusion into Granary using a technique we call \emph{reifying instrumentation}.
-Granary statically analyses the Linux kernel source code to bootstrap learning about  the interfaces that enable modules and the kernel to interact.
+Granary statically analyzes the Linux kernel source code to bootstrap learning about  the interfaces that enable modules and the kernel to interact.
 Granary exposes this information to instrumentation tools in the form of type and function \emph{wrappers}.
@@
 This information allows us to model things like how a typical file system operates on kernel data structures in the process of opening a file.
-From these models, we can construct behavioural patterns about how modules use kernel data structures and functions.
-These patterns can be used to classify modules \cite{DeviceDriverClassification} and identify spurious behaviour \cite{LXFI}.
+From these models, we can construct behavioral patterns about how modules use kernel data structures and functions.
+These patterns can be used to classify modules \cite{DeviceDriverClassification} and identify spurious behavior \cite{LXFI}.
 
@@
 	\item some kernel code is only distributed in a binary format (e.g. proprietary device drivers); and
-	\item virtualization-based approaches are unable to analyse non-virtualizable device drivers \cite{DRK}.
+	\item virtualization-based approaches are unable to analyze non-virtualizable device drivers \cite{DRK}.
 \end{inparaenum}
@@
 %Granary interposes on the system call and interrupt entrypoints, located in the x86-64 \texttt{MSR\_LSTAR} model-specific register and interrupt descriptor table (IDT), respectively. 
-Granary boostraps by duplicating the early entrypoint instructions that are responsible for switching execution onto a kernel stack, and then injects a jump to translated kernel code after the stack switch.
-The translation process continues when a system call or interrupt occurs and follows the jump from duplicated into translated code.
-We only translate code that is known to operate on a kernel stack so that instrumentation tools can safely save transient data in a re-entrant way by spilling that data to the stack.
 %
 %\paragraph{Module}
-%Granary interposes on the Linux kernel's module loading process.
-When a kernel module is loaded, Granary bootstraps by translating the first basic block of the module's initialisation function.
-It then replaces the pointer to that function with a pointer to the translated basic block.
-The translation process continues when the kernel initialises the module by invoking the translated module code.
+%
 %
@@
 Similar to other DBT systems, Granary uses caching and hot code patching to reduce the number of context switches.
-In our experience, kernel code stabilises very quickly: context switches stop happening after the first few seconds of translation.
+In our experience, kernel code stabilizes very quickly: context switches stop happening after the first few seconds of translation.
 
@@
 %\paragraph{Module function wrappers} Granary does not maintain transparency when its type wrappers replace pointers to module functions with pointers to wrapped module functions.
-This is perhaps the riskiest break in transparency within Granary because a module might treat a function pointer as being representative of the module being in some particular state.\footnote{We have not yet encountered a module where changing pointers to module functions into module wrappers altered the module's behaviour under instrumentation.} If Granary is configured to use transparent module function pointer addresses in shared data structures then fast attaching is disabled and tools will not have access to the same level of static program information.
+This is perhaps the riskiest break in transparency within Granary because a module might treat a function pointer as being representative of the module being in some particular state.\footnote{We have not yet encountered a module where changing pointers to module functions into module wrappers altered the module's behavior under instrumentation.} If Granary is configured to use transparent module function pointer addresses in shared data structures then fast attaching is disabled and tools will not have access to the same level of static program information.
 
-\subsection{Re-entrancy}
+\subsection{Reentrancy}
 
 \paragraph{Code cache}
-Granary's code cache and JIT translation mechanism is fully re-entrant.
+Granary's code cache and JIT translation mechanism is fully reentrant.
 Context-switches into Granary disable interrupts and switch execution onto a CPU-private stack.
@@
 
-Granary's solution to this problem is two-fold. Like \texttt{btkernel}, Granary generates ``edge'' code for each direct control-flow instruction (e.g. \texttt{call}, \texttt{jmp}, \texttt{jcc}).
+Like btkernel, Granary solves this problem using  by generating ``edge'' code for each direct control-flow instruction (e.g.
+\texttt{call}, \texttt{jmp}, \texttt{jcc}).
+When a basic block is first translated, its direct control-flow instructions (CTIs) are converted into jumps to CTI-specific edge code.
+Edge code lives outside the code cache and is responsible for transferring control from a basic block ($B_{pred}$) to Granary so that the next basic block ($B_{succ}$) can be found/translated, and then connected ($B_{pred} \underset{\texttt{jz}}{\to} B_{succ}$) via hot code patching.
+Reentrancy of edge code is obtained by spilling registers onto the runtime call stack, and later restoring them before jumping back to the now-patched CTI (\Cref{fig:direct_edge_code}).
+
+Granary also generates edge code for indirect control-flow instructions; however, this edge code is split into two separate components: entry and exit edge code.
+Entry edge code is generated on a per-indirect CTI instruction basis, and spills registers onto the runtim call stack.
+Entry edge is responsible for indirect branch lookup, and will indirectly jump to the exit edge code for a given basic block.
+Exit edge code is generated for each basic block that is targeted by an indirect CTI.
+Exit edge code restores the register state saved by the entry edge code, and then directly jumps to the beginning of the targeted basic block.
 
@@
 
-\section{Mixed-Mode Execution}\label{sec:modes}
-Granary supports two modes of execution: instrumented and native.
-Module code is instrumented and executes from Granary's code cache, which is under Granary's control.
-Non-module kernel code runs natively.
-A mode switch occurs when execution transfers between native and instrumented code.
-Some mode switches happen naturally (e.g., when instrumented code returns to native code) and other mode switches are mediated by Granary (e.g., when instrumented module code invokes a kernel function).
+% Code block showing an example JZ (jump if ZF=1).
+\lstset{language=[x64]Assembler}
+\newsavebox\nativejcc
+\begin{lrbox}{\nativejcc}
+\begin{minipage}[b]{4cm}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+cmp $0x0, %rax
+jz value_is_zero 
+...
+( fall through )
+\end{lstlisting}
+\end{minipage}
+\end{lrbox}
 
-The mode switch from instrumented module code to native kernel code is easier to detect since the instrumented code runs under Granary's control.
-Granary treats all kernel functions as \emph{detach} points, where a mode switch from instrumented to native code occurs.
-At these points, Granary stops executing.
+\newsavebox\translatedjcc
+\begin{lrbox}{\translatedjcc}
+\begin{minipage}[b]{4cm}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+  cmp $0x0, %rax
+jz_to_patch:
+  jmp edge_value_is_zero
+  ...
+edge_value_is_zero:
+  ( spill 2 registers )
+  ( save value_is_zero )
+  ( save target policy )
+  call patcher_for_jz
+  ( restore 2 registers )
+  jmp .jz_to_patch
+\end{lstlisting}
+\end{minipage}
+\end{lrbox}
 
-However, Granary needs a way to regain control when native kernel code invokes module code.
-Our fallback solution for regaining control uses hardware page protection to trap attempts by the kernel to execute native module code.
-We handle these traps by redirecting execution to instrumented module code.
-While comprehensive, this approach is not ideal because: \begin{inparaenum}[i)]
-	\item trapping on every execution attempt introduces overhead; and
-	\item the trap does not provide sufficient information about which interfaces were being used by the kernel to invoke the module.
-\end{inparaenum}
+\begin{figure}[t!]
+\subfloat[Native Code]{\usebox\nativejcc}
+\hfill
+ \subfloat[Translated Code]{\usebox\translatedjcc}
+\caption{\label{fig:direct_edge_code}Example translation of a conditional control flow instruction (\texttt{jz}) and its edge code.
+The edge code saves the intended target of the jump (\texttt{value\_is\_zero}) and the instrumentation policy to apply when translating the targeted code, and then context-switches into Granary by invoking \texttt{patcher\_for\_jz}.
+The \texttt{patcher\_for\_jz} function atomically replaces the {\bf\texttt{ jmp}}\texttt{ edge\_value\_is\_zero} instruction with a fully resolved \texttt{jz} instruction, and upon return, executes the newly resolved \texttt{jz} instruction.}
+\end{figure}
 
-Granary's main approach to regaining control is based on the observation that modules tell the kernel about their interfaces by registering functions with the kernel.
+One challenge with any DBT system is how it copes with non-reentrant instrumentation.
+For example, instrumentation that modifies CPU-private data structures cannot be safely executed by pre-emptive kernel running on a multi-core system.
+If the task being instrumented is interrupted, then the kernel might decide to re-schedule the task to resume its execution on a different CPU.
+Any modifications to CPU-private state would lead to undefined behavior.
+
+DRK's solution to this problem assumes that all injected instrumentation instructions are potentially non-reentrant, and \emph{delays} interrupts occuring within instrumented code \cite{DRK}.
+Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86: interrupts only arrive on \emph{logical} instruction boundaries, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
+This approach introduces subtle complexity when creating instrumentation that injects instructions both before (pre) and after (post) native instructions.
+For example, an interrupt occuring in the pre-instrumentation of a native instruction will be delayed only until the native instruction.
+When the interrupt is handled and execution attempts to return to the interrupted instructions, DRK re-instruments the ``tail" of interrupted basic block, starting from the native instruction.
+Instrumenting this tail involves injecting pre- and post-instrumentation around the first native instruction.
+If the pre-instrumentation is not idempotent, then its re-execution will result in undefined behavior.
+We encountered this issue when using DRK for kernel module instrumentation, where our pre-instrumentation spilled registers to the runtime call stack, and our post-instrumentation restored the spilled registers.
+Because of interrupt delaying, the spilling code was sometimes repeated many times, whereas the restoring was only executed once.
+This led to hard-to-debug kernel crashes.
+
+Btkernel provides no solution to the problem of handling non-reentrant instrumentation, beyond requiring that instrumentation tools disable and re-enable interrupts around non-reentrant instrumentation code \cite{btkernel}.
+Unfortunately, disabling and re-enabling interrupt is costly.
+
+Granary charts a middle ground between the two aforementioned approaches.
+Interrupt delaying has value because it enables non-reentrant instrumentation to be injected without costly disabling/enabling of interrupts, while not supporting delaying has value because not all instrumentation is non-reentrant.
+By default, Granary does not delay interrupts.
+However, instrumentation tools can declare \emph{interrupt delay ranges}: sequences of instructions (which can contain internal control flow) inside of a basic block that must be executed atomically.
+Granary encodes bounds information about interrupt delay ranges inside of each basic block's meta-data.
+When a basic block is interrupted, Granary inspects the meta-data and determines if the interrupt must be delayed.
+If a delay is required, then the instructions belonging to the delay range are copied into a CPU-private buffer, followed by code that re-constructs the interrupt stack frame and re-issues the interrupt to Granary.
+Granary then emulates a return from the current interrupt handler (without issuing an \texttt{iret} instruction), but leaves interrupts disabled and redirects execution into the copied code.
+When the copied code finishes, Granary regains control and observes the reconstructed interrupt stack frame.
+At this point, Granary or the tool can choose to handle the interrupt or defer the handling of the interrupt to the kernel.
+
+%, but introduces implementation complexities and high overheads for interrupt-heavy workloads. Precise interrupt delivery is one of the ways which DRK maintains transparency: interrupt handlers only 
+%The interaction between DRK's interrupt delaying mechanism and how also makes designing instrumentation subtlely harder.
+For example, Granary's \texttt{RCUdbg} tool injects instrumentation inside of read-side critical sections that looks for loads/stores from tainted addresses.
+When necessary, \texttt{RCUdbg} spills and restores registers around each checked load/store operation, so that the registers can be safely used by the checking code.
+This is an example of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after.
+DRK is oblivious to the relationship between pre/post instrumentation, and so an interrupt that occurs within the pre-instrumentation will only be delayed until the native instruction.
+
+\section{Choosing What to Instrument}
+
+\begin{figure*}[ht!]
+\lstset{language=C, tabsize=2, stepnumber=1}
+\begin{multicols}{2}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+struct device_driver {
+	...
+	int (*probe)(struct device *);
+	int (*remove)(struct device *);
+	void (*shutdown)(struct device *);
+	int (*suspend)(struct device *, pm_message_t);
+	int (*resume)(struct device *);
+	...
+	const struct dev_pm_ops *pm;
+	...
+};
+\end{lstlisting}
+\columnbreak
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+TYPE_WRAPPER(struct device_driver, {
+    PRE_OUT {
+        ABORT_IF_FUNCTION_IS_WRAPPED(arg.probe)
+        WRAP_FUNCTION(arg.probe);
+        WRAP_FUNCTION(arg.remove);
+        ...
+    }
+    POST_OUT {
+        POST_WRAP(arg.pm);
+    }
+})
+\end{lstlisting}
+\end{multicols}
+\caption{The Linux device driver structure is shown on the left.
+The automatically generated type wrapper for this structure is shown on the right.
+In the wrapper code, \texttt{arg} is a reference to a \texttt{struct device\_driver} object passed as, or referenced by, an argument to a kernel or module wrapper.
+Code in the \texttt{PRE\_OUT} section is applied to arguments of the wrapped type before a kernel wrapper is invoked.
+Similarly, code in the \texttt{POST\_OUT} section is applied to arguments of the wrapped type after a kernel wrapper is invoked.
+\texttt{POST\_WRAP} invokes the type wrapper that is specific to the value to which it is applied (\texttt{arg.pm}).
+Type wrappers also support \texttt{\_IN} suffixes instead of \texttt{\_OUT} suffixes, which apply to data going into modules (i.e., over module wrappers).
+Finally, the \texttt{RETURN\_} prefix is used to apply some code to return values of either kernel or module wrappers.}
+\label{fig:type_wrapper}
+\end{figure*}
+
+Granary allows tool developers to precisely specify what code should be instrumented.
+Example granularities of instrumentation include the whole kernel, only specific modules, only specific functions, or on-demand instrumentation based on events (e.g.
+interrupts or exceptions).
+
+\subsection{Kernel}
+Like DRK and btkernel, Granary can instrument the whole kernel.
+Instrumenting the whole kernel requires interposing on the system call entrypoint (x86-64 \texttt{MSR\_LSTAR} model-specific register) and on the interrupt vector entrypoints (stored in the interrupt descriptor table).
+For whole-kernel instrumentation, Granary boostraps by duplicating the the early entrypoint instructions that are responsible for switching execution onto a kernel stack, and then injects a jump to translate/instrumented kernel code after the stack switch.
+The translation process continues when a system call or interrupt occurs and follows the jump from duplicated into translated code.
+Granary only translate code that is known to operate on a kernel stack so that instrumentation tools can safely save transient data in a reentrant way by spilling that data to the stack.
+
+\subsection{Modules}
+
+Granary can target instrumentation only at specific kernel modules, while executing the rest of the kernel natively and without overhead.
+Instrumenting only specific kernel modules is particularly useful for debugging tools because modules represent the bulk of new kernel code under development, and contain the most bugs \cite{FaultsInLinux}.
+
+Granary interposes on an existing module by using hardware page protection to mark an existing module as non-executable and trap attempts to execute that module's code.
+Granary transparently recovers from these traps by returning execution to instrumented code.
+Granary takes control of dynamically loaded modules by interposing on the Linux kernel's module loading process.
+When a kernel module is loaded, Granary bootstraps by translating the first basic block of the module's initialization function.
+It then replaces the pointer to that function with a pointer to the translated basic block.
+The translation process continues when the kernel initializes the module by invoking the translated module code.
+As with existing modules, the native code segments of dynamically loaded modules are page protected to prevent and recover from accidental execution.
+
+While sufficient for comprehensiveness, hardware page protection is an inefficient mechanism for gaining control of module code execution.
+That is, after the (instrumented) module initializer runs to completion, Granary needs an efficient way to regain control of future invocations of the module's code.
+Our approach to regaining control is based on the observation that modules tell the kernel about their interfaces by registering functions with the kernel.
 We expect that at least some of the registered functions will be executed by the kernel because this is the mechanism by which modules extend the kernel's functionality.
-To Granary, registered module functions represent potential future \emph{attach} points, where a mode switch from native to instrumented code will occur.
+To Granary, registered module functions represent potential future \emph{attach} points, where a switch from native to instrumented execution will occur.
 Furthermore, Granary discovers additional attach points at detach points by observing pointers to module functions that are passed by modules to the kernel.
@@
 A type wrapper is a function that recursively  traverses the in-memory object graph and converts pointers to module code into pointers to wrapped module functions.
-Type and kernel wrappers are automatically generated at Granary's compile time by scripts that statically analyse the kernel source code.
+Type and kernel wrappers are automatically generated at Granary's compile time by scripts that statically analyze the kernel source code.
 Granary automatically matches any variable in a kernel wrapper to a type wrapper if the base type (absent pointers, specifiers, and qualifiers) of that variable matches the type wrapper's wrapped type.
 Similarly, Granary automatically generates module wrappers using a combination of compile-time meta-programming and runtime code generation to match type wrappers to the declared arguments of \texttt{C} function pointer types.
-Granary's wrapper approach has two benefits: (i) it is more efficient than the trap-based alternative, and (ii) it gives instrumentation tools more information about the executing module (\Cref{sec:reify}).
+Granary's wrapper approach has two benefits: (i) it is more efficient than the trap-based alternative, and (ii) it gives instrumentation tools direct and semantically meaningful access to data crossing the kernel/module boundary.
 
-After Granary bootstraps on the module initialisation process, \emph{attaching} occurs in one of three ways: \begin{enumerate}
+After Granary bootstraps on the module initialization process, \emph{attaching} occurs in one of three ways: \begin{enumerate}
 	\item {\bf Implicit attaching:} the kernel returns to instrumented module code in the code cache.
+This occurs when instrumented moduel code is interrupted, or when the module invokes a kernel function (detaching) and that function returns to the instrumented module code (attaching).
 	\item {\bf Fast attaching:} the kernel invokes a wrapped module function.
+The first instance of fast attaching for any kernel module is when the kernel's module loader invokes the instrumented module initializer function.
 	\item {\bf Slow attaching:} the kernel invokes unwrapped module code.
@@
 Granary \emph{detaches} when control transfers from instrumented code to native (uninstrumented) code. Detaching occurs in one of two ways: \begin{enumerate}
-	\item {\bf Implicit detaching:} instrumented module code returns to the original kernel, or is interrupted (initial interrupt handling is done by the kernel).
+	\item {\bf Implicit detaching:} instrumented module code returns to the kernel code which originally invoked the module, or is interrupted (initial interrupt handling is done by the kernel).
 	\item {\bf Wrapped detaching:} instrumented module code invokes a kernel wrapper, which later transfers control to the kernel.
 \end{enumerate}
+
+Granary's wrappers expose a wealth of static program information to instrumentation tools.
+Instrumentation tools can use information present in wrappers to alter how code is instrumented, or to simply inspect the runtime object graph in a type-safe way.
+For example, we are actively using type information present in wrappers to create a tool that generates models of typical module behavior.
+The wrapper information serves two roles in our tool.
+First, we use type wrappers to assign type-specific IDs to module-allocated memory that is shared across the module/kernel interface.
+Type ID assignment is critical because it allows us to match memory reads and writes to specific kernel data structure fields.
+Second, we use the kernel and module function wrappers to generate call graphs of a module's execution.
+This call graph models module behavior (according to the kernel) because we label module code nodes with kernel data structure and function pointer field names (derived from module wrappers).
+This labelling allows us to generalize across similar modules.
+For example, code reached by calling the \texttt{ext4\_mount} and \texttt{btrfs\_mount} functions from the \texttt{ext4} and \texttt{btrfs} file system modules, respectively, are both labelled as \texttt{file\_system\_type::mount}, because the function addresses are stored (and later replaced by module wrappers) in the \texttt{mount} field of a \texttt{file\_system\_type} data structure.
+The combined records from multiple ``trusted'' modules of the same class (e.g., mature, open-source file systems) models the behavior of typical kernel modules.
+We hope that these models will help us build tools that classify and identify spurious module behavior.
+
+\subsection{Functions}
+
+\subsection{On-demand}
+
+%\section{Mixed-Mode Execution}\label{sec:modes}
+%Granary supports two modes of execution: instrumented and native.
+Module code is instrumented and executes from Granary's code cache, which is under Granary's control.
+Non-module kernel code runs natively.
+A mode switch occurs when execution transfers between native and instrumented code.
+Some mode switches happen naturally (e.g., when instrumented code returns to native code) and other mode switches are mediated by Granary (e.g., when instrumented module code invokes a kernel function).
+%
+%The mode switch from instrumented module code to native kernel code is easier to detect since the instrumented code runs under Granary's control.
+Granary treats all kernel functions as \emph{detach} points, where a mode switch from instrumented to native code occurs.
+At these points, Granary stops executing.
+%
+%However, Granary needs a way to regain control when native kernel code invokes module code.
+Our fallback solution for regaining control uses hardware page protection to trap attempts by the kernel to execute native module code.
+We handle these traps by redirecting execution to instrumented module code.
+While comprehensive, this approach is not ideal because: \begin{inparaenum}[i)]
+%	\item trapping on every execution attempt introduces overhead; and
+%	\item the trap does not provide sufficient information about which interfaces were being used by the kernel to invoke the module.
+%\end{inparaenum}
 
@@
 %Granary uses a technique we call \emph{reifying instrumentation} to provide the benefits of high-level static analysis information to dynamic instrumentation tools.
-Reifying instrumentation bootstraps on Granary's mixed-mode execution approach, which exposes static type information to instrumentation tools.
-For example, we are actively using type information present in wrappers to create a tool that generates models of typical module behaviour.
-The wrapper information serves two roles in our tool.
-First, we use type wrappers to assign type-specific IDs to module-allocated memory that is shared across the module/kernel interface.
-Type ID assignment is critical because it allows us to match memory reads and writes to specific kernel data structure fields.
-Second, we use the kernel and module function wrappers to generate call graphs of a module's execution.
-This call graph models module behaviour (according to the kernel) because we label module code nodes with kernel data structure and function pointer field names (derived from module wrappers).
-This labelling allows us to generalise across similar modules.
-For example, code reached by calling the \texttt{ext4\_mount} and \texttt{btrfs\_mount} functions from the \texttt{ext4} and \texttt{btrfs} modules, respectively, are both labelled as \texttt{file\_system\_type::mount}, because the function addresses are stored (and later replaced by module wrappers) in the \texttt{mount} field of a \texttt{file\_system\_type} data structure.
-The combined records from multiple ``trusted'' modules of the same class (e.g., mature, open-source file systems) will model the behaviour of typical kernel modules.
-We hope that these models will help us build tools that classify and identify spurious module behaviour.
+Reifying instrumentation bootstraps on
 

### 2013-11-11 VEE-14/paper.tex
About to work on the on-demand instrumentation subsection of the what to instrument section.

@@
 Context-switches into Granary disable interrupts and switch execution onto a CPU-private stack.
-Context switches out of Granary and into the code cache or native code returns executuin to the native stack and re-enables interrupts.
-JIT translation, however, presents a problem: context-switches save or restore the entire register state, so as not to disturb the execution of the instrumented program.
-This raises the question: when resolving the target of a direct or indirect control-flow instruction, how should execution be directed to the correct location?
+Context switches out of Granary and into the code cache or native code returns execution to the native stack and re-enables interrupts.
+This approach presents a minor challenge: context-switches save or restore the entire register state, so as not to disturb the execution of the instrumented program.
+Therefore, if a context switch into Granary occurs because the next basic block must be found/translated, then the address of that next basic block must be stored \emph{somewhere} in order for execution to eventually reach that basic block when the context switches back to the code cache.
+But context switching is not allowed to perturb the register state of the instrumented program, and so the target address must be resolved by another means.
 
-Like btkernel, Granary solves this problem using  by generating ``edge'' code for each direct control-flow instruction (e.g.
+Like btkernel, Granary solves this problem of figuring out how to get to where the code needs to go  by generating ``edge'' code for each direct control-flow instruction (e.g.
 \texttt{call}, \texttt{jmp}, \texttt{jcc}).
@@
 
-\subsection{Modules}
+\subsection{Modules}\label{sec:module}
 
@@
 \subsection{Functions}
+
+Granary can instrument or not instrument functions in several different ways.
+For example, \emph{probe-based} instrumentation allows Granary to selectively instrument only a particular function, while \emph{wrapper-based} instrumentation allows Granary to isolate a particular function that is invoked by instrumented code, and choose to instrument that function or run it natively (all while having the flexibility to arbitrarily inspect that function's arguments and return values in a type-safe way).
+
+\paragraph{Probe-based instrumentation}
+
+Probe-based instrumentation allows an instrumentation tool to replace \emph{all} executions of a particular kernel function, regardless of whether or not the replaced function is executed by native or instrumented code \cite{KProbes, AnyWhereAnyTimeDBT, KernInst}.
+Probe-based instrumentation can be used to wrap a particular function without instrumenting its code, or to wrap and instrument the probed function's code.
+
+Probe-based instrumentation provides visibility into specific kernel events (e.g.
+memory allocations) to instrumentation tools that don't instrument the whole kernel (e.g.
+module-only).
+One example tool that combines module-only and probe-based instrumentation is Granary's kernel module memory leak detector.
+This leak detector finds leaks of module-owned memory.
+We say that the some memory is owned by a module iff that memory is deallocated by a module.
+Unfortunately, ownership of module-allocated memory (for an arbitrary, binary module) is not known \emph{a priori}.
+Therefore, our tool conservatively assumes that all module-allocated memory is module-owned, unless it is freed by the kernel (observed via probe-based instrumentation of a kernel memory deallocator).
+We also use probe-based instrumentation to assign module ownership to kernel-allocated \texttt{sk\_buff} data structures, which are allocated by the kernel but deallocated by network device driver modules.
+
+\paragraph{Wrapper-based instrumentation}
+
+Wrapper-based instrumentation makes use of the same wrapper capabilities used by module-only instrumentation (\Cref{sec:module}), but is available regardless of the instrumentation granularity.
+Wrapper-based instrumentation is similar to probe-based instrumentation, but permits more selectivity in terms of the context in which the function will be wrapped.
+That is, a function can be wrapped in one way when executed by instrumented kernel code, and in another way when executed by instrumented module code.
+Finer-grained selectivity can be achieved by having individual instrumentation policies manually decide how to wrap a given function.
+
+%Granary can instrument, or not instrument, particular function
 

### 2013-11-11 VEE-14/paper.tex
Added an example function wrapper.

@@
 
+\begin{figure}[t!]
+\lstset{language=C, tabsize=2, stepnumber=1}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+FUNCTION_WRAPPER(MODULE, __kmalloc, (size, gfp), {
+  void *addr = __kmalloc(size, gfp);
+  if(is_valid_address(addr)) {
+    add_watchpoint(addr, addr, size);
+  }
+  return addr;
+})
+\end{lstlisting}
+\caption{Example module function wrapper for the \texttt{\_\_kmalloc} memory allocator.
+This allocator adds an ``address watchpoint", which Granary's module memory leak detector uses to track the liveness of module-allocated memory.
+Every access of a ``watched address" marks the associated memory as live within the current epoch.
+Memory that has not been live for several consecutive epochs is treated as a potential leak, and a is checked for reachability using a heap memory scanner.}
+\label{fig:malloc_wrapper}
+\end{figure}
+
+Granary's module memory leak detector is an example application that uses the selectivity of wrappers (with respect to their calling context) to distinguish between module-allocated and kernel-allocated memory (\Cref{fig:malloc_wrapper}).
+
 %Granary can instrument, or not instrument, particular function
@@
 \subsection{On-demand}
+
+On-demand instrumentation makes use
 

### 2013-11-11 VEE-14/paper.tex
Worked on on-demand and function wrappers sub sections

--- 2013-11-11 VEE-14/paper.tex
@@
 
+\begin{figure}[t!]
+\lstset{language=C, tabsize=2, stepnumber=1}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+MODULE_FUNCTION_WRAPPER(__kmalloc, (size, gfp), {
+  void *addr = __kmalloc(size, gfp);
+  if(is_valid_address(addr)) {
+    add_watchpoint(addr, addr, size);
+  }
+  return addr;
+})
+\end{lstlisting}
+\caption{Example module function wrapper for the \texttt{\_\_kmalloc} function.
+This wrapper adds an ``address watchpoint" to the address returned by the kernel's \texttt{\_\_kmalloc} memory allocator.
+Granary's module memory leak detector uses address watchpoints to track the liveness of module-allocated memory.
+Every access of a ``watched address" marks the associated memory as live within the current epoch.
+Memory that has not been live for several consecutive epochs is treated as a potential leak, and a is checked for reachability using a heap memory scanner.}
+\label{fig:malloc_wrapper}
+\end{figure}
+
+Granary's module memory leak detector is one example application that uses the selectivity of wrappers (with respect to their calling context) to distinguish between module-allocated and kernel-allocated memory.
+\Cref{fig:malloc_wrapper} shows a module function wrapper that adds ``address watchpoints"  \cite{BehaveOrBeWatched} to module-allocated memory.
+
 %Granary can instrument, or not instrument, particular function
@@
 \subsection{On-demand}
+
+On-demand instrumentation makes use of Granary's control over interrupt and exception handling.
+When a specific exception or interrupt occurs, a Granary tool can turn on instrumentation for the interrupted code.
+This ties in well with Granary's address watchpoints system, which taints memory addresses by changing their high-order 16 bits into a ``non-canonical" form.
+Watched addresses trap when accessed by native code, which introduces an opportunity for on-demand instrumentation to be enabled.
+Another application of on-demand instrumentation is data watchpoints based on hardware page protection and memory shadowing.
+
+Turning on instrumentation when specific data is accessed makes on-demand instrumentation inherently \emph{data-centric} as opposed to \emph{code-centric}.
+That is, instrumentation is targeted at instructions accessing specific data, instead of the typical ``dragnet" approach employed by most DBT systems, which instrument all memory loads/stores so as to capture the select few that access key data.
+
+% instrumentation turns on when specific data is accessed, and can be turned off at the granularity of an instruction, a basic block, the current executing function, or the current executing task/thread.
 

### 2013-11-11 VEE-14/paper.tex
Minor change to a figure

@@
 \begin{minipage}[b]{4cm}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
-cmp $0x0, %rax
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]cmp $0x0, %rax
+
 jz value_is_zero 
 ...
+
 ( fall through )
@@
 \newsavebox\translatedjcc
-\begin{lrbox}{\translatedjcc}
-\begin{minipage}[b]{4cm}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
-  cmp $0x0, %rax
+\begin{lrbox}{\translatedjcc}\begin{minipage}[b]{4cm}\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]  cmp $0x0, %rax
 jz_to_patch:

### 2013-11-11 VEE-14/paper.tex
Last night's work before leaving for home

--- 2013-11-11 VEE-14/paper.tex
@@
 
-\begin{figure}[t!]
-\lstset{language=C, tabsize=2, stepnumber=1}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
-FUNCTION_WRAPPER(MODULE, __kmalloc, (size, gfp), {
-  void *addr = __kmalloc(size, gfp);
-  if(is_valid_address(addr)) {
-    add_watchpoint(addr, addr, size);
-  }
-  return addr;
-})
-\end{lstlisting}
-\caption{Example module function wrapper for the \texttt{\_\_kmalloc} memory allocator.
-This allocator adds an ``address watchpoint", which Granary's module memory leak detector uses to track the liveness of module-allocated memory.
-Every access of a ``watched address" marks the associated memory as live within the current epoch.
-Memory that has not been live for several consecutive epochs is treated as a potential leak, and a is checked for reachability using a heap memory scanner.}
-\label{fig:malloc_wrapper}
-\end{figure}
-
-Granary's module memory leak detector is an example application that uses the selectivity of wrappers (with respect to their calling context) to distinguish between module-allocated and kernel-allocated memory (\Cref{fig:malloc_wrapper}).
-
-%Granary can instrument, or not instrument, particular function
-
 \subsection{On-demand}
 
-On-demand instrumentation makes use
+Demand-based instrumenetation blah blah blah
 

### 2013-11-11 EuroSys'14/abstract.tex

\documentclass[11pt,preprint,onecolumn,nocopyrightspace]{sigplanconf}
\usepackage{amsmath,epsfig,subfigure,multirow}

\usepackage{graphicx}
\usepackage{refstyle}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{framed}
\usepackage{lipsum}
\usepackage{multicol}
\usepackage[english]{babel}
\usepackage{numprint}
\usepackage[dvipsnames]{color}
\definecolor{shadecolor}{RGB}{250,250,250}

\begin{document}

\titlebanner{}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.

\title{Trap or Trace: Implementing Kernel-space Data Watchpoints Efficiently}
\subtitle{}

\authorinfo{Peter Goodman\and Angela Demke Brown}
           {Department of Computer Science\\ University of Toronto}
           {\{pag, demke\}@cs.toronto.edu}

\authorinfo{Akshay Kumar\and Ashvin Goel}
           {Department of Electrical and Computer Engineering\\ University of Toronto}
           {\{akshay, ashvin\}@eecg.toronto.edu}

\maketitle

\begin{abstract}

Finding, understanding, and fixing bugs in operating system kernels is challenging. Watchpoints (also called data breakpoints) are a powerful tool that help with this problem; however, hardware-based watchpoints are too scarce for many uses and existing software-based watchpoints are either slow, or lack contextual information about the watched memory.

In this paper, we introduce \emph{behavioral watchpoints}: a new form of software-based watchpoints. Behavioral watchpoints extend on traditional watchpoints in novel ways: they watch ranges of addresses and they maintain context-specific information about watched memory, enabling debugging applications that cannot be easily implemented using traditional hardware or software watchpoints. For example, our framework can specialize instrumentation based on the context in which a watchpoint is triggered, which makes it possible to limit the scope of heavy-weight instrumentation to only code operating on watched data.

We discuss and evaluate several implementation strategies for our behavioral watchpoints framework, ranging from trap-based watchpoints (data-centric) to whole-kernel instrumentation based watchpoints (code-centric). We show that behavioral watchpoints, as implemented within a kernel dynamic binary translation framework, have reasonable overheads for their intended use in analysing and debugging operating system kernels.
\end{abstract}
\end{document}

%need to explain how the novelty helps (which is what I have tried above but the text is not strong).

%also, we complained earlier that software watchpoints are slow but we don't say that behavioural watchpoints are fast, so there is a disconnect there.

%also, the lack of contextual information, versus having contextual information issue, is hard to understand. Perhaps, saying that behavioral watchpoints enable specializing instrumentation based on context (and giving a simple example about light versus heavy weight instrumentation) might be easier for this audience to understand.

### 2013-11-11 VEE-14/paper.tex
Working on the HOW section

--- 2013-11-11 VEE-14/paper.tex
@@
 Part of building CFGs requires tracking whether or not a given basic block is the entrypoint to a function.
-This is challenging with existing DBT systems because some optimised functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs.
+This is challenging with existing DBT systems because some optimized functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs.
 Granary solves this problem using demand-based runtime code specialization: every entrypoint basic block is instrumented in one way, and later re-executions of any entrypoint basic blocks are specialized and instrumented in another way.
@@
 Granary uses dynamic binary translation (DBT) to instrument the Linux kernel.
-DBT is used for emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimisation}, and runtime instrumentation (analysis \cite{DynamoRIO, DRK, btkernel, ProfilingSimics}, security \cite{Vx32,NaCl,ProgramShepherding}, and debugging \cite{Valgrind}).
+DBT is used for emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimization}, and runtime instrumentation (analysis \cite{DynamoRIO, DRK, btkernel, ProfilingSimics}, security \cite{Vx32,NaCl,ProgramShepherding}, and debugging \cite{Valgrind}).
 We chose DBT because \begin{inparaenum}[i)]
@@
 
-\section{Choosing What to Instrument}
+\section{Specializing on What to Instrument}
 
@@
 
-\section{Context-Driven Instrumentation}\label{sec:policies}
+\section{Specializing on How to Instrument}\label{sec:how}
 
-A natural extension of mixed-mode execution is to give Granary tools the ability to dynamically switch the policy used to instrument code.
-Instrumenting only module code allows Granary to specialize the execution of native code.
-In the same way, policy switching allows tools to specialize the execution of instrumented code.
-In fact, optimising the performance of an early Granary tool was the original motivation for policy switching.
+Granary gives tools the ability to be selective about what is instrumented.
+A natural extension of this selectivity is to allow tools to specialize how code is instrumented by dynamically switching the policy used to instrument code.
+Specialization is desirable in several cases.
+For example, some instrumentation is high-overhead, but does not need to be applied to all code.
+One example of this is Granary's \texttt{RCUdbg} tool, which instruments read-side and write-side critical sections more heavily than the rest of code.
+Unfortunately, existing DBT systems fall short in their support for runtime code specialization \cite{DRK,btkernel,Pin,DynamoRIO}.
 
-We developed a Granary tool that detects several Read-Copy-Update (RCU) API misuses in module code.
+To implement the equivalent of \texttt{RCUdbg}'s runtime code specialization using an existing DBT system requires that a tool developer maintain, update, and check runtime state to know whether or not code is executing inside or outside of a read-side critical section.
+If a basic block is executing in a read-side critical section, then instrumentation should be enabled, and if it's not in a read-side critical section then it should be disabled.
+Enabling/disabling instrumentation might be done at the instruction granularity, which introduces at least one branch instruction per native instruction, or at the granularity of basic blocks, which requires generating two copies of each native basic block, where a single branch at the beginning of a basic block chooses between the instrumented or uninstrumented version.
+Up until now, it has been assumed that there is a mechanism for maintaing state that consistently tracks the current execution context of kernel code (inside vs.
+outside of a read-side critical section).
+However, maintaining state in a way that is reentrant and resilient again arbitrary pre-emption and resumption is onerous.
+
+%Existing systems move the burden of code specialization to instrumentation tool developers.
+
+Fortunately for tool developers, Granary has built-in support for runtime code specialization that does not require manual state management or cause code cache bloat.
+Granary approach to runtime code specialization uses a technique called \emph{context-driven instrumentation}.
+
+%For example, instrumenting only module code allows Granary to specialize the execution of native code.
+In the same way, policy switching allows Granary's \texttt{RCUdbg} tool to specialize the execution of instrumented code executing inside of read-side and write-side critical sections.
+In fact, optimizing the performance of an early Granary tool was the original motivation for policy switching.
+
+\subsection{Context-Driven Instrumentation}\label{sec:policies}
+
+%We developed a Granary tool that detects several Read-Copy-Update (RCU) API misuses in module code.
 Our tool focused on read-side critical sections (delimited by calls to \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock} in the code), however, most code executes outside of read-side critical sections.
-As an optimisation, we wanted our heavyweight API-checking instrumentation to apply only when the code was executing within a read-side critical section.
-Implementing this optimisation was challenging because each basic block was only translated once, and we had no way to know whether or not the heavyweight instrumentation should be applied to it.
+As an optimization, we wanted our heavyweight API-checking instrumentation to apply only when the code was executing within a read-side critical section.
+Implementing this optimization was challenging because each basic block was only translated once, and we had no way to know whether or not the heavyweight instrumentation should be applied to it.
 Knowing the state (within or outside a read-side critical section) at translation time was not sufficient, since the same translated block could later be executed in the opposite context.
@@
 
-To implement this optimisation, Granary enables tailoring instrumentation to the context in which the instrumented code will execute.
-To do so, Granary allows different versions of the same module code to co-exist within Granary's code cache.
-That is, a different version of each of a module's basic blocks exists in Granary's code cache  for each encountered execution context.
-In the case of our RCU checker tool, there were two execution contexts: within and outside of a read-side critical section.
+%To implement this optimization,
+
+Granary enables specializing instrumentation to the context in which the instrumented code will execute.
+Specialization is achieved by allowing multiple, differently instrumented versions of the same code to co-exist within Granary's code cache.
+That is, if the next basic block to execute has not yet been translated/instrumented for the current execution context, then a new version of that basic block will be created  specifically for the the current execution context.
+
+%In the case of our \texttt{RCUdbg}, there were three execution contexts: within and outside of a read-side critical section.
 If the same module basic block is  executed in the two contexts, then Granary's code cache would contain two different instrumented versions of that basic block.
@@
 
+Granary tools identify execution contexts by creating instrumentation policies.
 An instrumentation policy is both a name for an execution context, as well as a function that decides how to instrument basic blocks that will execute within that context.
@@
 Tools are not limited to one policy though: any policy can declare a policy switch that will take effect when a selected control-transferring instruction (CTI) is executed.
-For example, our RCU tool invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.
+The effect of a tool specifying a policy switch on a CTI is that the basic block(s) targeted by that CTI will be instrumented according to the specified policy.
+CTIs with unspecified policies inherit their policies from their containing basic blocks.
+
+For example, our \texttt{RCUdbg} invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.
+A similar switch occurs from $P_{\mathit{read\_critical}}$ back to $P_{\mathit{null}}$ at a \texttt{rcu\_read\_unlock}.
+Specialization that distinguishes between the inside and outside of read-side critical sections allows \texttt{RCUdbg} to do heavy-weight checks on functions like \texttt{memset}, but only when \texttt{memset} is invoked from within a read-side critical section.
 
 Because policies name an execution context, they also represent states in a finite state machine.
-That is, a module's execution is in the state named by a policy if the code executing was instrumented by that policy.
+That is, code cache execution is in the state named by a policy if the  executing code was instrumented by that policy.
 A state transition occurs when control transfers from code instrumented by one policy to code instrumented by another policy.
@@
 If two read-side critical sections are nested then switching from $P_{\mathit{read\_critical}}$ to $P_{\mathit{null}}$ on the first \texttt{call} to \texttt{rcu\_read\_unlock} meant that our tool would lose track of being in the context of the outer read-side critical section.
-We solved this limitation by observing that, in most cases, function returns are natural policy reverting points.
-This is because the runtime call stack encodes both the context in which the current code is executing, as well as the continuation of the current executing code.
+\texttt{RCUdbg} solves this problem by tracking the nesting depth of RCU read-side critical sections using policies, e.g.
+$P_{\mathit{read\_critical}_1}$, $P_{\mathit{read\_critical}_2}$, etc.
 
-The effect of a tool specifying a policy switch on a CTI is that the basic block(s) targeted by that CTI will be instrumented according to the specified policy.
-CTIs with unspecified policies inherit their policies from their containing basic blocks.
-As hinted at above, \texttt{ret} instructions cannot explicitly switch policies, which allows a function instrumented by a different policy than its caller to return to the caller's context (i.e., policy).
-The asymmetry between \texttt{call}s and \texttt{ret}s is intentional: \texttt{call}s place contextual breadcrumbs (in the form of return addresses) on the runtime call stack, and \texttt{ret}s read these breadcrumbs to return to a previous context.
-Under this lens, the instruction pointer tracks our current policy, and return addresses form a stack of previous policies.
-Policy switches behave similarly to state transitions in a pushdown automaton; \texttt{call} instructions push a new state onto the stack for each function, \texttt{ret} instructions pop the current function's state from the stack, and other CTIs induce state transitions within the current function by altering control flow.
+An alternative and sometimes more powerful model for policy switching is to prevent policy switching on function return instructions (e.g.
+either through manually specifying a policy switch, or by inheriting the policy from the basic block containing the \texttt{ret} instruction).
+In most cases, function returns are natural policy reverting points.
+This is because the current instruction pointer (program counter) encodes the current execution context (by being in a basic block instrumented by some policy), and the code cache return addresses stored on the runtime call stack encode all previous execution contexts.
+This asymmetry between \texttt{call}s and \texttt{ret}s yields a more powerful context-tracking state machine: \texttt{call}s place contextual breadcrumbs (in the form of return addresses) on the runtime call stack, and \texttt{ret}s read these breadcrumbs to return to a previous context.
+Policy switches under this lens behave similarly to state transitions in a pushdown automaton; \texttt{call} instructions push a new state onto the stack for each function, \texttt{ret} instructions pop the current function's state from the stack, and other CTIs induce state transitions within the current function by altering control flow.
 
 Granary implements policy tracking and switching by encoding policy information into the meta-data and CTIs of basic blocks.
+\Cref{fig:direct_edge_code} shows an example of how policy information is directly recorded into control-transfer edge code.
 If a policy switch is not specified on an instrumented CTI then that CTI inherits the policy used to instrument the basic block containing the CTI.
@@
 Granary decodes and instruments the targeted instructions according to the input policy information.
-Because every non-\texttt{ret} CTI encodes policy information, and because Granary's translation mechanism depends only on CTI policy and target information, Granary is able to ensure that policy information is never lost or corrupted, even in the face of concurrent executions of the same module code, arbitrary pre-emption, and arbitrary resumption.
+Because Granary's translation mechanism depends only on CTI policy and target information, Granary is able to ensure that policy information is never lost or corrupted, even in the face of concurrent executions of the same module code, arbitrary pre-emption, and arbitrary resumption.
 
@@
 Prior work can be classified as either whole-kernel/system DBT, or probe-based kernel instrumentation. We discuss how Granary differs from these approaches.
-\vspace{-3pt}\paragraph{Whole-kernel/system DBT} DRK \cite{DRK} is a kernel space port of the DynamoRIO \cite{DynamoRIO} DBT framework.
-DRK instruments the entire kernel, including modules, and follows a strict transparency model, which limits the flexibility of instrumentation and increases overheads.
+\paragraph{Whole-kernel/system DBT}
 PinOS \cite{PinOS} is a port of the Pin \cite{Pin} DBT framework that performs whole-system instrumentation.
@@
 
-\vspace{-3pt}\paragraph{Probe-based Instrumentation} Several systems support injecting code at specific locations within kernel code.
-For example, KernInst \cite{KernInst} can inject code at almost any location in an unmodified commodity OS.
+QMEU is a flexible, whole-system emulator with kernel-level paravirtualization support (using KVM) \cite{QEMU}. Like PinOS, QEMU cannot instrument non-virtualizable device drivers.
+
+DRK \cite{DRK} is a kernel space port of the DynamoRIO \cite{DynamoRIO} DBT framework.
+DRK instruments the entire kernel and all device drivers/modules.
+DRK follows a strict transparency model, which limits the flexibility of instrumentation and increases overheads.
+
+Btkernel is a fast and scalable system for the kernel \cite{btkernel}.
+Like Granary, but unlike DRK, btkernel applies a relaxed transparency model when instrumenting the kernel.
+For exmaple, code cache addresses are exposed to the instrumented kernel in the form of function call return addresses and interrupt return addresses.
+Unlike Granary, btkernel operates using an instruction translation ``rulebook", which limits instrumentation to being defined ahead-of-time, on a per-instruction basis.
+Btkernel does not support any form of runtime code specialization.
+That is, like DRK, btkernel instruments all code, but cannot change the granularity of instrumentation or the instrumentation policy applied to code.
+
+\paragraph{Probe-based Instrumentation} Several systems support injecting code at specific locations within kernel code.
+
+KernInst \cite{KernInst} can inject code at almost any location in an unmodified commodity OS.
+By default, KernInst uses debugging information normally present in kernel binaries to specify probe-points; however, absolute memory addresses can also be provided.
+
 Other examples with similar or more restricted functionality include LTTng \cite{LTTng}, KProbes \cite{KProbes}, DProbes \cite{DProbes}, and ftrace \cite{ftrace}.

### 2013-11-11 VEE-14/paper.tex
Added a key sentence into the intro, described what has been evaluated.

@@
 In practice, we would like to target instrumentation at specific code for efficiency, while also allowing rich and detailed instrumentation of the targeted code.
-This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary tools to instrument only specific code while the majority of code executes with zero overhead.
+Limiting the scope of what is instrumented is  important because the overhead introduced by using runtime code instrumentation is usually dominated by the instrumentation itself and not the runtime system supporting that instrumentation.
+
+%This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary tools to instrument only specific code while the majority of code executes with zero overhead.
 
@@
 
+What has been evaluated:
+
+\paragraph{lmbench}
+\begin{itemize}
+	\item Btkernel
+	\item Granary, full-kernel, null
+	\item Granary, full-kernel, even/odd
+	\item Granary, full-kernel, even/odd, where even/odd is a runtime check and branch.
+\end{itemize}
+
+\paragraph{fileserver}
+Setup: 1GB RAM, 3GB partition, ext3+jbd, >1GB mounted file system.
+\begin{itemize}
+	\item full-kernel
+	\begin{itemize}
+		\item Btkernel
+		\item Granary,null
+		\item Granary, even/odd
+		\item Granary, even/odd, where even/odd is a runtime check and branch.
+	\end{itemize}
+	\item module-only, (ext3+jbd)
+	\begin{itemize}
+		\item Granary,null
+		\item Granary, even/odd
+		\item Granary, even/odd, where even/odd is a runtime check and branch.
+	\end{itemize}
+\end{itemize}
+
+\paragraph{kernel compile}
+Same as fileserver, but with 8GB RAM instead of 3GB.
+
+\paragraph{recursive copy of kernel source tree}
+Same as fileserver, but with 8GB RAM instead of 3GB.
+
 \section{Related Work}\label{sec:related}

### 2013-11-13 VEE-14/paper.tex
Working on adding some descriptions of RCU in.

@@
 \usepackage{graphicx}
-\usepackage{multicol}
+\usepackage{multicol,multirow}
+\usepackage{tabularx}
 
@@
 In write-side critical sections, \texttt{RCUdbg} detects when writers write to shared data without using the correct API.
-
-%is a synchronization mechanism that permits readers to operate concurrently with writers, but does not guarantee that readers observe the most up-to-date changes (by writers) to shared state.
-RCU readers
-
-%When used correctly, RCU readers must declare that they will operate on shared state by entering into a read-side critical section, delimited by \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock}, respectively.
-When inside a read-side critical section, a pointer to shared state \emph{must} be obtained by invoking \texttt{rcu\_dereference}.
-Thereafter, a reader can freely access the shared state through this pointer, but only up until the
 
@@
 
-%\section{Reifying Instrumentation}\label{sec:reify}
+\section{Application: RCUdbg}
+
+Read-Copy-Update (RCU) is a scalable, lock-free synchronization mechanism used in the Linux kernel \cite{RCU,RCUInLinux}.
+RCU is used to synchronize accesses (reads) and modifications (writes) of shared state.
+When correctly used, RCU permits concurrent reads and writes to shared data structures, but only if the following properties are satisfied: \begin{enumerate}
+	\item Readers must tolerate an inconsistent view of the shared data structure.
+That is, a reader operating concurrently with a writer might observe either the version of the data before or after the writer's changes have been published.
+	\item Readers must operate within \emph{read-side} critical sections, which are delimited by \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock}, respectively.
+	\item Within read-side critical sections, readers can only read a shared data structure after obtaining a pointer to that shared data structure using the \texttt{rcu\_dereference} API function.
+\texttt{rcu\_dereference} operates on a pointer to a shared data structure, and returns a pointer to that can be used to safely access the shared data structure within the remainder of the reader's read-side critical section.
+Notably, accessing the original data structure and not the dereferenced data structure pointer--even after an \texttt{rcu\_dereference}--is undefined behavior.
+That is, within a read-side critical section, RCU only guarantees the existence of data structures pointed to by RCU-dereferenced pointers.
+	\item Writers must synchronize among themselves, e.g.
+via mutual exclusion locks or spin locks, before any writes to a shared data structure are performed.
+We say writes to the shared data structure occur within \emph{write-side} critical sections.
+	\item RCU only defines how writers can update pointers within a shared data structure.
+Modification of non-pointer fields within an RCU-protected data structure is beyond the scope of RCU's correctness guarantees.
+	\item Writers wishing to update a pointer in a shared data structure must use the \texttt{rcu\_assign\_pointer} API function.
+\texttt{rcu\_assign\_pointer} assigns a new value to a pointer in a shared data structure.
+	\item Readers must access any data structure pointer field operated on by an \texttt{rcu\_\linebreak[0]assign\_pointer} using \texttt{rcu\_\linebreak[0]dereference}.
+For example, if the \texttt{next} pointer field in \emph{any} list element of an RCU-protected list has been assigned to by \texttt{rcu\_assign\_\linebreak[0]pointer}, then \emph{all} \texttt{next} pointer fields must be accessed by readers of that list by using \texttt{rcu\_dereference}.
+\end{enumerate}
+
+The above list of requirements is not exhaustive.
+For example, we have omitted a description of how elements within an RCU-protected collection can be garbage collected after their removal from the collection.
+\texttt{RCUdbg} is primarily concerned with finding bugs related to the interaction of the above list of requirements.
+\Cref{fig:rcu_bugs} shows some example usages of RCU, and some of the more subtle pointer-related bugs that can happen because of the interaction of rules (1), (2), (3), and (7).
+
+% NOTE!!!!
 %
-%Granary uses a technique we call \emph{reifying instrumentation} to provide the benefits of high-level static analysis information to dynamic instrumentation tools.
-Reifying instrumentation bootstraps on
+% All of those spaces before /* BUG! */ and such are so that the lstlisting will be *close* to the columnwidth!! Somehow linewidth param of lstset didn't work for me :-(
+
+\lstset{
+	language=C,
+	tabsize=2,
+	stepnumber=1,
+	morekeywords={rcu_read_lock,rcu_dereference,rcu_read_unlock,rcu_assign_pointer,spin_lock,spin_unlock}}
+
+% Example code showing an existence bug.
+\newsavebox\rcuexistencebug
+\begin{lrbox}{\rcuexistencebug}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+rcu_read_lock();
+  elm = rcu_dereference(list_head);
+  data = elm->data;                  /* Valid */
+rcu_read_unlock();
+data = elm->data;                     /* BUG! */
+\end{lstlisting}
+\end{lrbox}
+
+% Example code showing another existence bug.
+\newsavebox\rcuderefbug
+\begin{lrbox}{\rcuderefbug}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+rcu_read_lock();
+  elm = rcu_dereference(list_head);
+  if(elm != NULL)
+    data = list_head->data;           /* BUG! */
+rcu_read_unlock();
+\end{lstlisting}
+\end{lrbox}
+
+% Example code showing another existence bug.
+\newsavebox\rcuassignderefbug
+\begin{lrbox}{\rcuassignderefbug}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+spin_lock();
+  elm = kmalloc(...);
+  elm->next = list_head->next;
+  rcu_assign_pointer(list_head->next, elm);
+spin_unlock();
+...
+rcu_read_lock();
+  elm = rcu_dereference(list_head);
+  while(elm != NULL) {
+    data += elm->data;
+    elm = elm->next;                  /* BUG! */
+  }
+rcu_read_unlock();
+\end{lstlisting}
+\end{lrbox}
+
+\begin{figure}[ht!] %
+\subfloat[Despite correctly obtaining a pointer to a shared data structure using \texttt{rcu\_dereference} within a read-side critical section, properties (2) and (3) tell us that the data pointed to by \texttt{elm} is not guaranteed to exist after the read-side critical section ends.]{\fbox{\usebox\rcuexistencebug}} %
+
+\subfloat[Despite the call to \texttt{rcu\_dereference}, properties (1) and (3) tell us that \texttt{list\_head} is not guaranteed to be equivalent to \texttt{elm} because a writer might have concurrently changed the value of the \texttt{list\_head} pointer.
+That is, if \texttt{elm} is non-\texttt{NULL}, then the data pointed to by \texttt{elm} is guaranteed to exist for the remainder of the read-side critical section.
+However, the value of \texttt{list\_head} is \emph{not} guaranteed to remain consistent--even within read-side critical sections--and so \texttt{list\_head} might point to a \texttt{NULL} pointer when \texttt{elm} does not.]{\fbox{\usebox\rcuderefbug}} %
+
+\subfloat[This example shows the code for a writer thread adding a new element into the second position of an RCU-protected list.
+The code for reader threads is also shown, where readers iterate over the list by traversing \texttt{next} pointers and sum up the \texttt{data} fields of each visited list element.
+The access of \texttt{elm->next} violates of properties (3) and (7).
+That is, the next element of the list is only guaranteed to exist within the read-side critical section if the reader executes \texttt{rcu\_dereference(elm->next)}.
+Directly reading \texttt{elm->next} might yield a \emph{temporarily} valid pointer to the next list element, or it might return an inconsistent value for the pointer that is a mix of the two versions of the pointer.]{\fbox{\usebox\rcuassignderefbug}} %
+\caption{\label{fig:rcu_bugs}Example RCU bugs related to misuses of pointers to shared data.}
+\end{figure}
 
@@
 		\item Btkernel
-		\item Granary,null
+		\item Granary, null
 		\item Granary, even/odd
@@
 	\begin{itemize}
-		\item Granary,null
+		\item Granary, null
 		\item Granary, even/odd

### 2013-11-13 VEE-14/paper.tex
Working on characterizing RCU bugs

@@
 Read-Copy-Update (RCU) is a scalable, lock-free synchronization mechanism used in the Linux kernel \cite{RCU,RCUInLinux}.
-RCU is used to synchronize accesses (reads) and modifications (writes) of shared state.
-When correctly used, RCU permits concurrent reads and writes to shared data structures, but only if the following properties are satisfied: \begin{enumerate}
+When correctly used, RCU permits concurrent reads and writes to shared data structures, but requires that writers synchronize among themselves.
+However, like other synchronization mechanisms, RCU has a deceptively simple API.
+Misuses of the RCU API often result in hard-to-find bugs that can be implementation-dependent (i.e.
+bugs that manifest in one version of RCU but not another) or architecture-specific (e.g.
+bugs that only exhibit on architectures with a relaxed memory model).
+For example, RCU bugs sometimes manifest as use-after-free bugs, which can result in silent data corruption (if the underlying memory remains used) or kernel panics.
+
+Unfortunately, approaching RCU from the perspective of finding use-after-free errors is insufficient.
+A better characterization of this class of RCU bugs is proof-of-existence bugs: a data structure cannot be proven to exist at some potentially allocated memory.
+Existence in this case relates to semantic reachability.
+For example, if a list element $E_n$ from an RCU-protected list is reachable from the head of the list by traversing zero-or-more intermediate elements $E_0, ..., E_{n-1}$, then $E_n$ is said to exist.
+While related, memory leak detection (e.g.
+via conservate mark and sweep) is also insufficient for ladidadidaaaa TODO TODO TODO
+
+% (assuming that pointer reachability is sufficient to determine semantic reachability in this case).
+
+Granary's \texttt{RCUdbg} tool is designed to catch RCU bugs related to kernel code using pointers to RCU-protected data structures that haven't been proven to exist.
+
+%the associated memory may or may not be free, but the data structure occupying that memory can't be proven to exist. This
+
+%  However, this class of RCU bugs is not well-suited to being detected by a use-after-free memory checker tool
+%RCU is notoriously hard to use, despite its relatively simple API.
+RCU usage bugs, like bugs with other synchronization mechanisms, often stay hidden and only exhibit themselves for very specific thread interleavings/schedules.
+
+% go silently unnoticed because the memory
+
+%While bugs related to mutual exlusion locks might result in easily observable deadlocks, bugs related to RCU typically result in use-after-free errors, which sometimes go silently unnoticed, sometimes result in kernel panics, and most of the time work as expected because the associated data is .
+
+but only if the following properties are satisfied: \begin{enumerate}
 	\item Readers must tolerate an inconsistent view of the shared data structure.
@@
 \begin{figure}[ht!] %
-\subfloat[Despite correctly obtaining a pointer to a shared data structure using \texttt{rcu\_dereference} within a read-side critical section, properties (2) and (3) tell us that the data pointed to by \texttt{elm} is not guaranteed to exist after the read-side critical section ends.]{\fbox{\usebox\rcuexistencebug}} %
+\subfloat[\label{fig:rcu_existence_bug}Despite correctly obtaining a pointer to a shared data structure using \texttt{rcu\_dereference} within a read-side critical section, properties (2) and (3) tell us that the data pointed to by \texttt{elm} is not guaranteed to exist after the read-side critical section ends.]{\fbox{\usebox\rcuexistencebug}} %
 
-\subfloat[Despite the call to \texttt{rcu\_dereference}, properties (1) and (3) tell us that \texttt{list\_head} is not guaranteed to be equivalent to \texttt{elm} because a writer might have concurrently changed the value of the \texttt{list\_head} pointer.
+\subfloat[\label{fig:rcu_deref_bug}Despite the call to \texttt{rcu\_dereference}, properties (1) and (3) tell us that \texttt{list\_head} is not guaranteed to be equivalent to \texttt{elm} because a writer might have concurrently changed the value of the \texttt{list\_head} pointer.
 That is, if \texttt{elm} is non-\texttt{NULL}, then the data pointed to by \texttt{elm} is guaranteed to exist for the remainder of the read-side critical section.
@@
 
-\subfloat[This example shows the code for a writer thread adding a new element into the second position of an RCU-protected list.
+\subfloat[\label{fig:rcu_assign_bug}This example shows the code for a writer thread adding a new element into the second position of an RCU-protected list.
 The code for reader threads is also shown, where readers iterate over the list by traversing \texttt{next} pointers and sum up the \texttt{data} fields of each visited list element.
@@
 \end{figure}
+
+The \texttt{RCUdbg} tool detects the the following bugs: \begin{enumerate}
+	\item Reads through an RCU-dereferenced pointer outside of a read-side critical section (\Cref{fig:rcu_existence_bug}).
+	\item Reads through an RCU-assigned pointer that don't first use \texttt{rcu\_dereference} (\Cref{fig:rcu_deref_bug}).
+	\item Reads through fields of RCU-protected data structures that should be accessed via \texttt{rcu\_dereference} (\Cref{fig:rcu_assign_bug}).
+	\item Reads through an RCU-dereferenced pointer in the wrong read-side critical section.
+	\item Writes through an RCU-dereferenced pointer.
+	\item Writes through an RCU-assigned pointer.
+\end{enumerate}
+
+\texttt{RCUdbg} uses Granary's address watchpoints feature \cite{BehaveOrBeWatched} to ``taint" addresses participating in RCU-protected data structures.
 
@@
 
+TODO: JIFL
+
 \paragraph{Probe-based Instrumentation} Several systems support injecting code at specific locations within kernel code.
+
+TODO: SystemTap
 

### 2013-11-13 VEE-14/paper.tex
Updated the RCU stuff with a lot of details and bug fixes.

@@
 However, like other synchronization mechanisms, RCU has a deceptively simple API.
-Misuses of the RCU API often result in hard-to-find bugs that can be implementation-dependent (i.e.
+Misuses of the RCU API often result in hard-to-find bugs that can be implementation-specific (i.e.
 bugs that manifest in one version of RCU but not another) or architecture-specific (e.g.
@@
 
-Unfortunately, approaching RCU from the perspective of finding use-after-free errors is insufficient.
-A better characterization of this class of RCU bugs is proof-of-existence bugs: a data structure cannot be proven to exist at some potentially allocated memory.
-Existence in this case relates to semantic reachability.
-For example, if a list element $E_n$ from an RCU-protected list is reachable from the head of the list by traversing zero-or-more intermediate elements $E_0, ..., E_{n-1}$, then $E_n$ is said to exist.
-While related, memory leak detection (e.g.
-via conservate mark and sweep) is also insufficient for ladidadidaaaa TODO TODO TODO
-
 % (assuming that pointer reachability is sufficient to determine semantic reachability in this case).
 
-Granary's \texttt{RCUdbg} tool is designed to catch RCU bugs related to kernel code using pointers to RCU-protected data structures that haven't been proven to exist.
+Granary's \texttt{RCUdbg} tool is designed to catch RCU bugs related to kernel code misusing pointers to RCU-protected data structures.
+The following is a non-exhaustive list of requirements that must be met by reader and writer threads operating on an RCU-protected data structure.
+The list omits a description of how RCU-protected data structures are garbage collected.
+Details concerning garbage collection are beyond the scope of the bugs caught by \texttt{RCUdbg}.
 
@@
 
-but only if the following properties are satisfied: \begin{enumerate}
+\begin{enumerate}
 	\item Readers must tolerate an inconsistent view of the shared data structure.
 That is, a reader operating concurrently with a writer might observe either the version of the data before or after the writer's changes have been published.
+This can be complicated when writers perform multiple, discrete updates to a data structure.
+One example of this is a doubly linked list: a writer that is adding a new entry into a doubly linked list must update at least two RCU-protected pointers.
+
 	\item Readers must operate within \emph{read-side} critical sections, which are delimited by \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock}, respectively.
+
 	\item Within read-side critical sections, readers can only read a shared data structure after obtaining a pointer to that shared data structure using the \texttt{rcu\_dereference} API function.
-\texttt{rcu\_dereference} operates on a pointer to a shared data structure, and returns a pointer to that can be used to safely access the shared data structure within the remainder of the reader's read-side critical section.
+An RCU-protected pointer is like a normal pointer, but points to a potentially incomplete or partially-constructed data structure.
+\texttt{rcu\_dereference} operates on an RCU-protected pointer and returns an ``unprotected" version of the pointer.
+The reader can use the unprotected pointer to safely access the shared data structure within the remainder of the its read-side critical section.
+If \texttt{rcu\_dereference} is not used, then the reader might observe a partially-constructed (i.e.
+incomplete) version of the shared data structure.
 Notably, accessing the original data structure and not the dereferenced data structure pointer--even after an \texttt{rcu\_dereference}--is undefined behavior.
-That is, within a read-side critical section, RCU only guarantees the existence of data structures pointed to by RCU-dereferenced pointers.
-	\item Writers must synchronize among themselves, e.g.
-via mutual exclusion locks or spin locks, before any writes to a shared data structure are performed.
+That is, within a read-side critical section, RCU only guarantees the ``completeness" of data structures pointed to by RCU-dereferenced pointers.
+Data accessed within a read-side critical section is \emph{not} guaranteed to exist outside the read-side critical section.
+
+	\item Writers must synchronize among themselves before any writes to a shared data structure are performed.
+Synchronization is typically achieved via mutual exclusion locks, spin locks, or by having a single, designated writer thread.
 We say writes to the shared data structure occur within \emph{write-side} critical sections.
-	\item RCU only defines how writers can update pointers within a shared data structure.
-Modification of non-pointer fields within an RCU-protected data structure is beyond the scope of RCU's correctness guarantees.
-	\item Writers wishing to update a pointer in a shared data structure must use the \texttt{rcu\_assign\_pointer} API function.
-\texttt{rcu\_assign\_pointer} assigns a new value to a pointer in a shared data structure.
-	\item Readers must access any data structure pointer field operated on by an \texttt{rcu\_\linebreak[0]assign\_pointer} using \texttt{rcu\_\linebreak[0]dereference}.
-For example, if the \texttt{next} pointer field in \emph{any} list element of an RCU-protected list has been assigned to by \texttt{rcu\_assign\_\linebreak[0]pointer}, then \emph{all} \texttt{next} pointer fields must be accessed by readers of that list by using \texttt{rcu\_dereference}.
+
+	\item RCU only defines how writers can update RCU-protected pointer within a shared data structure.
+Modification of non-RCU-protected pointer fields within an RCU-protected data structure is beyond the scope of RCU's correctness guarantees.
+
+	\item Writers wishing to update an RCU-protected pointer in a shared data structure must use the \texttt{rcu\_assign\_pointer} API function.
+\texttt{rcu\_assign\_pointer} assigns a new value to an RCU-protected pointer in a shared data structure.
+
+	\item Readers must access any pointer field operated on by an \texttt{rcu\_\linebreak[0]assign\_pointer} using \texttt{rcu\_\linebreak[0]dereference}.
+For example, if the \texttt{next} pointer field in \emph{any} list element of an RCU-protected list has been assigned to by \texttt{rcu\_assign\_\linebreak[0]pointer}, then \emph{all} \texttt{next} pointer fields must be accessed by readers of that list by using \texttt{rcu\_\linebreak[0]dereference}.
+In this sense, using \texttt{rcu\_\linebreak[0]dereference} and \texttt{rcu\_assign\_\linebreak[0]pointer} marks specific variables and data type fields as containing RCU-protected pointers.
+
 \end{enumerate}
-
-The above list of requirements is not exhaustive.
-For example, we have omitted a description of how elements within an RCU-protected collection can be garbage collected after their removal from the collection.
-\texttt{RCUdbg} is primarily concerned with finding bugs related to the interaction of the above list of requirements.
-\Cref{fig:rcu_bugs} shows some example usages of RCU, and some of the more subtle pointer-related bugs that can happen because of the interaction of rules (1), (2), (3), and (7).
 
@@
 The access of \texttt{elm->next} violates of properties (3) and (7).
-That is, the next element of the list is only guaranteed to exist within the read-side critical section if the reader executes \texttt{rcu\_dereference(elm->next)}.
-Directly reading \texttt{elm->next} might yield a \emph{temporarily} valid pointer to the next list element, or it might return an inconsistent value for the pointer that is a mix of the two versions of the pointer.]{\fbox{\usebox\rcuassignderefbug}} %
+That is, the next element of the list is only guaranteed to be fully constructed within the read-side critical section if the reader executes \texttt{rcu\_dereference(elm->next)}.
+Therefore, the first read of \texttt{elm->next} might yield a pointer to partially constructed data.
+Because that data is not guaranteed to be fully constructed, the next read of \texttt{elm->next} (i.e.
+\texttt{elm->next->next}) might read an invalid pointer from uninitialized memory.
+Treating the value of uninitialized memory as a valid pointer could trigger a kernel panic or silent data corruption.]{\fbox{\usebox\rcuassignderefbug}} %
 \caption{\label{fig:rcu_bugs}Example RCU bugs related to misuses of pointers to shared data.}
@@
 
-The \texttt{RCUdbg} tool detects the the following bugs: \begin{enumerate}
-	\item Reads through an RCU-dereferenced pointer outside of a read-side critical section (\Cref{fig:rcu_existence_bug}).
-	\item Reads through an RCU-assigned pointer that don't first use \texttt{rcu\_dereference} (\Cref{fig:rcu_deref_bug}).
-	\item Reads through fields of RCU-protected data structures that should be accessed via \texttt{rcu\_dereference} (\Cref{fig:rcu_assign_bug}).
-	\item Reads through an RCU-dereferenced pointer in the wrong read-side critical section.
-	\item Writes through an RCU-dereferenced pointer.
-	\item Writes through an RCU-assigned pointer.
-\end{enumerate}
+\texttt{RCUdbg} is primarily concerned with finding bugs related to RCU-rpotected pointer misuses by reader threads (both inside and outside of read-side critical sections).
+As mentioned above, RCU bugs sometimes manifest as use-after-free bugs.
+However, a pure use-after-free-based checking approach is insufficient (\Cref{fig:rcu_assign_bug} is one such example).
+\Cref{fig:rcu_existence_bug} is an example of a \emph{potential} use-after-free.
+That is, the example code might be interrupted after \texttt{rcu\_read\_unlock}, but before the access to \texttt{elm->data}.
+If \texttt{elm} is garbage collected and freed before the reader is re-scheduled then this code can result in a use-after-free bug.
+However, being unlucky enough to observe this thread schedule is unlikely.
+Fortunately, \texttt{RCUdbg} catches the bug in \Cref{fig:rcu_existence_bug} right when \texttt{elm->data} is accessed.
+
+%\Cref{fig:rcu_bugs} shows some example usages of RCU, and some of the more subtle pointer-related bugs that can happen because of the interaction of rules (1), (2), (3), and (7).
+
+The \texttt{RCUdbg} tool detects the the following class of RCU bugs: \begin{itemize}
+	\item[B1)] Reads through an RCU-dereferenced pointer outside of a read-side critical section (\Cref{fig:rcu_existence_bug}).
+	\item[B2)] Reads through an RCU-dereferenced pointer in the wrong read-side critical section.
+
+	\item[B2)] Reads through an RCU-assigned pointer that don't first use \texttt{rcu\_dereference} (\Cref{fig:rcu_deref_bug}).
+	\item[B3)] Reads through fields of RCU-protected data structures that should be accessed via \texttt{rcu\_dereference} (\Cref{fig:rcu_assign_bug}).
+
+	\item[B5)] Writes through an RCU-dereferenced pointer.
+	\item[B6)] Writes through an RCU-assigned pointer by a non-write thread.
+\end{itemize}
+
+Bugs B1 and B2 relate to whether or not we can prove that a given RCU protected data structure exists.
+Bugs B2 and B3 relate to whether or not a given RCU protected data structure has been fully constructed.
+Finally, bugs B5 and B6 relate to invalid uses RCU-protected and RCU-unprotected pointers, where correct synchronization among writers is not maintained.
+
+% or is constructed at a given program point. Bugs B2 through B
+
+% of this class of RCU bugs is proof-of-existence or proof-of-constructedness bugs.
+That is, an RCU-protected data structue exists if it is \emph{reachable} (i.e.
+by at least one reader thread), and if it is \emph{fully constructed} (i.e.
+all fields or components of the data structure have been initialized).
 
 \texttt{RCUdbg} uses Granary's address watchpoints feature \cite{BehaveOrBeWatched} to ``taint" addresses participating in RCU-protected data structures.
+
+%Unfortunately, approaching RCU bug detection from the perspective of finding use-after-free errors is insufficient.
+Use-after-free detection attacks a symptom of the problem that is highly dependent on thread interleavings and scheduling.
+That is, RCU can be misused in a way that results in very unlikely, but potential use-after-free bugs.
+An example of this case is when a writer thread, operating on a RCU-protected linked list, unlinks an element and queues its memory for later freeing.
+RCU provides a mechanism whereby readers can safely access this element for a brief period of time (called a read-side critical section).
+When the memory is eventually freed depends is implementation-defined, but depends on other reader activity.
+
+%How soon after a read-side critical section ends
+
+% There are many cases readers can access an RCU-protected data structure because they have not been freed \emph{yet}
+
+%
+
+% a data structure cannot be proven to exist at some potentially allocated memory.
+Existence in this case relates to semantic reachability.
+For example, if a list element $E_n$ from an RCU-protected list is reachable from the head of the list by traversing zero-or-more intermediate elements $E_0, ..., E_{n-1}$, then $E_n$ is said to exist.
+While related, memory leak detection (e.g.
+via conservate mark and sweep) is also insufficient for ladidadidaaaa TODO TODO TODO
 

### 2013-11-13 VEE-14/paper.tex
Minor edition of the rcu section, about to get into implementation details

@@
 When correctly used, RCU permits concurrent reads and writes to shared data structures, but requires that writers synchronize among themselves.
-However, like other synchronization mechanisms, RCU has a deceptively simple API.
-Misuses of the RCU API often result in hard-to-find bugs that can be implementation-specific (i.e.
+Like other synchronization mechanisms, RCU has a deceptively simple API.
+Misuses of the API can result in hard-to-find bugs that can be implementation-specific (i.e.
 bugs that manifest in one version of RCU but not another) or architecture-specific (e.g.
@@
 The following is a non-exhaustive list of requirements that must be met by reader and writer threads operating on an RCU-protected data structure.
-The list omits a description of how RCU-protected data structures are garbage collected.
-Details concerning garbage collection are beyond the scope of the bugs caught by \texttt{RCUdbg}.
+The list omits a description of how RCU-protected data structures are garbage collected, which is beyond the scope of the bugs caught by \texttt{RCUdbg}.
 
@@
 Treating the value of uninitialized memory as a valid pointer could trigger a kernel panic or silent data corruption.]{\fbox{\usebox\rcuassignderefbug}} %
-\caption{\label{fig:rcu_bugs}Example RCU bugs related to misuses of pointers to shared data.}
+\caption{\label{fig:rcu_bugs}Example RCU bugs related to misuses of RCU-protected pointers.}
 \end{figure}
@@
 As mentioned above, RCU bugs sometimes manifest as use-after-free bugs.
-However, a pure use-after-free-based checking approach is insufficient (\Cref{fig:rcu_assign_bug} is one such example).
-\Cref{fig:rcu_existence_bug} is an example of a \emph{potential} use-after-free.
-That is, the example code might be interrupted after \texttt{rcu\_read\_unlock}, but before the access to \texttt{elm->data}.
-If \texttt{elm} is garbage collected and freed before the reader is re-scheduled then this code can result in a use-after-free bug.
-However, being unlucky enough to observe this thread schedule is unlikely.
+Unfortunately, a pure use-after-free-based checking approach is insufficient (\Cref{fig:rcu_assign_bug} is one such example).
+\Cref{fig:rcu_existence_bug} is an example of an RCU usage bug that can manifest as a use-after-free bug if one is unlucky enough to observe a very specific thread schedule.
+Suppose that execution is interrupted after \texttt{rcu\_read\_unlock}, but before the access to \texttt{elm->data}.
+If \texttt{elm} is garbage collected and freed before the code resumes its execution, then the access to \texttt{elm->data} is a use-after-free bug.
+This kind of bug is a time bomb waiting to go off.
 Fortunately, \texttt{RCUdbg} catches the bug in \Cref{fig:rcu_existence_bug} right when \texttt{elm->data} is accessed.
@@
 
-The \texttt{RCUdbg} tool detects the the following class of RCU bugs: \begin{itemize}
+\texttt{RCUdbg} detects the the following RCU bugs related to misuses of RCU-protected pointers.
+\begin{itemize}[leftmargin=3.2em]
 	\item[B1)] Reads through an RCU-dereferenced pointer outside of a read-side critical section (\Cref{fig:rcu_existence_bug}).

### 2013-11-13 VEE-14/paper.tex
Last updates

@@
 
-\subsection{Functions}
+\subsection{Functions}\label{sec:function_wrapper}
 
@@
 
-\texttt{RCUdbg} uses Granary's address watchpoints feature \cite{BehaveOrBeWatched} to ``taint" addresses participating in RCU-protected data structures.
+\texttt{RCUdbg} gains visibility on RCU operations in two ways: \begin{enumerate}
+	\item The Linux kernel RCU implementation is annotated to provide more information to \texttt{RCUdbg}.
+This information includes source code file and line information for bug reporting, and additional memory references that \texttt{RCUdbg} uses to get better visibility on bugs.
+These annotations are strictly internal to the RCU API, and thus require no changes to code using this API.
+	\item The annotated RCU API functions are specialized using Granary's function wrapping feature (\Cref{sec:function_wrapper}). Specifically, 
+\end{enumerate}
+
+%\texttt{RCUdbg} interposes on an annotations are added into the Linux kernel's RCU implementation. These annotations
+
+%uses Granary's address watchpoints feature \cite{BehaveOrBeWatched} to ``taint" addresses participating in RCU-protected data structures.
+
+%Function wrappers are used to specialize the execution of \texttt{rcu\_\linebreak[0]dereference} and \texttt{rcu\_assign\_\linebreak[0]pointer}:
+
+\begin{figure}[t!]
+\lstset{language=C, tabsize=2, stepnumber=1}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+// As used in the code:
+//   p = rcu_dereference(q);
+// Internally annotated to become:
+//   p = __rcudbg_deref(&q, q);
+// Wrapper:
+FUNCTION_WRAPPER(__rcudbg_deref, (q_addr, q), {
+  if(is_watched_address(q_addr)) {
+    // Make sure q_addr
+  } else {
+
+  }
+  return add_watchoint(q, DEREF);
+})
+\end{lstlisting}
+\caption{}
+\label{fig:rcu_dereference_wrapper}
+\end{figure}
+
+%\begin{center}
+%\begin{minipage}{0.6\columnwidth}
+%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+%elm = rcu_dereference(list_head);
+%\end{lstlisting}
+%\end{minipage}
+%\end{center}
+%
+%Becomes:
+%
+%\begin{center}
+%\begin{minipage}{0.8\columnwidth}
+%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+%elm = deref_taint(rcu_dereference(list_head));
+%\end{lstlisting}
+%\end{minipage}
+%\end{center}
+%
+%Tainted addresses are introduced in two ways:
+%\begin{description}
+%	\item[\texttt{rcu\_assign\_pointer}] operates on two pointers: 
+%	\item[\texttt{rcu\_dereference}] operates on a potentially tainted pointer (as introduced by \texttt{rcu\_assign\_\linebreak[0]pointer}, and returns a deref-tainted pointer.
+%\end{description}
+%
 

### 2013-11-25 VEE-14/paper.tex
Working on optimizations section

--- 2013-11-13 VEE-14/paper.tex
@@
 \usepackage{graphicx}
-\usepackage{multicol,multirow}
-\usepackage{tabularx}
+\usepackage{multicol}
 
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{Flexible Binary Instrumentation for the Kernel}
+\title{Fast and Flexible Binary Instrumentation for the Kernel}
 
 %for single author (just remove % characters)
-\authorinfo{Peter Goodman \and Akshay Kumar \and Angela Demke Brown \and Ashvin Goel}
+\authorinfo{Peter Goodman \and Angela Demke Brown \and Ashvin Goel}
 {University of Toronto}{}
@@
 In practice, we would like to target instrumentation at specific code for efficiency, while also allowing rich and detailed instrumentation of the targeted code.
-Limiting the scope of what is instrumented is  important because the overhead introduced by using runtime code instrumentation is usually dominated by the instrumentation itself and not the runtime system supporting that instrumentation.
-
-%This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary tools to instrument only specific code while the majority of code executes with zero overhead.
+This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary tools to instrument only specific code while the majority of code executes with zero overhead.
 
@@
 
+%is a synchronization mechanism that permits readers to operate concurrently with writers, but does not guarantee that readers observe the most up-to-date changes (by writers) to shared state.
+RCU readers
+
+%When used correctly, RCU readers must declare that they will operate on shared state by entering into a read-side critical section, delimited by \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock}, respectively.
+When inside a read-side critical section, a pointer to shared state \emph{must} be obtained by invoking \texttt{rcu\_dereference}.
+Thereafter, a reader can freely access the shared state through this pointer, but only up until the
+
 %In reader-side code
@@
 Part of building CFGs requires tracking whether or not a given basic block is the entrypoint to a function.
-This is challenging with existing DBT systems because some optimized functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs.
+This is challenging with existing DBT systems because some optimised functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs.
 Granary solves this problem using demand-based runtime code specialization: every entrypoint basic block is instrumented in one way, and later re-executions of any entrypoint basic blocks are specialized and instrumented in another way.
@@
 Granary uses dynamic binary translation (DBT) to instrument the Linux kernel.
-DBT is used for emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimization}, and runtime instrumentation (analysis \cite{DynamoRIO, DRK, btkernel, ProfilingSimics}, security \cite{Vx32,NaCl,ProgramShepherding}, and debugging \cite{Valgrind}).
+DBT is used for emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimisation}, and runtime instrumentation (analysis \cite{DynamoRIO, DRK, btkernel, ProfilingSimics}, security \cite{Vx32,NaCl,ProgramShepherding}, and debugging \cite{Valgrind}).
 We chose DBT because \begin{inparaenum}[i)]
@@
 Granary also generates edge code for indirect control-flow instructions; however, this edge code is split into two separate components: entry and exit edge code.
-Entry edge code is generated on a per-indirect CTI instruction basis, and spills registers onto the runtim call stack.
+Entry edge code is generated on a per-indirect CTI instruction basis, and spills registers onto the runtime call stack.
 Entry edge is responsible for indirect branch lookup, and will indirectly jump to the exit edge code for a given basic block.
@@
 \begin{minipage}[b]{4cm}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]cmp $0x0, %rax
-
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+cmp $0x0, %rax
 jz value_is_zero 
 ...
-
 ( fall through )
@@
 \newsavebox\translatedjcc
-\begin{lrbox}{\translatedjcc}\begin{minipage}[b]{4cm}\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]  cmp $0x0, %rax
-jz_to_patch:
-  jmp edge_value_is_zero
+\begin{lrbox}{\translatedjcc}\begin{minipage}[b]{4cm}\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+cmp $0x0, %rax
+edge_jz:
+  jz edge_value_is_zero
   ...
 edge_value_is_zero:
-  ( spill 2 registers )
-  ( save value_is_zero )
-  ( save target policy )
-  call patcher_for_jz
-  ( restore 2 registers )
-  jmp .jz_to_patch
+  call (patch_edge_jz)
 \end{lstlisting}
@@
 \caption{\label{fig:direct_edge_code}Example translation of a conditional control flow instruction (\texttt{jz}) and its edge code.
-The edge code saves the intended target of the jump (\texttt{value\_is\_zero}) and the instrumentation policy to apply when translating the targeted code, and then context-switches into Granary by invoking \texttt{patcher\_for\_jz}.
-The \texttt{patcher\_for\_jz} function atomically replaces the {\bf\texttt{ jmp}}\texttt{ edge\_value\_is\_zero} instruction with a fully resolved \texttt{jz} instruction, and upon return, executes the newly resolved \texttt{jz} instruction.}
+The original \texttt{jz} is translated into a \texttt{jz} that uses a 32-bit relative displacement to transfer control to an edge code function call (\texttt{edge\_jz}).
+When executed, the function call reads its target from a data structure (\texttt{patch\_edge\_jz}), which contains all information necessary to resolve the \texttt{jz}'s target and patch the original \texttt{jz}.
+Patching is performed with a four byte atomic write to the relative displacement operand of the \texttt{jz} instruction.
+Finally, the return address of the \texttt{call} instruction is altered to transfer control to the newly resolved target.}
 \end{figure}
@@
 When necessary, \texttt{RCUdbg} spills and restores registers around each checked load/store operation, so that the registers can be safely used by the checking code.
-This is an example of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after.
+This is an examuple of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after.
 DRK is oblivious to the relationship between pre/post instrumentation, and so an interrupt that occurs within the pre-instrumentation will only be delayed until the native instruction.
 
-\section{Specializing on What to Instrument}
+\subsection{Optimizations}
+\paragraph{Ahead-of-time Translation} 
+Granary ends basic blocks at conditional branches (\texttt{jCC}), direct and indirect \texttt{jmp}s, and \texttt{ret}s and \texttt{iret}s.
+Basic blocks ending in conditional branches are extended to contain two trailing CTIs: a copy of the native \texttt{jCC}, and a synthesized direct \texttt{jmp} to the code executed when the condition code is not set (i.e.
+when the conditional jump is not taken).
+
+If a basic block ends in either a copy of a native direct \texttt{jmp} or synthesized direct \texttt{jmp}, then Granary will perform ahead-of-time translation and instrumentation of the targeted basic block.
+The repeated application of translating \texttt{jmp} targets forms traces in Granary's code cache.
+Granary optimizes traces by directly connecting basic blocks together when there is internal control flow within the blocks of the trace, and by eliding synthesized \texttt{jmp}s between adjacent basic blocks.
+
+Granary's tracing strategy is based on the observation that the most commonly executed path through code follows a straight line.
+That is, execution is more likely to fall through a conditional CTI than it is to take the conditional CTI.
+We experimentally verified Granary's tracing strategy by disabling tracing and relying on Granary's performance counters to count how many conditional branches were patched (by their edge code), versus how many synthesized fall-through \texttt{jmp}s were patched.
+
+%using internal performance counters within Granary
+
+%Static branch prediction hints (\texttt{likely}, \texttt{unlikely} macros) are heavily used in the Linux kernel source code.
+Although static branch prediction hints can be specified on individual conditional branch instructions, recent x86-based architectures ignore these hints \cite{AgnerMicroarchitecture}.
+Therefore, static branch prediction hints at the source code level mainly affect compiler optimizations, including intraprocedural basic block scheduling.
+Our tracing strategy is based on
+
+%Tr
+
+%We expect and have experimentally verified that the most common execution path through code is straight-line and does not following conditional branches.
+
+%Due to instruction prefetching, we expect that the likely or
+
+%On recent x86-based architectures, these hints only affect the schedule of basic blocks in the intraprocedural control-flow graph \cite{AgnerMicroarchitecture}.
+
+\paragraph{Patching and Alignment}
+
+For efficiency, basic blocks in Granary's code cache are directly connected through direct control flow instructions (\texttt{call}, \texttt{jmp}, \texttt{jCC}).
+That is, if there is a direct branch from one native instruction to another, then there will be a corresponding direct branch between two translated instructions in Granary's code cache.
+
+Granary opportunistically connects two basic blocks together using a direct branch if \begin{inparaenum}[i)]
+	\item the branch transfers control to another basic block within the current trace; or
+	\item the branch transfers control to an already translated basic block.
+\end{inparaenum} However, not all basic blocks can be connected at translation time, therefore Granary employs edge code and hot-code patching (\Cref{fig:direct_edge_code}) to connect basic blocks at runtime.
+
+Granary atomically patches direct CTIs that target edge code to instead point to their (now resolved) target basic block.
+Due to limitations of the x86 architecture, atomic operations cannot straddle cache lines.
+Therefore, patchable instructions (or parts thereof) must be aligned so as to not straddle two cache lines.
+
+The method used to pad basic blocks so as to align specific instructions must be transparent (i.e.
+not change the execution behavior).
+The typical approach to alignment is to introduce \texttt{nop} instructions, or \texttt{jmp} instructions which jump over the padding.
+Adding extra instructions, however, introduces added latency to the fetch/decode/execute cycle.
+Therefore, when possible, Granary pads instructions using otherwise ignored instruction prefixes.
+For example, Granary aligned patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86 processors \cite{AgnerMicroarchitecture}).
+
+%To reduce the overhead of alignment
+
+%In some cases, basic blocks can be directly connected during basic block translation time.
+For example, intra-trace CTIs, or CTIs targeting already translated basic blocks.
+However, not all basic blocks can be connected this way, therefore some runtime
+
+%Granary connects some basic blocks at translation time (e.g.
+intra-trace control-flow instructions).
+This is common when tracing is enabled, but also occurs when a branch targets an already translated basic block.
+However,
+
+%Some basic blocks are directly connected at basic block translation time.
+One example is branches that target other basic blocks within a trace.
+Another example is a branch to an already translated basic block.
+
+%Other basic blocks are only connected at runtime.
+For example, if the function \texttt{memset} has never been executed, then a \texttt{call} to \texttt{memset} will be translated to target edge code (\Cref{fig:direct_edge_code}) that, when executed, will translate \texttt{memset}'s first basic block and patch the \texttt{call} to target the translated basic block.
+
+%For example, the targets of function call instruction (\texttt{call}) are resolved just-in-time (unless the target is already known).
+For each translated \texttt{call} instruction, Granary generates edge code (\Cref{fig:direct_edge_code}) that will find or translate the first basic block of the function being called, and patch the translated \texttt{call} instruction to point to the function's first basic block.
+
+%For example, if a direct branch in one basic block targets another block within the trace under construction, then that
+
+%Basic blocks are directly connected at block translation time when branches in traces re-enter a trace, or when branches target previously translated basic blocks.
+If a branch does not target another basic block within its trace, or does not target a previously translated basic block, then the branch will target edge code that will resolve that branches target at runtime.
+
+% However, when a given basic block is translated, not all of its targets are known (e.g.
+\texttt{call} and \texttt{jCC} targets).
+Therefore, edge code (\Cref{fig:direct_edge_code}) is generated that translates/resolves a branch target when that branch is first executed, and patches the translated branch to avoi
+
+% The targets of some of these direct CTIs are only resolved at runtime, by executing edge code.
+As an optimization, jumps to edge code are dynamically patched to transfer control to their intended targeted basic blocks.
+
+%edge code use hot code patching to update a translated CTI to redirect execution
+
+\paragraph{Trace Allocation}
+
+\paragraph{Indirect Control Flow}
+
+%\subsection{Reproducability}
+%Hello
+
+\section{Choosing What to Instrument}
 
@@
 
-\section{Specializing on How to Instrument}\label{sec:how}
+\section{Context-Driven Instrumentation}\label{sec:policies}
 
-Granary gives tools the ability to be selective about what is instrumented.
-A natural extension of this selectivity is to allow tools to specialize how code is instrumented by dynamically switching the policy used to instrument code.
-Specialization is desirable in several cases.
-For example, some instrumentation is high-overhead, but does not need to be applied to all code.
-One example of this is Granary's \texttt{RCUdbg} tool, which instruments read-side and write-side critical sections more heavily than the rest of code.
-Unfortunately, existing DBT systems fall short in their support for runtime code specialization \cite{DRK,btkernel,Pin,DynamoRIO}.
+A natural extension of mixed-mode execution is to give Granary tools the ability to dynamically switch the policy used to instrument code.
+Instrumenting only module code allows Granary to specialize the execution of native code.
+In the same way, policy switching allows tools to specialize the execution of instrumented code.
+In fact, optimising the performance of an early Granary tool was the original motivation for policy switching.
 
-To implement the equivalent of \texttt{RCUdbg}'s runtime code specialization using an existing DBT system requires that a tool developer maintain, update, and check runtime state to know whether or not code is executing inside or outside of a read-side critical section.
-If a basic block is executing in a read-side critical section, then instrumentation should be enabled, and if it's not in a read-side critical section then it should be disabled.
-Enabling/disabling instrumentation might be done at the instruction granularity, which introduces at least one branch instruction per native instruction, or at the granularity of basic blocks, which requires generating two copies of each native basic block, where a single branch at the beginning of a basic block chooses between the instrumented or uninstrumented version.
-Up until now, it has been assumed that there is a mechanism for maintaing state that consistently tracks the current execution context of kernel code (inside vs.
-outside of a read-side critical section).
-However, maintaining state in a way that is reentrant and resilient again arbitrary pre-emption and resumption is onerous.
-
-%Existing systems move the burden of code specialization to instrumentation tool developers.
-
-Fortunately for tool developers, Granary has built-in support for runtime code specialization that does not require manual state management or cause code cache bloat.
-Granary approach to runtime code specialization uses a technique called \emph{context-driven instrumentation}.
-
-%For example, instrumenting only module code allows Granary to specialize the execution of native code.
-In the same way, policy switching allows Granary's \texttt{RCUdbg} tool to specialize the execution of instrumented code executing inside of read-side and write-side critical sections.
-In fact, optimizing the performance of an early Granary tool was the original motivation for policy switching.
-
-\subsection{Context-Driven Instrumentation}\label{sec:policies}
-
-%We developed a Granary tool that detects several Read-Copy-Update (RCU) API misuses in module code.
+We developed a Granary tool that detects several Read-Copy-Update (RCU) API misuses in module code.
 Our tool focused on read-side critical sections (delimited by calls to \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock} in the code), however, most code executes outside of read-side critical sections.
-As an optimization, we wanted our heavyweight API-checking instrumentation to apply only when the code was executing within a read-side critical section.
-Implementing this optimization was challenging because each basic block was only translated once, and we had no way to know whether or not the heavyweight instrumentation should be applied to it.
+As an optimisation, we wanted our heavyweight API-checking instrumentation to apply only when the code was executing within a read-side critical section.
+Implementing this optimisation was challenging because each basic block was only translated once, and we had no way to know whether or not the heavyweight instrumentation should be applied to it.
 Knowing the state (within or outside a read-side critical section) at translation time was not sufficient, since the same translated block could later be executed in the opposite context.
@@
 
-%To implement this optimization,
-
-Granary enables specializing instrumentation to the context in which the instrumented code will execute.
-Specialization is achieved by allowing multiple, differently instrumented versions of the same code to co-exist within Granary's code cache.
-That is, if the next basic block to execute has not yet been translated/instrumented for the current execution context, then a new version of that basic block will be created  specifically for the the current execution context.
-
-%In the case of our \texttt{RCUdbg}, there were three execution contexts: within and outside of a read-side critical section.
+To implement this optimisation, Granary enables tailoring instrumentation to the context in which the instrumented code will execute.
+To do so, Granary allows different versions of the same module code to co-exist within Granary's code cache.
+That is, a different version of each of a module's basic blocks exists in Granary's code cache  for each encountered execution context.
+In the case of our RCU checker tool, there were two execution contexts: within and outside of a read-side critical section.
 If the same module basic block is  executed in the two contexts, then Granary's code cache would contain two different instrumented versions of that basic block.
@@
 
-Granary tools identify execution contexts by creating instrumentation policies.
 An instrumentation policy is both a name for an execution context, as well as a function that decides how to instrument basic blocks that will execute within that context.
@@
 Tools are not limited to one policy though: any policy can declare a policy switch that will take effect when a selected control-transferring instruction (CTI) is executed.
-The effect of a tool specifying a policy switch on a CTI is that the basic block(s) targeted by that CTI will be instrumented according to the specified policy.
-CTIs with unspecified policies inherit their policies from their containing basic blocks.
-
-For example, our \texttt{RCUdbg} invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.
-A similar switch occurs from $P_{\mathit{read\_critical}}$ back to $P_{\mathit{null}}$ at a \texttt{rcu\_read\_unlock}.
-Specialization that distinguishes between the inside and outside of read-side critical sections allows \texttt{RCUdbg} to do heavy-weight checks on functions like \texttt{memset}, but only when \texttt{memset} is invoked from within a read-side critical section.
+For example, our RCU tool invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.
 
 Because policies name an execution context, they also represent states in a finite state machine.
-That is, code cache execution is in the state named by a policy if the  executing code was instrumented by that policy.
+That is, a module's execution is in the state named by a policy if the code executing was instrumented by that policy.
 A state transition occurs when control transfers from code instrumented by one policy to code instrumented by another policy.
@@
 If two read-side critical sections are nested then switching from $P_{\mathit{read\_critical}}$ to $P_{\mathit{null}}$ on the first \texttt{call} to \texttt{rcu\_read\_unlock} meant that our tool would lose track of being in the context of the outer read-side critical section.
-\texttt{RCUdbg} solves this problem by tracking the nesting depth of RCU read-side critical sections using policies, e.g.
-$P_{\mathit{read\_critical}_1}$, $P_{\mathit{read\_critical}_2}$, etc.
+We solved this limitation by observing that, in most cases, function returns are natural policy reverting points.
+This is because the runtime call stack encodes both the context in which the current code is executing, as well as the continuation of the current executing code.
 
-An alternative and sometimes more powerful model for policy switching is to prevent policy switching on function return instructions (e.g.
-either through manually specifying a policy switch, or by inheriting the policy from the basic block containing the \texttt{ret} instruction).
-In most cases, function returns are natural policy reverting points.
-This is because the current instruction pointer (program counter) encodes the current execution context (by being in a basic block instrumented by some policy), and the code cache return addresses stored on the runtime call stack encode all previous execution contexts.
-This asymmetry between \texttt{call}s and \texttt{ret}s yields a more powerful context-tracking state machine: \texttt{call}s place contextual breadcrumbs (in the form of return addresses) on the runtime call stack, and \texttt{ret}s read these breadcrumbs to return to a previous context.
-Policy switches under this lens behave similarly to state transitions in a pushdown automaton; \texttt{call} instructions push a new state onto the stack for each function, \texttt{ret} instructions pop the current function's state from the stack, and other CTIs induce state transitions within the current function by altering control flow.
+The effect of a tool specifying a policy switch on a CTI is that the basic block(s) targeted by that CTI will be instrumented according to the specified policy.
+CTIs with unspecified policies inherit their policies from their containing basic blocks.
+As hinted at above, \texttt{ret} instructions cannot explicitly switch policies, which allows a function instrumented by a different policy than its caller to return to the caller's context (i.e., policy).
+The asymmetry between \texttt{call}s and \texttt{ret}s is intentional: \texttt{call}s place contextual breadcrumbs (in the form of return addresses) on the runtime call stack, and \texttt{ret}s read these breadcrumbs to return to a previous context.
+Under this lens, the instruction pointer tracks our current policy, and return addresses form a stack of previous policies.
+Policy switches behave similarly to state transitions in a pushdown automaton; \texttt{call} instructions push a new state onto the stack for each function, \texttt{ret} instructions pop the current function's state from the stack, and other CTIs induce state transitions within the current function by altering control flow.
 
 Granary implements policy tracking and switching by encoding policy information into the meta-data and CTIs of basic blocks.
-\Cref{fig:direct_edge_code} shows an example of how policy information is directly recorded into control-transfer edge code.
 If a policy switch is not specified on an instrumented CTI then that CTI inherits the policy used to instrument the basic block containing the CTI.
@@
 Granary decodes and instruments the targeted instructions according to the input policy information.
-Because Granary's translation mechanism depends only on CTI policy and target information, Granary is able to ensure that policy information is never lost or corrupted, even in the face of concurrent executions of the same module code, arbitrary pre-emption, and arbitrary resumption.
+Because every non-\texttt{ret} CTI encodes policy information, and because Granary's translation mechanism depends only on CTI policy and target information, Granary is able to ensure that policy information is never lost or corrupted, even in the face of concurrent executions of the same module code, arbitrary pre-emption, and arbitrary resumption.
 
-\section{Application: RCUdbg}
-
-Read-Copy-Update (RCU) is a scalable, lock-free synchronization mechanism used in the Linux kernel \cite{RCU,RCUInLinux}.
-When correctly used, RCU permits concurrent reads and writes to shared data structures, but requires that writers synchronize among themselves.
-However, like other synchronization mechanisms, RCU has a deceptively simple API.
-Misuses of the RCU API often result in hard-to-find bugs that can be implementation-specific (i.e.
-bugs that manifest in one version of RCU but not another) or architecture-specific (e.g.
-bugs that only exhibit on architectures with a relaxed memory model).
-For example, RCU bugs sometimes manifest as use-after-free bugs, which can result in silent data corruption (if the underlying memory remains used) or kernel panics.
-
-% (assuming that pointer reachability is sufficient to determine semantic reachability in this case).
-
-Granary's \texttt{RCUdbg} tool is designed to catch RCU bugs related to kernel code misusing pointers to RCU-protected data structures.
-The following is a non-exhaustive list of requirements that must be met by reader and writer threads operating on an RCU-protected data structure.
-The list omits a description of how RCU-protected data structures are garbage collected.
-Details concerning garbage collection are beyond the scope of the bugs caught by \texttt{RCUdbg}.
-
-%the associated memory may or may not be free, but the data structure occupying that memory can't be proven to exist. This
-
-%  However, this class of RCU bugs is not well-suited to being detected by a use-after-free memory checker tool
-%RCU is notoriously hard to use, despite its relatively simple API.
-RCU usage bugs, like bugs with other synchronization mechanisms, often stay hidden and only exhibit themselves for very specific thread interleavings/schedules.
-
-% go silently unnoticed because the memory
-
-%While bugs related to mutual exlusion locks might result in easily observable deadlocks, bugs related to RCU typically result in use-after-free errors, which sometimes go silently unnoticed, sometimes result in kernel panics, and most of the time work as expected because the associated data is .
-
-\begin{enumerate}
-	\item Readers must tolerate an inconsistent view of the shared data structure.
-That is, a reader operating concurrently with a writer might observe either the version of the data before or after the writer's changes have been published.
-This can be complicated when writers perform multiple, discrete updates to a data structure.
-One example of this is a doubly linked list: a writer that is adding a new entry into a doubly linked list must update at least two RCU-protected pointers.
-
-	\item Readers must operate within \emph{read-side} critical sections, which are delimited by \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock}, respectively.
-
-	\item Within read-side critical sections, readers can only read a shared data structure after obtaining a pointer to that shared data structure using the \texttt{rcu\_dereference} API function.
-An RCU-protected pointer is like a normal pointer, but points to a potentially incomplete or partially-constructed data structure.
-\texttt{rcu\_dereference} operates on an RCU-protected pointer and returns an ``unprotected" version of the pointer.
-The reader can use the unprotected pointer to safely access the shared data structure within the remainder of the its read-side critical section.
-If \texttt{rcu\_dereference} is not used, then the reader might observe a partially-constructed (i.e.
-incomplete) version of the shared data structure.
-Notably, accessing the original data structure and not the dereferenced data structure pointer--even after an \texttt{rcu\_dereference}--is undefined behavior.
-That is, within a read-side critical section, RCU only guarantees the ``completeness" of data structures pointed to by RCU-dereferenced pointers.
-Data accessed within a read-side critical section is \emph{not} guaranteed to exist outside the read-side critical section.
-
-	\item Writers must synchronize among themselves before any writes to a shared data structure are performed.
-Synchronization is typically achieved via mutual exclusion locks, spin locks, or by having a single, designated writer thread.
-We say writes to the shared data structure occur within \emph{write-side} critical sections.
-
-	\item RCU only defines how writers can update RCU-protected pointer within a shared data structure.
-Modification of non-RCU-protected pointer fields within an RCU-protected data structure is beyond the scope of RCU's correctness guarantees.
-
-	\item Writers wishing to update an RCU-protected pointer in a shared data structure must use the \texttt{rcu\_assign\_pointer} API function.
-\texttt{rcu\_assign\_pointer} assigns a new value to an RCU-protected pointer in a shared data structure.
-
-	\item Readers must access any pointer field operated on by an \texttt{rcu\_\linebreak[0]assign\_pointer} using \texttt{rcu\_\linebreak[0]dereference}.
-For example, if the \texttt{next} pointer field in \emph{any} list element of an RCU-protected list has been assigned to by \texttt{rcu\_assign\_\linebreak[0]pointer}, then \emph{all} \texttt{next} pointer fields must be accessed by readers of that list by using \texttt{rcu\_\linebreak[0]dereference}.
-In this sense, using \texttt{rcu\_\linebreak[0]dereference} and \texttt{rcu\_assign\_\linebreak[0]pointer} marks specific variables and data type fields as containing RCU-protected pointers.
-
-\end{enumerate}
-
-% NOTE!!!!
+%\section{Reifying Instrumentation}\label{sec:reify}
 %
-% All of those spaces before /* BUG! */ and such are so that the lstlisting will be *close* to the columnwidth!! Somehow linewidth param of lstset didn't work for me :-(
-
-\lstset{
-	language=C,
-	tabsize=2,
-	stepnumber=1,
-	morekeywords={rcu_read_lock,rcu_dereference,rcu_read_unlock,rcu_assign_pointer,spin_lock,spin_unlock}}
-
-% Example code showing an existence bug.
-\newsavebox\rcuexistencebug
-\begin{lrbox}{\rcuexistencebug}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
-rcu_read_lock();
-  elm = rcu_dereference(list_head);
-  data = elm->data;                  /* Valid */
-rcu_read_unlock();
-data = elm->data;                     /* BUG! */
-\end{lstlisting}
-\end{lrbox}
-
-% Example code showing another existence bug.
-\newsavebox\rcuderefbug
-\begin{lrbox}{\rcuderefbug}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
-rcu_read_lock();
-  elm = rcu_dereference(list_head);
-  if(elm != NULL)
-    data = list_head->data;           /* BUG! */
-rcu_read_unlock();
-\end{lstlisting}
-\end{lrbox}
-
-% Example code showing another existence bug.
-\newsavebox\rcuassignderefbug
-\begin{lrbox}{\rcuassignderefbug}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
-spin_lock();
-  elm = kmalloc(...);
-  elm->next = list_head->next;
-  rcu_assign_pointer(list_head->next, elm);
-spin_unlock();
-...
-rcu_read_lock();
-  elm = rcu_dereference(list_head);
-  while(elm != NULL) {
-    data += elm->data;
-    elm = elm->next;                  /* BUG! */
-  }
-rcu_read_unlock();
-\end{lstlisting}
-\end{lrbox}
-
-\begin{figure}[ht!] %
-\subfloat[\label{fig:rcu_existence_bug}Despite correctly obtaining a pointer to a shared data structure using \texttt{rcu\_dereference} within a read-side critical section, properties (2) and (3) tell us that the data pointed to by \texttt{elm} is not guaranteed to exist after the read-side critical section ends.]{\fbox{\usebox\rcuexistencebug}} %
-
-\subfloat[\label{fig:rcu_deref_bug}Despite the call to \texttt{rcu\_dereference}, properties (1) and (3) tell us that \texttt{list\_head} is not guaranteed to be equivalent to \texttt{elm} because a writer might have concurrently changed the value of the \texttt{list\_head} pointer.
-That is, if \texttt{elm} is non-\texttt{NULL}, then the data pointed to by \texttt{elm} is guaranteed to exist for the remainder of the read-side critical section.
-However, the value of \texttt{list\_head} is \emph{not} guaranteed to remain consistent--even within read-side critical sections--and so \texttt{list\_head} might point to a \texttt{NULL} pointer when \texttt{elm} does not.]{\fbox{\usebox\rcuderefbug}} %
-
-\subfloat[\label{fig:rcu_assign_bug}This example shows the code for a writer thread adding a new element into the second position of an RCU-protected list.
-The code for reader threads is also shown, where readers iterate over the list by traversing \texttt{next} pointers and sum up the \texttt{data} fields of each visited list element.
-The access of \texttt{elm->next} violates of properties (3) and (7).
-That is, the next element of the list is only guaranteed to be fully constructed within the read-side critical section if the reader executes \texttt{rcu\_dereference(elm->next)}.
-Therefore, the first read of \texttt{elm->next} might yield a pointer to partially constructed data.
-Because that data is not guaranteed to be fully constructed, the next read of \texttt{elm->next} (i.e.
-\texttt{elm->next->next}) might read an invalid pointer from uninitialized memory.
-Treating the value of uninitialized memory as a valid pointer could trigger a kernel panic or silent data corruption.]{\fbox{\usebox\rcuassignderefbug}} %
-\caption{\label{fig:rcu_bugs}Example RCU bugs related to misuses of pointers to shared data.}
-\end{figure}
-
-\texttt{RCUdbg} is primarily concerned with finding bugs related to RCU-rpotected pointer misuses by reader threads (both inside and outside of read-side critical sections).
-As mentioned above, RCU bugs sometimes manifest as use-after-free bugs.
-However, a pure use-after-free-based checking approach is insufficient (\Cref{fig:rcu_assign_bug} is one such example).
-\Cref{fig:rcu_existence_bug} is an example of a \emph{potential} use-after-free.
-That is, the example code might be interrupted after \texttt{rcu\_read\_unlock}, but before the access to \texttt{elm->data}.
-If \texttt{elm} is garbage collected and freed before the reader is re-scheduled then this code can result in a use-after-free bug.
-However, being unlucky enough to observe this thread schedule is unlikely.
-Fortunately, \texttt{RCUdbg} catches the bug in \Cref{fig:rcu_existence_bug} right when \texttt{elm->data} is accessed.
-
-%\Cref{fig:rcu_bugs} shows some example usages of RCU, and some of the more subtle pointer-related bugs that can happen because of the interaction of rules (1), (2), (3), and (7).
-
-The \texttt{RCUdbg} tool detects the the following class of RCU bugs: \begin{itemize}
-	\item[B1)] Reads through an RCU-dereferenced pointer outside of a read-side critical section (\Cref{fig:rcu_existence_bug}).
-	\item[B2)] Reads through an RCU-dereferenced pointer in the wrong read-side critical section.
-
-	\item[B2)] Reads through an RCU-assigned pointer that don't first use \texttt{rcu\_dereference} (\Cref{fig:rcu_deref_bug}).
-	\item[B3)] Reads through fields of RCU-protected data structures that should be accessed via \texttt{rcu\_dereference} (\Cref{fig:rcu_assign_bug}).
-
-	\item[B5)] Writes through an RCU-dereferenced pointer.
-	\item[B6)] Writes through an RCU-assigned pointer by a non-write thread.
-\end{itemize}
-
-Bugs B1 and B2 relate to whether or not we can prove that a given RCU protected data structure exists.
-Bugs B2 and B3 relate to whether or not a given RCU protected data structure has been fully constructed.
-Finally, bugs B5 and B6 relate to invalid uses RCU-protected and RCU-unprotected pointers, where correct synchronization among writers is not maintained.
-
-% or is constructed at a given program point. Bugs B2 through B
-
-% of this class of RCU bugs is proof-of-existence or proof-of-constructedness bugs.
-That is, an RCU-protected data structue exists if it is \emph{reachable} (i.e.
-by at least one reader thread), and if it is \emph{fully constructed} (i.e.
-all fields or components of the data structure have been initialized).
-
-\texttt{RCUdbg} uses Granary's address watchpoints feature \cite{BehaveOrBeWatched} to ``taint" addresses participating in RCU-protected data structures.
-
-%Unfortunately, approaching RCU bug detection from the perspective of finding use-after-free errors is insufficient.
-Use-after-free detection attacks a symptom of the problem that is highly dependent on thread interleavings and scheduling.
-That is, RCU can be misused in a way that results in very unlikely, but potential use-after-free bugs.
-An example of this case is when a writer thread, operating on a RCU-protected linked list, unlinks an element and queues its memory for later freeing.
-RCU provides a mechanism whereby readers can safely access this element for a brief period of time (called a read-side critical section).
-When the memory is eventually freed depends is implementation-defined, but depends on other reader activity.
-
-%How soon after a read-side critical section ends
-
-% There are many cases readers can access an RCU-protected data structure because they have not been freed \emph{yet}
-
-%
-
-% a data structure cannot be proven to exist at some potentially allocated memory.
-Existence in this case relates to semantic reachability.
-For example, if a list element $E_n$ from an RCU-protected list is reachable from the head of the list by traversing zero-or-more intermediate elements $E_0, ..., E_{n-1}$, then $E_n$ is said to exist.
-While related, memory leak detection (e.g.
-via conservate mark and sweep) is also insufficient for ladidadidaaaa TODO TODO TODO
+%Granary uses a technique we call \emph{reifying instrumentation} to provide the benefits of high-level static analysis information to dynamic instrumentation tools.
+Reifying instrumentation bootstraps on
 
@@
 
-What has been evaluated:
-
-\paragraph{lmbench}
-\begin{itemize}
-	\item Btkernel
-	\item Granary, full-kernel, null
-	\item Granary, full-kernel, even/odd
-	\item Granary, full-kernel, even/odd, where even/odd is a runtime check and branch.
-\end{itemize}
-
-\paragraph{fileserver}
-Setup: 1GB RAM, 3GB partition, ext3+jbd, >1GB mounted file system.
-\begin{itemize}
-	\item full-kernel
-	\begin{itemize}
-		\item Btkernel
-		\item Granary, null
-		\item Granary, even/odd
-		\item Granary, even/odd, where even/odd is a runtime check and branch.
-	\end{itemize}
-	\item module-only, (ext3+jbd)
-	\begin{itemize}
-		\item Granary, null
-		\item Granary, even/odd
-		\item Granary, even/odd, where even/odd is a runtime check and branch.
-	\end{itemize}
-\end{itemize}
-
-\paragraph{kernel compile}
-Same as fileserver, but with 8GB RAM instead of 3GB.
-
-\paragraph{recursive copy of kernel source tree}
-Same as fileserver, but with 8GB RAM instead of 3GB.
-
 \section{Related Work}\label{sec:related}
 Prior work can be classified as either whole-kernel/system DBT, or probe-based kernel instrumentation. We discuss how Granary differs from these approaches.
-\paragraph{Whole-kernel/system DBT}
+\vspace{-3pt}\paragraph{Whole-kernel/system DBT} DRK \cite{DRK} is a kernel space port of the DynamoRIO \cite{DynamoRIO} DBT framework.
+DRK instruments the entire kernel, including modules, and follows a strict transparency model, which limits the flexibility of instrumentation and increases overheads.
 PinOS \cite{PinOS} is a port of the Pin \cite{Pin} DBT framework that performs whole-system instrumentation.
@@
 
-QMEU is a flexible, whole-system emulator with kernel-level paravirtualization support (using KVM) \cite{QEMU}. Like PinOS, QEMU cannot instrument non-virtualizable device drivers.
-
-DRK \cite{DRK} is a kernel space port of the DynamoRIO \cite{DynamoRIO} DBT framework.
-DRK instruments the entire kernel and all device drivers/modules.
-DRK follows a strict transparency model, which limits the flexibility of instrumentation and increases overheads.
-
-Btkernel is a fast and scalable system for the kernel \cite{btkernel}.
-Like Granary, but unlike DRK, btkernel applies a relaxed transparency model when instrumenting the kernel.
-For exmaple, code cache addresses are exposed to the instrumented kernel in the form of function call return addresses and interrupt return addresses.
-Unlike Granary, btkernel operates using an instruction translation ``rulebook", which limits instrumentation to being defined ahead-of-time, on a per-instruction basis.
-Btkernel does not support any form of runtime code specialization.
-That is, like DRK, btkernel instruments all code, but cannot change the granularity of instrumentation or the instrumentation policy applied to code.
-
-TODO: JIFL
-
-\paragraph{Probe-based Instrumentation} Several systems support injecting code at specific locations within kernel code.
-
-TODO: SystemTap
-
-KernInst \cite{KernInst} can inject code at almost any location in an unmodified commodity OS.
-By default, KernInst uses debugging information normally present in kernel binaries to specify probe-points; however, absolute memory addresses can also be provided.
-
+\vspace{-3pt}\paragraph{Probe-based Instrumentation} Several systems support injecting code at specific locations within kernel code.
+For example, KernInst \cite{KernInst} can inject code at almost any location in an unmodified commodity OS.
 Other examples with similar or more restricted functionality include LTTng \cite{LTTng}, KProbes \cite{KProbes}, DProbes \cite{DProbes}, and ftrace \cite{ftrace}.

### 2013-11-25 VEE-14/paper.tex

--- 2013-11-13 VEE-14/paper.tex
@@
 When necessary, \texttt{RCUdbg} spills and restores registers around each checked load/store operation, so that the registers can be safely used by the checking code.
-This is an example of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after.
+This is an examuple of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after.
 DRK is oblivious to the relationship between pre/post instrumentation, and so an interrupt that occurs within the pre-instrumentation will only be delayed until the native instruction.
+
+\subsection{Optimizations}
+\paragraph{Ahead-of-time Translation} 
+Granary ends basic blocks at conditional branches (\texttt{jCC}), direct and indirect \texttt{jmp}s, and \texttt{ret}s and \texttt{iret}s.
+Basic blocks ending in conditional branches are extended to contain two trailing CTIs: a copy of the native \texttt{jCC}, and a synthesized direct \texttt{jmp} to the code executed when the condition code is not set (i.e.
+when the conditional jump is not taken).
+
+If a basic block ends in either a copy of a native direct \texttt{jmp} or synthesized direct \texttt{jmp}, then Granary will perform ahead-of-time translation and instrumentation of the targeted basic block.
+The repeated application of translating \texttt{jmp} targets forms traces in Granary's code cache.
+Granary optimizes traces by directly connecting basic blocks together when there is internal control flow within the blocks of the trace, and by eliding synthesized \texttt{jmp}s between adjacent basic blocks.
+
+Granary's tracing strategy is based on the observation that the most commonly executed path through code follows a straight line.
+That is, execution is more likely to fall through a conditional CTI than it is to take the conditional CTI.
+We experimentally verified Granary's tracing strategy by disabling tracing and relying on Granary's performance counters to count how many conditional branches were patched (by their edge code), versus how many synthesized fall-through \texttt{jmp}s were patched.
+
+%using internal performance counters within Granary
+
+%Static branch prediction hints (\texttt{likely}, \texttt{unlikely} macros) are heavily used in the Linux kernel source code.
+Although static branch prediction hints can be specified on individual conditional branch instructions, recent x86-based architectures ignore these hints \cite{AgnerMicroarchitecture}.
+Therefore, static branch prediction hints at the source code level mainly affect compiler optimizations, including intraprocedural basic block scheduling.
+Our tracing strategy is based on
+
+%Tr
+
+%We expect and have experimentally verified that the most common execution path through code is straight-line and does not following conditional branches.
+
+%Due to instruction prefetching, we expect that the likely or
+
+%On recent x86-based architectures, these hints only affect the schedule of basic blocks in the intraprocedural control-flow graph \cite{AgnerMicroarchitecture}.
+
+\paragraph{Patching and Alignment}
+
+For efficiency, basic blocks in Granary's code cache are directly connected through direct control flow instructions (\texttt{call}, \texttt{jmp}, \texttt{jCC}).
+That is, if there is a direct branch from one native instruction to another, then there will be a corresponding direct branch between two translated instructions in Granary's code cache.
+
+Granary opportunistically connects two basic blocks together using a direct branch if \begin{inparaenum}[i)]
+	\item the branch transfers control to another basic block within the current trace; or
+	\item the branch transfers control to an already translated basic block.
+\end{inparaenum} However, not all basic blocks can be connected at translation time, therefore Granary employs edge code and hot-code patching (\Cref{fig:direct_edge_code}) to connect basic blocks at runtime.
+
+Granary atomically patches direct CTIs that target edge code to instead point to their (now resolved) target basic block.
+Due to limitations of the x86 architecture, atomic operations cannot straddle cache lines.
+Therefore, patchable instructions (or parts thereof) must be aligned so as to not straddle two cache lines.
+
+The method used to pad basic blocks so as to align specific instructions must be transparent (i.e.
+not change the execution behavior).
+The typical approach to alignment is to introduce \texttt{nop} instructions, or \texttt{jmp} instructions which jump over the padding.
+Adding extra instructions, however, introduces added latency to the fetch/decode/execute cycle.
+Therefore, when possible, Granary pads instructions using otherwise ignored instruction prefixes.
+For example, Granary aligned patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86 processors \cite{AgnerMicroarchitecture}).
+
+%To reduce the overhead of alignment
+
+%In some cases, basic blocks can be directly connected during basic block translation time.
+For example, intra-trace CTIs, or CTIs targeting already translated basic blocks.
+However, not all basic blocks can be connected this way, therefore some runtime
+
+%Granary connects some basic blocks at translation time (e.g.
+intra-trace control-flow instructions).
+This is common when tracing is enabled, but also occurs when a branch targets an already translated basic block.
+However,
+
+%Some basic blocks are directly connected at basic block translation time.
+One example is branches that target other basic blocks within a trace.
+Another example is a branch to an already translated basic block.
+
+%Other basic blocks are only connected at runtime.
+For example, if the function \texttt{memset} has never been executed, then a \texttt{call} to \texttt{memset} will be translated to target edge code (\Cref{fig:direct_edge_code}) that, when executed, will translate \texttt{memset}'s first basic block and patch the \texttt{call} to target the translated basic block.
+
+%For example, the targets of function call instruction (\texttt{call}) are resolved just-in-time (unless the target is already known).
+For each translated \texttt{call} instruction, Granary generates edge code (\Cref{fig:direct_edge_code}) that will find or translate the first basic block of the function being called, and patch the translated \texttt{call} instruction to point to the function's first basic block.
+
+%For example, if a direct branch in one basic block targets another block within the trace under construction, then that
+
+%Basic blocks are directly connected at block translation time when branches in traces re-enter a trace, or when branches target previously translated basic blocks.
+If a branch does not target another basic block within its trace, or does not target a previously translated basic block, then the branch will target edge code that will resolve that branches target at runtime.
+
+% However, when a given basic block is translated, not all of its targets are known (e.g.
+\texttt{call} and \texttt{jCC} targets).
+Therefore, edge code (\Cref{fig:direct_edge_code}) is generated that translates/resolves a branch target when that branch is first executed, and patches the translated branch to avoi
+
+% The targets of some of these direct CTIs are only resolved at runtime, by executing edge code.
+As an optimization, jumps to edge code are dynamically patched to transfer control to their intended targeted basic blocks.
+
+%edge code use hot code patching to update a translated CTI to redirect execution
+
+\paragraph{Trace Allocation}
+
+\paragraph{Indirect Control Flow}
+
+%\subsection{Reproducability}
+%Hello
 

### 2013-11-25 VEE-14/paper.tex
Working on masters paper; writing about DBT optimizations in Granary.

@@
 \subsection{Optimizations}
-\paragraph{Ahead-of-time Translation} 
+\paragraph{Ahead-of-time Translation} \label{sec:aot}
 Granary ends basic blocks at conditional branches (\texttt{jCC}), direct and indirect \texttt{jmp}s, and \texttt{ret}s and \texttt{iret}s.
@@
 
-The method used to pad basic blocks so as to align specific instructions must be transparent (i.e.
-not change the execution behavior).
-The typical approach to alignment is to introduce \texttt{nop} instructions, or \texttt{jmp} instructions which jump over the padding.
-Adding extra instructions, however, introduces added latency to the fetch/decode/execute cycle.
-Therefore, when possible, Granary pads instructions using otherwise ignored instruction prefixes.
-For example, Granary aligned patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86 processors \cite{AgnerMicroarchitecture}).
+%Padding basic blocks so as to align specific instructions must not change the execution behavior.
+
+One approach to change an instruction's alignment is to pad the basic block with additional \texttt{nop} instructions.
+An alternative when more than a few bytes of padding are needed is to introduce a single control-flow instruction (\texttt{jmp}) that jumps around the padded area.
+However, adding extra instructions introduces latency into the fetch/decode/execute cycle.
+Therefore, when possible, Granary uses ignored instruction prefixes to alter an instruction's alignment.
+For example, Granary aligns patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86 processors \cite{AgnerMicroarchitecture}.
+Finally, most basic blocks are small (i.e.
+fit in one cache line), therefore, aligning basic blocks on a cache line boundary naturally avoids the potential for a patchable instruction to straddle two cache lines.
 
@@
 \paragraph{Trace Allocation}
+Arranging for commonly executed sequences of basic blocks to appear adjacent in memory, also called tracing, typically improves the runtime performance of instrumented programs \cite{DynamoRIO}.
+Typical approaches to tracing require a runtime profiling infrastructure (e.g.
+recording basic block execution counts).
+Granary attempts to reap the benefits of tracing without the runtime profiling overhead by opportunistically forming traces using ahead-of-time translation (\Cref{sec:aot}), and by arranging for related basic blocks to be allocated sequentially in memory.
+
+Two basic blocks $B_1$ and $B_2$ are ``related" ($B_1 \to B_2$) if there is a control flow instruction from $B_1$ to $B_2$.
+Grnary arranges for $B_2$ to be allocated nearby or adjacent to $B_1$ by allocating $B_2$'s instructions from the same pool of memory used to allocate $B_1$'s instructions.
+The base case of this induction are kernel entrypoints.
+For example, each system call is an entrypoint into the kernel, and is associated with a private memory pool.
+When a system call is executed for the first time, the basic blocks needed to fulfill the system call's computation are allocated from that system call's memory pool.
+
+To avoid translation conflicts (e.g.
+on utility functions being simultaneously executed by distinct system calls), and to improve the predictability and reproducability of a given code cache layout, Granary can delay the takeover of kernel entrypoints.
+That is, Granary can take over one entrypoint at a time, and wait for a specified period of time before taking over the next entrypoint.
+The effect of delaying the takeover process is that it increases the chances that a given entrypoint (e.g.
+a blocking system call) will execute to completion, and thus have all of its basic blocks  arranged sequentially in memory.
+
+%, and is associated with a private memory pool from which to allocate memory for basic blocks.
+The net result is that most of the basic blocks eecuted in service of a single system call are allocated from that system call's memory pool.
+
+%The intuition for this approach is that kernel code is not very ``loopy", and so
+
+%Granary services memory allocation requests for basic block instructions from memory pools. By default, Granary has as many memory pools as there are CPUs. However, when trace allocation is enabled
+
+%Tracing is also beneficial for
+
+%However, tracing of this sort requires mechnisms for evicting stale basic blocks,
+
+%Constructing traces of ``hot'' 
+%JIT generally results in better performance .
+
+%By default, Granary allocates the memory needed to hold the instructions of a basic block from CPU-private memory pools.
+Each memory pool serves basic block allocation requests out of fixed-size memory blocks, ranging from 4K to 24K in size.
 

### 2013-11-25 VEE-14/paper.tex
Added some diagrams

@@
 Two basic blocks $B_1$ and $B_2$ are ``related" ($B_1 \to B_2$) if there is a control flow instruction from $B_1$ to $B_2$.
-Grnary arranges for $B_2$ to be allocated nearby or adjacent to $B_1$ by allocating $B_2$'s instructions from the same pool of memory used to allocate $B_1$'s instructions.
+Granary arranges for $B_2$ to be allocated nearby or adjacent to $B_1$ by allocating $B_2$'s instructions from the same pool of memory used to allocate $B_1$'s instructions.
 The base case of this induction are kernel entrypoints.
@@
 a blocking system call) will execute to completion, and thus have all of its basic blocks  arranged sequentially in memory.
+
+Trace allocation is implemented by recording the memory pool used to allocate a given basic block in that basic block's metadata.
+When Granary needs to translate a new basic block, it inspects the predecessor basic block's metadata and uses the predecessor basic block's memory pool to service allocation requests for the basic block in translation.
 
@@
 
+\begin{figure}
+\captionsetup[subfloat]{width=0.45\columnwidth}
+\begin{tabularx}{\columnwidth}{XX}%
+\subfloat[Full kernel instrumentation.]{\hspace{2.5em}\epsfig{file=diagrams/kernel.eps,width=0.125\textwidth}} & %
+\subfloat[Module-only instrumentation]{\hspace{2.5em}\epsfig{file=diagrams/module.eps,width=0.125\textwidth}} \\
+\subfloat[Module-only instrumentation, with a function specialized to run natively by using a module function wrapper.]{\hspace{2.5em}\epsfig{file=diagrams/module_wrapper.eps,width=0.125\textwidth}} & %
+\subfloat[\label{fig:module_probe}Module-only instrumentation, with a kernel function specialized to also be instrumented by using a probe-based wrapper.]{\hspace{2.5em}\epsfig{file=diagrams/module_probe.eps,width=0.125\textwidth}}
+\end{tabularx}
+%
+%\begin{multicols}{2}
+%\subfloat[Full kernel instrumentation.]{\epsfig{file=diagrams/kernel.eps,width=0.125\textwidth}}
+%\columnbreak
+%\subfloat[Module-only instrumentation]{\epsfig{file=diagrams/module.eps,width=0.125\textwidth}}
+%\end{multicols}
+%\begin{multicols}{2}
+%\subfloat[Module-only instrumentation, with a function specialized to run natively by using a module function wrapper.]{\epsfig{file=diagrams/module_wrapper.eps,width=0.125\textwidth}}
+%\columnbreak
+%\subfloat[Module-only instrumentation, with a kernel function specialized to also be instrumented by using a probe-based wrapper.]{\epsfig{file=diagrams/module_probe.eps,width=0.125\textwidth}}
+%\end{multicols}
+\caption{Example of using different granularities of instrumentation to selectively instrument or not instrument specific code.}
+\end{figure}
+
 Granary can instrument or not instrument functions in several different ways.
@@
 memory allocations) to instrumentation tools that don't instrument the whole kernel (e.g.
-module-only).
+module-only, \Cref{fig:module_probe}).
 One example tool that combines module-only and probe-based instrumentation is Granary's kernel module memory leak detector.

### 2013-11-27 VEE-14/paper.tex
Did some stuff.. which I don't remember
Described indirect branch lookup, resectioned some stuff.

@@
 
-\paragraph{Function return addresses}\label{para:return_address_transparency} Unlike most DBT frameworks, Granary inlines \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses.
+\subsubsection{Function return addresses}\label{para:return_address_transparency} Unlike most DBT frameworks, Granary inlines \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses.
 Inlining \texttt{call} instructions avoids return-address mispredictions and unnecessary control-flow by building longer basic blocks.
 
-\paragraph{Interrupt return addresses} Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts.
+\subsubsection{Interrupt return addresses} Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts.
 A special case arises for page fault exceptions that occur within specific kernel functions that access user space data \cite{btkernel}.
@@
 \subsection{Reentrancy}
-
-\paragraph{Code cache}
-Granary's code cache and JIT translation mechanism is fully reentrant.
+Reentrancy is an important concern for all DBT systems, and is especially important for DBT systems operating in the kernel environment.
+For example, hardware interrupts can redirect control flow at any instruction, and interrupt delivery cannot be delayed past the execution of memory instructions that might alter the interrupt delivery state \cite{DRK}.
+Preserving interrupt transparency, however, requires that interrupts be delayed until the end of any injected instrumentation instructions.
+Together, these requirements make it very difficult to write instrumentation tools.
+
+\subsubsection{Code cache}
+Granary's code cache and JIT translation mechanisms are fully reentrant.
 Context-switches into Granary disable interrupts and switch execution onto a CPU-private stack.
@@
 This approach presents a minor challenge: context-switches save or restore the entire register state, so as not to disturb the execution of the instrumented program.
-Therefore, if a context switch into Granary occurs because the next basic block must be found/translated, then the address of that next basic block must be stored \emph{somewhere} in order for execution to eventually reach that basic block when the context switches back to the code cache.
+
+%Therefore, if a context switch into Granary occurs because the next basic block must be found/translated, then the address of that next basic block must be stored \emph{somewhere} in order for execution to eventually reach that basic block when the context switches back to the code cache.
 But context switching is not allowed to perturb the register state of the instrumented program, and so the target address must be resolved by another means.
 
-Like btkernel, Granary solves this problem of figuring out how to get to where the code needs to go  by generating ``edge'' code for each direct control-flow instruction (e.g.
+%Like btkernel, Granary solves this problem of figuring out how to get to where the code needs to go  by generating ``edge'' code for each direct control-flow instruction (e.g.
 \texttt{call}, \texttt{jmp}, \texttt{jcc}).
-When a basic block is first translated, its direct control-flow instructions (CTIs) are converted into jumps to CTI-specific edge code.
-Edge code lives outside the code cache and is responsible for transferring control from a basic block ($B_{pred}$) to Granary so that the next basic block ($B_{succ}$) can be found/translated, and then connected ($B_{pred} \underset{\texttt{jz}}{\to} B_{succ}$) via hot code patching.
-Reentrancy of edge code is obtained by spilling registers onto the runtime call stack, and later restoring them before jumping back to the now-patched CTI (\Cref{fig:direct_edge_code}).
-
-Granary also generates edge code for indirect control-flow instructions; however, this edge code is split into two separate components: entry and exit edge code.
+
+%Granary also generates edge code for indirect control-flow instructions; however, this edge code is split into two separate components: entry and exit edge code.
 Entry edge code is generated on a per-indirect CTI instruction basis, and spills registers onto the runtime call stack.
@@
 
-\paragraph{Instrumentation}
-
+\subsubsection{Instrumentation}
+
+One challenge with any DBT system is how it copes with non-reentrant instrumentation.
+For example, instrumentation that modifies CPU-private data structures cannot be safely executed by pre-emptive kernel running on a multi-core system.
+If the task being instrumented is interrupted, then the kernel might decide to re-schedule the task to resume its execution on a different CPU.
+Any modifications to CPU-private state would lead to undefined behavior.
+
+DRK's solution to this problem assumes that all injected instrumentation instructions are potentially non-reentrant, and \emph{delays} interrupts occuring within instrumented code \cite{DRK}.
+Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86: interrupts only arrive on \emph{logical} instruction boundaries, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
+This approach introduces subtle complexity when creating instrumentation that injects instructions both before (pre) and after (post) native instructions.
+For example, an interrupt occuring in the pre-instrumentation of a native instruction will be delayed only until the native instruction.
+When the interrupt is handled and execution attempts to return to the interrupted instructions, DRK re-instruments the ``tail" of interrupted basic block, starting from the native instruction.
+Instrumenting this tail involves injecting pre- and post-instrumentation around the first native instruction.
+If the pre-instrumentation is not idempotent, then its re-execution will result in undefined behavior.
+We encountered this issue when using DRK for kernel module instrumentation, where our pre-instrumentation spilled registers to the runtime call stack, and our post-instrumentation restored the spilled registers.
+Because of interrupt delaying, the spilling code was sometimes repeated many times, whereas the restoring was only executed once.
+This led to hard-to-debug kernel crashes.
+
+Btkernel provides no solution to the problem of handling non-reentrant instrumentation, beyond requiring that instrumentation tools disable and re-enable interrupts around non-reentrant instrumentation code \cite{btkernel}.
+Unfortunately, disabling and re-enabling interrupt is costly.
+
+Granary charts a middle ground between the two aforementioned approaches.
+Interrupt delaying has value because it enables non-reentrant instrumentation to be injected without costly disabling/enabling of interrupts, while not supporting delaying has value because not all instrumentation is non-reentrant.
+By default, Granary does not delay interrupts.
+However, instrumentation tools can declare \emph{interrupt delay ranges}: sequences of instructions (which can contain internal control flow) inside of a basic block that must be executed atomically.
+Granary encodes bounds information about interrupt delay ranges inside of each basic block's meta-data.
+When a basic block is interrupted, Granary inspects the meta-data and determines if the interrupt must be delayed.
+If a delay is required, then the instructions belonging to the delay range are copied into a CPU-private buffer, followed by code that re-constructs the interrupt stack frame and re-issues the interrupt to Granary.
+Granary then emulates a return from the current interrupt handler (without issuing an \texttt{iret} instruction), but leaves interrupts disabled and redirects execution into the copied code.
+When the copied code finishes, Granary regains control and observes the reconstructed interrupt stack frame.
+At this point, Granary or the tool can choose to handle the interrupt or defer the handling of the interrupt to the kernel.
+
+%, but introduces implementation complexities and high overheads for interrupt-heavy workloads. Precise interrupt delivery is one of the ways which DRK maintains transparency: interrupt handlers only 
+%The interaction between DRK's interrupt delaying mechanism and how also makes designing instrumentation subtlely harder.
+For example, Granary's \texttt{RCUdbg} tool injects instrumentation inside of read-side critical sections that looks for loads/stores from tainted addresses.
+When necessary, \texttt{RCUdbg} spills and restores registers around each checked load/store operation, so that the registers can be safely used by the checking code.
+This is an examuple of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after.
+DRK is oblivious to the relationship between pre/post instrumentation, and so an interrupt that occurs within the pre-instrumentation will only be delayed until the native instruction.
+
+\subsection{Implementation}
+\subsubsection{Direct Control Flow}
 % Code block showing an example JZ (jump if ZF=1).
@@
 \end{figure}
-
-One challenge with any DBT system is how it copes with non-reentrant instrumentation.
-For example, instrumentation that modifies CPU-private data structures cannot be safely executed by pre-emptive kernel running on a multi-core system.
-If the task being instrumented is interrupted, then the kernel might decide to re-schedule the task to resume its execution on a different CPU.
-Any modifications to CPU-private state would lead to undefined behavior.
-
-DRK's solution to this problem assumes that all injected instrumentation instructions are potentially non-reentrant, and \emph{delays} interrupts occuring within instrumented code \cite{DRK}.
-Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86: interrupts only arrive on \emph{logical} instruction boundaries, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
-This approach introduces subtle complexity when creating instrumentation that injects instructions both before (pre) and after (post) native instructions.
-For example, an interrupt occuring in the pre-instrumentation of a native instruction will be delayed only until the native instruction.
-When the interrupt is handled and execution attempts to return to the interrupted instructions, DRK re-instruments the ``tail" of interrupted basic block, starting from the native instruction.
-Instrumenting this tail involves injecting pre- and post-instrumentation around the first native instruction.
-If the pre-instrumentation is not idempotent, then its re-execution will result in undefined behavior.
-We encountered this issue when using DRK for kernel module instrumentation, where our pre-instrumentation spilled registers to the runtime call stack, and our post-instrumentation restored the spilled registers.
-Because of interrupt delaying, the spilling code was sometimes repeated many times, whereas the restoring was only executed once.
-This led to hard-to-debug kernel crashes.
-
-Btkernel provides no solution to the problem of handling non-reentrant instrumentation, beyond requiring that instrumentation tools disable and re-enable interrupts around non-reentrant instrumentation code \cite{btkernel}.
-Unfortunately, disabling and re-enabling interrupt is costly.
-
-Granary charts a middle ground between the two aforementioned approaches.
-Interrupt delaying has value because it enables non-reentrant instrumentation to be injected without costly disabling/enabling of interrupts, while not supporting delaying has value because not all instrumentation is non-reentrant.
-By default, Granary does not delay interrupts.
-However, instrumentation tools can declare \emph{interrupt delay ranges}: sequences of instructions (which can contain internal control flow) inside of a basic block that must be executed atomically.
-Granary encodes bounds information about interrupt delay ranges inside of each basic block's meta-data.
-When a basic block is interrupted, Granary inspects the meta-data and determines if the interrupt must be delayed.
-If a delay is required, then the instructions belonging to the delay range are copied into a CPU-private buffer, followed by code that re-constructs the interrupt stack frame and re-issues the interrupt to Granary.
-Granary then emulates a return from the current interrupt handler (without issuing an \texttt{iret} instruction), but leaves interrupts disabled and redirects execution into the copied code.
-When the copied code finishes, Granary regains control and observes the reconstructed interrupt stack frame.
-At this point, Granary or the tool can choose to handle the interrupt or defer the handling of the interrupt to the kernel.
-
-%, but introduces implementation complexities and high overheads for interrupt-heavy workloads. Precise interrupt delivery is one of the ways which DRK maintains transparency: interrupt handlers only 
-%The interaction between DRK's interrupt delaying mechanism and how also makes designing instrumentation subtlely harder.
-For example, Granary's \texttt{RCUdbg} tool injects instrumentation inside of read-side critical sections that looks for loads/stores from tainted addresses.
-When necessary, \texttt{RCUdbg} spills and restores registers around each checked load/store operation, so that the registers can be safely used by the checking code.
-This is an examuple of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after.
-DRK is oblivious to the relationship between pre/post instrumentation, and so an interrupt that occurs within the pre-instrumentation will only be delayed until the native instruction.
+Like btkernel, Granary translates direct control-flow instructions (CTIs) into equivalent jumps into CTI-specific ``edge code" (\Cref{fig:direct_edge_code}).
+Edge code lives outside the code cache, so as to not pollute the icache \cite{btkernel}.
+
+When control flows through a basic block $B_{pred}$ for the first time, any jumps taken will transfer control to edge code that is specific to the jump taken.
+Edge code contains only a single indirect \texttt{call} instruction, thus allowing multiple edge codes to occupy a single cache line.
+The indirect \texttt{call} transfers control to Granary via a function pointer located in an ``edge data structure", which describes how to translate the target basic block, $B_{succ}$.
+Granary gains control and looks up or translates the target code according instrumentation policy information stored in the edge data structure.
+Then, Granary uses hot-code patching (\Cref{sec:patching}) to modify the original CTI that led to the edge code \texttt{call}, so that it now transfers control to $B_{succ}$ instead.
+Finally, Granary releases control and returns to $B_{succ}$.
+
+% and is responsible for transferring control from a basic block ($B_{pred}$) to Granary so that the next basic block ($B_{succ}$) can be found/translated, and then connected ($B_{pred} \underset{\texttt{jz}}{\to} B_{succ}$) via hot code patching.
+
+\subsubsection{Indirect Control Flow}
+
+%Indirect branch lookup through a global, fixed-size hash table.
+A basic block making an indirect control flow transfer jumps to an in-edge that knows where to find the indirect CTI target (e.g.
+in a specific register or memory location).
+The in-edge then uses the target as an index into the hash table, and jumps to the target address stored in the indexed hash table entry.
+That entry will either be the address of an out-edge, or of Granary's indirect branch resolver.
+Each out-edge is specific to a target basic block, and first checks if the intended jump target matches the out-edge's target.
+If so, then control is transfered to the target basic block, otherwise control is transfered either to Granary or to the next out-edge.
+Granary chains together out-edges through conditional jumps just as
+
+%Hash table entries are code addresses of out-edge code stub.
+Each out-edge code stub is specific to a target basic block, and is responsible for checking if the intended control-flow target is the out-edge's target or not.
+If not, the out-edge either directly jumps to the next out-edge whose target also fits into the same hash table bucket, or jumps to Granary to determine the target basic block.
+
+Granary translates indirect control flow instructions (\texttt{call}, \texttt{jmp}) into their equivalent direct control flow versions, where the direct versions jump to ``in-edge" code.
+In-edge code loads the indirect jump target (e.g.
+a specific register or memory location), saves the flags register, and several scratch registers.
+Finally, the in-edge uses the indirect target as an index into a global, fixed-size hash table of indirect branch targets.
+The entries of the hash table either point to Granary or to ``out-edge" code (\Cref{fig:ibl}).
+
+\begin{figure}[t!]
+\hspace{4em}\includegraphics[width=0.9\columnwidth]{diagrams/ibl.pdf}\vspace{-3em}
+\caption{\label{fig:ibl}Process of resolving an indirect branch.
+Control transfers from a source basic block, to the in-edge code specific to where the indirect target is stored (e.g.
+register, memory address), through the global hash table and on to either one or more out-edge code stubs, or to Granary.}
+\end{figure}
+
+Out-edge code is specific to a target basic block, and performs the function of separate chaining in the indirect branch lookup hash table.
+When control transfers to out-edge code, the out-edge checks to see if the indirect target matches the out-edge code's target basic block.
+If they match, then the out-edge code restores scratch and flags registers and jumps directly to the target basic block.
+Otherwise, the out-edge code jumps to another out-edge code or to Granary.
+If one out-edge code block jumps to another one, then the targets of the two out-edge code blocks collide in the global hash table.
+In this respect, out-edge code is chained together through control flow, just as colliding entries in a data hash table can be chained together through a linked list.
+
+%If an entry points to Granary, then no as-of-yet resolved indirect jump target hashes to that entry.
+
+%Granary implements separate chaining through code (i.e. control flow jumps, as opposed to date, e.g. linked list pointers)
+%In the event that one-or-more indirect jump targets do hash to an existing entry
+%In-edge code is responsible for 
+% to first search for their targets in a global, fixed-size hash table, and failing that, context-switch into Granary to perform a full lookup and potential translation.
+%using a fixed-size hash table, where the native target of the indirect
 
 \subsection{Optimizations}
-\paragraph{Ahead-of-time Translation} \label{sec:aot}
+\subsubsection{Ahead-of-time Translation} \label{sec:aot}
 Granary ends basic blocks at conditional branches (\texttt{jCC}), direct and indirect \texttt{jmp}s, and \texttt{ret}s and \texttt{iret}s.
@@
 
-\paragraph{Patching and Alignment}
+\subsubsection{Patching and Alignment}\label{sec:patching}
 
@@
 
-\paragraph{Trace Allocation}
+\subsubsection{Trace Allocation}
 Arranging for commonly executed sequences of basic blocks to appear adjacent in memory, also called tracing, typically improves the runtime performance of instrumented programs \cite{DynamoRIO}.
@@
 The net result is that most of the basic blocks eecuted in service of a single system call are allocated from that system call's memory pool.
-
 %The intuition for this approach is that kernel code is not very ``loopy", and so 
-
 %Granary services memory allocation requests for basic block instructions from memory pools. By default, Granary has as many memory pools as there are CPUs. However, when trace allocation is enabled
-
 %Tracing is also beneficial for 
-
 %However, tracing of this sort requires mechnisms for evicting stale basic blocks, 
-
 %Constructing traces of ``hot'' 
 %JIT generally results in better performance .
-
 %By default, Granary allocates the memory needed to hold the instructions of a basic block from CPU-private memory pools.
@@
 
-\paragraph{Indirect Control Flow}
+\subsubsection{System Calls}
 
@@
 
-\paragraph{Probe-based instrumentation}
+\subsubsection{Probe-based instrumentation}
 
@@
 
-\paragraph{Wrapper-based instrumentation}
+\subsubsection{Wrapper-based instrumentation}
 

### 2013-11-27 VEE-14/paper.tex
Fixed up some sub-sub-sections, and described syscall optimizations.

@@
   cmp $0x0, %rax
-edge_jz:
   jz edge_value_is_zero
-  ...
+  jmp edge_fall_through
+
 edge_value_is_zero:
-  call (patch_edge_jz)
+  call *(jz_patch_data)
+edge_fall_through:
+  call *(jmp_patch_data)
 \end{lstlisting}
@@
 \caption{\label{fig:direct_edge_code}Example translation of a conditional control flow instruction (\texttt{jz}) and its edge code.
-The original \texttt{jz} is translated into a \texttt{jz} that uses a 32-bit relative displacement to transfer control to an edge code function call (\texttt{edge\_jz}).
-When executed, the function call reads its target from a data structure (\texttt{patch\_edge\_jz}), which contains all information necessary to resolve the \texttt{jz}'s target and patch the original \texttt{jz}.
+The original \texttt{jz} is translated into a \texttt{jz} that uses a 32-bit relative displacement to transfer control to an edge code function call (\texttt{edge\_value\_is\_zero}).
+When executed, the function call reads its target from an edge data structure (\texttt{jz\_patch\_data}), which contains all information necessary to resolve the \texttt{jz}'s target and patch the original \texttt{jz}.
 Patching is performed with a four byte atomic write to the relative displacement operand of the \texttt{jz} instruction.
@@
 
-\subsubsection{Indirect Control Flow}
+\subsubsection{Indirect Control Flow}\label{sec:ibl}
 
@@
 If one out-edge code block jumps to another one, then the targets of the two out-edge code blocks collide in the global hash table.
-In this respect, out-edge code is chained together through control flow, just as colliding entries in a data hash table can be chained together through a linked list.
+In this respect, out-edge code is chained together through control flow, just as colliding entries in a typical hash table can be chained together through a linked list.
 
@@
 \subsubsection{System Calls}
+
+Granary treats indirect \texttt{jmp}s or \texttt{call}s through the Linux kernel system call table specially.
+The Linux kernel (as of Linux 3.12) has approximately 300 distinct system call targets, some of which are executed often.
+To avoid polluting the indirect branch lookup hash table with these targets (\Cref{sec:ibl}), and to avoid the indirect branch lookup overhead, Granary creates a separate, \emph{shadow} system call table.
+
+%re-direct jumps through the system call table
+
+Granary's translation post-processing mechanism identifies indirect control flow through the system call table and alters those CTIs to direct control flow through the shadow system call table instead.
+The entries of the shadow system call table are ``lazy" entrypoint translator functions.
+Upon their first execution, these entrypoint translators find or translate the first basic block of the system call handler, and update the shadow system call table accordingly.
+
+%Directing control through the shadow system call table avoids the overhead of going through the indirect branch lookup (IBL) table (\Cref{sec:ibl}), and reduces the number of potential conflicts in the IBL table.
+
+% whose entries are code cache basic block addresses translated according to the starting instrumentation policy.
 
@@
 
-Granary can target instrumentation only at specific kernel modules, while executing the rest of the kernel natively and without overhead.
+Unlike DRK and btkernel, Granary can target instrumentation only at specific kernel modules, while executing the rest of the kernel natively and without overhead.
 Instrumenting only specific kernel modules is particularly useful for debugging tools because modules represent the bulk of new kernel code under development, and contain the most bugs \cite{FaultsInLinux}.

### 2013-12-05 VEE-14/paper.tex
Updated the intro.

@@
 \newcommand{\comment}[1]{}
+
+% For writing Granary tool names.
+\newcommand{\toolname}[1]{{\scshape #1}}
 
@@
 In practice, we would like to target instrumentation at specific code for efficiency, while also allowing rich and detailed instrumentation of the targeted code.
-Limiting the scope of what is instrumented is  important because the overhead introduced by using runtime code instrumentation is usually dominated by the instrumentation itself and not the runtime system supporting that instrumentation.
+Limiting the scope of what is instrumented is  important because the overhead introduced by using runtime code instrumentation is usually dominated by the instrumentation itself and not the baseline overhead of the runtime system supporting that instrumentation.
+Also, specializing based on what is instrumented allows instrumentation tools to avoid unnecessary filtering.
+That is, if one wants to instrument RCU read-side and write-side critical sections, as is the case with Granary's \toolname{RCUdbg} tool, then instrumenting all code just to filter out / find those sections is wasteful.
 
@@
 they are unable to change how code is instrumented based on the context in which that code will run.
-For example, Granary's Read-Copy-Update (RCU) debugging tool, \texttt{RCUdbg}, specializes instrumentation at different stages of use of the RCU synchronization API.
+This is unfortunate because it means that tools wanting to instrument code differently must create their own solution that works around this deficiency of existing systems.
+A typical solution is for tools to generate as many instrumented versions of the code as there are execution contexts.
+When instrumented code is  executing, runtime checks query some thread-private or CPU-private state, and jump to the correct version of the code for the current context.
+Runtime checks of this form are inefficient, and generating many versions of the code introduces bloat.
+The machinery needed to multiplex between different versions of instrumented code is also orthogonal to the goals of the instrumentation tool, thus distracting tool-developers from their primary objectives.
+
+%Using an existing system, 
+% get around the problem of needing different instrumentation in different contexts, existing systems need to create as many versions of the code as there are contexts.
+Runtime checks must also be inserted into the code that coordinate with the runtime to track and detect the current execution context and jump to the correct version of the code.
+Runtime checks and state tracking impose unnecessary complications to tool-authors.
+The added code needed to query the current context and switch to a different version of the instrumentation is code that is not actively achieving the primary goal of the tool, and is instead making up for a defeciency in the runtime system.
+%are orthogonal to the main goals of an analysis tool:
+% Aside from the additional ineffeciency of these runtime checks, there is also the 
+% However, generating many differently instrumented versions of the code introduces bloat, especially when not all versions are well exercised.
+Also, generating many different versions of the code introduces state tracking and switching costs.
+% This imposes unnecessary performance overheads because of the need to additional instrumentation that is unrelated to 
+% due to having to perform runtime detection and swichting
+% to the right version of the code for every basic block or every instruction.
+% according to \emph{all} contexts. Instrumenting code according to all contexts imposes unnecessary performance overheads and requires state tracking.
+%,  dynamically tracking the current context, and dynamically choosing which 
+% commit to instrumenting code to handle \emph{all} contexts. For example, if
+
+One example of a tool that uses runtime code specialization is Granary's Read-Copy-Update (RCU) debugging tool, \toolname{RCUdbg}.
+\toolname{RCUdbg} catches bugs related to invalid uses of RCU-protected pointers (\Cref{sec:rcudbg}).
+\toolname{RCUdbg} distinguishes between two kinds of RCU-protected pointers (reader-side and writer-side) and plain old pointers, and three execution contexts (inside of a read-side critical section, inside of a write-side critical section, and in non-RCU-related code).
+The actions taken when a pointer is de-referenced varies depending on the kind of pointer and the execution context.
+Code executing in read-side critical sections is specialized to enforce that only plain-old and read-side pointers are accessed, and that all accesses through read-side pointers are memory reads.
+Write-side critical sections are specialized in a similar way.
+
+%There are similar checks for write-side pointers in write-side critical sections.
+% The instrumentation performed inside of read- and write-side critical sections is heavy weight, while code running in the third execution context can either run with lightweight or no instrumentation.
+%RCUdbg distinguishes between RCU-protected pointers and plain old pointers.
+There are two kinds of RCU-protected pointer (read-only, and read/write), and every access to an RCU-protected pointer is checked.
+The checks performed on an access to an RCU-protected pointer are context-specific.
+That is, a read-only pointer can only be read from within an RCU read-side critical section.
+Read/write pointers have similar restrictions.
+RCUdbg needs to know in which context some code is executing, i.e.
+in a read-side or write-side critical section, or outside of all critical section, and change instrumentation to check for different kinds of bugs based on the context.
+%A read/write pointer has different restrictions.
+In read-side critical sections, a read/write pointer can only be accessed by means of a special API function.
+In write-side critical sections, a read/write pointer can only be written to
+% One kind of RCU-protected pointer can be read from within ``RCU read-side critical sections", while another kind can (almost) arbitrarily accessed, but only inside of an ``RCU write-side critical section".
+Outside of these two kinds of sections, accessing RCU-protected pointers is illegal.
+The distinction between different contexts of code (read-side, write-side, outside), and different kins of pointers makes
+%Creating a tool to catch these kinds of bugs with an existing DBT system would be difficult because that tool would need to  
+%For example, Granary's Read-Copy-Update (RCU) debugging tool, \toolname{RCUdbg}, specializes instrumentation at different stages of use of the RCU synchronization API.
 Correct usage of the RCU API allows programmers to create scalable, shared data structures that permit concurrent access and modification.
 The ``trick" of RCU is that readers might not see the latest write to the data structure, and writers must synchronize to modify the data structure.
-\texttt{RCUdbg} distinguishes between code executing inside \emph{read-side} and \emph{write-side} critical sections, and instruments each code context differently.
-In read-side critical sections, \texttt{RCUdbg} detects bugs when readers access potentially non-existent data, do not the correct accessor API, or attempt to write to shared data.
-In write-side critical sections, \texttt{RCUdbg} detects when writers write to shared data without using the correct API.
-
+\toolname{RCUdbg} distinguishes between code executing inside \emph{read-side} and \emph{write-side} critical sections, and instruments each code context differently.
+In read-side critical sections, \toolname{RCUdbg} detects bugs when readers access potentially non-existent data, do not the correct accessor API, or attempt to write to shared data.
+In write-side critical sections, \toolname{RCUdbg} detects when writers write to shared data without using the correct API.
 %In reader-side code
-
 %For example, Granary has a tool that  dynamically constructs inter- and intra-procedural control flow graphs (CFGs).
@@
 
-The \texttt{RCUdbg} tool is an example of \emph{context-driven code specialization}: code executing inside of a read-side critical section is instrumented differently from code executing inside of a write-side critical section, and code execution outside of both contexts executes with minimal instrumentation.
-In a Granary tool, each tracked execution context is associated with its own instrumentation policy.
+The \toolname{RCUdbg} tool is an example of \emph{context-driven code specialization}: code executing inside of a read-side critical section is instrumented differently from code executing inside of a write-side critical section, and code executing outside of both contexts runs natively or with lightweight instrumentation.
+Achieving this level of specialization using and existing system requires complicated runtime state management and checks to switch between different versions of the code.
+With Granary, a tool like \toolname{RCUdbg} can specialize code using instrumentation policies.
+An instrumentation policy is a function that names an execution context and decides how to instrument instructions that will execute within that context.
+When a policy is used to instrument a sequence of instructions (called a basic block), a new version of the basic block is generated and cached.
+
+\toolname{RCUdbg} defines three policies, and declares transitions between these policies when control flows in and out of RCU read- and write-side critical sections.
+When a function like \texttt{memset} is invoked in one context, the policy associated with that context is used to generate an instrumented version of \texttt{memset} for that context.
+If, for example, \texttt{memset} is never invoked within a read-side critical section, then \texttt{memset}'s code will never be specialized to perform the heavyweight checks needed to detect RCU-protected pointer misuses in this context.
+Similarly, if \texttt{memset} is executed both in a write-side critical section and outside of any RCU-related code, then two differently instrumented versions of \texttt{memset} will be generated.
+
+The key insight of Granary's approach to runtime code specialization is that code versioning is a context-tracking mechanism.
+That is, if an instruction from a basic block instrumented by \toolname{RCUdbg}'s read-side critical section policy is being executed, then the current thread must be executing inside of a read-side critical section.
+Similarly, if the read-side policy is used to instrument a basic block, then \texttt{RCUdbg} knows at instrumentation time\footnote{Instrumentation time is analogous to ``compile time", in that a policy is generally used to instrument a basic block of code once, and then that basic block may be executed many times.} that this basic block will only execute within a read-side critical section.
+If the basic block contains a \texttt{call} to the \texttt{memset} function, then \texttt{memset} is also known to execute within a read-side critical section because it does not terminate a read-side critical section.
+By deciding which version of \texttt{memset} to execute at instrumentation time, \toolname{RCUdbg} avoids expensive runtime checks that other systems impose in order to achieve similar kinds of specialization.
+\toolname{RCUdbg} can specify that \texttt{memset} will also execute within a read-side critical section by telling Granary that it should use the read-side critical section policy to instrument \texttt{memset}.
+
+%Granary records 
+%Policy information is recorded inside of Granary's
+
+% instruction by telling Granary what policy to use to instrument the code 
+%, and when the \texttt{call} is eventually executed, 
+%This simplifies tool creation becase 
+%By propagating policy information through control-flow transferring instructions, 
+%Therefore, we can know at translation and instrumentation time that
+% No additional state needs to be maintained by \texttt{RCUdbg} to determine which version of the basic block to execute.
+%Any control-flow instructions the transfer control to other basic blocks that aren't the \texttt{rcu\_read\_unlock} function must also 
+%That is, the failure of existing systems is that they have no built-in support for maintaining multiple versions of code.
+Tools built on existing systems must emulate versioning by duplicating their instrumentation for each desired version, and add code that executes the right version at the right time.
+%, as would be done using existing systems, 
+%What makes Granary's runtime code specialization efficient is that it ties into Granary's \texttt{just-in-time} instrumentation model.
+That is, Granary only instruments code that will execute.
+Code that never executes is neither translated nor instrumented.
+Also, only one version of a basic block is generated at a time.
+Therefore, even though \toolname{RCUdbg} defines three policies, only
+%If some native code is about to execute, then Granary will only instrument it with the policy associated with 
+%, and an instrumentation policy is only applied to code if that code will execute within the context associated with that context.
+%defines three different instrumentation policies--one for each execution context--and declares switches 
+% instrumentation policies to 
+%only needs to define three instrumentation policies (one for each execution context), and declare 
+%different versions of the code are generated \emph{just-in-time}.
+Therefore, if the \texttt{memset} function is never executed inside of a read-side critical section, then \texttt{memset}'s code will never be specialized to perform the heavyweight checks needed to detect RCU-protected pointer misuses in this context.
+Similarly, if \texttt{memset} is executed both in a write-side critical section and outside of any RCU-related code, then two differently instrumented versions of \texttt{memset} will be generated.
+%However, unlike other systems, no runtime check that chooses to dispatch to one version of \texttt{memset} or another depending on the runtime context. Instead, 
+% if some code never executes in a particular context, then Granary will not generate an instrumented version of that code for that context.
+The second benefit of this approach is that it is stateless, and requires no runtime checks to switch between different versions of the code depending on the execution context.
+%Granary tracks the current execution context with zero runtime overhead by 
+%In a Granary tool, each tracked execution context is associated with its own instrumentation policy.
 An instrumentation policy decides how to instrument code that will execute within the associated context.
-Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks.
+%Different versions of the code are instrumented \emph{just-in-time}. That is, there will 
+%Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks.
 This avoids typical poblems with tracking context in just-in-time DBT systems (e.g., maintaining runtime context-tracking state, reentrancy, and the inability to predict later execution contexts at translation time).
-
 % In the previous example, executing code is classified into two contexts: entrypoint and non-entrypoint basic blocks.
@@
 $P_{\mathit{after\_entry}}$ connects basic blocks in the intra-procedural CFG.
-
 %Finally, an important missing feature of existing low-level instrumentation systems is the ``big picture'' view that one gets from static analysis information.
@@
 Tools can use wrappers to apply context-specific policies when instrumenting module code as well as to inspect and manipulate parts of kernel/module memory in a type-safe way.
-
 %For example, we are actively developing a tool that uses static type information to learn about all the fields of all kernel data structures that are read from or written to by modules.
@@
 
-The rest of the paper describes Granary in more detail. \Cref{sec:dbt} describes how Granary is implemented.
+The rest of the paper describes Granary in more detail. \Cref{sec:dbt} describes how Granary is implemented, \Cref{sec:what} describes how Granary can specialize instrumentation
 
@@
 
+\subsection{Basic Blocks}
+
 Basic blocks in Granary's code cache contain x86-64 binary instructions.
@@
 %The interaction between DRK's interrupt delaying mechanism and how also makes designing instrumentation subtlely harder.
-For example, Granary's \texttt{RCUdbg} tool injects instrumentation inside of read-side critical sections that looks for loads/stores from tainted addresses.
-When necessary, \texttt{RCUdbg} spills and restores registers around each checked load/store operation, so that the registers can be safely used by the checking code.
+For example, Granary's \toolname{RCUdbg} tool injects instrumentation inside of read-side critical sections that looks for loads/stores from tainted addresses.
+When necessary, \toolname{RCUdbg} spills and restores registers around each checked load/store operation, so that the registers can be safely used by the checking code.
 This is an examuple of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after.
@@
 For example, some instrumentation is high-overhead, but does not need to be applied to all code.
-One example of this is Granary's \texttt{RCUdbg} tool, which instruments read-side and write-side critical sections more heavily than the rest of code.
+One example of this is Granary's \toolname{RCUdbg} tool, which instruments read-side and write-side critical sections more heavily than the rest of code.
 Unfortunately, existing DBT systems fall short in their support for runtime code specialization \cite{DRK,btkernel,Pin,DynamoRIO}.
 
-To implement the equivalent of \texttt{RCUdbg}'s runtime code specialization using an existing DBT system requires that a tool developer maintain, update, and check runtime state to know whether or not code is executing inside or outside of a read-side critical section.
+To implement the equivalent of \toolname{RCUdbg}'s runtime code specialization using an existing DBT system requires that a tool developer maintain, update, and check runtime state to know whether or not code is executing inside or outside of a read-side critical section.
 If a basic block is executing in a read-side critical section, then instrumentation should be enabled, and if it's not in a read-side critical section then it should be disabled.
@@
 %For example, instrumenting only module code allows Granary to specialize the execution of native code.
-In the same way, policy switching allows Granary's \texttt{RCUdbg} tool to specialize the execution of instrumented code executing inside of read-side and write-side critical sections.
+In the same way, policy switching allows Granary's \toolname{RCUdbg} tool to specialize the execution of instrumented code executing inside of read-side and write-side critical sections.
 In fact, optimizing the performance of an early Granary tool was the original motivation for policy switching.
@@
 
-%In the case of our \texttt{RCUdbg}, there were three execution contexts: within and outside of a read-side critical section.
+%In the case of our \toolname{RCUdbg}, there were three execution contexts: within and outside of a read-side critical section.
 If the same module basic block is  executed in the two contexts, then Granary's code cache would contain two different instrumented versions of that basic block.
@@
 
-For example, our \texttt{RCUdbg} invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.
+For example, our \toolname{RCUdbg} invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.
 A similar switch occurs from $P_{\mathit{read\_critical}}$ back to $P_{\mathit{null}}$ at a \texttt{rcu\_read\_unlock}.
-Specialization that distinguishes between the inside and outside of read-side critical sections allows \texttt{RCUdbg} to do heavy-weight checks on functions like \texttt{memset}, but only when \texttt{memset} is invoked from within a read-side critical section.
+Specialization that distinguishes between the inside and outside of read-side critical sections allows \toolname{RCUdbg} to do heavy-weight checks on functions like \texttt{memset}, but only when \texttt{memset} is invoked from within a read-side critical section.
 
@@
 If two read-side critical sections are nested then switching from $P_{\mathit{read\_critical}}$ to $P_{\mathit{null}}$ on the first \texttt{call} to \texttt{rcu\_read\_unlock} meant that our tool would lose track of being in the context of the outer read-side critical section.
-\texttt{RCUdbg} solves this problem by tracking the nesting depth of RCU read-side critical sections using policies, e.g.
+\toolname{RCUdbg} solves this problem by tracking the nesting depth of RCU read-side critical sections using policies, e.g.
 $P_{\mathit{read\_critical}_1}$, $P_{\mathit{read\_critical}_2}$, etc.
@@
 
-\section{Application: RCUdbg}
+\section{Application: RCUdbg}\label{sec:rcudbg}
 
@@
 
-Granary's \texttt{RCUdbg} tool is designed to catch RCU bugs related to kernel code misusing pointers to RCU-protected data structures.
+Granary's \toolname{RCUdbg} tool is designed to catch RCU bugs related to kernel code misusing pointers to RCU-protected data structures.
 The following is a non-exhaustive list of requirements that must be met by reader and writer threads operating on an RCU-protected data structure.
-The list omits a description of how RCU-protected data structures are garbage collected, which is beyond the scope of the bugs caught by \texttt{RCUdbg}.
+The list omits a description of how RCU-protected data structures are garbage collected, which is beyond the scope of the bugs caught by \toolname{RCUdbg}.
 
@@
 
-\texttt{RCUdbg} is primarily concerned with finding bugs related to RCU-rpotected pointer misuses by reader threads (both inside and outside of read-side critical sections).
+\toolname{RCUdbg} is primarily concerned with finding bugs related to RCU-rpotected pointer misuses by reader threads (both inside and outside of read-side critical sections).
 As mentioned above, RCU bugs sometimes manifest as use-after-free bugs.
@@
 This kind of bug is a time bomb waiting to go off.
-Fortunately, \texttt{RCUdbg} catches the bug in \Cref{fig:rcu_existence_bug} right when \texttt{elm->data} is accessed.
+Fortunately, \toolname{RCUdbg} catches the bug in \Cref{fig:rcu_existence_bug} right when \texttt{elm->data} is accessed.
 
@@
 
-\texttt{RCUdbg} detects the the following RCU bugs related to misuses of RCU-protected pointers.
+\toolname{RCUdbg} detects the the following RCU bugs related to misuses of RCU-protected pointers.
 \begin{itemize}[leftmargin=3.2em]
@@
 
-\texttt{RCUdbg} gains visibility on RCU operations in two ways: \begin{enumerate}
-	\item The Linux kernel RCU implementation is annotated to provide more information to \texttt{RCUdbg}.
-This information includes source code file and line information for bug reporting, and additional memory references that \texttt{RCUdbg} uses to get better visibility on bugs.
+\toolname{RCUdbg} gains visibility on RCU operations in two ways: \begin{enumerate}
+	\item The Linux kernel RCU implementation is annotated to provide more information to \toolname{RCUdbg}.
+This information includes source code file and line information for bug reporting, and additional memory references that \toolname{RCUdbg} uses to get better visibility on bugs.
 These annotations are strictly internal to the RCU API, and thus require no changes to code using this API.
@@
 
-%\texttt{RCUdbg} interposes on an annotations are added into the Linux kernel's RCU implementation. These annotations
+%\toolname{RCUdbg} interposes on an annotations are added into the Linux kernel's RCU implementation. These annotations
 

### 2013-12-05 VEE-14/paper.tex
Addressed Ashvin's first set of comments on pages 1 and 2. Next set of comments will be more challenging to address :-/

@@
 
-The rest of the paper describes Granary in more detail. \Cref{sec:dbt} describes how Granary is implemented, \Cref{sec:what} describes how Granary can specialize instrumentation
+The rest of the paper describes Granary in more detail.
+\Cref{sec:dbt} describes how Granary is implemented, \Cref{sec:what} describes how a Granary tool can choose to instrument only specific code, \Cref{sec:how} describes how Granary change the policy used to instrument code, etc.
 
@@
 Granary translates and instruments kernel binaries one basic block at a time.
-In Granary, a basic block is a sequence of instructions ending in a conditional branch, \texttt{ret}, or \texttt{jmp}, but not a \texttt{call} instruction.
+In Granary, a basic block is a sequence of x86-64 instructions ending in a conditional branch, \texttt{ret}, or a \texttt{jmp} instruction.
+Unlike DRK and DynamoRIO, Granary does not terminate basic blocks at \texttt{call} instructions.
+This enables longer basic blocks and native-like proceccor return address prediction, but makes Granary non-transparent (\Cref{sec:transparency}).
 Entrypoint basic blocks (e.g.
 interrupt vectors, system call entrypoints, module initializers) are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as execution ``discovers'' them.
-Translated basic blocks are linked together and stored in a globally accessible \emph{code cache}.
+
+Translated basic blocks are linked together and stored in a \emph{code cache}.
+Basic blocks are indexed by $(pc, policy)$ pairs, where $pc$ is the native program counter associated with the beginning of the basic block's instructions, and $policy$ is the instrumentation policy used to instrument the native instructions starting at $pc$.
+The index is maintained by a globally accessible hash table.
 
@@
 When instrumented code yields to Granary, a ``context switch'' occurs that transfers execution to a CPU-private stack where Granary operates.
-Granary context-switches back to the code cache when the next basic block has been found or translated so that instrumented execution may continue.
+Granary operates on a CPU-private stack for two reasons.
+First, if Granary did not switch stacks, then context switches into Granary risk overflowing the runtime call stack.
+This can happen because kernel stacks are typically small (4K to 8K in size), and because a context switch could happen when the native stack is nearly exhausted.
+Second, on an SMP, multiple processors can simultaneously execute Granary code (e.g.
+to translate code that will later execute).
+In such cases, Granary depends on CPU-private data structures to avoid contention, and the stack is one such data structure.
+A context-switch back to the code cache occurs when the next basic block has been found or translated.
+At this point, instrumented execution continues.
 Similar to other DBT systems, Granary uses caching and hot code patching to reduce the number of context switches.
@@
 
-\subsection{Basic Blocks}
-
-Basic blocks in Granary's code cache contain x86-64 binary instructions.
-Associated with each basic block is meta-data describing the instructions of the basic block, and the instrumentation policy used to instrument those instructions.
-Basic block meta-data includes: \begin{inparaenum}[i)]
-	\item the length in bytes of the original and translated basic blocks;
-	\item the address of the first instruction in the original basic block; and
-	\item the instrumentation policy.
-\end{inparaenum} Basic block meta-data is queried and extended by Granary instrumentation tools.
-For example, Granary's control-flow graph build tool, \texttt{cfg}, extends the basic block meta-data to track per-block execution counts.
-Block meta-data is also queried by interrupt handlers when deciding how to handle interrupts and exceptions in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block.
+%Granary code executes with interrupts disabled and depends on CPU-private data to avo
+%Granary's translation mechanism depends on CPU-private data to avoid contention when two or more processors simultaneously execute Granary code.
+%because kernel stacks are typically small (4K to 8K in size), and because context-switches into Granary can occur at arbitrary program points, Granary cannot risk 
+%and a context switch into Granary might occur at any point. If most of the stack is already used by current thread before it yields control to Granary then
+
+\subsection{Basic Block Meta-Data}
+
+Associated with each basic block is meta-data that describes the basic block. Basic block meta-data includes: \begin{itemize}
+	\item The length in bytes of the original and translated basic blocks, and the address of the first instruction in the original basic block and the instrumented basic block.
+This information helps provide a rich debugging interface for instrumentation tool authors.
+Granary is unique among DBT tools in that the debugging experience is actually pleasant.
+For example, Granary has a built-in debugging mechanism for stepping back in time through previously executed basic blocks, as well as querying past register state.
+	\item Linux kernel-specific exception table meta-data about instructions contained within the current basic block.
+This kernel-specific meta-data is used by Granary to implement transparent page fault exception handling (\Cref{sec:interrupt_return_address_transparency}).
+	\item Ranges of instructions that must be executed atomically. Granary uses this information to implement selective interrupt delaying (\Cref{sec:interrupt_delay}).
+	\item The allocator used to allocate the instructions of the basic block. Granary uses this information to implement its trace-allocation optimization (\Cref{sec:trace_alloc}).
+	\item The instrumentation policy used to instrument the basic block.
+Granary uses this information to allow instrumentation tools to define policy-specific interrupt handling routines.
+For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if a code executing within a read-side critical section is interrupted.
+\end{itemize} %
+%Granary uses basic block meta-data for trace allocation (\Cref{sec:trace_alloc}), interrupt delaying (\Cref{sec:interrupt_delay}), and policy-specific interrupt handling.
+
+Basic block meta-data can be also be extended by instrumentation tools.
+This makes developing instrumentation tools that record per-basic block statistics very easy to make.
+For example, Granary's control-flow graph tool, \toolname{cfg}, extends the basic block meta-data to track per-block and per-edge execution counts.
+This data is useful because it provides precise code coverage information (block and branch coverage), as well as useful data for profile-guided optimization.
+This meta-data is specified by \toolname{cfg} and automatically allocated for \toolname{cfg} by Granary.
+
+%Basic block meta-data can also contain Granary's intermediate representation for x86-64 instructions.
+This is a unique feature that enables instrumentation tools to implement their own form of hot-code patching (among other things), as the intermediate representation remains tied to an emitted x86-64 instruction once the block is committed to the code cache.
+
+%Unlike any other instrumentation tool, Granary allows the x86-64 instruction intermediate representation
+
+%Another
+
+%novelty of Granary is that basic block meta-data enables a good
+
+%Basic block meta-data is queried and extended by Granary instrumentation tools.
+
+%// TODO!
+
+%Block meta-data is also queried by interrupt handlers when deciding how to handle interrupts and exceptions in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block.
+This enables policy-specific interrupt handling.
+That is, when a basic block instrumented by a specific policy is interrupted, that policy can choose how to handle the interrupt.
+For example, Granary's address watchpoints framework can instrument code using a lightweight, \toolname{null} instrumentation.
 
 \subsection{Transparency}\label{sec:transparency}
-
 An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}.
@@
 
-\subsubsection{Function return addresses}\label{para:return_address_transparency} Unlike most DBT frameworks, Granary inlines \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses.
-Inlining \texttt{call} instructions avoids return-address mispredictions and unnecessary control-flow by building longer basic blocks.
-
-\subsubsection{Interrupt return addresses} Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts.
+\subsubsection{Function return addresses}\label{sec:return_address_transparency}
+Unlike most DBT frameworks, Granary includes \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses.
+The benefit of including \texttt{call} instructions is better return-address prediction by the processor and longer basic blocks, which provide more code visibility to instrumentation tools at instrumentation time.
+
+\subsubsection{Interrupt return addresses}\label{sec:interrupt_return_address_transparency}
+Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts.
 A special case arises for page fault exceptions that occur within specific kernel functions that access user space data \cite{btkernel}.
-The Linux kernel records ranges of code addresses that are permitted to fault in an ``exception table'' data structure, and on a page fault exception, checks if the interrupt return address belongs to one of the pre-defined ranges.
+The Linux kernel records ranges of code addresses that are permitted to fault in an ``exception table'' data structure, and on a page fault exception, checks if the exception return address belongs to one of the pre-defined ranges.
 Granary records any exception table entries as part of each basic block's meta-data, so that a page fault in a basic block can be mapped to the correct native faulting program counter.
@@
 For example, hardware interrupts can redirect control flow at any instruction, and interrupt delivery cannot be delayed past the execution of memory instructions that might alter the interrupt delivery state \cite{DRK}.
-Preserving interrupt transparency, however, requires that interrupts be delayed until the end of any injected instrumentation instructions.
+Preserving interrupt transparency, however, requires that interrupts either be delayed until the end of any injected instrumentation instructions.
 Together, these requirements make it very difficult to write instrumentation tools.
 
-\subsubsection{Code cache}
-Granary's code cache and JIT translation mechanisms are fully reentrant.
-Context-switches into Granary disable interrupts and switch execution onto a CPU-private stack.
-Context switches out of Granary and into the code cache or native code returns execution to the native stack and re-enables interrupts.
-This approach presents a minor challenge: context-switches save or restore the entire register state, so as not to disturb the execution of the instrumented program.
+%Another concern is that instrumentation itself is not re-entrant. For example, instrumentation instructions added by a tool that access CPU-private information
+
+\subsubsection{Granary}
+For efficiency, most kernel-space instrumentation systems (Granary included) operate on CPU-private state (e.g.
+Granary's CPU-private stacks).
+However, accessing CPU-private without first disabling interrupts results in undefined behavior.
+Therefore, context-switches into Granary first disable interrupts and then switch execution onto one of Granary's CPU-private stacks.
+Context switches out of Granary and into native code (\Cref{sec:detach}) or the code cache returns execution to the native stack and re-enable interrupts.
+Execution behavior is preserved because Granary saves the machine state (i.e.
+registers, flags) in a reentrant way.
+All other parts of Granary are fully reentrant.
 
@@
 
-\subsubsection{Instrumentation}
+\subsubsection{Instrumentation}\label{sec:interrupt_delay}
 
 One challenge with any DBT system is how it copes with non-reentrant instrumentation.
-For example, instrumentation that modifies CPU-private data structures cannot be safely executed by pre-emptive kernel running on a multi-core system.
-If the task being instrumented is interrupted, then the kernel might decide to re-schedule the task to resume its execution on a different CPU.
+For example, instrumentation that modifies CPU-private data structures cannot be safely executed by a pre-emptive kernel running on an SMP.
+If the thread being instrumented is interrupted, then the kernel might decide to re-schedule the task to resume its execution on a different CPU.
 Any modifications to CPU-private state would lead to undefined behavior.
@@
 \begin{figure}[t!]
-\hspace{4em}\includegraphics[width=0.9\columnwidth]{diagrams/ibl.pdf}\vspace{-3em}
+\vspace{-1.5em}\hspace{4em}\includegraphics[width=0.9\columnwidth]{diagrams/ibl.pdf}\vspace{-2.5em}
 \caption{\label{fig:ibl}Process of resolving an indirect branch.
@@
 
-\subsubsection{Trace Allocation}
+\subsubsection{Trace Allocation}\label{sec:trace_alloc}
 Arranging for commonly executed sequences of basic blocks to appear adjacent in memory, also called tracing, typically improves the runtime performance of instrumented programs \cite{DynamoRIO}.
@@
 
-Trace allocation is implemented by recording the memory pool used to allocate a given basic block in that basic block's metadata.
-When Granary needs to translate a new basic block, it inspects the predecessor basic block's metadata and uses the predecessor basic block's memory pool to service allocation requests for the basic block in translation.
+Trace allocation is implemented by recording the memory pool used to allocate a given basic block in that basic block's meta-data.
+When Granary needs to translate a new basic block, it inspects the predecessor basic block's meta-data and uses the predecessor basic block's memory pool to service allocation requests for the basic block in translation.
 
@@
 
-\section{Specializing on What to Instrument}
+\section{Specializing on What to Instrument}\label{sec:what}
 
@@
 
+\label{sec:detach}
 Granary \emph{detaches} when control transfers from instrumented code to native (uninstrumented) code. Detaching occurs in one of two ways: \begin{enumerate}

### 2013-12-05 VEE-14/paper.tex
Decided to elaborate on return address transparency. Specifically, on why Granary does not end basic blocks at CALL instructions.

@@
 Granary uses this information to allow instrumentation tools to define policy-specific interrupt handling routines.
-For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if a code executing within a read-side critical section is interrupted.
+For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if a code executing within a read-side critical section is interrupted (\Cref{sec:rcudbg}).
 \end{itemize} %
@@
 This makes developing instrumentation tools that record per-basic block statistics very easy to make.
-For example, Granary's control-flow graph tool, \toolname{cfg}, extends the basic block meta-data to track per-block and per-edge execution counts.
+For example, Granary's control-flow graph tool, \toolname{cfg}, extends the basic block meta-data to track per-block and per-edge execution counts (\Cref{sec:cfg}).
 This data is useful because it provides precise code coverage information (block and branch coverage), as well as useful data for profile-guided optimization.
@@
 \subsubsection{Function return addresses}\label{sec:return_address_transparency}
-Unlike most DBT frameworks, Granary includes \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses.
-The benefit of including \texttt{call} instructions is better return-address prediction by the processor and longer basic blocks, which provide more code visibility to instrumentation tools at instrumentation time.
+Unlike most DBT frameworks, Granary's basic blocks do not terminate at \texttt{call} instructions.
+This exposes code cache addresses to instrumented module code in the form of return addresses stored on the runtime call stack.
+The benefit of including \texttt{call} instructions are: \begin{description}[labelindent=\parindent]
+
+	\item[i) Zero overhead procedure returns.] Unlike Granary, DRK emulates procedure calls and returns.
+DRK emulates \texttt{call} instructions by pushing the intended native return address onto the runtime call stack, followed by a direct direct jump to the next (instrumented) basic block \cite{DRK}.
+This approach enables return address transparency, but introduces additional overheads on procedure returns, which must do costly indirect branch lookups to find the intended return target.
+
+Btkernel's approach is similar in spirit to Granary's; however, btkernel ends basic blocks at \texttt{call} instructions.
+This introduces unnecessary control-flow on the return path from procedures.
+That is, btkernel's ``call/ret optimization" ostensibly enables zero-overhead procedure returns by leaving \texttt{ret} instructions as-is; however, the benefits are more apparent than real.
+To transfer control to the next logical basic block after a procedure return, btkernel's \texttt{ret}s first transfer control to the end of the basic block that performed the \texttt{call}, and then \texttt{jmp} to the next basic block.
+Granary avoids this unnecessary \texttt{jmp} because basic blocks do not end at \texttt{call} instructions.
+
+In some cases, emulating DRK's return-address lookup mechanism is desirable.
+For example, if Granary attaches while some program's execution is underway (and so the runtime call stack contains native return addresses), then a direct \texttt{ret} back to one of the native return addresses might be undesirable (e.g.
+if the tool wants comprehensive instrumentation of all code).
+In such cases, a tool can partially specialize instrumentation according to whether or not the return address might be in the code cache.
+Partial specialization is enabled by \emph{policy properties}, which are transparently inherited along with normal instrumentation policy information.
+In this case, the policy property can be used to tell Granary to emulate \texttt{ret} instructions when the return address is potentially a native address, and directly \texttt{ret} when the return address is known to be in the code cache.\footnote{This feature is used when Granary is configured to instrument user space programs.
+User space Granary depends on the Linux dynamic loader to inject Granary into a running process.
+When injected, a Granary takeover function is executed on an already active native runtime call stack.
+Granary's takeover procedure transparently emulates returns, thus taking over the previously fully-native call stack.
+This feature allows instrumentation tools to focus on their particular objectives and not the intricacies of thread takeover.}
+
+%performing a \texttt{ret} requires an additional \texttt{jmp} to 
+%Neither of DRK's nor btkernel's approaches enable native-like return address prediction by the CPU.
+%and then jumping to the next basic block.
+This in turn .
+Btkernel's approach to procedure calls introduces similar inefficiencies, despite the fact that does not provide return address transparency, because it also emulates procedure calls, but by manually pushing the instrumented return address onto the stack instead of the native one.
+%Granary's approach improves on DRK and btkernel by allowing 
+%Better return-address prediction by the processor.
+
+	\item[iii) Improved visibility.] Not ending basic blocks on \texttt{call} instructions results in longer basic blocks, which in turn provide additional visibility to tools at instrumentation time into what will be executed by the current thread.
+\toolname{RCUdbg} benefits from this additional visibility because it can observe the end of read-side critical sections (\texttt{call}s to \texttt{rcu\_read\_unlock}), and optimistically assume that functions executing before the read-side critical section ends will execute within the read-side critical section.
+
+	\item[ii) Zero-overhead attaching.] (\Cref{sec:implicit_attach})
+
+	%longer basic blocks, which provide more code visibility to instrumentation tools at instrumentation time.
+\end{description}
 
@@
 
+\label{sec:implicit_attach}
 After Granary bootstraps on the module initialization process, \emph{attaching} occurs in one of three ways: \begin{enumerate}
@@
 Because Granary's translation mechanism depends only on CTI policy and target information, Granary is able to ensure that policy information is never lost or corrupted, even in the face of concurrent executions of the same module code, arbitrary pre-emption, and arbitrary resumption.
+
+\section{Application: Control-Flow Graphs}\label{sec:cfg}
 

### 2013-12-05 VEE-14/paper.tex
Working on the first few comments on ashvin's second review of the paper.

@@
 
-\subsection{Basic Block Meta-Data}
+\subsection{Basic Block Meta-Data}\label{sec:metadata}
 
@@
 
-	\item[iii) Improved visibility.] Not ending basic blocks on \texttt{call} instructions results in longer basic blocks, which in turn provide additional visibility to tools at instrumentation time into what will be executed by the current thread.
+	\item[ii) Improved visibility.] Not ending basic blocks on \texttt{call} instructions results in longer basic blocks, which in turn provide additional visibility to tools at instrumentation time into what will be executed by the current thread.
 \toolname{RCUdbg} benefits from this additional visibility because it can observe the end of read-side critical sections (\texttt{call}s to \texttt{rcu\_read\_unlock}), and optimistically assume that functions executing before the read-side critical section ends will execute within the read-side critical section.
-
-	\item[ii) Zero-overhead attaching.] (\Cref{sec:implicit_attach})
 
@@
 
+\begin{figure}[t!]
+\begin{multicols}{2}
+\vspace{-1.5em}\subfloat[\label{fig:interrupt_delay_drk}Pre and post instrumentation in DRK.
+Injected instrumentation blocks are treated as atomic with respect to interrupts.]{\includegraphics[height=10em,clip=true,trim=0 5em 10em 0]{diagrams/drk_instrumented_mov.pdf}}\vspace{-2.5em}
+\columnbreak
+\vspace{-1.5em}\hspace{-0.5em}\subfloat[\label{fig:interrupt_delay_granary}Pre and post instrumentation in Granary.
+The two instrumentation blocks, as well as the native instruction are wrapped in a delay region.]{\includegraphics[height=10em,clip=true,trim=0 5em 10em 0]{diagrams/granary_instrumented_mov.pdf}}\vspace{-2.5em}
+\end{multicols}
+\caption{Example showing equivalent pre- and post-instruction instrumentation with DRK (left) and Granary (right).
+In DRK, the injected instrumentation instructions (grey) are guaranteed to execute atomically, up until the next native instruction (white).
+In Granary, no guarantees are made on injected instrumentation, unless otherwise instructed.
+In this case, a delay region surrounds both the pre- and post-instrumentation, as well as the native instruction, guaranteeing that the entire sequence of instructions execute on the same CPU.}
+\end{figure}
+
 One challenge with any DBT system is how it copes with non-reentrant instrumentation.
@@
 DRK's solution to this problem assumes that all injected instrumentation instructions are potentially non-reentrant, and \emph{delays} interrupts occuring within instrumented code \cite{DRK}.
-Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86: interrupts only arrive on \emph{logical} instruction boundaries, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
+Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86: interrupts and exceptions only arrive on logical instruction boundaries in the native code, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
 This approach introduces subtle complexity when creating instrumentation that injects instructions both before (pre) and after (post) native instructions.
-For example, an interrupt occuring in the pre-instrumentation of a native instruction will be delayed only until the native instruction.
-When the interrupt is handled and execution attempts to return to the interrupted instructions, DRK re-instruments the ``tail" of interrupted basic block, starting from the native instruction.
-Instrumenting this tail involves injecting pre- and post-instrumentation around the first native instruction.
-If the pre-instrumentation is not idempotent, then its re-execution will result in undefined behavior.
+
+\Cref{fig:interrupt_delay_drk} shows an example of pre- and post-instrumentation around a \texttt{mov} instruction.
+If an interrupt arrives within the pre-instrumentation, then DRK delays they delivery of the interrupt until the native instruction (\texttt{mov}).
+After the interrupt is handled and execution attempts to return to the interrupted instructions, DRK will re-instrument the ``tail" of interrupted basic block, starting from the \texttt{mov} instruction.
+This involves re-creating the pre- and post-instrumentation of the \texttt{mov} in the tail basic block.
+When executed, this will result in an unbalanced number of \texttt{push} and \texttt{pop} instructions.
+That is, \texttt{push \%rax} is executed in both the original and tail basic block, but \texttt{pop \%rax} is only executed in the tail basic block.
+This is a specific example of a more general problem: DRK's approach to interrupt delaying does not play well with idempotent instrumentation.
+
+%If the pre-instrumentation is not idempotent, then its re-execution will result in undefined behavior.
 We encountered this issue when using DRK for kernel module instrumentation, where our pre-instrumentation spilled registers to the runtime call stack, and our post-instrumentation restored the spilled registers.
@@
 Granary charts a middle ground between the two aforementioned approaches.
-Interrupt delaying has value because it enables non-reentrant instrumentation to be injected without costly disabling/enabling of interrupts, while not supporting delaying has value because not all instrumentation is non-reentrant.
+Most instrumentation is re-entrant, partially because kernel instrumentation tools can depend on the runtime call stack for scratch storage space.
+However, interrupt delaying is sometimes useful because it enables non-reentrant instrumentation to avoid to continually disable and re-enable interrupts.
+For example, Granary's basic block tracing tool writes to a CPU-private buffer containing the sequence of executed basic block program counters.
+Operating on a CPU-private buffer is more efficient than operating on a global buffer because it avoids cache line contention and locking on an SMP.
+However, writing to a CPU-private buffer is only safe if the entire write executes on the same CPU.
+An equivalent tool developed for btkernel---which does not support interrupt delaying---would have to disable and the re-enable interrupts around the write.
+The is inefficient because it introduces flag stalls into the instruction pipeline \cite{AgnerMicroarchitecture}, and because interrupts arrive relatively infrequently.
+
+Granary conveniently solves this problem by allowing instrumentation tools to declare \emph{interrupt delay regions}.
+An interrupt delay region is a sequence of instructions (which can contain internal control flow) inside of a basic block that must be executed atomically.
+\Cref{fig:interrupt_delay_granary} shows an interrut delay region that surrounds both the pre- and post-instrumentation of a native \texttt{mov} instruction.
+If an interrupt arrives anywhere within this region of code, Granary will delay the interrupt to the end of the region.
+The declarative nature of interrupt delay regions allows instrumentation developers to opt-in to interrupt delaying, and control exactly when, where, and for how long interrupts can be delayed.
+
+Granary implements interrupt delay regions by recording bounds information about interrupt delay regions inside of each basic block's meta-data (\Cref{sec:metadata}).
+When a basic block is interrupted, Granary inspects the block's meta-data and determines if the interrupt must be delayed.
+If a delay is required, then all instructions belonging to the delay range are copied into a CPU-private buffer.
+All instructions are copied because a delay region might contain internal control-flow.
+A CPU-private buffer is used because the number of CPUs is a natural upper bound on the number of simultaneously executing delay regions.
+After the copied instructions, Granary emits code that code re-creates the interrupt stack frame, followed by a jump to the interrupt vector (which is controlled by Granary).
+This jump ensures that Granary's interrupt handler regains control after the instructions within the delay region fully execute, but before the kernel observes the interrupt.
+When Granary regains control, it can then delegate to the instrumentation tool to handle the interrupt, or defer handling of the interrupt to the kernel.
+
+When all instructions have been emitted to the CPU-private buffer, Granary emulates a return from the interrupt (without issuing an \texttt{iret} instruction).
+The emulated return leaves interrupts disabled and transfers control to the CPU-private buffer.
+Leaving interrupts disabled ensures that the remainder of the delay region will execute on the current CPU, and avoids reentrancy issues related to the copied region being located in a CPU-private buffer.
+
+% and re-issues the interrupt to Granary.
+Granary then emulates a return from the current interrupt handler (without issuing an \texttt{iret} instruction), but leaves interrupts disabled and redirects execution into the copied code.\footnote{Leaving interrupts disabled is required to ensure that executing the copied instructions in the CPU-private buffer is reentrant.
+Granary performs some sanity checks on the instructions being copied to ensure that they don't re-enable interrupts.} When the copied code finishes, Granary regains control and observes the reconstructed interrupt stack frame.
+At this point, Granary or the tool can choose to handle the interrupt or defer the handling of the interrupt to the kernel.
+
+%while not supporting delaying has value because not all instrumentation is non-reentrant.
 By default, Granary does not delay interrupts.
-However, instrumentation tools can declare \emph{interrupt delay ranges}: sequences of instructions (which can contain internal control flow) inside of a basic block that must be executed atomically.
-Granary encodes bounds information about interrupt delay ranges inside of each basic block's meta-data.
-When a basic block is interrupted, Granary inspects the meta-data and determines if the interrupt must be delayed.
-If a delay is required, then the instructions belonging to the delay range are copied into a CPU-private buffer, followed by code that re-constructs the interrupt stack frame and re-issues the interrupt to Granary.
-Granary then emulates a return from the current interrupt handler (without issuing an \texttt{iret} instruction), but leaves interrupts disabled and redirects execution into the copied code.
-When the copied code finishes, Granary regains control and observes the reconstructed interrupt stack frame.
-At this point, Granary or the tool can choose to handle the interrupt or defer the handling of the interrupt to the kernel.
+However, instrumentation tools can declare \emph{interrupt delay ranges}:
 
@@
 
-\subsection{Implementation}
+\subsection{Control Flow Implementation}
 \subsubsection{Direct Control Flow}
@@
 Like btkernel, Granary translates direct control-flow instructions (CTIs) into equivalent jumps into CTI-specific ``edge code" (\Cref{fig:direct_edge_code}).
-Edge code lives outside the code cache, so as to not pollute the icache \cite{btkernel}.
+Edge code lives outside the code cache, so as to not pollute the processor's instruction cache \cite{btkernel}.
 
@@
 The indirect \texttt{call} transfers control to Granary via a function pointer located in an ``edge data structure", which describes how to translate the target basic block, $B_{succ}$.
-Granary gains control and looks up or translates the target code according instrumentation policy information stored in the edge data structure.
-Then, Granary uses hot-code patching (\Cref{sec:patching}) to modify the original CTI that led to the edge code \texttt{call}, so that it now transfers control to $B_{succ}$ instead.
-Finally, Granary releases control and returns to $B_{succ}$.
+Granary gains control and looks up or translates the target code according instrumentation policy stored in the edge data structure.
+Then, Granary uses modifies the original CTI in $B_{pred}$ in place so that it transfers control to $B_{succ}$  (\Cref{sec:patching}).
+Finally, Granary releases control and context switches to $B_{succ}$.
 
@@
 \subsection{Optimizations}
+
 \subsubsection{Ahead-of-time Translation} \label{sec:aot}
@@
 \subsubsection{Patching and Alignment}\label{sec:patching}
-
 For efficiency, basic blocks in Granary's code cache are directly connected through direct control flow instructions (\texttt{call}, \texttt{jmp}, \texttt{jCC}).
@@
 %To reduce the overhead of alignment 
-
 %In some cases, basic blocks can be directly connected during basic block translation time.
@@
 However, not all basic blocks can be connected this way, therefore some runtime
-
 %Granary connects some basic blocks at translation time (e.g.
@@
 However,
-
 %Some basic blocks are directly connected at basic block translation time.
@@
 Another example is a branch to an already translated basic block.
-
 %Other basic blocks are only connected at runtime.
 For example, if the function \texttt{memset} has never been executed, then a \texttt{call} to \texttt{memset} will be translated to target edge code (\Cref{fig:direct_edge_code}) that, when executed, will translate \texttt{memset}'s first basic block and patch the \texttt{call} to target the translated basic block.
-
 %For example, the targets of function call instruction (\texttt{call}) are resolved just-in-time (unless the target is already known).
 For each translated \texttt{call} instruction, Granary generates edge code (\Cref{fig:direct_edge_code}) that will find or translate the first basic block of the function being called, and patch the translated \texttt{call} instruction to point to the function's first basic block.
-
 %For example, if a direct branch in one basic block targets another block within the trace under construction, then that 
-
 %Basic blocks are directly connected at block translation time when branches in traces re-enter a trace, or when branches target previously translated basic blocks.
 If a branch does not target another basic block within its trace, or does not target a previously translated basic block, then the branch will target edge code that will resolve that branches target at runtime.
-
 % However, when a given basic block is translated, not all of its targets are known (e.g.
@@
 Therefore, edge code (\Cref{fig:direct_edge_code}) is generated that translates/resolves a branch target when that branch is first executed, and patches the translated branch to avoi
-
 % The targets of some of these direct CTIs are only resolved at runtime, by executing edge code.
 As an optimization, jumps to edge code are dynamically patched to transfer control to their intended targeted basic blocks.
-
 %edge code use hot code patching to update a translated CTI to redirect execution
@@
 %Directing control through the shadow system call table avoids the overhead of going through the indirect branch lookup (IBL) table (\Cref{sec:ibl}), and reduces the number of potential conflicts in the IBL table.
-
 % whose entries are code cache basic block addresses translated according to the starting instrumentation policy.
-
 %\subsection{Reproducability}
@@
 \section{Specializing on What to Instrument}\label{sec:what}
+
+\begin{figure}
+\captionsetup[subfloat]{width=0.45\columnwidth}
+\begin{tabularx}{\columnwidth}{XX}%
+\subfloat[Full kernel instrumentation.]{\hspace{2.5em}\epsfig{file=diagrams/kernel.eps,width=0.125\textwidth}} & %
+\subfloat[Module-only instrumentation]{\hspace{2.5em}\epsfig{file=diagrams/module.eps,width=0.125\textwidth}} \\
+\subfloat[Module-only instrumentation, with a function specialized to run natively by using a module function wrapper.]{\hspace{2.5em}\epsfig{file=diagrams/module_wrapper.eps,width=0.125\textwidth}} & %
+\subfloat[\label{fig:module_probe}Module-only instrumentation, with a kernel function specialized to also be instrumented by using a probe-based wrapper.]{\hspace{2.5em}\epsfig{file=diagrams/module_probe.eps,width=0.125\textwidth}}
+\end{tabularx}
+\caption{Example of using different granularities of instrumentation to selectively instrument or not instrument specific code.}
+\end{figure}
 
@@
 
-While sufficient for comprehensiveness, hardware page protection is an inefficient mechanism for gaining control of module code execution.
+While correct, hardware page protection is an inefficient mechanism for gaining control of module code execution.
 That is, after the (instrumented) module initializer runs to completion, Granary needs an efficient way to regain control of future invocations of the module's code.
@@
 \subsection{Functions}\label{sec:function_wrapper}
-
-\begin{figure}
-\captionsetup[subfloat]{width=0.45\columnwidth}
-\begin{tabularx}{\columnwidth}{XX}%
-\subfloat[Full kernel instrumentation.]{\hspace{2.5em}\epsfig{file=diagrams/kernel.eps,width=0.125\textwidth}} & %
-\subfloat[Module-only instrumentation]{\hspace{2.5em}\epsfig{file=diagrams/module.eps,width=0.125\textwidth}} \\
-\subfloat[Module-only instrumentation, with a function specialized to run natively by using a module function wrapper.]{\hspace{2.5em}\epsfig{file=diagrams/module_wrapper.eps,width=0.125\textwidth}} & %
-\subfloat[\label{fig:module_probe}Module-only instrumentation, with a kernel function specialized to also be instrumented by using a probe-based wrapper.]{\hspace{2.5em}\epsfig{file=diagrams/module_probe.eps,width=0.125\textwidth}}
-\end{tabularx}
-%
-%\begin{multicols}{2}
-%\subfloat[Full kernel instrumentation.]{\epsfig{file=diagrams/kernel.eps,width=0.125\textwidth}}
-%\columnbreak
-%\subfloat[Module-only instrumentation]{\epsfig{file=diagrams/module.eps,width=0.125\textwidth}}
-%\end{multicols}
-%\begin{multicols}{2}
-%\subfloat[Module-only instrumentation, with a function specialized to run natively by using a module function wrapper.]{\epsfig{file=diagrams/module_wrapper.eps,width=0.125\textwidth}}
-%\columnbreak
-%\subfloat[Module-only instrumentation, with a kernel function specialized to also be instrumented by using a probe-based wrapper.]{\epsfig{file=diagrams/module_probe.eps,width=0.125\textwidth}}
-%\end{multicols}
-\caption{Example of using different granularities of instrumentation to selectively instrument or not instrument specific code.}
-\end{figure}
 

### 2013-12-05 VEE-14/paper.tex
Explained the reasoning behing out-edges, as a contrast to DRK and btkernel.

@@
 This exposes code cache addresses to instrumented module code in the form of return addresses stored on the runtime call stack.
-The benefit of including \texttt{call} instructions are: \begin{description}[labelindent=\parindent]
-
-	\item[i) Zero overhead procedure returns.] Unlike Granary, DRK emulates procedure calls and returns.
+The benefit of including \texttt{call} instructions are: \begin{enumerate}
+
+	\item {\bf Zero overhead procedure returns.} Unlike Granary, DRK emulates procedure calls and returns.
 DRK emulates \texttt{call} instructions by pushing the intended native return address onto the runtime call stack, followed by a direct direct jump to the next (instrumented) basic block \cite{DRK}.
@@
 
-	\item[ii) Improved visibility.] Not ending basic blocks on \texttt{call} instructions results in longer basic blocks, which in turn provide additional visibility to tools at instrumentation time into what will be executed by the current thread.
+	\item {\bf Improved visibility.} Not ending basic blocks on \texttt{call} instructions results in longer basic blocks, which in turn provide additional visibility to tools at instrumentation time into what will be executed by the current thread.
 \toolname{RCUdbg} benefits from this additional visibility because it can observe the end of read-side critical sections (\texttt{call}s to \texttt{rcu\_read\_unlock}), and optimistically assume that functions executing before the read-side critical section ends will execute within the read-side critical section.
@@
 	%longer basic blocks, which provide more code visibility to instrumentation tools at instrumentation time.
-\end{description}
+\end{enumerate}
 
@@
 \begin{minipage}[b]{4cm}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 cmp $0x0, %rax
@@
 \newsavebox\translatedjcc
-\begin{lrbox}{\translatedjcc}\begin{minipage}[b]{4cm}\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lrbox}{\translatedjcc}\begin{minipage}[b]{4cm}\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
   cmp $0x0, %rax
@@
 
-Granary translates indirect control flow instructions (\texttt{call}, \texttt{jmp}) into their equivalent direct control flow versions, where the direct versions jump to ``in-edge" code.
+Granary translates indirect CTIs (\texttt{call}, \texttt{jmp}) into their equivalent direct control flow versions, where the direct versions jump to ``in-edge" code.
 In-edge code loads the indirect jump target (e.g.
@@
 In this respect, out-edge code is chained together through control flow, just as colliding entries in a typical hash table can be chained together through a linked list.
+
+Granary's use of out-edge code avoids re-entrancy issues related to indirect control flow that present in DRK and btkernel.
+Both DRK and btkernel store the target of an indirect CTI in a CPU-private memory location.
+To complete an indirect control transfer, both systems perform (roughly) the following steps: \begin{enumerate}[itemindent=\parindent]%
+\setlength{\itemsep}{2pt}%
+\setlength{\parskip}{0pt}%
+	\item Save the flags and disable interrupts.
+	\item Save registers.
+	\item Lookup the next basic block address.
+	\item Store that address in a CPU private memory location.
+	\item Restore registers.
+	\item Restore the flags and interrupt state.
+	\item Indirectly jump to the target basic block through the CPU-private location.
+\end{enumerate} %
+However, the ordering of steps (6) and (7) introduce a reentrancy problem: the access to the CPU-private location is not safe because an interrupt might arrive before the indirect jump in step (7).
+This unusual design is motivated by the constraint that the entire lookup procedure must save (2) and restore (5) the register state for correctness, but must also save the target pc \emph{somewhere} so that execution can eventually be redirected.
+Granary's out-edge code avoids this reentrancy issue by having the out edge code itself (in the form of a direct \texttt{jmp}) contain the target pc.
 
@@
 \begin{multicols}{2}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 struct device_driver {
@@
 \columnbreak
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 TYPE_WRAPPER(struct device_driver, {
@@
 \lstset{language=C, tabsize=2, stepnumber=1}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 MODULE_FUNCTION_WRAPPER(__kmalloc, (size, gfp), {
@@
 \begin{lrbox}{\rcuexistencebug}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 rcu_read_lock();
@@
 \begin{lrbox}{\rcuderefbug}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 rcu_read_lock();
@@
 \begin{lrbox}{\rcuassignderefbug}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 spin_lock();
@@
 \lstset{language=C, tabsize=2, stepnumber=1}
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 // As used in the code:
@@
 %\begin{minipage}{0.6\columnwidth}
-%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 %elm = rcu_dereference(list_head);
@@
 %\begin{minipage}{0.8\columnwidth}
-%\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
 %elm = deref_taint(rcu_dereference(list_head));

### 2013-12-06 VEE-14/paper.tex
Working on a table describing different kinds of instrumentation scopes / strategies.

@@
 We chose DBT because \begin{inparaenum}[i)]
-	\item static binary analysis tools are unable to cope with \texttt{x86}'s mix of variably-sized instructions and data; 
+	\item static binary analysis tools are unable to cope with \texttt{x86-64}'s mix of variably-sized instructions and data; 
 	\item some kernel code is only distributed in a binary format (e.g. proprietary device drivers); and
@@
 DRK's solution to this problem assumes that all injected instrumentation instructions are potentially non-reentrant, and \emph{delays} interrupts occuring within instrumented code \cite{DRK}.
-Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86: interrupts and exceptions only arrive on logical instruction boundaries in the native code, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
+Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86-64: interrupts and exceptions only arrive on logical instruction boundaries in the native code, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
 This approach introduces subtle complexity when creating instrumentation that injects instructions both before (pre) and after (post) native instructions.
@@
 \begin{figure}[t!]
-\vspace{-1.5em}\hspace{4em}\includegraphics[width=0.9\columnwidth]{diagrams/ibl.pdf}\vspace{-2.5em}
+\hspace{4em}\includegraphics[height=13em,clip=true,trim=0 5em 10em 0]{diagrams/ibl.pdf}
 \caption{\label{fig:ibl}Process of resolving an indirect branch.
@@
 Granary atomically patches direct CTIs that target edge code to instead point to their (now resolved) target basic block.
-Due to limitations of the x86 architecture, atomic operations cannot straddle cache lines.
+Due to limitations of the x86-64 architecture, atomic operations cannot straddle cache lines.
 Therefore, patchable instructions (or parts thereof) must be aligned so as to not straddle two cache lines.
@@
 Therefore, when possible, Granary uses ignored instruction prefixes to alter an instruction's alignment.
-For example, Granary aligns patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86 processors \cite{AgnerMicroarchitecture}.
+For example, Granary aligns patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86-64 processors \cite{AgnerMicroarchitecture}.
 Finally, most basic blocks are small (i.e.
@@
 
-\begin{figure}
+\begin{table*}[ht!]
+\scriptsize
+\begin{tabularx}{\linewidth}{| l | >{\raggedright}m{0.3\linewidth} | X |}
+\hline
+\bf Name & \bf Description & \bf Example Tool \\
+\hline
+Full-kernel & All kernel code is instrumented.
+& Granary's \toolname{cfg} tool constructs inter- and intra-procedural control flow graphs and records per-basic block and per-edge performance counters.
+This data is useful for debugging an performance analysis.
+Applying the \toolname{cfg} tool to the whole kernel allows us to perform profile-guided optimization of kernel code, including a form of de-virtualization.
+That is, most indirect control flow instructions (\texttt{call}, \texttt{jmp}) within the kernel target only one destination.
+Fast checks can be inlined into kernel code that use direct control transfer instructions when possible, and fall-back on indirect control transfer instructions when necessary.\\
+\hline
+Module-only & All code from a specific kernel module or driver is instrumented.
+Granary detaches when module code invokes kernel code, and re-attaches when kernel code invokes module code.
+All non-module kernel code executes with zero overhead.
+& TODO \\
+\hline
+Probe-based & All code within a particular function is instrumented. Granary attaches when the function is invoked, and detaches when the function returns or invokes another function. & TODO \\
+\hline
+\end{tabularx}
+\caption{\label{tab:what_kinds}Different kinds to instrumentation strategies supported by Granary.}
+\end{table*}
+
+\begin{figure}[t]
 \captionsetup[subfloat]{width=0.45\columnwidth}
 \begin{tabularx}{\columnwidth}{XX}%
-\subfloat[Full kernel instrumentation.]{\hspace{2.5em}\epsfig{file=diagrams/kernel.eps,width=0.125\textwidth}} & %
+\subfloat[Full-kernel instrumentation.]{\hspace{2.5em}\epsfig{file=diagrams/kernel.eps,width=0.125\textwidth}} & %
 \subfloat[Module-only instrumentation]{\hspace{2.5em}\epsfig{file=diagrams/module.eps,width=0.125\textwidth}} \\
@@
 The translation process continues when a system call or interrupt occurs and follows the jump from duplicated into translated code.
-Granary only translate code that is known to operate on a kernel stack so that instrumentation tools can safely save transient data in a reentrant way by spilling that data to the stack.
+Granary only translate code that is known to operate on a kernel stack\footnote{Not all kernel code executes on a kernel stack.
+For example, when a user space application invokes the \texttt{syscall} instruction, control is transferred to the kernel, but no stack switch occurs.
+The kernel's early system call entrypoint code is responsible for switching execution onto a kernel stack.} so that instrumentation tools can safely save transient data in a reentrant way by spilling that data to the stack.
 

### 2013-12-06 VEE-14/paper.tex
Minor edits. Moved one of the sentences that Ashvin considered important up as the intro to the DBT subsection on direct control flow. It is perhaps more appropriate there.

@@
 Limiting the scope of what is instrumented is  important because the overhead introduced by using runtime code instrumentation is usually dominated by the instrumentation itself and not the baseline overhead of the runtime system supporting that instrumentation.
-Also, specializing based on what is instrumented allows instrumentation tools to avoid unnecessary filtering.
-That is, if one wants to instrument RCU read-side and write-side critical sections, as is the case with Granary's \toolname{RCUdbg} tool, then instrumenting all code just to filter out / find those sections is wasteful.
+Also, specializing based on what is instrumented allows instrumentation tools focus instrumentation on the code of interest.
+For example, if one wants to instrument RCU read-side and write-side critical sections, as is the case with Granary's \toolname{RCUdbg} tool, then instrumenting all code just to filter out / find those sections is wasteful.
 
@@
 
-Another limitation of existing systems is that they are unable to perform efficient runtime code specialization, i.e.
-they are unable to change how code is instrumented based on the context in which that code will run.
-This is unfortunate because it means that tools wanting to instrument code differently must create their own solution that works around this deficiency of existing systems.
-A typical solution is for tools to generate as many instrumented versions of the code as there are execution contexts.
-When instrumented code is  executing, runtime checks query some thread-private or CPU-private state, and jump to the correct version of the code for the current context.
-Runtime checks of this form are inefficient, and generating many versions of the code introduces bloat.
-The machinery needed to multiplex between different versions of instrumented code is also orthogonal to the goals of the instrumentation tool, thus distracting tool-developers from their primary objectives.
+Another limitation of existing systems is that they are unable to perform efficient runtime code specialization.
+That is, they are unable to change how code is instrumented based on the context in which that code will run.
+This is unfortunate because it means that tools wanting to instrument code differently must create their own solution that works around this deficiency.
+For example, one way that existing tools can specialize code is to generate as many versions of the code as there are potential specializations.
+When instrumented code is executing, runtime checks continually check and jump to the correct version of the code for the current execution context.
+However, runtime checks are inefficient, and generating many versions of the code introduces bloat and pollutes the CPU's instruction cache.
+Also, the runtime checks needed to multiplex between different versions of instrumented code is orthogonal to the true goals of the instrumentation tool, thus distracting tool-developers from their primary objectives.
 
@@
 Achieving this level of specialization using and existing system requires complicated runtime state management and checks to switch between different versions of the code.
-With Granary, a tool like \toolname{RCUdbg} can specialize code using instrumentation policies.
+With Granary, a tool like \toolname{RCUdbg} can specialize code using \emph{instrumentation policies}.
 An instrumentation policy is a function that names an execution context and decides how to instrument instructions that will execute within that context.
@@
 \subsection{Control Flow Implementation}
+
+Granary's mechanism of looking up branch targets and resolving control flow depends only on CTI policy and target information.
+This information is either directly embedded in emitted executable code, or in immutable data structures directly referenced by emitted code.
+This approach ensures that policy information is never lost or corrupted, even in the face of concurrent executions of the same emitted code, arbitrary pre-emption, and arbitrary resumption.
+Because Granary's translation mechanism is driven by these two inputs, Granary is otherwise stateless\footnote{This is not quite true, as Granary maintains persistent state in the form of the code cache index.
+However, the code cache index behaves similarly to memoization in a purely functional programming language: it is an optimization, not a necessity.} across context switches.
+
 \subsubsection{Direct Control Flow}
@@
 
-\section{Specializing on What to Instrument}\label{sec:what}
+\section{Scope-Driven Specialization}\label{sec:what}
 
@@
 \scriptsize
-\begin{tabularx}{\linewidth}{| l | >{\raggedright}m{0.3\linewidth} | X |}
+\begin{tabularx}{\linewidth}{| c | >{\raggedright}m{0.3\linewidth} | X |}
 \hline
-\bf Name & \bf Description & \bf Example Tool \\
+\bf Scope & \bf Description & \bf Example Tool \\
 \hline
-Full-kernel & All kernel code is instrumented.
+Kernel & All kernel code is instrumented.
 & Granary's \toolname{cfg} tool constructs inter- and intra-procedural control flow graphs and records per-basic block and per-edge performance counters.
@@
 \hline
-Module-only & All code from a specific kernel module or driver is instrumented.
+Module & All code from a specific kernel module or driver is instrumented.
 Granary detaches when module code invokes kernel code, and re-attaches when kernel code invokes module code.
@@
 \hline
-Probe-based & All code within a particular function is instrumented. Granary attaches when the function is invoked, and detaches when the function returns or invokes another function. & TODO \\
+Function & All code within a particular function is instrumented. Granary attaches when the function is invoked, and detaches when the function returns or invokes another function. & TODO \\
+\hline
+Basic Block & Individual basic blocks are instrumented on demand.
+& Granary's address watchpoints framework can be tuned to instrument only code accessing specific data.
+A general protection fault occurs when native code dereferences a  ``watched address".
+Granary can transparently handle this fault by returning execution to an instrumented version of the faulting code (where the instrumentation prevents such faults).
+Native execution can resume at the end of the basic block.
+\\
 \hline
@@
 
-Granary allows tool developers to precisely specify what code should be instrumented.
+Granary allows tool developers to precisely specify the scope or granularity of instrumentation.
+For example, Granary can instrument all kernel code, specific module code, or only specific function code.
+
+%what code should be instrumented.
 Example granularities of instrumentation include the whole kernel, only specific modules, only specific functions, or on-demand instrumentation based on events (e.g.
@@
 
-\section{Specializing on How to Instrument}\label{sec:how}
+\section{Context-Driven Instrumentation}\label{sec:how}
 
@@
 One example of this is Granary's \toolname{RCUdbg} tool, which instruments read-side and write-side critical sections more heavily than the rest of code.
-Unfortunately, existing DBT systems fall short in their support for runtime code specialization \cite{DRK,btkernel,Pin,DynamoRIO}.
+Unfortunately, existing DBT systems do not support runtime code specialization \cite{DRK,btkernel,Pin,DynamoRIO}.
 
@@
 Granary decodes and instruments the targeted instructions according to the input policy information.
-Because Granary's translation mechanism depends only on CTI policy and target information, Granary is able to ensure that policy information is never lost or corrupted, even in the face of concurrent executions of the same module code, arbitrary pre-emption, and arbitrary resumption.
 

### 2013-12-07 VEE-14/paper.tex
Not much progress today. Decided to either start shaving off the wrapper stuff, or maybe I will talk about wrappers and probes uniformly (given their similarily). Also brought in Guardrail as a motivating example for adjusting the granularity of instrumentation based on an analysis tool's scope.

@@
 
-\section{Scope-Driven Specialization}\label{sec:what}
-
 \begin{table*}[ht!]
 \scriptsize
-\begin{tabularx}{\linewidth}{| c | >{\raggedright}m{0.3\linewidth} | X |}
+\begin{tabularx}{\linewidth}{| l | >{\raggedright}p{0.3\linewidth} | X |}
 \hline
-\bf Scope & \bf Description & \bf Example Tool \\
+\bf Granularity & \bf Description & \bf Example Tool \\
 \hline
@@
 All non-module kernel code executes with zero overhead.
-& TODO \\
+& Granary's Guardrail tool records traces all instructions and memory operations performed by device driver code \cite{Guardrail}.
+A sideline thread analyzes the recorded traces to detect data race bugs and memory access bugs.
+\\
 \hline
@@
 \end{tabularx}
-\caption{\label{tab:what_kinds}Different kinds to instrumentation strategies supported by Granary.}
+\caption{\label{tab:granularities}Different granularities of instrumentation supported by Granary.}
 \end{table*}
+
+\section{Scope-Driven Specialization}\label{sec:what}
+
+Some instrumentation tools have a limited scope.
+That is, they are made to solve a specific problem within a specific domain.
+For example, Granary's Guardrail tool is concerned with finding bugs in Linux device drivers \cite{Guardrail}.
+Guardrail records and analyzes traces of device driver code to find bugs.
+However, Guardrail's analysis can be heavyweight, and so limiting the analysis to only device driver code results in better performance.
+Granary allows Guardrail to set the granularity of instrumentation to only instrument a specific module, while leaving the rest of the kernel to execute without overhead.
+
+%The recording and analysis phases of Guardrail are heavyweight, therefore, limiting
+
+% Instrumenting the whole kernel to observer driver (module) execution is unnecessary and potentially very inefficient (especially if the tool is meant for online analysis). That is,
+
+Guardrail is an example of how tools can choose an instrumentation granularity that better matches their scope and needs.
+That is, a Granary tool can choose to instrument some code, while leaving other code to execute natively.
+This is useful when the instrumentation in question has a high overhead or when only a subset of the kernel's code is of interest.
+\Cref{tab:granularities} provides a high-level overview of the different granularities of instrumentation supported by Granary.
+Granary tools, however, are not restricted to only the Granularities listed in the table.
+
+%instrument all kernel code, specific module code, or only specific function code.
+%what code should be instrumented.
+Example granularities of instrumentation include the whole kernel, only specific modules, only specific functions, or on-demand instrumentation based on events (e.g.
+interrupts or exceptions).
+
+\subsection{Instrumenting All Kernel Code}
+Like DRK and btkernel, Granary can instrument the whole kernel.
+Instrumenting the whole kernel requires interposing on the system call entrypoint (x86-64 \texttt{MSR\_LSTAR} model-specific register) and on the interrupt vector entrypoints (stored in the interrupt descriptor table).
+For whole-kernel instrumentation, Granary boostraps by duplicating the the early entrypoint instructions that are responsible for switching execution onto a kernel stack, and then injects a jump to translate/instrumented kernel code after the stack switch.
+The translation process continues when a system call or interrupt occurs and follows the jump from duplicated into translated code.
+Granary only translate code that is known to operate on a kernel stack\footnote{Not all kernel code executes on a kernel stack.
+For example, when a user space application invokes the \texttt{syscall} instruction, control is transferred to the kernel, but no stack switch occurs.
+The kernel's early system call entrypoint code is responsible for switching execution onto a kernel stack.} so that instrumentation tools can safely save transient data in a reentrant way by spilling that data to the stack.
 
@@
 \end{tabularx}
-\caption{Example of using different granularities of instrumentation to selectively instrument or not instrument specific code.}
+\caption{Visual example of using different granularities of instrumentation to selectively instrument or not instrument specific code.}
 \end{figure}
+
+\subsection{Instrumenting Only Module Code}\label{sec:module}
 
@@
 
-Granary allows tool developers to precisely specify the scope or granularity of instrumentation.
-For example, Granary can instrument all kernel code, specific module code, or only specific function code.
-
-%what code should be instrumented.
-Example granularities of instrumentation include the whole kernel, only specific modules, only specific functions, or on-demand instrumentation based on events (e.g.
-interrupts or exceptions).
-
-\subsection{Kernel}
-Like DRK and btkernel, Granary can instrument the whole kernel.
-Instrumenting the whole kernel requires interposing on the system call entrypoint (x86-64 \texttt{MSR\_LSTAR} model-specific register) and on the interrupt vector entrypoints (stored in the interrupt descriptor table).
-For whole-kernel instrumentation, Granary boostraps by duplicating the the early entrypoint instructions that are responsible for switching execution onto a kernel stack, and then injects a jump to translate/instrumented kernel code after the stack switch.
-The translation process continues when a system call or interrupt occurs and follows the jump from duplicated into translated code.
-Granary only translate code that is known to operate on a kernel stack\footnote{Not all kernel code executes on a kernel stack.
-For example, when a user space application invokes the \texttt{syscall} instruction, control is transferred to the kernel, but no stack switch occurs.
-The kernel's early system call entrypoint code is responsible for switching execution onto a kernel stack.} so that instrumentation tools can safely save transient data in a reentrant way by spilling that data to the stack.
-
-\subsection{Modules}\label{sec:module}
-
 Unlike DRK and btkernel, Granary can target instrumentation only at specific kernel modules, while executing the rest of the kernel natively and without overhead.
@@
 
-\subsection{Functions}\label{sec:function_wrapper}
-
-Granary can instrument or not instrument functions in several different ways.
+\subsection{Instrumenting Specific Functions}\label{sec:function_wrapper}
+
+%Granary can instrument or not instrument functions in several different ways.
 For example, \emph{probe-based} instrumentation allows Granary to selectively instrument only a particular function, while \emph{wrapper-based} instrumentation allows Granary to isolate a particular function that is invoked by instrumented code, and choose to instrument that function or run it natively (all while having the flexibility to arbitrarily inspect that function's arguments and return values in a type-safe way).
 
-\subsubsection{Probe-based instrumentation}
+%\subsubsection{Probe-based instrumentation}
 
@@
 
-\subsubsection{Wrapper-based instrumentation}
-
-Wrapper-based instrumentation makes use of the same wrapper capabilities used by module-only instrumentation (\Cref{sec:module}), but is available regardless of the instrumentation granularity.
+%\subsubsection{Wrapper-based instrumentation}
+%
+%Wrapper-based instrumentation makes use of the same wrapper capabilities used by module-only instrumentation (\Cref{sec:module}), but is available regardless of the instrumentation granularity.
 Wrapper-based instrumentation is similar to probe-based instrumentation, but permits more selectivity in terms of the context in which the function will be wrapped.
@@
 Finer-grained selectivity can be achieved by having individual instrumentation policies manually decide how to wrap a given function.
-
-\begin{figure}[t!]
-\lstset{language=C, tabsize=2, stepnumber=1}
-\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-MODULE_FUNCTION_WRAPPER(__kmalloc, (size, gfp), {
-  void *addr = __kmalloc(size, gfp);
-  if(is_valid_address(addr)) {
-    add_watchpoint(addr, addr, size);
-  }
-  return addr;
-})
-\end{lstlisting}
-\caption{Example module function wrapper for the \texttt{\_\_kmalloc} function.
+%
+%\begin{figure}[t!]
+%\lstset{language=C, tabsize=2, stepnumber=1}
+%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
+%MODULE_FUNCTION_WRAPPER(__kmalloc, (size, gfp), {
+%  void *addr = __kmalloc(size, gfp);
+%  if(is_valid_address(addr)) {
+%    add_watchpoint(addr, addr, size);
+%  }
+%  return addr;
+%})
+%\end{lstlisting}
+%\caption{Example module function wrapper for the \texttt{\_\_kmalloc} function.
 This wrapper adds an ``address watchpoint" to the address returned by the kernel's \texttt{\_\_kmalloc} memory allocator.
@@
 Memory that has not been live for several consecutive epochs is treated as a potential leak, and a is checked for reachability using a heap memory scanner.}
-\label{fig:malloc_wrapper}
-\end{figure}
-
-Granary's module memory leak detector is one example application that uses the selectivity of wrappers (with respect to their calling context) to distinguish between module-allocated and kernel-allocated memory.
+%\label{fig:malloc_wrapper}
+%\end{figure}
+%
+%Granary's module memory leak detector is one example application that uses the selectivity of wrappers (with respect to their calling context) to distinguish between module-allocated and kernel-allocated memory.
 \Cref{fig:malloc_wrapper} shows a module function wrapper that adds ``address watchpoints"  \cite{BehaveOrBeWatched} to module-allocated memory.
@@
 
-\subsection{On-demand}
+\subsection{On-Demand Instrumentation}
 

### 2013-12-08 VEE-14/paper.tex
Hacking on the module wrappers section. Really chopping it down to the essentials.

@@
 
+% Make text in figure captions small
+\usepackage{caption3} % load caption package kernel first
+\DeclareCaptionOption{parskip}[]{} % disable "parskip" caption option
+\usepackage[small]{caption}
+
 % For drawing FSMs
@@
 \usepackage{epsfig}
+
+% Compressed enumerations
+\newenvironment{enumerate*}%
+  {\begin{enumerate}%
+    \setlength{\itemsep}{2pt}%
+    \setlength{\parskip}{0pt}}%
+  {\end{enumerate}}
 
@@
 
-\section{Scope-Driven Specialization}\label{sec:what}
-
-Some instrumentation tools have a limited scope.
-That is, they are made to solve a specific problem within a specific domain.
+\section{Only Instrument What is Needed}\label{sec:what}
+
+Some instrumentation tools have a limited scope, and therefore do not need to instrument all code.
 For example, Granary's Guardrail tool is concerned with finding bugs in Linux device drivers \cite{Guardrail}.
-Guardrail records and analyzes traces of device driver code to find bugs.
-However, Guardrail's analysis can be heavyweight, and so limiting the analysis to only device driver code results in better performance.
-Granary allows Guardrail to set the granularity of instrumentation to only instrument a specific module, while leaving the rest of the kernel to execute without overhead.
+Guardrail records traces of executed instructions, and analyzes those traces to find memory access and concurrency bugs.
+Both phases are heavyweight, and so limiting the scope of what is instrumented to what is of interest (module code) results in better performance.
+Unfortunately, existing systems are unable to change the granularity of instrumentation.
+Fortunately, Granary allows Guardrail to set the granularity of instrumentation to only instrument a specific module, while leaving the rest of the kernel to execute without overhead.
+
+%The closest that an existing system can get to
+
+% Therefore, if Guardrail were implemented using DRK or btkernel, then it would have to implement its own form of runtime code specialization
+
+%. Recording and analyzing every instruction executed by the kernel is
+
+%Guardrail records and analyzes traces of device driver code to find bugs.
+However, Guardrail's analysis can be heavyweight.
+If Guardrail were forced to instrument all code (e.g.
+with DRK or btkernel), then Guardrail would impose significant performance overheads.
+However, if Guardrail's recording were limited to only module code (
+
+%and so limiting the analysis to only device driver code results in better performance.
 
@@
 
-Guardrail is an example of how tools can choose an instrumentation granularity that better matches their scope and needs.
+Guardrail is an example of how tools can choose the instrumentation granularity that best matches their needs.
 That is, a Granary tool can choose to instrument some code, while leaving other code to execute natively.
 This is useful when the instrumentation in question has a high overhead or when only a subset of the kernel's code is of interest.
-\Cref{tab:granularities} provides a high-level overview of the different granularities of instrumentation supported by Granary.
-Granary tools, however, are not restricted to only the Granularities listed in the table.
+\Cref{tab:granularities} provides a non-exhaustive list of the different granularities of instrumentation supported by Granary.
 
@@
 \subsection{Instrumenting All Kernel Code}
+
 Like DRK and btkernel, Granary can instrument the whole kernel.
-Instrumenting the whole kernel requires interposing on the system call entrypoint (x86-64 \texttt{MSR\_LSTAR} model-specific register) and on the interrupt vector entrypoints (stored in the interrupt descriptor table).
-For whole-kernel instrumentation, Granary boostraps by duplicating the the early entrypoint instructions that are responsible for switching execution onto a kernel stack, and then injects a jump to translate/instrumented kernel code after the stack switch.
+Instrumenting the whole kernel requires interposing on the system call entrypoint\footnote{The system call entrypoint is stored in the x86-64 \texttt{MSR\_LSTAR} model-specific register.} and on the interrupt vector entrypoints\footnote{Interrupt vector entrypoints are stored in the interrupt descriptor table, which can be read from and written to using the \texttt{sidt} and \texttt{lidt} instructions, respectively.}.
+For whole-kernel instrumentation, Granary bootstraps by duplicating the the early entrypoint instructions that are responsible for switching execution onto a kernel stack, and then injects a jump to translated/instrumented kernel code after the stack switch.
 The translation process continues when a system call or interrupt occurs and follows the jump from duplicated into translated code.
-Granary only translate code that is known to operate on a kernel stack\footnote{Not all kernel code executes on a kernel stack.
+Granary only translates code that is known to operate on a kernel stack\footnote{Not all kernel code executes on a kernel stack.
 For example, when a user space application invokes the \texttt{syscall} instruction, control is transferred to the kernel, but no stack switch occurs.
@@
 \subfloat[Module-only instrumentation]{\hspace{2.5em}\epsfig{file=diagrams/module.eps,width=0.125\textwidth}} \\
-\subfloat[Module-only instrumentation, with a function specialized to run natively by using a module function wrapper.]{\hspace{2.5em}\epsfig{file=diagrams/module_wrapper.eps,width=0.125\textwidth}} & %
-\subfloat[\label{fig:module_probe}Module-only instrumentation, with a kernel function specialized to also be instrumented by using a probe-based wrapper.]{\hspace{2.5em}\epsfig{file=diagrams/module_probe.eps,width=0.125\textwidth}}
+\subfloat[Module-only instrumentation, with a module function wrapped to run natively.]{\hspace{2.5em}\epsfig{file=diagrams/module_wrapper.eps,width=0.125\textwidth}} & %
+\subfloat[\label{fig:module_probe}Module-only instrumentation, with a kernel function wrapped to also be instrumented.]{\hspace{2.5em}\epsfig{file=diagrams/module_probe.eps,width=0.125\textwidth}}
 \end{tabularx}
@@
 
-\begin{figure*}[ht!]
-\lstset{language=C, tabsize=2, stepnumber=1}
-\begin{multicols}{2}
-\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-struct device_driver {
-  ...
-  int  (*probe)(struct device *);
-  int  (*remove)(struct device *);
-  void (*shutdown)(struct device *);
-  int  (*suspend)(struct device *, pm_message_t);
-  int  (*resume)(struct device *);
-  ...
-  const struct dev_pm_ops *pm;
-  ...
-};
-\end{lstlisting}
-\columnbreak
-\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-TYPE_WRAPPER(struct device_driver, {
-  PRE_OUT {
-    ABORT_IF_FUNCTION_IS_WRAPPED(arg.probe)
-    WRAP_FUNCTION(arg.probe);
-    WRAP_FUNCTION(arg.remove);
-    ...
-  }
-  POST_OUT {
-    POST_WRAP(arg.pm);
-  }
-})
-\end{lstlisting}
-\end{multicols}
-\caption{The Linux device driver structure is shown on the left.
+Granary is the first kernel-space instrumentation system able to target instrumentation only at specific kernel modules, while executing the rest of the kernel natively and without overhead.
+Instrumenting only specific kernel modules is particularly useful for debugging tools because modules represent the bulk of new kernel code under development, and contain the most bugs \cite{FaultsInLinux}.
+
+The basic technique that Granary uses to interpose on module code is hardware page protection.
+First, Granary page protects existing module code, marking it as non-executable.
+Then, Granary traps attempts to execute the protected module code and transparently recovers from these traps by returning execution to instrumented code.
+Granary also interposes on the Linux kernel's module loading process.
+When a kernel module is dynamically loaded, Granary bootstraps by translating the first basic block of the module's initialization function.
+It then replaces the pointer to that function with a pointer to the translated basic block.
+The translation process continues when the kernel initializes the module by invoking the translated module code.
+As with existing modules, the native code segments of dynamically loaded modules are page protected to prevent and recover from accidental execution.
+
+%Granary interposes on an existing module by using hardware page protection to mark the module's code as non-executable and trapping attempts to execute the protected code.
+
+While correct, hardware page protection is an inefficient mechanism for gaining control of module code execution.
+That is, after the (instrumented) module initializer runs to completion, Granary needs an efficient way to regain control of future invocations of the module's code.
+Our approach to regaining control is based on the observation that modules tell the kernel about their interfaces by registering interface functions with the kernel.
+These functions are the only supported mechanism for extending the kernel's functionality in a standardized way.
+When the kernel executes registered functions, control transfers from the kernel into the module.
+To Granary, registered module functions represent potential future \emph{attach} points, where a switch from native to instrumented execution will occur.
+Furthermore, Granary discovers additional attach points at detach points by observing pointers to module functions that are passed by modules to the kernel.
+
+Granary uses static analysis to dynamically discover potential attach points. Potential attach points include: \begin{enumerate*}
+	\item Function pointers passed as arguments to kernel functions.
+	\item Function pointers stored in structures passed to kernel functions.
+This rule applies transitively.
+For example, if one structure has a field that points to another structure, and the latter structure contains a function pointer, then the latter structure's function pointers are also potential attach points.
+This is a common pattern in the Linux kernel, and mimics virtual tables in object-oriented languages.
+	\item Function pointers returned from attach points.
+\end{enumerate*}
+
+%Any function pointer passed as an argument to a kernel function, or stored in a structure that can be passed as an argument to a kernel function is treated as a potential attach point.
+
+The static analysis phase of Granary generates ``wrappers'' for the kernel/module interface.
+Interface wrappers come in two forms: kernel function wrappers and data-structure wrappers (called type wrappers).
+Kernel function wrappers are transparently invoked when module code attempts to invoke kernel code.
+Function wrappers delegate to type wrappers to find attach points passed to the kernel.
+
+%\begin{figure*}[t!]
+%\lstset{language=C, tabsize=2, stepnumber=1}
+%\begin{multicols}{2}
+%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
+%struct device_driver {
+%  ...
+%  int  (*probe)(struct device *);
+%  int  (*remove)(struct device *);
+%  void (*shutdown)(struct device *);
+%  int  (*suspend)(struct device *, pm_message_t);
+%  int  (*resume)(struct device *);
+%  ...
+%  const struct dev_pm_ops *pm;
+%  ...
+%};
+%\end{lstlisting}
+%\columnbreak
+%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
+%TYPE_WRAPPER(struct device_driver, {
+%  PRE_OUT {
+%    ABORT_IF_FUNCTION_IS_WRAPPED(arg.probe)
+%    WRAP_FUNCTION(arg.probe);
+%    WRAP_FUNCTION(arg.remove);
+%    ...
+%  }
+%  POST_OUT {
+%    POST_WRAP(arg.pm);
+%  }
+%})
+%\end{lstlisting}
+%\end{multicols}
+%\caption{The Linux device driver structure is shown on the left.
 The automatically generated type wrapper for this structure is shown on the right.
@@
 Finally, the \texttt{RETURN\_} prefix is used to apply some code to return values of either kernel or module wrappers.}
-\label{fig:type_wrapper}
-\end{figure*}
-
-Unlike DRK and btkernel, Granary can target instrumentation only at specific kernel modules, while executing the rest of the kernel natively and without overhead.
-Instrumenting only specific kernel modules is particularly useful for debugging tools because modules represent the bulk of new kernel code under development, and contain the most bugs \cite{FaultsInLinux}.
-
-Granary interposes on an existing module by using hardware page protection to mark an existing module as non-executable and trap attempts to execute that module's code.
-Granary transparently recovers from these traps by returning execution to instrumented code.
-Granary takes control of dynamically loaded modules by interposing on the Linux kernel's module loading process.
-When a kernel module is loaded, Granary bootstraps by translating the first basic block of the module's initialization function.
-It then replaces the pointer to that function with a pointer to the translated basic block.
-The translation process continues when the kernel initializes the module by invoking the translated module code.
-As with existing modules, the native code segments of dynamically loaded modules are page protected to prevent and recover from accidental execution.
-
-While correct, hardware page protection is an inefficient mechanism for gaining control of module code execution.
-That is, after the (instrumented) module initializer runs to completion, Granary needs an efficient way to regain control of future invocations of the module's code.
-Our approach to regaining control is based on the observation that modules tell the kernel about their interfaces by registering functions with the kernel.
-We expect that at least some of the registered functions will be executed by the kernel because this is the mechanism by which modules extend the kernel's functionality.
-To Granary, registered module functions represent potential future \emph{attach} points, where a switch from native to instrumented execution will occur.
-Furthermore, Granary discovers additional attach points at detach points by observing pointers to module functions that are passed by modules to the kernel.
-
-Granary uses static analysis to dynamically discover attach points by wrapping the kernel/module interface.
-Kernel functions are wrapped by \emph{kernel wrappers} that inspect and traverse argument pointers in search of pointers to module functions.
-\Cref{fig:type_wrapper} shows an example of a type wrapper for the Linux device driver structure.
-If a pointer to a module function (a future attach point) is discovered, then Granary replaces that pointer with a function-specific \emph{module wrapper}.
+%\label{fig:type_wrapper}
+%\end{figure*}
+
+\begin{figure}[t]
+\lstset{language=C, tabsize=2, stepnumber=1}
+\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
+// struct inode *
+// iget_locked(struct super_block *, unsigned long);
+FUNCTION_WRAPPER(MODULE, iget_locked, (sb, ino), {
+  PRE_OUT_WRAP(sb);
+  struct inode *inode = iget_locked(sb, ino);
+  POST_OUT_WRAP(sb);
+  RETURN_IN_WRAP(inode);
+  return inode;
+})
+\end{lstlisting}
+\caption{Atomatically generated wrapper for the \texttt{iget\_locked} kernel function.
+\texttt{iget\_locked} obtains an inode from a mounted file system.
+The super block argument (\texttt{sb}) is first pre-wrapped before it is passed to the original \texttt{iget\_locked} function.
+After \texttt{iget\_locked} returns, \texttt{sb} is then post-wrapped.
+Finally, the inode data structure returned by the kernel (\texttt{inode}) is return-wrapped and then returned to instrumented module code.
+The \texttt{OUT} and \texttt{IN} qualifiers of the wrappers denote the direction that data is travelling.
+\texttt{OUT}-wrappers apply to data passed from instrumented code out to the native code.
+\texttt{IN}-wrappers apply to data passed from native code into instrumented code.}
+\end{figure}
+
+%\Cref{fig:type_wrapper} shows an example of a type wrapper for the Linux device driver structure.
+
+If a pointer to a module function (a future attach point) is discovered, then Granary replaces that pointer with a function-specific \emph{module function wrapper}.
+Module function wrappers behave similarly to kernel function wrappers, in that they provide instrumentation tools a mechanism to observe a
+
+%are responsible for wrapping their arguments. When an argument
+
+%At runtime, these wrappers interpose on kernel functions and traverse in-memory data structures, find attach points, and convert them into
+
+%At runtime, Granary detects and interposes on potential attach points by wrapping the kernel/module interface.
+
+%Kernel functions are wrapped by \emph{kernel wrappers} that inspect and traverse argument pointers in search of pointers to module functions.
 Granary inspects and modifies arguments to module wrappers in the same way as for kernel wrappers.
@@
 
-Kernel and module wrappers invoke \emph{type wrappers} to find and wrap pointers to module functions that are directly or indirectly referenced by kernel/module function arguments.
+%Kernel and module wrappers invoke \emph{type wrappers} to find and wrap pointers to module functions that are directly or indirectly referenced by kernel/module function arguments.
 A type wrapper is a function that recursively  traverses the in-memory object graph and converts pointers to module code into pointers to wrapped module functions.
@@
 
-\label{sec:implicit_attach}
-After Granary bootstraps on the module initialization process, \emph{attaching} occurs in one of three ways: \begin{enumerate}
-	\item {\bf Implicit attaching:} the kernel returns to instrumented module code in the code cache.
+%\label{sec:implicit_attach}
+%After Granary bootstraps on the module initialization process, \emph{attaching} occurs in one of three ways: \begin{enumerate}
+%	\item {\bf Implicit attaching:} the kernel returns to instrumented module code in the code cache.
 This occurs when instrumented moduel code is interrupted, or when the module invokes a kernel function (detaching) and that function returns to the instrumented module code (attaching).
-	\item {\bf Fast attaching:} the kernel invokes a wrapped module function.
+%	\item {\bf Fast attaching:} the kernel invokes a wrapped module function.
 The first instance of fast attaching for any kernel module is when the kernel's module loader invokes the instrumented module initializer function.
-	\item {\bf Slow attaching:} the kernel invokes unwrapped module code.
+%	\item {\bf Slow attaching:} the kernel invokes unwrapped module code.
 If the kernel executes a module function which was passed to the kernel in a type-unsafe manner, then the processor will raise a fault because Granary uses hardware page protection to prevent module code from being executed.
 Granary handles these faults by returning execution to the instrumented version of the faulting module code.
-\end{enumerate}
-
-\label{sec:detach}
-Granary \emph{detaches} when control transfers from instrumented code to native (uninstrumented) code. Detaching occurs in one of two ways: \begin{enumerate}
-	\item {\bf Implicit detaching:} instrumented module code returns to the kernel code which originally invoked the module, or is interrupted (initial interrupt handling is done by the kernel).
-	\item {\bf Wrapped detaching:} instrumented module code invokes a kernel wrapper, which later transfers control to the kernel.
-\end{enumerate}
+%\end{enumerate}
+%
+%\label{sec:detach}
+%Granary \emph{detaches} when control transfers from instrumented code to native (uninstrumented) code. Detaching occurs in one of two ways: \begin{enumerate}
+%	\item {\bf Implicit detaching:} instrumented module code returns to the kernel code which originally invoked the module, or is interrupted (initial interrupt handling is done by the kernel).
+%	\item {\bf Wrapped detaching:} instrumented module code invokes a kernel wrapper, which later transfers control to the kernel.
+%\end{enumerate}
 

### 2013-12-10 VEE-14/paper.tex
Edited up to but not including section 4.

@@
 \hline
-Function & All code within a particular function is instrumented. Granary attaches when the function is invoked, and detaches when the function returns or invokes another function. & TODO \\
+Function & All code within a particular function is instrumented.
+Granary attaches when the function is invoked, and detaches when the function returns or invokes another function.
+& \toolname{RCUdbg} interposes on the \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} Linux kernel functions.
+When invoked by native code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} attaches instrumentation.
+When invoked by instrumented code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} updates the current nesting depth of the read-side critical section.
+\\
+
+% Granary's \toolname{Irrigator} tool find leaks on module-owned memory (\Cref{sec:function_wrapper}).
+\toolname{Irrigator} instruments only module code, but also interposes on specific kernel functions (kernel memory allocators and deallocators) to gain visibility into kernel code that would normally execute natively (using module-only instrumentation).\\
 \hline
-Basic Block & Individual basic blocks are instrumented on demand.
+%Basic Block & Individual basic blocks are instrumented on demand.
 & Granary's address watchpoints framework can be tuned to instrument only code accessing specific data.
@@
 \\
-\hline
+%\hline
 \end{tabularx}
@@
 Our approach to regaining control is based on the observation that modules tell the kernel about their interfaces by registering interface functions with the kernel.
-These functions are the only supported mechanism for extending the kernel's functionality in a standardized way.
-When the kernel executes registered functions, control transfers from the kernel into the module.
-To Granary, registered module functions represent potential future \emph{attach} points, where a switch from native to instrumented execution will occur.
-Furthermore, Granary discovers additional attach points at detach points by observing pointers to module functions that are passed by modules to the kernel.
-
-Granary uses static analysis to dynamically discover potential attach points. Potential attach points include: \begin{enumerate*}
+These functions are the standard mechanism for extending the kernel's functionality.
+When the kernel executes a registered module function, control transfers from the kernel into the module.
+The act of registering a function is a signal to Granary that instrumentation must ``attach'' to the registered function when the kernel executes that function.
+Granary's basic attaching mechanism, page protection, raises a trap when registered functions are executed.
+In practice, we would like to avoid the overhead of the trap by proactively finding attach points and replacing them with their instrumented equivalents.
+
+%Therefore, to Granary, any function pointer passed from the module to the kernel is a potential future attach point.
+%registered module functions represent potential future \emph{attach} points, where a switch from native to instrumented execution will occur. 
+%Furthermore, Granary discovers additional attach points at detach points by observing pointers to module functions that are passed by modules to the kernel.
+
+Granary uses static analysis of the kernel source code to dynamically discover potential attach points. Potential attach are recognized as being any of: \begin{enumerate*}
 	\item Function pointers passed as arguments to kernel functions.
-	\item Function pointers stored in structures passed to kernel functions.
+	\item Function pointers stored in structures passed (by pointer or value) to kernel functions.
 This rule applies transitively.
@@
 This is a common pattern in the Linux kernel, and mimics virtual tables in object-oriented languages.
-	\item Function pointers returned from attach points.
+	\item Function pointers and structures returned from attach points.
 \end{enumerate*}
@@
 
-The static analysis phase of Granary generates ``wrappers'' for the kernel/module interface.
-Interface wrappers come in two forms: kernel function wrappers and data-structure wrappers (called type wrappers).
-Kernel function wrappers are transparently invoked when module code attempts to invoke kernel code.
-Function wrappers delegate to type wrappers to find attach points passed to the kernel.
+Granary's static analysis phase generates ``wrappers'' for the kernel/module interface. Two kinds of wrappers are generated for the kernel/module interface: \begin{inparaenum}[i)]
+	\item kernel function wrappers (\Cref{fig:function_wrapper}), and
+	\item data-structure wrappers (called type wrappers).
+\end{inparaenum} Kernel function wrappers are transparently invoked when module code attempts to invoke kernel code.
+Function wrappers delegate to type wrappers to find and wrap attach points passed to the kernel.
 
@@
 \end{lstlisting}
-\caption{Atomatically generated wrapper for the \texttt{iget\_locked} kernel function.
+\caption{\label{fig:function_wrapper}Atomatically generated wrapper for the \texttt{iget\_locked} kernel function.
 \texttt{iget\_locked} obtains an inode from a mounted file system.
@@
 If a pointer to a module function (a future attach point) is discovered, then Granary replaces that pointer with a function-specific \emph{module function wrapper}.
-Module function wrappers behave similarly to kernel function wrappers, in that they provide instrumentation tools a mechanism to observe a
+Module function wrappers behave similarly to kernel function wrappers, in that they provide Granary more visibility into potential attach points, and provide instrumentation tools visibility into the runtime object graph.
+
+% instrumentation tools a mechanism to observe a
 
@@
 %\end{enumerate}
+
+\label{sec:wrappers}
 
@@
 
+Instrumenting individual functions is useful in when the code of interest is only a function or group of functions, or when the tool does not instrument all code, but wants visibility into functions that would otherwise execute natively.
+For example, \toolname{RCUdbg} interposes on the \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} kernel functions.
+When \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} is invoked by native kernel code, Granary attaches and begins instrumenting code using the read-side critical section instrumentation policy.
+When \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} is invoked by instrumented code, Granary updates the nesting depth of the current read-side critical section.
+\texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} behaves similarly.
+
+%For example, Granary's \toolname{Irrigator} tool detects memory leaks in kernel modules.
+\toolname{Irrigator}  finds leaks of ``module-owned" memory.
+We say that the some memory is owned by a module iff that memory is deallocated by a module.
+Unfortunately, ownership of module-allocated memory (for an arbitrary, binary module) is not known \emph{a priori}.
+Therefore, \toolname{Irrigator} conservatively assumes that all module-allocated memory is module-owned, unless it is deallocated by the kernel.
+\toolname{}
+
+%However, the instrumentation granularity of \toolname{Irrigator} is module code, and so (almost all) kernel code executes natively.
+Fortunately, Granary allows \toolname{Irrigator} to instrument specific kernel functions---in this case memory deallocators---and detect when the kernel frees module-allocated memory.
+\toolname{Irrigator} also interposes on specific kernel memory allocators (e.g.
+for the \texttt{sk\_buff} data structures), which are allocated by the kernel but deallocated by modules (e.g.
+network device drivers deallocate \texttt{sk\_buff} data structures).
+
+%We also use probe-based instrumentation to assign module ownership to kernel-allocated \texttt{sk\_buff} data structures, which are allocated by the kernel but deallocated by network device driver modules.
+
+%: \begin{inparaenum}[i)]
+%	\item the instrumentation tool is only interested in that particular function; or
+%	\item when the instrumentation tool needs visibility into that function
+%\end{inparaenum}
+
 %Granary can instrument or not instrument functions in several different ways.
@@
 
-Probe-based instrumentation allows an instrumentation tool to replace \emph{all} executions of a particular kernel function, regardless of whether or not the replaced function is executed by native or instrumented code \cite{KProbes, AnyWhereAnyTimeDBT, KernInst}.
+%Like KernInst \cite{KernInst,AnyWhereAnyTimeDBT}, Granary can take over the execution of individual kernel functions.
+
+Granary uses code splicing to allow tools to interpose on the execution of individual functions invoked by native code (which nornmally execute outside of Granary's control) \cite{KernInst,AnyWhereAnyTimeDBT,KProbes}.
+Granary allows tools to interpose on specific functions invoked by instrumented code using kernel or module function wrappers (\Cref{sec:wrappers}).
+\Cref{fig:module_probe} is a visual example of how interposing on individual functions can be combined with module-only instrumentation to provide an instrumentation tool visibility into otherwise native kernel events.
+
+% Instrumenting invidual functions, called \emph{probe-based instrumentation}, giv \cite{KProbes}.
+
+%Probe-based instrumentation allows an instrumentation tool to replace \emph{all} executions of a particular kernel function, regardless of whether or not the replaced function is executed by native or instrumented code \cite{KProbes, AnyWhereAnyTimeDBT, KernInst}.
 Probe-based instrumentation can be used to wrap a particular function without instrumenting its code, or to wrap and instrument the probed function's code.
 
-Probe-based instrumentation provides visibility into specific kernel events (e.g.
+%Probe-based instrumentation provides visibility into specific kernel events (e.g.
 memory allocations) to instrumentation tools that don't instrument the whole kernel (e.g.
 module-only, \Cref{fig:module_probe}).
-One example tool that combines module-only and probe-based instrumentation is Granary's kernel module memory leak detector.
-This leak detector finds leaks of module-owned memory.
-We say that the some memory is owned by a module iff that memory is deallocated by a module.
-Unfortunately, ownership of module-allocated memory (for an arbitrary, binary module) is not known \emph{a priori}.
-Therefore, our tool conservatively assumes that all module-allocated memory is module-owned, unless it is freed by the kernel (observed via probe-based instrumentation of a kernel memory deallocator).
-We also use probe-based instrumentation to assign module ownership to kernel-allocated \texttt{sk\_buff} data structures, which are allocated by the kernel but deallocated by network device driver modules.
+One example tool that combines module-only and probe-based instrumentation is
 
@@
 
-\subsection{On-Demand Instrumentation}
-
-On-demand instrumentation makes use of Granary's control over interrupt and exception handling.
+%\subsection{On-Demand Instrumentation}
+%
+%On-demand instrumentation makes use of Granary's control over interrupt and exception handling.
 When a specific exception or interrupt occurs, a Granary tool can turn on instrumentation for the interrupted code.
@@
 Another application of on-demand instrumentation is data watchpoints based on hardware page protection and memory shadowing.
-
-Turning on instrumentation when specific data is accessed makes on-demand instrumentation inherently \emph{data-centric} as opposed to \emph{code-centric}.
+%
+%Turning on instrumentation when specific data is accessed makes on-demand instrumentation inherently \emph{data-centric} as opposed to \emph{code-centric}.
 That is, instrumentation is targeted at instructions accessing specific data, instead of the typical ``dragnet" approach employed by most DBT systems, which instrument all memory loads/stores so as to capture the select few that access key data.

### 2013-12-10 VEE-14/paper.tex
Minor edits to sections 1 .. 3. Fixed typos, grammar, and flow issues.

@@
 Limiting the scope of what is instrumented is  important because the overhead introduced by using runtime code instrumentation is usually dominated by the instrumentation itself and not the baseline overhead of the runtime system supporting that instrumentation.
-Also, specializing based on what is instrumented allows instrumentation tools focus instrumentation on the code of interest.
-For example, if one wants to instrument RCU read-side and write-side critical sections, as is the case with Granary's \toolname{RCUdbg} tool, then instrumenting all code just to filter out / find those sections is wasteful.
+Also, specializing based on what is instrumented allows instrumentation tools to focus instrumentation on the code of interest.
+For example, if one wants to instrument RCU read-side critical sections, as is the case with Granary's \toolname{RCUdbg} tool, then instrumenting all code just to filter out and find those critical sections is wasteful.
 
@@
 However, runtime checks are inefficient, and generating many versions of the code introduces bloat and pollutes the CPU's instruction cache.
-Also, the runtime checks needed to multiplex between different versions of instrumented code is orthogonal to the true goals of the instrumentation tool, thus distracting tool-developers from their primary objectives.
+Granary avoids these issues by supporting context-specific runtime code specialization that generates multiple versions of basic blocks \emph{just-in-time} (JIT), instead ahead-of-time, and without requiring runtime checks.
+
+%Also, the runtime checks needed to multiplex between different versions of instrumented code are orthogonal to the true goals of the instrumentation tool, thus distracting tool-developers from their primary objectives.
 
@@
 
-One example of a tool that uses runtime code specialization is Granary's Read-Copy-Update (RCU) debugging tool, \toolname{RCUdbg}.
+One example tool that uses Granary's JIT-based runtime code specialization feature is \toolname{RCUdbg}: Granary's Read-Copy-Update (RCU) debugging tool.
 \toolname{RCUdbg} catches bugs related to invalid uses of RCU-protected pointers (\Cref{sec:rcudbg}).
@@
 Granary translates and instruments kernel binaries one basic block at a time.
-In Granary, a basic block is a sequence of x86-64 instructions ending in a conditional branch, \texttt{ret}, or a \texttt{jmp} instruction.
-Unlike DRK and DynamoRIO, Granary does not terminate basic blocks at \texttt{call} instructions.
+In Granary, a basic block is a sequence of x86-64 instructions ending in a conditional branch (\texttt{jCC}), \texttt{ret}, or a \texttt{jmp} instruction.
+Unlike DRK and btkernel, Granary does not terminate basic blocks at \texttt{call} instructions.
 This enables longer basic blocks and native-like proceccor return address prediction, but makes Granary non-transparent (\Cref{sec:transparency}).
-Entrypoint basic blocks (e.g.
-interrupt vectors, system call entrypoints, module initializers) are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as execution ``discovers'' them.
+Granary takes control of execution starting from ``natural" program entrypoints (e.g.
+interrupt vectors, system call entrypoints, module initializers).
+Entrypoint basic blocks  are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as execution ``discovers'' them.
 
@@
 
-Granary's just-in-time translation approach means that code executing from the code cache may yield control to Granary to request the address of the next basic block to execute.
+Granary's JIT-based translation approach means that code executing from the code cache may yield control to Granary to request the address of the next basic block to execute.
+This happens when there is a control flow transfer in a basic block, and the target of that control flow transfer is not already indexed in Granary's code cache.
 When instrumented code yields to Granary, a ``context switch'' occurs that transfers execution to a CPU-private stack where Granary operates.
@@
 to translate code that will later execute).
-In such cases, Granary depends on CPU-private data structures to avoid contention, and the stack is one such data structure.
+In such cases, Granary depends on CPU-private data structures to avoid contention, and the private stack is one such data structure.
 A context-switch back to the code cache occurs when the next basic block has been found or translated.
 At this point, instrumented execution continues.
-Similar to other DBT systems, Granary uses caching and hot code patching to reduce the number of context switches.
+Similar to other DBT systems, Granary uses hot code patching (also called code splicing) to reduce the number of context switches.
 In our experience, kernel code stabilizes very quickly: context switches stop happening after the first few seconds of translation.
@@
 Granary uses this information to allow instrumentation tools to define policy-specific interrupt handling routines.
-For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if a code executing within a read-side critical section is interrupted (\Cref{sec:rcudbg}).
+For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if code executing within a read-side critical section is interrupted (\Cref{sec:rcudbg}).
 \end{itemize} %
@@
 This data is useful because it provides precise code coverage information (block and branch coverage), as well as useful data for profile-guided optimization.
-This meta-data is specified by \toolname{cfg} and automatically allocated for \toolname{cfg} by Granary.
+This meta-data is specified by \toolname{cfg} and automatically allocated by Granary.
 
@@
 Granary includes configurable levels of transparency to address the trade-off between transparency and overhead.
-By default, Granary uses a relaxed transparency model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility.
+By default, Granary uses a relaxed transparency \cite{btkernel} model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility.
 By default, Granary exposes the following two artifacts: code cache addresses as function and interrupt return addresses.
@@
 Granary's CPU-private stacks).
-However, accessing CPU-private without first disabling interrupts results in undefined behavior.
+However, accessing CPU-private data without first disabling interrupts results in undefined behavior.
 Therefore, context-switches into Granary first disable interrupts and then switch execution onto one of Granary's CPU-private stacks.
-Context switches out of Granary and into native code (\Cref{sec:detach}) or the code cache returns execution to the native stack and re-enable interrupts.
+Context switches out of Granary and into native code or the code cache returns execution to the native stack and re-enable interrupts.
 Execution behavior is preserved because Granary saves the machine state (i.e.
@@
 One challenge with any DBT system is how it copes with non-reentrant instrumentation.
-For example, instrumentation that modifies CPU-private data structures cannot be safely executed by a pre-emptive kernel running on an SMP.
-If the thread being instrumented is interrupted, then the kernel might decide to re-schedule the task to resume its execution on a different CPU.
+For example, instrumentation that modifies CPU-private data structures (without disabling interrupts) cannot be safely executed by a pre-emptive kernel running on an SMP.
+If the thread being instrumented is interrupted, then the kernel might decide to re-schedule the thread to resume its execution on a different CPU.
 Any modifications to CPU-private state would lead to undefined behavior.
+An example of this is instrumentation that acquires a CPU-private lock on one CPU, is interrupted, and later releases the wrong lock on the wrong CPU.
 
 DRK's solution to this problem assumes that all injected instrumentation instructions are potentially non-reentrant, and \emph{delays} interrupts occuring within instrumented code \cite{DRK}.
-Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86-64: interrupts and exceptions only arrive on logical instruction boundaries in the native code, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
-This approach introduces subtle complexity when creating instrumentation that injects instructions both before (pre) and after (post) native instructions.
+Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86-64: interrupts and exceptions only arrive on native instruction boundaries, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
+This approach introduces subtle complexity when injecting non-idempotent instrumentation before (pre) and after (post) native instructions.
 
@@
 That is, \texttt{push \%rax} is executed in both the original and tail basic block, but \texttt{pop \%rax} is only executed in the tail basic block.
-This is a specific example of a more general problem: DRK's approach to interrupt delaying does not play well with idempotent instrumentation.
+This is a specific example of a more general problem: DRK's approach to interrupt delaying does not play well with non-idempotent instrumentation.
 
@@
 Btkernel provides no solution to the problem of handling non-reentrant instrumentation, beyond requiring that instrumentation tools disable and re-enable interrupts around non-reentrant instrumentation code \cite{btkernel}.
-Unfortunately, disabling and re-enabling interrupt is costly.
+The is inefficient because it introduces flag stalls into the instruction pipeline \cite{AgnerMicroarchitecture}, and because interrupts arrive relatively infrequently.
 
@@
 Most instrumentation is re-entrant, partially because kernel instrumentation tools can depend on the runtime call stack for scratch storage space.
-However, interrupt delaying is sometimes useful because it enables non-reentrant instrumentation to avoid to continually disable and re-enable interrupts.
-For example, Granary's basic block tracing tool writes to a CPU-private buffer containing the sequence of executed basic block program counters.
+However, interrupt delaying is sometimes useful because it enables non-reentrant instrumentation to avoid having to continually disable and re-enable interrupts.
+For example, Granary's \toolname{Guardrail} tool records traces of executed module instructions into CPU-private buffers \cite{Guardrail}.
+These traces are concurrently analyzed for bugs by another thread.
 Operating on a CPU-private buffer is more efficient than operating on a global buffer because it avoids cache line contention and locking on an SMP.
@@
 An equivalent tool developed for btkernel---which does not support interrupt delaying---would have to disable and the re-enable interrupts around the write.
-The is inefficient because it introduces flag stalls into the instruction pipeline \cite{AgnerMicroarchitecture}, and because interrupts arrive relatively infrequently.
 
@@
 
-Granary's mechanism of looking up branch targets and resolving control flow depends only on CTI policy and target information.
+Granary's mechanism of looking up branch targets and resolving control flow depends only on control flow instruction (CTI) policy and target information.
 This information is either directly embedded in emitted executable code, or in immutable data structures directly referenced by emitted code.
@@
 \end{figure}
-Like btkernel, Granary translates direct control-flow instructions (CTIs) into equivalent jumps into CTI-specific ``edge code" (\Cref{fig:direct_edge_code}).
+Like btkernel, Granary translates direct CTIs into equivalent jumps into CTI-specific ``edge code" (\Cref{fig:direct_edge_code}).
 Edge code lives outside the code cache, so as to not pollute the processor's instruction cache \cite{btkernel}.
@@
 Granary gains control and looks up or translates the target code according instrumentation policy stored in the edge data structure.
-Then, Granary uses modifies the original CTI in $B_{pred}$ in place so that it transfers control to $B_{succ}$  (\Cref{sec:patching}).
+Then, Granary modifies the original CTI in $B_{pred}$ so that it transfers control to $B_{succ}$  (\Cref{sec:patching}).
 Finally, Granary releases control and context switches to $B_{succ}$.
@@
 
-Granary translates indirect CTIs (\texttt{call}, \texttt{jmp}) into their equivalent direct control flow versions, where the direct versions jump to ``in-edge" code.
+Granary translates indirect CTIs (\texttt{call}, \texttt{jmp}) into their equivalent direct control flow versions, where the direct versions branch to ``in-edge" code.
 In-edge code loads the indirect jump target (e.g.
@@
 
-Granary's use of out-edge code avoids re-entrancy issues related to indirect control flow that present in DRK and btkernel.
+Granary's use of out-edge code avoids re-entrancy issues related to indirect control flow that are present in DRK and btkernel.
 Both DRK and btkernel store the target of an indirect CTI in a CPU-private memory location.
@@
 
-\subsubsection{Ahead-of-time Translation} \label{sec:aot}
+\subsubsection{Ahead-of-time Translation Optimization} \label{sec:aot}
 Granary ends basic blocks at conditional branches (\texttt{jCC}), direct and indirect \texttt{jmp}s, and \texttt{ret}s and \texttt{iret}s.
@@
 The repeated application of translating \texttt{jmp} targets forms traces in Granary's code cache.
-Granary optimizes traces by directly connecting basic blocks together when there is internal control flow within the blocks of the trace, and by eliding synthesized \texttt{jmp}s between adjacent basic blocks.
+Granary optimizes traces by directly connecting basic blocks together when there is internal control flow within the blocks of the trace, and by eliding \texttt{jmp}s between adjacent basic blocks.
 
@@
 That is, execution is more likely to fall through a conditional CTI than it is to take the conditional CTI.
-We experimentally verified Granary's tracing strategy by disabling tracing and relying on Granary's performance counters to count how many conditional branches were patched (by their edge code), versus how many synthesized fall-through \texttt{jmp}s were patched.
+We experimentally verified Granary's tracing strategy by disabling tracing and relying on Granary's internal performance counters to count how many conditional branches were patched (by their edge code, \texttildelow 25\%), versus how many synthesized fall-through \texttt{jmp}s were patched (\texttildelow 80\%).
 
@@
 
-\subsubsection{Patching and Alignment}\label{sec:patching}
+\subsubsection{Patching and Alignment  Optimizations}\label{sec:patching}
 For efficiency, basic blocks in Granary's code cache are directly connected through direct control flow instructions (\texttt{call}, \texttt{jmp}, \texttt{jCC}).
-That is, if there is a direct branch from one native instruction to another, then there will be a corresponding direct branch between two translated instructions in Granary's code cache.
-
-Granary opportunistically connects two basic blocks together using a direct branch if \begin{inparaenum}[i)]
+That is, if there is a direct branch from one native instruction to another, then there will be a corresponding direct branch in the translated code.
+
+Granary opportunistically connects two basic blocks together using a direct branch when: \begin{inparaenum}[i)]
 	\item the branch transfers control to another basic block within the current trace; or
@@
 An alternative when more than a few bytes of padding are needed is to introduce a single control-flow instruction (\texttt{jmp}) that jumps around the padded area.
-However, adding extra instructions introduces latency into the fetch/decode/execute cycle.
+However, adding extra instructions introduces latency into the processor's fetch/decode/execute cycle.
 Therefore, when possible, Granary uses ignored instruction prefixes to alter an instruction's alignment.
@@
 Finally, most basic blocks are small (i.e.
-fit in one cache line), therefore, aligning basic blocks on a cache line boundary naturally avoids the potential for a patchable instruction to straddle two cache lines.
+fit in one cache line), therefore, aligning basic blocks on cache line boundaries naturally avoids the potential for a patchable instruction to straddle two cache lines.
 
@@
 
-\subsubsection{Trace Allocation}\label{sec:trace_alloc}
+\subsubsection{Trace Allocation Optimization}\label{sec:trace_alloc}
 Arranging for commonly executed sequences of basic blocks to appear adjacent in memory, also called tracing, typically improves the runtime performance of instrumented programs \cite{DynamoRIO}.
-Typical approaches to tracing require a runtime profiling infrastructure (e.g.
-recording basic block execution counts).
+Typical approaches to tracing require a runtime profiling infrastructure \cite{DynamoRIOOptimization}.
 Granary attempts to reap the benefits of tracing without the runtime profiling overhead by opportunistically forming traces using ahead-of-time translation (\Cref{sec:aot}), and by arranging for related basic blocks to be allocated sequentially in memory.
@@
 
-Trace allocation is implemented by recording the memory pool used to allocate a given basic block in that basic block's meta-data.
+Trace allocation is implemented by recording the memory pool used to allocate a given basic block in that basic block's meta-data (\Cref{sec:metadata}).
 When Granary needs to translate a new basic block, it inspects the predecessor basic block's meta-data and uses the predecessor basic block's memory pool to service allocation requests for the basic block in translation.
@@
 
-\subsubsection{System Calls}
+\subsubsection{System Call Table Optimization}
 
@@
 Some instrumentation tools have a limited scope, and therefore do not need to instrument all code.
-For example, Granary's Guardrail tool is concerned with finding bugs in Linux device drivers \cite{Guardrail}.
-Guardrail records traces of executed instructions, and analyzes those traces to find memory access and concurrency bugs.
+For example, Granary's \toolname{Guardrail} tool is concerned with finding bugs in Linux device drivers \cite{Guardrail}.
+\toolname{Guardrail} records traces of executed instructions, and analyzes those traces to find memory access and concurrency bugs.
 Both phases are heavyweight, and so limiting the scope of what is instrumented to what is of interest (module code) results in better performance.
 Unfortunately, existing systems are unable to change the granularity of instrumentation.
-Fortunately, Granary allows Guardrail to set the granularity of instrumentation to only instrument a specific module, while leaving the rest of the kernel to execute without overhead.
+Fortunately, Granary allows \toolname{Guardrail} to set the granularity of instrumentation to only instrument a specific module, while leaving the rest of the kernel to execute without overhead.
 
@@
 
-Guardrail is an example of how tools can choose the instrumentation granularity that best matches their needs.
+\toolname{Guardrail} is an example of how tools can choose the instrumentation granularity that best matches their needs.
 That is, a Granary tool can choose to instrument some code, while leaving other code to execute natively.
 This is useful when the instrumentation in question has a high overhead or when only a subset of the kernel's code is of interest.
-\Cref{tab:granularities} provides a non-exhaustive list of the different granularities of instrumentation supported by Granary.
+\Cref{tab:granularities} contains a non-exhaustive list of tools using Granary's different instrumentation granularities.
+\Cref{fig:granularities} visually shows how different granularities of instrumentation can be mixed and matched to achieve the desired instrumentation visibility.
 
@@
 \end{tabularx}
-\caption{Visual example of using different granularities of instrumentation to selectively instrument or not instrument specific code.}
+\caption{\label{fig:granularities}Visual example of using different granularities of instrumentation to selectively instrument or not instrument specific code.}
 \end{figure}
@@
 Granary also interposes on the Linux kernel's module loading process.
-When a kernel module is dynamically loaded, Granary bootstraps by translating the first basic block of the module's initialization function.
-It then replaces the pointer to that function with a pointer to the translated basic block.
+When a kernel module is dynamically loaded, Granary bootstraps by page protecting the new module's code, and by translating the first basic block of the module's initialization function.
+It then replaces the kernel's pointer to the native initialization function with a pointer to the translated basic block.
 The translation process continues when the kernel initializes the module by invoking the translated module code.
-As with existing modules, the native code segments of dynamically loaded modules are page protected to prevent and recover from accidental execution.
+
+%As with existing modules, the native code segments of dynamically loaded modules are page protected to prevent and recover from accidental execution.
 
@@
 The act of registering a function is a signal to Granary that instrumentation must ``attach'' to the registered function when the kernel executes that function.
-Granary's basic attaching mechanism, page protection, raises a trap when registered functions are executed.
+In practice, when presented with such a signal, we would like to avoid the overhead of the trap that would occur when the registered function is executed by proactively replacing the registered function pointer with an instrumented equivalent.
+
+%Granary's basic attaching mechanism, page protection, raises a trap when registered functions are executed.
 In practice, we would like to avoid the overhead of the trap by proactively finding attach points and replacing them with their instrumented equivalents.
@@
 
-Granary uses static analysis of the kernel source code to dynamically discover potential attach points. Potential attach are recognized as being any of: \begin{enumerate*}
+%Granary's use and exposition of static analysis makes it the first kernel-space DBT system to bridge the gap between static and dynamic analysis.
+
+Granary statically analyses of the kernel source code to discover potential attach points.
+Granary bridges the gap between static and dynamic analysis by using the static type information to generate code that dynamically discovers potential attach points.
+This code also gives tools type-safe access to the runtime object graph.
+The static analysis recognizes potential attach points as being: \begin{enumerate*}
 	\item Function pointers passed as arguments to kernel functions.
@@
 
-Granary's static analysis phase generates ``wrappers'' for the kernel/module interface. Two kinds of wrappers are generated for the kernel/module interface: \begin{inparaenum}[i)]
+The code generation phase of Granary's static analysis creates ``wrappers'' for the kernel/module interface. Two kinds of wrappers are generated for the kernel/module interface: \begin{inparaenum}[i)]
 	\item kernel function wrappers (\Cref{fig:function_wrapper}), and

### 2013-12-13 VEE-14/paper.tex
Updated section 4.

@@
 \subsection*{Abstract}
-%Kernel modules extend the functionality of operating systems.
-They represent the bulk of new kernel code in development and contain the majority of new kernel bugs.
-Unfortunately, analyzing and debugging modules is very challenging.
-Static analysis of module source code is difficult because of the tight interaction between modules and the kernel.
-Some modules, however, are only distributed in a binary format, which makes static analysis intractable.
-Existing dynamic analysis tools capable of instrumenting modules are not suited toward module analysis.
-They either impose unnecessary performance overheads, lack flexibility and are too coarse-grained, or do not provide a means of understanding the ``big picture'' view of what an arbitrary module does as it executes.
-
-%We created Granary to address the challenges of module analysis.
-Granary is a framework for creating flexible and efficient tools that analyze or debug arbitrary, binary Linux kernel modules.
-Granary uses dynamic binary translation to dynamically rewrite and comprehensively instrument kernel modules.
-Granary makes it easy to create flexible tools by supporting context-aware runtime code specialization, and by integrating high-level static analysis information with low-level instruction manipulation.
-Module analysis tools built on Granary are efficient because of Granary's adoption of a relaxed transparency model, and because of Granary's ability to instrument module code without imposing overhead on non-module kernel code.
+TODO
 
@@
 
-%Operating system (OS) kernels present an apparently insane environment for a transparent binary instrumentation framework.
-For example, hardware interrupts can redirect control flow at any instruction, and interrupt delivery cannot be delayed past the execution of memory instructions that might alter the interrupt delivery state~\cite{DRK}.
-Preserving interrupt transparency, however, requires that interrupts be delayed until the end of any injected instrumentation instructions.
-Together, these requirements make it very difficult to write instrumentation tools.
-The insane environment of OS kernels also creates several opportunities for analysis tools.
-For example, interrupts can be ``tamed'' and put to new uses; privileged hardware features can be used; and the kernel ABI and source code can be relied upon to be a good predictor of module behavior at the kernel/module boundary.
-We also found that the kernel is largely insensitive to transparency issues, allowing us to relax transparency in exchange for both improved performance and greater visibility into the execution of instrumented code.
-This increased visibility is particularly important for analysis and debugging tools.
-
-%The challenges and opportunities of the kernel environment motivated the creation of Granary.
-Granary is a new binary instrumentation framework for the Linux kernel environment, designed specifically for analyzing kernel modules.
-Granary targets modules because they represent the bulk of new bugs and new code in development in the Linux kernel \cite{FaultsInLinux}.
-Granary allows developers to create flexible and dynamic analysis tools for arbitrary Linux kernel module binaries.
-Granary has two main goals: \begin{inparaenum}[i)]
-%	\item make it easy to create efficient, low-level analysis tools; and
-%	\item provide high-level static analysis information to the analysis tools.
-%\end{inparaenum} Granary achieves the first goal by using dynamic binary translation (DBT) to provide low-level access to instructions, memory, and interrupts.
-Granary achieves the second goal by giving tools direct access to the results of its own static analysis of the Linux kernel source code.
-By meeting these goals, Granary tools are able to do things like  figure out which data structure fields are associated with each memory access.
-
-%Granary addresses three limitations of existing program analysis systems: \begin{inparaenum}[i)]
-%	\item they have focused on instrumenting all code \cite{DRK,btkernel,QEMU};
-%	\item they do not allow for flexible runtime code specialization; and
-%	\item they do not provide static type information to the instrumented code.
-%\end{inparaenum}
-
 Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analyzed (e.g., modules, functions).
@@
 For example, if one wants to instrument RCU read-side critical sections, as is the case with Granary's \toolname{RCUdbg} tool, then instrumenting all code just to filter out and find those critical sections is wasteful.
-
-%This requirement motivates the design of \emph{mixed-mode execution}, which allows Granary tools to instrument only specific code while the majority of code executes with zero overhead.
-
-%This is challenging because Granary must be comprehensive: all code that is meant to be instrumented should be instrumented.
-In other words, relinquishing control to execute kernel code natively must not cause Granary to miss the execution of any module code.
 
@@
 
-%Also, the runtime checks needed to multiplex between different versions of instrumented code are orthogonal to the true goals of the instrumentation tool, thus distracting tool-developers from their primary objectives.
-
-%Using an existing system, 
-% get around the problem of needing different instrumentation in different contexts, existing systems need to create as many versions of the code as there are contexts.
-Runtime checks must also be inserted into the code that coordinate with the runtime to track and detect the current execution context and jump to the correct version of the code.
-Runtime checks and state tracking impose unnecessary complications to tool-authors.
-The added code needed to query the current context and switch to a different version of the instrumentation is code that is not actively achieving the primary goal of the tool, and is instead making up for a defeciency in the runtime system.
-%are orthogonal to the main goals of an analysis tool:
-% Aside from the additional ineffeciency of these runtime checks, there is also the 
-% However, generating many differently instrumented versions of the code introduces bloat, especially when not all versions are well exercised.
-Also, generating many different versions of the code introduces state tracking and switching costs.
-% This imposes unnecessary performance overheads because of the need to additional instrumentation that is unrelated to 
-% due to having to perform runtime detection and swichting
-% to the right version of the code for every basic block or every instruction.
-% according to \emph{all} contexts. Instrumenting code according to all contexts imposes unnecessary performance overheads and requires state tracking.
-%,  dynamically tracking the current context, and dynamically choosing which 
-% commit to instrumenting code to handle \emph{all} contexts. For example, if
-
 One example tool that uses Granary's JIT-based runtime code specialization feature is \toolname{RCUdbg}: Granary's Read-Copy-Update (RCU) debugging tool.
@@
 
-%There are similar checks for write-side pointers in write-side critical sections.
-% The instrumentation performed inside of read- and write-side critical sections is heavy weight, while code running in the third execution context can either run with lightweight or no instrumentation.
-%RCUdbg distinguishes between RCU-protected pointers and plain old pointers.
-There are two kinds of RCU-protected pointer (read-only, and read/write), and every access to an RCU-protected pointer is checked.
-The checks performed on an access to an RCU-protected pointer are context-specific.
-That is, a read-only pointer can only be read from within an RCU read-side critical section.
-Read/write pointers have similar restrictions.
-RCUdbg needs to know in which context some code is executing, i.e.
-in a read-side or write-side critical section, or outside of all critical section, and change instrumentation to check for different kinds of bugs based on the context.
-%A read/write pointer has different restrictions.
-In read-side critical sections, a read/write pointer can only be accessed by means of a special API function.
-In write-side critical sections, a read/write pointer can only be written to
-% One kind of RCU-protected pointer can be read from within ``RCU read-side critical sections", while another kind can (almost) arbitrarily accessed, but only inside of an ``RCU write-side critical section".
-Outside of these two kinds of sections, accessing RCU-protected pointers is illegal.
-The distinction between different contexts of code (read-side, write-side, outside), and different kins of pointers makes
-%Creating a tool to catch these kinds of bugs with an existing DBT system would be difficult because that tool would need to  
-%For example, Granary's Read-Copy-Update (RCU) debugging tool, \toolname{RCUdbg}, specializes instrumentation at different stages of use of the RCU synchronization API.
-Correct usage of the RCU API allows programmers to create scalable, shared data structures that permit concurrent access and modification.
-The ``trick" of RCU is that readers might not see the latest write to the data structure, and writers must synchronize to modify the data structure.
-\toolname{RCUdbg} distinguishes between code executing inside \emph{read-side} and \emph{write-side} critical sections, and instruments each code context differently.
-In read-side critical sections, \toolname{RCUdbg} detects bugs when readers access potentially non-existent data, do not the correct accessor API, or attempt to write to shared data.
-In write-side critical sections, \toolname{RCUdbg} detects when writers write to shared data without using the correct API.
-%In reader-side code
-%For example, Granary has a tool that  dynamically constructs inter- and intra-procedural control flow graphs (CFGs).
-The outputs of this tool enable Granary to do better runtime register allocation, which improves tool efficiency.
-Part of building CFGs requires tracking whether or not a given basic block is the entrypoint to a function.
-This is challenging with existing DBT systems because some optimized functions will re-execute their first basic block, which might incorrectly appear as new function calls in the CFGs.
-Granary solves this problem using demand-based runtime code specialization: every entrypoint basic block is instrumented in one way, and later re-executions of any entrypoint basic blocks are specialized and instrumented in another way.
-
 The \toolname{RCUdbg} tool is an example of \emph{context-driven code specialization}: code executing inside of a read-side critical section is instrumented differently from code executing inside of a write-side critical section, and code executing outside of both contexts runs natively or with lightweight instrumentation.
@@
 
-%Granary records 
-%Policy information is recorded inside of Granary's
-
-% instruction by telling Granary what policy to use to instrument the code 
-%, and when the \texttt{call} is eventually executed, 
-%This simplifies tool creation becase 
-%By propagating policy information through control-flow transferring instructions, 
-%Therefore, we can know at translation and instrumentation time that
-% No additional state needs to be maintained by \texttt{RCUdbg} to determine which version of the basic block to execute.
-%Any control-flow instructions the transfer control to other basic blocks that aren't the \texttt{rcu\_read\_unlock} function must also 
-%That is, the failure of existing systems is that they have no built-in support for maintaining multiple versions of code.
-Tools built on existing systems must emulate versioning by duplicating their instrumentation for each desired version, and add code that executes the right version at the right time.
-%, as would be done using existing systems, 
-%What makes Granary's runtime code specialization efficient is that it ties into Granary's \texttt{just-in-time} instrumentation model.
-That is, Granary only instruments code that will execute.
-Code that never executes is neither translated nor instrumented.
-Also, only one version of a basic block is generated at a time.
-Therefore, even though \toolname{RCUdbg} defines three policies, only
-%If some native code is about to execute, then Granary will only instrument it with the policy associated with 
-%, and an instrumentation policy is only applied to code if that code will execute within the context associated with that context.
-%defines three different instrumentation policies--one for each execution context--and declares switches 
-% instrumentation policies to 
-%only needs to define three instrumentation policies (one for each execution context), and declare 
-%different versions of the code are generated \emph{just-in-time}.
-Therefore, if the \texttt{memset} function is never executed inside of a read-side critical section, then \texttt{memset}'s code will never be specialized to perform the heavyweight checks needed to detect RCU-protected pointer misuses in this context.
-Similarly, if \texttt{memset} is executed both in a write-side critical section and outside of any RCU-related code, then two differently instrumented versions of \texttt{memset} will be generated.
-%However, unlike other systems, no runtime check that chooses to dispatch to one version of \texttt{memset} or another depending on the runtime context. Instead, 
-% if some code never executes in a particular context, then Granary will not generate an instrumented version of that code for that context.
-The second benefit of this approach is that it is stateless, and requires no runtime checks to switch between different versions of the code depending on the execution context.
-%Granary tracks the current execution context with zero runtime overhead by 
-%In a Granary tool, each tracked execution context is associated with its own instrumentation policy.
-An instrumentation policy decides how to instrument code that will execute within the associated context.
-%Different versions of the code are instrumented \emph{just-in-time}. That is, there will 
-%Granary is able to track execution contexts because it explicitly maintains policy information in translated basic blocks.
-This avoids typical poblems with tracking context in just-in-time DBT systems (e.g., maintaining runtime context-tracking state, reentrancy, and the inability to predict later execution contexts at translation time).
-% In the previous example, executing code is classified into two contexts: entrypoint and non-entrypoint basic blocks.
-In the case of our CFG tool, two policies track the two execution contexts: $P_{\mathit{call\_entry}}$ and $P_{\mathit{after\_entry}}$.
-$P_{\mathit{call\_entry}}$ applies to the first execution of the first basic block of each function, and connects functions in the inter-procedural CFG.
-$P_{\mathit{after\_entry}}$ applies to all other executed basic blocks within each function, including re-executions of the first basic block.
-$P_{\mathit{after\_entry}}$ connects basic blocks in the intra-procedural CFG.
-%Finally, an important missing feature of existing low-level instrumentation systems is the ``big picture'' view that one gets from static analysis information.
-Similarly, static analysis tools lack access to instructions, memory, and processor-specific behaviors that are only available to low-level runtime instrumentation tools.
-Existing mixed static/dynamic analysis tools \cite{NaCl,AddressSanitizer,ThreadSanitizer} appear to get the best of both worlds, but sacrifice on runtime flexibility by committing to a single instrumentation policy at a program's compile time.
-The benefits of high-level static analysis motivated its inclusion into Granary using a technique we call \emph{reifying instrumentation}.
-Granary statically analyzes the Linux kernel source code to bootstrap learning about  the interfaces that enable modules and the kernel to interact.
-Granary exposes this information to instrumentation tools in the form of type and function \emph{wrappers}.
-Tools can use wrappers to apply context-specific policies when instrumenting module code as well as to inspect and manipulate parts of kernel/module memory in a type-safe way.
-%For example, we are actively developing a tool that uses static type information to learn about all the fields of all kernel data structures that are read from or written to by modules.
-This information allows us to model things like how a typical file system operates on kernel data structures in the process of opening a file.
-From these models, we can construct behavioral patterns about how modules use kernel data structures and functions.
-These patterns can be used to classify modules \cite{DeviceDriverClassification} and identify spurious behavior \cite{LXFI}.
-
 The rest of the paper describes Granary in more detail.
 \Cref{sec:dbt} describes how Granary is implemented, \Cref{sec:what} describes how a Granary tool can choose to instrument only specific code, \Cref{sec:how} describes how Granary change the policy used to instrument code, etc.
-
-% \Cref{sec:modes} describes how to efficiently implement mixed-mode execution.
-\Cref{sec:policies} describes a new method for dynamically changing how code is instrumented.
-\Cref{sec:reify} describes how to integrate static analysis information into a DBT system.
-Finally, \Cref{sec:eval} evaluates Granary's performance.
 
@@
 
-%Granary translates and instruments code at one of three default granularities: kernel, module, and function (also called probe-based instrumentation). 
-%
-%At runtime, the granularity of instrumentation can be altered based on the execution context.
-For example, a tool analyzing modules for bugs can choose not to instrument a particular function (e.g.
-whitelist the function) if that function is known to not contain bugs, or known to generate many false-positives.
-%
-%\paragraph{Kernel}
-%Granary interposes on the system call and interrupt entrypoints, located in the x86-64 \texttt{MSR\_LSTAR} model-specific register and interrupt descriptor table (IDT), respectively. 
-%
-%\paragraph{Module}
-%
-%
-%\paragraph{Function}
-%
-%Granary can interpose on the execution of specific kernel functions in three ways:
-%\begin{enumerate}
-%	\item {\bf Wrapping:} Granary can substitute the execution of a kernel function when that function is invoked from instrumented code.
-A different function can be substituted depending on the context from which the original function is invoked (instrumented kernel code, instrumented module code).
-We call this type of interposition ``wrapping'' because it is commonly used to inspect the substituted function's arguments and return value, while still invoking the original function.
-%	\item {\bf Substituting:} Granary can substitutes the execution of a kernel function, even when that function is executed by native kernel code.
-This is useful when tools only instrument part of the kernel (e.g.
-modules or functions) but require visibility on specific kernel events happening outside of the instrumented code (e.g.
-particular memory allocations).
-%	\item {\bf Instrumented Wrapping/Substituting:} Granary can optionally instrument wrapped or substituted code.
-This is useful when tools want to observe function arguments/return values using high-level C/C++ code, but where that function must be instrumented for visibility.
-%\end{enumerate}
-
 Granary translates and instruments kernel binaries one basic block at a time.
@@
 interrupt vectors, system call entrypoints, module initializers).
-Entrypoint basic blocks  are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as execution ``discovers'' them.
+Entrypoint basic blocks  are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as control flow reaches them for the first time.
 
@@
 
-%Granary code executes with interrupts disabled and depends on CPU-private data to avo
-%Granary's translation mechanism depends on CPU-private data to avoid contention when two or more processors simultaneously execute Granary code.
-%because kernel stacks are typically small (4K to 8K in size), and because context-switches into Granary can occur at arbitrary program points, Granary cannot risk 
-%and a context switch into Granary might occur at any point. If most of the stack is already used by current thread before it yields control to Granary then
-
 \subsection{Basic Block Meta-Data}\label{sec:metadata}
@@
 Granary is unique among DBT tools in that the debugging experience is actually pleasant.
-For example, Granary has a built-in debugging mechanism for stepping back in time through previously executed basic blocks, as well as querying past register state.
+For example, Granary has a built-in reverse debugging mechanism, which allows developers to inspect previously executed basic blocks and past register state.
 	\item Linux kernel-specific exception table meta-data about instructions contained within the current basic block.
@@
 For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if code executing within a read-side critical section is interrupted (\Cref{sec:rcudbg}).
-\end{itemize} %
-%Granary uses basic block meta-data for trace allocation (\Cref{sec:trace_alloc}), interrupt delaying (\Cref{sec:interrupt_delay}), and policy-specific interrupt handling.
-
-Basic block meta-data can be also be extended by instrumentation tools.
+\end{itemize}
+
+Basic block meta-data can also be extended by instrumentation tools.
 This makes developing instrumentation tools that record per-basic block statistics very easy to make.
@@
 
-%Basic block meta-data can also contain Granary's intermediate representation for x86-64 instructions.
-This is a unique feature that enables instrumentation tools to implement their own form of hot-code patching (among other things), as the intermediate representation remains tied to an emitted x86-64 instruction once the block is committed to the code cache.
-
-%Unlike any other instrumentation tool, Granary allows the x86-64 instruction intermediate representation
-
-%Another
-
-%novelty of Granary is that basic block meta-data enables a good
-
-%Basic block meta-data is queried and extended by Granary instrumentation tools.
-
-%// TODO!
-
-%Block meta-data is also queried by interrupt handlers when deciding how to handle interrupts and exceptions in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block.
-This enables policy-specific interrupt handling.
-That is, when a basic block instrumented by a specific policy is interrupted, that policy can choose how to handle the interrupt.
-For example, Granary's address watchpoints framework can instrument code using a lightweight, \toolname{null} instrumentation.
-
 \subsection{Transparency}\label{sec:transparency}
@@
 By default, Granary uses a relaxed transparency \cite{btkernel} model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility.
-By default, Granary exposes the following two artifacts: code cache addresses as function and interrupt return addresses.
+By default, the only artifact Granary exposes is that interrupt and function return addresses can point into the code cache.
 
@@
 
-%performing a \texttt{ret} requires an additional \texttt{jmp} to 
-%Neither of DRK's nor btkernel's approaches enable native-like return address prediction by the CPU.
-%and then jumping to the next basic block.
-This in turn .
-Btkernel's approach to procedure calls introduces similar inefficiencies, despite the fact that does not provide return address transparency, because it also emulates procedure calls, but by manually pushing the instrumented return address onto the stack instead of the native one.
-%Granary's approach improves on DRK and btkernel by allowing 
-%Better return-address prediction by the processor.
-
 	\item {\bf Improved visibility.} Not ending basic blocks on \texttt{call} instructions results in longer basic blocks, which in turn provide additional visibility to tools at instrumentation time into what will be executed by the current thread.
@@
 
-	%longer basic blocks, which provide more code visibility to instrumentation tools at instrumentation time.
 \end{enumerate}
@@
 
-%\paragraph{Module function wrappers} Granary does not maintain transparency when its type wrappers replace pointers to module functions with pointers to wrapped module functions.
-This is perhaps the riskiest break in transparency within Granary because a module might treat a function pointer as being representative of the module being in some particular state.\footnote{We have not yet encountered a module where changing pointers to module functions into module wrappers altered the module's behavior under instrumentation.} If Granary is configured to use transparent module function pointer addresses in shared data structures then fast attaching is disabled and tools will not have access to the same level of static program information.
-
 \subsection{Reentrancy}
@@
 Together, these requirements make it very difficult to write instrumentation tools.
-
-%Another concern is that instrumentation itself is not re-entrant. For example, instrumentation instructions added by a tool that access CPU-private information
 
@@
 All other parts of Granary are fully reentrant.
-
-%Therefore, if a context switch into Granary occurs because the next basic block must be found/translated, then the address of that next basic block must be stored \emph{somewhere} in order for execution to eventually reach that basic block when the context switches back to the code cache.
-But context switching is not allowed to perturb the register state of the instrumented program, and so the target address must be resolved by another means.
-
-%Like btkernel, Granary solves this problem of figuring out how to get to where the code needs to go  by generating ``edge'' code for each direct control-flow instruction (e.g.
-\texttt{call}, \texttt{jmp}, \texttt{jcc}).
-
-%Granary also generates edge code for indirect control-flow instructions; however, this edge code is split into two separate components: entry and exit edge code.
-Entry edge code is generated on a per-indirect CTI instruction basis, and spills registers onto the runtime call stack.
-Entry edge is responsible for indirect branch lookup, and will indirectly jump to the exit edge code for a given basic block.
-Exit edge code is generated for each basic block that is targeted by an indirect CTI.
-Exit edge code restores the register state saved by the entry edge code, and then directly jumps to the beginning of the targeted basic block.
 
@@
 
-%If the pre-instrumentation is not idempotent, then its re-execution will result in undefined behavior.
-We encountered this issue when using DRK for kernel module instrumentation, where our pre-instrumentation spilled registers to the runtime call stack, and our post-instrumentation restored the spilled registers.
-Because of interrupt delaying, the spilling code was sometimes repeated many times, whereas the restoring was only executed once.
-This led to hard-to-debug kernel crashes.
-
 Btkernel provides no solution to the problem of handling non-reentrant instrumentation, beyond requiring that instrumentation tools disable and re-enable interrupts around non-reentrant instrumentation code \cite{btkernel}.
@@
 
-% and re-issues the interrupt to Granary.
-Granary then emulates a return from the current interrupt handler (without issuing an \texttt{iret} instruction), but leaves interrupts disabled and redirects execution into the copied code.\footnote{Leaving interrupts disabled is required to ensure that executing the copied instructions in the CPU-private buffer is reentrant.
-Granary performs some sanity checks on the instructions being copied to ensure that they don't re-enable interrupts.} When the copied code finishes, Granary regains control and observes the reconstructed interrupt stack frame.
-At this point, Granary or the tool can choose to handle the interrupt or defer the handling of the interrupt to the kernel.
-
-%while not supporting delaying has value because not all instrumentation is non-reentrant.
-By default, Granary does not delay interrupts.
-However, instrumentation tools can declare \emph{interrupt delay ranges}:
-
-%, but introduces implementation complexities and high overheads for interrupt-heavy workloads. Precise interrupt delivery is one of the ways which DRK maintains transparency: interrupt handlers only 
-%The interaction between DRK's interrupt delaying mechanism and how also makes designing instrumentation subtlely harder.
-For example, Granary's \toolname{RCUdbg} tool injects instrumentation inside of read-side critical sections that looks for loads/stores from tainted addresses.
-When necessary, \toolname{RCUdbg} spills and restores registers around each checked load/store operation, so that the registers can be safely used by the checking code.
-This is an examuple of pre and post instruction instrumentation: some instructions are injected before a native instruction, and some instructions are injected after.
-DRK is oblivious to the relationship between pre/post instrumentation, and so an interrupt that occurs within the pre-instrumentation will only be delayed until the native instruction.
-
 \subsection{Control Flow Implementation}
@@
 
-\subsubsection{Direct Control Flow}
-% Code block showing an example JZ (jump if ZF=1).
+\subsubsection{Direct Control Flow}\label{sec:direct_control_flow}
 \lstset{language=[x64]Assembler}
@@
 
-% and is responsible for transferring control from a basic block ($B_{pred}$) to Granary so that the next basic block ($B_{succ}$) can be found/translated, and then connected ($B_{pred} \underset{\texttt{jz}}{\to} B_{succ}$) via hot code patching.
-
 \subsubsection{Indirect Control Flow}\label{sec:ibl}
-
-%Indirect branch lookup through a global, fixed-size hash table.
-A basic block making an indirect control flow transfer jumps to an in-edge that knows where to find the indirect CTI target (e.g.
-in a specific register or memory location).
-The in-edge then uses the target as an index into the hash table, and jumps to the target address stored in the indexed hash table entry.
-That entry will either be the address of an out-edge, or of Granary's indirect branch resolver.
-Each out-edge is specific to a target basic block, and first checks if the intended jump target matches the out-edge's target.
-If so, then control is transfered to the target basic block, otherwise control is transfered either to Granary or to the next out-edge.
-Granary chains together out-edges through conditional jumps just as
-
-%Hash table entries are code addresses of out-edge code stub.
-Each out-edge code stub is specific to a target basic block, and is responsible for checking if the intended control-flow target is the out-edge's target or not.
-If not, the out-edge either directly jumps to the next out-edge whose target also fits into the same hash table bucket, or jumps to Granary to determine the target basic block.
 
@@
 
-%If an entry points to Granary, then no as-of-yet resolved indirect jump target hashes to that entry.
-
-%Granary implements separate chaining through code (i.e. control flow jumps, as opposed to date, e.g. linked list pointers)
-%In the event that one-or-more indirect jump targets do hash to an existing entry
-%In-edge code is responsible for 
-% to first search for their targets in a global, fixed-size hash table, and failing that, context-switch into Granary to perform a full lookup and potential translation.
-%using a fixed-size hash table, where the native target of the indirect
-
 \subsection{Optimizations}
@@
 
-%using internal performance counters within Granary
-
-%Static branch prediction hints (\texttt{likely}, \texttt{unlikely} macros) are heavily used in the Linux kernel source code.
-Although static branch prediction hints can be specified on individual conditional branch instructions, recent x86-based architectures ignore these hints \cite{AgnerMicroarchitecture}.
-Therefore, static branch prediction hints at the source code level mainly affect compiler optimizations, including intraprocedural basic block scheduling.
-Our tracing strategy is based on
-
-%Tr
-
-%We expect and have experimentally verified that the most common execution path through code is straight-line and does not following conditional branches.
-
-%Due to instruction prefetching, we expect that the likely or
-
-%On recent x86-based architectures, these hints only affect the schedule of basic blocks in the intraprocedural control-flow graph \cite{AgnerMicroarchitecture}.
-
 \subsubsection{Patching and Alignment  Optimizations}\label{sec:patching}
@@
 Therefore, patchable instructions (or parts thereof) must be aligned so as to not straddle two cache lines.
-
-%Padding basic blocks so as to align specific instructions must not change the execution behavior.
 
@@
 
-%To reduce the overhead of alignment 
-%In some cases, basic blocks can be directly connected during basic block translation time.
-For example, intra-trace CTIs, or CTIs targeting already translated basic blocks.
-However, not all basic blocks can be connected this way, therefore some runtime
-%Granary connects some basic blocks at translation time (e.g.
-intra-trace control-flow instructions).
-This is common when tracing is enabled, but also occurs when a branch targets an already translated basic block.
-However,
-%Some basic blocks are directly connected at basic block translation time.
-One example is branches that target other basic blocks within a trace.
-Another example is a branch to an already translated basic block.
-%Other basic blocks are only connected at runtime.
-For example, if the function \texttt{memset} has never been executed, then a \texttt{call} to \texttt{memset} will be translated to target edge code (\Cref{fig:direct_edge_code}) that, when executed, will translate \texttt{memset}'s first basic block and patch the \texttt{call} to target the translated basic block.
-%For example, the targets of function call instruction (\texttt{call}) are resolved just-in-time (unless the target is already known).
-For each translated \texttt{call} instruction, Granary generates edge code (\Cref{fig:direct_edge_code}) that will find or translate the first basic block of the function being called, and patch the translated \texttt{call} instruction to point to the function's first basic block.
-%For example, if a direct branch in one basic block targets another block within the trace under construction, then that 
-%Basic blocks are directly connected at block translation time when branches in traces re-enter a trace, or when branches target previously translated basic blocks.
-If a branch does not target another basic block within its trace, or does not target a previously translated basic block, then the branch will target edge code that will resolve that branches target at runtime.
-% However, when a given basic block is translated, not all of its targets are known (e.g.
-\texttt{call} and \texttt{jCC} targets).
-Therefore, edge code (\Cref{fig:direct_edge_code}) is generated that translates/resolves a branch target when that branch is first executed, and patches the translated branch to avoi
-% The targets of some of these direct CTIs are only resolved at runtime, by executing edge code.
-As an optimization, jumps to edge code are dynamically patched to transfer control to their intended targeted basic blocks.
-%edge code use hot code patching to update a translated CTI to redirect execution
-
 \subsubsection{Trace Allocation Optimization}\label{sec:trace_alloc}
@@
 
-%, and is associated with a private memory pool from which to allocate memory for basic blocks.
-The net result is that most of the basic blocks eecuted in service of a single system call are allocated from that system call's memory pool.
-%The intuition for this approach is that kernel code is not very ``loopy", and so 
-%Granary services memory allocation requests for basic block instructions from memory pools. By default, Granary has as many memory pools as there are CPUs. However, when trace allocation is enabled
-%Tracing is also beneficial for 
-%However, tracing of this sort requires mechnisms for evicting stale basic blocks, 
-%Constructing traces of ``hot'' 
-%JIT generally results in better performance .
-%By default, Granary allocates the memory needed to hold the instructions of a basic block from CPU-private memory pools.
-Each memory pool serves basic block allocation requests out of fixed-size memory blocks, ranging from 4K to 24K in size.
-
 \subsubsection{System Call Table Optimization}
@@
 
-%re-direct jumps through the system call table
-
 Granary's translation post-processing mechanism identifies indirect control flow through the system call table and alters those CTIs to direct control flow through the shadow system call table instead.
@@
 Upon their first execution, these entrypoint translators find or translate the first basic block of the system call handler, and update the shadow system call table accordingly.
-
-%Directing control through the shadow system call table avoids the overhead of going through the indirect branch lookup (IBL) table (\Cref{sec:ibl}), and reduces the number of potential conflicts in the IBL table.
-% whose entries are code cache basic block addresses translated according to the starting instrumentation policy.
-%\subsection{Reproducability}
-%Hello
 
@@
 All non-module kernel code executes with zero overhead.
-& Granary's Guardrail tool records traces all instructions and memory operations performed by device driver code \cite{Guardrail}.
+& Granary's \toolname{Guardrail} tool records traces all instructions and memory operations performed by device driver code \cite{Guardrail}.
 A sideline thread analyzes the recorded traces to detect data race bugs and memory access bugs.
@@
 \\
-
-% Granary's \toolname{Irrigator} tool find leaks on module-owned memory (\Cref{sec:function_wrapper}).
-\toolname{Irrigator} instruments only module code, but also interposes on specific kernel functions (kernel memory allocators and deallocators) to gain visibility into kernel code that would normally execute natively (using module-only instrumentation).\\
 \hline
-%Basic Block & Individual basic blocks are instrumented on demand.
-& Granary's address watchpoints framework can be tuned to instrument only code accessing specific data.
-A general protection fault occurs when native code dereferences a  ``watched address".
-Granary can transparently handle this fault by returning execution to an instrumented version of the faulting code (where the instrumentation prevents such faults).
-Native execution can resume at the end of the basic block.
-\\
-%\hline
 \end{tabularx}
@@
 
-%The closest that an existing system can get to
-
-% Therefore, if Guardrail were implemented using DRK or btkernel, then it would have to implement its own form of runtime code specialization
-
-%. Recording and analyzing every instruction executed by the kernel is
-
-%Guardrail records and analyzes traces of device driver code to find bugs.
-However, Guardrail's analysis can be heavyweight.
-If Guardrail were forced to instrument all code (e.g.
-with DRK or btkernel), then Guardrail would impose significant performance overheads.
-However, if Guardrail's recording were limited to only module code (
-
-%and so limiting the analysis to only device driver code results in better performance.
-
-%The recording and analysis phases of Guardrail are heavyweight, therefore, limiting
-
-% Instrumenting the whole kernel to observer driver (module) execution is unnecessary and potentially very inefficient (especially if the tool is meant for online analysis). That is,
-
 \toolname{Guardrail} is an example of how tools can choose the instrumentation granularity that best matches their needs.
@@
 \Cref{fig:granularities} visually shows how different granularities of instrumentation can be mixed and matched to achieve the desired instrumentation visibility.
-
-%instrument all kernel code, specific module code, or only specific function code.
-%what code should be instrumented.
-Example granularities of instrumentation include the whole kernel, only specific modules, only specific functions, or on-demand instrumentation based on events (e.g.
-interrupts or exceptions).
 
@@
 
-%As with existing modules, the native code segments of dynamically loaded modules are page protected to prevent and recover from accidental execution.
-
-%Granary interposes on an existing module by using hardware page protection to mark the module's code as non-executable and trapping attempts to execute the protected code.
-
 While correct, hardware page protection is an inefficient mechanism for gaining control of module code execution.
@@
 In practice, when presented with such a signal, we would like to avoid the overhead of the trap that would occur when the registered function is executed by proactively replacing the registered function pointer with an instrumented equivalent.
-
-%Granary's basic attaching mechanism, page protection, raises a trap when registered functions are executed.
-In practice, we would like to avoid the overhead of the trap by proactively finding attach points and replacing them with their instrumented equivalents.
-
-%Therefore, to Granary, any function pointer passed from the module to the kernel is a potential future attach point.
-%registered module functions represent potential future \emph{attach} points, where a switch from native to instrumented execution will occur. 
-%Furthermore, Granary discovers additional attach points at detach points by observing pointers to module functions that are passed by modules to the kernel.
-
-%Granary's use and exposition of static analysis makes it the first kernel-space DBT system to bridge the gap between static and dynamic analysis.
 
@@
 \end{enumerate*}
-
-%Any function pointer passed as an argument to a kernel function, or stored in a structure that can be passed as an argument to a kernel function is treated as a potential attach point.
 
@@
 
-%\Cref{fig:type_wrapper} shows an example of a type wrapper for the Linux device driver structure.
-
 If a pointer to a module function (a future attach point) is discovered, then Granary replaces that pointer with a function-specific \emph{module function wrapper}.
 Module function wrappers behave similarly to kernel function wrappers, in that they provide Granary more visibility into potential attach points, and provide instrumentation tools visibility into the runtime object graph.
-
-% instrumentation tools a mechanism to observe a
-
-%are responsible for wrapping their arguments. When an argument
-
-%At runtime, these wrappers interpose on kernel functions and traverse in-memory data structures, find attach points, and convert them into
-
-%At runtime, Granary detects and interposes on potential attach points by wrapping the kernel/module interface.
-
-%Kernel functions are wrapped by \emph{kernel wrappers} that inspect and traverse argument pointers in search of pointers to module functions.
-Granary inspects and modifies arguments to module wrappers in the same way as for kernel wrappers.
-This allows Granary to discover kernel entry points that will cause instrumentation to detach.
-Finally, Granary redirects execution to the appropriate kernel or instrumented module function after wrapping has occurred.
-
-%Kernel and module wrappers invoke \emph{type wrappers} to find and wrap pointers to module functions that are directly or indirectly referenced by kernel/module function arguments.
-A type wrapper is a function that recursively  traverses the in-memory object graph and converts pointers to module code into pointers to wrapped module functions.
-Type and kernel wrappers are automatically generated at Granary's compile time by scripts that statically analyze the kernel source code.
-Granary automatically matches any variable in a kernel wrapper to a type wrapper if the base type (absent pointers, specifiers, and qualifiers) of that variable matches the type wrapper's wrapped type.
-Similarly, Granary automatically generates module wrappers using a combination of compile-time meta-programming and runtime code generation to match type wrappers to the declared arguments of \texttt{C} function pointer types.
-Granary's wrapper approach has two benefits: (i) it is more efficient than the trap-based alternative, and (ii) it gives instrumentation tools direct and semantically meaningful access to data crossing the kernel/module boundary.
-
-%\label{sec:implicit_attach}
-%After Granary bootstraps on the module initialization process, \emph{attaching} occurs in one of three ways: \begin{enumerate}
-%	\item {\bf Implicit attaching:} the kernel returns to instrumented module code in the code cache.
-This occurs when instrumented moduel code is interrupted, or when the module invokes a kernel function (detaching) and that function returns to the instrumented module code (attaching).
-%	\item {\bf Fast attaching:} the kernel invokes a wrapped module function.
-The first instance of fast attaching for any kernel module is when the kernel's module loader invokes the instrumented module initializer function.
-%	\item {\bf Slow attaching:} the kernel invokes unwrapped module code.
-If the kernel executes a module function which was passed to the kernel in a type-unsafe manner, then the processor will raise a fault because Granary uses hardware page protection to prevent module code from being executed.
-Granary handles these faults by returning execution to the instrumented version of the faulting module code.
-%\end{enumerate}
-%
-%\label{sec:detach}
-%Granary \emph{detaches} when control transfers from instrumented code to native (uninstrumented) code. Detaching occurs in one of two ways: \begin{enumerate}
-%	\item {\bf Implicit detaching:} instrumented module code returns to the kernel code which originally invoked the module, or is interrupted (initial interrupt handling is done by the kernel).
-%	\item {\bf Wrapped detaching:} instrumented module code invokes a kernel wrapper, which later transfers control to the kernel.
-%\end{enumerate}
 
@@
 
-Instrumenting individual functions is useful in when the code of interest is only a function or group of functions, or when the tool does not instrument all code, but wants visibility into functions that would otherwise execute natively.
+Instrumenting individual functions is useful when the code of interest is only a function or group of functions, or when the tool does not instrument all code, but wants visibility into functions that would otherwise execute natively.
 For example, \toolname{RCUdbg} interposes on the \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} kernel functions.
@@
 
-%For example, Granary's \toolname{Irrigator} tool detects memory leaks in kernel modules.
-\toolname{Irrigator}  finds leaks of ``module-owned" memory.
-We say that the some memory is owned by a module iff that memory is deallocated by a module.
-Unfortunately, ownership of module-allocated memory (for an arbitrary, binary module) is not known \emph{a priori}.
-Therefore, \toolname{Irrigator} conservatively assumes that all module-allocated memory is module-owned, unless it is deallocated by the kernel.
-\toolname{}
-
-%However, the instrumentation granularity of \toolname{Irrigator} is module code, and so (almost all) kernel code executes natively.
-Fortunately, Granary allows \toolname{Irrigator} to instrument specific kernel functions---in this case memory deallocators---and detect when the kernel frees module-allocated memory.
-\toolname{Irrigator} also interposes on specific kernel memory allocators (e.g.
-for the \texttt{sk\_buff} data structures), which are allocated by the kernel but deallocated by modules (e.g.
-network device drivers deallocate \texttt{sk\_buff} data structures).
-
-%We also use probe-based instrumentation to assign module ownership to kernel-allocated \texttt{sk\_buff} data structures, which are allocated by the kernel but deallocated by network device driver modules.
-
-%: \begin{inparaenum}[i)]
-%	\item the instrumentation tool is only interested in that particular function; or
-%	\item when the instrumentation tool needs visibility into that function
-%\end{inparaenum}
-
-%Granary can instrument or not instrument functions in several different ways.
-For example, \emph{probe-based} instrumentation allows Granary to selectively instrument only a particular function, while \emph{wrapper-based} instrumentation allows Granary to isolate a particular function that is invoked by instrumented code, and choose to instrument that function or run it natively (all while having the flexibility to arbitrarily inspect that function's arguments and return values in a type-safe way).
-
-%\subsubsection{Probe-based instrumentation}
-
-%Like KernInst \cite{KernInst,AnyWhereAnyTimeDBT}, Granary can take over the execution of individual kernel functions.
-
-Granary uses code splicing to allow tools to interpose on the execution of individual functions invoked by native code (which nornmally execute outside of Granary's control) \cite{KernInst,AnyWhereAnyTimeDBT,KProbes}.
+Granary uses code splicing \cite{KernInst,AnyWhereAnyTimeDBT,KProbes} to allow tools to interpose on the execution of individual functions invoked by native code (which nornmally execute outside of Granary's control).
 Granary allows tools to interpose on specific functions invoked by instrumented code using kernel or module function wrappers (\Cref{sec:wrappers}).
@@
 
-% Instrumenting invidual functions, called \emph{probe-based instrumentation}, giv \cite{KProbes}.
-
-%Probe-based instrumentation allows an instrumentation tool to replace \emph{all} executions of a particular kernel function, regardless of whether or not the replaced function is executed by native or instrumented code \cite{KProbes, AnyWhereAnyTimeDBT, KernInst}.
-Probe-based instrumentation can be used to wrap a particular function without instrumenting its code, or to wrap and instrument the probed function's code.
-
-%Probe-based instrumentation provides visibility into specific kernel events (e.g.
-memory allocations) to instrumentation tools that don't instrument the whole kernel (e.g.
-module-only, \Cref{fig:module_probe}).
-One example tool that combines module-only and probe-based instrumentation is
-
-%\subsubsection{Wrapper-based instrumentation}
-%
-%Wrapper-based instrumentation makes use of the same wrapper capabilities used by module-only instrumentation (\Cref{sec:module}), but is available regardless of the instrumentation granularity.
-Wrapper-based instrumentation is similar to probe-based instrumentation, but permits more selectivity in terms of the context in which the function will be wrapped.
-That is, a function can be wrapped in one way when executed by instrumented kernel code, and in another way when executed by instrumented module code.
-Finer-grained selectivity can be achieved by having individual instrumentation policies manually decide how to wrap a given function.
-%
-%\begin{figure}[t!]
-%\lstset{language=C, tabsize=2, stepnumber=1}
-%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-%MODULE_FUNCTION_WRAPPER(__kmalloc, (size, gfp), {
-%  void *addr = __kmalloc(size, gfp);
-%  if(is_valid_address(addr)) {
-%    add_watchpoint(addr, addr, size);
-%  }
-%  return addr;
-%})
-%\end{lstlisting}
-%\caption{Example module function wrapper for the \texttt{\_\_kmalloc} function.
-This wrapper adds an ``address watchpoint" to the address returned by the kernel's \texttt{\_\_kmalloc} memory allocator.
-Granary's module memory leak detector uses address watchpoints to track the liveness of module-allocated memory.
-Every access of a ``watched address" marks the associated memory as live within the current epoch.
-Memory that has not been live for several consecutive epochs is treated as a potential leak, and a is checked for reachability using a heap memory scanner.}
-%\label{fig:malloc_wrapper}
-%\end{figure}
-%
-%Granary's module memory leak detector is one example application that uses the selectivity of wrappers (with respect to their calling context) to distinguish between module-allocated and kernel-allocated memory.
-\Cref{fig:malloc_wrapper} shows a module function wrapper that adds ``address watchpoints"  \cite{BehaveOrBeWatched} to module-allocated memory.
-
-%Granary can instrument, or not instrument, particular function
-
-%\subsection{On-Demand Instrumentation}
-%
-%On-demand instrumentation makes use of Granary's control over interrupt and exception handling.
-When a specific exception or interrupt occurs, a Granary tool can turn on instrumentation for the interrupted code.
-This ties in well with Granary's address watchpoints system, which taints memory addresses by changing their high-order 16 bits into a ``non-canonical" form.
-Watched addresses trap when accessed by native code, which introduces an opportunity for on-demand instrumentation to be enabled.
-Another application of on-demand instrumentation is data watchpoints based on hardware page protection and memory shadowing.
-%
-%Turning on instrumentation when specific data is accessed makes on-demand instrumentation inherently \emph{data-centric} as opposed to \emph{code-centric}.
-That is, instrumentation is targeted at instructions accessing specific data, instead of the typical ``dragnet" approach employed by most DBT systems, which instrument all memory loads/stores so as to capture the select few that access key data.
-
-% instrumentation turns on when specific data is accessed, and can be turned off at the granularity of an instruction, a basic block, the current executing function, or the current executing task/thread.
-
-%\section{Mixed-Mode Execution}\label{sec:modes}
-%Granary supports two modes of execution: instrumented and native.
-Module code is instrumented and executes from Granary's code cache, which is under Granary's control.
-Non-module kernel code runs natively.
-A mode switch occurs when execution transfers between native and instrumented code.
-Some mode switches happen naturally (e.g., when instrumented code returns to native code) and other mode switches are mediated by Granary (e.g., when instrumented module code invokes a kernel function).
-%
-%The mode switch from instrumented module code to native kernel code is easier to detect since the instrumented code runs under Granary's control.
-Granary treats all kernel functions as \emph{detach} points, where a mode switch from instrumented to native code occurs.
-At these points, Granary stops executing.
-%
-%However, Granary needs a way to regain control when native kernel code invokes module code.
-Our fallback solution for regaining control uses hardware page protection to trap attempts by the kernel to execute native module code.
-We handle these traps by redirecting execution to instrumented module code.
-While comprehensive, this approach is not ideal because: \begin{inparaenum}[i)]
-%	\item trapping on every execution attempt introduces overhead; and
-%	\item the trap does not provide sufficient information about which interfaces were being used by the kernel to invoke the module.
-%\end{inparaenum}
-
-\section{Context-Driven Instrumentation}\label{sec:how}
-
-Granary gives tools the ability to be selective about what is instrumented.
+\section{Runtime Instrumentation Code Specialization}\label{sec:how}
+
+Existing instrumentation systems make it challenging to change how code is instrumented based on the runtime context.
+In practice, not all code needs or should be instrumented in the same way.
+Selectivity is on example of this: not all code needs to be instrumented, which is why Granary lets tools run some code natively and other code under instrumentation.
+Just as selectivity allows a tool to choose which code to instrument, specialization allows a tool to choose how to instrument that code.
+Specialization is especially useful for analysis tools that distinguish between light and heavyweight analyses, and for debugging tools that do different bug checks in different contexts.
+For example, \toolname{RCUdbg} instruments code running inside RCU read-side critical sections differently from code running outside of read-side critical sections.
+Code is instrumented differently in these two contexts because a memory access that is valid inside of a read-side critical section can be invalid outside of a read-side critical section.
+If the instrumentation could not be specialized for these two different contexts, then \toolname{RCUdbg} would need insert instrumentation on every memory access that first determines the current context, and then performs the appropriate bug check.
+This additional check in the instrumentation introduces performance and space overheads (additional instructions) for \emph{all} memory instructions, as well as state management complexities for the runtime itself.
+That is, \toolname{RCUdbg} would need to track the current context in a reentrant way.
+%These additional overheads and complexities
+
+%n doesn't just introduce performance and space (additional instructions) overheads on \emph{all} memory accesses, it also introduces complexity in terms of state management.
+That is, \toolname{RCUdbg} would need
+
+%Not all code needs to be or should be instrumented in the same way, just as not all code needs to be instrumented. 
+%For code that is instrumented, having similar flexibility to selectivity is also useful: just
+%just as instrumented code is a ``heavyweight" version of native code, so can 
+%. In some cases, tools can distinguish between light and heavyweight instrumentation policies, just as instrumentation itself is a heavyweight version of native execution.
+% However, not all code that is instrumented should have to be instrumented in the same way.
+%For example, a security application could perform lightweight instrumentation of trusted kernel code, and more heavyweight instrumentation of untrusted modules.
+%we want to change how code is instrumented when: \begin{enumerate}
+%	\item {\bf Not all code needs to be instrumented in the same way.} This is an important concern when comprehensiveness isn't required, or when there are timing or performance constraints.
+For example,
+%	\item {\bf Different code should be instrumented differently.} For example, DBT-based sandboxing \cite{Vx32,ProgramShepherding} can be implemented by instrumenting trusted and untrusted code differently.
+In the case of the Linux, the kernel could be instrumented with a lightweight sandboxing policy for comprehensiveness, and untrusted modules could be instrumented with heavyweight policies for correctness.
+%	\item {\bf The same code should be instrumented differently depending on the context.}
+%\end{enumerate} 
+%This is unfortunate, especially when the instrumentation is heavyweight, because then all instrumented code is slowed down equally. Some bugs are sensitive to 
+%For analyses that must be run on production systems,
+
+%Granary gives tools the ability to be selective about what is instrumented.
 A natural extension of this selectivity is to allow tools to specialize how code is instrumented by dynamically switching the policy used to instrument code.
@@
 For example, some instrumentation is high-overhead, but does not need to be applied to all code.
-One example of this is Granary's \toolname{RCUdbg} tool, which instruments read-side and write-side critical sections more heavily than the rest of code.
+One example of this is Granary's \toolname{RCUdbg} tool, which instruments read-side critical sections more heavily than the rest of code.
 Unfortunately, existing DBT systems do not support runtime code specialization \cite{DRK,btkernel,Pin,DynamoRIO}.
 
-To implement the equivalent of \toolname{RCUdbg}'s runtime code specialization using an existing DBT system requires that a tool developer maintain, update, and check runtime state to know whether or not code is executing inside or outside of a read-side critical section.
+%To implement the equivalent of \toolname{RCUdbg}'s runtime code specialization using an existing DBT system requires that a tool developer maintain, update, and check runtime state to know whether or not code is executing inside or outside of a read-side critical section.
 If a basic block is executing in a read-side critical section, then instrumentation should be enabled, and if it's not in a read-side critical section then it should be disabled.
@@
 
-%Existing systems move the burden of code specialization to instrumentation tool developers.
-
-Fortunately for tool developers, Granary has built-in support for runtime code specialization that does not require manual state management or cause code cache bloat.
-Granary approach to runtime code specialization uses a technique called \emph{context-driven instrumentation}.
-
-%For example, instrumenting only module code allows Granary to specialize the execution of native code.
-In the same way, policy switching allows Granary's \toolname{RCUdbg} tool to specialize the execution of instrumented code executing inside of read-side and write-side critical sections.
-In fact, optimizing the performance of an early Granary tool was the original motivation for policy switching.
-
-\subsection{Context-Driven Instrumentation}\label{sec:policies}
+Fortunately for tool developers, Granary has built-in support for runtime code specialization. Granary's implementation of runtime code specialization is unique in that: \begin{enumerate}
+
+	\item {\bf It does not require manual state management.} Reentrant state management in the kernel environment is extremely complex.
+For example, using thread-private memory is a natural way to track the current execution context.
+However, there is no well-defined thread-private storage class in the Linux kernel.
+This is because interrupt handlers can, but don't always, share stacks with interrupted code, and interrupt handler code can swap out the current thread and swap in another thread.
+CPU-private storage is another option, but introduces an impedence mismatch on SMPs because contextual state is typically thread-specific, but a given thread can be rescheduled (e.g.
+by an interrupt) to execute on a different CPU.
+Finally, global state introduces potential performance bottlenecks (locking, false/true sharing) and sometimes tricky synchronization protocols.
+
+	\item {\bf It does require any runtime checking of the current runtime context.} If a tool like \toolname{RCUdbg} were implemented with DRK or btkernel, then \toolname{RCUdbg} would need to instrument code to continually check the current context and perform the correct memory access check given the current context.
+Checking the current context introduces performance overheads (additional instructions must be executed; branch mispredictions), and memory overheads (additional instructions are needed just to figure out what to check; icache misses).
+
+\end{enumerate}
+
+Granary's approach to runtime code specialization works by generating multiple versions of each basic block \emph{just-in-time}, where each version of a basic block is specialized to a different execution context.
+
+% For example, in the Linux kernel, interrupt handling code can execute on the interrupted stack.
+If Granary is instrumenting all kernel code, the a tool developer needs to decide if the instrumented interrupt handling code should inherit the runtime context from the interrupted thread.
+Making such a judgement is further complicated by the fact that interrupt arrival and semantics are hardware-defined, but thread scheduling is software defined.
+That is, if an interrupt handler inherits
+%This results in less code cache bloat and more focused and simple instrumentation tools. That is, tool developers can focus on what the tool should do, rather than how to fit their idea of the tool
+%	\item It works just-in-time, using basic block versioning, thus generating the minimum number of versio
+% , and does not cause performance overheads by continually checking .
+%Granary's approach to runtime code specialization 
+ %uses a technique called \emph{context-driven instrumentation}.
+
+\subsection{Specialization Using Basic Block Versions}\label{sec:policies}
 
@@
 
-%To implement this optimization,
-
-Granary enables specializing instrumentation to the context in which the instrumented code will execute.
+Granary tools can specialize instrumentation to the context in which the instrumented code will execute.
 Specialization is achieved by allowing multiple, differently instrumented versions of the same code to co-exist within Granary's code cache.
@@
 An instrumentation policy is both a name for an execution context, as well as a function that decides how to instrument basic blocks that will execute within that context.
-All tools define an initial policy that Granary uses to instrument module code.
-Tools are not limited to one policy though: any policy can declare a policy switch that will take effect when a selected control-transferring instruction (CTI) is executed.
-The effect of a tool specifying a policy switch on a CTI is that the basic block(s) targeted by that CTI will be instrumented according to the specified policy.
-CTIs with unspecified policies inherit their policies from their containing basic blocks.
-
-For example, our \toolname{RCUdbg} invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.
+All tools define an initial policy that Granary uses to instrument code.
+Tools are not limited to one policy though: any policy can declare a ``policy switch" that will take effect when a selected control-transferring instruction (CTI) is executed.
+Every policy function must declare what the ``next" policy to switch to will be.
+Policy switches can also be specified on a per-CTI basis.
+If a CTI has a specified policy, then any basic blocks targeted by that CTI will be instrumented by the specified policy.
+If a CTI does not have a specified policy, then it automatically inherits the next policy.
+
+%The effect of a tool specifying a policy switch on a CTI is that the basic block(s) targeted by that CTI will be instrumented according to the specified policy.
+A CTI with an unspecified policy inherits its policy from its containing basic block.
+That is, if no policy is specified on a CTI, then
+
+For example, \toolname{RCUdbg} defines three policies: $P_{read}$ for code executing within read-side critical sections, $P_{write}$ for code executing within write-side critical sections, and $P_{outside}$ for code executing outside of RCU read- and write-side critical sections.
+The following sequence of events describe some of the policy switching behavior of \toolname{RCUdbg} and how it interacts with block versioning.
+\begin{enumerate*}
+	\item Instrumentation begins with $P_{outside}$.
+	\item $P_{outside}$ instruments a basic block containing a \texttt{call} to the \texttt{memset} kernel function. The \texttt{call} inherits the next policy, which is $P_{outside}$.
+	\item When executed, the \texttt{call} to \texttt{memset} will cause \texttt{memset} to be instrumented by $P_{outside}$.
+No other instrumented versions of \texttt{memset} exist in Granary's code cache.
+Granary knows to use $P_{outside}$ and not another policy because $P_{outside}$ is stored in the edge data structure used by the first execution of the \texttt{call} (\Cref{sec:direct_control_flow}).
+	\item Later, $P_{outside}$ instruments a basic block containing a \texttt{call} to \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock}.
+$P_{outside}$ sets next policy to $P_{read}$ and uses $P_{read}$ to instrument the tail of the basic block following the \texttt{call}.
+	\item Execution naturally switches from being in one context (outside) into another (read) as execution returns from the \texttt{call} to \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock}. 
+	\item $P_{read}$ instruments a basic block containing a \texttt{call} to \texttt{memset}. The \texttt{call} inherits the next policy, which is $P_{read}$.
+	\item When executed, the \texttt{call} to \texttt{memset} within the read-side critical section will cause \texttt{memset} to be instrumented by $P_{read}$.
+There are now two versions of \texttt{memset} in the code cache: one specialized for executing outside of RCU-related code, and the other specialized to executing within read-side critical sections.
+\end{enumerate*}
+
+%For example, our \toolname{RCUdbg} invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.
 A similar switch occurs from $P_{\mathit{read\_critical}}$ back to $P_{\mathit{null}}$ at a \texttt{rcu\_read\_unlock}.
@@
 
-Because policies name an execution context, they also represent states in a finite state machine.
+The \toolname{RCUdbg} example highlights how the current execution context is implicity tracked by the current program counter or instruction pointer.
+Before the \texttt{call} to \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock}, the instructions were instrumented by $P_{outside}$.
+However, after the \texttt{call}, the instructions were instrumented by $P_{read}$.
+This insight reveals that policy switches and control-flow transfers behave like state transitions in a finite state machine, and that the program counter, by virtue of pointing at an instruction instrumented by a specific policy, tracks the current pogram state.
+This analogy, however, reveals a potential limitation: if RCU read-side critical sections are nested, then \toolname{RCUdbg}'s 3-state policy switching approach (as described) cannot distinguish a \texttt{call} to \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} from within an unnested and a nested read-side critical section.
+\toolname{RCUdbg} solves this problem by having more than one read policies: $P_{\mathit{{read}_0}}$, $P_{\mathit{{read}_1}}$, ..., $P_{\mathit{{read}_N}}$, which are generated on demand.
+In practice, this does not lead to code cache bloat because RCU read-side critical sections do not nest more than two times.
+
+By default, Granary doesn't allow tools to specify policy switches on procedure return instructions (\texttt{ret}).
+This increases the ``power" of the formal model of policy switching from being a deterministic finite state automaton to being a restricted form of pushdown automaton.
+The program counter continues to track the current execution state, and saved program counters (returns addresses stored on the runtime call stack) record prior execution states.
+In this model, \texttt{call}s place contextual breadcrumbs (in the form of return addresses) on the runtime call stack, and \texttt{ret}s read these breadcrumbs to return to a previous context.
+For many tools, this is a natural model of state tracking as it is bound to both the current thread.
+
+%\toolname{RCUdbg} solves this problem by tracking the nesting depth using different read-policies ($P_{read}_0$, $P_{read}_1$, ...).
+
+%Because policies name an execution context, they also represent states in a finite state machine.
 That is, code cache execution is in the state named by a policy if the  executing code was instrumented by that policy.
@@
 
-An alternative and sometimes more powerful model for policy switching is to prevent policy switching on function return instructions (e.g.
+%An alternative and sometimes more powerful model for policy switching is to prevent policy switching on function return instructions (e.g.
 either through manually specifying a policy switch, or by inheriting the policy from the basic block containing the \texttt{ret} instruction).
@@
 Granary decodes and instruments the targeted instructions according to the input policy information.
+
+\subsection{Partial Specialization Using Policy Properties}
+
+Granary internally specializes code at a finer granularity than that described by a tool's instrumentation policies.
+Granary will sometimes generate multiple versions of a given basic block, even for the same instrumentation policy.
+Other times, Granary will only generate extra entries in the code cache index that are eventually disambiguated.
+When apparently redundant code generation occurs, it means that certain properties that are internally tracked by Granary are present, and that these properties affect how Granary or its tools can optimize or instrument code.
+
+For example, Granary tracks whether code may or may not access user space data, because this code can potentially result in a page fault (\Cref{sec:interrupt_return_address_transparency}).
+Granary can detect this code using a combination of pattern recognition, and policy property propagation.
+This internal mechanism is needed for correctly handling page faults in the kernel, and is also useful for optimizing code generated by tools (e.g.
+\toolname{BoundsChecker}) using Granary's address watchpoints framework \cite{BehaveOrBeWatched}.
+
+Another example of partial specialization that benefits instrumentation tools is Granary's SIMD register tracking property.
+Granary detects and tracks the usage of SIMD registers (\texttt{\%xmm0} through \texttt{\%xmm16}).
+This informs instrumentation tools on when its safe to save transient data to the SIMD registers instead of to memory, which can result in improved performance \cite{Minemu}.
+
+%An access to user space data can potentially cause a fault in the kernel .
+Code that accesses user space data contains specify binary patterns; however, sometimes those patterns are not present in the individual basic blocks that dereference the user space address.
+
+%For example, Granary's address watchpoints framework uses address tainting to easily associate meta-data to allocated objects, and to make it easy to distinguish between tracked and untracked objects \cite{BehaveOrBeWatched}.
+The mechanism used to taint an address depends on x86-64 ``non-canonical" addresses.
+A canonical address has all of its 16 high-order bits set to 0 or set to 1.
+In Linux, the 16 high-order bits of kernel addresses are all 1, and the 16 high-order
+
+%These propeties, called policy properties, transparently persist across control-flow instructions, even if the CTI changes the current policy. Granary tracks the following policy properties:
+
+%\begin{description}
+%	\item[Temporary]
+%	\item {\bf Kernel vs. module code.}
+%	\item 
+%	\item {\bf SIMD register usage.}
+%\end{description}
 
@@
 %the associated memory may or may not be free, but the data structure occupying that memory can't be proven to exist. This 
-
 %  However, this class of RCU bugs is not well-suited to being detected by a use-after-free memory checker tool
@@
 RCU usage bugs, like bugs with other synchronization mechanisms, often stay hidden and only exhibit themselves for very specific thread interleavings/schedules.
-
 % go silently unnoticed because the memory 
-
 %While bugs related to mutual exlusion locks might result in easily observable deadlocks, bugs related to RCU typically result in use-after-free errors, which sometimes go silently unnoticed, sometimes result in kernel panics, and most of the time work as expected because the associated data is .

### 2013-12-13 VEE-14/paper.tex
Minor updates>

@@
 
-Granary implements policy tracking and switching by encoding policy information into the meta-data and CTIs of basic blocks.
+Granary implements policy tracking and switching by encoding policy information into the meta-data of basic blocks, the instructions of in-edge code generated for indirect control flow, and the edge data structures used by direct control flow.
+When an instrumented CTI executes for the first time, it yields control to Granary with the CTI target and policy information as inputs.
+Granary decodes and instruments the targeted instructions according to the input policy information.
+
+% CTIs of basic blocks.
 \Cref{fig:direct_edge_code} shows an example of how policy information is directly recorded into control-transfer edge code.
 If a policy switch is not specified on an instrumented CTI then that CTI inherits the policy used to instrument the basic block containing the CTI.
-When an instrumented CTI executes for the first time, it yields control to Granary with the CTI target and policy information as inputs.
-Granary decodes and instruments the targeted instructions according to the input policy information.
 

### 2013-12-13 VEE-14/paper.tex
Fixes and update of partial specialization subsection.

@@
 
-	\item {\bf It does require any runtime checking of the current runtime context.} If a tool like \toolname{RCUdbg} were implemented with DRK or btkernel, then \toolname{RCUdbg} would need to instrument code to continually check the current context and perform the correct memory access check given the current context.
+	\item {\bf It does not require any runtime checking of the current runtime context.} If a tool like \toolname{RCUdbg} were implemented with DRK or btkernel, then \toolname{RCUdbg} would need to instrument code to continually check the current context and perform the correct memory access check given the current context.
 Checking the current context introduces performance overheads (additional instructions must be executed; branch mispredictions), and memory overheads (additional instructions are needed just to figure out what to check; icache misses).
@@
 In this model, \texttt{call}s place contextual breadcrumbs (in the form of return addresses) on the runtime call stack, and \texttt{ret}s read these breadcrumbs to return to a previous context.
-For many tools, this is a natural model of state tracking as it is bound to both the current thread.
+For many tools, this is a natural model of state tracking as it is bound to the current thread, and because kernel code tends to follow a good call/return discipline.
 
@@
 
-For example, Granary tracks whether code may or may not access user space data, because this code can potentially result in a page fault (\Cref{sec:interrupt_return_address_transparency}).
+For example, Granary tracks whether code may or may not access user space data, because this code can potentially result in a page fault.
 Granary can detect this code using a combination of pattern recognition, and policy property propagation.
-This internal mechanism is needed for correctly handling page faults in the kernel, and is also useful for optimizing code generated by tools (e.g.
-\toolname{BoundsChecker}) using Granary's address watchpoints framework \cite{BehaveOrBeWatched}.
+This internal mechanism is needed to correctly handle page faults in instrumented kernel code (\Cref{sec:interrupt_return_address_transparency}), and is also useful for optimizing code generated by tools as well.
+For example, Granary's \toolname{BoundsChecker} tool alters the 16 high order bits of addresses returned from kernel allocators, to make it easy to associate bounds information with each allocated object \cite{BehaveOrBeWatched}.
+Checked and unchecked kernel addresses can distinguished by testing only a single bit.
+However, user space addresses and checked addresses are harder to distinguish, requiring two bits to be checked.
+Fortunately,  few instructions dereference user space addresses, and Granary tracks whichs basic blocks might contain those instructions, thus allowing the \toolname{BoundsChecker} to avoid many unnecessary bit tests.
 

### 2013-12-17 VEE-14/paper.tex
Working on lmbench evaluation charts.

@@
 
-\subsection{Optimizations}
+\subsection{Optimizations}\label{sec:optimize}
 
@@
 
-What has been evaluated:
-
-\paragraph{lmbench}
-\begin{itemize}
-	\item Btkernel
-	\item Granary, full-kernel, null
-	\item Granary, full-kernel, even/odd
-	\item Granary, full-kernel, even/odd, where even/odd is a runtime check and branch.
-\end{itemize}
-
-\paragraph{fileserver}
-Setup: 1GB RAM, 3GB partition, ext3+jbd, >1GB mounted file system.
-\begin{itemize}
-	\item full-kernel
-	\begin{itemize}
-		\item Btkernel
-		\item Granary, null
-		\item Granary, even/odd
-		\item Granary, even/odd, where even/odd is a runtime check and branch.
-	\end{itemize}
-	\item module-only, (ext3+jbd)
-	\begin{itemize}
-		\item Granary, null
-		\item Granary, even/odd
-		\item Granary, even/odd, where even/odd is a runtime check and branch.
-	\end{itemize}
-\end{itemize}
-
-\paragraph{kernel compile}
-Same as fileserver, but with 8GB RAM instead of 3GB.
-
-\paragraph{recursive copy of kernel source tree}
-Same as fileserver, but with 8GB RAM instead of 3GB.
+\begin{figure*}
+\centering\epsfig{file=charts/lmbench_legend.eps,height=6em} \\
+\subfloat[Fast operations]{\epsfig{file=charts/lmbench_fast.eps,height=18em}}
+\subfloat[Fork operations]{\epsfig{file=charts/lmbench_fork.eps,height=18em}}
+\caption{Microbenchmark results from \texttt{lmbench} that shows the effect of Granary's code cache optimizations (\Cref{sec:optimize}).}
+\end{figure*}
 

### 2013-12-18 VEE-14/paper.tex
Updated the charts and the CFG section.

@@
 \section{Application: Control-Flow Graphs}\label{sec:cfg}
+Granary's \toolname{cfg} tool constructs inter- and intra-procedural control flow graphs from the instrumented kernel execution.
+The \toolname{cfg} tool can also act as a code coverage and profiling tool by recording basic block and conditional branch execution counts.
+\Cref{fig:lmbench} evaluates the performance impact of using the \toolname{cfg} tool for profile-guided optimization.
+First, \toolname{cfg} instruments a single run of the \texttt{lmbench} microbenchmark suite.
+Then, the recorded profile data is used to guide the ahead-of-time tracing infrastructure (\Cref{sec:aot}) on the evaluated runs of \texttt{lmbench}.
+
+\toolname{cfg}'s instrumentation introduces counter increments into basic blocks, counter increments following conditional branches, and target recorders for indirect CTIs (\texttt{call}, \texttt{jmp}).
+Each indirect CTI recorder maintains a variable-sized array of all observed targets for the given CTI.
+The recording process follows either a fast or a slow path.
+In the fast path, the target is compared with a cached version of the most recent target.
+If they are the same then no further action is taken, as this implies that the target has already been recorded.
+Most executions ``hit" in this cache because most indirect branches in the kernel only have a single target.
+If the target and the cached target differ, then the cached target is immediately updated and the slow path is executed.
+The slow path checks to see if the target has been permanently recorded, and if not, records it.
+
+Control-flow graphs are reported in terms of basic blocks (nodes) and edges (control flow transfers).
+The reporting operates by iterating over each basic block and performing the following for each basic block: \begin{enumerate*}
+	\item Decode the native instructions associated with the basic block. The location and number of instructions to decode is found in the basic block's meta-data (\Cref{sec:metadata}).
+	\item For each decoded direct CTI, print out the target PC of the CTI.
+If the CTI is a conditional CTI, then print out two PCs: the target PC, and the fall-through PC.
+If conditional branch counting is enabled, then the value of the counter associated with the fall-through edge is also printed.
+	\item For each decoded indirect CTI, print out all target PCs from the associated indirect CTI recorder.
+\end{enumerate*}
+
+The reported data can be re-integrated into Granary to drive profile-guided optimizations, or post-processed into a format for use by tools like KCacheGrind for profile data visualization.
+
+%A single-element cache is also maintained that records the most recent indirect branch target.
+Most indirect CTIs in the kernel only have one target, and so once the target has been permanently recorded, all future executions hit the cache.
+
+%As an optimization, the indirect target recorders maintain a one-element cache of the most recent target of a given indirect branch
+
+% Basic blocks in the graph can be labeled with execution counts, and control-flow edges between basic blocks can also be labeled with execution counts.
+Counting basic block and edge executions enables Granary's \toolname{cfg} tool implicitly provides code coverage and profiling information.
+\Cref{fig:lmbench} evaluates how profiling information from Granary's \toolname{cfg} tool is integrated with Granary's ahead-of-time translation optimization (\Cref{sec:aot}).
 
@@
 
-\begin{figure*}
-\centering\epsfig{file=charts/lmbench_legend.eps,height=6em} \\
-\subfloat[Fast operations]{\epsfig{file=charts/lmbench_fast.eps,height=18em}}
-\subfloat[Fork operations]{\epsfig{file=charts/lmbench_fork.eps,height=18em}}
-\caption{Microbenchmark results from \texttt{lmbench} that shows the effect of Granary's code cache optimizations (\Cref{sec:optimize}).}
+\begin{figure*}[ht!]
+\includegraphics[height=14em,clip=true,trim=0 -5em 10em -5em]{charts/lmbench_legend.eps}
+\subfloat[Fast operations]{\epsfig{file=charts/lmbench_fast.eps,height=14em}}
+\subfloat[Fork operations]{\epsfig{file=charts/lmbench_fork.eps,height=14em}}
+\caption{\label{fig:lmbench}Microbenchmark results from \texttt{lmbench} that shows the effect of Granary's code cache optimizations (\Cref{sec:optimize}).}
 \end{figure*}
+
+Before examining applications that are enabled by Granary, \Cref{sec:microbench} evaluates the overhead of the baseline instrumentation.
+Then, \Cref{sec:macrobenchmark} does something with the evaluation TODO.
+
+\subsection{Microbenchmark}\label{sec:microbench}
+For these experiments, we operate with a large enough code cache (\texttildelow 4MB) so that flushes do not occur.
+We perform a number of stress tests with a high level of kernel activity, to show the robustness of Granary and to illustrate worst-case behavior of the baseline instrumentation system.
+These tests ran on a Lenovo X230 laptop with 8GB of RAM and four Intel\textregistered\ Core\texttrademark\ i5 CPUs operating at 2.50GHz.
+
+\Cref{fig:lmbench} shows the results of running the \texttt{lmbench} microbenchmark, which compares the overhead of Granary's baseline instrumentation (with and without optimizations) to native execution.
+The results highlight that the biggest performance boost is achieved with aggressive ahead-of-time tracing (\Cref{sec:aot}), and that profiling information from Granary's \toolname{cfg} tool can improve on these results by guiding which branches the tracer follows.
+
+\subsection{Macrobenchmark}\label{sec:macrobenchmark}
+
+% when profiling data from Granary's \toolname{cfg} tool are used to predict and potentially exchange the targets of conditional branches.
+
+% We evaluated behavioral watchpoints using a microbenchmark and measured the performance of common filesystem operations on an in-memory disk.
+Our tests ran on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU with 4GB physical memory.
+%Watchpoint instrumentation was enabled for every memory operation.
+We measured the overhead of the watchpoint framework using a null client and heap-based buffer overflow detector.
+%We also evaluated the performance of heap-based buffer overflow detector developed using watchpoints.
 

### 2013-12-18 VEE-14/paper.tex
Working on the RCU section.

@@
 
-The reported data can be re-integrated into Granary to drive profile-guided optimizations, or post-processed into a format for use by tools like KCacheGrind for profile data visualization.
+The reported data can be re-integrated into Granary to drive profile-guided optimizations, or post-processed into a format for use by tools like KCacheGrind for profile data visualization \cite{KCacheGrind}.
 
@@
 
-Granary's \toolname{RCUdbg} tool is designed to catch RCU bugs related to kernel code misusing pointers to RCU-protected data structures.
-The following is a non-exhaustive list of requirements that must be met by reader and writer threads operating on an RCU-protected data structure.
-The list omits a description of how RCU-protected data structures are garbage collected, which is beyond the scope of the bugs caught by \toolname{RCUdbg}.
+Granary's \toolname{RCUdbg} tool is designed to catch bugs related to misuses of pointers to RCU-protected data structures.
+These pointers, called RCU-protected pointers, link together discrete objects within larger RCU-protected data structure.
+For example, in an RCU-protected linked list, the RCU-protected pointers are the \texttt{next} pointers that connect list elements.
+An RCU-protected pointer is like a normal pointer, but points to a potentially incomplete or partially-constructed data structure.
+Data structures can be potentially incomplete or partially constructed because RCU permits concurrent reads and writes to the shared data structure.
+Because of this, reads can sometimes see an inconsistent version of the data structure.
+For example, if a writer modifies the \texttt{next} pointer of an RCU-protected linked list, then a concurrently executing reader may see either the old value or the new value of the \texttt{next} pointer.
+This is further complicated by multi-step updates (e.g.
+adding or removing an element from a doubly-linked list).
+%For example, adding an element to a doubly-linked list requires two separate pointers to be updated.
+Concurrently executing readers can therefore observe the list in one of several intermediate states.
+
+%
+%\toolname{RCUdbg} infers that some pointers are RCU-protected pointers based on how specific RCU API functions are used by the instrumented code.
+When certain API functions are invoked, \toolname{RCUdbg} knows that some of the function's arguments or return values are RCU-protected pointers, and therefore treats them as such.
+For example, the RCU API includes two functions for operating on RCU-protected pointers: \begin{inparaenum}[i)]
+%	\item \texttt{rcu\_\linebreak[0]dereference} for reading the value of an RCU-protected pointer, and;
+%	\item \texttt{rcu\_\linebreak[0]assign\_\linebreak[0]pointer} for assigning a value to an RCU-protected pointer. 
+%\end{inparaenum}
+%
+The following requirements describe how code can use the RCU API to read from and modify a data structure like an RCU-protected linked list.
+
+% with eachother and operate on RCU-protected pointers.
+
+%There are two kinds of RCU-protected pointers: read-only pointers, and read/write pointers.
+Read-only pointers can only be read from within RCU read-side critical sections, delimited by  \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock}.
+Read/write pointers can be written to
+
+%Code operating on RCU-protected pointers must use specific API functions: , and By assuming that these functions are correctly used some of the time, we can
+
+%Therefore, a use of \texttt{rcu\_\linebreak[0]dereference} implies that the pointer being read should be RCU-protected, and the the value returned should also be RCU-protected.
+Without an \emph{a priori} understanding of the data structures being operated on, these two API functions give \toolname{RCUdbg} enough information about RCU protected pointers to find bugs.
+
+%Without a prior understanding of what pointers are RCU-protected pointers, we can assume that any pointer operated on by one the aforementioned API functions is an RCU-protected pointer.
+
+%Without an \emph{a priori} understanding of the structure of the data being syncrhonized using RCU, we can conservatively assert that RCU-protected pointers.
 
@@
 \begin{enumerate}
-	\item Readers must tolerate an inconsistent view of the shared data structure.
+%	\item Readers must tolerate an inconsistent view of the shared data structure.
 That is, a reader operating concurrently with a writer might observe either the version of the data before or after the writer's changes have been published.
@@
 	\item Within read-side critical sections, readers can only read a shared data structure after obtaining a pointer to that shared data structure using the \texttt{rcu\_dereference} API function.
-An RCU-protected pointer is like a normal pointer, but points to a potentially incomplete or partially-constructed data structure.
-\texttt{rcu\_dereference} operates on an RCU-protected pointer and returns an ``unprotected" version of the pointer.
-The reader can use the unprotected pointer to safely access the shared data structure within the remainder of the its read-side critical section.
+This API function operates on an RCU-protected pointer and returns a read-access-only version of the pointer.
+The reader can use the read-access pointer to safely read from the shared data structure within the remainder of its read-side critical section.
 If \texttt{rcu\_dereference} is not used, then the reader might observe a partially-constructed (i.e.
@@
 Notably, accessing the original data structure and not the dereferenced data structure pointer--even after an \texttt{rcu\_dereference}--is undefined behavior.
-That is, within a read-side critical section, RCU only guarantees the ``completeness" of data structures pointed to by RCU-dereferenced pointers.
+That is, within a read-side critical section, RCU only guarantees the ``completeness" of data structures pointed to by the read-access-only pointers obtained from  \texttt{rcu\_dereference}.
 Data accessed within a read-side critical section is \emph{not} guaranteed to exist outside the read-side critical section.
@@
 
-	\item RCU only defines how writers can update RCU-protected pointer within a shared data structure.
+%	\item RCU defines how writers can update RCU-protected pointers within a shared data structure.
 Modification of non-RCU-protected pointer fields within an RCU-protected data structure is beyond the scope of RCU's correctness guarantees.
@@
 \begin{figure}[ht!] %
-\subfloat[\label{fig:rcu_existence_bug}Despite correctly obtaining a pointer to a shared data structure using \texttt{rcu\_dereference} within a read-side critical section, properties (2) and (3) tell us that the data pointed to by \texttt{elm} is not guaranteed to exist after the read-side critical section ends.]{\fbox{\usebox\rcuexistencebug}} %
-
-\subfloat[\label{fig:rcu_deref_bug}Despite the call to \texttt{rcu\_dereference}, properties (1) and (3) tell us that \texttt{list\_head} is not guaranteed to be equivalent to \texttt{elm} because a writer might have concurrently changed the value of the \texttt{list\_head} pointer.
+\subfloat[\label{fig:rcu_existence_bug}Despite correctly obtaining a pointer to a shared data structure using \texttt{rcu\_dereference} within a read-side critical section, the data pointed to by \texttt{elm} is not guaranteed to exist after the read-side critical section ends.]{\fbox{\usebox\rcuexistencebug}} %
+
+\subfloat[\label{fig:rcu_deref_bug}Despite the call to \texttt{rcu\_dereference}, \texttt{list\_head} is not guaranteed to be equivalent to \texttt{elm} because a writer might have concurrently changed the value of the \texttt{list\_head} pointer.
 That is, if \texttt{elm} is non-\texttt{NULL}, then the data pointed to by \texttt{elm} is guaranteed to exist for the remainder of the read-side critical section.
@@
 
-\subfloat[\label{fig:rcu_assign_bug}This example shows the code for a writer thread adding a new element into the second position of an RCU-protected list.
+\subfloat[\label{fig:rcu_assign_bug}This example shows the code for a writer thread adding a new element into the second position of an RCU-protected linked list.
 The code for reader threads is also shown, where readers iterate over the list by traversing \texttt{next} pointers and sum up the \texttt{data} fields of each visited list element.
-The access of \texttt{elm->next} violates of properties (3) and (7).
-That is, the next element of the list is only guaranteed to be fully constructed within the read-side critical section if the reader executes \texttt{rcu\_dereference(elm->next)}.
+Unfortunately, the access of \texttt{elm->next} results in undefined behavior.
+This is because the next element of the list is only guaranteed to be fully constructed within the read-side critical section if the reader executes \texttt{rcu\_dereference(elm->next)}.
 Therefore, the first read of \texttt{elm->next} might yield a pointer to partially constructed data.
@@
 
-\toolname{RCUdbg} is primarily concerned with finding bugs related to RCU-rpotected pointer misuses by reader threads (both inside and outside of read-side critical sections).
+\toolname{RCUdbg} is primarily concerned with finding bugs related to RCU-protected pointer misuses by reader threads (both inside and outside of read-side critical sections).
 As mentioned above, RCU bugs sometimes manifest as use-after-free bugs.
-Unfortunately, a pure use-after-free-based checking approach is insufficient (\Cref{fig:rcu_assign_bug} is one such example).
-\Cref{fig:rcu_existence_bug} is an example of an RCU usage bug that can manifest as a use-after-free bug if one is unlucky enough to observe a very specific thread schedule.
-Suppose that execution is interrupted after \texttt{rcu\_read\_unlock}, but before the access to \texttt{elm->data}.
-If \texttt{elm} is garbage collected and freed before the code resumes its execution, then the access to \texttt{elm->data} is a use-after-free bug.
+Unfortunately, a pure use-after-free-based checking approach is insufficient.
+\Cref{fig:rcu_assign_bug} is an example where a use-after-free based checking approach will miss an RCU-protected pointer misuse because the bug does not manifest as a use-after-free.
+\Cref{fig:rcu_existence_bug} is an example of an RCU usage bug that can manifest as a use-after-free bug, but only if one is unlucky enough to observe a very specific thread schedule.
+That is, suppose that execution is interrupted after \texttt{rcu\_read\_unlock}, but before the access to \texttt{elm->data}.
+If \texttt{elm} is freed before the interrupted code resumes its execution, then the access to \texttt{elm->data} is a use-after-free bug.
 This kind of bug is a time bomb waiting to go off.
-Fortunately, \toolname{RCUdbg} catches the bug in \Cref{fig:rcu_existence_bug} right when \texttt{elm->data} is accessed.
+Fortunately, \toolname{RCUdbg} catches this bug and others in a scheduling-independent way.
 
@@
 
-\begin{figure}[t!]
-\lstset{language=C, tabsize=2, stepnumber=1}
-\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-// As used in the code:
-//   p = rcu_dereference(q);
-// Internally annotated to become:
-//   p = __rcudbg_deref(&q, q);
-// Wrapper:
-FUNCTION_WRAPPER(__rcudbg_deref, (q_addr, q), {
-  if(is_watched_address(q_addr)) {
-    // Make sure q_addr
-  } else {
-
-  }
-  return add_watchoint(q, DEREF);
-})
-\end{lstlisting}
-\caption{}
-\label{fig:rcu_dereference_wrapper}
-\end{figure}
+%\begin{figure}[t!]
+%\lstset{language=C, tabsize=2, stepnumber=1}
+%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
+%// As used in the code:
+%//   p = rcu_dereference(q);
+%// Internally annotated to become:
+%//   p = __rcudbg_deref(&q, q);
+%// Wrapper:
+%FUNCTION_WRAPPER(__rcudbg_deref, (q_addr, q), {
+%  if(is_watched_address(q_addr)) {
+%    // Make sure q_addr
+%  } else {
+%
+%  }
+%  return add_watchoint(q, DEREF);
+%})
+%\end{lstlisting}
+%\caption{}
+%\label{fig:rcu_dereference_wrapper}
+%\end{figure}
 
@@
 
-\section{Evaluation}\label{sec:eval}
-
 \begin{figure*}[ht!]
@@
 
+\section{Evaluation}\label{sec:eval}
+
 Before examining applications that are enabled by Granary, \Cref{sec:microbench} evaluates the overhead of the baseline instrumentation.

### 2013-12-18 VEE-14/paper.tex
Minor updates to the optimizations section.

@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{Fast and Flexible Binary Instrumentation for the Kernel}
+\title{Fast and Flexible Dynamic Binary Translation for the Kernel}
 
@@
 That is, execution is more likely to fall through a conditional CTI than it is to take the conditional CTI.
-We experimentally verified Granary's tracing strategy by disabling tracing and relying on Granary's internal performance counters to count how many conditional branches were patched (by their edge code, \texttildelow 25\%), versus how many synthesized fall-through \texttt{jmp}s were patched (\texttildelow 80\%).
+We experimentally verified Granary's tracing strategy by disabling tracing and relying on Granary's internal performance counters to count how many conditional branches were patched (by their edge code, \texttildelow 35\%), versus how many synthesized fall-through \texttt{jmp}s were patched (\texttildelow 80\%).
+
+Profile-guided ahead-of-time tracing further improves on Granary's otherwise na\"{i}ve tracing strategy by integrating profile information from Granary's \toolname{cfg} tool (\Cref{sec:cfg}).
 
@@
 For example, Granary aligns patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86-64 processors \cite{AgnerMicroarchitecture}.
+When possible, alignment requirements are reduced by allowing a patchable instruction to straddle a cache line, but for the part to be patched to be fully contained by a single cache line.
 Finally, most basic blocks are small (i.e.
@@
 \begin{itemize}[leftmargin=3.2em]
-	\item[B1)] Reads through an RCU-dereferenced pointer outside of a read-side critical section (\Cref{fig:rcu_existence_bug}).
-	\item[B2)] Reads through an RCU-dereferenced pointer in the wrong read-side critical section.
-
-	\item[B2)] Reads through an RCU-assigned pointer that don't first use \texttt{rcu\_dereference} (\Cref{fig:rcu_deref_bug}).
-	\item[B3)] Reads through fields of RCU-protected data structures that should be accessed via \texttt{rcu\_dereference} (\Cref{fig:rcu_assign_bug}).
-
-	\item[B5)] Writes through an RCU-dereferenced pointer.
-	\item[B6)] Writes through an RCU-assigned pointer by a non-write thread.
+	\item[\bf B1] Reads through an RCU-dereferenced pointer outside of a read-side critical section (\Cref{fig:rcu_existence_bug}).
+	\item[\bf B2] Reads through an RCU-dereferenced pointer in the wrong read-side critical section.
+
+	\item[\bf B2] Reads through an RCU-assigned pointer that don't first use \texttt{rcu\_dereference} (\Cref{fig:rcu_deref_bug}).
+	\item[\bf B3] Reads through fields of RCU-protected data structures that should be accessed via \texttt{rcu\_dereference} (\Cref{fig:rcu_assign_bug}).
+
+	\item[\bf B5] Writes through an RCU-dereferenced pointer.
+	\item[\bf B6] Writes through an RCU-assigned pointer by a non-write thread.
 \end{itemize}

### 2013-12-18 VEE-14/paper.tex
Added some comments about the trace allocator to the evaluation section.

@@
 
-\begin{figure*}[ht!]
+\begin{figure*}
 \includegraphics[height=14em,clip=true,trim=0 -5em 10em -5em]{charts/lmbench_legend.eps}
@@
 The results highlight that the biggest performance boost is achieved with aggressive ahead-of-time tracing (\Cref{sec:aot}), and that profiling information from Granary's \toolname{cfg} tool can improve on these results by guiding which branches the tracer follows.
+
+The trace allocation optimization (\Cref{sec:trace_alloc}) does not result in significant performance improvements, and sometimes results in worse performance.
+This is because Granary's default allocation strategy (allocating space for instructions from per-CPU memory pools) already results in similar memory layouts to the trace allocator.
+A more significant difference would  be observed if Granary defaulted to a global allocation mechanism instead.
+Also, Granary's inclusion of function call instructions (\texttt{call}) into basic blocks likely works against the trace allocator.
+If basic blocks ended at \texttt{call} instructions, then the trace allocation strategy would arrange for the returning basic block and the returned-to basic blocks to be adjacent in memory.
+
+%This is because the memory layout of basic blocks with and without the trace allocation optimization ends up being similar in Granary.
 

### 2013-12-19 VEE-14/paper.tex
Some minor updates to the RCU section.

@@
 
-\toolname{RCUdbg} gains visibility on RCU operations in two ways: \begin{enumerate}
+\toolname{RCUdbg} gains visibility on RCU operations in three ways: \begin{enumerate}
 	\item The Linux kernel RCU implementation is annotated to provide more information to \toolname{RCUdbg}.
@@
 These annotations are strictly internal to the RCU API, and thus require no changes to code using this API.
-	\item The annotated RCU API functions are specialized using Granary's function wrapping feature (\Cref{sec:function_wrapper}). Specifically, 
+	\item The annotated RCU API functions are specialized using Granary's function wrapping feature (\Cref{sec:function_wrapper}).
+Interposing on these functions allows \toolname{RCUdbg} to add an address watchpoint to the address returned by \texttt{rcu\_\linebreak[0]dereference}, and to correlate uses of \texttt{rcu\_\linebreak[0]dereference} and \texttt{rcu\_\linebreak[0]assign\_\linebreak[0]pointer}.
+	\item Read-side critical sections are discovered at instrumentation time (rather than run time) by observing \texttt{call}s to \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock}, respectively.
+This allows \toolname{RCUdbg} to instrument code running inside of read-side critical sections using one policy ($P_{read}$), and code running outside of read-side critical sections using another policy ($P_{outside}$).
 \end{enumerate}
@@
 
+Perhaps I should do one...
+
 % when profiling data from Granary's \toolname{cfg} tool are used to predict and potentially exchange the targets of conditional branches.

### 2013-12-19 VEE-14/paper.tex
Added normalized charts for lmbench.

@@
 This allows \toolname{RCUdbg} to instrument code running inside of read-side critical sections using one policy ($P_{read}$), and code running outside of read-side critical sections using another policy ($P_{outside}$).
+\toolname{RCUdbg} switches between the two policies on entry to and exit from a read-side critical section.
 \end{enumerate}
@@
 \subfloat[Fast operations]{\epsfig{file=charts/lmbench_fast.eps,height=14em}}
-\subfloat[Fork operations]{\epsfig{file=charts/lmbench_fork.eps,height=14em}}
+\subfloat[Fork operations]{\epsfig{file=charts/lmbench_fork.eps,height=14em}}\\
+\subfloat[Fast operations (Normalized)]{\epsfig{file=charts/lmbench_fast_norm.eps,height=14em}}
+\subfloat[Fork operations (Normalized)]{\epsfig{file=charts/lmbench_fork_norm.eps,height=14em}}
 \caption{\label{fig:lmbench}Microbenchmark results from \texttt{lmbench} that shows the effect of Granary's code cache optimizations (\Cref{sec:optimize}).}
@@
 \section{Conclusion}\label{sec:conclusion}
+In summary, we conclude.
+
 %We created Granary to address the challenges of instrumenting binary kernel modules.

### 2014-11-04 HotOS'15/paper.tex
Initial commit of the paper for Ashvin's course project.

%--------sigalternate----------
\documentclass{sig-alternate}
%--------/sigalternate----------

%\documentclass[letterpaper,twocolumn,10pt]{article}
%\usepackage{endnotes,multirow}
\usepackage{refstyle,amsmath,chngcntr}
\usepackage{epsfig,subfigure,framed}
\usepackage{textcomp} % for tilde
\usepackage{paralist} % for in-paragraph lists
\usepackage{graphicx}

% Used for code snippets
\usepackage{listings,courier}

% Make text in figure captions small
\usepackage{caption3} % load caption package kernel first
%\DeclareCaptionOption{parskip}[]{} % disable "parskip" caption option
\usepackage{caption}

% Make sure that figure numbers are continuous through the document
\counterwithout{figure}{section}
\counterwithout{figure}{subsection}

% Settings on code listings.
\lstset{language=C,
		xleftmargin=0pt,
		xrightmargin=0pt,
		framexbottommargin=0pt,
        framextopmargin=0pt,
        framesep=0pt}

% Modify spacing before/after title, sections, and subsections.
%\setlength{\droptitle}{-3em}  %--------sigalternate----------
%\posttitle{\par\end{center}\vspace{-1.5em}} %--------sigalternate----------
\newcommand{\Sref}[1]{Section~\ref{#1}}

%--------sigalternate----------\titlespacing\section{0pt}{5pt plus 4pt minus 2pt}{5pt plus 2pt minus 2pt}        
%--------sigalternate----------\titlespacing\subsection{0pt}{5pt plus 4pt minus 2pt}{5pt plus 2pt minus 2pt}

% Add a horizontal rule above captions, and make captions closer to their figures
\DeclareCaptionFormat{ruled_caption}{#1#2#3\hrulefill}
\captionsetup[figure]{format=ruled_caption}

% Prevent footnotes from spanning multiple pages
\interfootnotelinepenalty=10000

% For leaving some comments in the draft.
\newcommand{\comment}[1]{}
\newcommand{\Toolname}{DataReactor}

% For getting rid of copyright box
\makeatletter
\def\@copyrightspace{\relax}
\makeatother

\begin{document}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf \Toolname: Precise Detection of False Sharing in COTS Binaries}
\numberofauthors{3}
\author{
% 1st. author
\alignauthor
Peter Goodman\\
       \email{pag@cs.toronto.edu}
\and
% 2nd. author
\alignauthor
Ashvin Goel\\
       \email{asvhin@eecg.toronto.edu}
% 3rd. author
\alignauthor
Angela Demke Brown\\
       \email{demke@cs.toronto.edu}
}
\maketitle
\begin{abstract}

Hello
\end{abstract}

\section{Introduction}\label{sec:intro}

In this work, we present \Toolname, a system for precisely detecting the sources of false sharing in
multi-threaded, commercial off-the-shelf (COTS) binaries. This is important because false sharing
can limit the scalability of multi-threaded applications on cache-coherent multiprocessors
\cite{ImpactOfFalseSharing}. The key insight of our approach is that it begins by focusing on what
data is being accessed, and then uses that information to progressively narrow down on what code
is participating in a data race.

% TODO: Define precise
% TODO: Define false-sharing

\section{Related Work}\label{sec:background}
\Toolname{} draws on techniques from existing race detectors and false sharing detectors. This
is appropriate because some forms of false-positives for race detectors are caused by false
sharing. That is, where race detectors focus on two threads concurrently accessing the same
data, false sharing detectors focus on two threads concurrently accessing different data, but
on the same cache line.

\subsection{Race Detection}

Race detectors can be categorized according to the following parameters:
\begin{description}
\item[Approach] Static or dynamic. \hfill \\
Static race detectors analyze source code or binaries, typically performing interprocedural control-
and data-flow analyses. The benefit of static analysis is that they can inspect all program paths.

Dynamic race detectors monitor memory accesses for races. The benefit of this approach is that
races can be caught "in the act". The drawbacks of this approach is that most dynamic race detectors
are too slow to run in production environments, and therefore must be tested on simulated workloads.
If these workloads do not exercise all code paths than many races can be missed. Dynamic race detectors
can be further broken down into on-the-fly detectors or post-mortem detectors, which analyze memory
access logs.

\item[Model] Precise or imprecise.  \hfill \\
Precise race detectors report no false-positives and are usually implemented using Lamport's
happens-before relationship \cite{VectorClocks}. Precise race detectors often trade efficiency
for completeness in order to achieve their precision. That is, they sometime miss true races in
code.

Imprecise race detectors can have high false positive rates, but tend to be more complete in terms of
including races not detected by precise detectors. They tend to be implemented using a variant of the
lockset algorithm \cite{Eraser}.

\item[Granularity] Fine-grained or coarse-grained. \hfill \\
Fine-grained race detectors operate at or close to the granularity of memory accesses. Coarse-grained
race detectors operate at an object granularity, or at a fixed granularity (e.g. cache line, page).

The granularity of a race detector can affect its precision. For example, a happens-before race detector 
can be imprecise (i.e. report false positives) by operating at a coarse granularity. If the granularity were
the size of a cache line, then the race detector would be a cache line contention detector: it would hint
at races as well as false-sharing.

\item[Mode] Comprehensive or sampling-based. \hfill \\
Comprehensive race detectors typically analyze every memory read and write. They are usually implemented
by instrumenting all program code.

Sampling-based race detectors analyze either a subset of the code or a subset of the data. In the former
case, they focus on detecting races by specific code.
\end{description}

\Toolname{} is a dynamic, precise, variable-granularity, mixed-mode false-sharing detector. It draws inspiration
from a number of existing race detectors, including LiteRace \cite{LiteRace}, MultiRace \cite{MultiRace},
RaceMob \cite{RaceMob}, RaceTrack \cite{RaceTrack}, and DataCollider \cite{DataCollider}.

LiteRace \cite{LiteRace} is a precise, sampling-based dynamic race detector implemented on top of the Microsoft
Pheonix binary translator. It samples at a function granularity, and executes either one specialized version of the
code or another. The sampling rate for a function is inversely proportional to how often it executes. This is based
on the insight that races are likely to occur in infrequently executed code, because otherwise they are benign or
would have been noticed and resolved. This insight, called the ``cold code hypothesis", motivated \Toolname's
post-mortem ranking scheme: cases of false sharing are more highly ranked when one---but \emph{not} both---of
the participating threads are executing cold code. This variant on the cold code hypothesis emphasizes the similarly
hard-to-find case of cold code pessimizing hot code.

MultiRace \cite{MultiRace} is a comprehensive, dynamic race detector that employs a hybrid happens-before and
lockset-based approach. It operates at an object granularity, and therefore reports false-positives. It is unique
in that it uses ``pointer swizzing" as a means of remapping memory to ``minipages" that track the dominating
access pattern of a thread (no-acces, read-only, read-write). \Toolname{} uses a similar pointer swizzing approach, but
in the form of address watchpoints \cite{AddressWatchpoints}. Pointers to heap-allocated objects are ``swizzled" to
include a type identifier that is unique to a $\langle PC,size \rangle$ pair. This improves the diagnosis of false-sharing,
because it allows \Toolname{} to distinguish between two threads accessing different fields of the same object, and
two threads accessing two nearby objects of the same (or different) types.

RaceMob \cite{RaceMob} and RaceTrack \cite{RaceTrack} are two hybrid happens-before and lockset-based race detectors.
RaceMob begins by statically analyzing source code to find potential races (lockset), then dynamically verifies those races
(happens-before). RaceTrack is a JIT-based race-detector for Microsoft's common language runtime (CLR). It uses the
concept of threadsets to efficiently maintain happens-before relationships for memory locations and detect the possible
presence of a race and uses lockets to confirm predicted races. The novelty of RaceTrack is that it adapts the detection
granularity from object-granularity to field-granularity. \Toolname{} follows the same high-level approach as both RaceMob
and RaceTrack in that it attempts to predict the source of false-sharing at one granularity, then verify those predictions 
at another granularity.

DataCollider \cite{DataCollider} is a dynamic, precise, sampling-based race detector for the kernel. The key strengths of
DataCollider is that it has low overhead, it is independent of the synchronization mechanisms being employed, and that
it often catches both threads in the act. DataCollider uniformly samples code by introducing breakpoints into the code.
When a breakpoint is hit, a hardware watchpoint is added to data that was about to be accessed by the interrupted instruction.
Then, the interrupted thread is paused for a short period of time to give another thread the opportunity to access the
same data, thus triggering the watchpoint and detecting a race. This brilliantly simple approach was the main inspiration
for \Toolname{}. \Toolname{} differs from DataCollider, in that the latter is code-centric insofar as it samples code, whereas
\Toolname{} is data-centric insofar as it samples data.

\subsection{False Sharing Detection}

\section{Design}\label{sec:design}

\section{Status}\label{sec:status}

\section{Evaluation}\label{sec:evaluation}

\bibliographystyle{abbrv}
\bibliography{library}
\end{document}

### 2014-11-04 HotOS'15/paper.tex
Second commit. Done most of the related work, although I am missing some things that would probably be useful things to have, but that I don't diectly draw inspiration from (e.g. Pluto, Callgrind, Simics). Fleshed out the intro a bit. Now onto design section.

@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{\Large \bf \Toolname: Precise Detection of False Sharing in COTS Binaries}
+\title{\Large \bf \Toolname: Precise False Sharing Detection for COTS Binaries}
 \numberofauthors{3}
@@
 \maketitle
-\begin{abstract}
 
-Hello
-\end{abstract}
+%\begin{abstract}
+%\end{abstract}
 
@@
 
-In this work, we present \Toolname, a system for precisely detecting the sources of false sharing in
-multi-threaded, commercial off-the-shelf (COTS) binaries. This is important because false sharing
-can limit the scalability of multi-threaded applications on cache-coherent multiprocessors
-\cite{ImpactOfFalseSharing}. The key insight of our approach is that it begins by focusing on what
-data is being accessed, and then uses that information to progressively narrow down on what code
-is participating in a data race.
+% Key ideas:
+%		1) 	Shadow memory as a proxy for memory accesses, not as a meta-data store.
+%		2) 	Tagged reads as a way to avoid cache contention on shadow memory, but still
+%			highlight the essential memory access behaviour.
+%		3)	Proportional heap sampling, by using type-specific information to decide where
+%			to place hardware watchpoints.
+
+False-sharing is a performance-degrading problem that limits the scalability of multithreaded applications
+\cite{ImpactOfFalseSharing}. This problem is increasingly important in the multi-core era. The level of
+concurrency in applications is growing to match the additional processing power provided by increasing
+core counts on modern processors. Despite this rise in importance, detecting and diagnosing false
+sharing in large systems remains challenging. Due to the nature of cache coherent multiprocessors, false
+sharing is often indistinguishable from true sharing (including race conditions). Also, diagnosing the cause
+of false sharing requires that one distinguish accesses to the same objects from ``pseudo sharing", which
+occurs when two nearby objects are separately but concurrently accessed \cite{DegenerateSharingAndFalseCoherence}.
+
+In this work, we present \Toolname{}, a system for precisely detecting and diagnosing false sharing
+in multi-threaded, commercial off-the-shelf (COTS) binaries. The key insights of our approach is that
+it iteratively discovers and specializes instrumentation to detect false sharing at increasingly finer
+granularities. Our work addresses three limitations of prior work:
+\begin{itemize}
+\item They introduce extra cache coherency traffic, which limits scalability and hides performance
+problems. The extra traffic introduced comes from the mechanisms used to track cache line ownership.
+
+\item They are unable to selectively detect false sharing on only specific types of objects or specific
+code.
+
+\item They are imprecise: they cannot distinguish between false sharing, true sharing (including
+data races), and and pseudo sharing. This can result in high false-positive rates that decrease the
+usefulness of the tools.
+
+\end{itemize}
+
+\paragraph{Contributions}
+This paper makes the following contributions:
+\begin{itemize}
+\item It presents \Toolname{}, a software-only framework that dynamically and precisely diagnoses
+false sharing  in multi-threaded, CTOS binaries. We believe that operating directly on binaries is
+important because it allows one to verify observed performance problems using the same binaries
+that are running in a production environment.
+
+\item It presents two new techniques, \emph{proxy memory} and \emph{tagged reads}, that enable
+efficient determination of memory access behavior using shadow memory. Unlike existing shadow
+memory systems, proxy memory is not treated as a mutable meta-data store. This is important
+for false sharing detection as it avoids the extra serialization and coherence traffic that is introduced
+by dynamically manipulating shadow state.
+
+\item It describes a two new, data-centric sampling approaches: \begin{inparaenum}[i)]
+\item uniform heap sampling, and
+\item selective heap sampling.
+\end{inparaenum} Uniform heap sampling enables \Toolname{} to sample objects in the heap uniformly
+according to their type. This is beneficial because it improves coverage: if 90\% of false sharing occurs because
+of accesses to only 10\% of the data, then \Toolname{} will not miss those races by blindly focusing on the
+other 90\% of data. Once data of interesting has been narrowed down, \Toolname{} can apply selective heap
+sampling, which focuses sampling on specific types of objects or data accessed by specific code (or both).
+\end{itemize}
 
@@
 
+% --------------------------------------------------------------------------------------
+
 \section{Related Work}\label{sec:background}
-\Toolname{} draws on techniques from existing race detectors and false sharing detectors. This
-is appropriate because some forms of false-positives for race detectors are caused by false
+\Toolname{} primarily draws on techniques from existing race detectors and false sharing detectors.
+This is appropriate because some forms of false-positives for race detectors are caused by false
 sharing. That is, where race detectors focus on two threads concurrently accessing the same
@@
 
-\Toolname{} is a dynamic, precise, variable-granularity, mixed-mode false-sharing detector. It draws inspiration
-from a number of existing race detectors, including LiteRace \cite{LiteRace}, MultiRace \cite{MultiRace},
-RaceMob \cite{RaceMob}, RaceTrack \cite{RaceTrack}, and DataCollider \cite{DataCollider}.
+Under this categorization scheme, \Toolname{} is a dynamic, precise, variable-granularity, mixed-mode
+false-sharing detector. It draws inspiration from a number of existing race detectors, including LiteRace
+\cite{LiteRace}, MultiRace \cite{MultiRace}, RaceMob \cite{RaceMob}, RaceTrack \cite{RaceTrack}, and
+DataCollider \cite{DataCollider}.
 
@@
 in that it uses ``pointer swizzing" as a means of remapping memory to ``minipages" that track the dominating
-access pattern of a thread (no-acces, read-only, read-write). \Toolname{} uses a similar pointer swizzing approach, but
+access pattern of a thread (read-write > read-only > no-access). \Toolname{} uses a similar pointer swizzing approach, but
 in the form of address watchpoints \cite{AddressWatchpoints}. Pointers to heap-allocated objects are ``swizzled" to
@@
 
+Similar to race detectors, false-sharing detectors can be sampling-based or comprehensive. Comprehensive approaches
+can be categorized  into simulation-based and instrumentation-based approaches. Simulation-based approaches focus on
+simulating the full cache hierarchy and can distinguish fine-grained events like false sharing due to set-associativity and
+false sharing due to different data being accessed on the same cache line. Instrumentation-based approaches focus more
+on mechanisms for estimating or detecting cache line contention.
+
+\paragraph{Sampling-based}
+Intel's VTune \cite{IntelVTune} an an imprecise, sampling-based profiler that can detect false sharing using hardware performance
+counters and monitoring events. It represents the state-of-the-art in performance analysis tools and is able to provide detailed,
+architecture-specific performance-tuning suggestions. Unlike \Toolname{}, VTune is unable to analyze only a portion of the code,
+or only the portion of code that accesses specific shared data structures.
+
+\paragraph{Comprehensive}
+
+% Simulation based approaches:
+% TODO: Callgrind/Cachegrind?
+% TODO: Simics?
+% TODO: Pluto?
+
+\newcommand{\DrContention}{Zhao \emph{et al.}}
+
+\DrContention{} \cite{DrContention} implement a dynamic false and true sharing detector using the DynamoRIO \cite{DynamoRIO}
+DBT system and Umbra shadow memory framework \cite{Umbra}. Unlike \Toolname{}, \DrContention{} use fixed-granularity
+shadow memory to track cache line ownership by processor cores. The drawback of this approach is that it necessitates checking
+and updating the shadow memory on every memory access, which introduces its own form of cache line contention. \Toolname{}
+avoids this issue by using shadow memory and tagged reads as a proxy for memory access behavior instead of as a meta-data store.
+
+Plastic \cite{Plastic} is a Xen-based dynamic false sharing detector and repairer. Like \Toolname{}, it takes a pipeline-based
+approach that successively narrows down on the root causes of false sharing. Plastic also uses DBT as part of its
+repair mechanism. Plastic's pipeline begins by using hardware performance counters related to cache contention
+events to detect the presence of false-sharing. Then, it applies a similar technique to MultiRace \cite{MultiRace}, in
+that it tracks the dominant access pattern by each thread to memory on a shared page. This information informs
+their repair mechanism, which transparently re-maps memory accesses at a byte granularity.
+
+SHERIFF \cite{SHERIFF} is a false-sharing detector and repairer that works in a similar way to MultiRace and Plastic. Initially,
+it applies read-only hardware page protection to all shared memory. Then, as threads attempt to write to protected pages, those
+pages are privately mapped as read/write. When the same page is privately mapped by two or more threads, SHERIFF applies
+the twinning and diffing approach of TreadMarks \cite{TreadMarks} to detect local changes to shared pages and narrow down
+on instances of false sharing. Like \Toolname{}, SHERIFF uses a coarse-grained mechanism to hint at the potential for races.
+
+% --------------------------------------------------------------------------------------
 \section{Design}\label{sec:design}
+
+% --------------------------------------------------------------------------------------
+\section{Implementation}\label{sec:implementation}
+\Toolname{} is implemented using the Granary dynamic binary translator \cite{Granary}, as well as the address watchpoints
+\cite{AddressWatchpoints} framework.
 
@@
 
+% Answer the following questions:
+%		1)	What are DataReactor's overheads over time?
+%		2) 	How sensitive is DataReactor to different sampling rates?s
+%		3)	How effective is DataReactor as discovering false sharing?
+%		4)	Are the cases of detected false sharing high impact?
+
 \bibliographystyle{abbrv}

### 2014-11-05 HotOS'15/paper.tex
Working on the design section.

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 
@@
 granularities. Our work addresses three limitations of prior work:
-\begin{itemize}
+\begin{enumerate}
 \item They introduce extra cache coherency traffic, which limits scalability and hides performance
@@
 
-\end{itemize}
+\end{enumerate}
 
@@
 
+\newcommand\TypeIdPair{$\langle PC,size \rangle$ }
+
 MultiRace \cite{MultiRace} is a comprehensive, dynamic race detector that employs a hybrid happens-before and
@@
 in the form of address watchpoints \cite{AddressWatchpoints}. Pointers to heap-allocated objects are ``swizzled" to
-include a type identifier that is unique to a $\langle PC,size \rangle$ pair. This improves the diagnosis of false-sharing,
+include a type identifier that is unique to a \TypeIdPair{} pair. This improves the diagnosis of false-sharing,
 because it allows \Toolname{} to distinguish between two threads accessing different fields of the same object, and
@@
 
+At a high-level, \Toolname{} follows the same principle as Plastic \cite{Plastic} and SHERIFF \cite{SHERIFF}: it detects
+false sharing at coarse granularity and then verifies cases of false sharing at a finer granularity. Unlike other approaches,
+\Toolname{} repeats this process by incrementally refining the detection granularity, and by selectively instrumenting only
+that code that appears to trigger false sharing.
+
+As described, this design presents two challenges. First, because our proportional sampling approach will sample
+data of all types, we might find instances of false sharing that have no significant performance impact. Therefore,
+we need a means of estimating the performance impact of false sharing so that we can rank detected instances of
+false sharing from most to least important. We accomplish this as part of a separate analysis step. This step
+instruments all code to count how often each basic block is executed. We correlate this information with detected
+instances of false sharing to infer cases where one or both threads are executing hot code and contending on a single
+cache line. In the former case, we apply a variant of the cold code hypothesis \cite{LiteRace} to rank cases of cold
+code ``pessimizing" hot code more highly.
+
+Second, our sampling approach can miss false sharing, simply by sampling the right data at the wrong time, or by
+sampling unshared data. Our approach to solving these these problems is to infer beliefs about the sharing of
+objects of specific types, and to propagate these beliefs to code operating on those objects. By propagating these
+beliefs to code, we can selectively instrument only specific code and sample specific data in order to discover
+false sharing on specific types of objects. We discuss this further in \Secref{beliefs}.
+
+\subsection{Type-specific Sampling}\label{sec:beliefs}
+
+%First, a sampling based approach implies the potential for false-negatives. That is, \Toolname{} can report that no
+%false sharing was detected in some particular code. This, however, does not imply the absense of false sharing. Merely,
+%it suggests that if there is false sharing, then it was missed because the wrong data was sampled at the wrong time.
+
+%\begin{description}
+%\item[Code Hotness Analysis] \hfill \\
+%Here we instrument all code to count how often each basic block is executed. The results of this analysis are used when 
+%
+%The purpose of this
+%stage is to measure code hotness for the sake of ranking detected instances of false sharing. The accuracy of this stage depends
+%on the workload used to exercise the code. If this workload does not resemble production workloads then the recorded metrics
+%might rank false sharing instances incorrectly.
+%
+%\newcommand\TypeId{$TypeId$}
+%
+%\item[Type Analysis] \hfill \\
+%This stage instruments all code and identifies what types of objects are accessed by what code. Object types are loosely
+%defined as \TypeIdPair{} pairs, where $PC$ is the return address of an allocation function, and $size$ is the number of bytes
+%being allocated. To each \TypeIdPair pair, we associate a unique \TypeId{}. This \TypeId{} is used to ``taint" allocated
+%addresses \cite{AddressWatchpoints}, and when accessed, is used to taint the code accessing the data with the \TypeId{}.
+%The overarching purpose of this step is to map every program counter to a set of 
+%
+%\end{description}
+
+%\subsection{Code Hotness}
+%\Toolname{} instruments all code and counts the number of times each basic block is executed.
+
 % --------------------------------------------------------------------------------------
@@
 
+% --------------------------------------------------------------------------------------
 \section{Status}\label{sec:status}
 
+% --------------------------------------------------------------------------------------
 \section{Evaluation}\label{sec:evaluation}
@@
 
+% --------------------------------------------------------------------------------------
 \bibliographystyle{abbrv}

### 2014-11-05 HotOS'15/paper.tex
improved the 'key insight' portion of the intro. Still a lot of work to do, but heading to bed.

@@
 In this work, we present \Toolname{}, a system for precisely detecting and diagnosing false sharing
-in multi-threaded, commercial off-the-shelf (COTS) binaries. The key insights of our approach is that
-it iteratively discovers and specializes instrumentation to detect false sharing at increasingly finer
-granularities. Our work addresses three limitations of prior work:
+in multi-threaded applications. \Toolname{} uses dynamic binary translation (DBT) and shadow memory
+to instrument and detect false sharing in commercial off-the-shelf (COTS) binaries.  The key insight of our
+approach is that accesses to shadow memory can be used as a proxy for accesses to actual memory. By 
+detecting contention on proxied accesses, we can establish beliefs about the sources of false sharing. These
+beliefs are strenghtened and eventually verified by specializing code to proxy memory accesses at increasingly
+finer granularities.
+
+Our work addresses three limitations of prior work:
 \begin{enumerate}
@@
 \paragraph{Contributions}
-This paper makes the following contributions:
+This paper makes the following significant contributions:
 \begin{itemize}
@@
 At a high-level, \Toolname{} follows the same principle as Plastic \cite{Plastic} and SHERIFF \cite{SHERIFF}: it detects
-false sharing at coarse granularity and then verifies cases of false sharing at a finer granularity. Unlike other approaches,
+false sharing at a coarse granularity and then verifies detected cases at a finer granularity. Unlike other approaches,
 \Toolname{} repeats this process by incrementally refining the detection granularity, and by selectively instrumenting only
-that code that appears to trigger false sharing.
+that code that appears to trigger false sharing. Whereas Plastic and SHERIFF use page-granularity detection, \Toolname{}
+is able to vary the granularity over time using a shadow memory. The key insight of our approach is that accesses to
+shadow memory at varying granularities can be used as a proxy for accesses to actual memory.
+
+\Toolname{} operates as follows:
+\begin{enumerate}
+\item Choose 
+\end{enumerate}
 
@@
 false sharing on specific types of objects. We discuss this further in \Secref{beliefs}.
+
+\subsection{Proxy Memory and Tagged Reads}\label{sec:proxymem}
 

### 2014-11-05 HotOS'15/paper.tex
Renamed design to methodology, used a bit of the same style as engler.. mwahahaha. Strengthening the idea of beliefs about code, as I think that accurately describes what is being done.

--- 2014-11-04 HotOS'15/paper.tex
@@
 sharing in large systems remains challenging. Due to the nature of cache coherent multiprocessors, false
-sharing is often indistinguishable from true sharing (including race conditions). Also, diagnosing the cause
+sharing is often indistinguishable from true sharing (including data races). Also, diagnosing the cause
 of false sharing requires that one distinguish accesses to the same objects from ``pseudo sharing", which
@@
 In this work, we present \Toolname{}, a system for precisely detecting and diagnosing false sharing
-in multi-threaded, commercial off-the-shelf (COTS) binaries. The key insights of our approach is that
-it iteratively discovers and specializes instrumentation to detect false sharing at increasingly finer
-granularities. Our work addresses three limitations of prior work:
-\begin{itemize}
+in multi-threaded applications. \Toolname{} uses dynamic binary translation (DBT) and shadow memory
+to detect false sharing in commercial off-the-shelf (COTS) binaries.  The key insight of our
+approach is that accesses to shadow memory can be used as a proxy for accesses to actual memory. By 
+detecting contention on proxied accesses, we can establish beliefs about what code causes false sharing. These
+beliefs are strenghtened and eventually verified by specializing code to proxy memory accesses at increasingly
+finer granularities.
+
+Our work addresses three limitations of prior work:
+\begin{enumerate}
 \item They introduce extra cache coherency traffic, which limits scalability and hides performance
@@
 
-\end{itemize}
+\end{enumerate}
 
 \paragraph{Contributions}
-This paper makes the following contributions:
+This paper makes the following significant contributions:
 \begin{itemize}
@@
 efficient determination of memory access behavior using shadow memory. Unlike existing shadow
-memory systems, proxy memory is not treated as a mutable meta-data store. This is important
-for false sharing detection as it avoids the extra serialization and coherence traffic that is introduced
-by dynamically manipulating shadow state.
+memory systems, proxy memory is not treated as a mutable meta-data store. This avoids extra
+serialization and coherence traffic introduced by prior work when dynamically manipulating shadow state.
 
@@
 % --------------------------------------------------------------------------------------
-\section{Design}\label{sec:design}
+\section{Methodology}\label{sec:methodology}
+
+This section introduces our approach to finding false sharing bugs. The goal of our approach is extract beliefs about
+where false sharing might be, then iteratively strenghten those beliefs, either verifying or discarding them in the process.
+
+We begin with the same intuition as SHERIFF: if false sharing occurs on a single cache line, then that implies that two
+threads are simultaneously accessing the same page of memory \cite{SHERIFF}. Therefore, by detecting contention on
+a single page, we can establish a belief that false sharing might occur on the contended page.
+
+Unfortunately, a page-granularity detection of false sharing is na{\"i}ve: it provides no guidance as to the location of the
+contended cache line, if any. Our insight is that we can narrow down on instances of false sharing by varying the
+granularity of the mechanism used to detect thread contention. That is, we can start with a coarse-grained but
+innacurate approach that tells us what code might cause false sharing, thus establishing an initial belief set. Then, we
+can strengthen or discard beliefs by repeating this same process, but using finer-grained detection mechanisms that
+apply only to code thought to contain false sharing.
+
+The remainder of this section explains how we vary the detection granularity (\Secref{proxymem}), how we use sampling
+to detect contending accesses to shared memory (\Secref{heapsample}), how we reduce false negatives with our
+sampling-based approach (\Secref{falsenegs}), and finally how we rank discovered instances of false sharing (\Secref{ranking})
+
+\subsection{Proxy Memory and Tagged Reads}\label{sec:proxymem}
+
+\subsection{Heap Sampling}\label{sec:heapsample}
+
+\subsection{Reducing False Negatives}\label{sec:falsenegs}
+
+One challenge with sampling-based approaches is their ability to get ``unlucky" and either not sample
+the right thing, or sample the right thing at the wrong time. In our case, this results in false-negatives:
+we might report that no false sharing is detected in some code, but this does not imply the absence of
+false sharing.
+
+To address this challenge, we propagate existing beliefs about potential false sharing on a single instruction
+to nearby instructions, as well as to instructions that operate on data of the same type. The intuition is
+two-fold. First, we assume locality of accesses to shared data: if we observe a shared access on one instruction,
+then nearby instructions probably access shared data as well. Second, a belief of false sharing is a belief of
+sharing: if we observe an object of a particular type being concurrently accessed in one place, then it might
+imply that concurrent accesses can also happen in other places.
+
+\newcommand\TypeId{$TypeId$}
+
+We perform two separate analyses to extend belief sets: \begin{description}
+\item[Type Analysis] \hfill \\
+This analysis instruments all code and identifies what types of objects are accessed by what code. Object types are loosely
+defined as \TypeIdPair{} pairs, where $PC$ is the return address of an allocation function, and $size$ is the number of bytes
+being allocated. To each \TypeIdPair pair, we associate a unique \TypeId{}. We taint \cite{AddressWatchpoints} allocated
+addresses with their \TypeId{}, and propagate this taint to code accessing typed data.
+
+\item[Concurrency Analysis] \hfill \\
+This analysis instruments all code and identifies what functions can execute concurrently. Each instrumented function
+is associated with a core set and a threadset \cite{DrContention}. When executed, the coreset is updated to include the
+identifier of the CPU core executing the function. Similarly, the function's threadset is updated to include an approximation
+of the thread identifier ($SP >> 13$). We anlayse the coresets and threadsets of each function to determine whether or
+not a function can execute concurrently (across CPU cores, or in two threads on the same core).
+
+\end{description}
+
+We combine the type and concurrency analyses to propagate beliefs in the following way: if sampling detects potential
+false sharing on a single instruction operating on a specific type of object, then we mark all instructions in the intersection
+of those identified to operate on the same type with the set of concurrent functions.
+
+\subsection{Ranking}\label{sec:ranking}
+
+Because we sample data of all types, we might find instances of false sharing that have no significant impact on performance.
+Therefore, we need a means of estimating the performance impact of false sharing so that we can rank detected instances of
+false sharing from most to least important. We accomplish this as part of a separate analysis step. This step
+instruments all code to count how often each basic block is executed. We correlate this information with detected
+instances of false sharing to infer cases where one or both threads are executing hot code and contending on a single
+cache line. In the former case, we apply a variant of the cold code hypothesis \cite{LiteRace} to rank cases of cold
+code ``pessimizing" hot code more highly.
+
+%At a high-level, \Toolname{} follows the same principle as Plastic \cite{Plastic} and SHERIFF \cite{SHERIFF}: it detects
+%false sharing at a coarse granularity and then verifies detected cases at a finer granularity. Unlike other approaches,
+%\Toolname{} repeats this process by incrementally refining the detection granularity, and by selectively instrumenting only
+%that code that appears to trigger false sharing. Whereas Plastic and SHERIFF use page-granularity detection, \Toolname{}
+%is able to vary the granularity over time using a \emph{proxy memory} (\Secref{proxymem}).
 
 % --------------------------------------------------------------------------------------
-\section{Implementation}\label{sec:implementation}
-\Toolname{} is implemented using the Granary dynamic binary translator \cite{Granary}, as well as the address watchpoints
+\section{Status}\label{sec:status}
+\Toolname{} is being implemented using the Granary dynamic binary translator \cite{Granary}, as well as the address watchpoints
 \cite{AddressWatchpoints} framework.
 
-\section{Status}\label{sec:status}
+The following has been implemented:
+% TODO...
 
+% --------------------------------------------------------------------------------------
 \section{Evaluation}\label{sec:evaluation}

### 2014-11-05 HotOS'15/paper.tex
Still need to do the subsection on proxy memory.

@@
 \item It presents \Toolname{}, a software-only framework that dynamically and precisely diagnoses
-false sharing  in multi-threaded, CTOS binaries. We believe that operating directly on binaries is
+false sharing  in multi-threaded, COTS binaries. We believe that operating directly on binaries is
 important because it allows one to verify observed performance problems using the same binaries
@@
 
+% --------------------------------------
 \subsection{Race Detection}
@@
 
+% --------------------------------------
 \subsection{False Sharing Detection}
@@
 
-The remainder of this section explains how we vary the detection granularity (\Secref{proxymem}), how we use sampling
+The remainder of this section explains how we proxy memory access behavior (\Secref{proxymem}), how we use sampling
 to detect contending accesses to shared memory (\Secref{heapsample}), how we reduce false negatives with our
-sampling-based approach (\Secref{falsenegs}), and finally how we rank discovered instances of false sharing (\Secref{ranking})
+sampling-based approach (\Secref{falsenegs}), how we achieve precision by eliminating false positives (\Secref{falsepos}),
+and finally how we rank discovered instances of false sharing (\Secref{ranking})
 
+% --------------------------------------
 \subsection{Proxy Memory and Tagged Reads}\label{sec:proxymem}
 
+Proxy memory is a form of read-only shadow memory that summarizes or proxies actual memory access behavior. The key
+idea  .... TODO!!
+
+% --------------------------------------
 \subsection{Heap Sampling}\label{sec:heapsample}
 
+\Toolname{} follows DataCollider's sampling based approach, but samples proxy memory instead of code \cite{DataCollider}.
+\Toolname{} applies the roughly the following high-level algorithm to sample proxy memory and detect potential false sharing:
+\begin{enumerate}
+\item Choose a proxy memory address $P$ uniformly at random. Add a hardware watchpoint $w(P)$ to $P$.
+\item If the watchpoint $w(P)$ is not triggered within $t$ units of time, then remove $w(P)$ and go to step (1).
+\item If $w(P)$ is triggered for thread $T_i$ then pause $T_i$ for $t$ units of time. The hope is that within this
+time slice, another thread, $T_j$ will be caught in the act of false sharing by triggering $w(P)$.
+\item If $w(P)$ is not triggered again for another $t$ time units, then remove $w(P)$ and go to step (1).
+\item If another thread $T_j$ triggers $w(P)$, then record stack traces for $T_i$ and $T_j$, and mark the associated program
+counters of each of $T_i$ and $T_j$ as potentially having false sharing.
+\item Go to step (1) to repeat the process for a new sample location.
+\end{enumerate}
+
+As described, our sampling algorithm is susceptible to bias: if only a small percentage of the application's data is concurrently
+modified, or ``hot", then our algorithm might miss potential false sharing by wasting time sampling from the cold data.
+To avoid this issue, we apply two new techniques: \emph{uniform heap sampling}, and \emph{selective heap sampling}.
+
+\newcommand\TypeId{$TypeId$}
+
+\paragraph{Uniform Heap Sampling}
+Uniform heap sampling is a round-robin algorithm that uniformly selects proxy memory locations to sample according to
+the types of data proxied. We interpose on memory allocation and deallocation routines (e.g. \texttt{malloc} and \texttt{free})
+and associate unique \TypeId{}s to each \TypeIdPair{}, where $PC$ is the return address of an allocation function, and $size$ is the
+number of bytes allocated.
+
+For each proxy memory location $P$, we maintain a counting \emph{typeset}: a mapping of \TypeId{}s to the number of objects of type
+\TypeId{} that are fully or partially proxied by $P$.
+
+\paragraph{Selective Heap Sampling}
+Selective heap sampling is a biased variant of uniform heap sampling that allows a user to guide the sampler to only focus
+on specific types of data. This is beneficial when one is trying to solve specific performance problems related to a subset
+of the data being accessed by the application.
+
+% --------------------------------------
 \subsection{Reducing False Negatives}\label{sec:falsenegs}
@@
 
-\newcommand\TypeId{$TypeId$}
-
 We perform two separate analyses to extend belief sets: \begin{description}
 \item[Type Analysis] \hfill \\
-This analysis instruments all code and identifies what types of objects are accessed by what code. Object types are loosely
-defined as \TypeIdPair{} pairs, where $PC$ is the return address of an allocation function, and $size$ is the number of bytes
-being allocated. To each \TypeIdPair pair, we associate a unique \TypeId{}. We taint \cite{AddressWatchpoints} allocated
-addresses with their \TypeId{}, and propagate this taint to code accessing typed data.
+This analysis instruments all code and identifies what types of objects are accessed by what code. We apply the same
+notion of typesets as used in the uniform heap sampler; however, in this case typesets are associated with basic
+blocks of code instead of with proxy memory. We use address watchpoints \cite{AddressWatchpoints} to taint
+allocated addresses with their \TypeId{}, and then propagate these \TypeId{}s to the typesets of the blocks operating
+of data of each type.
 
@@
 
+% --------------------------------------
+\subsection{Eliminating False Positives}\label{sec:falsepos}
+
+We apply two techniques to reduce and finally eliminate false positives:
+\begin{enumerate}
+\item We iteratively refine the granularity of proxy memory, starting from the page granularity and going down to the cache
+line granularity. As the granularity is refined and potential false sharing continues to be detected, our confidence in beliefs
+of false sharing improves.
+
+\item We apply a precise data race detector \cite{DataCollider} to instructions that demonstrate cache line contention. This
+allows us to distinguish between true and false sharing.
+\end{enumerate}
+
+Distinguishing between false sharing and pseudo sharing is more challenging. We approximate this detection by extending
+the type analysis phase to record the relative byte offset within each type of accessed object for each instruction. If false
+sharing is detected between two instructions operating on the same offset ($\pm$ the memory access size), then we report
+pseudo sharing, otherwise false sharing. Another valid heuristic would include the natural alignment and sizes of object
+types.
+
+% TODO: How to make it precise? Apply DataCollider to places where we've found potential false sharing at the granularity
+% of a cache line.
+
+% How do we distinguish between false sharing and pseudo sharing? Use the type analysis feature to figure out what offsets
+% into what data were being accessed.h
+
+% --------------------------------------
 \subsection{Ranking}\label{sec:ranking}

### 2014-11-05 HotOS'15/paper.tex
Done report for ashvin's course.

@@
 \subsection{Proxy Memory and Tagged Reads}\label{sec:proxymem}
+Proxy memory is a form of variable-granularity shadow memory that proxies actual memory access behavior. The key
+idea is to instrument all loads and stores in such a way that their load/store behavior is reflected in the shadow
+memory in a way that can be externally observed. For example, one way to achieve this is to instrument every load
+to read from shadow memory, every store to modify shadow memory, and trap loads and stores by adding hardware
+watchpoints to shadow memory locations. Such heavyweight instrumentation would be inefficient, and so we apply
+two optimizations: belief-based selective instrumentation, and tagged reads.
 
-Proxy memory is a form of read-only shadow memory that summarizes or proxies actual memory access behavior. The key
-idea  .... TODO!!
+\paragraph{Selective Instrumentation}
+The first optimization to proxy memory is to not instrument all code to operate on proxy memory. Initially, \Toolname{}
+comprehensively instruments all code, thus incurring significant overheads. However, as beliefs about false sharing
+are refined, less and less code is instrumented, thus allowing code to run faster when precision during identification of
+potential false sharing is needed.
+
+\paragraph{Tagged Reads}
+The second optimization avoids the problem of introducing unnecessary cache coherency traffic by treating proxy memory
+as read-only, and disambiguating the intention of an operation (load, store) by using a \emph{tagged read}. In our implementation,
+a read operations (on the proxy memory) that intend to communicate loads (of actual memory) are distinguished from
+read operations that intend to communicate a store by using different encodings of the logical \texttt{OR} instruction.
 
@@
 \Toolname{} follows DataCollider's sampling based approach, but samples proxy memory instead of code \cite{DataCollider}.
-\Toolname{} applies the roughly the following high-level algorithm to sample proxy memory and detect potential false sharing:
+\Toolname{} applies roughly the following high-level algorithm to sample proxy memory and detect potential false sharing:
 \begin{enumerate}
@@
 \item If $w(P)$ is triggered for thread $T_i$ then pause $T_i$ for $t$ units of time. The hope is that within this
-time slice, another thread, $T_j$ will be caught in the act of false sharing by triggering $w(P)$.
+time slice, another thread, $T_j$ will be caught in the act of contending on $P$ by triggering $w(P)$.
 \item If $w(P)$ is not triggered again for another $t$ time units, then remove $w(P)$ and go to step (1).
@@
 \section{Status}\label{sec:status}
-\Toolname{} is being implemented using the Granary dynamic binary translator \cite{Granary}, as well as the address watchpoints
-\cite{AddressWatchpoints} framework.
+\Toolname{} is being implemented using the Granary dynamic binary translator \cite{Granary} for the x86-64 architecture.
 
-The following has been implemented:
+A limited form of proxy memory has been implemented. Work is progressing on making it possible to generically interpose
+on memory allocators. Once this is done, the next step will be to work on the type analysis, as well as to associate typesets
+with proxy memory locations. Code hotness analysis is expected to be an easy-to-do task. Concurrency analysis will be trickier.
+
+Sampling and hardware watchpoint management will be orchestrated using GDB scripts.
+
 % TODO...
@@
 \section{Evaluation}\label{sec:evaluation}
+Our evaluation aims to answer the following questions:
+\begin{itemize}
+\item What are \Toolname{}'s overheads? For example, what was the performance impact of the selective instrumentation
+and tagged read optimizations?
 
-% Answer the following questions:
-%		1)	What are DataReactor's overheads over time?
-%		2) 	How sensitive is DataReactor to different sampling rates?s
-%		3)	How effective is DataReactor as discovering false sharing?
-%		4)	Are the cases of detected false sharing high impact?
+\item How sensitive is \Toolname{} to different sampling rates, delays, and approaches? How does the thread
+pause time during sampling affect our ability to detect false sharing? Can we quantify the benefits (if any) of
+uniform heap sampling over na{\"i}ve heap sampling?
+
+\item How effective is \Toolname{} at discovering false sharing? Is it able to distinguish between true sharing, false sharing,
+and pseudo sharing?
+
+\item How impactful are the instances of false sharing detected by \Toolname{}? Is code hotness a good metric for ranking
+false sharing?
+\end{itemize}
+
+\subsection{Evaluation Challenges}
+Granary is a medium weight DBT system, therefore its overheads will likely be significant. It is also new, and so no time has been
+spent optimizing the code produced by Granary. Therefore, the overheads are not expected to be particularly good. The major
+challenge will be convincing readers that:
+\begin{enumerate}
+\item Most of the analysis can be performed in a non-production environment, where we can claim that speed doesn't matter.
+This will obviously be a contentious claim, because speed always matters, especially in false sharing detection. Here, I hope
+that the belief propagation steps of the type and concurrency analysis stages will be beneficial in allaying those concerns.
+
+\item The application of DataCollider to distinguish true and false sharing can be performed in production given the results of
+all prior stages of the pipeline.
+\end{enumerate}
 

### 2014-11-06 HotOS'15/paper.tex
Minor typo fixes.

@@
 
-\item It describes a two new, data-centric sampling approaches: \begin{inparaenum}[i)]
-\item uniform heap sampling, and
-\item selective heap sampling.
-\end{inparaenum} Uniform heap sampling enables \Toolname{} to sample objects in the heap uniformly
-according to their type. This is beneficial because it improves coverage: if 90\% of false sharing occurs because
-of accesses to only 10\% of the data, then \Toolname{} will not miss those races by blindly focusing on the
-other 90\% of data. Once data of interesting has been narrowed down, \Toolname{} can apply selective heap
-sampling, which focuses sampling on specific types of objects or data accessed by specific code (or both).
+\item It describes two new, data-centric sampling approaches: uniform heap sampling, and selective heap
+sampling. Uniform heap sampling enables \Toolname{} to sample objects in the heap uniformly according to
+their type. This is beneficial because it improves coverage: if 90\% of false sharing occurs because of accesses
+to only 10\% of the data, then \Toolname{} will not miss those races by blindly focusing on the other 90\% of data.
+Once data of interesting has been narrowed down, \Toolname{} can apply selective heap sampling, which
+focuses sampling on specific types of objects or data accessed by specific code (or both).
 \end{itemize}

### 2014-11-06 HotOS'15/paper.tex
Changed the described purpose of the selective heap shadowing a bit.

@@
 \paragraph{Selective Heap Sampling}
-Selective heap sampling is a biased variant of uniform heap sampling that allows a user to guide the sampler to only focus
-on specific types of data. This is beneficial when one is trying to solve specific performance problems related to a subset
-of the data being accessed by the application.
+Selective heap sampling is a biased variant of uniform heap sampling that enables \Toolname{} to only focus on the subset
+of the heap that is believed to contain objects that participate in false sharing. Selective heap sampling builds typesets
+for only a subset of proxy memory associated with specific allocation points.
+
+Selective heap sampling also allows a user to guide the sampler to only focus on specific types of data. This is beneficial
+when one is trying to solve specific performance problems related to a subset of the data being accessed by the application.
 

### 2014-12-13 HotOS'15/paper.tex
Working on related work section. Re-did most of the intro. No abstract yet.

@@
 \newcommand{\comment}[1]{}
-\newcommand{\Toolname}{DataReactor}
+\newcommand{\TextToolname}{CatcheScan}
+\newcommand{\Toolname}{\textsc{\TextToolname{}}}
+\newcommand{\BoldToolname}{\textsc{\bf\TextToolname{}}}
 
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{\Large \bf \Toolname: Precise False Sharing Detection for COTS Binaries}
+\title{\Large \bf \Toolname: Cache Line Contention Detection for COTS Binaries}
 \numberofauthors{3}
@@
 
-% Key ideas:
-%		1) 	Shadow memory as a proxy for memory accesses, not as a meta-data store.
-%		2) 	Tagged reads as a way to avoid cache contention on shadow memory, but still
-%			highlight the essential memory access behaviour.
-%		3)	Proportional heap sampling, by using type-specific information to decide where
-%			to place hardware watchpoints.
+In multi-threaded applications, cache-line contention is a symptom of several kinds of bugs: data races, true sharing,
+false sharing, and pseudo sharing. Data races are potential correctness bugs that introduce non-determinism into running
+programs. True, false, and pseudo sharing are performance bugs that limit the scalability of multi-threaded applications
+\cite{ImpactOfFalseSharing}. Both problems are important in the multi-core era, where the level of concurrency in
+applications is growing to match the additional processing power provided by increasing core counts on modern processors.
 
-False-sharing is a performance-degrading problem that limits the scalability of multithreaded applications
-\cite{ImpactOfFalseSharing}. This problem is increasingly important in the multi-core era. The level of
-concurrency in applications is growing to match the additional processing power provided by increasing
-core counts on modern processors. Despite this rise in importance, detecting and diagnosing false
-sharing in large systems remains challenging. Due to the nature of cache coherent multiprocessors, false
-sharing is often indistinguishable from true sharing (including data races). Also, diagnosing the cause
-of false sharing requires that one distinguish accesses to the same objects from ``pseudo sharing", which
-occurs when two nearby objects are separately but concurrently accessed \cite{DegenerateSharingAndFalseCoherence}.
+% Despite this rise in importance, detecting and diagnosing false
+%sharing in large systems remains challenging. Due to the nature of cache coherent multiprocessors, false
+%sharing is often indistinguishable from true sharing (including data races). Also, diagnosing the cause
+%of false sharing requires that one distinguish accesses to the same objects from ``pseudo sharing", which
+%occurs when two nearby objects are separately but concurrently accessed \cite{DegenerateSharingAndFalseCoherence}.
 
-In this work, we present \Toolname{}, a system for precisely detecting and diagnosing false sharing
-in multi-threaded applications. \Toolname{} uses dynamic binary translation (DBT) and shadow memory
-to detect false sharing in commercial off-the-shelf (COTS) binaries.  The key insight of our
-approach is that accesses to shadow memory can be used as a proxy for accesses to actual memory. By 
-detecting contention on proxied accesses, we can establish beliefs about what code causes false sharing. These
-beliefs are strenghtened and eventually verified by specializing code to proxy memory accesses at increasingly
-finer granularities.
+This paper presents \BoldToolname{}, a software-only runtime for precisely detecting and diagnosting cache line
+contention in commercial, off-the-shelf (COTS) x86-64 program binaries. \Toolname{} addresses three limitations of prior
+work: \begin{enumerate}
+\item They are not general: they require an understanding of the semantics of all synchronization primitives used
+by the program. This is unsatisfactory for COTS binaries that employ lock-free algorithms and home-grown synchronization
+primitives.
 
-Our work addresses three limitations of prior work:
-\begin{enumerate}
-\item They introduce extra cache coherency traffic, which limits scalability and hides performance
-problems. The extra traffic introduced comes from the mechanisms used to track cache line ownership.
+\item They are imprecise: they cannot distinguish between data races, true sharing, false sharing, and pseudo sharing
+\cite{DegenerateSharingAndFalseCoherence}. This can result in high false-positive rates or confusing reports that decrease
+the usefulness of these tools for practitioners.
 
-\item They are unable to selectively detect false sharing on only specific types of objects or specific
-code.
+\item They are not comprehensive: they miss actual cases of cache line contention because they use heuristics to sample 
+or instrument only a subset of the code, and therefore lose visibility on accesses to data that might cause bugs.
 
-\item They are imprecise: they cannot distinguish between false sharing, true sharing (including
-data races), and and pseudo sharing. This can result in high false-positive rates that decrease the
-usefulness of the tools.
+%\item They are not selective. While apparently at odds with comprehensiveness, they track or instrument \emph{too much}, 
+%and therefore waste time looking for problems where none exist.
+
+%\item They introduce extra cache coherency traffic, which limits scalability and hides performance
+%problems. The extra traffic introduced comes from the mechanisms used to track cache line ownership.
+
+%\item They are unable to selectively detect false sharing on only specific types of objects or specific
+%code.
 
 \end{enumerate}
+
+\Toolname{} uses dynamic binary translation (DBT) to instrument a program binary and detect cache line contention
+at runtime. Our approach combines unlimited watchpoints \cite{UnlimitedWatchpoints,DynamoRIOWatchpoints}, data
+sampling, and a lightweight cache line ownership tracking mechanism to detect cache line contention. The key insight
+of our approach is that if we know \begin{inparaenum}[i)]
+\item what code is executed by more than one thread,
+\item what types of data are accessed by more than one thread, and
+\item what types of data are accessed by what code, then
+\end{inparaenum}
+we can efficiently detect cache line contention by selectively instrumenting only the code and the data that matters.
+
+%we can extend on and improve existing techniques to detecting
+%cache line contention by combining solutions the following problems:
+%\begin{enumerate}
+%\item Code concurrency: Detecting what code is executed by more than one thread.
+%\item Data concurrency: Detecting what types of data are accessed by more than one thread.
+%\item Code polymorphism: Detecting what code accesses what types of data.
+%\end{enumerate}
+
+% infinite software watchpoints, as implemented within a DBT system, as
+%well as being selective about what parts of the program to
+
+%The key insight of our
+%approach is that accesses to shadow memory can be used as a proxy for accesses to actual memory. By 
+%detecting contention on proxied accesses, we can establish beliefs about what code causes false sharing. These
+%beliefs are strenghtened and eventually verified by specializing code to proxy memory accesses at increasingly
+%finer granularities.
 
@@
 \begin{itemize}
-\item It presents \Toolname{}, a software-only framework that dynamically and precisely diagnoses
-false sharing  in multi-threaded, COTS binaries. We believe that operating directly on binaries is
-important because it allows one to verify observed performance problems using the same binaries
-that are running in a production environment.
+\item It describes \Toolname{}, a software-only system that dynamically and precisely diagnoses cache line
+contention in multi-threaded, COTS binaries. This enables \Toolname{} to detect performance
+and correctness bugs in programs using home-grown synchronization primitives and lock-free algorithms.
 
-\item It presents two new techniques, \emph{proxy memory} and \emph{tagged reads}, that enable
-efficient determination of memory access behavior using shadow memory. Unlike existing shadow
-memory systems, proxy memory is not treated as a mutable meta-data store. This avoids extra
-serialization and coherence traffic introduced by prior work when dynamically manipulating shadow state.
+\item It describes a new data-centric sampling approach, \emph{comprehensive data sampling}, that improves
+sampling coverage and contention detection accuracy. Comprehensive sampling uniformly samples objects in
+memory  according to their type. This avoids problems related to sampling bias in situations where a
+relatively small amount of shared data is crowded out by a much larger amount of unshared data (as happens
+when multi-threaded programs operate on partitions of a data set).
 
-\item It describes two new, data-centric sampling approaches: uniform heap sampling, and selective heap
-sampling. Uniform heap sampling enables \Toolname{} to sample objects in the heap uniformly according to
-their type. This is beneficial because it improves coverage: if 90\% of false sharing occurs because of accesses
-to only 10\% of the data, then \Toolname{} will not miss those races by blindly focusing on the other 90\% of data.
-Once data of interesting has been narrowed down, \Toolname{} can apply selective heap sampling, which
-focuses sampling on specific types of objects or data accessed by specific code (or both).
+%from wasting cycles looking for contention on
+%unshared data
+
+%This is beneficial because it improves coverage: if 90\% of false sharing occurs because of accesses
+%to only 10\% of the data, then \Toolname{} will not miss those races by blindly focusing on the other 90\% of data
+
+\item It describes a novel separation of cache line contention detection into two phases: \begin{enumerate}
+\item[1) \emph{Training:}] an ahead-of-time, heavyweight analysis of the program's execution that
+discovers what code can contain cache line contention, and what data is shared by multiple threads. A key
+aspect of this phase is that it \emph{does not} need to exhibit any cache line contention; the only requirement
+is that training and production workloads exercise/cover the code in similar ways.
+
+% TODO: Evaluate by showing the percentage of code that doesn't get instrumented at all, or is only
+% instrumented with lightweight instrumentation.
+
+\item[2) \emph{Detection:}] an online analysis that detects cache line contention by selectively
+instrumenting only the relevant code and sampling only the relevant data as recognized by the
+offline training phase.
+\end{enumerate}
+
+This separation is motivated by the need to reduce instrumentation overheads of the detection phase by
+as much as possible in order to maximize the detection accuracy (e.g. data races are time-sensitive). The key
+insight enabling this separation is that time--and therefore the cost of instrumentation--does not matter to
+the accuracy of the results produced by the training phase.
+
+%A key aspect of the training phase is that it must exhibit approximately the same code coverage as a
+%production
+%that it does not need to exhibit a buggy execution (i.e. contain actual
+%cache line contention)
+
+%It describes a novel separation of cache line contention detection into two phases: \begin{enumerate}
+%\item[1) \emph{Training:}] This phase uses heavyweight instrumentation to discover \begin{inparaenum}
+%\item what code is hot;
+%\item what code is executed by more than one thread;
+%\item what code accesses what data; and,
+%\item function call graphs.
+%\end{inparaenum}
+%
+%This information enables the detection phase to selectively prune the set of code that is instrumented,
+%and the set of data that is sampled in order to reduce overheads. 
+%
+%
+%
+%\item[2) \emph{Detection:}] b
+%\end{enumerate}
+
+%ahead-of-time training phase that discovers code and data that can be ignored
+%during online analysis, thus reducing online instrumentation overheads. The key insight is that all training
+%data gathered is 
+%time--and
+%time spent--doing offline analysis does not affect the accuracy of the results produced. This is important
+%because discovering what code can be pruned and what data does
+
+%\begin{enumerate}
+%\item Detecting what code is important to the 
+%\end{enumerate}
+%
+%analyzing
+%the 
+%
+%the detection
+%of what code should be instrumented and what data should be sampled is \emph{not} time-sensitive, and
+%therefore can be performed any time. 
+%
+%using results gathered from
+%an offline learning and analysis phase. This separation is motivated by recognizing that
+
+%\item It presents two new techniques, \emph{proxy memory} and \emph{tagged reads}, that enable
+%efficient determination of memory access behavior using shadow memory. Unlike existing shadow
+%memory systems, proxy memory is not treated as a mutable meta-data store. This avoids extra
+%serialization and coherence traffic introduced by prior work when dynamically manipulating shadow state.
+
+%\item It describes two new, data-centric sampling approaches: uniform heap sampling, and selective heap
+%sampling. .
+%Once data of interesting has been narrowed down, \Toolname{} can apply selective heap sampling, which
+%focuses sampling on specific types of objects or data accessed by specific code (or both).
 \end{itemize}
@@
 
-\section{Related Work}\label{sec:background}
-\Toolname{} primarily draws on techniques from existing race detectors and false sharing detectors.
-This is appropriate because some forms of false-positives for race detectors are caused by false
-sharing. That is, where race detectors focus on two threads concurrently accessing the same
-data, false sharing detectors focus on two threads concurrently accessing different data, but
-on the same cache line.
+\section{Overview}\label{sec:overview}
 
-% --------------------------------------
-\subsection{Race Detection}
+\Toolname{} uses DBT to instrument the memory accesses of a program while it is running. During the offline
+learning phase, DBT is used to figure out what code is executed by more than one thread, what data is accessed
+by more than one thread, and what code accesses what data. During its online detection phase, \Toolname{} uses
+an implementation of unlimited watchpoints watch for accesses to a representative sample of cache lines, assign
+owning threads to those cache lines, and finally detect transfers in ownership.
 
-Race detectors can be categorized according to the following parameters:
+\subsection{Definitions}
+\Toolname{} looks for cache line contention bugs of the following forms:
 \begin{description}
-\item[Approach] Static or dynamic. \hfill \\
-Static race detectors analyze source code or binaries, typically performing interprocedural control-
-and data-flow analyses. The benefit of static analysis is that they can inspect all program paths.
+\item[Data races] Data races occur when the same program variable is concurrently modified by two threads, or 
+modified by one thread and concurrently accessed by another thread, where neither access is atomic (e.g. using
+and atomic read-modify-write instruction).
 
-Dynamic race detectors monitor memory accesses for races. The benefit of this approach is that
-races can be caught "in the act". The drawbacks of this approach is that most dynamic race detectors
-are too slow to run in production environments, and therefore must be tested on simulated workloads.
-If these workloads do not exercise all code paths than many races can be missed. Dynamic race detectors
-can be further broken down into on-the-fly detectors or post-mortem detectors, which analyze memory
-access logs.
+\item[True sharing] True sharing has the same access pattern as a data race, except that one or both operations
+are atomic.
 
-\item[Model] Precise or imprecise.  \hfill \\
-Precise race detectors report no false-positives and are usually implemented using Lamport's
-happens-before relationship \cite{VectorClocks}. Precise race detectors often trade efficiency
-for completeness in order to achieve their precision. That is, they sometime miss true races in
-code.
+\item[False sharing] False sharing occurs when two semantically related variables (e.g. two fields in a struct) occupy
+the same cache line and are concurrently accessed, thus introducing contention on the cache line.
 
-Imprecise race detectors can have high false positive rates, but tend to be more complete in terms of
-including races not detected by precise detectors. They tend to be implemented using a variant of the
-lockset algorithm \cite{Eraser}.
-
-\item[Granularity] Fine-grained or coarse-grained. \hfill \\
-Fine-grained race detectors operate at or close to the granularity of memory accesses. Coarse-grained
-race detectors operate at an object granularity, or at a fixed granularity (e.g. cache line, page).
-
-The granularity of a race detector can affect its precision. For example, a happens-before race detector 
-can be imprecise (i.e. report false positives) by operating at a coarse granularity. If the granularity were
-the size of a cache line, then the race detector would be a cache line contention detector: it would hint
-at races as well as false-sharing.
-
-\item[Mode] Comprehensive or sampling-based. \hfill \\
-Comprehensive race detectors typically analyze every memory read and write. They are usually implemented
-by instrumenting all program code.
-
-Sampling-based race detectors analyze either a subset of the code or a subset of the data. In the former
-case, they focus on detecting races by specific code.
+\item[Pseudo sharing] Pseudo sharing occurs when two semantically unrelated variables (e.g. fields in adjacent structs,
+adjacent array entries) occupy the same cache line and are concurrently accessed \cite{DegenerateSharingAndFalseCoherence}.
 \end{description}
 
-Under this categorization scheme, \Toolname{} is a dynamic, precise, variable-granularity, mixed-mode
-false-sharing detector. It draws inspiration from a number of existing race detectors, including LiteRace
-\cite{LiteRace}, MultiRace \cite{MultiRace}, RaceMob \cite{RaceMob}, RaceTrack \cite{RaceTrack}, and
-DataCollider \cite{DataCollider}.
+%\subsection{Mechanisms}
+
+% --------------------------------------------------------------------------------------
+
+\section{Related Work}\label{sec:background}
+\Toolname{} draws on and extends/refines techniques from existing race detectors and false sharing detectors.
+
+%% --------------------------------------
+\paragraph{Dr. Memory}
+Dr. Memory \cite{DrContention} implements a set of tools for detecting cache line contention, thread correlations,
+and true/false sharing detection in COTS binary programs. Dr. Memory uses the DynamoRIO \cite{DynamoRIO} DBT system
+and the Umbra shadow memory framework \cite{Umbra} to interpose on memory accesses and implement their
+detection tools. Dr. Memory's contention tool implements the following algorithm for detecting contention. Before each
+memory write,  instrumentation performs a ``racy" check to see if the current thread owns the cache line being accessed.
+If the cache line is owned by another thread, then a report is generated, and ownership is transferred. If the cache line is
+unowned, then ownership is established.
+
+This approach has two drawbacks. First, every time ownership is transferred/established, there is an additional
+memory write, which itself introduces CPU overheads due to contention on the cache lines of the shadow memory. Second,
+ownership is tracked on data that is never shared. This introduces extra CPU and memory overheads by instrumenting
+too much code and tracking too much data.
+
+\Toolname{} addresses both problems. First, its offline learning phase prunes the set of code that should be instrumented.
+This has the side effect of reducing the physical memory requirements of shadow memory because fewer instructions will
+introduce shadow accesses, and therefore less shadow is accessed. Second, our sampling approach means that ownership
+is only tracked for ``watched" objects. This introduces an additional racy read at the beginning of instrumentation, but the
+benefit is that the majority of shadow accesses are read-only, and therefore do not generate as much additional contention
+on the cache lines associated with shadow memory.
+
+%% --------------------------------------
+\paragraph{DataCollider}
+DataCollider \cite{DataCollider} is a precise, sampling-based race detector for the WIndows kernel. DataCollider's key
+novelties are its extremely low overhead, as well as its generality: it detects data races without any understanding of
+a system's underlying synchronization primitives. DataCollider uniformly samples code by introducing breakpoints into the code.
+When a breakpoint is hit, a hardware watchpoint is added to data that was about to be accessed by the interrupted instruction.
+Then, the interrupted thread is paused for a short period of time to give another thread the opportunity to access the
+same data. If the data is accessed by a second thread while the first thread is paused, then the watchpoint will detect the
+access, thus allowing DataCollider to detect a race ``in the act.""
+
+While this approach is elegant in its simplicity, it has two major drawbacks. First, its dependence on hardware watchpoints
+limits it to being able to detect races on at most four distinct 8-byte memory locations simultaneously \cite{IntelSDManual}.
+Second, it cannot detect guarded cache line line conflicts: if the same cache line is concurrently accessed by two threads, but
+the accesses are guarded by a lock, then the pausing introduced by the first thread will prevent the watchpoint from being hit
+in the second thread (because the second thread will block on the lock).
+
+%% --------------------------------------
+\paragraph{Other}
+
+Intel's VTune \cite{IntelVTune} an an imprecise, sampling-based profiler that can detect false sharing using hardware performance
+counters and monitoring events. It provides detailed, architecture-specific performance-tuning suggestions. Unlike \Toolname{},
+VTune can only report individual program counters that general cache coherency events, but cannot correlate threads by showing
+two contending stack traces.
+
+%, thus triggering the watchpoint and detecting a race ``in the act".
+%
+%%sampling based approach enables it to be general
+%
+% The key strengths of
+%DataCollider is that it has low overhead, it is independent of the synchronization mechanisms being employed, and that
+%it often catches both threads in the act.  This brilliantly simple approach was the main inspiration
+%for \Toolname{}. \Toolname{} differs from DataCollider, in that the latter is code-centric insofar as it samples code, whereas
+%\Toolname{} is data-centric insofar as it samples data.
+
+%In either case, a memory write occurs to take ownership.
+
+%\DrContention{} comprehensively interposes on
+%all memory accesses to track ownership of all cache lines. Every instruction is instrumented to first check if the current
+%thread owns the cache line. If not, then the thread is added to the cache line's \emph{threadset}.
+
+% TODO: Evaluate by disabling is_watched check, i.e. perform ownership tracking on all cache lines, vs. only a subset
+% of cache lines.
+
+%Several tracking mechanisms are discussed, including ones that
+%track strict ownership transers, and others that attempt to measure thread correlations and the impact of of cache line
+%invalidations. The drawback of this approach is that it necessitates checking and updating the shadow memory on every
+%memory access, which introduces performance overhead due to extra cache line contention.
+
+%Like \DrContention{}, \Toolname{} uses DBT and shadow memory to track ownership of cache lines. However, \Toolname{}
+%only tracks cache line ownership for a subset of the cache lines (those currently being sampled). This makes a big difference
+%because, although potentiall all instructions are instrumented for shadow memory, only a small subset of th
+
+\paragraph{Plastic}
+Plastic \cite{Plastic} is a Xen-based dynamic false sharing detector and repairer. Like \Toolname{}, it takes a pipeline-based
+approach that successively narrows down on the root causes of false sharing. Plastic also uses DBT as part of its
+repair mechanism. Plastic's pipeline begins by using hardware performance counters related to cache contention
+events to detect the presence of false-sharing. Then, it applies a similar technique to MultiRace \cite{MultiRace}, in
+that it tracks the dominant access pattern by each thread to memory on a shared page. This information informs
+their repair mechanism, which transparently re-maps memory accesses at a byte granularity.
+
+\paragraph{SHERIFF}
+SHERIFF \cite{SHERIFF} is a false-sharing detector and repairer that works in a similar way to MultiRace and Plastic. Initially,
+it applies read-only hardware page protection to all shared memory. Then, as threads attempt to write to protected pages, those
+pages are privately mapped as read/write. When the same page is privately mapped by two or more threads, SHERIFF applies
+the twinning and diffing approach of TreadMarks \cite{TreadMarks} to detect local changes to shared pages and narrow down
+on instances of false sharing. Like \Toolname{}, SHERIFF uses a coarse-grained mechanism to hint at the potential for races.
+
+%% --------------------------------------
+%\subsection{Race Detection}
+%
+%Race detectors can be categorized according to the following parameters:
+%\begin{description}
+%\item[Approach] Static or dynamic. \hfill \\
+%Static race detectors analyze source code or binaries, typically performing interprocedural control-
+%and data-flow analyses. The benefit of static analysis is that they can inspect all program paths.
+%
+%Dynamic race detectors monitor memory accesses for races. The benefit of this approach is that
+%races can be caught "in the act". The drawbacks of this approach is that most dynamic race detectors
+%are too slow to run in production environments, and therefore must be tested on simulated workloads.
+%If these workloads do not exercise all code paths than many races can be missed. Dynamic race detectors
+%can be further broken down into on-the-fly detectors or post-mortem detectors, which analyze memory
+%access logs.
+%
+%\item[Model] Precise or imprecise.  \hfill \\
+%Precise race detectors report no false-positives and are usually implemented using Lamport's
+%happens-before relationship \cite{VectorClocks}. Precise race detectors often trade efficiency
+%for completeness in order to achieve their precision. That is, they sometime miss true races in
+%code.
+%
+%Imprecise race detectors can have high false positive rates, but tend to be more complete in terms of
+%including races not detected by precise detectors. They tend to be implemented using a variant of the
+%lockset algorithm \cite{Eraser}.
+%
+%\item[Granularity] Fine-grained or coarse-grained. \hfill \\
+%Fine-grained race detectors operate at or close to the granularity of memory accesses. Coarse-grained
+%race detectors operate at an object granularity, or at a fixed granularity (e.g. cache line, page).
+%
+%The granularity of a race detector can affect its precision. For example, a happens-before race detector 
+%can be imprecise (i.e. report false positives) by operating at a coarse granularity. If the granularity were
+%the size of a cache line, then the race detector would be a cache line contention detector: it would hint
+%at races as well as false-sharing.
+%
+%\item[Mode] Comprehensive or sampling-based. \hfill \\
+%Comprehensive race detectors typically analyze every memory read and write. They are usually implemented
+%by instrumenting all program code.
+%
+%Sampling-based race detectors analyze either a subset of the code or a subset of the data. In the former
+%case, they focus on detecting races by specific code.
+%\end{description}
+
+%Under this categorization scheme, \Toolname{} is a dynamic, precise, variable-granularity, mixed-mode
+%false-sharing detector. It draws inspiration from a number of existing race detectors, including LiteRace
+%\cite{LiteRace}, MultiRace \cite{MultiRace}, RaceMob \cite{RaceMob}, RaceTrack \cite{RaceTrack}, and
+%DataCollider \cite{DataCollider}.
 
@@
 
-DataCollider \cite{DataCollider} is a dynamic, precise, sampling-based race detector for the kernel. The key strengths of
-DataCollider is that it has low overhead, it is independent of the synchronization mechanisms being employed, and that
-it often catches both threads in the act. DataCollider uniformly samples code by introducing breakpoints into the code.
-When a breakpoint is hit, a hardware watchpoint is added to data that was about to be accessed by the interrupted instruction.
-Then, the interrupted thread is paused for a short period of time to give another thread the opportunity to access the
-same data, thus triggering the watchpoint and detecting a race. This brilliantly simple approach was the main inspiration
-for \Toolname{}. \Toolname{} differs from DataCollider, in that the latter is code-centric insofar as it samples code, whereas
-\Toolname{} is data-centric insofar as it samples data.
+%% --------------------------------------
+%\subsection{False Sharing Detection}
+%
+%Similar to race detectors, false-sharing detectors can be sampling-based or comprehensive. Comprehensive approaches
+%can be categorized  into simulation-based and instrumentation-based approaches. Simulation-based approaches focus on
+%simulating the full cache hierarchy and can distinguish fine-grained events like false sharing due to set-associativity and
+%false sharing due to different data being accessed on the same cache line. Instrumentation-based approaches focus more
+%on mechanisms for estimating or detecting cache line contention.
 
-% --------------------------------------
-\subsection{False Sharing Detection}
+%\paragraph{Sampling-based}
 
-Similar to race detectors, false-sharing detectors can be sampling-based or comprehensive. Comprehensive approaches
-can be categorized  into simulation-based and instrumentation-based approaches. Simulation-based approaches focus on
-simulating the full cache hierarchy and can distinguish fine-grained events like false sharing due to set-associativity and
-false sharing due to different data being accessed on the same cache line. Instrumentation-based approaches focus more
-on mechanisms for estimating or detecting cache line contention.
-
-\paragraph{Sampling-based}
-Intel's VTune \cite{IntelVTune} an an imprecise, sampling-based profiler that can detect false sharing using hardware performance
-counters and monitoring events. It represents the state-of-the-art in performance analysis tools and is able to provide detailed,
-architecture-specific performance-tuning suggestions. Unlike \Toolname{}, VTune is unable to analyze only a portion of the code,
-or only the portion of code that accesses specific shared data structures.
-
-\paragraph{Comprehensive}
+%\paragraph{Comprehensive}
 
@@
 % TODO: Pluto?
-
-\newcommand{\DrContention}{Zhao \emph{et al.}}
-
-\DrContention{} \cite{DrContention} implement a dynamic false and true sharing detector using the DynamoRIO \cite{DynamoRIO}
-DBT system and Umbra shadow memory framework \cite{Umbra}. Unlike \Toolname{}, \DrContention{} use fixed-granularity
-shadow memory to track cache line ownership by processor cores. The drawback of this approach is that it necessitates checking
-and updating the shadow memory on every memory access, which introduces its own form of cache line contention. \Toolname{}
-avoids this issue by using shadow memory and tagged reads as a proxy for memory access behavior instead of as a meta-data store.
-
-Plastic \cite{Plastic} is a Xen-based dynamic false sharing detector and repairer. Like \Toolname{}, it takes a pipeline-based
-approach that successively narrows down on the root causes of false sharing. Plastic also uses DBT as part of its
-repair mechanism. Plastic's pipeline begins by using hardware performance counters related to cache contention
-events to detect the presence of false-sharing. Then, it applies a similar technique to MultiRace \cite{MultiRace}, in
-that it tracks the dominant access pattern by each thread to memory on a shared page. This information informs
-their repair mechanism, which transparently re-maps memory accesses at a byte granularity.
-
-SHERIFF \cite{SHERIFF} is a false-sharing detector and repairer that works in a similar way to MultiRace and Plastic. Initially,
-it applies read-only hardware page protection to all shared memory. Then, as threads attempt to write to protected pages, those
-pages are privately mapped as read/write. When the same page is privately mapped by two or more threads, SHERIFF applies
-the twinning and diffing approach of TreadMarks \cite{TreadMarks} to detect local changes to shared pages and narrow down
-on instances of false sharing. Like \Toolname{}, SHERIFF uses a coarse-grained mechanism to hint at the potential for races.
 

### 2014-12-15 HotOS'15/paper.tex
Working on the methodology paper. Having some issues on what the right amount of detail is for describing the implementation of ownership tracking. There are many valid designs, and there is a single correct design. The correct design is heavyweight, requiring a CMPXCHG. I chose one of the valid designs that seemed to have a reasonable state transition diagram.

@@
-%--------sigalternate----------
-\documentclass{sig-alternate}
-%--------/sigalternate----------
+\documentclass[letterpaper,twocolumn,10pt]{article}
 
-%\documentclass[letterpaper,twocolumn,10pt]{article}
+\usepackage[margin=1in]{geometry}
 %\usepackage{endnotes,multirow}
@@
 \usepackage{paralist} % for in-paragraph lists
+\usepackage{parskip} % For making paragraphs split by new lines instead of indents
 \usepackage{graphicx}
 \usepackage{hyperref}
+
+% Used for changing the spacing after the title, sections, and subsections.
+\usepackage{titlesec,titling}
 
@@
 
+\lstdefinelanguage
+   [x64]{Assembler}     % add a "x64" dialect of Assembler
+   [x86masm]{Assembler} % based on the "x86masm" dialect
+   % with these extra keywords:
+   {morekeywords={LOCK,XCHG,CMP,JZ,JNZ,CALL, %
+                  rax,rdx,rcx,rbx,rsi,rdi,rsp,rbp, %
+                  r8,r8d,r8w,r8b,r9,r9d,r9w,r9b}} % etc.
+
 % Modify spacing before/after title, sections, and subsections.
@@
 \newcommand{\Sref}[1]{Section~\ref{#1}}
-
-%--------sigalternate----------\titlespacing\section{0pt}{5pt plus 4pt minus 2pt}{5pt plus 2pt minus 2pt}        
-%--------sigalternate----------\titlespacing\subsection{0pt}{5pt plus 4pt minus 2pt}{5pt plus 2pt minus 2pt}
 
@@
 \newcommand{\comment}[1]{}
-\newcommand{\TextToolname}{CatcheScan}
+\newcommand{\TextToolname}{Malcontent}
 \newcommand{\Toolname}{\textsc{\TextToolname{}}}
@@
 
+% Modify spacing before the title
+\setlength{\droptitle}{-3em}
+
 \begin{document}
+
+%don't want date printed
+\date{}
 
@@
 \title{\Large \bf \Toolname: Cache Line Contention Detection for COTS Binaries}
-\numberofauthors{3}
 \author{
-% 1st. author
-\alignauthor
-Peter Goodman\\
-       \email{pag@cs.toronto.edu}
-\and
-% 2nd. author
-\alignauthor
-Ashvin Goel\\
-       \email{asvhin@eecg.toronto.edu}
-% 3rd. author
-\alignauthor
-Angela Demke Brown\\
-       \email{demke@cs.toronto.edu}
-}
+{\rm Peter Goodman} \hspace{1.5em} {\rm Angela Demke Brown} \hspace{1.5em} {\rm Ashvin Goel}\\
+University of Toronto
+} % end author
 \maketitle
@@
 %\end{abstract}
+
+\subsection*{Abstract}
+Cache-line contention is an increasingly important problem in multi-threaded applications running on cache coherent SMPs.
+Contention on cache lines is symptomatic of both performance and correctness bugs. In terms of performance, cache line
+contention limits scalability by introducing extra latency when accessing shared memory.  In terms of correctness, cache
+line contention can also be caused by data races, which introduce non-determinism into one's program.
+
+In this paper, we describe \Toolname, a system that detects and diagnoses cache line contention in arbitrary binary programs.
+\Toolname{} refines existing contention detection approaches in several novel ways. First, we describe a method training
+a detection system to focus on only the code and data that matter to the problem. Second, we describe a new form of
+data-centric sampling: comprehensive data sampling. Our sampling approach uses unlimited watchpoints to uniformly
+sample memory locations according to the type of data stored in memory. This avoids sampling bias where a large portion
+of a program's memory is unshared.
 
@@
 of our approach is that if we know \begin{inparaenum}[i)]
-\item what code is executed by more than one thread,
 \item what types of data are accessed by more than one thread, and
@@
 
-\paragraph{Contributions}
+% --------------------------------------
+\subsection{Contributions}
 This paper makes the following significant contributions:
@@
 %unshared data
-
 %This is beneficial because it improves coverage: if 90\% of false sharing occurs because of accesses
@@
 
-\item It describes a novel separation of cache line contention detection into two phases: \begin{enumerate}
-\item[1) \emph{Training:}] an ahead-of-time, heavyweight analysis of the program's execution that
+\item It describes a novel separation of cache line contention detection into two phases:
+
+1. {\bf Training:} an ahead-of-time, heavyweight analysis of the program's execution that
 discovers what code can contain cache line contention, and what data is shared by multiple threads. A key
@@
 
-\item[2) \emph{Detection:}] an online analysis that detects cache line contention by selectively
+2. {\bf Detection:} an online analysis that detects cache line contention by selectively
 instrumenting only the relevant code and sampling only the relevant data as recognized by the
 offline training phase.
-\end{enumerate}
 
@@
 %
-%
-%
-%\item[2) \emph{Detection:}] b
+%\item[2) \emph{Detection:}] 
 %\end{enumerate}
@@
 
-%\begin{enumerate}
-%\item Detecting what code is important to the 
-%\end{enumerate}
-%
-%analyzing
-%the 
-%
-%the detection
-%of what code should be instrumented and what data should be sampled is \emph{not} time-sensitive, and
+%the detection of what code should be instrumented and what data should be sampled is \emph{not} time-sensitive, and
 %therefore can be performed any time. 
 %
-%using results gathered from
-%an offline learning and analysis phase. This separation is motivated by recognizing that
+%using results gathered from an offline learning and analysis phase. This separation is motivated by recognizing that
 
@@
 
-\Toolname{} uses DBT to instrument the memory accesses of a program while it is running. During the offline
-learning phase, DBT is used to figure out what code is executed by more than one thread, what data is accessed
-by more than one thread, and what code accesses what data. During its online detection phase, \Toolname{} uses
-an implementation of unlimited watchpoints watch for accesses to a representative sample of cache lines, assign
-owning threads to those cache lines, and finally detect transfers in ownership.
+\Toolname{} uses the Granary DBT \cite{Granary} framework to instrument the memory accesses of a program while
+it is running. During the offline training phase, DBT is used to figure out what data is accessed by more than one thread,
+and what code accesses that data. During its online detection phase, \Toolname{} uses Granary's implementation of unlimited
+watchpoints to watch for accesses to a representative sample of cache lines, assign owning threads to those cache lines,
+and finally detect transfers in ownership.
 
@@
 \Toolname{} looks for cache line contention bugs of the following forms:
-\begin{description}
-\item[Data races] Data races occur when the same program variable is concurrently modified by two threads, or 
-modified by one thread and concurrently accessed by another thread, where neither access is atomic (e.g. using
+
+{\bf Data races} occur when the same program variable is concurrently modified by two threads, or 
+modified by one thread and concurrently accessed by another thread, and where neither access is atomic (e.g. using
 and atomic read-modify-write instruction).
 
-\item[True sharing] True sharing has the same access pattern as a data race, except that one or both operations
+{\bf True sharing} has the same access pattern as a data race, except that one or both operations
 are atomic.
 
-\item[False sharing] False sharing occurs when two semantically related variables (e.g. two fields in a struct) occupy
+{\bf False sharing} occurs when two semantically related variables (e.g. two fields in a struct) occupy
 the same cache line and are concurrently accessed, thus introducing contention on the cache line.
 
-\item[Pseudo sharing] Pseudo sharing occurs when two semantically unrelated variables (e.g. fields in adjacent structs,
+{\bf Pseudo sharing} occurs when two semantically unrelated variables (e.g. fields in adjacent structs,
 adjacent array entries) occupy the same cache line and are concurrently accessed \cite{DegenerateSharingAndFalseCoherence}.
-\end{description}
 
-%\subsection{Mechanisms}
+% --------------------------------------
+\subsection{Mechanisms}
+
+% --------------------------------------
+\paragraph{Training}
+%\paragraph{Code Concurrency Analysis }
+%\Toolname{} analyzes code to determine what code is executed by more than one thread. This analysis allows \Toolname{}
+%to either not instrument, or only very lightly instrument, non-concurrent code during its detection phase. The insight is
+%that code that does get executed by more than one thread is more likely to operate on shared data than code that is
+%only ever observed (during training) to have been executed by a single thread.
+%
+%There are some important cases where this insight does not hold. For example, in a single writer, multiple reader queue,
+%the code implementing writes to the queue might only be executed by a designated writer thread, and therefore never
+%be marked as concurrent. Obviously, there is contention bet
+
+%\paragraph{Data Concurrency Analysis}
+\Toolname's training phase focuses on reducing detection overheads by finding code that doesn't need to be instrumented
+during the detection phase. The key challenge of this phase is determining what code can safely be pruned without negatively
+affecting the accuracy of the detection phase. Our insight is that the detection phase should only look for cache contention
+on accesses to shared data structures. To do this, we must know \begin{inparaenum}
+\item what data structures are shared, and
+\item what code accesses those data structures.
+\end{inparaenum}
+
+\newcommand\TypeId{$TypeID$}
+\newcommand\TypeIdPair{$\langle PC,size \rangle$ }
+
+We implemented these analyses using Granary's address watchpoints \cite{AddressWatchpoints} system, which allows us to
+``taint" addresses and then interpose on instructions using tainted addresses to access memory. Using Granary, we interpose on
+all dynamic memory allocator functions and taint every address returned by those functions with an abstract \TypeId{}.
+Each \TypeId{} uniquely identifies an allocation context in terms of a \TypeIdPair{} pair, where $PC$ is the program counter
+to which the allocation function will return, and $size$ is the number of bytes being allocated. This arrangement allows us to
+approximate the actual number of distinct types of program data structures for arbitrary binaries, even in the absence of
+static source code information or debug information.
+
+We use per-type shadow memory to detect data structure sharing accross multiple threads.
+We associate each \TypeId{} with a small amount of shadow memory: $\langle InitialThreadId,\linebreak[0]IsShared \rangle$. This
+arrangement means that every allocated address tainted by the same \TypeId{} is indirectly associated with the same
+shadow memory. If this is the first allocation for a \TypeId, then the shadow memory is initialized to
+$\langle CurrThreadId,\linebreak[0]false \rangle$. Sharing is detected and marked either by subsequent allocations for the
+same \TypeId{} but by different threads than the initial thread, or by accesses to tainted addresses by threads other
+than the initial thread.
+
+To detect what code accesses what types of data, we propagate the {\TypeId}s to code on every memory access to
+a tainted address. Finally, to determine what code accesses shared data, we combine the output of both analyses to
+produce, for each basic block of code executed during training, we report on whether or not it contained accesses
+to shared data structures. During the detection phase, we heavily instrument only those blocks that contain accesses
+to shared data structures, or (to be conservative) were never executed during the training phase.
+
+%taint from the address to the code accessing the memory associated with tainted address.
+%This allows us to identify what code is accessing what types of data.
+%Later allocations do not modify the shadow, unless the thread performing the
+%allocation does not 
+%When an
+%object is allocated, its address is tainted with \TypeId, and, if this is a new \TypeId, then the sha
+%and which is initialized 
+%Every time a tainted address is accessed, we perform the following: \begin{enumerate}
+%\item Propagate 
+%
+%\item Check if the current thread matches the $AllocThread$. If not, then set $IsShared \gets true$. Future accesses to an
+%object with $IsShared = true$ d
+%
+%\end{enumerate}
+% identifier that is
+%uniquely associated with the allocation context 
+%associated with the allocation context, as represented by a \TypeIdPair{} pair.
+%For each dynamically allocated data structure, we ``taint'' the address of the 
+%Pruning code 
+%This analysis determines what types of data are shared, i.e. accessed by more than one thread, so that the 
+%\Toolname{}'s training phase analysis data access patterns to determine what types of data are accessed by more than one
+%thread. The insight is that we should only look
+%\paragraph{Code Polymorphism Analysis}
+
+% --------------------------------------
+\paragraph{Detection}
+
+The detection phase is composed of three components:
+
+1. {\bf The interposition layer}. We interpose on dynamic memory allocators and record the most recently allocated
+address for each \TypeId{}. The set of all most recently allocated objects forms the population of addresses from which
+we sample.
+
+As a minor optimization, we restrict ourselves to only those {\TypeId}s associated that were found during training
+to be associated with shared data structures. Also, unlike during the training phase, we \emph{do not} taint any
+addresses.
+
+2.  {\bf The monitor}. The monitor is a separate thread, spawned by \Toolname{} at program startup, and is responsible
+for implementing \Toolname's \emph{comprehensive data sampling}. The monitor thread periodically
+wakes up, chooses a sample set using a round-robin algorithm over the population of recently allocated addresses by type, and
+then enables software watchpoints on all sampled addresses.
+
+After enabling the watchpoints, the monitor goes to sleep. When it next wakes up, it removes any active watchpoints
+and generates reports on any cache line contention that was detected. Finally, it restarts the sampling process.
+
+3. {\bf The instrumentation}. There are two tiers of instrumentation: lightweight and heavyweight. Lightweight
+instrumentation applies to all basic blocks of code identified during training as \emph{not} containing accesses to shared
+data. Lightweight instrumentation exists to maintain Granary's control over the program binary, and therefore to maintain
+the comprehensiveness of the analysis.
+
+The heavyweight instrumentation implements a form of ``unlimited watchpoints" \cite{UnlimitedWatchpoints}. The
+granularity of each watchpoint is a cache line, but can easily be adjusted in order to model systems with smaller or
+larger cache line sizes\footnote{The granularity of shadow memory is tunable via a command-line argument}.
+In our setup, every 64 bytes of memory has its own shadow memory and watchpoint. Note that
+this does not model actual hardware limits on cache sizes and cache contention caused by associativity.
+
+When a memory address is accessed, the instrumentation checks the shadow memory to see if the watchpoint for
+the current address is set. If it is, then the instrumentation code further checks to shadow to see if another thread
+owns the cache line (in which case contention is detected), or if the current thread can take ownership of the cache line.
 
@@
 \paragraph{DataCollider}
-DataCollider \cite{DataCollider} is a precise, sampling-based race detector for the WIndows kernel. DataCollider's key
+DataCollider \cite{DataCollider} is a precise, sampling-basbased oned race detector for the WIndows kernel. DataCollider's key
 novelties are its extremely low overhead, as well as its generality: it detects data races without any understanding of
@@
 limits it to being able to detect races on at most four distinct 8-byte memory locations simultaneously \cite{IntelSDManual}.
-Second, it cannot detect guarded cache line line conflicts: if the same cache line is concurrently accessed by two threads, but
-the accesses are guarded by a lock, then the pausing introduced by the first thread will prevent the watchpoint from being hit
-in the second thread (because the second thread will block on the lock).
+Second, in some cases, it cannot be generalized to detecting arbitrary cache contention when the one contending access
+\emph{happens-before} the other \cite{VectorClocks}. For example, if true/false sharing frequently occurs within critical sections;
+however, this would be missed by a naiive generalization of the DataCollider algorithm to this problem.
+
+%% --------------------------------------
+%\paragraph{Predator}
+
+% TODO: I should really write about this, but time :-(
 
@@
 
-Intel's VTune \cite{IntelVTune} an an imprecise, sampling-based profiler that can detect false sharing using hardware performance
-counters and monitoring events. It provides detailed, architecture-specific performance-tuning suggestions. Unlike \Toolname{},
-VTune can only report individual program counters that general cache coherency events, but cannot correlate threads by showing
-two contending stack traces.
+%Intel's VTune \cite{IntelVTune} an an imprecise, sampling-based profiler that can detect false sharing using hardware performance
+%counters and monitoring events. It provides detailed, architecture-specific performance-tuning suggestions. Unlike \Toolname{},
+%VTune can only report individual program counters that general cache coherency events, but cannot correlate threads by showing
+%two contending stack traces.
 
@@
 
-\paragraph{Plastic}
-Plastic \cite{Plastic} is a Xen-based dynamic false sharing detector and repairer. Like \Toolname{}, it takes a pipeline-based
-approach that successively narrows down on the root causes of false sharing. Plastic also uses DBT as part of its
-repair mechanism. Plastic's pipeline begins by using hardware performance counters related to cache contention
-events to detect the presence of false-sharing. Then, it applies a similar technique to MultiRace \cite{MultiRace}, in
-that it tracks the dominant access pattern by each thread to memory on a shared page. This information informs
-their repair mechanism, which transparently re-maps memory accesses at a byte granularity.
+Plastic \cite{Plastic} is a Xen-based dynamic false sharing detector and repairer. Plastic uses hardware performance counters
+to detect when cache line eviction events due to contention exceed some threshold, and then turns on page-granularity
+ownership tracking using hardware page protection. When Plastic discovers sharing of a page, it turns on heavyweight recording
+of memory access instructions. Later, it analyzes this record to determine what--if any--cache lines on the page had contention,
+and then uses a limited form of DBT to ``patch" the virtualized execution so as to remap previously contended accesses into
+non-contending accesses. Plastic's dependence on virtualization limit its ability to find false sharing in arbitrary, binary device
+drivers that cannot be virtualized.
 
-\paragraph{SHERIFF}
-SHERIFF \cite{SHERIFF} is a false-sharing detector and repairer that works in a similar way to MultiRace and Plastic. Initially,
-it applies read-only hardware page protection to all shared memory. Then, as threads attempt to write to protected pages, those
-pages are privately mapped as read/write. When the same page is privately mapped by two or more threads, SHERIFF applies
-the twinning and diffing approach of TreadMarks \cite{TreadMarks} to detect local changes to shared pages and narrow down
-on instances of false sharing. Like \Toolname{}, SHERIFF uses a coarse-grained mechanism to hint at the potential for races.
+%
+%SHERIFF \cite{SHERIFF} is a precise false-sharing detector and repairer that works in a similar way to Plastic. Instead of
+%tracking page ownership at the VMM level, it tracks page ownership on a per-application basis by turning threads into
+%processes, thus allowing it to bootstrap off of existing OS-provided process memory isolation. SHERIFF detects contention
+%using twinning and diffing of shared pages \cite{TreadMarks}, and fixes false-sharing using a combination of delaying and
+
+Eraser \cite{Eraser} is a dynamic race detector that uses shadow memory to track the set of locks that must be held in order
+to access shared data. Eraser initially associates the set of all locks with every shared variable. As shared variables are accessed,
+their lock sets are refined based on observing the current set of locks held by a particular thread. If a shared variable's lockset
+is refined to the empty set, then a data race has potentially been discovered. The benefit of this approach is that it has the
+ability to predict data races, even if two accesses were never truly concurrent. The drawback is that a lockset approach is
+challenging to generalize: it requires a complete understanding of a program's locking discipline.
+
+Other work based on Lamport's \emph{happens-before} relationship \cite{VectorClocks} can precisely distinguish data races
+from true sharing by recognizing that synchronizing operations (such as locks and condition variables) introduce a strict ordering of
+accesses. Under this lens, true sharing occurs when two threads access the same variable, but where one of the accesses happens-%
+before the other due to the presence of a synchronizing operation. A data race occurs where neither operation logically happens-%
+before the other. Like Eraser, generalizing happens-before to arbitrary COTS binaries can be challenging as it requires an
+understanding of the sycnrhonization primitives used by a program.
+
+%Initially,
+%it applies read-only hardware page protection to all shared memory. Then, as threads attempt to write to protected pages, those
+%pages are privately mapped as read/write. When the same page is privately mapped by two or more threads, SHERIFF applies
+%the twinning and diffing approach of TreadMarks  to detect local changes to shared pages and narrow down
+%on instances of false sharing. Like \Toolname{}, SHERIFF uses a coarse-grained mechanism to hint at the potential for races.
+
+%
+%Like \Toolname{}, it takes a pipeline-based
+%approach that successively narrows down on the root causes of false sharing. Plastic also uses DBT as part of its
+%repair mechanism. Plastic's pipeline begins by using hardware performance counters related to cache contention
+%events to detect the presence of false-sharing. Then, it applies a similar technique to MultiRace \cite{MultiRace}, in
+%that it tracks the dominant access pattern by each thread to memory on a shared page. This information informs
+%their repair mechanism, which transparently re-maps memory accesses at a byte granularity.
+
+%
+%SHERIFF \cite{SHERIFF} is a false-sharing detector and repairer that works in a similar way to MultiRace and Plastic. Initially,
+%it applies read-only hardware page protection to all shared memory. Then, as threads attempt to write to protected pages, those
+%pages are privately mapped as read/write. When the same page is privately mapped by two or more threads, SHERIFF applies
+%the twinning and diffing approach of TreadMarks \cite{TreadMarks} to detect local changes to shared pages and narrow down
+%on instances of false sharing. Like \Toolname{}, SHERIFF uses a coarse-grained mechanism to hint at the potential for races.
 
@@
 %DataCollider \cite{DataCollider}.
+%
+%LiteRace \cite{LiteRace} is a precise, sampling-based dynamic race detector implemented on top of the Microsoft
+%Pheonix binary translator. It samples at a function granularity, and executes either one specialized version of the
+%code or another. The sampling rate for a function is inversely proportional to how often it executes. This is based
+%on the insight that races are likely to occur in infrequently executed code, because otherwise they are benign or
+%would have been noticed and resolved. This insight, called the ``cold code hypothesis", motivated \Toolname's
+%post-mortem ranking scheme: cases of false sharing are more highly ranked when one---but \emph{not} both---of
+%the participating threads are executing cold code. This variant on the cold code hypothesis emphasizes the similarly
+%hard-to-find case of cold code pessimizing hot code.
 
-LiteRace \cite{LiteRace} is a precise, sampling-based dynamic race detector implemented on top of the Microsoft
-Pheonix binary translator. It samples at a function granularity, and executes either one specialized version of the
-code or another. The sampling rate for a function is inversely proportional to how often it executes. This is based
-on the insight that races are likely to occur in infrequently executed code, because otherwise they are benign or
-would have been noticed and resolved. This insight, called the ``cold code hypothesis", motivated \Toolname's
-post-mortem ranking scheme: cases of false sharing are more highly ranked when one---but \emph{not} both---of
-the participating threads are executing cold code. This variant on the cold code hypothesis emphasizes the similarly
-hard-to-find case of cold code pessimizing hot code.
-
-\newcommand\TypeIdPair{$\langle PC,size \rangle$ }
-
-MultiRace \cite{MultiRace} is a comprehensive, dynamic race detector that employs a hybrid happens-before and
-lockset-based approach. It operates at an object granularity, and therefore reports false-positives. It is unique
-in that it uses ``pointer swizzing" as a means of remapping memory to ``minipages" that track the dominating
-access pattern of a thread (read-write > read-only > no-access). \Toolname{} uses a similar pointer swizzing approach, but
-in the form of address watchpoints \cite{AddressWatchpoints}. Pointers to heap-allocated objects are ``swizzled" to
-include a type identifier that is unique to a \TypeIdPair{} pair. This improves the diagnosis of false-sharing,
-because it allows \Toolname{} to distinguish between two threads accessing different fields of the same object, and
-two threads accessing two nearby objects of the same (or different) types.
-
-RaceMob \cite{RaceMob} and RaceTrack \cite{RaceTrack} are two hybrid happens-before and lockset-based race detectors.
-RaceMob begins by statically analyzing source code to find potential races (lockset), then dynamically verifies those races
-(happens-before). RaceTrack is a JIT-based race-detector for Microsoft's common language runtime (CLR). It uses the
-concept of threadsets to efficiently maintain happens-before relationships for memory locations and detect the possible
-presence of a race and uses lockets to confirm predicted races. The novelty of RaceTrack is that it adapts the detection
-granularity from object-granularity to field-granularity. \Toolname{} follows the same high-level approach as both RaceMob
-and RaceTrack in that it attempts to predict the source of false-sharing at one granularity, then verify those predictions 
-at another granularity.
+%
+%MultiRace \cite{MultiRace} is a comprehensive, dynamic race detector that employs a hybrid happens-before and
+%lockset-based approach. It operates at an object granularity, and therefore reports false-positives. It is unique
+%in that it uses ``pointer swizzing" as a means of remapping memory to ``minipages" that track the dominating
+%access pattern of a thread (read-write > read-only > no-access). \Toolname{} uses a similar pointer swizzing approach, but
+%in the form of address watchpoints \cite{AddressWatchpoints}. Pointers to heap-allocated objects are ``swizzled" to
+%include a type identifier that is unique to a \TypeIdPair{} pair. This improves the diagnosis of false-sharing,
+%because it allows \Toolname{} to distinguish between two threads accessing different fields of the same object, and
+%two threads accessing two nearby objects of the same (or different) types.
+%
+%RaceMob \cite{RaceMob} and RaceTrack \cite{RaceTrack} are two hybrid happens-before and lockset-based race detectors.
+%RaceMob begins by statically analyzing source code to find potential races (lockset), then dynamically verifies those races
+%(happens-before). RaceTrack is a JIT-based race-detector for Microsoft's common language runtime (CLR). It uses the
+%concept of threadsets to efficiently maintain happens-before relationships for memory locations and detect the possible
+%presence of a race and uses lockets to confirm predicted races. The novelty of RaceTrack is that it adapts the detection
+%granularity from object-granularity to field-granularity. \Toolname{} follows the same high-level approach as both RaceMob
+%and RaceTrack in that it attempts to predict the source of false-sharing at one granularity, then verify those predictions 
+%at another granularity.
 
@@
 %\paragraph{Sampling-based}
-
 %\paragraph{Comprehensive}
@@
 
-This section introduces our approach to finding false sharing bugs. The goal of our approach is extract beliefs about
-where false sharing might be, then iteratively strenghten those beliefs, either verifying or discarding them in the process.
+\Toolname{} can be thought of as combining DataCollider's sampling based approach \cite{DataCollider}, and Dr. Memory's
+cache line ownership tracking approach \cite{DrContention}. The major distinction between \Toolname, DataCollider, and
+Dr. Memory are that:
+\begin{itemize}
+\item \Toolname's sampling approach is data-centric, as opposed to DataCollider's, which is code-centric. That is, DataCollider
+interposes on code first in order to find data that should be watched. \Toolname{} starts with the data that it expects to
+be shared (based on training), and then waits for code to access that data.
 
-We begin with the same intuition as SHERIFF: if false sharing occurs on a single cache line, then that implies that two
-threads are simultaneously accessing the same page of memory \cite{SHERIFF}. Therefore, by detecting contention on
-a single page, we can establish a belief that false sharing might occur on the contended page.
-
-Unfortunately, a page-granularity detection of false sharing is na{\"i}ve: it provides no guidance as to the location of the
-contended cache line, if any. Our insight is that we can narrow down on instances of false sharing by varying the
-granularity of the mechanism used to detect thread contention. That is, we can start with a coarse-grained but
-innacurate approach that tells us what code might cause false sharing, thus establishing an initial belief set. Then, we
-can strengthen or discard beliefs by repeating this same process, but using finer-grained detection mechanisms that
-apply only to code thought to contain false sharing.
-
-The remainder of this section explains how we proxy memory access behavior (\Secref{proxymem}), how we use sampling
-to detect contending accesses to shared memory (\Secref{heapsample}), how we reduce false negatives with our
-sampling-based approach (\Secref{falsenegs}), how we achieve precision by eliminating false positives (\Secref{falsepos}),
-and finally how we rank discovered instances of false sharing (\Secref{ranking})
+\item Unlike Dr. Memory, \Toolname{} does not heavily instrument all code for cache line ownership tracking, nor does it
+persistently track the ownership of all cache lines. Instead, the training phase informs the detection phase on what
+instructions to instrument, and sampling turns on temporary tracking of a subset of cache lines. When the \emph{first}
+instance of contention is detected, then the issue is reported and tracking is disabled.
+\end{itemize}
 
 % --------------------------------------
-\subsection{Proxy Memory and Tagged Reads}\label{sec:proxymem}
-Proxy memory is a form of variable-granularity shadow memory that proxies actual memory access behavior. The key
-idea is to instrument all loads and stores in such a way that their load/store behavior is reflected in the shadow
-memory in a way that can be externally observed. For example, one way to achieve this is to instrument every load
-to read from shadow memory, every store to modify shadow memory, and trap loads and stores by adding hardware
-watchpoints to shadow memory locations. Such heavyweight instrumentation would be inefficient, and so we apply
-two optimizations: belief-based selective instrumentation, and tagged reads.
+\subsection{Ownership Tracking}
+\begin{figure}
+\begin{center}
+\epsfig{file=states.eps, height=30em}
+\end{center}
+\caption{\label{fig:states_of_shadow}States of individual shadow memory cells. Each cell tracks the watched and ownership %
+status of a cache line. Yellow states are unwatched. Green states are watched. Red states represent contention that was missed %
+due to races in the instrumentation. Transitions between states are based on (potentially stale) observed prior states. For %
+example, the spurious acquire transition is possible because the decision to make the transition is based on a racy \texttt{CMP} %
+against the $SampleId$ (\figref{inline_assembly}). If two threads simultaneously see that the cache line it watched but not %
+owned, then both will execute the atomic \texttt{XCHG}, and only one thread will ``win". The ignore transition is induced, %
+either because the sampling timeout expires, or more likely, because the report transition occurs.}
+\end{figure}
 
-\paragraph{Selective Instrumentation}
-The first optimization to proxy memory is to not instrument all code to operate on proxy memory. Initially, \Toolname{}
-comprehensively instruments all code, thus incurring significant overheads. However, as beliefs about false sharing
-are refined, less and less code is instrumented, thus allowing code to run faster when precision during identification of
-potential false sharing is needed.
+\begin{figure}
+\lstset{language=[x64]Assembler}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+;; Racy watchpoint check.
+CMP [ShadowAddr + Offs(SampleId)], 0
+JZ ignore
+;; Racy ownership check.
+CMP [ShadowAddr + Offs(ThreadId)], CurrThreadId
+JZ ignore
+;; Take ownership, or transfer ownership
+LOCK XCHG [ShadowAddr], OldState
+CALL CheckContention(OldState,Addr,PC)
+ignore:
+\end{lstlisting}
+\caption{\label{fig:inline_assembly}Instrumentation instructions injected for each shared memory access. The algorithm optimizes on %
+instruction footprint by using a single \texttt{XCHG} instruction to perform multiple state transitions from %
+\Figref{states_of_shadow}, and then figuring out what state we should be in based on the value of $OldState$.}
+\end{figure}
 
-\paragraph{Tagged Reads}
-The second optimization avoids the problem of introducing unnecessary cache coherency traffic by treating proxy memory
-as read-only, and disambiguating the intention of an operation (load, store) by using a \emph{tagged read}. In our implementation,
-a read operations (on the proxy memory) that intend to communicate loads (of actual memory) are distinguished from
-read operations that intend to communicate a store by using different encodings of the logical \texttt{OR} instruction.
+Initially, a every cache line is unwatched, as represented by the absence of a $SampleId$ in \Figref{states_of_shadow}, and by the value
+\texttt{0} in shadow memory. When a cache line is sampled by the monitor thread, it is assigned a unique 16-bit $SampleId$. $SampleId$s
+are distinct from $TypeId$ because multiple samples for a given type might be taken. This happens when objects of a type span multiple
+cache lines, and when more samples than types are requested\footnote{The number of samples to take per sampling period is tunable %
+via a command-line argument.}. The distinction between $SampleId$ and $TypeId$ is primarily useful for reporting: at the program's
+initialization, each $SampleId$ is allocated space to describe a detected case of cache line contention. The description includes information
+about what types of memory was being sampled, what offsets into that memory were being sampled, stack traces for the contending
+threads, addresses accessed by each thread, and details about each thread's memory access type (read/write, atomic, width).
+
+The limitation to 16 bits for $SampleId$s was a deliberate decision. We take advantage of the structure of ``canonical addresses" on the
+x86-64 architecture, whereby the high-order 16 bits of a 64-bit address must have the same value as the 47th bit. In user space, all 16
+high-order bits are 0. We represent the $ThreadId$ using the thread's ``base pointer'' (a pointer thread-local storage). Thus, the
+\texttt{XCHG} in \Figref{inline_assembly} has the side-effect of temporarily disabling the watchpoint.
+
+% TODO: describe why this is reasonable, what alternatives there are, and the complications with those alternatives (atomicity,
+% re-watching, etc)
+
+% TODO: Should we have done it where we read the value of shadow memory into a reg, then used that as the old state to pass on
+% (plus the old value of the thread pointer)? This is valid, and might be better. But this has weird transitions.
+
+% Essentially: the "right" solution using a CMPXCHG, but that requires a lot of spills/fills of regs  === heavyweight.
+
+%, the shadow is initialized with a $SampleId$. The first read or write to a cache
+
+\section{Optimization}\label{sec:optimizations}
+
+% TODO: Make the code accessing data context-sensitive. I.e. so that we can avoid cases like memset always
+% being heavily instrumented.
+
+% TODO: Have a third tier of instrumentation, i.e. going native on tail functions that don't access shared data,
+% Or on chains of such functions.
+
+%
+%This section introduces our approach to finding false sharing bugs. The goal of our approach is extract beliefs about
+%where false sharing might be, then iteratively strenghten those beliefs, either verifying or discarding them in the process.
+%
+%We begin with the same intuition as SHERIFF: if false sharing occurs on a single cache line, then that implies that two
+%threads are simultaneously accessing the same page of memory \cite{SHERIFF}. Therefore, by detecting contention on
+%a single page, we can establish a belief that false sharing might occur on the contended page.
+%
+%Unfortunately, a page-granularity detection of false sharing is na{\"i}ve: it provides no guidance as to the location of the
+%contended cache line, if any. Our insight is that we can narrow down on instances of false sharing by varying the
+%granularity of the mechanism used to detect thread contention. That is, we can start with a coarse-grained but
+%innacurate approach that tells us what code might cause false sharing, thus establishing an initial belief set. Then, we
+%can strengthen or discard beliefs by repeating this same process, but using finer-grained detection mechanisms that
+%apply only to code thought to contain false sharing.
+%
+%The remainder of this section explains how we proxy memory access behavior (\Secref{proxymem}), how we use sampling
+%to detect contending accesses to shared memory (\Secref{heapsample}), how we reduce false negatives with our
+%sampling-based approach (\Secref{falsenegs}), how we achieve precision by eliminating false positives (\Secref{falsepos}),
+%and finally how we rank discovered instances of false sharing (\Secref{ranking})
+%
+%% --------------------------------------
+%\subsection{Proxy Memory and Tagged Reads}\label{sec:proxymem}
+%Proxy memory is a form of variable-granularity shadow memory that proxies actual memory access behavior. The key
+%idea is to instrument all loads and stores in such a way that their load/store behavior is reflected in the shadow
+%memory in a way that can be externally observed. For example, one way to achieve this is to instrument every load
+%to read from shadow memory, every store to modify shadow memory, and trap loads and stores by adding hardware
+%watchpoints to shadow memory locations. Such heavyweight instrumentation would be inefficient, and so we apply
+%two optimizations: belief-based selective instrumentation, and tagged reads.
+%
+%\paragraph{Selective Instrumentation}
+%The first optimization to proxy memory is to not instrument all code to operate on proxy memory. Initially, \Toolname{}
+%comprehensively instruments all code, thus incurring significant overheads. However, as beliefs about false sharing
+%are refined, less and less code is instrumented, thus allowing code to run faster when precision during identification of
+%potential false sharing is needed.
+%
+%\paragraph{Tagged Reads}
+%The second optimization avoids the problem of introducing unnecessary cache coherency traffic by treating proxy memory
+%as read-only, and disambiguating the intention of an operation (load, store) by using a \emph{tagged read}. In our implementation,
+%a read operations (on the proxy memory) that intend to communicate loads (of actual memory) are distinguished from
+%read operations that intend to communicate a store by using different encodings of the logical \texttt{OR} instruction.
 
 % --------------------------------------
-\subsection{Heap Sampling}\label{sec:heapsample}
-
-\Toolname{} follows DataCollider's sampling based approach, but samples proxy memory instead of code \cite{DataCollider}.
-\Toolname{} applies roughly the following high-level algorithm to sample proxy memory and detect potential false sharing:
-\begin{enumerate}
-\item Choose a proxy memory address $P$ uniformly at random. Add a hardware watchpoint $w(P)$ to $P$.
-\item If the watchpoint $w(P)$ is not triggered within $t$ units of time, then remove $w(P)$ and go to step (1).
-\item If $w(P)$ is triggered for thread $T_i$ then pause $T_i$ for $t$ units of time. The hope is that within this
-time slice, another thread, $T_j$ will be caught in the act of contending on $P$ by triggering $w(P)$.
-\item If $w(P)$ is not triggered again for another $t$ time units, then remove $w(P)$ and go to step (1).
-\item If another thread $T_j$ triggers $w(P)$, then record stack traces for $T_i$ and $T_j$, and mark the associated program
-counters of each of $T_i$ and $T_j$ as potentially having false sharing.
-\item Go to step (1) to repeat the process for a new sample location.
-\end{enumerate}
-
-As described, our sampling algorithm is susceptible to bias: if only a small percentage of the application's data is concurrently
-modified, or ``hot", then our algorithm might miss potential false sharing by wasting time sampling from the cold data.
-To avoid this issue, we apply two new techniques: \emph{uniform heap sampling}, and \emph{selective heap sampling}.
-
-\newcommand\TypeId{$TypeId$}
-
-\paragraph{Uniform Heap Sampling}
-Uniform heap sampling is a round-robin algorithm that uniformly selects proxy memory locations to sample according to
-the types of data proxied. We interpose on memory allocation and deallocation routines (e.g. \texttt{malloc} and \texttt{free})
-and associate unique \TypeId{}s to each \TypeIdPair{}, where $PC$ is the return address of an allocation function, and $size$ is the
-number of bytes allocated.
-
-For each proxy memory location $P$, we maintain a counting \emph{typeset}: a mapping of \TypeId{}s to the number of objects of type
-\TypeId{} that are fully or partially proxied by $P$.
-
-\paragraph{Selective Heap Sampling}
-Selective heap sampling is a biased variant of uniform heap sampling that enables \Toolname{} to only focus on the subset
-of the heap that is believed to contain objects that participate in false sharing. Selective heap sampling builds typesets
-for only a subset of proxy memory associated with specific allocation points.
-
-Selective heap sampling also allows a user to guide the sampler to only focus on specific types of data. This is beneficial
-when one is trying to solve specific performance problems related to a subset of the data being accessed by the application.
-
-% --------------------------------------
-\subsection{Reducing False Negatives}\label{sec:falsenegs}
-
-One challenge with sampling-based approaches is their ability to get ``unlucky" and either not sample
-the right thing, or sample the right thing at the wrong time. In our case, this results in false-negatives:
-we might report that no false sharing is detected in some code, but this does not imply the absence of
-false sharing.
-
-To address this challenge, we propagate existing beliefs about potential false sharing on a single instruction
-to nearby instructions, as well as to instructions that operate on data of the same type. The intuition is
-two-fold. First, we assume locality of accesses to shared data: if we observe a shared access on one instruction,
-then nearby instructions probably access shared data as well. Second, a belief of false sharing is a belief of
-sharing: if we observe an object of a particular type being concurrently accessed in one place, then it might
-imply that concurrent accesses can also happen in other places.
-
-We perform two separate analyses to extend belief sets: \begin{description}
-\item[Type Analysis] \hfill \\
-This analysis instruments all code and identifies what types of objects are accessed by what code. We apply the same
-notion of typesets as used in the uniform heap sampler; however, in this case typesets are associated with basic
-blocks of code instead of with proxy memory. We use address watchpoints \cite{AddressWatchpoints} to taint
-allocated addresses with their \TypeId{}, and then propagate these \TypeId{}s to the typesets of the blocks operating
-of data of each type.
-
-\item[Concurrency Analysis] \hfill \\
-This analysis instruments all code and identifies what functions can execute concurrently. Each instrumented function
-is associated with a core set and a threadset \cite{DrContention}. When executed, the coreset is updated to include the
-identifier of the CPU core executing the function. Similarly, the function's threadset is updated to include an approximation
-of the thread identifier ($SP >> 13$). We anlayse the coresets and threadsets of each function to determine whether or
-not a function can execute concurrently (across CPU cores, or in two threads on the same core).
-
-\end{description}
-
-We combine the type and concurrency analyses to propagate beliefs in the following way: if sampling detects potential
-false sharing on a single instruction operating on a specific type of object, then we mark all instructions in the intersection
-of those identified to operate on the same type with the set of concurrent functions.
-
-% --------------------------------------
-\subsection{Eliminating False Positives}\label{sec:falsepos}
-
-We apply two techniques to reduce and finally eliminate false positives:
-\begin{enumerate}
-\item We iteratively refine the granularity of proxy memory, starting from the page granularity and going down to the cache
-line granularity. As the granularity is refined and potential false sharing continues to be detected, our confidence in beliefs
-of false sharing improves.
-
-\item We apply a precise data race detector \cite{DataCollider} to instructions that demonstrate cache line contention. This
-allows us to distinguish between true and false sharing.
-\end{enumerate}
-
-Distinguishing between false sharing and pseudo sharing is more challenging. We approximate this detection by extending
-the type analysis phase to record the relative byte offset within each type of accessed object for each instruction. If false
-sharing is detected between two instructions operating on the same offset ($\pm$ the memory access size), then we report
-pseudo sharing, otherwise false sharing. Another valid heuristic would include the natural alignment and sizes of object
-types.
+%\subsection{Heap Sampling}\label{sec:heapsample}
+%
+%\Toolname{} applies roughly the following high-level algorithm to sample proxy memory and detect potential false sharing:
+%\begin{enumerate}
+%\item Choose a proxy memory address $P$ uniformly at random. Add a hardware watchpoint $w(P)$ to $P$.
+%\item If the watchpoint $w(P)$ is not triggered within $t$ units of time, then remove $w(P)$ and go to step (1).
+%\item If $w(P)$ is triggered for thread $T_i$ then pause $T_i$ for $t$ units of time. The hope is that within this
+%time slice, another thread, $T_j$ will be caught in the act of contending on $P$ by triggering $w(P)$.
+%\item If $w(P)$ is not triggered again for another $t$ time units, then remove $w(P)$ and go to step (1).
+%\item If another thread $T_j$ triggers $w(P)$, then record stack traces for $T_i$ and $T_j$, and mark the associated program
+%counters of each of $T_i$ and $T_j$ as potentially having false sharing.
+%\item Go to step (1) to repeat the process for a new sample location.
+%\end{enumerate}
+%
+%As described, our sampling algorithm is susceptible to bias: if only a small percentage of the application's data is concurrently
+%modified, or ``hot", then our algorithm might miss potential false sharing by wasting time sampling from the cold data.
+%To avoid this issue, we apply two new techniques: \emph{uniform heap sampling}, and \emph{selective heap sampling}.
+%
+%\newcommand\TypeId{$TypeId$}
+%
+%\paragraph{Uniform Heap Sampling}
+%Uniform heap sampling is a round-robin algorithm that uniformly selects proxy memory locations to sample according to
+%the types of data proxied. We interpose on memory allocation and deallocation routines (e.g. \texttt{malloc} and \texttt{free})
+%and associate unique \TypeId{}s to each \TypeIdPair{}, where $PC$ is the return address of an allocation function, and $size$ is the
+%number of bytes allocated.
+%
+%For each proxy memory location $P$, we maintain a counting \emph{typeset}: a mapping of \TypeId{}s to the number of objects of type
+%\TypeId{} that are fully or partially proxied by $P$.
+%
+%\paragraph{Selective Heap Sampling}
+%Selective heap sampling is a biased variant of uniform heap sampling that enables \Toolname{} to only focus on the subset
+%of the heap that is believed to contain objects that participate in false sharing. Selective heap sampling builds typesets
+%for only a subset of proxy memory associated with specific allocation points.
+%
+%Selective heap sampling also allows a user to guide the sampler to only focus on specific types of data. This is beneficial
+%when one is trying to solve specific performance problems related to a subset of the data being accessed by the application.
+%
+%% --------------------------------------
+%\subsection{Reducing False Negatives}\label{sec:falsenegs}
+%
+%One challenge with sampling-based approaches is their ability to get ``unlucky" and either not sample
+%the right thing, or sample the right thing at the wrong time. In our case, this results in false-negatives:
+%we might report that no false sharing is detected in some code, but this does not imply the absence of
+%false sharing.
+%
+%To address this challenge, we propagate existing beliefs about potential false sharing on a single instruction
+%to nearby instructions, as well as to instructions that operate on data of the same type. The intuition is
+%two-fold. First, we assume locality of accesses to shared data: if we observe a shared access on one instruction,
+%then nearby instructions probably access shared data as well. Second, a belief of false sharing is a belief of
+%sharing: if we observe an object of a particular type being concurrently accessed in one place, then it might
+%imply that concurrent accesses can also happen in other places.
+%
+%We perform two separate analyses to extend belief sets: \begin{description}
+%\item[Type Analysis] \hfill \\
+%This analysis instruments all code and identifies what types of objects are accessed by what code. We apply the same
+%notion of typesets as used in the uniform heap sampler; however, in this case typesets are associated with basic
+%blocks of code instead of with proxy memory. We use address watchpoints \cite{AddressWatchpoints} to taint
+%allocated addresses with their \TypeId{}, and then propagate these \TypeId{}s to the typesets of the blocks operating
+%of data of each type.
+%
+%\item[Concurrency Analysis] \hfill \\
+%This analysis instruments all code and identifies what functions can execute concurrently. Each instrumented function
+%is associated with a core set and a threadset \cite{DrContention}. When executed, the coreset is updated to include the
+%identifier of the CPU core executing the function. Similarly, the function's threadset is updated to include an approximation
+%of the thread identifier ($SP >> 13$). We anlayse the coresets and threadsets of each function to determine whether or
+%not a function can execute concurrently (across CPU cores, or in two threads on the same core).
+%
+%\end{description}
+%
+%We combine the type and concurrency analyses to propagate beliefs in the following way: if sampling detects potential
+%false sharing on a single instruction operating on a specific type of object, then we mark all instructions in the intersection
+%of those identified to operate on the same type with the set of concurrent functions.
+%
+%% --------------------------------------
+%\subsection{Eliminating False Positives}\label{sec:falsepos}
+%
+%We apply two techniques to reduce and finally eliminate false positives:
+%\begin{enumerate}
+%\item We iteratively refine the granularity of proxy memory, starting from the page granularity and going down to the cache
+%line granularity. As the granularity is refined and potential false sharing continues to be detected, our confidence in beliefs
+%of false sharing improves.
+%
+%\item We apply a precise data race detector \cite{DataCollider} to instructions that demonstrate cache line contention. This
+%allows us to distinguish between true and false sharing.
+%\end{enumerate}
+%
+%Distinguishing between false sharing and pseudo sharing is more challenging. We approximate this detection by extending
+%the type analysis phase to record the relative byte offset within each type of accessed object for each instruction. If false
+%sharing is detected between two instructions operating on the same offset ($\pm$ the memory access size), then we report
+%pseudo sharing, otherwise false sharing. Another valid heuristic would include the natural alignment and sizes of object
+%types.
 
@@
 % --------------------------------------
-\subsection{Ranking}\label{sec:ranking}
-
-Because we sample data of all types, we might find instances of false sharing that have no significant impact on performance.
-Therefore, we need a means of estimating the performance impact of false sharing so that we can rank detected instances of
-false sharing from most to least important. We accomplish this as part of a separate analysis step. This step
-instruments all code to count how often each basic block is executed. We correlate this information with detected
-instances of false sharing to infer cases where one or both threads are executing hot code and contending on a single
-cache line. In the former case, we apply a variant of the cold code hypothesis \cite{LiteRace} to rank cases of cold
-code ``pessimizing" hot code more highly.
+%\subsection{Ranking}\label{sec:ranking}
+%
+%Because we sample data of all types, we might find instances of false sharing that have no significant impact on performance.
+%Therefore, we need a means of estimating the performance impact of false sharing so that we can rank detected instances of
+%false sharing from most to least important. We accomplish this as part of a separate analysis step. This step
+%instruments all code to count how often each basic block is executed. We correlate this information with detected
+%instances of false sharing to infer cases where one or both threads are executing hot code and contending on a single
+%cache line. In the former case, we apply a variant of the cold code hypothesis \cite{LiteRace} to rank cases of cold
+%code ``pessimizing" hot code more highly.
 
@@
 % --------------------------------------------------------------------------------------
-\section{Status}\label{sec:status}
-\Toolname{} is being implemented using the Granary dynamic binary translator \cite{Granary} for the x86-64 architecture.
-
-A limited form of proxy memory has been implemented. Work is progressing on making it possible to generically interpose
-on memory allocators. Once this is done, the next step will be to work on the type analysis, as well as to associate typesets
-with proxy memory locations. Code hotness analysis is expected to be an easy-to-do task. Concurrency analysis will be trickier.
-
-Sampling and hardware watchpoint management will be orchestrated using GDB scripts.
-
-% TODO...
+%\section{Status}\label{sec:status}
+%\Toolname{} is being implemented using the Granary dynamic binary translator \cite{Granary} for the x86-64 architecture.
+%
+%A limited form of proxy memory has been implemented. Work is progressing on making it possible to generically interpose
+%on memory allocators. Once this is done, the next step will be to work on the type analysis, as well as to associate typesets
+%with proxy memory locations. Code hotness analysis is expected to be an easy-to-do task. Concurrency analysis will be trickier.
+%
+%Sampling and hardware watchpoint management will be orchestrated using GDB scripts.
+%
+%% TODO...
 
@@
 \begin{itemize}
-\item What are \Toolname{}'s overheads? For example, what was the performance impact of the selective instrumentation
-and tagged read optimizations?
+\item Does the training phase significant reduce the performance overheads of the detection phase?
 
-\item How sensitive is \Toolname{} to different sampling rates, delays, and approaches? How does the thread
-pause time during sampling affect our ability to detect false sharing? Can we quantify the benefits (if any) of
-uniform heap sampling over na{\"i}ve heap sampling?
-
-\item How effective is \Toolname{} at discovering false sharing? Is it able to distinguish between true sharing, false sharing,
-and pseudo sharing?
-
-\item How impactful are the instances of false sharing detected by \Toolname{}? Is code hotness a good metric for ranking
-false sharing?
 \end{itemize}
 
-\subsection{Evaluation Challenges}
-Granary is a medium weight DBT system, therefore its overheads will likely be significant. It is also new, and so no time has been
-spent optimizing the code produced by Granary. Therefore, the overheads are not expected to be particularly good. The major
-challenge will be convincing readers that:
-\begin{enumerate}
-\item Most of the analysis can be performed in a non-production environment, where we can claim that speed doesn't matter.
-This will obviously be a contentious claim, because speed always matters, especially in false sharing detection. Here, I hope
-that the belief propagation steps of the type and concurrency analysis stages will be beneficial in allaying those concerns.
-
-\item The application of DataCollider to distinguish true and false sharing can be performed in production given the results of
-all prior stages of the pipeline.
-\end{enumerate}
+%Our evaluation aims to answer the following questions:
+%\begin{itemize}
+%\item What are \Toolname{}'s overheads? For example, what was the performance impact of the selective instrumentation
+%and tagged read optimizations?
+%
+%\item How sensitive is \Toolname{} to different sampling rates, delays, and approaches? How does the thread
+%pause time during sampling affect our ability to detect false sharing? Can we quantify the benefits (if any) of
+%uniform heap sampling over na{\"i}ve heap sampling?
+%
+%\item How effective is \Toolname{} at discovering false sharing? Is it able to distinguish between true sharing, false sharing,
+%and pseudo sharing?
+%
+%\item How impactful are the instances of false sharing detected by \Toolname{}? Is code hotness a good metric for ranking
+%false sharing?
+%\end{itemize}
+%
+%\subsection{Evaluation Challenges}
+%Granary is a medium weight DBT system, therefore its overheads will likely be significant. It is also new, and so no time has been
+%spent optimizing the code produced by Granary. Therefore, the overheads are not expected to be particularly good. The major
+%challenge will be convincing readers that:
+%\begin{enumerate}
+%\item Most of the analysis can be performed in a non-production environment, where we can claim that speed doesn't matter.
+%This will obviously be a contentious claim, because speed always matters, especially in false sharing detection. Here, I hope
+%that the belief propagation steps of the type and concurrency analysis stages will be beneficial in allaying those concerns.
+%
+%\item The application of DataCollider to distinguish true and false sharing can be performed in production given the results of
+%all prior stages of the pipeline.
+%\end{enumerate}
 

### 2014-12-17 HotOS'15/paper.tex
Working on the methodology section still. Kind of boring so far.

@@
 \usepackage{hyperref}
+
+% For algorithms
+\usepackage{program}
 
@@
 \end{center}
-\caption{\label{fig:states_of_shadow}States of individual shadow memory cells. Each cell tracks the watched and ownership %
-status of a cache line. Yellow states are unwatched. Green states are watched. Red states represent contention that was missed %
-due to races in the instrumentation. Transitions between states are based on (potentially stale) observed prior states. For %
-example, the spurious acquire transition is possible because the decision to make the transition is based on a racy \texttt{CMP} %
-against the $SampleId$ (\figref{inline_assembly}). If two threads simultaneously see that the cache line it watched but not %
-owned, then both will execute the atomic \texttt{XCHG}, and only one thread will ``win". The ignore transition is induced, %
-either because the sampling timeout expires, or more likely, because the report transition occurs.}
+\caption{\label{fig:states_of_shadow}States of individual shadow memory cells. Shadow memory cells begin as unwatched. %
+Then, when a cache line is selected for sampling, it becomes watched. This is represented by the presence of a $SampleId$. %
+The first thread that tries to access the shadow memory of the cache line will become the cache line's owner, as represented %
+by the presence of a $ThreadId$. If multiple threads attempt to simultaneously acquire ownership of the cache line, then %
+contention will be detected and reported. If not, then execution will continue until a second thread accesses the cache line, %
+thus triggering an ownership transfer.}
 \end{figure}
@@
 ;; Racy watchpoint check.
-CMP [ShadowAddr + Offs(SampleId)], 0
+CMP [Shadow(Addr) + Offs(SampleId)], 0
 JZ ignore
 ;; Racy ownership check.
-CMP [ShadowAddr + Offs(ThreadId)], CurrThreadId
+CMP [Shadow(Addr) + Offs(ThreadId)], MyThreadId
 JZ ignore
 ;; Take ownership, or transfer ownership
-LOCK XCHG [ShadowAddr], OldState
-CALL CheckContention(OldState,Addr,PC)
+CALL CheckContention(Addr,PC)
 ignore:
 \end{lstlisting}
-\caption{\label{fig:inline_assembly}Instrumentation instructions injected for each shared memory access. The algorithm optimizes on %
-instruction footprint by using a single \texttt{XCHG} instruction to perform multiple state transitions from %
-\Figref{states_of_shadow}, and then figuring out what state we should be in based on the value of $OldState$.}
+\caption{\label{fig:inline_assembly}Instrumentation instructions injected for each shared memory access. The algorithm %
+optimizes on instruction footprint by performing lightweight, racy checks up front as a way of avoiding more heavyweight %
+contention detection. If the cache line appears watched, and if the race checks fail, then the $CheckContention$ function %
+is invoked. This function handles the state transitions of \Figref{states_of_shadow} internally using an atomic compare and %
+swap.}
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}[language=C,basicstyle=\footnotesize\ttfamily]
+void CheckContention(uintptr_t addr, PC pc) {
+  // Figure out address of shadow memory,
+  // given address of memory being accessed.
+  WP *wp = (addr / 64) + gShadowBase;
+  uint16_t sample_id = wp->sample_id;
+  uint16_t thread_id = GetThreadId();
+  if (!sample_id) return;
+
+  // Assume unowned; try to take ownership. We
+  // check the return value to figure out if
+  // the prior state was unowned.
+  WP state = CAS(wp, WP(sample_id,0),
+                     WP(sample_id,thread_id));
+
+  // Contention already detected, or
+  // sampling disabled due to sample
+  // period expiring.
+  if (!state.sample_id) return;
+
+  // Acquired ownership.
+  if (!state.thread_id) {
+    ReportOwner(sample_id,addr, pc);
+    PauseThread(gConfigSleepTime);
+  } else {
+    wp->sample_id = 0;  // Unwatch.
+    ReportContender(sample_id,addr, pc)
+  }
+}
+\end{lstlisting}
+\caption{\label{fig:algorithm}Algorithm for detecting contention on a cache line. Contention reports are updated as the threads %
+are discovered, and finalized at the end of each sampling period before new samples are taken. The \texttt{PauseThread} function %
+is notable. By passing a configurable sleep duration to the \texttt{PauseThread} function, we can turn on DataCollider-like detection, %
+or by not pausing, turn on Dr. Memory-like detection. Not shown in this algorithm is the extent of information passed to the reports %
+(memory operand sizes and types, call stacks).}
 \end{figure}
@@
 
-The limitation to 16 bits for $SampleId$s was a deliberate decision. We take advantage of the structure of ``canonical addresses" on the
-x86-64 architecture, whereby the high-order 16 bits of a 64-bit address must have the same value as the 47th bit. In user space, all 16
-high-order bits are 0. We represent the $ThreadId$ using the thread's ``base pointer'' (a pointer thread-local storage). Thus, the
-\texttt{XCHG} in \Figref{inline_assembly} has the side-effect of temporarily disabling the watchpoint.
-
-% TODO: describe why this is reasonable, what alternatives there are, and the complications with those alternatives (atomicity,
-% re-watching, etc)
-
-% TODO: Should we have done it where we read the value of shadow memory into a reg, then used that as the old state to pass on
-% (plus the old value of the thread pointer)? This is valid, and might be better. But this has weird transitions.
-
-% Essentially: the "right" solution using a CMPXCHG, but that requires a lot of spills/fills of regs  === heavyweight.
+If a sampled memory location is accessed, then the instrumentation instructions of \Figref{inline_assembly} will detect the non-zero
+$SampleId$, and call out to the $CheckContention$ function, but only if the current thread does not appear to own the cache line
+already. The $CheckContention$ algorithm (\Figref{algorithm}) performs both the ``Acquire" and the ``Detect" state transitions of
+\Figref{states_of_shadow} using a single atomic compare-and-swap operation (\texttt{CAS}).
 

### 2014-12-17 VEE-15/paper.tex
old thing

--- 2013-12-18 VEE-14/paper.tex
@@
 \usepackage{tabularx}
-
-% Make text in figure captions small
-\usepackage{caption3} % load caption package kernel first
-\DeclareCaptionOption{parskip}[]{} % disable "parskip" caption option
-\usepackage[small]{caption}
 
@@
 \usepackage{epsfig}
-
-% Compressed enumerations
-\newenvironment{enumerate*}%
-  {\begin{enumerate}%
-    \setlength{\itemsep}{2pt}%
-    \setlength{\parskip}{0pt}}%
-  {\end{enumerate}}
 
@@
 
-% For writing Granary tool names.
-\newcommand{\toolname}[1]{{\scshape #1}}
-
 % Customize cleverref
@@
 %make title bold and 14 pt font (Latex default is non-bold, 16 pt)
-\title{Fast and Flexible Dynamic Binary Translation for the Kernel}
+\title{Virtual Register Scheduling}
 
@@
 \subsection*{Abstract}
-TODO
 
-\section{Introduction}\label{sec:intro}
-
-Granary is a framework for creating flexible and efficient tools that instrument the Linux kernel.
-Granary's key novelty is its \emph{runtime code specialization} feature, which enables tools to decide what code to instrument, and how to instrument that code.
-Flexible specialization allows tools to be picky about instrumentation.
-For example, a tool can choose to instrument every load and store operation, but only inside nested critical sections that are executing in module code.
-Granary's runtime code specialization feature addresses two limitations of existing kernel analysis systems: \begin{inparaenum}[i)]
-	\item they have focused on instrumenting all code \cite{DRK,btkernel,QEMU}; and
-	\item they can only instrument code using a single policy.
+A common challenge faced by static and dynamic binary instrumentation
+systems is how to manage transient state used by injected or modified binary
+instructions. Virtual register systems solve this problem by providing
+instrumentation tool writers with an infinite set of registers for use by
+instrumented code. Unfortunately, generic virtual register systems are hard
+to implement in practice. In this work, we present a generic virtual
+register system for a complex instruction set (x86-64). Our virtual register
+system addresses three limitations of prior work: % \cite{Pin,Valgrind}
+\begin{inparaenum}[i)]
+	\item they require that all code be comprehensively instrumented;
+	\item they require a strong notion of threads and the availability of
+ thread-local storage, and;
+	\item they require a runtime monitor, and therefore cannot be applied
+by both static and dynamic binary instrumentation systems.
 \end{inparaenum}
 
-Instrumenting all code (e.g., the whole kernel) imposes unnecessary performance overheads when only a subset of the code is being analyzed (e.g., modules, functions).
-In practice, we would like to target instrumentation at specific code for efficiency, while also allowing rich and detailed instrumentation of the targeted code.
-Limiting the scope of what is instrumented is  important because the overhead introduced by using runtime code instrumentation is usually dominated by the instrumentation itself and not the baseline overhead of the runtime system supporting that instrumentation.
-Also, specializing based on what is instrumented allows instrumentation tools to focus instrumentation on the code of interest.
-For example, if one wants to instrument RCU read-side critical sections, as is the case with Granary's \toolname{RCUdbg} tool, then instrumenting all code just to filter out and find those critical sections is wasteful.
-
-Another limitation of existing systems is that they are unable to perform efficient runtime code specialization.
-That is, they are unable to change how code is instrumented based on the context in which that code will run.
-This is unfortunate because it means that tools wanting to instrument code differently must create their own solution that works around this deficiency.
-For example, one way that existing tools can specialize code is to generate as many versions of the code as there are potential specializations.
-When instrumented code is executing, runtime checks continually check and jump to the correct version of the code for the current execution context.
-However, runtime checks are inefficient, and generating many versions of the code introduces bloat and pollutes the CPU's instruction cache.
-Granary avoids these issues by supporting context-specific runtime code specialization that generates multiple versions of basic blocks \emph{just-in-time} (JIT), instead ahead-of-time, and without requiring runtime checks.
-
-One example tool that uses Granary's JIT-based runtime code specialization feature is \toolname{RCUdbg}: Granary's Read-Copy-Update (RCU) debugging tool.
-\toolname{RCUdbg} catches bugs related to invalid uses of RCU-protected pointers (\Cref{sec:rcudbg}).
-\toolname{RCUdbg} distinguishes between two kinds of RCU-protected pointers (reader-side and writer-side) and plain old pointers, and three execution contexts (inside of a read-side critical section, inside of a write-side critical section, and in non-RCU-related code).
-The actions taken when a pointer is de-referenced varies depending on the kind of pointer and the execution context.
-Code executing in read-side critical sections is specialized to enforce that only plain-old and read-side pointers are accessed, and that all accesses through read-side pointers are memory reads.
-Write-side critical sections are specialized in a similar way.
-
-The \toolname{RCUdbg} tool is an example of \emph{context-driven code specialization}: code executing inside of a read-side critical section is instrumented differently from code executing inside of a write-side critical section, and code executing outside of both contexts runs natively or with lightweight instrumentation.
-Achieving this level of specialization using and existing system requires complicated runtime state management and checks to switch between different versions of the code.
-With Granary, a tool like \toolname{RCUdbg} can specialize code using \emph{instrumentation policies}.
-An instrumentation policy is a function that names an execution context and decides how to instrument instructions that will execute within that context.
-When a policy is used to instrument a sequence of instructions (called a basic block), a new version of the basic block is generated and cached.
-
-\toolname{RCUdbg} defines three policies, and declares transitions between these policies when control flows in and out of RCU read- and write-side critical sections.
-When a function like \texttt{memset} is invoked in one context, the policy associated with that context is used to generate an instrumented version of \texttt{memset} for that context.
-If, for example, \texttt{memset} is never invoked within a read-side critical section, then \texttt{memset}'s code will never be specialized to perform the heavyweight checks needed to detect RCU-protected pointer misuses in this context.
-Similarly, if \texttt{memset} is executed both in a write-side critical section and outside of any RCU-related code, then two differently instrumented versions of \texttt{memset} will be generated.
-
-The key insight of Granary's approach to runtime code specialization is that code versioning is a context-tracking mechanism.
-That is, if an instruction from a basic block instrumented by \toolname{RCUdbg}'s read-side critical section policy is being executed, then the current thread must be executing inside of a read-side critical section.
-Similarly, if the read-side policy is used to instrument a basic block, then \texttt{RCUdbg} knows at instrumentation time\footnote{Instrumentation time is analogous to ``compile time", in that a policy is generally used to instrument a basic block of code once, and then that basic block may be executed many times.} that this basic block will only execute within a read-side critical section.
-If the basic block contains a \texttt{call} to the \texttt{memset} function, then \texttt{memset} is also known to execute within a read-side critical section because it does not terminate a read-side critical section.
-By deciding which version of \texttt{memset} to execute at instrumentation time, \toolname{RCUdbg} avoids expensive runtime checks that other systems impose in order to achieve similar kinds of specialization.
-\toolname{RCUdbg} can specify that \texttt{memset} will also execute within a read-side critical section by telling Granary that it should use the read-side critical section policy to instrument \texttt{memset}.
-
-The rest of the paper describes Granary in more detail.
-\Cref{sec:dbt} describes how Granary is implemented, \Cref{sec:what} describes how a Granary tool can choose to instrument only specific code, \Cref{sec:how} describes how Granary change the policy used to instrument code, etc.
-
-\section{Dynamic Binary Translation}\label{sec:dbt}
-
-Granary uses dynamic binary translation (DBT) to instrument the Linux kernel.
-DBT is used for emulation \cite{QEMU}, runtime optimization \cite{DynamoRIOOptimization}, and runtime instrumentation (analysis \cite{DynamoRIO, DRK, btkernel, ProfilingSimics}, security \cite{Vx32,NaCl,ProgramShepherding}, and debugging \cite{Valgrind}).
-We chose DBT because \begin{inparaenum}[i)]
-	\item static binary analysis tools are unable to cope with \texttt{x86-64}'s mix of variably-sized instructions and data; 
-	\item some kernel code is only distributed in a binary format (e.g. proprietary device drivers); and
-	\item virtualization-based approaches are unable to analyze non-virtualizable device drivers \cite{DRK}.
-\end{inparaenum}
-
-Granary translates and instruments kernel binaries one basic block at a time.
-In Granary, a basic block is a sequence of x86-64 instructions ending in a conditional branch (\texttt{jCC}), \texttt{ret}, or a \texttt{jmp} instruction.
-Unlike DRK and btkernel, Granary does not terminate basic blocks at \texttt{call} instructions.
-This enables longer basic blocks and native-like proceccor return address prediction, but makes Granary non-transparent (\Cref{sec:transparency}).
-Granary takes control of execution starting from ``natural" program entrypoints (e.g.
-interrupt vectors, system call entrypoints, module initializers).
-Entrypoint basic blocks  are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as control flow reaches them for the first time.
-
-Translated basic blocks are linked together and stored in a \emph{code cache}.
-Basic blocks are indexed by $(pc, policy)$ pairs, where $pc$ is the native program counter associated with the beginning of the basic block's instructions, and $policy$ is the instrumentation policy used to instrument the native instructions starting at $pc$.
-The index is maintained by a globally accessible hash table.
-
-Granary's JIT-based translation approach means that code executing from the code cache may yield control to Granary to request the address of the next basic block to execute.
-This happens when there is a control flow transfer in a basic block, and the target of that control flow transfer is not already indexed in Granary's code cache.
-When instrumented code yields to Granary, a ``context switch'' occurs that transfers execution to a CPU-private stack where Granary operates.
-Granary operates on a CPU-private stack for two reasons.
-First, if Granary did not switch stacks, then context switches into Granary risk overflowing the runtime call stack.
-This can happen because kernel stacks are typically small (4K to 8K in size), and because a context switch could happen when the native stack is nearly exhausted.
-Second, on an SMP, multiple processors can simultaneously execute Granary code (e.g.
-to translate code that will later execute).
-In such cases, Granary depends on CPU-private data structures to avoid contention, and the private stack is one such data structure.
-A context-switch back to the code cache occurs when the next basic block has been found or translated.
-At this point, instrumented execution continues.
-Similar to other DBT systems, Granary uses hot code patching (also called code splicing) to reduce the number of context switches.
-In our experience, kernel code stabilizes very quickly: context switches stop happening after the first few seconds of translation.
-
-\subsection{Basic Block Meta-Data}\label{sec:metadata}
-
-Associated with each basic block is meta-data that describes the basic block. Basic block meta-data includes: \begin{itemize}
-	\item The length in bytes of the original and translated basic blocks, and the address of the first instruction in the original basic block and the instrumented basic block.
-This information helps provide a rich debugging interface for instrumentation tool authors.
-Granary is unique among DBT tools in that the debugging experience is actually pleasant.
-For example, Granary has a built-in reverse debugging mechanism, which allows developers to inspect previously executed basic blocks and past register state.
-	\item Linux kernel-specific exception table meta-data about instructions contained within the current basic block.
-This kernel-specific meta-data is used by Granary to implement transparent page fault exception handling (\Cref{sec:interrupt_return_address_transparency}).
-	\item Ranges of instructions that must be executed atomically. Granary uses this information to implement selective interrupt delaying (\Cref{sec:interrupt_delay}).
-	\item The allocator used to allocate the instructions of the basic block. Granary uses this information to implement its trace-allocation optimization (\Cref{sec:trace_alloc}).
-	\item The instrumentation policy used to instrument the basic block.
-Granary uses this information to allow instrumentation tools to define policy-specific interrupt handling routines.
-For example, \toolname{RCUdbg} implements a read-side policy-specific interrupt handling routine that reports an error if code executing within a read-side critical section is interrupted (\Cref{sec:rcudbg}).
-\end{itemize}
-
-Basic block meta-data can also be extended by instrumentation tools.
-This makes developing instrumentation tools that record per-basic block statistics very easy to make.
-For example, Granary's control-flow graph tool, \toolname{cfg}, extends the basic block meta-data to track per-block and per-edge execution counts (\Cref{sec:cfg}).
-This data is useful because it provides precise code coverage information (block and branch coverage), as well as useful data for profile-guided optimization.
-This meta-data is specified by \toolname{cfg} and automatically allocated by Granary.
-
-\subsection{Transparency}\label{sec:transparency}
-An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}.
-Granary includes configurable levels of transparency to address the trade-off between transparency and overhead.
-By default, Granary uses a relaxed transparency \cite{btkernel} model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility.
-By default, the only artifact Granary exposes is that interrupt and function return addresses can point into the code cache.
-
-\subsubsection{Function return addresses}\label{sec:return_address_transparency}
-Unlike most DBT frameworks, Granary's basic blocks do not terminate at \texttt{call} instructions.
-This exposes code cache addresses to instrumented module code in the form of return addresses stored on the runtime call stack.
-The benefit of including \texttt{call} instructions are: \begin{enumerate}
-
-	\item {\bf Zero overhead procedure returns.} Unlike Granary, DRK emulates procedure calls and returns.
-DRK emulates \texttt{call} instructions by pushing the intended native return address onto the runtime call stack, followed by a direct direct jump to the next (instrumented) basic block \cite{DRK}.
-This approach enables return address transparency, but introduces additional overheads on procedure returns, which must do costly indirect branch lookups to find the intended return target.
-
-Btkernel's approach is similar in spirit to Granary's; however, btkernel ends basic blocks at \texttt{call} instructions.
-This introduces unnecessary control-flow on the return path from procedures.
-That is, btkernel's ``call/ret optimization" ostensibly enables zero-overhead procedure returns by leaving \texttt{ret} instructions as-is; however, the benefits are more apparent than real.
-To transfer control to the next logical basic block after a procedure return, btkernel's \texttt{ret}s first transfer control to the end of the basic block that performed the \texttt{call}, and then \texttt{jmp} to the next basic block.
-Granary avoids this unnecessary \texttt{jmp} because basic blocks do not end at \texttt{call} instructions.
-
-In some cases, emulating DRK's return-address lookup mechanism is desirable.
-For example, if Granary attaches while some program's execution is underway (and so the runtime call stack contains native return addresses), then a direct \texttt{ret} back to one of the native return addresses might be undesirable (e.g.
-if the tool wants comprehensive instrumentation of all code).
-In such cases, a tool can partially specialize instrumentation according to whether or not the return address might be in the code cache.
-Partial specialization is enabled by \emph{policy properties}, which are transparently inherited along with normal instrumentation policy information.
-In this case, the policy property can be used to tell Granary to emulate \texttt{ret} instructions when the return address is potentially a native address, and directly \texttt{ret} when the return address is known to be in the code cache.\footnote{This feature is used when Granary is configured to instrument user space programs.
-User space Granary depends on the Linux dynamic loader to inject Granary into a running process.
-When injected, a Granary takeover function is executed on an already active native runtime call stack.
-Granary's takeover procedure transparently emulates returns, thus taking over the previously fully-native call stack.
-This feature allows instrumentation tools to focus on their particular objectives and not the intricacies of thread takeover.}
-
-	\item {\bf Improved visibility.} Not ending basic blocks on \texttt{call} instructions results in longer basic blocks, which in turn provide additional visibility to tools at instrumentation time into what will be executed by the current thread.
-\toolname{RCUdbg} benefits from this additional visibility because it can observe the end of read-side critical sections (\texttt{call}s to \texttt{rcu\_read\_unlock}), and optimistically assume that functions executing before the read-side critical section ends will execute within the read-side critical section.
-
-\end{enumerate}
-
-\subsubsection{Interrupt return addresses}\label{sec:interrupt_return_address_transparency}
-Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts.
-A special case arises for page fault exceptions that occur within specific kernel functions that access user space data \cite{btkernel}.
-The Linux kernel records ranges of code addresses that are permitted to fault in an ``exception table'' data structure, and on a page fault exception, checks if the exception return address belongs to one of the pre-defined ranges.
-Granary records any exception table entries as part of each basic block's meta-data, so that a page fault in a basic block can be mapped to the correct native faulting program counter.
-
-\subsection{Reentrancy}
-Reentrancy is an important concern for all DBT systems, and is especially important for DBT systems operating in the kernel environment.
-For example, hardware interrupts can redirect control flow at any instruction, and interrupt delivery cannot be delayed past the execution of memory instructions that might alter the interrupt delivery state \cite{DRK}.
-Preserving interrupt transparency, however, requires that interrupts either be delayed until the end of any injected instrumentation instructions.
-Together, these requirements make it very difficult to write instrumentation tools.
-
-\subsubsection{Granary}
-For efficiency, most kernel-space instrumentation systems (Granary included) operate on CPU-private state (e.g.
-Granary's CPU-private stacks).
-However, accessing CPU-private data without first disabling interrupts results in undefined behavior.
-Therefore, context-switches into Granary first disable interrupts and then switch execution onto one of Granary's CPU-private stacks.
-Context switches out of Granary and into native code or the code cache returns execution to the native stack and re-enable interrupts.
-Execution behavior is preserved because Granary saves the machine state (i.e.
-registers, flags) in a reentrant way.
-All other parts of Granary are fully reentrant.
-
-\subsubsection{Instrumentation}\label{sec:interrupt_delay}
-
-\begin{figure}[t!]
-\begin{multicols}{2}
-\vspace{-1.5em}\subfloat[\label{fig:interrupt_delay_drk}Pre and post instrumentation in DRK.
-Injected instrumentation blocks are treated as atomic with respect to interrupts.]{\includegraphics[height=10em,clip=true,trim=0 5em 10em 0]{diagrams/drk_instrumented_mov.pdf}}\vspace{-2.5em}
-\columnbreak
-\vspace{-1.5em}\hspace{-0.5em}\subfloat[\label{fig:interrupt_delay_granary}Pre and post instrumentation in Granary.
-The two instrumentation blocks, as well as the native instruction are wrapped in a delay region.]{\includegraphics[height=10em,clip=true,trim=0 5em 10em 0]{diagrams/granary_instrumented_mov.pdf}}\vspace{-2.5em}
-\end{multicols}
-\caption{Example showing equivalent pre- and post-instruction instrumentation with DRK (left) and Granary (right).
-In DRK, the injected instrumentation instructions (grey) are guaranteed to execute atomically, up until the next native instruction (white).
-In Granary, no guarantees are made on injected instrumentation, unless otherwise instructed.
-In this case, a delay region surrounds both the pre- and post-instrumentation, as well as the native instruction, guaranteeing that the entire sequence of instructions execute on the same CPU.}
-\end{figure}
-
-One challenge with any DBT system is how it copes with non-reentrant instrumentation.
-For example, instrumentation that modifies CPU-private data structures (without disabling interrupts) cannot be safely executed by a pre-emptive kernel running on an SMP.
-If the thread being instrumented is interrupted, then the kernel might decide to re-schedule the thread to resume its execution on a different CPU.
-Any modifications to CPU-private state would lead to undefined behavior.
-An example of this is instrumentation that acquires a CPU-private lock on one CPU, is interrupted, and later releases the wrong lock on the wrong CPU.
-
-DRK's solution to this problem assumes that all injected instrumentation instructions are potentially non-reentrant, and \emph{delays} interrupts occuring within instrumented code \cite{DRK}.
-Interrupt delaying allows DRK to maintain the precise interrupt semantics of x86-64: interrupts and exceptions only arrive on native instruction boundaries, and kernel interrupt handlers only observe faults or exceptions at native instruction boundaries.
-This approach introduces subtle complexity when injecting non-idempotent instrumentation before (pre) and after (post) native instructions.
-
-\Cref{fig:interrupt_delay_drk} shows an example of pre- and post-instrumentation around a \texttt{mov} instruction.
-If an interrupt arrives within the pre-instrumentation, then DRK delays they delivery of the interrupt until the native instruction (\texttt{mov}).
-After the interrupt is handled and execution attempts to return to the interrupted instructions, DRK will re-instrument the ``tail" of interrupted basic block, starting from the \texttt{mov} instruction.
-This involves re-creating the pre- and post-instrumentation of the \texttt{mov} in the tail basic block.
-When executed, this will result in an unbalanced number of \texttt{push} and \texttt{pop} instructions.
-That is, \texttt{push \%rax} is executed in both the original and tail basic block, but \texttt{pop \%rax} is only executed in the tail basic block.
-This is a specific example of a more general problem: DRK's approach to interrupt delaying does not play well with non-idempotent instrumentation.
-
-Btkernel provides no solution to the problem of handling non-reentrant instrumentation, beyond requiring that instrumentation tools disable and re-enable interrupts around non-reentrant instrumentation code \cite{btkernel}.
-The is inefficient because it introduces flag stalls into the instruction pipeline \cite{AgnerMicroarchitecture}, and because interrupts arrive relatively infrequently.
-
-Granary charts a middle ground between the two aforementioned approaches.
-Most instrumentation is re-entrant, partially because kernel instrumentation tools can depend on the runtime call stack for scratch storage space.
-However, interrupt delaying is sometimes useful because it enables non-reentrant instrumentation to avoid having to continually disable and re-enable interrupts.
-For example, Granary's \toolname{Guardrail} tool records traces of executed module instructions into CPU-private buffers \cite{Guardrail}.
-These traces are concurrently analyzed for bugs by another thread.
-Operating on a CPU-private buffer is more efficient than operating on a global buffer because it avoids cache line contention and locking on an SMP.
-However, writing to a CPU-private buffer is only safe if the entire write executes on the same CPU.
-An equivalent tool developed for btkernel---which does not support interrupt delaying---would have to disable and the re-enable interrupts around the write.
-
-Granary conveniently solves this problem by allowing instrumentation tools to declare \emph{interrupt delay regions}.
-An interrupt delay region is a sequence of instructions (which can contain internal control flow) inside of a basic block that must be executed atomically.
-\Cref{fig:interrupt_delay_granary} shows an interrut delay region that surrounds both the pre- and post-instrumentation of a native \texttt{mov} instruction.
-If an interrupt arrives anywhere within this region of code, Granary will delay the interrupt to the end of the region.
-The declarative nature of interrupt delay regions allows instrumentation developers to opt-in to interrupt delaying, and control exactly when, where, and for how long interrupts can be delayed.
-
-Granary implements interrupt delay regions by recording bounds information about interrupt delay regions inside of each basic block's meta-data (\Cref{sec:metadata}).
-When a basic block is interrupted, Granary inspects the block's meta-data and determines if the interrupt must be delayed.
-If a delay is required, then all instructions belonging to the delay range are copied into a CPU-private buffer.
-All instructions are copied because a delay region might contain internal control-flow.
-A CPU-private buffer is used because the number of CPUs is a natural upper bound on the number of simultaneously executing delay regions.
-After the copied instructions, Granary emits code that code re-creates the interrupt stack frame, followed by a jump to the interrupt vector (which is controlled by Granary).
-This jump ensures that Granary's interrupt handler regains control after the instructions within the delay region fully execute, but before the kernel observes the interrupt.
-When Granary regains control, it can then delegate to the instrumentation tool to handle the interrupt, or defer handling of the interrupt to the kernel.
-
-When all instructions have been emitted to the CPU-private buffer, Granary emulates a return from the interrupt (without issuing an \texttt{iret} instruction).
-The emulated return leaves interrupts disabled and transfers control to the CPU-private buffer.
-Leaving interrupts disabled ensures that the remainder of the delay region will execute on the current CPU, and avoids reentrancy issues related to the copied region being located in a CPU-private buffer.
-
-\subsection{Control Flow Implementation}
-
-Granary's mechanism of looking up branch targets and resolving control flow depends only on control flow instruction (CTI) policy and target information.
-This information is either directly embedded in emitted executable code, or in immutable data structures directly referenced by emitted code.
-This approach ensures that policy information is never lost or corrupted, even in the face of concurrent executions of the same emitted code, arbitrary pre-emption, and arbitrary resumption.
-Because Granary's translation mechanism is driven by these two inputs, Granary is otherwise stateless\footnote{This is not quite true, as Granary maintains persistent state in the form of the code cache index.
-However, the code cache index behaves similarly to memoization in a purely functional programming language: it is an optimization, not a necessity.} across context switches.
-
-\subsubsection{Direct Control Flow}\label{sec:direct_control_flow}
-\lstset{language=[x64]Assembler}
-\newsavebox\nativejcc
-\begin{lrbox}{\nativejcc}
-\begin{minipage}[b]{4cm}
-\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-cmp $0x0, %rax
-jz value_is_zero 
-...
-( fall through )
-\end{lstlisting}
-\end{minipage}
-\end{lrbox}
-
-\newsavebox\translatedjcc
-\begin{lrbox}{\translatedjcc}\begin{minipage}[b]{4cm}\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-  cmp $0x0, %rax
-  jz edge_value_is_zero
-  jmp edge_fall_through
-
-edge_value_is_zero:
-  call *(jz_patch_data)
-edge_fall_through:
-  call *(jmp_patch_data)
-\end{lstlisting}
-\end{minipage}
-\end{lrbox}
-
-\begin{figure}[t!]
-\subfloat[Native Code]{\usebox\nativejcc}
-\hfill
-\subfloat[Translated Code]{\usebox\translatedjcc}
-\caption{\label{fig:direct_edge_code}Example translation of a conditional control flow instruction (\texttt{jz}) and its edge code.
-The original \texttt{jz} is translated into a \texttt{jz} that uses a 32-bit relative displacement to transfer control to an edge code function call (\texttt{edge\_value\_is\_zero}).
-When executed, the function call reads its target from an edge data structure (\texttt{jz\_patch\_data}), which contains all information necessary to resolve the \texttt{jz}'s target and patch the original \texttt{jz}.
-Patching is performed with a four byte atomic write to the relative displacement operand of the \texttt{jz} instruction.
-Finally, the return address of the \texttt{call} instruction is altered to transfer control to the newly resolved target.}
-\end{figure}
-Like btkernel, Granary translates direct CTIs into equivalent jumps into CTI-specific ``edge code" (\Cref{fig:direct_edge_code}).
-Edge code lives outside the code cache, so as to not pollute the processor's instruction cache \cite{btkernel}.
-
-When control flows through a basic block $B_{pred}$ for the first time, any jumps taken will transfer control to edge code that is specific to the jump taken.
-Edge code contains only a single indirect \texttt{call} instruction, thus allowing multiple edge codes to occupy a single cache line.
-The indirect \texttt{call} transfers control to Granary via a function pointer located in an ``edge data structure", which describes how to translate the target basic block, $B_{succ}$.
-Granary gains control and looks up or translates the target code according instrumentation policy stored in the edge data structure.
-Then, Granary modifies the original CTI in $B_{pred}$ so that it transfers control to $B_{succ}$  (\Cref{sec:patching}).
-Finally, Granary releases control and context switches to $B_{succ}$.
-
-\subsubsection{Indirect Control Flow}\label{sec:ibl}
-
-Granary translates indirect CTIs (\texttt{call}, \texttt{jmp}) into their equivalent direct control flow versions, where the direct versions branch to ``in-edge" code.
-In-edge code loads the indirect jump target (e.g.
-a specific register or memory location), saves the flags register, and several scratch registers.
-Finally, the in-edge uses the indirect target as an index into a global, fixed-size hash table of indirect branch targets.
-The entries of the hash table either point to Granary or to ``out-edge" code (\Cref{fig:ibl}).
-
-\begin{figure}[t!]
-\hspace{4em}\includegraphics[height=13em,clip=true,trim=0 5em 10em 0]{diagrams/ibl.pdf}
-\caption{\label{fig:ibl}Process of resolving an indirect branch.
-Control transfers from a source basic block, to the in-edge code specific to where the indirect target is stored (e.g.
-register, memory address), through the global hash table and on to either one or more out-edge code stubs, or to Granary.}
-\end{figure}
-
-Out-edge code is specific to a target basic block, and performs the function of separate chaining in the indirect branch lookup hash table.
-When control transfers to out-edge code, the out-edge checks to see if the indirect target matches the out-edge code's target basic block.
-If they match, then the out-edge code restores scratch and flags registers and jumps directly to the target basic block.
-Otherwise, the out-edge code jumps to another out-edge code or to Granary.
-If one out-edge code block jumps to another one, then the targets of the two out-edge code blocks collide in the global hash table.
-In this respect, out-edge code is chained together through control flow, just as colliding entries in a typical hash table can be chained together through a linked list.
-
-Granary's use of out-edge code avoids re-entrancy issues related to indirect control flow that are present in DRK and btkernel.
-Both DRK and btkernel store the target of an indirect CTI in a CPU-private memory location.
-To complete an indirect control transfer, both systems perform (roughly) the following steps: \begin{enumerate}[itemindent=\parindent]%
-\setlength{\itemsep}{2pt}%
-\setlength{\parskip}{0pt}%
-	\item Save the flags and disable interrupts.
-	\item Save registers.
-	\item Lookup the next basic block address.
-	\item Store that address in a CPU private memory location.
-	\item Restore registers.
-	\item Restore the flags and interrupt state.
-	\item Indirectly jump to the target basic block through the CPU-private location.
-\end{enumerate} %
-However, the ordering of steps (6) and (7) introduce a reentrancy problem: the access to the CPU-private location is not safe because an interrupt might arrive before the indirect jump in step (7).
-This unusual design is motivated by the constraint that the entire lookup procedure must save (2) and restore (5) the register state for correctness, but must also save the target pc \emph{somewhere} so that execution can eventually be redirected.
-Granary's out-edge code avoids this reentrancy issue by having the out edge code itself (in the form of a direct \texttt{jmp}) contain the target pc.
-
-\subsection{Optimizations}\label{sec:optimize}
-
-\subsubsection{Ahead-of-time Translation Optimization} \label{sec:aot}
-Granary ends basic blocks at conditional branches (\texttt{jCC}), direct and indirect \texttt{jmp}s, and \texttt{ret}s and \texttt{iret}s.
-Basic blocks ending in conditional branches are extended to contain two trailing CTIs: a copy of the native \texttt{jCC}, and a synthesized direct \texttt{jmp} to the code executed when the condition code is not set (i.e.
-when the conditional jump is not taken).
-
-If a basic block ends in either a copy of a native direct \texttt{jmp} or synthesized direct \texttt{jmp}, then Granary will perform ahead-of-time translation and instrumentation of the targeted basic block.
-The repeated application of translating \texttt{jmp} targets forms traces in Granary's code cache.
-Granary optimizes traces by directly connecting basic blocks together when there is internal control flow within the blocks of the trace, and by eliding \texttt{jmp}s between adjacent basic blocks.
-
-Granary's tracing strategy is based on the observation that the most commonly executed path through code follows a straight line.
-That is, execution is more likely to fall through a conditional CTI than it is to take the conditional CTI.
-We experimentally verified Granary's tracing strategy by disabling tracing and relying on Granary's internal performance counters to count how many conditional branches were patched (by their edge code, \texttildelow 35\%), versus how many synthesized fall-through \texttt{jmp}s were patched (\texttildelow 80\%).
-
-Profile-guided ahead-of-time tracing further improves on Granary's otherwise na\"{i}ve tracing strategy by integrating profile information from Granary's \toolname{cfg} tool (\Cref{sec:cfg}).
-
-\subsubsection{Patching and Alignment  Optimizations}\label{sec:patching}
-For efficiency, basic blocks in Granary's code cache are directly connected through direct control flow instructions (\texttt{call}, \texttt{jmp}, \texttt{jCC}).
-That is, if there is a direct branch from one native instruction to another, then there will be a corresponding direct branch in the translated code.
-
-Granary opportunistically connects two basic blocks together using a direct branch when: \begin{inparaenum}[i)]
-	\item the branch transfers control to another basic block within the current trace; or
-	\item the branch transfers control to an already translated basic block.
-\end{inparaenum} However, not all basic blocks can be connected at translation time, therefore Granary employs edge code and hot-code patching (\Cref{fig:direct_edge_code}) to connect basic blocks at runtime.
-
-Granary atomically patches direct CTIs that target edge code to instead point to their (now resolved) target basic block.
-Due to limitations of the x86-64 architecture, atomic operations cannot straddle cache lines.
-Therefore, patchable instructions (or parts thereof) must be aligned so as to not straddle two cache lines.
-
-One approach to change an instruction's alignment is to pad the basic block with additional \texttt{nop} instructions.
-An alternative when more than a few bytes of padding are needed is to introduce a single control-flow instruction (\texttt{jmp}) that jumps around the padded area.
-However, adding extra instructions introduces latency into the processor's fetch/decode/execute cycle.
-Therefore, when possible, Granary uses ignored instruction prefixes to alter an instruction's alignment.
-For example, Granary aligns patchable conditional jumps using static branch prediction hint prefixes, which are ignored by recent x86-64 processors \cite{AgnerMicroarchitecture}.
-When possible, alignment requirements are reduced by allowing a patchable instruction to straddle a cache line, but for the part to be patched to be fully contained by a single cache line.
-Finally, most basic blocks are small (i.e.
-fit in one cache line), therefore, aligning basic blocks on cache line boundaries naturally avoids the potential for a patchable instruction to straddle two cache lines.
-
-\subsubsection{Trace Allocation Optimization}\label{sec:trace_alloc}
-Arranging for commonly executed sequences of basic blocks to appear adjacent in memory, also called tracing, typically improves the runtime performance of instrumented programs \cite{DynamoRIO}.
-Typical approaches to tracing require a runtime profiling infrastructure \cite{DynamoRIOOptimization}.
-Granary attempts to reap the benefits of tracing without the runtime profiling overhead by opportunistically forming traces using ahead-of-time translation (\Cref{sec:aot}), and by arranging for related basic blocks to be allocated sequentially in memory.
-
-Two basic blocks $B_1$ and $B_2$ are ``related" ($B_1 \to B_2$) if there is a control flow instruction from $B_1$ to $B_2$.
-Granary arranges for $B_2$ to be allocated nearby or adjacent to $B_1$ by allocating $B_2$'s instructions from the same pool of memory used to allocate $B_1$'s instructions.
-The base case of this induction are kernel entrypoints.
-For example, each system call is an entrypoint into the kernel, and is associated with a private memory pool.
-When a system call is executed for the first time, the basic blocks needed to fulfill the system call's computation are allocated from that system call's memory pool.
-
-To avoid translation conflicts (e.g.
-on utility functions being simultaneously executed by distinct system calls), and to improve the predictability and reproducability of a given code cache layout, Granary can delay the takeover of kernel entrypoints.
-That is, Granary can take over one entrypoint at a time, and wait for a specified period of time before taking over the next entrypoint.
-The effect of delaying the takeover process is that it increases the chances that a given entrypoint (e.g.
-a blocking system call) will execute to completion, and thus have all of its basic blocks  arranged sequentially in memory.
-
-Trace allocation is implemented by recording the memory pool used to allocate a given basic block in that basic block's meta-data (\Cref{sec:metadata}).
-When Granary needs to translate a new basic block, it inspects the predecessor basic block's meta-data and uses the predecessor basic block's memory pool to service allocation requests for the basic block in translation.
-
-\subsubsection{System Call Table Optimization}
-
-Granary treats indirect \texttt{jmp}s or \texttt{call}s through the Linux kernel system call table specially.
-The Linux kernel (as of Linux 3.12) has approximately 300 distinct system call targets, some of which are executed often.
-To avoid polluting the indirect branch lookup hash table with these targets (\Cref{sec:ibl}), and to avoid the indirect branch lookup overhead, Granary creates a separate, \emph{shadow} system call table.
-
-Granary's translation post-processing mechanism identifies indirect control flow through the system call table and alters those CTIs to direct control flow through the shadow system call table instead.
-The entries of the shadow system call table are ``lazy" entrypoint translator functions.
-Upon their first execution, these entrypoint translators find or translate the first basic block of the system call handler, and update the shadow system call table accordingly.
-
-\begin{table*}[ht!]
-\scriptsize
-\begin{tabularx}{\linewidth}{| l | >{\raggedright}p{0.3\linewidth} | X |}
-\hline
-\bf Granularity & \bf Description & \bf Example Tool \\
-\hline
-Kernel & All kernel code is instrumented.
-& Granary's \toolname{cfg} tool constructs inter- and intra-procedural control flow graphs and records per-basic block and per-edge performance counters.
-This data is useful for debugging an performance analysis.
-Applying the \toolname{cfg} tool to the whole kernel allows us to perform profile-guided optimization of kernel code, including a form of de-virtualization.
-That is, most indirect control flow instructions (\texttt{call}, \texttt{jmp}) within the kernel target only one destination.
-Fast checks can be inlined into kernel code that use direct control transfer instructions when possible, and fall-back on indirect control transfer instructions when necessary.\\
-\hline
-Module & All code from a specific kernel module or driver is instrumented.
-Granary detaches when module code invokes kernel code, and re-attaches when kernel code invokes module code.
-All non-module kernel code executes with zero overhead.
-& Granary's \toolname{Guardrail} tool records traces all instructions and memory operations performed by device driver code \cite{Guardrail}.
-A sideline thread analyzes the recorded traces to detect data race bugs and memory access bugs.
-\\
-\hline
-Function & All code within a particular function is instrumented.
-Granary attaches when the function is invoked, and detaches when the function returns or invokes another function.
-& \toolname{RCUdbg} interposes on the \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} Linux kernel functions.
-When invoked by native code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} attaches instrumentation.
-When invoked by instrumented code, \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} updates the current nesting depth of the read-side critical section.
-\\
-\hline
-\end{tabularx}
-\caption{\label{tab:granularities}Different granularities of instrumentation supported by Granary.}
-\end{table*}
-
-\section{Only Instrument What is Needed}\label{sec:what}
-
-Some instrumentation tools have a limited scope, and therefore do not need to instrument all code.
-For example, Granary's \toolname{Guardrail} tool is concerned with finding bugs in Linux device drivers \cite{Guardrail}.
-\toolname{Guardrail} records traces of executed instructions, and analyzes those traces to find memory access and concurrency bugs.
-Both phases are heavyweight, and so limiting the scope of what is instrumented to what is of interest (module code) results in better performance.
-Unfortunately, existing systems are unable to change the granularity of instrumentation.
-Fortunately, Granary allows \toolname{Guardrail} to set the granularity of instrumentation to only instrument a specific module, while leaving the rest of the kernel to execute without overhead.
-
-\toolname{Guardrail} is an example of how tools can choose the instrumentation granularity that best matches their needs.
-That is, a Granary tool can choose to instrument some code, while leaving other code to execute natively.
-This is useful when the instrumentation in question has a high overhead or when only a subset of the kernel's code is of interest.
-\Cref{tab:granularities} contains a non-exhaustive list of tools using Granary's different instrumentation granularities.
-\Cref{fig:granularities} visually shows how different granularities of instrumentation can be mixed and matched to achieve the desired instrumentation visibility.
-
-\subsection{Instrumenting All Kernel Code}
-
-Like DRK and btkernel, Granary can instrument the whole kernel.
-Instrumenting the whole kernel requires interposing on the system call entrypoint\footnote{The system call entrypoint is stored in the x86-64 \texttt{MSR\_LSTAR} model-specific register.} and on the interrupt vector entrypoints\footnote{Interrupt vector entrypoints are stored in the interrupt descriptor table, which can be read from and written to using the \texttt{sidt} and \texttt{lidt} instructions, respectively.}.
-For whole-kernel instrumentation, Granary bootstraps by duplicating the the early entrypoint instructions that are responsible for switching execution onto a kernel stack, and then injects a jump to translated/instrumented kernel code after the stack switch.
-The translation process continues when a system call or interrupt occurs and follows the jump from duplicated into translated code.
-Granary only translates code that is known to operate on a kernel stack\footnote{Not all kernel code executes on a kernel stack.
-For example, when a user space application invokes the \texttt{syscall} instruction, control is transferred to the kernel, but no stack switch occurs.
-The kernel's early system call entrypoint code is responsible for switching execution onto a kernel stack.} so that instrumentation tools can safely save transient data in a reentrant way by spilling that data to the stack.
-
-\begin{figure}[t]
-\captionsetup[subfloat]{width=0.45\columnwidth}
-\begin{tabularx}{\columnwidth}{XX}%
-\subfloat[Full-kernel instrumentation.]{\hspace{2.5em}\epsfig{file=diagrams/kernel.eps,width=0.125\textwidth}} & %
-\subfloat[Module-only instrumentation]{\hspace{2.5em}\epsfig{file=diagrams/module.eps,width=0.125\textwidth}} \\
-\subfloat[Module-only instrumentation, with a module function wrapped to run natively.]{\hspace{2.5em}\epsfig{file=diagrams/module_wrapper.eps,width=0.125\textwidth}} & %
-\subfloat[\label{fig:module_probe}Module-only instrumentation, with a kernel function wrapped to also be instrumented.]{\hspace{2.5em}\epsfig{file=diagrams/module_probe.eps,width=0.125\textwidth}}
-\end{tabularx}
-\caption{\label{fig:granularities}Visual example of using different granularities of instrumentation to selectively instrument or not instrument specific code.}
-\end{figure}
-
-\subsection{Instrumenting Only Module Code}\label{sec:module}
-
-Granary is the first kernel-space instrumentation system able to target instrumentation only at specific kernel modules, while executing the rest of the kernel natively and without overhead.
-Instrumenting only specific kernel modules is particularly useful for debugging tools because modules represent the bulk of new kernel code under development, and contain the most bugs \cite{FaultsInLinux}.
-
-The basic technique that Granary uses to interpose on module code is hardware page protection.
-First, Granary page protects existing module code, marking it as non-executable.
-Then, Granary traps attempts to execute the protected module code and transparently recovers from these traps by returning execution to instrumented code.
-Granary also interposes on the Linux kernel's module loading process.
-When a kernel module is dynamically loaded, Granary bootstraps by page protecting the new module's code, and by translating the first basic block of the module's initialization function.
-It then replaces the kernel's pointer to the native initialization function with a pointer to the translated basic block.
-The translation process continues when the kernel initializes the module by invoking the translated module code.
-
-While correct, hardware page protection is an inefficient mechanism for gaining control of module code execution.
-That is, after the (instrumented) module initializer runs to completion, Granary needs an efficient way to regain control of future invocations of the module's code.
-Our approach to regaining control is based on the observation that modules tell the kernel about their interfaces by registering interface functions with the kernel.
-These functions are the standard mechanism for extending the kernel's functionality.
-When the kernel executes a registered module function, control transfers from the kernel into the module.
-The act of registering a function is a signal to Granary that instrumentation must ``attach'' to the registered function when the kernel executes that function.
-In practice, when presented with such a signal, we would like to avoid the overhead of the trap that would occur when the registered function is executed by proactively replacing the registered function pointer with an instrumented equivalent.
-
-Granary statically analyses of the kernel source code to discover potential attach points.
-Granary bridges the gap between static and dynamic analysis by using the static type information to generate code that dynamically discovers potential attach points.
-This code also gives tools type-safe access to the runtime object graph.
-The static analysis recognizes potential attach points as being: \begin{enumerate*}
-	\item Function pointers passed as arguments to kernel functions.
-	\item Function pointers stored in structures passed (by pointer or value) to kernel functions.
-This rule applies transitively.
-For example, if one structure has a field that points to another structure, and the latter structure contains a function pointer, then the latter structure's function pointers are also potential attach points.
-This is a common pattern in the Linux kernel, and mimics virtual tables in object-oriented languages.
-	\item Function pointers and structures returned from attach points.
-\end{enumerate*}
-
-The code generation phase of Granary's static analysis creates ``wrappers'' for the kernel/module interface. Two kinds of wrappers are generated for the kernel/module interface: \begin{inparaenum}[i)]
-	\item kernel function wrappers (\Cref{fig:function_wrapper}), and
-	\item data-structure wrappers (called type wrappers).
-\end{inparaenum} Kernel function wrappers are transparently invoked when module code attempts to invoke kernel code.
-Function wrappers delegate to type wrappers to find and wrap attach points passed to the kernel.
-
-%\begin{figure*}[t!]
-%\lstset{language=C, tabsize=2, stepnumber=1}
-%\begin{multicols}{2}
-%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-%struct device_driver {
-%  ...
-%  int  (*probe)(struct device *);
-%  int  (*remove)(struct device *);
-%  void (*shutdown)(struct device *);
-%  int  (*suspend)(struct device *, pm_message_t);
-%  int  (*resume)(struct device *);
-%  ...
-%  const struct dev_pm_ops *pm;
-%  ...
-%};
-%\end{lstlisting}
-%\columnbreak
-%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-%TYPE_WRAPPER(struct device_driver, {
-%  PRE_OUT {
-%    ABORT_IF_FUNCTION_IS_WRAPPED(arg.probe)
-%    WRAP_FUNCTION(arg.probe);
-%    WRAP_FUNCTION(arg.remove);
-%    ...
-%  }
-%  POST_OUT {
-%    POST_WRAP(arg.pm);
-%  }
-%})
-%\end{lstlisting}
-%\end{multicols}
-%\caption{The Linux device driver structure is shown on the left.
-The automatically generated type wrapper for this structure is shown on the right.
-In the wrapper code, \texttt{arg} is a reference to a \texttt{struct device\_driver} object passed as, or referenced by, an argument to a kernel or module wrapper.
-Code in the \texttt{PRE\_OUT} section is applied to arguments of the wrapped type before a kernel wrapper is invoked.
-Similarly, code in the \texttt{POST\_OUT} section is applied to arguments of the wrapped type after a kernel wrapper is invoked.
-\texttt{POST\_WRAP} invokes the type wrapper that is specific to the value to which it is applied (\texttt{arg.pm}).
-Type wrappers also support \texttt{\_IN} suffixes instead of \texttt{\_OUT} suffixes, which apply to data going into modules (i.e., over module wrappers).
-Finally, the \texttt{RETURN\_} prefix is used to apply some code to return values of either kernel or module wrappers.}
-%\label{fig:type_wrapper}
-%\end{figure*}
-
-\begin{figure}[t]
-\lstset{language=C, tabsize=2, stepnumber=1}
-\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-// struct inode *
-// iget_locked(struct super_block *, unsigned long);
-FUNCTION_WRAPPER(MODULE, iget_locked, (sb, ino), {
-  PRE_OUT_WRAP(sb);
-  struct inode *inode = iget_locked(sb, ino);
-  POST_OUT_WRAP(sb);
-  RETURN_IN_WRAP(inode);
-  return inode;
-})
-\end{lstlisting}
-\caption{\label{fig:function_wrapper}Atomatically generated wrapper for the \texttt{iget\_locked} kernel function.
-\texttt{iget\_locked} obtains an inode from a mounted file system.
-The super block argument (\texttt{sb}) is first pre-wrapped before it is passed to the original \texttt{iget\_locked} function.
-After \texttt{iget\_locked} returns, \texttt{sb} is then post-wrapped.
-Finally, the inode data structure returned by the kernel (\texttt{inode}) is return-wrapped and then returned to instrumented module code.
-The \texttt{OUT} and \texttt{IN} qualifiers of the wrappers denote the direction that data is travelling.
-\texttt{OUT}-wrappers apply to data passed from instrumented code out to the native code.
-\texttt{IN}-wrappers apply to data passed from native code into instrumented code.}
-\end{figure}
-
-If a pointer to a module function (a future attach point) is discovered, then Granary replaces that pointer with a function-specific \emph{module function wrapper}.
-Module function wrappers behave similarly to kernel function wrappers, in that they provide Granary more visibility into potential attach points, and provide instrumentation tools visibility into the runtime object graph.
-
-\label{sec:wrappers}
-
-Granary's wrappers expose a wealth of static program information to instrumentation tools.
-Instrumentation tools can use information present in wrappers to alter how code is instrumented, or to simply inspect the runtime object graph in a type-safe way.
-For example, we are actively using type information present in wrappers to create a tool that generates models of typical module behavior.
-The wrapper information serves two roles in our tool.
-First, we use type wrappers to assign type-specific IDs to module-allocated memory that is shared across the module/kernel interface.
-Type ID assignment is critical because it allows us to match memory reads and writes to specific kernel data structure fields.
-Second, we use the kernel and module function wrappers to generate call graphs of a module's execution.
-This call graph models module behavior (according to the kernel) because we label module code nodes with kernel data structure and function pointer field names (derived from module wrappers).
-This labelling allows us to generalize across similar modules.
-For example, code reached by calling the \texttt{ext4\_mount} and \texttt{btrfs\_mount} functions from the \texttt{ext4} and \texttt{btrfs} file system modules, respectively, are both labelled as \texttt{file\_system\_type::mount}, because the function addresses are stored (and later replaced by module wrappers) in the \texttt{mount} field of a \texttt{file\_system\_type} data structure.
-The combined records from multiple ``trusted'' modules of the same class (e.g., mature, open-source file systems) models the behavior of typical kernel modules.
-We hope that these models will help us build tools that classify and identify spurious module behavior.
-
-\subsection{Instrumenting Specific Functions}\label{sec:function_wrapper}
-
-Instrumenting individual functions is useful when the code of interest is only a function or group of functions, or when the tool does not instrument all code, but wants visibility into functions that would otherwise execute natively.
-For example, \toolname{RCUdbg} interposes on the \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} kernel functions.
-When \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} is invoked by native kernel code, Granary attaches and begins instrumenting code using the read-side critical section instrumentation policy.
-When \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} is invoked by instrumented code, Granary updates the nesting depth of the current read-side critical section.
-\texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} behaves similarly.
-
-Granary uses code splicing \cite{KernInst,AnyWhereAnyTimeDBT,KProbes} to allow tools to interpose on the execution of individual functions invoked by native code (which nornmally execute outside of Granary's control).
-Granary allows tools to interpose on specific functions invoked by instrumented code using kernel or module function wrappers (\Cref{sec:wrappers}).
-\Cref{fig:module_probe} is a visual example of how interposing on individual functions can be combined with module-only instrumentation to provide an instrumentation tool visibility into otherwise native kernel events.
-
-\section{Runtime Instrumentation Code Specialization}\label{sec:how}
-
-Existing instrumentation systems make it challenging to change how code is instrumented based on the runtime context.
-In practice, not all code needs or should be instrumented in the same way.
-Selectivity is on example of this: not all code needs to be instrumented, which is why Granary lets tools run some code natively and other code under instrumentation.
-Just as selectivity allows a tool to choose which code to instrument, specialization allows a tool to choose how to instrument that code.
-Specialization is especially useful for analysis tools that distinguish between light and heavyweight analyses, and for debugging tools that do different bug checks in different contexts.
-For example, \toolname{RCUdbg} instruments code running inside RCU read-side critical sections differently from code running outside of read-side critical sections.
-Code is instrumented differently in these two contexts because a memory access that is valid inside of a read-side critical section can be invalid outside of a read-side critical section.
-If the instrumentation could not be specialized for these two different contexts, then \toolname{RCUdbg} would need insert instrumentation on every memory access that first determines the current context, and then performs the appropriate bug check.
-This additional check in the instrumentation introduces performance and space overheads (additional instructions) for \emph{all} memory instructions, as well as state management complexities for the runtime itself.
-That is, \toolname{RCUdbg} would need to track the current context in a reentrant way.
-%These additional overheads and complexities
-
-%n doesn't just introduce performance and space (additional instructions) overheads on \emph{all} memory accesses, it also introduces complexity in terms of state management.
-That is, \toolname{RCUdbg} would need
-
-%Not all code needs to be or should be instrumented in the same way, just as not all code needs to be instrumented. 
-%For code that is instrumented, having similar flexibility to selectivity is also useful: just
-%just as instrumented code is a ``heavyweight" version of native code, so can 
-%. In some cases, tools can distinguish between light and heavyweight instrumentation policies, just as instrumentation itself is a heavyweight version of native execution.
-% However, not all code that is instrumented should have to be instrumented in the same way.
-%For example, a security application could perform lightweight instrumentation of trusted kernel code, and more heavyweight instrumentation of untrusted modules.
-%we want to change how code is instrumented when: \begin{enumerate}
-%	\item {\bf Not all code needs to be instrumented in the same way.} This is an important concern when comprehensiveness isn't required, or when there are timing or performance constraints.
-For example,
-%	\item {\bf Different code should be instrumented differently.} For example, DBT-based sandboxing \cite{Vx32,ProgramShepherding} can be implemented by instrumenting trusted and untrusted code differently.
-In the case of the Linux, the kernel could be instrumented with a lightweight sandboxing policy for comprehensiveness, and untrusted modules could be instrumented with heavyweight policies for correctness.
-%	\item {\bf The same code should be instrumented differently depending on the context.}
-%\end{enumerate} 
-%This is unfortunate, especially when the instrumentation is heavyweight, because then all instrumented code is slowed down equally. Some bugs are sensitive to 
-%For analyses that must be run on production systems,
-
-%Granary gives tools the ability to be selective about what is instrumented.
-A natural extension of this selectivity is to allow tools to specialize how code is instrumented by dynamically switching the policy used to instrument code.
-Specialization is desirable in several cases.
-For example, some instrumentation is high-overhead, but does not need to be applied to all code.
-One example of this is Granary's \toolname{RCUdbg} tool, which instruments read-side critical sections more heavily than the rest of code.
-Unfortunately, existing DBT systems do not support runtime code specialization \cite{DRK,btkernel,Pin,DynamoRIO}.
-
-%To implement the equivalent of \toolname{RCUdbg}'s runtime code specialization using an existing DBT system requires that a tool developer maintain, update, and check runtime state to know whether or not code is executing inside or outside of a read-side critical section.
-If a basic block is executing in a read-side critical section, then instrumentation should be enabled, and if it's not in a read-side critical section then it should be disabled.
-Enabling/disabling instrumentation might be done at the instruction granularity, which introduces at least one branch instruction per native instruction, or at the granularity of basic blocks, which requires generating two copies of each native basic block, where a single branch at the beginning of a basic block chooses between the instrumented or uninstrumented version.
-Up until now, it has been assumed that there is a mechanism for maintaing state that consistently tracks the current execution context of kernel code (inside vs.
-outside of a read-side critical section).
-However, maintaining state in a way that is reentrant and resilient again arbitrary pre-emption and resumption is onerous.
-
-Fortunately for tool developers, Granary has built-in support for runtime code specialization. Granary's implementation of runtime code specialization is unique in that: \begin{enumerate}
-
-	\item {\bf It does not require manual state management.} Reentrant state management in the kernel environment is extremely complex.
-For example, using thread-private memory is a natural way to track the current execution context.
-However, there is no well-defined thread-private storage class in the Linux kernel.
-This is because interrupt handlers can, but don't always, share stacks with interrupted code, and interrupt handler code can swap out the current thread and swap in another thread.
-CPU-private storage is another option, but introduces an impedence mismatch on SMPs because contextual state is typically thread-specific, but a given thread can be rescheduled (e.g.
-by an interrupt) to execute on a different CPU.
-Finally, global state introduces potential performance bottlenecks (locking, false/true sharing) and sometimes tricky synchronization protocols.
-
-	\item {\bf It does not require any runtime checking of the current runtime context.} If a tool like \toolname{RCUdbg} were implemented with DRK or btkernel, then \toolname{RCUdbg} would need to instrument code to continually check the current context and perform the correct memory access check given the current context.
-Checking the current context introduces performance overheads (additional instructions must be executed; branch mispredictions), and memory overheads (additional instructions are needed just to figure out what to check; icache misses).
-
-\end{enumerate}
-
-Granary's approach to runtime code specialization works by generating multiple versions of each basic block \emph{just-in-time}, where each version of a basic block is specialized to a different execution context.
-
-% For example, in the Linux kernel, interrupt handling code can execute on the interrupted stack.
-If Granary is instrumenting all kernel code, the a tool developer needs to decide if the instrumented interrupt handling code should inherit the runtime context from the interrupted thread.
-Making such a judgement is further complicated by the fact that interrupt arrival and semantics are hardware-defined, but thread scheduling is software defined.
-That is, if an interrupt handler inherits
-%This results in less code cache bloat and more focused and simple instrumentation tools. That is, tool developers can focus on what the tool should do, rather than how to fit their idea of the tool
-%	\item It works just-in-time, using basic block versioning, thus generating the minimum number of versio
-% , and does not cause performance overheads by continually checking .
-%Granary's approach to runtime code specialization 
- %uses a technique called \emph{context-driven instrumentation}.
-
-\subsection{Specialization Using Basic Block Versions}\label{sec:policies}
-
-%We developed a Granary tool that detects several Read-Copy-Update (RCU) API misuses in module code.
-Our tool focused on read-side critical sections (delimited by calls to \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock} in the code), however, most code executes outside of read-side critical sections.
-As an optimization, we wanted our heavyweight API-checking instrumentation to apply only when the code was executing within a read-side critical section.
-Implementing this optimization was challenging because each basic block was only translated once, and we had no way to know whether or not the heavyweight instrumentation should be applied to it.
-Knowing the state (within or outside a read-side critical section) at translation time was not sufficient, since the same translated block could later be executed in the opposite context.
-That is, code executing \emph{within} a read-side critical section may have originally been translated while executing \emph{outside} of a read-side critical section, and would thus omit the heavyweight API-checking instrumentation.
-This omission could cause our tool to miss bugs (i.e., it would not be comprehensive) when the basic block executes within a read-side critical section.
-
-Granary tools can specialize instrumentation to the context in which the instrumented code will execute.
-Specialization is achieved by allowing multiple, differently instrumented versions of the same code to co-exist within Granary's code cache.
-That is, if the next basic block to execute has not yet been translated/instrumented for the current execution context, then a new version of that basic block will be created  specifically for the the current execution context.
-
-%In the case of our \toolname{RCUdbg}, there were three execution contexts: within and outside of a read-side critical section.
-If the same module basic block is  executed in the two contexts, then Granary's code cache would contain two different instrumented versions of that basic block.
-The way that Granary distinguishes between different execution contexts is with \emph{instrumentation policies}.
-
-Granary tools identify execution contexts by creating instrumentation policies.
-An instrumentation policy is both a name for an execution context, as well as a function that decides how to instrument basic blocks that will execute within that context.
-All tools define an initial policy that Granary uses to instrument code.
-Tools are not limited to one policy though: any policy can declare a ``policy switch" that will take effect when a selected control-transferring instruction (CTI) is executed.
-Every policy function must declare what the ``next" policy to switch to will be.
-Policy switches can also be specified on a per-CTI basis.
-If a CTI has a specified policy, then any basic blocks targeted by that CTI will be instrumented by the specified policy.
-If a CTI does not have a specified policy, then it automatically inherits the next policy.
-
-%The effect of a tool specifying a policy switch on a CTI is that the basic block(s) targeted by that CTI will be instrumented according to the specified policy.
-A CTI with an unspecified policy inherits its policy from its containing basic block.
-That is, if no policy is specified on a CTI, then
-
-For example, \toolname{RCUdbg} defines three policies: $P_{read}$ for code executing within read-side critical sections, $P_{write}$ for code executing within write-side critical sections, and $P_{outside}$ for code executing outside of RCU read- and write-side critical sections.
-The following sequence of events describe some of the policy switching behavior of \toolname{RCUdbg} and how it interacts with block versioning.
-\begin{enumerate*}
-	\item Instrumentation begins with $P_{outside}$.
-	\item $P_{outside}$ instruments a basic block containing a \texttt{call} to the \texttt{memset} kernel function. The \texttt{call} inherits the next policy, which is $P_{outside}$.
-	\item When executed, the \texttt{call} to \texttt{memset} will cause \texttt{memset} to be instrumented by $P_{outside}$.
-No other instrumented versions of \texttt{memset} exist in Granary's code cache.
-Granary knows to use $P_{outside}$ and not another policy because $P_{outside}$ is stored in the edge data structure used by the first execution of the \texttt{call} (\Cref{sec:direct_control_flow}).
-	\item Later, $P_{outside}$ instruments a basic block containing a \texttt{call} to \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock}.
-$P_{outside}$ sets next policy to $P_{read}$ and uses $P_{read}$ to instrument the tail of the basic block following the \texttt{call}.
-	\item Execution naturally switches from being in one context (outside) into another (read) as execution returns from the \texttt{call} to \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock}. 
-	\item $P_{read}$ instruments a basic block containing a \texttt{call} to \texttt{memset}. The \texttt{call} inherits the next policy, which is $P_{read}$.
-	\item When executed, the \texttt{call} to \texttt{memset} within the read-side critical section will cause \texttt{memset} to be instrumented by $P_{read}$.
-There are now two versions of \texttt{memset} in the code cache: one specialized for executing outside of RCU-related code, and the other specialized to executing within read-side critical sections.
-\end{enumerate*}
-
-%For example, our \toolname{RCUdbg} invokes a policy switch from policy $P_{\mathit{null}}$ to policy $P_{\mathit{read\_critical}}$ when $P_{\mathit{null}}$'s instrumentation function observes a \texttt{call} instruction to the \texttt{rcu\_read\_lock} Linux kernel function.
-A similar switch occurs from $P_{\mathit{read\_critical}}$ back to $P_{\mathit{null}}$ at a \texttt{rcu\_read\_unlock}.
-Specialization that distinguishes between the inside and outside of read-side critical sections allows \toolname{RCUdbg} to do heavy-weight checks on functions like \texttt{memset}, but only when \texttt{memset} is invoked from within a read-side critical section.
-
-The \toolname{RCUdbg} example highlights how the current execution context is implicity tracked by the current program counter or instruction pointer.
-Before the \texttt{call} to \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock}, the instructions were instrumented by $P_{outside}$.
-However, after the \texttt{call}, the instructions were instrumented by $P_{read}$.
-This insight reveals that policy switches and control-flow transfers behave like state transitions in a finite state machine, and that the program counter, by virtue of pointing at an instruction instrumented by a specific policy, tracks the current pogram state.
-This analogy, however, reveals a potential limitation: if RCU read-side critical sections are nested, then \toolname{RCUdbg}'s 3-state policy switching approach (as described) cannot distinguish a \texttt{call} to \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock} from within an unnested and a nested read-side critical section.
-\toolname{RCUdbg} solves this problem by having more than one read policies: $P_{\mathit{{read}_0}}$, $P_{\mathit{{read}_1}}$, ..., $P_{\mathit{{read}_N}}$, which are generated on demand.
-In practice, this does not lead to code cache bloat because RCU read-side critical sections do not nest more than two times.
-
-By default, Granary doesn't allow tools to specify policy switches on procedure return instructions (\texttt{ret}).
-This increases the ``power" of the formal model of policy switching from being a deterministic finite state automaton to being a restricted form of pushdown automaton.
-The program counter continues to track the current execution state, and saved program counters (returns addresses stored on the runtime call stack) record prior execution states.
-In this model, \texttt{call}s place contextual breadcrumbs (in the form of return addresses) on the runtime call stack, and \texttt{ret}s read these breadcrumbs to return to a previous context.
-For many tools, this is a natural model of state tracking as it is bound to the current thread, and because kernel code tends to follow a good call/return discipline.
-
-%\toolname{RCUdbg} solves this problem by tracking the nesting depth using different read-policies ($P_{read}_0$, $P_{read}_1$, ...).
-
-%Because policies name an execution context, they also represent states in a finite state machine.
-That is, code cache execution is in the state named by a policy if the  executing code was instrumented by that policy.
-A state transition occurs when control transfers from code instrumented by one policy to code instrumented by another policy.
-A limitation with this approach is that a single state does not encode the sequence of previous states that led to execution being in the current state.
-For example, RCU permits nested read-side critical sections.
-If two read-side critical sections are nested then switching from $P_{\mathit{read\_critical}}$ to $P_{\mathit{null}}$ on the first \texttt{call} to \texttt{rcu\_read\_unlock} meant that our tool would lose track of being in the context of the outer read-side critical section.
-\toolname{RCUdbg} solves this problem by tracking the nesting depth of RCU read-side critical sections using policies, e.g.
-$P_{\mathit{read\_critical}_1}$, $P_{\mathit{read\_critical}_2}$, etc.
-
-%An alternative and sometimes more powerful model for policy switching is to prevent policy switching on function return instructions (e.g.
-either through manually specifying a policy switch, or by inheriting the policy from the basic block containing the \texttt{ret} instruction).
-In most cases, function returns are natural policy reverting points.
-This is because the current instruction pointer (program counter) encodes the current execution context (by being in a basic block instrumented by some policy), and the code cache return addresses stored on the runtime call stack encode all previous execution contexts.
-This asymmetry between \texttt{call}s and \texttt{ret}s yields a more powerful context-tracking state machine: \texttt{call}s place contextual breadcrumbs (in the form of return addresses) on the runtime call stack, and \texttt{ret}s read these breadcrumbs to return to a previous context.
-Policy switches under this lens behave similarly to state transitions in a pushdown automaton; \texttt{call} instructions push a new state onto the stack for each function, \texttt{ret} instructions pop the current function's state from the stack, and other CTIs induce state transitions within the current function by altering control flow.
-
-Granary implements policy tracking and switching by encoding policy information into the meta-data of basic blocks, the instructions of in-edge code generated for indirect control flow, and the edge data structures used by direct control flow.
-When an instrumented CTI executes for the first time, it yields control to Granary with the CTI target and policy information as inputs.
-Granary decodes and instruments the targeted instructions according to the input policy information.
-
-% CTIs of basic blocks.
-\Cref{fig:direct_edge_code} shows an example of how policy information is directly recorded into control-transfer edge code.
-If a policy switch is not specified on an instrumented CTI then that CTI inherits the policy used to instrument the basic block containing the CTI.
-
-\subsection{Partial Specialization Using Policy Properties}
-
-Granary internally specializes code at a finer granularity than that described by a tool's instrumentation policies.
-Granary will sometimes generate multiple versions of a given basic block, even for the same instrumentation policy.
-Other times, Granary will only generate extra entries in the code cache index that are eventually disambiguated.
-When apparently redundant code generation occurs, it means that certain properties that are internally tracked by Granary are present, and that these properties affect how Granary or its tools can optimize or instrument code.
-
-For example, Granary tracks whether code may or may not access user space data, because this code can potentially result in a page fault.
-Granary can detect this code using a combination of pattern recognition, and policy property propagation.
-This internal mechanism is needed to correctly handle page faults in instrumented kernel code (\Cref{sec:interrupt_return_address_transparency}), and is also useful for optimizing code generated by tools as well.
-For example, Granary's \toolname{BoundsChecker} tool alters the 16 high order bits of addresses returned from kernel allocators, to make it easy to associate bounds information with each allocated object \cite{BehaveOrBeWatched}.
-Checked and unchecked kernel addresses can distinguished by testing only a single bit.
-However, user space addresses and checked addresses are harder to distinguish, requiring two bits to be checked.
-Fortunately,  few instructions dereference user space addresses, and Granary tracks whichs basic blocks might contain those instructions, thus allowing the \toolname{BoundsChecker} to avoid many unnecessary bit tests.
-
-Another example of partial specialization that benefits instrumentation tools is Granary's SIMD register tracking property.
-Granary detects and tracks the usage of SIMD registers (\texttt{\%xmm0} through \texttt{\%xmm16}).
-This informs instrumentation tools on when its safe to save transient data to the SIMD registers instead of to memory, which can result in improved performance \cite{Minemu}.
-
-%An access to user space data can potentially cause a fault in the kernel .
-Code that accesses user space data contains specify binary patterns; however, sometimes those patterns are not present in the individual basic blocks that dereference the user space address.
-
-%For example, Granary's address watchpoints framework uses address tainting to easily associate meta-data to allocated objects, and to make it easy to distinguish between tracked and untracked objects \cite{BehaveOrBeWatched}.
-The mechanism used to taint an address depends on x86-64 ``non-canonical" addresses.
-A canonical address has all of its 16 high-order bits set to 0 or set to 1.
-In Linux, the 16 high-order bits of kernel addresses are all 1, and the 16 high-order
-
-%These propeties, called policy properties, transparently persist across control-flow instructions, even if the CTI changes the current policy. Granary tracks the following policy properties:
-
-%\begin{description}
-%	\item[Temporary]
-%	\item {\bf Kernel vs. module code.}
-%	\item 
-%	\item {\bf SIMD register usage.}
-%\end{description}
-
-\section{Application: Control-Flow Graphs}\label{sec:cfg}
-Granary's \toolname{cfg} tool constructs inter- and intra-procedural control flow graphs from the instrumented kernel execution.
-The \toolname{cfg} tool can also act as a code coverage and profiling tool by recording basic block and conditional branch execution counts.
-\Cref{fig:lmbench} evaluates the performance impact of using the \toolname{cfg} tool for profile-guided optimization.
-First, \toolname{cfg} instruments a single run of the \texttt{lmbench} microbenchmark suite.
-Then, the recorded profile data is used to guide the ahead-of-time tracing infrastructure (\Cref{sec:aot}) on the evaluated runs of \texttt{lmbench}.
-
-\toolname{cfg}'s instrumentation introduces counter increments into basic blocks, counter increments following conditional branches, and target recorders for indirect CTIs (\texttt{call}, \texttt{jmp}).
-Each indirect CTI recorder maintains a variable-sized array of all observed targets for the given CTI.
-The recording process follows either a fast or a slow path.
-In the fast path, the target is compared with a cached version of the most recent target.
-If they are the same then no further action is taken, as this implies that the target has already been recorded.
-Most executions ``hit" in this cache because most indirect branches in the kernel only have a single target.
-If the target and the cached target differ, then the cached target is immediately updated and the slow path is executed.
-The slow path checks to see if the target has been permanently recorded, and if not, records it.
-
-Control-flow graphs are reported in terms of basic blocks (nodes) and edges (control flow transfers).
-The reporting operates by iterating over each basic block and performing the following for each basic block: \begin{enumerate*}
-	\item Decode the native instructions associated with the basic block. The location and number of instructions to decode is found in the basic block's meta-data (\Cref{sec:metadata}).
-	\item For each decoded direct CTI, print out the target PC of the CTI.
-If the CTI is a conditional CTI, then print out two PCs: the target PC, and the fall-through PC.
-If conditional branch counting is enabled, then the value of the counter associated with the fall-through edge is also printed.
-	\item For each decoded indirect CTI, print out all target PCs from the associated indirect CTI recorder.
-\end{enumerate*}
-
-The reported data can be re-integrated into Granary to drive profile-guided optimizations, or post-processed into a format for use by tools like KCacheGrind for profile data visualization \cite{KCacheGrind}.
-
-%A single-element cache is also maintained that records the most recent indirect branch target.
-Most indirect CTIs in the kernel only have one target, and so once the target has been permanently recorded, all future executions hit the cache.
-
-%As an optimization, the indirect target recorders maintain a one-element cache of the most recent target of a given indirect branch
-
-% Basic blocks in the graph can be labeled with execution counts, and control-flow edges between basic blocks can also be labeled with execution counts.
-Counting basic block and edge executions enables Granary's \toolname{cfg} tool implicitly provides code coverage and profiling information.
-\Cref{fig:lmbench} evaluates how profiling information from Granary's \toolname{cfg} tool is integrated with Granary's ahead-of-time translation optimization (\Cref{sec:aot}).
-
-\section{Application: RCUdbg}\label{sec:rcudbg}
-
-Read-Copy-Update (RCU) is a scalable, lock-free synchronization mechanism used in the Linux kernel \cite{RCU,RCUInLinux}.
-When correctly used, RCU permits concurrent reads and writes to shared data structures, but requires that writers synchronize among themselves.
-Like other synchronization mechanisms, RCU has a deceptively simple API.
-Misuses of the API can result in hard-to-find bugs that can be implementation-specific (i.e.
-bugs that manifest in one version of RCU but not another) or architecture-specific (e.g.
-bugs that only exhibit on architectures with a relaxed memory model).
-For example, RCU bugs sometimes manifest as use-after-free bugs, which can result in silent data corruption (if the underlying memory remains used) or kernel panics.
-
-% (assuming that pointer reachability is sufficient to determine semantic reachability in this case).
-
-Granary's \toolname{RCUdbg} tool is designed to catch bugs related to misuses of pointers to RCU-protected data structures.
-These pointers, called RCU-protected pointers, link together discrete objects within larger RCU-protected data structure.
-For example, in an RCU-protected linked list, the RCU-protected pointers are the \texttt{next} pointers that connect list elements.
-An RCU-protected pointer is like a normal pointer, but points to a potentially incomplete or partially-constructed data structure.
-Data structures can be potentially incomplete or partially constructed because RCU permits concurrent reads and writes to the shared data structure.
-Because of this, reads can sometimes see an inconsistent version of the data structure.
-For example, if a writer modifies the \texttt{next} pointer of an RCU-protected linked list, then a concurrently executing reader may see either the old value or the new value of the \texttt{next} pointer.
-This is further complicated by multi-step updates (e.g.
-adding or removing an element from a doubly-linked list).
-%For example, adding an element to a doubly-linked list requires two separate pointers to be updated.
-Concurrently executing readers can therefore observe the list in one of several intermediate states.
-
-%
-%\toolname{RCUdbg} infers that some pointers are RCU-protected pointers based on how specific RCU API functions are used by the instrumented code.
-When certain API functions are invoked, \toolname{RCUdbg} knows that some of the function's arguments or return values are RCU-protected pointers, and therefore treats them as such.
-For example, the RCU API includes two functions for operating on RCU-protected pointers: \begin{inparaenum}[i)]
-%	\item \texttt{rcu\_\linebreak[0]dereference} for reading the value of an RCU-protected pointer, and;
-%	\item \texttt{rcu\_\linebreak[0]assign\_\linebreak[0]pointer} for assigning a value to an RCU-protected pointer. 
-%\end{inparaenum}
-%
-The following requirements describe how code can use the RCU API to read from and modify a data structure like an RCU-protected linked list.
-
-% with eachother and operate on RCU-protected pointers.
-
-%There are two kinds of RCU-protected pointers: read-only pointers, and read/write pointers.
-Read-only pointers can only be read from within RCU read-side critical sections, delimited by  \texttt{rcu\_\linebreak[0]read\_\linebreak[0]lock} and \texttt{rcu\_\linebreak[0]read\_\linebreak[0]unlock}.
-Read/write pointers can be written to
-
-%Code operating on RCU-protected pointers must use specific API functions: , and By assuming that these functions are correctly used some of the time, we can
-
-%Therefore, a use of \texttt{rcu\_\linebreak[0]dereference} implies that the pointer being read should be RCU-protected, and the the value returned should also be RCU-protected.
-Without an \emph{a priori} understanding of the data structures being operated on, these two API functions give \toolname{RCUdbg} enough information about RCU protected pointers to find bugs.
-
-%Without a prior understanding of what pointers are RCU-protected pointers, we can assume that any pointer operated on by one the aforementioned API functions is an RCU-protected pointer.
-
-%Without an \emph{a priori} understanding of the structure of the data being syncrhonized using RCU, we can conservatively assert that RCU-protected pointers.
-
-%the associated memory may or may not be free, but the data structure occupying that memory can't be proven to exist. This 
-%  However, this class of RCU bugs is not well-suited to being detected by a use-after-free memory checker tool
-%RCU is notoriously hard to use, despite its relatively simple API.
-RCU usage bugs, like bugs with other synchronization mechanisms, often stay hidden and only exhibit themselves for very specific thread interleavings/schedules.
-% go silently unnoticed because the memory 
-%While bugs related to mutual exlusion locks might result in easily observable deadlocks, bugs related to RCU typically result in use-after-free errors, which sometimes go silently unnoticed, sometimes result in kernel panics, and most of the time work as expected because the associated data is .
-
-\begin{enumerate}
-%	\item Readers must tolerate an inconsistent view of the shared data structure.
-That is, a reader operating concurrently with a writer might observe either the version of the data before or after the writer's changes have been published.
-This can be complicated when writers perform multiple, discrete updates to a data structure.
-One example of this is a doubly linked list: a writer that is adding a new entry into a doubly linked list must update at least two RCU-protected pointers.
-
-	\item Readers must operate within \emph{read-side} critical sections, which are delimited by \texttt{rcu\_read\_lock} and \texttt{rcu\_read\_unlock}, respectively.
-
-	\item Within read-side critical sections, readers can only read a shared data structure after obtaining a pointer to that shared data structure using the \texttt{rcu\_dereference} API function.
-This API function operates on an RCU-protected pointer and returns a read-access-only version of the pointer.
-The reader can use the read-access pointer to safely read from the shared data structure within the remainder of its read-side critical section.
-If \texttt{rcu\_dereference} is not used, then the reader might observe a partially-constructed (i.e.
-incomplete) version of the shared data structure.
-Notably, accessing the original data structure and not the dereferenced data structure pointer--even after an \texttt{rcu\_dereference}--is undefined behavior.
-That is, within a read-side critical section, RCU only guarantees the ``completeness" of data structures pointed to by the read-access-only pointers obtained from  \texttt{rcu\_dereference}.
-Data accessed within a read-side critical section is \emph{not} guaranteed to exist outside the read-side critical section.
-
-	\item Writers must synchronize among themselves before any writes to a shared data structure are performed.
-Synchronization is typically achieved via mutual exclusion locks, spin locks, or by having a single, designated writer thread.
-We say writes to the shared data structure occur within \emph{write-side} critical sections.
-
-%	\item RCU defines how writers can update RCU-protected pointers within a shared data structure.
-Modification of non-RCU-protected pointer fields within an RCU-protected data structure is beyond the scope of RCU's correctness guarantees.
-
-	\item Writers wishing to update an RCU-protected pointer in a shared data structure must use the \texttt{rcu\_assign\_pointer} API function.
-\texttt{rcu\_assign\_pointer} assigns a new value to an RCU-protected pointer in a shared data structure.
-
-	\item Readers must access any pointer field operated on by an \texttt{rcu\_\linebreak[0]assign\_pointer} using \texttt{rcu\_\linebreak[0]dereference}.
-For example, if the \texttt{next} pointer field in \emph{any} list element of an RCU-protected list has been assigned to by \texttt{rcu\_assign\_\linebreak[0]pointer}, then \emph{all} \texttt{next} pointer fields must be accessed by readers of that list by using \texttt{rcu\_\linebreak[0]dereference}.
-In this sense, using \texttt{rcu\_\linebreak[0]dereference} and \texttt{rcu\_assign\_\linebreak[0]pointer} marks specific variables and data type fields as containing RCU-protected pointers.
-
-\end{enumerate}
-
-% NOTE!!!!
-%
-% All of those spaces before /* BUG! */ and such are so that the lstlisting will be *close* to the columnwidth!! Somehow linewidth param of lstset didn't work for me :-(
-
-\lstset{
-	language=C,
-	tabsize=2,
-	stepnumber=1,
-	morekeywords={rcu_read_lock,rcu_dereference,rcu_read_unlock,rcu_assign_pointer,spin_lock,spin_unlock}}
-
-% Example code showing an existence bug.
-\newsavebox\rcuexistencebug
-\begin{lrbox}{\rcuexistencebug}
-\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-rcu_read_lock();
-  elm = rcu_dereference(list_head);
-  data = elm->data;                         /* Valid */
-rcu_read_unlock();
-data = elm->data;                            /* BUG! */
-\end{lstlisting}
-\end{lrbox}
-
-% Example code showing another existence bug.
-\newsavebox\rcuderefbug
-\begin{lrbox}{\rcuderefbug}
-\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-rcu_read_lock();
-  elm = rcu_dereference(list_head);
-  if(elm != NULL)
-    data = list_head->data;                  /* BUG! */
-rcu_read_unlock();
-\end{lstlisting}
-\end{lrbox}
-
-% Example code showing another existence bug.
-\newsavebox\rcuassignderefbug
-\begin{lrbox}{\rcuassignderefbug}
-\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-spin_lock();
-  elm = kmalloc(...);
-  elm->next = list_head->next;
-  rcu_assign_pointer(list_head->next, elm);
-spin_unlock();
-...
-rcu_read_lock();
-  elm = rcu_dereference(list_head);
-  while(elm != NULL) {
-    data += elm->data;
-    elm = elm->next;                         /* BUG! */
-  }
-rcu_read_unlock();
-\end{lstlisting}
-\end{lrbox}
-
-\begin{figure}[ht!] %
-\subfloat[\label{fig:rcu_existence_bug}Despite correctly obtaining a pointer to a shared data structure using \texttt{rcu\_dereference} within a read-side critical section, the data pointed to by \texttt{elm} is not guaranteed to exist after the read-side critical section ends.]{\fbox{\usebox\rcuexistencebug}} %
-
-\subfloat[\label{fig:rcu_deref_bug}Despite the call to \texttt{rcu\_dereference}, \texttt{list\_head} is not guaranteed to be equivalent to \texttt{elm} because a writer might have concurrently changed the value of the \texttt{list\_head} pointer.
-That is, if \texttt{elm} is non-\texttt{NULL}, then the data pointed to by \texttt{elm} is guaranteed to exist for the remainder of the read-side critical section.
-However, the value of \texttt{list\_head} is \emph{not} guaranteed to remain consistent--even within read-side critical sections--and so \texttt{list\_head} might point to a \texttt{NULL} pointer when \texttt{elm} does not.]{\fbox{\usebox\rcuderefbug}} %
-
-\subfloat[\label{fig:rcu_assign_bug}This example shows the code for a writer thread adding a new element into the second position of an RCU-protected linked list.
-The code for reader threads is also shown, where readers iterate over the list by traversing \texttt{next} pointers and sum up the \texttt{data} fields of each visited list element.
-Unfortunately, the access of \texttt{elm->next} results in undefined behavior.
-This is because the next element of the list is only guaranteed to be fully constructed within the read-side critical section if the reader executes \texttt{rcu\_dereference(elm->next)}.
-Therefore, the first read of \texttt{elm->next} might yield a pointer to partially constructed data.
-Because that data is not guaranteed to be fully constructed, the next read of \texttt{elm->next} (i.e.
-\texttt{elm->next->next}) might read an invalid pointer from uninitialized memory.
-Treating the value of uninitialized memory as a valid pointer could trigger a kernel panic or silent data corruption.]{\fbox{\usebox\rcuassignderefbug}} %
-\caption{\label{fig:rcu_bugs}Example RCU bugs related to misuses of RCU-protected pointers.}
-\end{figure}
-
-\toolname{RCUdbg} is primarily concerned with finding bugs related to RCU-protected pointer misuses by reader threads (both inside and outside of read-side critical sections).
-As mentioned above, RCU bugs sometimes manifest as use-after-free bugs.
-Unfortunately, a pure use-after-free-based checking approach is insufficient.
-\Cref{fig:rcu_assign_bug} is an example where a use-after-free based checking approach will miss an RCU-protected pointer misuse because the bug does not manifest as a use-after-free.
-\Cref{fig:rcu_existence_bug} is an example of an RCU usage bug that can manifest as a use-after-free bug, but only if one is unlucky enough to observe a very specific thread schedule.
-That is, suppose that execution is interrupted after \texttt{rcu\_read\_unlock}, but before the access to \texttt{elm->data}.
-If \texttt{elm} is freed before the interrupted code resumes its execution, then the access to \texttt{elm->data} is a use-after-free bug.
-This kind of bug is a time bomb waiting to go off.
-Fortunately, \toolname{RCUdbg} catches this bug and others in a scheduling-independent way.
-
-%\Cref{fig:rcu_bugs} shows some example usages of RCU, and some of the more subtle pointer-related bugs that can happen because of the interaction of rules (1), (2), (3), and (7).
-
-\toolname{RCUdbg} detects the the following RCU bugs related to misuses of RCU-protected pointers.
-\begin{itemize}[leftmargin=3.2em]
-	\item[\bf B1] Reads through an RCU-dereferenced pointer outside of a read-side critical section (\Cref{fig:rcu_existence_bug}).
-	\item[\bf B2] Reads through an RCU-dereferenced pointer in the wrong read-side critical section.
-
-	\item[\bf B2] Reads through an RCU-assigned pointer that don't first use \texttt{rcu\_dereference} (\Cref{fig:rcu_deref_bug}).
-	\item[\bf B3] Reads through fields of RCU-protected data structures that should be accessed via \texttt{rcu\_dereference} (\Cref{fig:rcu_assign_bug}).
-
-	\item[\bf B5] Writes through an RCU-dereferenced pointer.
-	\item[\bf B6] Writes through an RCU-assigned pointer by a non-write thread.
-\end{itemize}
-
-Bugs B1 and B2 relate to whether or not we can prove that a given RCU protected data structure exists.
-Bugs B2 and B3 relate to whether or not a given RCU protected data structure has been fully constructed.
-Finally, bugs B5 and B6 relate to invalid uses RCU-protected and RCU-unprotected pointers, where correct synchronization among writers is not maintained.
-
-% or is constructed at a given program point. Bugs B2 through B
-
-% of this class of RCU bugs is proof-of-existence or proof-of-constructedness bugs.
-That is, an RCU-protected data structue exists if it is \emph{reachable} (i.e.
-by at least one reader thread), and if it is \emph{fully constructed} (i.e.
-all fields or components of the data structure have been initialized).
-
-\toolname{RCUdbg} gains visibility on RCU operations in two ways: \begin{enumerate}
-	\item The Linux kernel RCU implementation is annotated to provide more information to \toolname{RCUdbg}.
-This information includes source code file and line information for bug reporting, and additional memory references that \toolname{RCUdbg} uses to get better visibility on bugs.
-These annotations are strictly internal to the RCU API, and thus require no changes to code using this API.
-	\item The annotated RCU API functions are specialized using Granary's function wrapping feature (\Cref{sec:function_wrapper}). Specifically, 
-\end{enumerate}
-
-%\toolname{RCUdbg} interposes on an annotations are added into the Linux kernel's RCU implementation. These annotations
-
-%uses Granary's address watchpoints feature \cite{BehaveOrBeWatched} to ``taint" addresses participating in RCU-protected data structures.
-
-%Function wrappers are used to specialize the execution of \texttt{rcu\_\linebreak[0]dereference} and \texttt{rcu\_assign\_\linebreak[0]pointer}:
-
-%\begin{figure}[t!]
-%\lstset{language=C, tabsize=2, stepnumber=1}
-%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-%// As used in the code:
-%//   p = rcu_dereference(q);
-%// Internally annotated to become:
-%//   p = __rcudbg_deref(&q, q);
-%// Wrapper:
-%FUNCTION_WRAPPER(__rcudbg_deref, (q_addr, q), {
-%  if(is_watched_address(q_addr)) {
-%    // Make sure q_addr
-%  } else {
-%
-%  }
-%  return add_watchoint(q, DEREF);
-%})
-%\end{lstlisting}
-%\caption{}
-%\label{fig:rcu_dereference_wrapper}
-%\end{figure}
-
-%\begin{center}
-%\begin{minipage}{0.6\columnwidth}
-%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-%elm = rcu_dereference(list_head);
-%\end{lstlisting}
-%\end{minipage}
-%\end{center}
-%
-%Becomes:
-%
-%\begin{center}
-%\begin{minipage}{0.8\columnwidth}
-%\begin{lstlisting}[basicstyle=\scriptsize\ttfamily]
-%elm = deref_taint(rcu_dereference(list_head));
-%\end{lstlisting}
-%\end{minipage}
-%\end{center}
-%
-%Tainted addresses are introduced in two ways:
-%\begin{description}
-%	\item[\texttt{rcu\_assign\_pointer}] operates on two pointers: 
-%	\item[\texttt{rcu\_dereference}] operates on a potentially tainted pointer (as introduced by \texttt{rcu\_assign\_\linebreak[0]pointer}, and returns a deref-tainted pointer.
-%\end{description}
-%
-
-%Unfortunately, approaching RCU bug detection from the perspective of finding use-after-free errors is insufficient.
-Use-after-free detection attacks a symptom of the problem that is highly dependent on thread interleavings and scheduling.
-That is, RCU can be misused in a way that results in very unlikely, but potential use-after-free bugs.
-An example of this case is when a writer thread, operating on a RCU-protected linked list, unlinks an element and queues its memory for later freeing.
-RCU provides a mechanism whereby readers can safely access this element for a brief period of time (called a read-side critical section).
-When the memory is eventually freed depends is implementation-defined, but depends on other reader activity.
-
-%How soon after a read-side critical section ends
-
-% There are many cases readers can access an RCU-protected data structure because they have not been freed \emph{yet}
+%\section{Introduction}\label{sec:intro}
 
@@
 
-% a data structure cannot be proven to exist at some potentially allocated memory.
-Existence in this case relates to semantic reachability.
-For example, if a list element $E_n$ from an RCU-protected list is reachable from the head of the list by traversing zero-or-more intermediate elements $E_0, ..., E_{n-1}$, then $E_n$ is said to exist.
-While related, memory leak detection (e.g.
-via conservate mark and sweep) is also insufficient for ladidadidaaaa TODO TODO TODO
+%Existing virtual register implementations  punt on generality and require
+%\begin{inparaenum}[i)]
+%	\item that all code be comprehensively instrumented,
+%	\item that the instruction set or OS be virtualized, and
+%	\item the availability of dedicated, thread-local storage.
+%\end{inparaenum} These re
 
-\begin{figure*}
-\includegraphics[height=14em,clip=true,trim=0 -5em 10em -5em]{charts/lmbench_legend.eps}
-\subfloat[Fast operations]{\epsfig{file=charts/lmbench_fast.eps,height=14em}}
-\subfloat[Fork operations]{\epsfig{file=charts/lmbench_fork.eps,height=14em}}
-\caption{\label{fig:lmbench}Microbenchmark results from \texttt{lmbench} that shows the effect of Granary's code cache optimizations (\Cref{sec:optimize}).}
-\end{figure*}
+%In this work, we present a generic virtual register system for a complex instruction set (x86-64). Our virtual register system is general in that it
+%\begin{inparaenum}[i)]
+%	\item does not require that all code be instrumented,
+%	\item works on user and kernel space,
+%	\item operates directely on x86-64 binary instruction,
+%	\item does not assume the availability of 
+%\end{inparaenum}
 
-\section{Evaluation}\label{sec:eval}
+%\section{Related Work}\label{sec:related}
 
-Before examining applications that are enabled by Granary, \Cref{sec:microbench} evaluates the overhead of the baseline instrumentation.
-Then, \Cref{sec:macrobenchmark} does something with the evaluation TODO.
+%\section{Evaluation}\label{sec:eval}
 
-\subsection{Microbenchmark}\label{sec:microbench}
-For these experiments, we operate with a large enough code cache (\texttildelow 4MB) so that flushes do not occur.
-We perform a number of stress tests with a high level of kernel activity, to show the robustness of Granary and to illustrate worst-case behavior of the baseline instrumentation system.
-These tests ran on a Lenovo X230 laptop with 8GB of RAM and four Intel\textregistered\ Core\texttrademark\ i5 CPUs operating at 2.50GHz.
+%\section{Conclusion}\label{sec:conclusion}
 
-\Cref{fig:lmbench} shows the results of running the \texttt{lmbench} microbenchmark, which compares the overhead of Granary's baseline instrumentation (with and without optimizations) to native execution.
-The results highlight that the biggest performance boost is achieved with aggressive ahead-of-time tracing (\Cref{sec:aot}), and that profiling information from Granary's \toolname{cfg} tool can improve on these results by guiding which branches the tracer follows.
-
-The trace allocation optimization (\Cref{sec:trace_alloc}) does not result in significant performance improvements, and sometimes results in worse performance.
-This is because Granary's default allocation strategy (allocating space for instructions from per-CPU memory pools) already results in similar memory layouts to the trace allocator.
-A more significant difference would  be observed if Granary defaulted to a global allocation mechanism instead.
-Also, Granary's inclusion of function call instructions (\texttt{call}) into basic blocks likely works against the trace allocator.
-If basic blocks ended at \texttt{call} instructions, then the trace allocation strategy would arrange for the returning basic block and the returned-to basic blocks to be adjacent in memory.
-
-%This is because the memory layout of basic blocks with and without the trace allocation optimization ends up being similar in Granary.
-
-\subsection{Macrobenchmark}\label{sec:macrobenchmark}
-
-% when profiling data from Granary's \toolname{cfg} tool are used to predict and potentially exchange the targets of conditional branches.
-
-% We evaluated behavioral watchpoints using a microbenchmark and measured the performance of common filesystem operations on an in-memory disk.
-Our tests ran on a desktop equipped with an Intel\textregistered\ Core\texttrademark\ 2 Duo 2.93 GHz CPU with 4GB physical memory.
-%Watchpoint instrumentation was enabled for every memory operation.
-We measured the overhead of the watchpoint framework using a null client and heap-based buffer overflow detector.
-%We also evaluated the performance of heap-based buffer overflow detector developed using watchpoints.
-
-\section{Related Work}\label{sec:related}
-Prior work can be classified as either whole-kernel/system DBT, or probe-based kernel instrumentation. We discuss how Granary differs from these approaches.
-\paragraph{Whole-kernel/system DBT}
-PinOS \cite{PinOS} is a port of the Pin \cite{Pin} DBT framework that performs whole-system instrumentation.
-It has high overheads and depends on hardware virtualization, which prevents it from instrumenting non-virtualizable kernel modules.
-
-QMEU is a flexible, whole-system emulator with kernel-level paravirtualization support (using KVM) \cite{QEMU}. Like PinOS, QEMU cannot instrument non-virtualizable device drivers.
-
-DRK \cite{DRK} is a kernel space port of the DynamoRIO \cite{DynamoRIO} DBT framework.
-DRK instruments the entire kernel and all device drivers/modules.
-DRK follows a strict transparency model, which limits the flexibility of instrumentation and increases overheads.
-
-Btkernel is a fast and scalable system for the kernel \cite{btkernel}.
-Like Granary, but unlike DRK, btkernel applies a relaxed transparency model when instrumenting the kernel.
-For exmaple, code cache addresses are exposed to the instrumented kernel in the form of function call return addresses and interrupt return addresses.
-Unlike Granary, btkernel operates using an instruction translation ``rulebook", which limits instrumentation to being defined ahead-of-time, on a per-instruction basis.
-Btkernel does not support any form of runtime code specialization.
-That is, like DRK, btkernel instruments all code, but cannot change the granularity of instrumentation or the instrumentation policy applied to code.
-
-TODO: JIFL
-
-\paragraph{Probe-based Instrumentation} Several systems support injecting code at specific locations within kernel code.
-
-TODO: SystemTap
-
-KernInst \cite{KernInst} can inject code at almost any location in an unmodified commodity OS.
-By default, KernInst uses debugging information normally present in kernel binaries to specify probe-points; however, absolute memory addresses can also be provided.
-
-Other examples with similar or more restricted functionality include LTTng \cite{LTTng}, KProbes \cite{KProbes}, DProbes \cite{DProbes}, and ftrace \cite{ftrace}.
-These systems are unable to perform fine-grained instrumentation of instructions or memory.
-
-\section{Conclusion}\label{sec:conclusion}
-%We created Granary to address the challenges of instrumenting binary kernel modules.
-Granary provides mixed-mode execution to remove the overheads of DBT for uninstrumented kernel code.
-A relaxed transparency model further improves performance and allows greater visibility into the interactions between module and kernel code.
-Granary also supports policy-driven instrumentation, which allows tools to specialize their instrumentation based on the context in which code executes.
-Finally, Granary exposes high-level static analysis information to dynamic instrumentation tools, making it easy to match low-level memory reads with specific fields in data structures at the source code level.
-
-%Together, Granary's features simplify the development of powerful kernel module analysis tools, while delivering lower overheads than previous kernel dynamic binary instrumentation solutions.
-
-\bibliographystyle{acm}
-\bibliography{library}
+%\bibliographystyle{acm}
+%\bibliography{library}
 

### 2014-12-17 HotOS'15/paper.tex
Improved the diagrams. Still working on methodology :-(

@@
 \usepackage{refstyle,amsmath,chngcntr}
+\usepackage{epstopdf}
 \usepackage{epsfig,subfigure,framed}
@@
 \usepackage{hyperref}
+\usepackage{dblfloatfix}
 
@@
 
-In this paper, we describe \Toolname, a system that detects and diagnoses cache line contention in arbitrary binary programs.
-\Toolname{} refines existing contention detection approaches in several novel ways. First, we describe a method training
-a detection system to focus on only the code and data that matter to the problem. Second, we describe a new form of
-data-centric sampling: comprehensive data sampling. Our sampling approach uses unlimited watchpoints to uniformly
-sample memory locations according to the type of data stored in memory. This avoids sampling bias where a large portion
-of a program's memory is unshared.
+In this paper, we describe \Toolname, a general framework for detecting and diagnosing cache line contention in arbitrary,
+binary programs. \Toolname{} refines existing contention detection approaches in several novel ways. First, we describe a
+method training a detection system to focus on only the code and data that matter to the problem. Second, we describe a
+new form of data-centric sampling: comprehensive data sampling. Our sampling approach uses unlimited watchpoints to
+uniformly sample memory locations according to the type of data stored in memory. This avoids sampling bias where a large
+portion of a program's memory is unshared.
 
@@
 % --------------------------------------
-\subsection{Ownership Tracking}
-\begin{figure}
+\subsection{Training}
+
+% --------------------------------------
+\subsection{Shadow Memory}
+\begin{figure*}
+\includegraphics[width=\textwidth]{shadow.pdf}
+\caption{\label{fig:layout_of_shadow}Layout of shadow memory. Native cache lines map to individual shadow state cells. %
+ In the above %
+diagram, the cache line associated with bytes [\texttt{0xF00},\texttt{0xF40}) are being sampled using $SampleId=11$. When %
+we look at entry $11$ in the sample table, we see that the monitor thread chose this sample using a recent allocation of an %
+object with $TypeId=4$.}
+\end{figure*}
+
+\Toolname{} reserves one sixteenth of the virtual address space for shadow memory \cite{Umbra}, which it uses to track the watched and
+ownership state of every cache line. The majority of this space is never backed by physical memory due to a lazy memory allocation
+scheme. Also, unlike prior shadow memory-based approaches \cite{DrContention}, our sampling approach enforces a strict upper
+bound on the amount of physical memory needed for shadow memory\footnote{The upper bound on shadow memory usage is one
+memory page per sample. Because each shadow page can cover up to 1024 cache lines, having multiple samples map to cells on the
+same shadow page is common.}.
+
+% TODO: This isn't really true, but it could be true if we we had an index of what pages have samples, and map all other pages to the
+% sample underlying zero memory somehow...
+
+For each sampled cache line of native memory, \Toolname{} stores a $\langle SampleId,ThreadId \rangle$ pair in shadow memory.
+Initially, every cache line is ``unwatched", as represented by $SampleId=0$ in \Figref{states_of_shadow}. When a new sampling period
+begins, the monitor thread ``watches`` caches lines by assigning those lines a unique, non-zero $SampleId$. \Figref{layout_of_shadow}
+shows how the $SampleId$ stored in shadow memory associates with information about what was sampled and what threads contended on
+the cache line. The information stored for each sample include: \begin{itemize}
+
+	\item Type of object (in terms of a $TypeId$) affected by contention on the cache line. Using a secondary lookup table, we can
+determine the size and allocation context of this object.
+
+	\item Address of allocated object, and range of bytes within the allocated object that are located on the sampled cache
+line. Sampling a subsequence of bytes happens when a single object spans multiple cache lines, either due to alignment or due to size. When
+this happens, multiple samples for the different parts of the same object are taken. This allows us to perform both object-granularity
+\emph{and} field-granularity contention detection \cite{RaceTrack}.
+
+	\item Stack traces of the thread that first acquires ownership of the cache line, and of the first detected contenter thread.
+
+	\item Accessed addresses, memory operand sizes, and memory operand types (read, write, atomic) of both threads.
+\end{itemize}
+
+% --------------------------------------
+\subsection{Instrumentation}
+If a sampled memory location is accessed, then the instrumentation instructions of \Figref{inline_assembly} will detect the non-zero
+$SampleId$, and invoke the $CheckContention$ function if the current thread does not already appear to own the sampled cache line.
+The $CheckContention$ algorithm (\Figref{algorithm}) is responsible for acquiring ownership of a cache line, as well as detecting
+contention. At a high level, the algorithm detects contention by failing to acquire ownership. The two key differences between the
+$CheckContention$ algorithm and prior work are: \begin{enumerate}
+	\item The owner thread can be \emph{optionally} paused. Pausing the owner thread emulates the race-detection capability of
+DataCollider. Using this approach, we can pinpoint two locations where contention (and potentially races) can
+absolutely occur.
+
+Not pausing the owner thread means that \Toolname{} will detect contention as an ownership transfer. The benefit of
+this approach is that it can find contention even within critical sections. The drawback is that accuracy will not be perfect: the last access of
+the sampled cache line by the owner thread might not be the instruction at which it obtained ownership of the cache line.
+
+Our sampling approach also enables a middle ground: pausing for a period that is shorter than the sampling period, thus converting
+to relaxed ownership transfer detection after the delay expires.
+
+% TODO: Measure these improvements!!
+	\item Unlike in Dr. Memory, sampling (and therefore detection) is immediately disabled once contention is detected. This improves
+performance in the contended case because it reduces contention on the shadow memory itself.
+\end{enumerate}
+
+% --------------------------------------
+\subsection{Reporting}
+
+% Once contention is detected, sampling
+%is disabled for the cache line, reporting information is collected, and execution continues.
+
+%$SampleId$s and $TypeId$s
+%are distinct because for two reasons: \begin{enumeration}
+%\item Multiple samples for the same $TypeId$ are taken. This happens when objects of a type span multiple cache lines, and when more
+%samples than types are requested\footnote{The number of samples to take per sampling period is tunable via a command-line argument.}.
+%\item For reporting purposes, we want to accurately report the type of object affected by cache line contention. The distinction between
+%samples and types enables each sample to record more elaborate information: address of the typed object, which components of the object
+%are located in the sampled cache line (in case an object spans multiple lines)
+%\end{enumeration}
+%multiple samples for a given type might be taken, and
+%
+%  The distinction between $SampleId$ and $TypeId$ is primarily useful for reporting: at the program's
+%initialization, each $SampleId$ is allocated space to describe a detected case of cache line contention. The description includes information
+%about what types of memory was being sampled, what offsets into that memory were being sampled, stack traces for the contending
+%threads, addresses accessed by each thread, and details about each thread's memory access type (read/write, atomic, width).
+%
+%
+%When a cache line is sampled by
+%the monitor thread, it is assigned a unique 16-bit $SampleId$.
+
+%\Figref{layout_of_shadow} shows the organi
+
+\begin{figure}[t]
 \begin{center}
-\epsfig{file=states.eps, height=30em}
+\includegraphics{states.pdf}
 \end{center}
-\caption{\label{fig:states_of_shadow}States of individual shadow memory cells. Shadow memory cells begin as unwatched. %
-Then, when a cache line is selected for sampling, it becomes watched. This is represented by the presence of a $SampleId$. %
-The first thread that tries to access the shadow memory of the cache line will become the cache line's owner, as represented %
-by the presence of a $ThreadId$. If multiple threads attempt to simultaneously acquire ownership of the cache line, then %
-contention will be detected and reported. If not, then execution will continue until a second thread accesses the cache line, %
-thus triggering an ownership transfer.}
+\caption{\label{fig:states_of_shadow}Valid states of shadow memory cells. Shadow cells are a $\langle SampleId,ThreadId \rangle$ %
+pair, where a $SampleId$ of $0$ represents unwatched. In this example, sampling of a cache line is enabled by assigning %
+$SampleId \gets 11$. The cache line becomes ``owned" when thread $5$ accesses the cache line. Owernship acquisition is %
+performed by the \texttt{CAS} operation in the $CheckContention$ algorithm (\Figref{algorithm}). If the \texttt{CAS} fails, %
+then contention is reported, and sampling of the line is disabled.}
+%
+%Shadow memory cells begin as unwatched. %
+%Then, when a cache line is selected for sampling, it becomes watched. This is represented by the presence of a $SampleId$. %
+%The first thread that tries to access the shadow memory of the cache line will become the cache line's owner, as represented %
+%by the presence of a $ThreadId$. If multiple threads attempt to simultaneously acquire ownership of the cache line, then %
+%contention will be detected and reported. If not, then execution will continue until a second thread accesses the cache line, %
+%thus triggering an ownership transfer.}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[t]
 \lstset{language=[x64]Assembler}
@@
 optimizes on instruction footprint by performing lightweight, racy checks up front as a way of avoiding more heavyweight %
-contention detection. If the cache line appears watched, and if the race checks fail, then the $CheckContention$ function %
-is invoked. This function handles the state transitions of \Figref{states_of_shadow} internally using an atomic compare and %
-swap.}
+contention detection. If the cache line appears watched, and if the current thread doesn't own the line, then the %
+$CheckContention$ algorithm is invoked. This algorithm either acquires ownership for the current thread, or detects and %
+reports contention.}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[t]
 \begin{lstlisting}[language=C,basicstyle=\footnotesize\ttfamily]
@@
 
-Initially, a every cache line is unwatched, as represented by the absence of a $SampleId$ in \Figref{states_of_shadow}, and by the value
-\texttt{0} in shadow memory. When a cache line is sampled by the monitor thread, it is assigned a unique 16-bit $SampleId$. $SampleId$s
-are distinct from $TypeId$ because multiple samples for a given type might be taken. This happens when objects of a type span multiple
-cache lines, and when more samples than types are requested\footnote{The number of samples to take per sampling period is tunable %
-via a command-line argument.}. The distinction between $SampleId$ and $TypeId$ is primarily useful for reporting: at the program's
-initialization, each $SampleId$ is allocated space to describe a detected case of cache line contention. The description includes information
-about what types of memory was being sampled, what offsets into that memory were being sampled, stack traces for the contending
-threads, addresses accessed by each thread, and details about each thread's memory access type (read/write, atomic, width).
-
-If a sampled memory location is accessed, then the instrumentation instructions of \Figref{inline_assembly} will detect the non-zero
-$SampleId$, and call out to the $CheckContention$ function, but only if the current thread does not appear to own the cache line
-already. The $CheckContention$ algorithm (\Figref{algorithm}) performs both the ``Acquire" and the ``Detect" state transitions of
-\Figref{states_of_shadow} using a single atomic compare-and-swap operation (\texttt{CAS}).
+% TODO: Evaluate actual memory overheads of shadow?
+%In shadow memory, \Toolname{} stores
+%s that are used to track the state of some native cache lines. 
+%Using instrumentation, native addresses %
+%are mapped to shadow addresses as follows: $(NativeAddr/64)+ShadowBaseAddr$.
+% to record four bytes of state for every 64-byte cache line.
 
@@
 
-\section{Optimization}\label{sec:optimizations}
+%\section{Optimizations}\label{sec:optimizations}
 

### 2014-12-18 HotOS'15/paper.tex
I think this is good enough for now.

@@
 
-% For leaving some comments in the draft.
-\newcommand{\comment}[1]{}
+% Name of the tool.
 \newcommand{\TextToolname}{Malcontent}
@@
 \newcommand{\BoldToolname}{\textsc{\bf\TextToolname{}}}
+
+\renewcommand{\Secref}[1]{Section~\ref{sec:#1}}
 
@@
 % --------------------------------------------------------------------------------------
-\section{Methodology}\label{sec:methodology}
+\section{Implementation}\label{sec:implementation}
 
@@
 % --------------------------------------
-\subsection{Training}
+\subsection{Interposition}
+
+\begin{figure}[t]
+\begin{lstlisting}[language=C,basicstyle=\footnotesize\ttfamily]
+WRAP_FUNCTION(libc, malloc, (void *),
+                            (size_t size)) {
+  auto malloc = WRAPPED_FUNCTION;
+  auto context = RETURN_ADDRESS;
+  auto addr = malloc(size);
+  if (addr) {
+    auto type_id = TypeIdFor(context, size);
+    AddSamplePoint(type_id, addr);
+  }
+  return addr;
+}
+\end{lstlisting}
+\caption{\label{fig:wrapper}Wrapper for the \texttt{malloc} function in \texttt{libc}. The wrapper invokes the native malloc %
+function, thus preserving native memory layouts, and determines the $TypeId$ for the allocated address. Finally, the }
+\end{figure}
+
+The interposition layer is responsible for assigning $TypeId$s to different types of allocated objects, and maintaining lists of
+the most recently allocated object addresses on a per-type basis. These lists are used by the monitor to comprehensively
+sample the heap.
+
+\Toolname{} uses Granary's wrapper mechanism \cite{Granary} to interpose on the dynamic memory allocation routines
+used by the application. By default, \Toolname{} interposes on \texttt{malloc} and other POSIX memory management
+routines  available from \texttt{libc}, as well as \texttt{new} from \texttt{libc++}. \Toolname{} can also be configured to
+interpose on non-standard allocators, either by using symbol information, or the addresses of the allocators within the
+binaries to be instrumented.
+
+\Figref{wrapper} shows an example wrapper for the \texttt{malloc} function. The wrapper determines the $TypeId$ for the
+allocated object given the size of the object and the allocation context (heuristically estimated using the return address into
+the caller of the \texttt{malloc} function). For finer-grained $TypeId$ associations, or more detailed reporting, calls stacks
+can be provided as the allocation context.
+
+{\bf Future work:} Like SHERIFF \cite{SHERIFF}, try to fix some cache contention issues by padding allocated objects.
 
 % --------------------------------------
-\subsection{Shadow Memory}
+\subsection{Monitor}
+
+The monitor thread is spawned at program startup and is responsible for periodically choosing memory locations to sample,
+then enabling their watchpoints. When a sampling period is over, the enabled watchpoints are disabled, and any detected
+contention is reported.
+
+The monitor employs a round-robin sampling algorithm. For each $TypeId$, it chooses the most recently allocated object, as
+recorded by the interposition layer. Then, it enables \emph{all} watchpoints associated with the cache lines spanned by
+the allocated object. When the maximum configured number of samples have been placed, it goes to sleep, so as to give the
+placed watchpoints time to be hit and detect contention. Eventually, it wakes up and reported any instances of detected
+contention.
+
+{\bf Future work: Biased sampling.} Bias sampling toward objects that are more typically accessed in \emph{cold} code.
+This would agree with the ``cold code hypothesis'' \cite{LiteRace}, wherein important sources of cache contention (in
+the case of \cite{LiteRace}, the focus was data races) are likely to be in colder code because otherwise those issues would
+have already been discovered, or are benign.
+
+{\bf Future work: Threshold sampling.} Disable sampling of some types when the number of detected issues exceeds some
+bound.
+
+% --------------------------------------
+\subsection{Instrumentation}
+\paragraph{Shadow Memory}
 \begin{figure*}
@@
 
-% --------------------------------------
-\subsection{Instrumentation}
+\begin{figure}[t]
+\begin{center}
+\includegraphics{states.pdf}
+\end{center}
+\caption{\label{fig:states_of_shadow}Valid states of shadow memory cells. Shadow cells are a $\langle SampleId,ThreadId \rangle$ %
+pair, where a $SampleId$ of $0$ represents unwatched. In this example, sampling of a cache line is enabled by assigning %
+$SampleId \gets 11$. The cache line becomes ``owned" when thread $5$ accesses the cache line. Owernship acquisition is %
+performed by the \texttt{CAS} operation in the $CheckContention$ algorithm (\Figref{algorithm}). If the \texttt{CAS} fails, %
+then contention is reported, and sampling of the line is disabled.}
+%
+%Shadow memory cells begin as unwatched. %
+%Then, when a cache line is selected for sampling, it becomes watched. This is represented by the presence of a $SampleId$. %
+%The first thread that tries to access the shadow memory of the cache line will become the cache line's owner, as represented %
+%by the presence of a $ThreadId$. If multiple threads attempt to simultaneously acquire ownership of the cache line, then %
+%contention will be detected and reported. If not, then execution will continue until a second thread accesses the cache line, %
+%thus triggering an ownership transfer.}
+\end{figure}
+
+\paragraph{Shadow Instrumentation}
+
+Shadow memory instrumentation is applied only to those memory-accessing instructions identified during the training phase. For all
+other instructions, Granary's ``null" instrumentation is applied. In some cases, \Toolname{} can even instruct Granary to execute code
+natively. The rest of this section focuses on shadow instrumentation.
+
 If a sampled memory location is accessed, then the instrumentation instructions of \Figref{inline_assembly} will detect the non-zero
@@
 
-Not pausing the owner thread means that \Toolname{} will detect contention as an ownership transfer. The benefit of
-this approach is that it can find contention even within critical sections. The drawback is that accuracy will not be perfect: the last access of
-the sampled cache line by the owner thread might not be the instruction at which it obtained ownership of the cache line.
+Not pausing the owner thread means that \Toolname{} can detect contention even within critical sections. The drawback
+is that accuracy will not be perfect: the last access of the sampled cache line by the owner thread might not be the
+instruction at which it obtained ownership of the cache line.
 
@@
 	\item Unlike in Dr. Memory, sampling (and therefore detection) is immediately disabled once contention is detected. This improves
-performance in the contended case because it reduces contention on the shadow memory itself.
+performance in the contended case because it reduces contention on the cache lines associated with shadow memory, as well as reduces
+CPU overheads due to executing fewer inject instrumentation instructions.
 \end{enumerate}
-
-% --------------------------------------
-\subsection{Reporting}
 
@@
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics{states.pdf}
-\end{center}
-\caption{\label{fig:states_of_shadow}Valid states of shadow memory cells. Shadow cells are a $\langle SampleId,ThreadId \rangle$ %
-pair, where a $SampleId$ of $0$ represents unwatched. In this example, sampling of a cache line is enabled by assigning %
-$SampleId \gets 11$. The cache line becomes ``owned" when thread $5$ accesses the cache line. Owernship acquisition is %
-performed by the \texttt{CAS} operation in the $CheckContention$ algorithm (\Figref{algorithm}). If the \texttt{CAS} fails, %
-then contention is reported, and sampling of the line is disabled.}
-%
-%Shadow memory cells begin as unwatched. %
-%Then, when a cache line is selected for sampling, it becomes watched. This is represented by the presence of a $SampleId$. %
-%The first thread that tries to access the shadow memory of the cache line will become the cache line's owner, as represented %
-%by the presence of a $ThreadId$. If multiple threads attempt to simultaneously acquire ownership of the cache line, then %
-%contention will be detected and reported. If not, then execution will continue until a second thread accesses the cache line, %
-%thus triggering an ownership transfer.}
-\end{figure}
-
-\begin{figure}[t]
+\begin{figure}
 \lstset{language=[x64]Assembler}
@@
 %\begin{enumerate}
-%\item We iteratively refine the granularity of proxy memory, starting from the page granularity and going down to the cache
+%\item We iteratively refine the granularity of proxy memory, starting from the page granularity and goig down to the cache
 %line granularity. As the granularity is refined and potential false sharing continues to be detected, our confidence in beliefs
@@
 \section{Evaluation}\label{sec:evaluation}
-Our evaluation aims to answer the following questions:
-\begin{itemize}
-\item Does the training phase significant reduce the performance overheads of the detection phase?
-
-\end{itemize}
+I want to evaluate the following things:
+
+\paragraph{Training} Here, I want to focus on measuring the reduction in overheads provided by the training phase. There will
+be two ways of showing this. The first is tabular, and will compare dynamic instruction counts. Specifically, it will measure how
+many instructions would execute with heavy instrumentation with training, how many without, and how many execute natively.
+
+Another thing that might be worth showing is that for some levels of sampling (e.g. 16 samples, 64 samples, 356 samples, etc.) how
+many of the heavily instrumented instructions go through a "heavyweight" instrumentation path. This could give insights on
+the actual coverage of sampling, and might inform us on whether or not we need to change our sampling strategies. Hopefully
+there will be some insights there.
+
+The next is way of showing the benefits are in terms of overhead. Here, we could either focus on a particular benchmark (e.g.
+Phoenix, PARSEC), or on a few applications (MySQL, Memcached, etc.). This would be a standard grouped bar chart, with three
+bars per application. Each bar is overhead relative to native. The first bar for each application shows baseline Granary overheads.
+The second bar shows overheads without training. The third bar shows overheads with training.
+
+\paragraph{Sampling}
+Here the purpose is to compare the performance benefits of ``comprehensive sampling" vs. ``comprehensive tracking". This serves
+as a direct comparison to Dr. Memory. Again, we want a grouped bar chart. The first bar for each application is a Granary version
+of Dr. Memory, which is a basically our tool without training without sampling. The second group of bars is for untrained \Toolname{}
+but using sampling, with increasing number of watchpoints. The third group of bars is for trained, with the same numbers of watchpoints.
+
+It would also be nice if we could have a small table that chooses a select few of the results from the graph, and shows data cache
+hit/miss ratios, as well as icache hit/miss ratios.
+
+\paragraph{Memory Overheads}
+This needs to show the memory overheads of \Toolname{}, both in terms of shadow and in terms of code cache. I've described that
+the shadow has an upper-bound on its memory overheads; however, I haven't actually implemented it as such. In practice though,
+that might not matter if none of our tests reach the theoretical upper bound.
+
+\paragraph{Effectiveness}
+For this to be useful, we need to find real bugs, and showing an example of a report for a real bug will be key I think. In the short
+time before HotOS, I don't expect to find much, as getting everything working will be its own challenge. This isn't discouraging me
+though as I feel that this is the first ``real" tool that Granary has, and for that reason, it's something that I'm very pleased with.
+
+\section{Conclusion}
+This paper presented \Toolname{}, a new cache contention detection framework that combines the strengths of DataCollider and
+Dr. Memory into a unified framework. The key novelties of \Toolname{} are the separation of the problem into two phases. The
+first phase figures out what code can contain bugs. The second phase detects bugs in that code. The tool, but more specifically
+Granary itself, is still under heavy development, and so performance and effectiveness evaluations could not be performed. It is
+hoped that Granary's stability will improve enough in time to get some kind of numbers.
 
